### 2021-11-04 paper/paper1/ms.tex
Scaffolding and template for a paper, minimal content

\documentclass[modern]{aastex631}
\bibliographystyle{aasjournal}
\turnoffedit
\usepackage[caption=false]{subfig}
\usepackage{booktabs}
\usepackage{censor}

\def\Teff{T_{\rm eff}}
\def\vsini{v\sin{i}}
\def\kmps{\mathrm{km}\;\mathrm{s}^{-1}}

\begin{document}
\shorttitle{blas\'e}
\shortauthors{Gully-Santiago}
\title{Transfer learning for echelle spectroscopy}

\author{Michael Gully-Santiago}
\affiliation{University of Texas at Austin Department of Astronomy}

\begin{abstract}

  We introduce blas\'e, a framework for transfer learning for high-grasp echelle spectroscopy.

\end{abstract}

\keywords{High resolution spectroscopy (2096)}

\section{Introduction}\label{sec:intro}

Here is an annotated bibliography.

\begin{deluxetable}{chc}
  \tablecaption{Annotated bibliography for intro\label{table1}}
  \tablehead{
    \colhead{Reference} & \nocolhead{two} & \colhead{Key idea}
  }
  \startdata
  \citet{2017ApJ...836..200G} & - & Extensions to Starfish
  \enddata
\end{deluxetable}

\newpage

\section{Methodology}

TBD

\section{Conclusions}

More placeholder text...

\begin{acknowledgements}
  The author acknowledges the Texas Advanced Computing Center (TACC, \url{http://www.tacc.utexas.edu}) at The University of Texas at Austin for providing HPC resources that have contributed to the research results reported within this paper.
\end{acknowledgements}

\clearpage

\facilities{HET (HPF)}

\software{  pandas \citep{mckinney10},
  matplotlib \citep{hunter07},
  astropy \citep{exoplanet:astropy13,exoplanet:astropy18},
  exoplanet \citep{exoplanet:joss}, %celerite
  numpy \citep{harris2020array},
  scipy \citep{2020SciPy-NMeth},
  ipython \citep{perez07},
  starfish \citep{czekala15},
  seaborn \citep{Waskom2021},
  pytorch \citep{2019arXiv191201703P}}

\bibliography{ms}

\clearpage

\appendix
\restartappendixnumbering

\section{Autodiff themes} \label{appendix:tools}

Here are some more details about autodiff
\end{document}

### 2021-11-04 paper/paper1/ms.tex
Writing prose in the paper

@@
 
-Here is an annotated bibliography.
+\subsection{Spectral fitting past and present}
 
+Below is a table exploring recent techniques for spectral fitting.
+
+\movetabledown=13mm
+\begin{rotatetable}
   \begin{deluxetable}{chc}
@@
     \startdata
-  \citet{2017ApJ...836..200G} & - & Extensions to Starfish
+    \citet{2015PhDT........82P} & - & \texttt{specmatch-syn} Synthetic spectral template matching \\
+    \citet{czekala15} & - & \texttt{starfish} robust spectral inference on precomputed models \\
+    \citet{2016ApJS..223....8C} & - & \texttt{sick} Spectral Inference Crank, similar to \texttt{Starfish}\\
+    \citet{2017ApJ...836..200G} & - & Modular extensions to \texttt{starfish}: starspots \\
+    \citet{2018ApJ...862...85G} & - & Modular extensions to \texttt{starfish}: veiling \\
+    \citet{2017ApJ...836...77Y} & - & \texttt{specmatch-emp} Empirical template matching for visible\\
+    \citet{exoplanet:astropy18} & - & \texttt{astropy} General purpose tools for spectral manipulation\\
+    \citet{2019AJ....158..164B} & - & \texttt{wobble}: data driven stellar and telluric models
     \enddata
   \end{deluxetable}
+
+\end{rotatetable}
+
+\subsection{Seeking physical self-consistency and high predictive accuracy}
+Ideally we seek models that are both physically self-consistent and highly accurate when compared to data.
+We want tunable models that can be evaluated quickly for use in likelihood calculations.
+We want likelihood calculations to accurately account for nearly-degenerate parameters, common in retrievals in the Planetary Science and Remote Sensing community.
+We want unbiased, noise-free templates for use in cross-correlation spectroscopy.
+We want precision and accuracy over a large spectral grasp: simultaneously wide bandwidth and high spectral resolution.
+We want high out-of-band prediction accuracy.
+Ideally we also want precision and accuracy in the modeling of--or at least separation of--affiliated telluric and instrumental signals.
+
+These goals form the backbone of spectral calibration, an increasingly valuable specialty as the deficits in our models become intolerable with greater data quantity and quality.
+
+\subsection{Current limitations to our spectral models}
+
+Several practical and ostensibly fundamental barriers limit the accuracy of spectral models.
+
+First, High latent dimensionality: Many effects matter at the precision level of the data...
+
+Second, Degeneracy among parameters
+
+- [ ]
+- [ ] Underlying atomic and molecular data may be wrong or approximate or missing
+- [ ] Underlying radiative transfer may be inaccurate or approximate (T-P profile / T-Tau)
+- [ ] Time is a dimension: Star is changing
+- [ ] Stellar surface inhomogeneities become important: Doppler Imaging and limb darkening
+- [ ] We do not know how to accurately parameterize some dimensions
+- [ ] High dynamic range is both a blessing and a curse:
+- [ ] Feeding back locations of model imperfections becomes unwieldy for many lines
+- [ ] Many lines essentially becomes a book-keeping and continuum assignment problem
+- [ ] Computational problem
+- [ ] Some lines have large line wings, blurring continuum and line
+- [ ] Line blanketing has almost no continuum
+- [ ] So in practice, tradeoffs between model flexibility and physical self-consistency
+- [ ] Semi-empirical models may be a middle ground: informed from models, but revised with data
+- [ ] The radiative transfer step is computationally expensive, so we want to avoid redoing that
+- [ ] Instead focus on pre-computed models that have that wisdom baked-in
+- [ ] We are essentially compressing pre-computed models into evaluable, interpretable models
+- [ ] Transfer learning is one plausible approach: pre-train on pre-computed models, transfer to data
+- [ ] Neural network frameworks make this possible with autodiff
+- [ ] We use the autodiff and GPU acceleration, but not the neural architectures
+- [ ] We do not use the neural architectures because we have good understand for how a spectrum is generated
+- [ ] Instead of tuning weights of a neural network, we tune the atomic and molecular properties of a spectrum
 
@@
   astropy \citep{exoplanet:astropy13,exoplanet:astropy18},
-  exoplanet \citep{exoplanet:joss}, %celerite
+  exoplanet \citep{exoplanet:joss}, %celerite?
   numpy \citep{harris2020array},

### 2021-11-05 paper/paper1/ms.tex
Minor prose tweaks

@@
 We want tunable models that can be evaluated quickly for use in likelihood calculations.
-We want likelihood calculations to accurately account for nearly-degenerate parameters, common in retrievals in the Planetary Science and Remote Sensing community.
+We want posterior inference to accurately explore correlations among parameters.
 We want unbiased, noise-free templates for use in cross-correlation spectroscopy.
-We want precision and accuracy over a large spectral grasp: simultaneously wide bandwidth and high spectral resolution.
+We want accuracy over a large spectral grasp: simultaneously wide bandwidth and high spectral resolution.
 We want high out-of-band prediction accuracy.
-Ideally we also want precision and accuracy in the modeling of--or at least separation of--affiliated telluric and instrumental signals.
+Ideally we also want precision and accuracy in the separation of affiliated telluric and instrumental signals.
 
@@
 
-Several practical and ostensibly fundamental barriers limit the accuracy of spectral models.
-
-First, High latent dimensionality: Many effects matter at the precision level of the data...
+Several practical barriers limit the accuracy of spectral models.
+First and foremost, real stars are usually more complicated than our simplified models of them.
+Real spectra often vary over more dimensions than specified.
+Conspicuous examples of these hidden variables can be found in protostars: starspots, accretion veiling, dust extinction, and magnetic Zeeman splitting.
+Jointly modeling all of these phenomena alongside the intrinsic stellar photosphere and  High latent dimensionality: Many effects matter at the precision level of the data...
 

### 2021-11-05 paper/paper1/ms.tex
Outline the entire paper in latex, with bullet points for themes

@@
 \usepackage{censor}
+
+\usepackage{outlines}
+\usepackage{enumitem}
 
@@
 
-- [ ]
-- [ ] Underlying atomic and molecular data may be wrong or approximate or missing
-- [ ] Underlying radiative transfer may be inaccurate or approximate (T-P profile / T-Tau)
-- [ ] Time is a dimension: Star is changing
-- [ ] Stellar surface inhomogeneities become important: Doppler Imaging and limb darkening
-- [ ] We do not know how to accurately parameterize some dimensions
-- [ ] High dynamic range is both a blessing and a curse:
-- [ ] Feeding back locations of model imperfections becomes unwieldy for many lines
-- [ ] Many lines essentially becomes a book-keeping and continuum assignment problem
-- [ ] Computational problem
-- [ ] Some lines have large line wings, blurring continuum and line
-- [ ] Line blanketing has almost no continuum
-- [ ] So in practice, tradeoffs between model flexibility and physical self-consistency
-- [ ] Semi-empirical models may be a middle ground: informed from models, but revised with data
-- [ ] The radiative transfer step is computationally expensive, so we want to avoid redoing that
-- [ ] Instead focus on pre-computed models that have that wisdom baked-in
-- [ ] We are essentially compressing pre-computed models into evaluable, interpretable models
-- [ ] Transfer learning is one plausible approach: pre-train on pre-computed models, transfer to data
-- [ ] Neural network frameworks make this possible with autodiff
-- [ ] We use the autodiff and GPU acceleration, but not the neural architectures
-- [ ] We do not use the neural architectures because we have good understand for how a spectrum is generated
-- [ ] Instead of tuning weights of a neural network, we tune the atomic and molecular properties of a spectrum
+\begin{itemize}
+  \item Underlying atomic and molecular data may be wrong or approximate or missing
+  \item Underlying radiative transfer may be inaccurate or approximate (T-P profile / T-Tau)
+  \item Time is a dimension: Star is changing
+  \item Stellar surface inhomogeneities become important: Doppler Imaging and limb darkening
+  \item We do not know how to accurately parameterize some dimensions
+\end{itemize}
 
-\newpage
+\subsection{High dynamic range is both a blessing and a curse:}
+\begin{itemize}
+  \item Feeding back locations of model imperfections becomes unwieldy for many lines
+  \item Many lines essentially becomes a book-keeping and continuum assignment problem
+  \item Computational problem
+  \item Some lines have large line wings, blurring continuum and line
+  \item Line blanketing has almost no continuum
+\end{itemize}
+
+\subsection{This paper: towards semi-empirical models}
+\begin{itemize}
+  \item So in practice, tradeoffs between model flexibility and physical self-consistency
+  \item Semi-empirical models may be a middle ground: informed from models, but revised with data
+  \item The radiative transfer step is computationally expensive, so we want to avoid redoing that
+  \item Instead focus on pre-computed models that have that wisdom baked-in
+  \item We are essentially compressing pre-computed models into evaluable, interpretable models
+  \item Transfer learning is one plausible approach: pre-train on pre-computed models, transfer to data
+  \item Neural network frameworks make this possible with autodiff
+  \item We use the autodiff and GPU acceleration, but not the neural architectures
+  \item We do not use the neural architectures because we have good understand for how a spectrum is generated
+  \item Instead of tuning weights of a neural network, we tune the atomic and molecular properties of a spectrum
+\end{itemize}
 
@@
 
-TBD
+\subsection{Overall Architecture and Design Choices}
+Here is an outline
+\begin{outline}
+  \1 Currently limit the bandwidth to a single high-bandwidth echelle spectrograph (plus buffer at edges)
+  \2 This choice is principally computational at the moment, no fundamental limit
+  \1 We fetch a high resolution pre-computed spectrum from PHOENIX
+  \1 We use the exact native wavelength sampling, yielding N pixels
+  \1 Tradeoff of working in log or linear flux
+  \2 First normalize by a scalar since we do not intend to use absolute flux calibration
+  \2 Log prevents negative fluxes
+  \2 Trickier to do Pseudo-Voigt approximation as sum in linear space
+\end{outline}
+
+\subsection{Initialization}
+\begin{outline}
+  \1 Initialize the model with preprocessing steps:
+  \2 Low-pass filter to remove narrow spectral lines
+  \2 (Optional) Place known Hydrogen, Sodium lines with large wings and non-standard line shapes
+  \2 Define a prominence threshold and find locations of peaks above that threshold
+  \2 Derive coarse properties about those peaks: widths and amplitudes
+  \1 We now have a coarse model that needs to be fine-tuned with autodiff
+\end{outline}
+
+\subsection{Continuum model}
+\begin{outline}
+  \1 How to describe the continuum:
+  \2 Black body / Rayleigh Jeans plus perturbation
+  \2 High order polynomial: Chebyshev? Somewhere between 3rd and 20th order?
+  \2 Use cross validation?
+  \1 GPs are also possible, but high Number of pixels--> need scalable GP
+  \2 Celerite not implemented for PyTorch?
+  \2 GPyTorch possible, but has certain restrictions on kernel and sampling
+\end{outline}
+
+\subsection{Lineshapes choice and tradeoffs}
+\begin{outline}
+  \1 Faced with a question: what lineshape to use
+  \2 To First order the lines are Gaussian or Lorentzian
+  \2 To Second order the lines are Voigt Profiles
+  \2 In detail the lines are the result of radiative transfer that can smear the Voigt profile
+  \2 In practice the high-res lines get convolved with instrumental profiles, so details don't matter too much
+  \1 Computational tradeoff of Lorentzian/Gaussian versus Pseudo-Voigt versus Voigt
+  \2 Additional Implementation challenge: Fadeeva function not implemented in PyTorch
+  \1 We choose a pseudo-Voigt as a balance between computation and adequate accuracy
+\end{outline}
+
+\subsection{GPU and Autodiff specific considerations}
+\begin{outline}
+  \1 Total number of parameters
+  \2 For the lines =  N-lines x 4 (lambda center, fwhm L, sigma G, Amplitude)
+  \2 For the continuum = Nth order polynomial = N+1 parameters, or Gaussian Process = 2 hyperparameters
+  \1 Each Tunable parameter gets `grad=True`
+  \1 We assemble the product of all lines and all continuum sources, this is the computational bottleneck
+  \1 This becomes a matrix of size N-lines x N-wavelength points
+  \1 We take the element-wise product over the lines axis, contracting to 1 x N-wavelength points
+  \1 For N-lines = X and N-wavelength-points = Y, we expect X*Y*k primative operations.
+  \1 For N-lines = X and N-wavelength-points = Y, we expect X*Y*32 bits of memory => XX GB
+  \1 NVIDIA GPUs can handle 8 - 40 GB of memory depending on the model
+\end{outline}
+
+\subsection{Minibatches}
+\begin{outline}
+  \1 In the parlance of Machine Learning, we have a big model and small data
+  \1 Really our model is too big: most of the entries are zero, so we should use a sparse implementation
+  \1 If the model is too big to fit in memory, we have to train in mini-batches
+  \1 We hand-in some fraction f ~ 1/20th of the entire wavelength range as a mini-batch
+  \1 The model is only evaluated at that 5\% of pixels,  $N_{lines} \times N_{pix} \cdot 0.05$)
+  \1 Technically all lines are updated
+  \1 In practice, only lines that are sufficiently informed by those pixels will get large updates
+  \1 Causes some stochasticity--- Stochastic Gradient Descent "SGD"
+  \1 Analogous but different from the notion of minibatch in ML: not a distinct axis, like $N_{images}$
+  \1 We use random minibatches
+\end{outline}
+
+\subsection{Optimization and training}
+\begin{outline}
+  \1 Run the training:
+  \1 Set the optimizer: Adam with a LR=0.004
+  \1 Set the loss function: Chi-squared with no per-pixel uncertainty
+  \1 Set the "mini-batch" fraction f~1-5%
+  \1 Choice of epochs or steps $N_steps = 300-10000$
+  \1 The training process sees 20 minibatches per epoch, sees $N_pixels$, with some repeats some misses
+  \1 Monitor with Tensorboard: see the improvement
+  \1 The training takes X minutes on an RTX2070 GPU with PyTorch v1.11.x, CUDA v11.x, and Intel X CPU
+  \1 We save the model parameters, refer to them collectively as the pre-trained model
+\end{outline}
+
+\section{Cloning Performance}
+
+\subsection{Results and scaling}
+
+\begin{outline}
+  \1 We compute the residual of native PHOENIX minus cloned model
+  \1 We see X percent residuals at native resolution
+  \1 We smooth the residual to HPF instrumental resolution and see Y percent residuals
+  \1 The main residuals come from missing line opacity due to our prominence threshold:
+  \1 Including smaller prominence lines means smaller residuals, but more lines and higher computational cost
+\end{outline}
+
+\subsection{Line data clustering and conspicuous flaws}
+\begin{outline}
+  \1 We see some lines that devolve into missing continuum opacity
+  \1 Especially acute near broad line wings if you don't handle the non-standard line shapes
+  \1 How do the lineshapes cluster in $\sigma_G$ vs $FWHM_L$?
+  \1 How do the lineshapes cluster in Amplitude vs FWHM? Can some lines be rejected?
+  \1 Does the pre-trained model round-trip effectively (model saved at end of training = model from loaded params)?
+\end{outline}
+
+\section{Refining the model from data with transfer learning}
+\begin{outline}
+  \1 Augment the pre-trained model with extrinsic parameters:
+  \2 vsini and RV
+  \1 How to re-sample into the data space?
+  \1 Option 1: Convolve the Pseudo-Voigt with a Instrumenal Gaussian
+  \2 The convolution of a Gaussian with a Pseudo-Voigt is simply a new Pseudo-Voigt!
+  \2 Simply evaluate on a new coarse wavelength grid!  Saves computation!
+  \1 Option 2: Numerically convolve output on a fine grid and sum pixels
+  \2 Computationally wasteful
+  \2 Requires generating full bandwidth spectrum at native resolution
+  \1 Both options should give the same answer (right?)
+  \2 Need to double-check commutivity of convolution and element-wise product...
+  \1 Ideally you need wavelength-dependent resolution, not just a single stock resolving power
+  \1 Need accurate per-pixel uncertainties
+  \1 We have to apply some amount of regularization to the pre-trained model
+  \2 Regularization tunes how much to trust the data versus the model in the semi-empirical model
+  \2 No Regularization means likely to overfit noise in the data
+  \2 Too much regularization resembles a static (fixed) PHOENIX model
+\end{outline}
+
+\section{Performance of transfer learning}
+\begin{outline}
+  \1 What is the residual level with bare PHOENIX (i.e. tuning continuum only, not lines)?
+  \1 What is the residual level with no regularization (i.e. tune all lines and continuum)?
+  \2 Should be near-zero except for missing lines
+  \2  What should we do about missing lines?
+  \1 What is the residual level with modest regularization?  What is the typical change to the line?
+  \2 Plot of cloned FWHM versus FWHM transferred
+  \2 The pseudo-voigt line properties can be analytically integrated to give an equivalent width
+  \2 Plot of cloned EW (before) versus transfer EW (after)
+\end{outline}
+
+\section{Discussion}
+
+\subsection{Questioning the assumptions of the method}
+\begin{outline}
+  \1 Is PseudoVoigt Adequate?
+  \1 What performance level is adequate? (depends on your application)
+  \1 What about PRV applications
+  \1 What about the choice of minibatch size?
+  \1 How would a sparse tensor change the computational performance?
+  \1 Are we using the GPU effectively?  Are we memory bandwidth limited?
+  \1 Should we update the line center positions?
+\end{outline}
+
+\subsection{Limitations}
+\begin{outline}
+  \1 What to do when continuum is absent even in at the native resolution of precomputed synthetic model?
+  \1 Relevant to brown dwarfs and possibly M dwarfs molecular bands
+  \1 Bandwidth limitations
+  \1 Interplay with tellurics
+\end{outline}
 
 \section{Conclusions}
+More placeholder text...
 
-More placeholder text...
+\
 
@@
 Here are some more details about autodiff
+
+% Table
+\begin{deluxetable}{cc}
+  \tablecaption{Notation used in this paper\label{table2}}
+  \tablehead{
+    \colhead{Symbol} & \nocolhead{Meaning}
+  }
+  \startdata
+  $N_{pix}$ & Number of pixels in the precomputed spectrum \\
+  $N_{lines}$ & Number of spectral lines \\
+  $\sigma_G$ & Gaussian line profile scale and standard deviation \\
+  $\gamma_L$ & Lorentzian line profile half width
+  \enddata
+\end{deluxetable}
+
 \end{document}

### 2021-11-05 paper/paper1/ms.tex
Outline to text, voila!

@@
 \usepackage{outlines}
-\usepackage{enumitem}
+\usepackage{amsmath,bm}
 
@@
 \subsection{Overall Architecture and Design Choices}
-Here is an outline
-\begin{outline}
-  \1 Currently limit the bandwidth to a single high-bandwidth echelle spectrograph (plus buffer at edges)
-  \2 This choice is principally computational at the moment, no fundamental limit
-  \1 We fetch a high resolution pre-computed spectrum from PHOENIX
-  \1 We use the exact native wavelength sampling, yielding N pixels
-  \1 Tradeoff of working in log or linear flux
-  \2 First normalize by a scalar since we do not intend to use absolute flux calibration
-  \2 Log prevents negative fluxes
-  \2 Trickier to do Pseudo-Voigt approximation as sum in linear space
-\end{outline}
+
+We start with a high resolution pre-computed stellar model spectrum, $\mathsf{M}(\bm{\lambda})$.
+For now, we require that the spectrum have sporadic regions of discernable continuum devoid of lines, and that the continuum varies smoothly in wavelength.
+We suspect most stellar spectra meet these criteria, with the exception of the coolest M dwarfs and brown dwarfs.
+The method is hypothetically capable of handling spectra without discernable continua, but we suspect some modifications to the preprocessing steps would be needed.
+We allow the spectrum to have any number of spectral lines, so long as the continuum criterion is met.
+We currently truncate the bandwidth of the precomputed synthetic spectra to match a high-bandwidth echelle spectrograph, extended with a buffer at the edges of size $\pm \Delta \lambda_{\mathrm{buffer}}$, chosen to account for plausible radial velocity and rotational broadening of real stars.
+A generous buffer of $v \sin{i} < 500 \;\kmps$ and $|RV|<500 \;\kmps$ yields a typical buffer of about 30~\AA.
+
+The choice of limiting the bandwidth to a region-of-interest around a single echelle spectrograph bandwidth stems from computational constraints.
+In principle there is no fundamental limit to the bandwidth one could clone with this method, up to and including the entire precomputed synthetic spectral model bandwidth.
+
+We adopt the exact native wavelength sampling with no smoothing or interpolation, yielding a wavelength vector $\bm{\lambda}$ with length $N_\mathrm{pixels}$ equal to the number of pixels within the extents of our region of interest including the buffers.
+
+At this stage we have a choice of whether to work in linear or log scale units.
+Adopting the log of the flux would ensure that the cloned model possesses only positive flux values, a desireable trait of any physical spectral model.
+We chose instead to stick with native linear flux scaling because of the availability of convenient approximations to astrophysically relevant line profiles in the linear scaling domain.
+We normalize the spectrum by its mean value since we do not intend to apply these methods to absolutely flux calibrated spectra.
 

### 2021-11-05 paper/paper1/ms.tex
writing prose from outline, continued

@@
 We start with a high resolution pre-computed stellar model spectrum, $\mathsf{M}(\bm{\lambda})$.
-For now, we require that the spectrum have sporadic regions of discernable continuum devoid of lines, and that the continuum varies smoothly in wavelength.
+This spectrum should have sporadic regions of discernable continuum devoid of lines, and the continuum must vary smoothly in wavelength.
 We suspect most stellar spectra meet these criteria, with the exception of the coolest M dwarfs and brown dwarfs.
-The method is hypothetically capable of handling spectra without discernable continua, but we suspect some modifications to the preprocessing steps would be needed.
-We allow the spectrum to have any number of spectral lines, so long as the continuum criterion is met.
-We currently truncate the bandwidth of the precomputed synthetic spectra to match a high-bandwidth echelle spectrograph, extended with a buffer at the edges of size $\pm \Delta \lambda_{\mathrm{buffer}}$, chosen to account for plausible radial velocity and rotational broadening of real stars.
+The method can hypothetically handle spectra without discernable continua, but we suspect some modifications to the preprocessing steps would be needed.
+We allow the spectrum to have any number of spectral lines.
+We truncate the red and blue limits of the precomputed synthetic spectrum to match a high-bandwidth echelle spectrograph, extended with a buffer at the edges of size $\pm \Delta \lambda_{\mathrm{buffer}}$, chosen to account for plausible radial velocity and rotational broadening of real stars.
 A generous buffer of $v \sin{i} < 500 \;\kmps$ and $|RV|<500 \;\kmps$ yields a typical buffer of about 30~\AA.
@@
 \subsection{Initialization}
+
+We initialize the cloned model with a series of preprocessing steps.
+We divide the entire spectrum by a black body of the same temperature as the model template.
+This step alone is usually adequate to move on to the spectral line finding.
+
+Occasionally a more advanced continuum flattening step may be needed to ensure that subsequent spectral line finding steps get applied uniformly.
+Here, we apply a low-pass Savitzky-Golay filter\authorcomment1{citation XX} to identify the continuum regions, and then fit a low-order polynomial to this pre-filtered continuum spectrum.
+We evaluate the low-order polynomial at all wavelength points, multiplied by the blackbody curve to assemble a ``perturbed black body'' that symbolizes the black body modulated by any broad-band continuum opacity that may reside in the model.
+We define a threshold flux loss of $1\%$, below which any lines are masked, erode that mask by $N_{\mathrm{iter}}=13$ pixels, and repeat the polynomial fitting process again.
+At this stage we fit a high order polynomial, with the polynomial order set through cross validation, typically $n_{\mathrm{poly}}\sim11$.
+The ``perturbed black body'' continuum model $\mathsf{B}$ contains  $n_{\mathrm{poly}}+1$ free parameters.
+
+In either continuum flattening case, we move on to find locations of spectral lines.
+We apply a local minimum finding algorithm on the spectrum by defining a prominence threshold $\delta f\sim0.01$.
+Testing demo with the real fixed static residual.
+
 \begin{outline}
-  \1 Initialize the model with preprocessing steps:
-  \2 Low-pass filter to remove narrow spectral lines
-  \2 (Optional) Place known Hydrogen, Sodium lines with large wings and non-standard line shapes
-  \2 Define a prominence threshold and find locations of peaks above that threshold
+  \1 Define a prominence threshold and find locations of peaks above that threshold
   \2 Derive coarse properties about those peaks: widths and amplitudes
@@
 \end{outline}
+
+The low-pass filter should be set to capture the genuine spectral shape, without over-fitting broad line-wings such as those in deep Hydrogen and sodium lines.
+We have experimentally found it helpful to hand-tune models of exceptionally broad and non-standard line-wings with with an optional pre-processing step.
 
@@
   $\sigma_G$ & Gaussian line profile scale and standard deviation \\
-  $\gamma_L$ & Lorentzian line profile half width
+  $\gamma_L$ & Lorentzian line profile half width\\
+  $\pm \Delta \lambda_{\mathrm{buffer}}$ & Buffer exceeding the red and blue limits of data spectrum
   \enddata

### 2021-11-08 examples/clone_PHOENIX_spectrum.py
working example of Phoenix cloning

@@ -4,6 +4,7 @@ from tqdm import trange
 import torch.optim as optim
 from blase.emulator import PhoenixEmulator
 import matplotlib.pyplot as plt
+from gollum.phoenix import PHOENIXSpectrum
 from torch.utils.tensorboard import SummaryWriter
 import webbrowser
 import numpy as np
@@ -48,55 +49,61 @@ writer = SummaryWriter(log_dir="runs/emulator1")
 webbrowser.open("http://localhost:6006/", new=2)
 
 
+# Pre-process the model as described in the paper
+spectrum = PHOENIXSpectrum(teff=4700, logg=4.5)
+spectrum = spectrum.divide_by_blackbody()
+spectrum = spectrum.normalize()
+
+# Numpy arrays: 1 x N_pix
+wl_native = spectrum.wavelength.value
+flux_native = spectrum.flux.value
+
 # Create the emulator
-emulator = PhoenixEmulator(4700, 4.5, prominence=0.1,)
+emulator = PhoenixEmulator(wl_native, flux_native, prominence=0.05)
 emulator.to(device)
 
 n_pix = len(emulator.wl_native)
 wl_native = emulator.wl_native.clone().detach().to(device)
 target = emulator.flux_native.clone().detach().to(device)
 
-if redo_training:
-    loss_fn = nn.MSELoss(reduction="mean")
-    optimizer = optim.Adam(emulator.parameters(), 0.004)
-    n_epochs = 300
-    sub_divisions = 20
-    losses = []
+# Training Loop
+loss_fn = nn.MSELoss(reduction="mean")
+optimizer = optim.Adam(emulator.parameters(), 0.004)
+n_epochs = 300
+sub_divisions = 20
+losses = []
 
-    t_iter = trange(n_epochs, desc="Training", leave=True)
-    for epoch in t_iter:
-        for i in range(sub_divisions):
-            emulator.train()
-            indices = torch.randint(0, n_pix, (n_pix // sub_divisions,)).to(device)
-            wl = wl_native[indices].to(device)
-            yhat = emulator.forward(wl)
-            loss = loss_fn(yhat, target[indices])
-            loss.backward()
-            optimizer.step()
-            optimizer.zero_grad()
-            losses.append(loss.item())
-            t_iter.set_description("Training Loss: {:0.8f}".format(loss.item()))
+t_iter = trange(n_epochs, desc="Training", leave=True)
+for epoch in t_iter:
+    for i in range(sub_divisions):
+        emulator.train()
[…]

### 2021-11-10 paper/paper1/ms.tex
Start writing prose, now on the Linux machine

@@
 
-At this stage we have a choice of whether to work in linear or log scale units.
+At this stage we have the choice of whether to work in linear or log scale flux units.
 Adopting the log of the flux would ensure that the cloned model possesses only positive flux values, a desireable trait of any physical spectral model.
 We chose instead to stick with native linear flux scaling because of the availability of convenient approximations to astrophysically relevant line profiles in the linear scaling domain.
-We normalize the spectrum by its mean value since we do not intend to apply these methods to absolutely flux calibrated spectra.
+We normalize the spectrum by its median value since we do not intend to apply these methods to absolutely flux calibrated spectra.
 
@@
 We initialize the cloned model with a series of preprocessing steps.
-We divide the entire spectrum by a black body of the same temperature as the model template.
-This step alone is usually adequate to move on to the spectral line finding.
+We divide the entire spectrum by a black body of the same effective temperature $T_{\mathrm{eff}}$ as the model template.
+The resulting signal typically possesses a smooth continuum variation, since
 

### 2021-11-10 src/blase/emulator.py
Tune the initialization based on observed scaling of lines from scipy heuristics

@@ -36,18 +36,23 @@ class PhoenixEmulator(nn.Module):
             self.wl_native, self.flux_native, prominence=prominence
         )
 
+        # Experimentally determined scale factors tweaks
+        amp_tweak = 0.14
+        sigma_width_tweak = 1.28
+        gamma_width_tweak = 1.52
+
         self.amplitudes = nn.Parameter(
-            torch.log(amplitudes).clone().detach().requires_grad_(True)
+            torch.log(amplitudes * amp_tweak).clone().detach().requires_grad_(True)
         )
         self.sigma_widths = nn.Parameter(
-            np.log(widths_angstroms / math.sqrt(2))
+            np.log(widths_angstroms / math.sqrt(2) * sigma_width_tweak)
             .clone()
             .detach()
             .requires_grad_(True)
         )
 
         self.gamma_widths = nn.Parameter(
-            np.log(widths_angstroms / math.sqrt(2))
+            np.log(widths_angstroms / math.sqrt(2) * gamma_width_tweak)
             .clone()
             .detach()
             .requires_grad_(True)
@@ -59,13 +64,13 @@ class PhoenixEmulator(nn.Module):
         )
 
         self.a_coeff = nn.Parameter(
-            torch.tensor(1.0, requires_grad=True, dtype=torch.float64)
+            torch.tensor(1.0, requires_grad=False, dtype=torch.float64)
         )
         self.b_coeff = nn.Parameter(
-            torch.tensor(0.0, requires_grad=True, dtype=torch.float64)
+            torch.tensor(0.0, requires_grad=False, dtype=torch.float64)
         )
         self.c_coeff = nn.Parameter(
-            torch.tensor(0.0, requires_grad=True, dtype=torch.float64)
+            torch.tensor(0.0, requires_grad=False, dtype=torch.float64)
         )
 
     def forward(self, wl):
@@ -76,7 +81,6 @@ class PhoenixEmulator(nn.Module):
         """
         # return self.product_of_lorentzian_model(wl)
         return self.product_of_pseudovoigt_model(wl)
-        # return self.sum_of_voigts_model(wl)
 
     def product_of_lorentzian_model(self, wl):
         """Return the Lorentzian-only forward model, modulated by Blackbody and slopes"""

### 2021-11-10 examples/clone_PHOENIX_spectrum.py
Freeze the a, b, and c linear coeffs; target line cores; continuum normalize

@@ -9,8 +9,6 @@ from torch.utils.tensorboard import SummaryWriter
 import webbrowser
 import numpy as np
 
-# Toggle below if you just want to read in a pre-trained model
-redo_training = True
 if torch.cuda.is_available():
     device = "cuda"
 else:
@@ -53,13 +51,15 @@ webbrowser.open("http://localhost:6006/", new=2)
 spectrum = PHOENIXSpectrum(teff=4700, logg=4.5)
 spectrum = spectrum.divide_by_blackbody()
 spectrum = spectrum.normalize()
+continuum_fit = spectrum.fit_continuum(polyorder=5)
+spectrum = spectrum.divide(continuum_fit, handle_meta="ff")
 
 # Numpy arrays: 1 x N_pix
 wl_native = spectrum.wavelength.value
 flux_native = spectrum.flux.value
 
 # Create the emulator
-emulator = PhoenixEmulator(wl_native, flux_native, prominence=0.05)
+emulator = PhoenixEmulator(wl_native, flux_native, prominence=0.02)
 emulator.to(device)
 
 n_pix = len(emulator.wl_native)
@@ -67,17 +67,33 @@ wl_native = emulator.wl_native.clone().detach().to(device)
 target = emulator.flux_native.clone().detach().to(device)
 
 # Training Loop
+emulator.lam_centers.requires_grad = False
+emulator.a_coeff.requires_grad = False
+emulator.b_coeff.requires_grad = False
+emulator.c_coeff.requires_grad = False
+
 loss_fn = nn.MSELoss(reduction="mean")
-optimizer = optim.Adam(emulator.parameters(), 0.004)
-n_epochs = 300
-sub_divisions = 20
+optimizer = optim.Adam(
+    filter(lambda p: p.requires_grad, emulator.parameters()), 0.004, amsgrad=True
+)
+n_epochs = 100
+sub_divisions = 30
 losses = []
+
+# Experiment: just optimize the line cores:
+indices = np.isin(emulator.wl_native.cpu(), emulator.lam_centers.detach().cpu())
+# from scipy.ndimage import binary_dilation
+# dilated_inds = binary_dilation(indices, iterations=2)
+peak_indices = torch.tensor(indices).to(device)
 
 t_iter = trange(n_epochs, desc="Training", leave=True)
 for epoch in t_iter:
     for i in range(sub_divisions):
         emulator.train()
-        indices = torch.randint(0, n_pix, (n_pix // sub_divisions,)).to(device)
[…]

### 2021-11-10 paper/paper1/ms.tex
writing prose

@@
 We start with a high resolution pre-computed stellar model spectrum, $\mathsf{M}(\bm{\lambda})$.
-This spectrum should have sporadic regions of discernable continuum devoid of lines, and the continuum must vary smoothly in wavelength.
+The procedure is largely agnostic to the exact details of how this spectrum was made.
+For the purposes of this paper, we will showcase examples from two well-known families of precomputed synthetic spectra: \texttt{PHOENIX} \citep{husser13} for stellar spectra $(T_{\mathrm{eff}}\in [2300, 10000]\;K)$ and \texttt{Sonora} \citep{2021ApJ...920...85M} for brown dwarf spectra $(T_{\mathrm{eff}}\in [1000, 2300]\;K)$.
+We place the following demands on the the precomputed spectra.
+They should have sporadic regions of discernable continuum devoid of lines, and the continuum must vary smoothly in wavelength.
+The spectral lines or pseudo-lines should be resolved, and not sub-sampled.
 We suspect most stellar spectra meet these criteria, with the exception of the coolest M dwarfs and brown dwarfs.
 The method can hypothetically handle spectra without discernable continua, but we suspect some modifications to the preprocessing steps would be needed.
-We allow the spectrum to have any number of spectral lines.
+We hypothetically allow the spectrum to have any number of spectral lines, though in practice there are computational tradeoffs as the number of lines grows.
 We truncate the red and blue limits of the precomputed synthetic spectrum to match a high-bandwidth echelle spectrograph, extended with a buffer at the edges of size $\pm \Delta \lambda_{\mathrm{buffer}}$, chosen to account for plausible radial velocity and rotational broadening of real stars.
@@
 We divide the entire spectrum by a black body of the same effective temperature $T_{\mathrm{eff}}$ as the model template.
-The resulting signal typically possesses a smooth continuum variation, since
+The resulting signal typically possesses a smooth continuum variation.
+An additional continuum flattening step ensures that subsequent spectral line finding steps get applied uniformly.
+Here, we identify the continuum regions as the most prominent peaks in the specturm separated by a pixel distance of at least $15000-90000$ pixels from adjacent peaks, yielding between 60 and 5 points.
+We then fit an $n_{\mathrm{poly}} = 3^{rd}-5^{th}$ order polynomial to these peaks, evaluate the polynomial on the entire bandwidth, and divide the spectrum by this trend.
+Visual inspection of this continuum-flattened spectrum typically possesses variations at the $<0.2\%$ level.
+Tuning the pixel distance and polynomial order can lower the undulations to $<0.05\%$ level for some input spectra.
 
-Occasionally a more advanced continuum flattening step may be needed to ensure that subsequent spectral line finding steps get applied uniformly.
-Here, we apply a low-pass Savitzky-Golay filter\authorcomment1{citation XX} to identify the continuum regions, and then fit a low-order polynomial to this pre-filtered continuum spectrum.
-We evaluate the low-order polynomial at all wavelength points, multiplied by the blackbody curve to assemble a ``perturbed black body'' that symbolizes the black body modulated by any broad-band continuum opacity that may reside in the model.
-We define a threshold flux loss of $1\%$, below which any lines are masked, erode that mask by $N_{\mathrm{iter}}=13$ pixels, and repeat the polynomial fitting process again.
-At this stage we fit a high order polynomial, with the polynomial order set through cross validation, typically $n_{\mathrm{poly}}\sim11$.
-The ``perturbed black body'' continuum model $\mathsf{B}$ contains  $n_{\mathrm{poly}}+1$ free parameters.
+We emphasize that a recreation of the unvarnished input spectrum can be obtained by multiplying by the blackbody curve times this polynomial shape, yielding the ``perturbed black body'', $\mathsf{B}(\bm{\lambda})$, that symbolizes the black body modulated by continuum opacity or broad-band radiative transfer effects.
+This smooth spectrum may be useful for applications that need to keep track of broad-band flux, such as low resolution, or molecular bandheads.
+The ``perturbed black body'' continuum model $\mathsf{B}$ contains  $n_{\mathrm{poly}}+1$ fixed lookup parameters, plus the input $T_{\mathrm{eff}}$.
 
-In either continuum flattening case, we move on to find locations of spectral lines.
-We apply a local minimum finding algorithm on the spectrum by defining a prominence threshold $\delta f\sim0.01$.
-Testing demo with the real fixed static residual.
+We apply a local-minimum-finding algorithm on the spectrum by defining a prominence threshold $P_{rom} \in (0.005, 0.1)$.
+This threshold dictates the number of lines that will be modeled: a lower prominence finding more, smaller lines, and a larger prominence finding fewer, deeper lines.
+The prominence algorithm successfully finds lines that reside on top of broad line wings, provided that they exceed the prominence threshold in their local region.
+Figure XX shows how
 

### 2021-11-10 paper/paper1/ms.tex
Writing prose for the methodology

@@
 The method can hypothetically handle spectra without discernable continua, but we suspect some modifications to the preprocessing steps would be needed.
-We hypothetically allow the spectrum to have any number of spectral lines, though in practice there are computational tradeoffs as the number of lines grows.
 We truncate the red and blue limits of the precomputed synthetic spectrum to match a high-bandwidth echelle spectrograph, extended with a buffer at the edges of size $\pm \Delta \lambda_{\mathrm{buffer}}$, chosen to account for plausible radial velocity and rotational broadening of real stars.
@@
 The choice of limiting the bandwidth to a region-of-interest around a single echelle spectrograph bandwidth stems from computational constraints.
-In principle there is no fundamental limit to the bandwidth one could clone with this method, up to and including the entire precomputed synthetic spectral model bandwidth.
-
+In principle there is no fundamental limit to the bandwidth one could clone with the method presented here, up to and including the entire precomputed synthetic spectral model bandwidth.
 We adopt the exact native wavelength sampling with no smoothing or interpolation, yielding a wavelength vector $\bm{\lambda}$ with length $N_\mathrm{pixels}$ equal to the number of pixels within the extents of our region of interest including the buffers.
@@
 
-We emphasize that a recreation of the unvarnished input spectrum can be obtained by multiplying by the blackbody curve times this polynomial shape, yielding the ``perturbed black body'', $\mathsf{B}(\bm{\lambda})$, that symbolizes the black body modulated by continuum opacity or broad-band radiative transfer effects.
-This smooth spectrum may be useful for applications that need to keep track of broad-band flux, such as low resolution, or molecular bandheads.
+We emphasize that a recreation of the unvarnished input spectrum can be obtained by multiplying the continuum-flattened signal by the ``perturbed black body'', $\mathsf{B}(\bm{\lambda})$, that symbolizes the black body modulated by continuum opacity or broad-band radiative transfer effects.
+This smooth spectrum may be useful for applications that need to keep track of broad-band flux, such as low resolution spectrographs, or regions with molecular bandheads.
 The ``perturbed black body'' continuum model $\mathsf{B}$ contains  $n_{\mathrm{poly}}+1$ fixed lookup parameters, plus the input $T_{\mathrm{eff}}$.
@@
 We apply a local-minimum-finding algorithm on the spectrum by defining a prominence threshold $P_{rom} \in (0.005, 0.1)$.
-This threshold dictates the number of lines that will be modeled: a lower prominence finding more, smaller lines, and a larger prominence finding fewer, deeper lines.
-The prominence algorithm successfully finds lines that reside on top of broad line wings, provided that they exceed the prominence threshold in their local region.
-Figure XX shows how
+This threshold dictates the number of lines that will be modeled: a lower prominence finds more, smaller lines, and a larger prominence finds fewer, deeper lines.
+The prominence algorithm successfully finds lines that reside on top of broad line wings, or unresolved bandheads provided that the individual lines exceed the prominence threshold in their local region.
+The number of lines $N_{\mathrm{lines}}$ depends on the bandwidth, prominence, and the intrinsic properties of the input spectrum, principally effective temperature and metallicity.
+
+For this paper, we illustrate examples for two \'echelle spectrographs with particularly large spectral grasp: the Habitable Zone Planet Finder \citep[HPF,][]{2014SPIE.9147E..1GM} on the Hobby-Eberly Telescope at McDonald Observatory in Fort Davis, Texas; and the Immersion Grating Infrared Spectrograph \citep[IGRINS,][]{park14} currently on the Gemini South Telescope on Cerro Pach\'on in Chile.
+The $R=55,000$ HPF has a native bandwidth of $x-y$, that we expand to $x-y^*$ including the edge buffers.
+IGRINS has two cameras for $H$ and $K$ band, with the combined spectrum spanning $z-y$ including the edge buffers and the region in-between the two cameras, all at a resolving power of $(R=45,000)$.
+The spectrograph acquisition, reduction, and post-processing steps yield data $\mathsf{D}(\bm{\lambda}_{d})$, where $\bm{\lambda}_{d}$ is the wavelength vector at the instrumental resolution and sampling of each instrument, generally much coarser than the resolution and sampling grid of the precomputed synthetic spectra.
+The data wavelength vector also contains pesky gaps between \'echelle orders, whereas the precomputed grid is usually contiguous.
+
+Figure XX shows how the number of detected lines $N_{\mathrm{lines}}$ scales with effective temperature and prominence threshold for the \texttt{PHOENIX} grid, truncated to the bandwidths-plus-buffers for HPF and IGRINS.
 

### 2021-11-10 paper/paper1/ms.tex
writing prose about the method, continued!

@@
 
-We start with a high resolution pre-computed stellar model spectrum, $\mathsf{M}(\bm{\lambda})$.
+We start with a high resolution pre-computed synthetic stellar model spectrum, $\mathsf{S}(\bm{\lambda})$.
 The procedure is largely agnostic to the exact details of how this spectrum was made.
@@
 In principle there is no fundamental limit to the bandwidth one could clone with the method presented here, up to and including the entire precomputed synthetic spectral model bandwidth.
-We adopt the exact native wavelength sampling with no smoothing or interpolation, yielding a wavelength vector $\bm{\lambda}$ with length $N_\mathrm{pixels}$ equal to the number of pixels within the extents of our region of interest including the buffers.
+We adopt the exact native wavelength sampling with no smoothing or interpolation, yielding a wavelength vector $\bm{\lambda}_s$ with length $N_\mathrm{s}$ equal to the number of pixels within the extents of our region of interest including the buffers.
 
@@
 Tuning the pixel distance and polynomial order can lower the undulations to $<0.05\%$ level for some input spectra.
+The high-pass filtering step should be set to capture the genuine spectral shape, without over-fitting broad line-wings such as those in deep Hydrogen and sodium lines.
+We have experimentally found it helpful to hand-tune models of exceptionally broad and non-standard line-wings with an optional pre-processing step, not described here.
 
-We emphasize that a recreation of the unvarnished input spectrum can be obtained by multiplying the continuum-flattened signal by the ``perturbed black body'', $\mathsf{B}(\bm{\lambda})$, that symbolizes the black body modulated by continuum opacity or broad-band radiative transfer effects.
+We emphasize that a recreation of the unvarnished input spectrum can be obtained by multiplying the continuum-flattened signal by the ``perturbed black body'', $\mathsf{B}(\bm{\lambda}_s)$, that symbolizes the black body modulated by continuum opacity or broad-band radiative transfer effects.
 This smooth spectrum may be useful for applications that need to keep track of broad-band flux, such as low resolution spectrographs, or regions with molecular bandheads.
@@
 
+Next we identify the spectral lines.
 We apply a local-minimum-finding algorithm on the spectrum by defining a prominence threshold $P_{rom} \in (0.005, 0.1)$.
-This threshold dictates the number of lines that will be modeled: a lower prominence finds more, smaller lines, and a larger prominence finds fewer, deeper lines.
+This threshold dictates the number of lines that will be modeled: a lower prominence finds more, weaker lines, and a larger prominence finds fewer, deeper lines.
 The prominence algorithm successfully finds lines that reside on top of broad line wings, or unresolved bandheads provided that the individual lines exceed the prominence threshold in their local region.
@@
 
-For this paper, we illustrate examples for two \'echelle spectrographs with particularly large spectral grasp: the Habitable Zone Planet Finder \citep[HPF,][]{2014SPIE.9147E..1GM} on the Hobby-Eberly Telescope at McDonald Observatory in Fort Davis, Texas; and the Immersion Grating Infrared Spectrograph \citep[IGRINS,][]{park14} currently on the Gemini South Telescope on Cerro Pach\'on in Chile.
+For this paper, we illustrate examples for two \'echelle spectrographs with particularly large spectral grasp: the Habitable Zone Planet Finder \citep[HPF,][]{2014SPIE.9147E..1GM} on the Hobby-Eberly Telescope at McDonald Observatory in Fort Davis, Texas; and the Immersion Grating Infrared Spectrograph \citep[IGRINS,][]{park14}
+currently on the Gemini South Telescope on Cerro Pach\'on in Chile.
 The $R=55,000$ HPF has a native bandwidth of $x-y$, that we expand to $x-y^*$ including the edge buffers.
@@
 The spectrograph acquisition, reduction, and post-processing steps yield data $\mathsf{D}(\bm{\lambda}_{d})$, where $\bm{\lambda}_{d}$ is the wavelength vector at the instrumental resolution and sampling of each instrument, generally much coarser than the resolution and sampling grid of the precomputed synthetic spectra.
-The data wavelength vector also contains pesky gaps between \'echelle orders, whereas the precomputed grid is usually contiguous.
+The data wavelength vector may also contains gaps between \'echelle orders, whereas the precomputed grid is usually contiguous.
+HPF may have up to $2048\times28=57,344$ pixels, and IGRINS has typically about $75,000$ pixels, after common truncation of noisy edge pixels.
 
-Figure XX shows how the number of detected lines $N_{\mathrm{lines}}$ scales with effective temperature and prominence threshold for the \texttt{PHOENIX} grid, truncated to the bandwidths-plus-buffers for HPF and IGRINS.
+Figure XX shows how the number of detected lines $N_{\mathrm{lines}}$ scales with effective temperature and prominence threshold $P_{rom}$ for the \texttt{PHOENIX} grid, truncated to the bandwidths-plus-buffers for HPF and IGRINS.
+The HPF-truncated spectra have $N_s=XX$ and the IGRINS-truncated spectra have $N_s=XX$.
 
-\begin{outline}
-  \1 Define a prominence threshold and find locations of peaks above that threshold
-  \2 Derive coarse properties about those peaks: widths and amplitudes
-  \1 We now have a coarse model that needs to be fine-tuned with autodiff
-\end{outline}
+So far we have only one piece of information about the peaks: their location.
+Next, we derive coarse properties about each detected peak: its amplitude and width, again using the prominence algorithms implemented in \texttt{scipy} \citep{2020SciPy-NMeth}.
 
-The low-pass filter should be set to capture the genuine spectral shape, without over-fitting broad line-wings such as those in deep Hydrogen and sodium lines.
-We have experimentally found it helpful to hand-tune models of exceptionally broad and non-standard line-wings with with an optional pre-processing step.
+There does not exist a general-purpose, single-shot algorithm for obtaining the lineshape in the presence of overlapping spectral lines: where do the wings of one line begin and the wings of another adjacent line end?
+We therefore do not attempt to determine anything about the lineshape at this stage, and instead assume that the lines resemble a Voigt profile, with a guess width about equally split between Lorentzian and Gaussian.
+
+We have now arrived at the \emph{blas\'e} clone model $\mathsf{M}(\bf{\lambda}_s)$ for a flattened synthetic spectrum $\mathsf{S}$: it is the cumulative product of transmission through the sea of all overlapping spectral lines:
+
+\begin{eqnarray}
+  \mathsf{M}(\bm{\lambda}_s) = {\displaystyle \prod_{j=1}^{N_{\mathrm{lines}}}  (1-a_j V(\bm{\lambda}_s-\lambda_{c,j}, \sigma_j, \gamma_j) )}
+\end{eqnarray}
+
+where $V(\lambda, \sigma_j, \gamma_j)$ is the normalized Voigt profile with Gaussian standard deviation $\sigma$, Lorentzian half-width $\gamma$, at line center position $\lambda_c$, for the $j^{th}$ spectral line.
+The amplitude $a$ is equal to the area under the curve of the line.
+
+The model evaluated with its coarse initial values would have terrible performance: it would only vaguely resemble the synthetic spectral model, with up to $\pm 50\%$ undulations from the inexact assignment of widths,lineshapes, and amplitudes.
+Instead, we tune the parameters of the model, starting from these coarse initial values.
+This model has $N_{\mathrm{lines}}\times 3$ free parameters, where the center wavelength is held fixed and the amplitude, width, and lineshape are allowed to vary.
+
+As seen in Figure XX, the number of lines can exceed $XX$ depending on the
 

### 2021-11-10 paper/paper1/ms.tex
Writing equations in the paper

@@
 \usepackage{censor}
+\usepackage{mathtools}
 
@@
 where $V(\lambda, \sigma_j, \gamma_j)$ is the normalized Voigt profile with Gaussian standard deviation $\sigma$, Lorentzian half-width $\gamma$, at line center position $\lambda_c$, for the $j^{th}$ spectral line.
-The amplitude $a$ is equal to the area under the curve of the line.
+The amplitude $a$ is always expected to be positive for absorption lines, and is equal to the area under the curve of the line since the Voigt profile is normalized.
 
@@
 This model has $N_{\mathrm{lines}}\times 3$ free parameters, where the center wavelength is held fixed and the amplitude, width, and lineshape are allowed to vary.
+We minimize a scalar loss function, which we define as the sum of the squares of the residual vector, $\mathsf{R} \equiv \mathsf{S}-\mathsf{M}$:
 
-As seen in Figure XX, the number of lines can exceed $XX$ depending on the
+\begin{eqnarray}
+  \mathcal{L} = \sum_i^{N_s} (S_i - M_i)^2 = \mathsf{R^\intercal}\cdot \mathsf{R}
+\end{eqnarray}
+
+As seen in Figure XX, the number of lines can exceed $XX$, meaning the model has over $XX \times 3 = ZZ$ free parameters.
+Fitting that large number of parameters is difficult with conventional optimizers, which struggle to converge after hundreds of parameters\authorcomment1{Citation to limits of optimizers}.
+
+Here we employ a variant of Stochastic Gradient Descent, an optimization technique that can scale to a virtually unlimited number of parameters\authorcomment1{provide a citation}.
+This technique computes the derivative of the loss scalar with respect to each of the parameters, so for example %\frac{\partial \mathcal{L}}{\partial a}
 

### 2021-11-10 paper/paper1/ms.tex
Some prose about autodiff and backprop

@@
 This model has $N_{\mathrm{lines}}\times 3$ free parameters, where the center wavelength is held fixed and the amplitude, width, and lineshape are allowed to vary.
-We minimize a scalar loss function, which we define as the sum of the squares of the residual vector, $\mathsf{R} \equiv \mathsf{S}-\mathsf{M}$:
+We minimize a scalar loss function, which we define as the mean squared error (MSE), proportional to $\chi^2$, the sum of the squares of the residual vector, $\mathsf{R} \equiv \mathsf{S}-\mathsf{M}$:
 
@@
 
-Here we employ a variant of Stochastic Gradient Descent, an optimization technique that can scale to a virtually unlimited number of parameters\authorcomment1{provide a citation}.
-This technique computes the derivative of the loss scalar with respect to each of the parameters, so for example %\frac{\partial \mathcal{L}}{\partial a}
+Here we employ a variant of Stochastic Gradient Descent (SGD), an optimization technique that can scale to a virtually unlimited number of parameters\authorcomment1{provide a citation}.
+This technique computes the derivative of the loss scalar with respect to each of the parameters, the Jacobian: $(\frac{\partial \mathcal{L}}{\partial a_j}, \frac{\partial \mathcal{L}}{\partial \sigma_j}, \frac{\partial \mathcal{L}}{\partial \gamma_j})$.
+The Jacobian indicates how the MSE would decrease with a change in the parameter-of-interest, or put simply which way and by how much you have to change the line properties to make a better fit.
+
+The optimizer updates the $a_j, \sigma_j, \gamma_j$ parameters by a small fraction of the Jacobian---called the learning rate LR---towards the direction that would improve the fit, for all $N_{\mathrm{lines}} \times 3$ parameters simultaneously.
+The Jacobian is calculated behind-the-scenes with automatic differentiation\authorcomment1{Citation} implemented as the so-called backwards-propagation algorithm or simply ``back-prop''.
+We choose the \texttt{PyTorch} framework that computes these Jacobians efficiently for all of the mathematical primatives in our \texttt{blas\'e} implementation.
+
+In principle we could compute the Jacobians analytically---since the partial derivatives of the Voigt function are known in closed-form---circumventing the vendor lock-in and memory overhead of an autodiff implementation with PyTorch.
+In practice, PyTorch also offers hardware acceleration (on GPU or TPU)\authorcomment1{citation needed}, mature support for a range of optimizers, and other perks that motivated its choice over analytic implementations or other machine learning frameworks.
+Still, the performance of an analytic implementation over the one presented here represents an interesting avenue of future research.
 

### 2021-11-10 paper/paper1/ms.tex
Significant prose writing today, done for the day/moment

@@
 
+\subsection{The \emph{blas\'e} clone model}
+
 We have now arrived at the \emph{blas\'e} clone model $\mathsf{M}(\bf{\lambda}_s)$ for a flattened synthetic spectrum $\mathsf{S}$: it is the cumulative product of transmission through the sea of all overlapping spectral lines:
@@
 \begin{eqnarray}
-  \mathsf{M}(\bm{\lambda}_s) = {\displaystyle \prod_{j=1}^{N_{\mathrm{lines}}}  (1-a_j V(\bm{\lambda}_s-\lambda_{c,j}, \sigma_j, \gamma_j) )}
+  \mathsf{M}(\bm{\lambda}_s) = {\displaystyle \prod_{j=1}^{N_{\mathrm{lines}}}  (1-a_j V(\bm{\lambda}_s-\lambda_{c,j}, \sigma_j, \gamma_j) )} \label{equation1}
 \end{eqnarray}
@@
 
-The model evaluated with its coarse initial values would have terrible performance: it would only vaguely resemble the synthetic spectral model, with up to $\pm 50\%$ undulations from the inexact assignment of widths,lineshapes, and amplitudes.
+The model evaluated with its coarse initial values would have terrible performance: it would only vaguely resemble the synthetic spectral model, with up to $\pm 50\%$ undulations from the inexact assignment of widths, lineshapes, and amplitudes.
 Instead, we tune the parameters of the model, starting from these coarse initial values.
 This model has $N_{\mathrm{lines}}\times 3$ free parameters, where the center wavelength is held fixed and the amplitude, width, and lineshape are allowed to vary.
-We minimize a scalar loss function, which we define as the mean squared error (MSE), proportional to $\chi^2$, the sum of the squares of the residual vector, $\mathsf{R} \equiv \mathsf{S}-\mathsf{M}$:
+We minimize a scalar ``goodness-of-fit'' metric, \emph{aka}loss scalar $\mathcal{L}$, chosen as the mean squared error (MSE), which is proportional to $\chi^2$, the sum of the squares of the residual vector $\mathsf{R} \equiv \mathsf{S}-\mathsf{M}$ but has no notion of per-pixel noise, since the precomputed synthetic spectrum has no uncertainty:
 
@@
 This technique computes the derivative of the loss scalar with respect to each of the parameters, the Jacobian: $(\frac{\partial \mathcal{L}}{\partial a_j}, \frac{\partial \mathcal{L}}{\partial \sigma_j}, \frac{\partial \mathcal{L}}{\partial \gamma_j})$.
-The Jacobian indicates how the MSE would decrease with a change in the parameter-of-interest, or put simply which way and by how much you have to change the line properties to make a better fit.
+The Jacobian indicates how the MSE would decrease with a change in the parameter-of-interest, or put simply which-way and by-how-much you have to change the line properties to get a better fit.
 
@@
 In principle we could compute the Jacobians analytically---since the partial derivatives of the Voigt function are known in closed-form---circumventing the vendor lock-in and memory overhead of an autodiff implementation with PyTorch.
+Futher, an autodiff-aware implementation of the \emph{celerit\'e} algorithm\authorcomment1{add DFM citation} does not yet exist in PyTorch, constricting our choices for simultaneous continuum fitting.
 In practice, PyTorch also offers hardware acceleration (on GPU or TPU)\authorcomment1{citation needed}, mature support for a range of optimizers, and other perks that motivated its choice over analytic implementations or other machine learning frameworks.
-Still, the performance of an analytic implementation over the one presented here represents an interesting avenue of future research.
-
-\subsection{Continuum model}
-\begin{outline}
-  \1 How to describe the continuum:
-  \2 Black body / Rayleigh Jeans plus perturbation
-  \2 High order polynomial: Chebyshev? Somewhere between 3rd and 20th order?
-  \2 Use cross validation?
-  \1 GPs are also possible, but high Number of pixels--> need scalable GP
-  \2 Celerite not implemented for PyTorch?
-  \2 GPyTorch possible, but has certain restrictions on kernel and sampling
-\end{outline}
-
-\subsection{Lineshapes choice and tradeoffs}
-\begin{outline}
-  \1 Faced with a question: what lineshape to use
-  \2 To First order the lines are Gaussian or Lorentzian
-  \2 To Second order the lines are Voigt Profiles
-  \2 In detail the lines are the result of radiative transfer that can smear the Voigt profile
-  \2 In practice the high-res lines get convolved with instrumental profiles, so details don't matter too much
-  \1 Computational tradeoff of Lorentzian/Gaussian versus Pseudo-Voigt versus Voigt
-  \2 Additional Implementation challenge: Fadeeva function not implemented in PyTorch
-  \1 We choose a pseudo-Voigt as a balance between computation and adequate accuracy
-\end{outline}
+The GPyTorch framework\authorcomment1{add citation} may offer a workaround to the absence of a \emph{celerit\'e} implementation.
+Still, the performance of an analytic Jacobian implementation over the autodiff route presented here presents an interesting avenue of future research.
 
 \subsection{GPU and Autodiff specific considerations}
+We make a few tweaks to the implementation for numerical purposes.
+First, we want all the parameters to be positive, forbidding negative amplitudes and negative widths.
+We therefore tune the natural log of the parameters, and exponentiate them before inclusion in Equation \ref{equation1}.
+Second, we found through iterative experimentation that the initialization amplitudes and widths were systematically shifted from the optimized values.
+We built-in a three scalar tweaks to the initialization amplitudes, Gaussian width, and Lorentzian width, which dramatically accelerated the optimization step.
+
+We set the \texttt{requires\_grad=True} property for any Torch tensor that we want to vary.
+This allows us to easily explore whether, say, allowing the $\lambda_c$ parameter to vary significantly improves the fit.
+It also allows us to tune or freeze an experimental wavelength-dependent pre-factor to Equation \ref{equation1} to correct for lingering imperfections in the otherwise-fixed continuum flattening procedure.
+
+The computational bottleneck occurs at...
+
 \begin{outline}
@@
   \2 For the continuum = Nth order polynomial = N+1 parameters, or Gaussian Process = 2 hyperparameters
-  \1 Each Tunable parameter gets `grad=True`
   \1 We assemble the product of all lines and all continuum sources, this is the computational bottleneck
@@
 
+\subsection{Lineshapes choice and tradeoffs}
+\begin{outline}
+  \1 Faced with a question: what lineshape to use
+  \2 To First order the lines are Gaussian or Lorentzian
+  \2 To Second order the lines are Voigt Profiles
+  \2 In detail the lines are the result of radiative transfer that can smear the Voigt profile
+  \2 In practice the high-res lines get convolved with instrumental profiles, so details don't matter too much
+  \1 Computational tradeoff of Lorentzian/Gaussian versus Pseudo-Voigt versus Voigt
+  \2 Additional Implementation challenge: Fadeeva function not implemented in PyTorch
+  \1 We choose a pseudo-Voigt as a balance between computation and adequate accuracy
+\end{outline}
+
 \subsection{Questioning the assumptions of the method}
+
 \begin{outline}

### 2021-11-15 paper/paper1/ms.tex
Prose and equation writing

@@
 \begin{eqnarray}
-  \mathsf{M}(\bm{\lambda}_s) = {\displaystyle \prod_{j=1}^{N_{\mathrm{lines}}}  (1-a_j V(\bm{\lambda}_s-\lambda_{c,j}, \sigma_j, \gamma_j) )} \label{equation1}
+  \mathsf{M}(\bm{\lambda}_s) = {\displaystyle \prod_{j=1}^{N_{\mathrm{lines}}}  (1-a_j \mathsf{V}(\bm{\lambda}_s-\lambda_{c,j}, \sigma_j, \gamma_j) )} \label{equation1}
 \end{eqnarray}
 
-where $V(\lambda, \sigma_j, \gamma_j)$ is the normalized Voigt profile with Gaussian standard deviation $\sigma$, Lorentzian half-width $\gamma$, at line center position $\lambda_c$, for the $j^{th}$ spectral line.
+where $\mathsf{V}(\lambda, \sigma_j, \gamma_j)$ is the normalized Voigt profile with Gaussian standard deviation $\sigma$, Lorentzian half-width $\gamma$, at line center position $\lambda_c$, for the $j^{th}$ spectral line.
 The amplitude $a$ is always expected to be positive for absorption lines, and is equal to the area under the curve of the line since the Voigt profile is normalized.
@@
 This model has $N_{\mathrm{lines}}\times 3$ free parameters, where the center wavelength is held fixed and the amplitude, width, and lineshape are allowed to vary.
-We minimize a scalar ``goodness-of-fit'' metric, \emph{aka}loss scalar $\mathcal{L}$, chosen as the mean squared error (MSE), which is proportional to $\chi^2$, the sum of the squares of the residual vector $\mathsf{R} \equiv \mathsf{S}-\mathsf{M}$ but has no notion of per-pixel noise, since the precomputed synthetic spectrum has no uncertainty:
+We minimize a scalar ``goodness-of-fit'' metric, \emph{aka} loss scalar $\mathcal{L}$, chosen as the mean squared error (MSE), which is proportional to $\chi^2$, the sum of the squares of the residual vector $\mathsf{R} \equiv \mathsf{S}-\mathsf{M}$ but has no notion of per-pixel noise, since the precomputed synthetic spectrum has no uncertainty:
 
@@
 The GPyTorch framework\authorcomment1{add citation} may offer a workaround to the absence of a \emph{celerit\'e} implementation.
-Still, the performance of an analytic Jacobian implementation over the autodiff route presented here presents an interesting avenue of future research.
+Still, the performance of an analytic Jacobian implementation over the autodiff route presented here offers an interesting avenue of future research.
 
@@
 Second, we found through iterative experimentation that the initialization amplitudes and widths were systematically shifted from the optimized values.
-We built-in a three scalar tweaks to the initialization amplitudes, Gaussian width, and Lorentzian width, which dramatically accelerated the optimization step.
+We built-in scalar tweaks to the initialization amplitudes, Gaussian width, and Lorentzian width, which dramatically accelerated the optimization step.
 
@@
 
-The computational bottleneck occurs at...
+The computational bottleneck occurs at the evaluation of Equation \ref{equation1}, which includes the assembly of a $N_{\mathrm{lines}}\times N_{s}$ matrix $\bm{\bar{V}}$ assembled by stacking each Voigt profile $\mathsf{V}_j(\bm{\lambda}_s)$ on top of each other:
+
+\begin{equation}
+  \begin{pmatrix}
+    \mathsf{V}_1(\bm{\lambda}_s)                    & \\
+    \mathsf{V}_2(\bm{\lambda}_s)                    & \\
+    \vdots                                          & \\
+    \mathsf{V}_{N_{\mathrm{lines}}}(\bm{\lambda}_s) &
+  \end{pmatrix}
+\end{equation}
+
+$\mathsf{V}_j(\bm{\lambda}_s)$ matrix  .
 

### 2021-11-15 paper/paper1/ms.tex
Writing about minibatches and sparsity

@@
 
-$\mathsf{V}_j(\bm{\lambda}_s)$ matrix  .
+Equation \ref{equation1} performs a type of tensor contraction, turning a $N_{\mathrm{lines}}\times N_{s}$ matrix into a $1\times N_{s}$ vector.
+The number of Floating Point Operations (FLOPS) scales with the number of entries in this matrix, with a scalar prefactor for the cost of evaluating a single line profile at a single wavelength point.
+Efficient GPU algorithms exist for tensor contractions such as this product along an axis of a matrix, allowing this computation to proceed quickly on modern machines\authorcomment1{add citation?}.
+In particular, the proprietary CUDA architecture for NVIDIA\textsuperscript{\tiny\textregistered} GPUs contains Tensor cores with specialized matrix math, although the algorithms place restrictions on floating precision depending on the particular hardware and software available.
+The chief bottleneck occurs when the storage of the $\bm{\bar{V}}$ matrix exceeds the available RAM of a GPU or CPU: the computation will fail with an ``Out of Memory'' exception.
+Modern NVIDIA GPUs have $8-40$ GB of RAM.
+The memory bottleneck is even more pernicious than mere storage of the $\bm{\bar{V}}$ matrix, since the CPU/GPU also has to store the computation graph for autodiff.
+The computation graph takes up more memory than the matrix itself, so it is generally not possible to evaluate Equation \ref{equation1} in its entirety in one-fell-swoop.
+A remedy is needed.
+
+Minibatches offer a natural solution to the line-by-line computational
+bottleneck\footnote{The Machine Learning and Artificial Intelligence (ML/AI) community employs minibatches to handle similar situations, in which the dataset volume exceeds the available GPU RAM.
+Our configuration departs slightly from the typical ML/AI situation, since their models are usually ``small'' relative to their datasets.
+Here our model is ``big'' relative to our dataset.}.
+You can choose to assemble and evaluate only a portion of the matrix at a time.
+The choice to evaluate only a portion of lines at a time would mean the model is inaccurately evaluated at those points.
+Instead we choose to evaluate only a subset $N_{\mathrm{subset}}$ of the total pixels $N_s$, so that the model is exact at those points.
+All lines update at each glimpse of a minibatch, but many lines get little information about how the loss scalar changes for their parameters.
+The value of $N_{\mathrm{subset}}$ should be set just below the threshold at which the computation runs out of memory.
+At present, this threshold is determined experimentally.
+The indices of the wavelength points can either be set to be contiguous in blocks, or random with possibly large gaps between indices, or else with some other scheme.
+We chose random sampling with replacement, meaning that some pixels get revisited more often than others.
+
+Another solution to the computation problem could be to take advantage of the sparsity of the $\bm{\bar{V}}$ matrix: most of the entries are vanishingly close to zero.
+Ideally we would only evaluate the model close to the line center position, some number of line cuts $N_{\mathrm{cut}}=\alpha \cdot \mathrm{FWHM}$ away.
+These line cuts produce a speedup by a factor of $\frac{N_s}{N_{\mathrm{cut}}}$, which can exceed $100\times$ for wide bandwidth spectra.
+Furthermore, efficient algorithms for assembling coalescing sparse matrices exist in \texttt{PyTorch}.
+A few practical considerations make sparse matrices peskier to work with than their dense counterparts.
+First, the $\alpha$ hyperparameter should be set large enough that the truncation effect is not seen for the deepest lines.
+The $\alpha$ should be preselected before the training process.
+If the FWHM initialization was poor enough, the question may arise how autodiff training will cope with the poor performance of a model, leading to the prospect of strange and difficult to diagnose artifacts near spurious line-cut truncation artifacts.
+At the time of writing, we do not implement sparse matrices in \emph{blas\'e}, but we find them the most compelling route for transformative performance improvements.
 

### 2021-11-15 paper/paper1/ms.tex
Prose is mostly in place up to the performance section

@@
 
-Minibatches offer a natural solution to the line-by-line computational
-bottleneck\footnote{The Machine Learning and Artificial Intelligence (ML/AI) community employs minibatches to handle similar situations, in which the dataset volume exceeds the available GPU RAM.
+\subsection{Minibatches}
+
+Minibatches offer a natural solution to the line-by-line computational bottleneck while also providing optimization performance improvements.
+Minibatches act as a form of regularization, the principal source of stochasticity in the Stochastic Gradient Descent algorithm, which tends to have better convergence than full-batch Gradient Descent\authorcomment1{citation}.
+The Machine Learning and Artificial Intelligence (ML/AI) community employs minibatches to handle similar situations, in which the dataset volume exceeds the available GPU RAM.
 Our configuration departs slightly from the typical ML/AI situation, since their models are usually ``small'' relative to their datasets.
-Here our model is ``big'' relative to our dataset.}.
-You can choose to assemble and evaluate only a portion of the matrix at a time.
-The choice to evaluate only a portion of lines at a time would mean the model is inaccurately evaluated at those points.
-Instead we choose to evaluate only a subset $N_{\mathrm{subset}}$ of the total pixels $N_s$, so that the model is exact at those points.
-All lines update at each glimpse of a minibatch, but many lines get little information about how the loss scalar changes for their parameters.
-The value of $N_{\mathrm{subset}}$ should be set just below the threshold at which the computation runs out of memory.
+Here our model is ``big'' relative to our dataset.
+
+Here we choose to assemble and evaluate only a portion of the matrix at a time, in minibatches.
+The choice to evaluate only a portion of lines at a time would mean the model is inaccurately evaluated at all wavelength pixels.
+Instead we choose to evaluate all lines, but only on a subset $N_{\mathrm{batch}}$ of the total pixels $N_s$, so that the model can eventually converge to exact at those points.
+All lines are allowed to update at each glimpse of a minibatch, but many lines with cores far from minibatch pixels will provide only weak information about how the loss scalar changes for their parameters.
+The value of $N_{\mathrm{batch}}$ should be set just below the threshold at which the computation runs out of memory.
 At present, this threshold is determined experimentally.
-The indices of the wavelength points can either be set to be contiguous in blocks, or random with possibly large gaps between indices, or else with some other scheme.
-We chose random sampling with replacement, meaning that some pixels get revisited more often than others.
+The indices of the wavelength points can either be set to be contiguous in blocks, or random with possibly large gaps between indices, or else with some other scheme tailored to the most informative wavelength regions.
+We experimented with several techniques and settled on random sampling with replacement, meaning that some pixels get revisited more often than others, with some minibatches tailored only to the line cores.
+
+\subsection{Sparsity}
 
@@
 These line cuts produce a speedup by a factor of $\frac{N_s}{N_{\mathrm{cut}}}$, which can exceed $100\times$ for wide bandwidth spectra.
-Furthermore, efficient algorithms for assembling coalescing sparse matrices exist in \texttt{PyTorch}.
+Furthermore, efficient algorithms for assembling and coalescing sparse matrices exist in \texttt{PyTorch}.
 A few practical considerations make sparse matrices peskier to work with than their dense counterparts.
@@
 
-\begin{outline}
-  \1 Total number of parameters
-  \2 For the lines =  N-lines x 4 (lambda center, fwhm L, sigma G, Amplitude)
-  \2 For the continuum = Nth order polynomial = N+1 parameters, or Gaussian Process = 2 hyperparameters
-  \1 We assemble the product of all lines and all continuum sources, this is the computational bottleneck
-  \1 This becomes a matrix of size N-lines x N-wavelength points
-  \1 We take the element-wise product over the lines axis, contracting to 1 x N-wavelength points
-  \1 For N-lines = X and N-wavelength-points = Y, we expect X*Y*k primative operations.
-  \1 For N-lines = X and N-wavelength-points = Y, we expect X*Y*32 bits of memory => XX GB
-  \1 NVIDIA GPUs can handle 8 - 40 GB of memory depending on the model
-\end{outline}
+\subsection{Optimization and training}
 
-\subsection{Minibatches}
-\begin{outline}
-  \1 In the parlance of Machine Learning, we have a big model and small data
-  \1 Really our model is too big: most of the entries are zero, so we should use a sparse implementation
-  \1 If the model is too big to fit in memory, we have to train in mini-batches
-  \1 We hand-in some fraction f ~ 1/20th of the entire wavelength range as a mini-batch
-  \1 The model is only evaluated at that 5\% of pixels,  $N_{lines} \times N_{pix} \cdot 0.05$)
-  \1 Technically all lines are updated
-  \1 In practice, only lines that are sufficiently informed by those pixels will get large updates
-  \1 Causes some stochasticity--- Stochastic Gradient Descent "SGD"
-  \1 Analogous but different from the notion of minibatch in ML: not a distinct axis, like $N_{images}$
-  \1 We use random minibatches
-\end{outline}
+We use the Adam optimizer\authorcomment1{citation} with a learning rate $LR=0.004$ and all the defaults for \texttt{PyTorch v1.11}.
+We set the minibatch fraction to typically $f\sim1-5\%$.
+We defined the number of training epochs $N_{epoch}=100-1000$.
+The optimizer steps $N_{batch}$ times in each epoch, so that on average each epoch sees the entire dataset.
+The minibatching indices may be chosen with replacement so that some pixels are repeated and some are not seen in a given epoch.
+The user can optionally monitor a live view of the training progress with Tensorboard\authorcomment1{citation} to gain an intuition for the training efficiency.
 
-\subsection{Optimization and training}
-\begin{outline}
-  \1 Run the training:
-  \1 Set the optimizer: Adam with a LR=0.004
-  \1 Set the loss function: Chi-squared with no per-pixel uncertainty
-  \1 Set the "mini-batch" fraction f~1-5%
-  \1 Choice of epochs or steps $N_steps = 300-10000$
-  \1 The training process sees 20 minibatches per epoch, sees $N_pixels$, with some repeats some misses
-  \1 Monitor with Tensorboard: see the improvement
-  \1 The training takes X minutes on an RTX2070 GPU with PyTorch v1.11.x, CUDA v11.x, and Intel X CPU
-  \1 We save the model parameters, refer to them collectively as the pre-trained model
-\end{outline}
+Figure XX shows a representative Loss curve (MSE versus epochs) for cloning the model in Figure ZZ.
+The NN epochs of training took X minutes on an RTX2070 GPU with PyTorch v1.11.x, CUDA v11.x, and Intel X CPU, with all tensors as FP64.
+That benchmark works out to X seconds per epoch, or about Z seconds per minibatch.
+
+We save the model parameters with \texttt{torch.save()} and refer to this collection of parameters as the \emph{pre-trained model}.
 

### 2021-11-15 paper/paper1/ms.tex
start the results section, some placeholder text

@@
 
-\begin{outline}
-  \1 We compute the residual of native PHOENIX minus cloned model
-  \1 We see X percent residuals at native resolution
-  \1 We smooth the residual to HPF instrumental resolution and see Y percent residuals
-  \1 The main residuals come from missing line opacity due to our prominence threshold:
-  \1 Including smaller prominence lines means smaller residuals, but more lines and higher computational cost
-\end{outline}
+We compute the residual $\mathsf{R}(\bm{\lambda}_s)$ of native PHOENIX minus cloned model, illustrated in the bottom panel of Figure XX\authorcomment1{make this figure}.
+We see an RMS of $\blackout{1}\%$ residuals at native resolution, with a long tail of residuals up to $\blackout{20}\%$.
+The greatest residuals surround the deepest and broadest lines---such as Hydrogen and neutral alkali metal lines---for which the non-Voigt lineshapes stand out.
+Figure ZZ\authorcomment1{make this figure} zooms in on the XX\authorcomment1{todo} line at $\censor{10200}\;$\AA.
+
+The other main source of large residuals is simply missing line opacity due to our finite prominence threshold.
+Lines with prominence less than $P_{rom}$ yield residuals that scale with $P_{rom}$.
+Including smaller prominence lines means smaller residuals, but more lines and higher computational cost.
 
 \subsection{Line data clustering and conspicuous flaws}
+We see some narrow lines devolve into missing continuum opacity, the favored tradeoff when the continuum estimate's poor performance outweighs the the pain of a narrow-but-tolerably-small spike.
+This problem can be seen in Figure XX where a line initialized at $\censor{10210}\;$\AA~ with an amplitude of ZZ and width of YY bloats to \censor{some other value}.
+These devolved lines tend to appear poorly immitated line wings.
+One can mitigate these devolved lines by adding more flexibility to the continuum or supporting more advanced line shapes.
+
 \begin{outline}
   \1 We see some lines that devolve into missing continuum opacity
-  \1 Especially acute near broad line wings if you don't handle the non-standard line shapes
   \1 How do the lineshapes cluster in $\sigma_G$ vs $FWHM_L$?

### 2021-11-16 paper/paper1/ms.tex
Missing word in the tex
Add placeholders for figures and pseudoVoigt subsection

@@
 \usepackage{mathtools}
-
 \usepackage{outlines}
@@
 
-Figure XX shows how the number of detected lines $N_{\mathrm{lines}}$ scales with effective temperature and prominence threshold $P_{rom}$ for the \texttt{PHOENIX} grid, truncated to the bandwidths-plus-buffers for HPF and IGRINS.
+\begin{figure}[hbt!]
+  \centering
+  \includegraphics[width=0.5\textwidth]{example-image-a}
+  \caption{Number of prominent spectral lines versus effective temperature for PHOENIX models truncated to HPF (\emph{left}) and IGRINS (\emph{right}) bandwidths, for different prominence thresholds of $0.05$ and $0.01$.}
+  \label{fig_Nlines_vs_teff}
+\end{figure}
+
+Figure \ref{fig_Nlines_vs_teff} shows how the number of detected lines $N_{\mathrm{lines}}$ scales with effective temperature and prominence threshold $P_{rom}$ for the \texttt{PHOENIX} grid, truncated to the bandwidths-plus-buffers for HPF and IGRINS.
 The HPF-truncated spectra have $N_s=XX$ and the IGRINS-truncated spectra have $N_s=XX$.
@@
 
+\subsection{Voigt profile evaluation}
+The Voigt profile $\mathsf{V}(\lambda, \sigma_j, \gamma_j)$ can be computed in exact closed form using the Fadeeva function\authorcomment1{citation}.
+Evaluation of the Fadeeva function can be computationally costly, and so approximate forms may be desirable.
+Here we adopt the pseudo-Voigt approximation described in \censor{some paper}\authorcomment1{look up reference to PseudoVoigt}.
+
+Two sampled Voigt approaches could be feasible.
+Direct convolution of a sampled Gaussian and Lorentzian profile may be too costly and inexact if pixels are coarsely sampled.
+Analogously, the convolution can be achieved in the Fourier domain through multiplication of the Fast Fourier Transforms of the Gaussian and Lorentzian profiles, followed by the subsequent Inverse Fourier Transforms.
+The main rationale for either of these two choices would be to anticipate and align with planned convolutions such as rotational broadening and instrumental kernels.
+We do not currently implement these sampled convolutions in \emph{blas\'e} due to their prohibitive computational cost.
+
+\subsection{Goodness of fit metric}
 The model evaluated with its coarse initial values would have terrible performance: it would only vaguely resemble the synthetic spectral model, with up to $\pm 50\%$ undulations from the inexact assignment of widths, lineshapes, and amplitudes.
@@
 
-As seen in Figure XX, the number of lines can exceed $XX$, meaning the model has over $XX \times 3 = ZZ$ free parameters.
+As seen in Figure \ref{fig_Nlines_vs_teff}, the number of lines can exceed $XX$, meaning the model has over $XX \times 3 = ZZ$ free parameters.
 Fitting that large number of parameters is difficult with conventional optimizers, which struggle to converge after hundreds of parameters\authorcomment1{Citation to limits of optimizers}.
@@
 In principle we could compute the Jacobians analytically---since the partial derivatives of the Voigt function are known in closed-form---circumventing the vendor lock-in and memory overhead of an autodiff implementation with PyTorch.
-Futher, an autodiff-aware implementation of the \emph{celerit\'e} algorithm\authorcomment1{add DFM citation} does not yet exist in PyTorch, constricting our choices for simultaneous continuum fitting.
+Further, an autodiff-aware implementation of the \emph{celerit\'e} algorithm\authorcomment1{add DFM citation} does not yet exist in PyTorch, constricting our choices for simultaneous continuum fitting.
 In practice, PyTorch also offers hardware acceleration (on GPU or TPU)\authorcomment1{citation needed}, mature support for a range of optimizers, and other perks that motivated its choice over analytic implementations or other machine learning frameworks.
@@
 
-Figure XX shows a representative Loss curve (MSE versus epochs) for cloning the model in Figure ZZ.
-The NN epochs of training took X minutes on an RTX2070 GPU with PyTorch v1.11.x, CUDA v11.x, and Intel X CPU, with all tensors as FP64.
+\begin{figure}[hbt!]
+  \centering
+  \includegraphics[width=0.6\textwidth]{example-image-golden}
+  \caption{PHOENIX spectrum cloned with blas\'e.}
+  \label{fig_cloned_spectrum_demo}
+\end{figure}
+
+Figure \ref{fig_cloned_spectrum_demo} shows a portion of a PHOENIX spectrum cloned with \emph{blas'e}.
+The $NN$ epochs of training took X minutes on an RTX2070 GPU with PyTorch v1.11.x, CUDA v11.x, and Intel X CPU, with all tensors as FP64.
 That benchmark works out to X seconds per epoch, or about Z seconds per minibatch.
@@
 
-We compute the residual $\mathsf{R}(\bm{\lambda}_s)$ of native PHOENIX minus cloned model, illustrated in the bottom panel of Figure XX\authorcomment1{make this figure}.
+We compute the residual $\mathsf{R}(\bm{\lambda}_s)$ of native PHOENIX minus cloned model, illustrated in the bottom panel of Figure \ref{fig_cloned_spectrum_demo}\authorcomment1{make this figure!}.
 We see an RMS of $\blackout{1}\%$ residuals at native resolution, with a long tail of residuals up to $\blackout{20}\%$.
 The greatest residuals surround the deepest and broadest lines---such as Hydrogen and neutral alkali metal lines---for which the non-Voigt lineshapes stand out.
-Figure ZZ\authorcomment1{make this figure} zooms in on the XX\authorcomment1{todo} line at $\censor{10200}\;$\AA.
+Figure \ref{fig_poor_cloning_performance}\authorcomment1{make this figure} zooms in on the \censor{XX} line at $\censor{10200}\;$\AA.
+
+\begin{figure}[hbt!]
+  \centering
+  \includegraphics[width=0.3\textwidth]{example-grid-100x100pt}
+  \caption{A \censor{Hydrogen} line with broad line wings exhibiting poor cloning performance.}
+  \label{fig_poor_cloning_performance}
+\end{figure}
 
@@
 We see some narrow lines devolve into missing continuum opacity, the favored tradeoff when the continuum estimate's poor performance outweighs the the pain of a narrow-but-tolerably-small spike.
-This problem can be seen in Figure XX where a line initialized at $\censor{10210}\;$\AA~ with an amplitude of ZZ and width of YY bloats to \censor{some other value}.
-These devolved lines tend to appear poorly immitated line wings.
+This problem can be seen in Figure \ref{fig_poor_cloning_performance} where a line initialized at $\censor{10210}\;$\AA~ with an amplitude of ZZ and width of YY bloats to \censor{some other value}.
+These devolved lines tend to appear at poorly immitated broad-and-deep line wings.
 One can mitigate these devolved lines by adding more flexibility to the continuum or supporting more advanced line shapes.

### 2021-11-16 paper/paper1/ms.tex
Add values to placeholders, prefactor discussion

@@
 An additional continuum flattening step ensures that subsequent spectral line finding steps get applied uniformly.
-Here, we identify the continuum regions as the most prominent peaks in the specturm separated by a pixel distance of at least $15000-90000$ pixels from adjacent peaks, yielding between 60 and 5 points.
-We then fit an $n_{\mathrm{poly}} = 3^{rd}-5^{th}$ order polynomial to these peaks, evaluate the polynomial on the entire bandwidth, and divide the spectrum by this trend.
-Visual inspection of this continuum-flattened spectrum typically possesses variations at the $<0.2\%$ level.
+Here, we identify the continuum regions by the most prominent peaks in the specturm separated by a distance of at least 5,000$-$90,000 pixels from adjacent peaks, yielding between 60 and 5 control points.
+We then fit an $n_{\mathrm{poly}} = 3^{rd}-15^{th}$ order polynomial to these peaks, evaluate the polynomial on the entire bandwidth $\mathsf{P}(\bm{\lambda}_s)$, and divide the spectrum by this trend.
+Visual inspection of this continuum-flattened spectrum typically possesses variations at the $<1\%$ level.
 Tuning the pixel distance and polynomial order can lower the undulations to $<0.05\%$ level for some input spectra.
-The high-pass filtering step should be set to capture the genuine spectral shape, without over-fitting broad line-wings such as those in deep Hydrogen and sodium lines.
+We found a pixel distance of 50,000 and $3^{rd}$ order polynomial worked across the range of effective temperatures for HPF's bandwidth.
+The IGRINS spectra had more significant continuum opacity and bandheads to accomodate and therefore needed to capture finer scale variation: 5,000 pixels and a $9^{th}$ order polynomial appeared adequate over most of the effective temperature range.
+This high-pass filtering step should be set to capture the genuine spectral shape, without over-fitting broad line-wings such as those in deep Hydrogen and sodium lines.
 We have experimentally found it helpful to hand-tune models of exceptionally broad and non-standard line-wings with an optional pre-processing step, not described here.
 
-We emphasize that a recreation of the unvarnished input spectrum can be obtained by multiplying the continuum-flattened signal by the ``perturbed black body'', $\mathsf{B}(\bm{\lambda}_s)$, that symbolizes the black body modulated by continuum opacity or broad-band radiative transfer effects.
-This smooth spectrum may be useful for applications that need to keep track of broad-band flux, such as low resolution spectrographs, or regions with molecular bandheads.
-The ``perturbed black body'' continuum model $\mathsf{B}$ contains  $n_{\mathrm{poly}}+1$ fixed lookup parameters, plus the input $T_{\mathrm{eff}}$.
+We emphasize that a recreation of the unvarnished input spectrum can be obtained by multiplying the continuum-flattened signal by the ``perturbed black body'', $\mathsf{B}(\bm{\lambda}_s)\cdot \mathsf{P}(\bm{\lambda}_s)$, that symbolizes the black body modulated by continuum opacity or broad-band radiative transfer effects.
+This smooth spectrum may be useful for applications that need to keep track of broad-band flux, such as low resolution spectra, or regions with molecular bandheads.
+The ``perturbed black body'' continuum model contains $n_{\mathrm{poly}}+1$ fixed-but-possibly tunable lookup parameters, plus the fixed input $T_{\mathrm{eff}}$.
 
@@
 currently on the Gemini South Telescope on Cerro Pach\'on in Chile.
-The $R=55,000$ HPF has a native bandwidth of $x-y$, that we expand to $x-y^*$ including the edge buffers.
-IGRINS has two cameras for $H$ and $K$ band, with the combined spectrum spanning $z-y$ including the edge buffers and the region in-between the two cameras, all at a resolving power of $(R=45,000)$.
+The $R=55,000$ HPF has a native bandwidth of $8079-12785$~\AA, that we expand to $8049-12815$~\AA~including the edge buffers.
+IGRINS has two cameras for $H$ and $K$ band, with the combined spectrum spanning $14267-25217\;$\AA~ including the edge buffers and the region in-between the two cameras, all at a resolving power of $(R=45,000)$.
 The spectrograph acquisition, reduction, and post-processing steps yield data $\mathsf{D}(\bm{\lambda}_{d})$, where $\bm{\lambda}_{d}$ is the wavelength vector at the instrumental resolution and sampling of each instrument, generally much coarser than the resolution and sampling grid of the precomputed synthetic spectra.
-The data wavelength vector may also contains gaps between \'echelle orders, whereas the precomputed grid is usually contiguous.
-HPF may have up to $2048\times28=57,344$ pixels, and IGRINS has typically about $75,000$ pixels, after common truncation of noisy edge pixels.
+The data wavelength vector may also contain gaps between \'echelle orders, whereas the precomputed wavelength coordinates are usually contiguous.
+HPF may have up to $2048\times28=$57,344 pixels, and IGRINS has typically about 75,000 pixels, after common trimming of noisy edge pixels and unusable telluric regions.
 
@@
 Figure \ref{fig_Nlines_vs_teff} shows how the number of detected lines $N_{\mathrm{lines}}$ scales with effective temperature and prominence threshold $P_{rom}$ for the \texttt{PHOENIX} grid, truncated to the bandwidths-plus-buffers for HPF and IGRINS.
-The HPF-truncated spectra have $N_s=XX$ and the IGRINS-truncated spectra have $N_s=XX$.
+The HPF-truncated spectra have $N_s=$335,849 and the IGRINS-truncated spectra have $N_s=$330,052.
 
@@
 \begin{eqnarray}
-  \mathsf{M}(\bm{\lambda}_s) = {\displaystyle \prod_{j=1}^{N_{\mathrm{lines}}}  (1-a_j \mathsf{V}(\bm{\lambda}_s-\lambda_{c,j}, \sigma_j, \gamma_j) )} \label{equation1}
+  \mathsf{M}(\bm{\lambda}_s) = {\displaystyle \mathsf{P} \prod_{j=1}^{N_{\mathrm{lines}}}  (1-a_j \mathsf{V}(\bm{\lambda}_s-\lambda_{c,j}, \sigma_j, \gamma_j) )} \label{equation1}
 \end{eqnarray}
@@
 The amplitude $a$ is always expected to be positive for absorption lines, and is equal to the area under the curve of the line since the Voigt profile is normalized.
+The optional Prefactor term $\mathsf{P}$ is fixed at $1$ by default, but can be allowed to vary to fine-adjust imperfections left over from the polynomial-based continuum flattening procedure.
 
@@
 
-As seen in Figure \ref{fig_Nlines_vs_teff}, the number of lines can exceed $XX$, meaning the model has over $XX \times 3 = ZZ$ free parameters.
+As seen in Figure \ref{fig_Nlines_vs_teff}, the number of lines can exceed 7,000, meaning the model has over 7,000 $\times 3 =$ 21,000 free parameters.
 Fitting that large number of parameters is difficult with conventional optimizers, which struggle to converge after hundreds of parameters\authorcomment1{Citation to limits of optimizers}.

### 2021-11-17 examples/clone_PHOENIX_spectrum.py
Move a bunch of notebooks to deprecated for now, add a bunch of old notebooks

@@ -77,7 +77,7 @@ optimizer = optim.Adam(
     filter(lambda p: p.requires_grad, emulator.parameters()), 0.004, amsgrad=True
 )
 n_epochs = 100
-sub_divisions = 30
+sub_divisions = 50
 losses = []
 
 # Experiment: just optimize the line cores:

### 2021-11-17 examples/gpytorch_demo.py
Add a GPyTorch demo as an experiment- can it find the right values?

@@ -1,0 +1,140 @@
+import torch
+from torch import nn
+from tqdm import trange
+import torch.optim as optim
+import matplotlib.pyplot as plt
+from torch.utils.tensorboard import SummaryWriter
+import webbrowser
+import numpy as np
+import celerite2
+from celerite2 import terms
+import gpytorch
+from tqdm import tqdm
+
+if torch.cuda.is_available():
+    device = "cuda"
+else:
+    device = "cpu"
+
+
+def plot_spectrum(data):
+    """
+    Generates matplotlib Figure using a trained network, along with images
+    and labels from a batch
+    """
+    fig, axes = plt.subplots(2, figsize=(8, 3))
+    axes[0].step(
+        data[0]["x"].numpy(), data[0]["y"].numpy(), label="Fit", lw=1,
+    )
+    axes[0].step(
+        data[1]["x"].numpy(), data[1]["y"].numpy(), label="Input Data", lw=1,
+    )
+    axes[0].set_ylim(-10, 10)
+    axes[0].legend(loc="upper right", ncol=2)
+    axes[1].step(
+        data[1]["x"].numpy(), data[0]["y"].numpy() - data[1]["y"].numpy(),
+    )
+    axes[1].set_ylim(-10, 10)
+    return fig
+
+
+# Set up parameterization
+n_samples = 10000
+x_vector = np.linspace(-1, 1, n_samples, endpoint=False)
+true_sigma, true_rho = 1.7, 0.1
+std_dev = 0.005
+variance = std_dev ** 2
+yerr = np.repeat(std_dev, n_samples)
+
+# Set up GP with celerite
+kernel = terms.SHOTerm(sigma=true_sigma, rho=true_rho, Q=0.25)
+gp = celerite2.GaussianProcess(kernel, mean=0.0)
+gp.compute(x_vector, yerr=None)
+
+# Make fake signal
+noise_free_signal = gp.sample()
+noise_draw = np.random.normal(scale=yerr)
+y_fake = noise_free_signal + noise_draw
+
+
+writer = SummaryWriter(log_dir="runs/gpytorch1")
+webbrowser.open("http://localhost:6006/", new=2)
+
+
+## GPyTorch
+train_x = torch.tensor(x_vector, dtype=torch.float64).to(device)
+train_y = torch.tensor(y_fake, dtype=torch.float64).to(device)
+train_yerr = torch.tensor(yerr, dtype=torch.float64).to(device)
+
+
+class GPRegressionModel(gpytorch.models.ExactGP):
[…]

### 2021-12-01 examples/clone_PHOENIX_spectrum.py
Add tensorboard profiling to the cloning script

@@ -2,6 +2,7 @@ import torch
 from torch import nn
 from tqdm import trange
 import torch.optim as optim
+import torch.profiler
 from blase.emulator import PhoenixEmulator
 import matplotlib.pyplot as plt
 from gollum.phoenix import PHOENIXSpectrum
@@ -43,7 +44,8 @@ def plot_spectrum(spectra):
     return fig
 
 
-writer = SummaryWriter(log_dir="runs/emulator1")
+log_dir = "runs/gpytorch1"
+writer = SummaryWriter(log_dir=log_dir)
 webbrowser.open("http://localhost:6006/", new=2)
 
 
@@ -86,6 +88,14 @@ indices = np.isin(emulator.wl_native.cpu(), emulator.lam_centers.detach().cpu())
 # dilated_inds = binary_dilation(indices, iterations=2)
 peak_indices = torch.tensor(indices).to(device)
 
+
+prof = torch.profiler.profile(
+    schedule=torch.profiler.schedule(wait=1, warmup=3, active=6, repeat=2),
+    on_trace_ready=torch.profiler.tensorboard_trace_handler(log_dir),
+    record_shapes=True,
+    with_stack=True,
+)
+prof.start()
 t_iter = trange(n_epochs, desc="Training", leave=True)
 for epoch in t_iter:
     for i in range(sub_divisions):
@@ -102,6 +112,7 @@ for epoch in t_iter:
         optimizer.zero_grad()
         losses.append(loss.item())
         t_iter.set_description("Training Loss: {:0.8f}".format(loss.item()))
+    prof.step()
 
     writer.add_scalar("loss", loss.item(), global_step=epoch)
     writer.add_scalar("a", emulator.a_coeff.item(), global_step=epoch)
@@ -121,5 +132,6 @@ for epoch in t_iter:
             "predictions vs. actuals", plot_spectrum(to_plot), global_step=epoch,
         )
 
+prof.stop()
 torch.save(emulator.state_dict(), "native_res_0p1prom.pt")
 

### 2021-12-01 paper/paper1/ms.tex
Add Figure 1 and text around it

@@
 Next we identify the spectral lines.
-We apply a local-minimum-finding algorithm on the spectrum by defining a prominence threshold $P_{rom} \in (0.005, 0.1)$.
+We apply a local-minimum-finding algorithm on the spectrum by defining a prominence threshold $P_{rom} \in (0.005, 0.02)$.
 This threshold dictates the number of lines that will be modeled: a lower prominence finds more, weaker lines, and a larger prominence finds fewer, deeper lines.
@@
   \centering
-  \includegraphics[width=0.5\textwidth]{example-image-a}
-  \caption{Number of prominent spectral lines versus effective temperature for PHOENIX models truncated to HPF (\emph{left}) and IGRINS (\emph{right}) bandwidths, for different prominence thresholds of $0.05$ and $0.01$.}
+  \includegraphics[width=0.5\textwidth]{figures/N_lines_vs_Teff_prom.png}
+  \caption{Number of prominent spectral lines versus effective temperature for PHOENIX models truncated to IGRINS (\emph{blue, connected} points) and HPF (\emph{orange, free-standing} points) bandwidths, for different prominence thresholds of $0.02$, $0.01$, and $0.005$.}
   \label{fig_Nlines_vs_teff}
@@
 Figure \ref{fig_Nlines_vs_teff} shows how the number of detected lines $N_{\mathrm{lines}}$ scales with effective temperature and prominence threshold $P_{rom}$ for the \texttt{PHOENIX} grid, truncated to the bandwidths-plus-buffers for HPF and IGRINS.
-The HPF-truncated spectra have $N_s=$335,849 and the IGRINS-truncated spectra have $N_s=$330,052.
+We see between about 2,000 and 20,000 lines depending on the $T_{\mathrm{eff}}$ and $P_{rom}$.
+HPF and IGRINS have comparable number of lines, and halving the prominence increases the number of lines by about $20-30\%$ in these ranges.
+The number of lines monotonically increases towards cooler effective temperatures.
+The HPF-truncated spectra have $N_s=$335,849 native resolution samples, comparable to the IGRINS-truncated spectra, $N_s=$330,052.
 
@@
 
-Figure \ref{fig_cloned_spectrum_demo} shows a portion of a PHOENIX spectrum cloned with \emph{blas'e}.
+Figure \ref{fig_cloned_spectrum_demo} shows a portion of a PHOENIX spectrum cloned with \emph{blas\'e}.
 The $NN$ epochs of training took X minutes on an RTX2070 GPU with PyTorch v1.11.x, CUDA v11.x, and Intel X CPU, with all tensors as FP64.

### 2022-01-03 paper/paper1/ms.tex
Add prose to sparse section, advanced lineshapes

@@
 
-Another solution to the computation problem could be to take advantage of the sparsity of the $\bm{\bar{V}}$ matrix: most of the entries are vanishingly close to zero.
+Another solution to the computation problem is to take advantage of the sparsity of the $\bm{\bar{V}}$ matrix: most of the Voigt profile's entries far from the line center are vanishingly close to zero.
 Ideally we would only evaluate the model close to the line center position, some number of line cuts $N_{\mathrm{cut}}=\alpha \cdot \mathrm{FWHM}$ away.
@@
 A few practical considerations make sparse matrices peskier to work with than their dense counterparts.
-First, the $\alpha$ hyperparameter should be set large enough that the truncation effect is not seen for the deepest lines.
+First, the $\alpha$ hyperparameter should be set large enough that the truncation effect is not seen for the deepest lines, or in post-processing steps such as line broadening.
 The $\alpha$ should be preselected before the training process.
 If the FWHM initialization was poor enough, the question may arise how autodiff training will cope with the poor performance of a model, leading to the prospect of strange and difficult to diagnose artifacts near spurious line-cut truncation artifacts.
-At the time of writing, we do not implement sparse matrices in \emph{blas\'e}, but we find them the most compelling route for transformative performance improvements.
+
+We implemented sparse matrices in \emph{blas\'e} using the \texttt{sparse\_coo\_tensor} function in \emph{PyTorch}.
+
+There are a few standardization strategies for applying the wing cuts.
+Here we coerce all wing cuts to be exactly 6000 pixels, with the middle pixel being at the line center position, and 3000 pixels to the red and blue side of the line.
+Most spectral lines only appreciably affect the signal in the central $\sim5-50$ pixels, so such a large wing cut may seem over-zealous.
+We emphasize that the line cut serves as the fixed pixel locations where the line will be evaluated with post-processing steps such as radial velocity shifts and line broadening.
+So these 6000 pixels must be capable of handling the fastest-moving and fast-spinning stars you expect to see in your observed spectra of interest.
+
+The PHOENIX spectra double their pixel sampling shortward of $\lambda = 1 \mu$m, so 6000 pixels in the visible is about 30 Angstroms, but 6000 pixels in the infrared is about 60 Angstroms.
+We \emph{could have} demanded that the wingcut be exactly 30 Å, or some other heuristic, but that choice would result in so-called \emph{jagged} or \emph{ragged} arrays: collections of 1D arrays that differ in their lengths.
+Some lines would have 5999 pixels, some would have 6000, and others would have 2999 or 3000 pixels, depending whether they occur to the red or blue side of the $1 \;\mu$m kink in sampling.
+Most efficient CPU and GPU algorithms are not equipped to deal with ragged arrays, therefore increasing the computation time.
+Instead, the choice of exactly 6000 pixels means we can use standard matrix or tensor arithmetic, which somewhat counterintuitively is faster and handles more pixels for all spectral lines: 6000 pixels everywhere is faster than 6000 pixels for some and 2999 for other lines.
+
+Locating the indicies of these 6000 ``nearest-neighbor'' pixels involves a sorting algorithm that takes about 0.01 seconds per spectral line, so for 6500 spectral lines the preprocessing step takes about 90 seconds.
+This initialization only occurs once and then is fixed, so we should not worry too much that it is slow.
+Heuristics for finding these indices faster are also feasible.
+
+We rewrite Equation \ref{equation1} as a sum by taking the log of both sides
+
+\begin{eqnarray}
+  \ln{\mathsf{F}(\bm{\lambda})} = \sum_{i=1}^{N_{lines}} \ln{\left(1-V(\bm{\lambda};A_i, \lambda_{c,i}, \sigma_i) \right)}
+\end{eqnarray}
+
+Sparse matrix methods generally support an operation known as \emph{coalescing}, which sums values with repeated indices.
+Each pixel gets computed on average about $\sim100$ times in this sparse implementation, which is about $50\times$ better than each pixel getting computed $N_{lines}\sim6500$ times in the dense approach.
+
+\subsection{Broad lines and advanced lineshapes}
+
+Some lines, such as Hydrogen, Sodium, potassium, neutral metal lines, and others have extremely broad line wings, approaching larger than the $\sim6000$ pixels we allocate for the sparse implementation.
+These special lines need to be handled separately from the weak lines, both from a computational performance perspective and an accuracy perspective.
+
+Extremely broad lines will exhibit truncation effects if the sparse window is small compared to the linewing size.
+The truncation effects will look like tophat functions severing the asymptotic wings, imbuing artificial step function kinks in the emulated spectrum.
+We can afford to increase the sparse window on a few, say $N_{broad}\sim20$ of the broadest spectral lines.
+We then construct and evaluate the entire dense matrix for those lines: $\sim 330\;000 \times 20$.
+The number of FLOPS in each category scales as about 6 Million for the 20 broad lines versus about 36 Million for the sea of about 6500 narrow lines, depending on the exact choices for wingcuts and number of lines.
+
+We introduce advanced lineshapes for these $\sim20$ broad lines, perturbing the Voigt line-wings with a smooth wavelength-dependent correction term $\mathsf{G}$:
+
+\begin{eqnarray}
+  \mathsf{\tilde{V}(\bm{\lambda})} &=& \mathsf{V} \odot \mathsf{G}\\
+  \mathsf{G} &=& 1 + (e^{a_j} - 1) \cdot \mathcal{S}\left(\frac{|\bm{\lambda}-\lambda_{c,j}| - \lambda_{t, j}}{b_j}\right)
+\end{eqnarray}
+
+where $\odot$ is the element-wise product (\emph{a.k.a} Hadamard product), $\mathcal{S}$ is the sigmoid function, and we have introduced three new tunable parameters for each of the $j$ broad lines; $\lambda_t$ is the truncation wavelength, $b$ is a scale parameter for how slowly or how rapidly in wavelength-space the transition from non-Lorentzian proceeds, and $a$ is a possibly negative stretch parameter that controls whether the line wing is sub- or super- Lorentzian.
+
+This functional form has a few advantages.
+It is smooth.
+The smoothness of the transition is controlled by a tunable parameter, $b$.
+It can handle either sub- or super-Lorentzian shapes.
+It is roughly based on the $\chi$-factor\authorcomment1{citation to Hartmann et al.
+1989, etc.}.
+In the limit $\lim_{a\to0} \mathsf{G}$, the lineshape becomes exactly Lorentzian.
+The sigmoid is efficiently implemented in PyTorch.
+It enforces that the perturbation only produces absorption and not emission profiles.
 

### 2022-01-10 src/blase/emulator.py
Some weird segfaults from torch dtypes and possibly sorting.

@@ -230,7 +230,7 @@ class SparsePhoenixEmulator(PhoenixEmulator):
     device (Torch Device or str): GPU or CPU?
     """
 
-    def __init__(self, wl_native, flux_native, prominence=0.03, device=None):
+    def __init__(self, wl_native, flux_native, prominence=0.01, device=None):
         super().__init__(wl_native, flux_native, prominence=prominence)
 
         if device is None:
@@ -245,19 +245,26 @@ class SparsePhoenixEmulator(PhoenixEmulator):
         # Currently defined in *pixels*
         wing_cut_pixels = 6000
 
+        print("-------------------------------------------------")
+        print("Initializing a sparse representation of the model")
         with torch.no_grad():
             list_of_nonzero_indices = []
             for line_center in tqdm(self.lam_centers.detach().numpy()):
                 distance = np.abs(wl_native - line_center)
                 these_pixels = np.argsort(distance)[0:wing_cut_pixels]
-                list_of_nonzero_indices.append(sorted(these_pixels))
+                list_of_nonzero_indices.append(np.sort(these_pixels))
 
-        self.indices_2D = torch.tensor(list_of_nonzero_indices, device=device)
-        self.indices_1D = self.indices_2D.reshape(-1)
-        self.indices = self.indices_1D.unsqueeze(0)
-        self.wl_2D = self.wl_native.to(device)[self.indices_2D]
-        self.wl_1D = self.wl_2D.reshape(-1)
-        self.active_mask = self.active_mask.to(device)
+            print("Successfully initialized a sparse representation of the spectrum")
+            self.indices_2D = torch.tensor(
+                list_of_nonzero_indices, dtype=torch.long, device=device
+            )
+            self.indices_1D = self.indices_2D.reshape(-1)
+            self.indices = self.indices_1D.unsqueeze(0)
+            self.wl_2D = self.wl_native.to(device)[self.indices_2D]
+            self.wl_1D = self.wl_2D.reshape(-1)
+            self.active_mask = self.active_mask.to(device)
+            print("Successfully reshaped and stored the sparse representation")
+            print("-------------------------------------------------")
 
     def forward(self):
         """The forward pass of the sparse implementation--- no wavelengths needed!

### 2022-01-10 paper/paper1/ms.tex
Start of some prose about the sparse implementation

@@
 The IGRINS spectra had more significant continuum opacity and bandheads to accomodate and therefore needed to capture finer scale variation: 5,000 pixels and a $9^{th}$ order polynomial appeared adequate over most of the effective temperature range.
+
 This high-pass filtering step should be set to capture the genuine spectral shape, without over-fitting broad line-wings such as those in deep Hydrogen and sodium lines.
-We have experimentally found it helpful to hand-tune models of exceptionally broad and non-standard line-wings with an optional pre-processing step, not described here.
+We have experimentally found it helpful to hand-tune models of exceptionally broad and non-standard line-wings with an optional pre-processing step that depends on the wavelength range of interest.
+Here we sketch the approach for common broad lines found in the long-wavelength visible and near-infrared wavelength regions.
+
+% TODO We pre-assign spectral lines for the conspicuous Hydrogen lines and neutral metal lines.  Table
 

### 2022-01-10 src/blase/emulator.py
Woohoo, dramatically speed up the sparse initialization

@@ -243,28 +243,40 @@ class SparsePhoenixEmulator(PhoenixEmulator):
 
         ## Define the wing cut
         # Currently defined in *pixels*
-        wing_cut_pixels = 6000
+        wing_cut_pixels = 1000
 
+        lines = self.lam_centers.detach().cpu().numpy()
+        wl_native = self.wl_native.cpu().numpy()
         print("-------------------------------------------------")
-        print("Initializing a sparse representation of the model")
-        with torch.no_grad():
-            list_of_nonzero_indices = []
-            for line_center in tqdm(self.lam_centers.detach().numpy()):
-                distance = np.abs(wl_native - line_center)
-                these_pixels = np.argsort(distance)[0:wing_cut_pixels]
-                list_of_nonzero_indices.append(np.sort(these_pixels))
-
-            print("Successfully initialized a sparse representation of the spectrum")
-            self.indices_2D = torch.tensor(
-                list_of_nonzero_indices, dtype=torch.long, device=device
+        print(
+            "Initializing a sparse representation of the model with {:} spectral lines".format(
+                len(lines)
             )
-            self.indices_1D = self.indices_2D.reshape(-1)
-            self.indices = self.indices_1D.unsqueeze(0)
-            self.wl_2D = self.wl_native.to(device)[self.indices_2D]
-            self.wl_1D = self.wl_2D.reshape(-1)
-            self.active_mask = self.active_mask.to(device)
-            print("Successfully reshaped and stored the sparse representation")
-            print("-------------------------------------------------")
+        )
+
+        # Find the index position of each spectral line
+        center_indices = np.searchsorted(wl_native, lines)
+
+        # From that, determine the beginning and ending indices
+        zero_indices = center_indices - (wing_cut_pixels // 2)
+        too_low = zero_indices < 0
+        zero_indices[too_low] = 0
+        end_indices = zero_indices + wing_cut_pixels
+        too_high = end_indices > self.n_pix
+        zero_indices[too_high] = len(wl_native) - wing_cut_pixels
+        end_indices[too_high] = len(wl_native)
+
+        # Make a 2D array of the indices
+        indices_2D = np.linspace(
+            zero_indices, end_indices, num=wing_cut_pixels, endpoint=True
+        )
+
+        self.indices_2D = torch.tensor(indices_2D.T, dtype=torch.long, device=device)
+        self.indices_1D = self.indices_2D.reshape(-1)
[…]

### 2022-01-11 src/blase/emulator.py
Add wingcut_pixels as a kwarg to the sparse emulator

@@ -33,9 +33,11 @@ class PhoenixEmulator(nn.Module):
         self.wl_native = torch.tensor(wl_native)
         self.flux_native = torch.tensor(flux_native)
 
-        (lam_centers, amplitudes, widths_angstroms,) = self.detect_lines(
-            self.wl_native, self.flux_native, prominence=prominence
-        )
+        (
+            lam_centers,
+            amplitudes,
+            widths_angstroms,
+        ) = self.detect_lines(self.wl_native, self.flux_native, prominence=prominence)
 
         self.n_pix = len(wl_native)
 
@@ -92,7 +94,7 @@ class PhoenixEmulator(nn.Module):
 
         # Fix the wavelength centers as gospel for now.
         self.lam_centers = nn.Parameter(
-            lam_centers.clone().detach().requires_grad_(False)
+            lam_centers.clone().detach().requires_grad_(True)
         )
 
         self.a_coeff = nn.Parameter(
@@ -228,9 +230,13 @@ class SparsePhoenixEmulator(PhoenixEmulator):
     flux_native (float vector): The native flux
     prominence (int scalar): The threshold for detecting lines
     device (Torch Device or str): GPU or CPU?
+    wing_cut_pixels (int scalar): the number of pixels centered on the line center
+        to evaluate in the sparse implementation, default: 1000 pixels
     """
 
-    def __init__(self, wl_native, flux_native, prominence=0.01, device=None):
+    def __init__(
+        self, wl_native, flux_native, prominence=0.01, device=None, wing_cut_pixels=None
+    ):
         super().__init__(wl_native, flux_native, prominence=prominence)
 
         if device is None:
@@ -243,16 +249,14 @@ class SparsePhoenixEmulator(PhoenixEmulator):
 
         ## Define the wing cut
         # Currently defined in *pixels*
-        wing_cut_pixels = 1000
+        if wing_cut_pixels is None:
+            wing_cut_pixels = 1000
+        else:
+            wing_cut_pixels = int(wing_cut_pixels)
 
         lines = self.lam_centers.detach().cpu().numpy()
         wl_native = self.wl_native.cpu().numpy()
-        print("-------------------------------------------------")
-        print(
-            "Initializing a sparse representation of the model with {:} spectral lines".format(
-                len(lines)
-            )
-        )
+        print("Initializing a sparse model with {:} spectral lines".format(len(lines)))
 
         # Find the index position of each spectral line
         center_indices = np.searchsorted(wl_native, lines)

### 2022-01-11 experiments/01_accuracy_vs_prom/accuracy_vs_prom.py
An experiment to measure accuracy vs line detection threshold, prom

@@ -1,0 +1,78 @@
+import torch
+from torch import nn
+from tqdm import trange
+import torch.optim as optim
+import torch.profiler
+from blase.emulator import SparsePhoenixEmulator
+import matplotlib.pyplot as plt
+from gollum.phoenix import PHOENIXSpectrum
+from torch.utils.tensorboard import SummaryWriter
+import webbrowser
+import numpy as np
+
+if torch.cuda.is_available():
+    device = "cuda"
+else:
+    device = "cpu"
+
+
+log_dir = "runs/sparse1"
+writer = SummaryWriter(log_dir=log_dir)
+webbrowser.open("http://localhost:6006/", new=2)
+
+
+# Pre-process the model as described in the paper
+spectrum = PHOENIXSpectrum(teff=4700, logg=4.5)
+spectrum = spectrum.divide_by_blackbody()
+spectrum = spectrum.normalize()
+continuum_fit = spectrum.fit_continuum(polyorder=5)
+spectrum = spectrum.divide(continuum_fit, handle_meta="ff")
+
+# Numpy arrays: 1 x N_pix
+wl_native = spectrum.wavelength.value
+flux_native = spectrum.flux.value
+
+# Vary the prominence parameter
+prominences = [0.2, 0.1, 0.06, 0.03, 0.01, 0.006, 0.003, 0.001]
+for i, prominence in enumerate(prominences):
+    emulator = SparsePhoenixEmulator(
+        wl_native, flux_native, prominence=prominence, device=None, wing_cut_pixels=100
+    )
+    emulator.to(device)
+    target = (
+        emulator.flux_native.clone()
+        .detach()
+        .to(device)[emulator.active_mask.cpu().numpy()]
+    )
+
+    n_lines = len(emulator.lam_centers.detach())
+
+    loss_fn = nn.MSELoss(reduction="mean")
+    optimizer = optim.Adam(
+        filter(lambda p: p.requires_grad, emulator.parameters()), 0.03, amsgrad=True
+    )
+    n_epochs = 100
+    losses = []
+    std_devs = []
+
+    t_iter = trange(n_epochs, desc="Training", leave=True)
+    for epoch in t_iter:
+        emulator.train()
+        yhat = emulator.forward()
+        loss = loss_fn(yhat, target)
+        loss.backward()
+        optimizer.step()
+        optimizer.zero_grad()
+        losses.append(loss.item())
[…]

### 2022-01-11 examples/clone_sparse.py
CPU is so much slower than GPU, but you can tweak the wingcut for speed

@@ -61,8 +61,10 @@ wl_native = spectrum.wavelength.value
 flux_native = spectrum.flux.value
 
 # Create the emulator
-prominence = 0.01
-emulator = SparsePhoenixEmulator(wl_native, flux_native, prominence=prominence)
+prominence = 0.03
+emulator = SparsePhoenixEmulator(
+    wl_native, flux_native, prominence=prominence, device=None, wing_cut_pixels=100
+)
 emulator.to(device)
 
 n_pix = len(emulator.wl_native)
@@ -73,7 +75,7 @@ target = (
 )
 
 # Training Loop
-emulator.lam_centers.requires_grad = False
+emulator.lam_centers.requires_grad = True
 emulator.a_coeff.requires_grad = False
 emulator.b_coeff.requires_grad = False
 emulator.c_coeff.requires_grad = False
@@ -82,10 +84,10 @@ loss_fn = nn.MSELoss(reduction="mean")
 optimizer = optim.Adam(
     filter(lambda p: p.requires_grad, emulator.parameters()), 0.03, amsgrad=True
 )
-n_epochs = 1000
+n_epochs = 200
 losses = []
 
-plot_every_N_steps = 45
+plot_every_N_steps = 25
 t_iter = trange(n_epochs, desc="Training", leave=True)
 for epoch in t_iter:
     emulator.train()
@@ -107,11 +109,16 @@ for epoch in t_iter:
         flux_clone = yhat.detach().cpu()
         flux_targ = target.cpu()
         to_plot = [
-            {"wl": wl_plot, "flux": flux_clone,},
+            {
+                "wl": wl_plot,
+                "flux": flux_clone,
+            },
             {"wl": wl_plot, "flux": flux_targ},
         ]
         writer.add_figure(
-            "predictions vs. actuals", plot_spectrum(to_plot), global_step=epoch,
+            "predictions vs. actuals",
+            plot_spectrum(to_plot),
+            global_step=epoch,
         )
 
 torch.save(emulator.state_dict(), "sparse_T4700g4p5_prom0p02_HPF.pt")

### 2022-01-20 examples/clone_sparse.py
We have been taking out and putting back in the emulator edge mask

@@ -84,15 +84,15 @@ loss_fn = nn.MSELoss(reduction="mean")
 optimizer = optim.Adam(
     filter(lambda p: p.requires_grad, emulator.parameters()), 0.03, amsgrad=True
 )
-n_epochs = 200
+n_epochs = 10_000
 losses = []
 
-plot_every_N_steps = 25
+plot_every_N_steps = 1000
 t_iter = trange(n_epochs, desc="Training", leave=True)
 for epoch in t_iter:
     emulator.train()
     yhat = emulator.forward()
-    loss = loss_fn(yhat, target)
+    loss = loss_fn(yhat[emulator.active_mask], target)
     loss.backward()
     optimizer.step()
     optimizer.zero_grad()
@@ -106,19 +106,14 @@ for epoch in t_iter:
     if (epoch % plot_every_N_steps) == 0:
         # torch.save(model.state_dict(), "model_coeffs.pt")
         wl_plot = wl_active.cpu()
-        flux_clone = yhat.detach().cpu()
+        flux_clone = yhat[emulator.active_mask].detach().cpu()
         flux_targ = target.cpu()
         to_plot = [
-            {
-                "wl": wl_plot,
-                "flux": flux_clone,
-            },
+            {"wl": wl_plot, "flux": flux_clone,},
             {"wl": wl_plot, "flux": flux_targ},
         ]
         writer.add_figure(
-            "predictions vs. actuals",
-            plot_spectrum(to_plot),
-            global_step=epoch,
+            "predictions vs. actuals", plot_spectrum(to_plot), global_step=epoch,
         )
 
 torch.save(emulator.state_dict(), "sparse_T4700g4p5_prom0p01_HPF.pt")

### 2022-01-20 experiments/02_synthetic_injection/02_synthetic_injection.py
Commit the injection/recovery code and parameters

@@ -1,0 +1,168 @@
+import os
+import torch
+from torch import nn
+from tqdm import trange
+import torch.optim as optim
+import torch.profiler
+from blase.emulator import SparsePhoenixEmulator
+import matplotlib.pyplot as plt
+from gollum.phoenix import PHOENIXSpectrum
+from torch.utils.tensorboard import SummaryWriter
+import webbrowser
+import numpy as np
+from muler.hpf import HPFSpectrum
+
+if torch.cuda.is_available():
+    device = "cuda"
+else:
+    device = "cpu"
+
+
+def plot_spectrum(spectra):
+    """
+    Generates matplotlib Figure using a trained network, along with images
+    and labels from a batch
+    """
+    fig, axes = plt.subplots(2, figsize=(8, 3))
+    axes[0].step(
+        spectra[0]["wl"].numpy(),
+        spectra[0]["flux"].numpy(),
+        label="Cloned Spectrum",
+        lw=1,
+    )
+    axes[0].step(
+        spectra[1]["wl"].numpy(),
+        spectra[1]["flux"].numpy(),
+        label="Native Spectrum",
+        lw=1,
+    )
+    axes[0].set_ylim(0, 1.5)
+    axes[0].legend(loc="upper right", ncol=2)
+    axes[1].step(
+        spectra[1]["wl"].numpy(),
+        spectra[0]["flux"].numpy() - spectra[1]["flux"].numpy(),
+    )
+    axes[1].set_ylim(-0.3, 0.3)
+    return fig
+
+
+log_dir = "runs/synthetic1"
+writer = SummaryWriter(log_dir=log_dir)
+webbrowser.open("http://localhost:6006/", new=2)
+
+
+# Pre-process the model as described in the paper
+spectrum = PHOENIXSpectrum(teff=4100, logg=3.5)
+spectrum = spectrum.divide_by_blackbody()
+spectrum = spectrum.normalize()
+continuum_fit = spectrum.fit_continuum(polyorder=5)
+spectrum = spectrum.divide(continuum_fit, handle_meta="ff")
+
+# Fetch the real HPF data
+# For now we will use some data from another project
+fn = "Goldilocks_20210919T084302_v1.0_0063.spectra.fits"
+dir = "~/GitHub/star-witness/data/HPF/goldilocks/UT21-3-015/"
+dir = os.path.expanduser(dir)
[…]

### 2022-01-20 experiments/02_synthetic_injection/02_synthetic_injection.py
GPU-specific changes to injection/recovery experiment 02

@@ -100,11 +100,11 @@ synthetic_spectrum = torch.load("./synthetic_perturbed_spectrum.pt")
 
 from blase.emulator import EchelleModel
 
-model = EchelleModel(data.spectral_axis.bin_edges.value, wl_native)
+model = EchelleModel(data.spectral_axis.bin_edges.value, wl_native.cpu())
 
 # Do not train on real data.  Train on Synthetic data with known properties!
 ## data_target = torch.tensor(data.flux.value, device=device, dtype=torch.float64)
-data_target = synthetic_spectrum
+data_target = synthetic_spectrum.to(device)
 
 data_wavelength = torch.tensor(
     data.wavelength.value, device=device, dtype=torch.float64
@@ -118,12 +118,12 @@ optimizer = optim.Adam(
     0.01,
     amsgrad=True,
 )
-n_epochs = 200
+n_epochs = 5000
 losses = []
 
 # high_res_model = emulator.flux_native.clone().detach().to(device)
 
-plot_every_N_steps = 1
+plot_every_N_steps = 100
 t_iter = trange(n_epochs, desc="Training", leave=True)
 for epoch in t_iter:
     model.train()
@@ -133,10 +133,10 @@ for epoch in t_iter:
     loss = loss_fn(yhat, data_target)
     loss.backward()
     optimizer.step()
-    for name, param in model.named_parameters():
-        print(name, param.grad)
-    for name, param in emulator.named_parameters():
-        print(name, param.grad)
+    # for name, param in model.named_parameters():
+    #    print(name, param.grad)
+    # for name, param in emulator.named_parameters():
+    #    print(name, param.grad)
     optimizer.zero_grad()
     losses.append(loss.item())
     t_iter.set_description("Training Loss: {:0.8f}".format(loss.item()))
@@ -153,16 +153,11 @@ for epoch in t_iter:
         flux_clone = yhat.detach().cpu()
         flux_targ = data_target.cpu()
         to_plot = [
-            {
-                "wl": wl_plot,
-                "flux": flux_clone,
-            },
+            {"wl": wl_plot, "flux": flux_clone,},
             {"wl": wl_plot, "flux": flux_targ},
         ]
         writer.add_figure(
-            "predictions vs. actuals",
-            plot_spectrum(to_plot),
-            global_step=epoch,
+            "predictions vs. actuals", plot_spectrum(to_plot), global_step=epoch,
         )
 
 torch.save(model.state_dict(), "model_T4100g3p5_prom0p01_HPF_recovery.pt")

### 2022-01-20 src/blase/emulator.py
Experimental new strategy for combining emulator into echelle model

@@ -34,11 +34,9 @@ class PhoenixEmulator(nn.Module):
         self.wl_native = torch.tensor(wl_native)
         self.flux_native = torch.tensor(flux_native)
 
-        (
-            lam_centers,
-            amplitudes,
-            widths_angstroms,
-        ) = self.detect_lines(self.wl_native, self.flux_native, prominence=prominence)
+        (lam_centers, amplitudes, widths_angstroms,) = self.detect_lines(
+            self.wl_native, self.flux_native, prominence=prominence
+        )
 
         self.n_pix = len(wl_native)
 
@@ -283,10 +281,6 @@ class SparsePhoenixEmulator(PhoenixEmulator):
         self.wl_1D = self.wl_2D.reshape(-1)
         self.active_mask = self.active_mask.to(device)
 
-        self.radial_velocity = nn.Parameter(
-            torch.tensor(10.0, requires_grad=True, dtype=torch.float64)
-        )
-
     def forward(self):
         """The forward pass of the sparse implementation--- no wavelengths needed!
 
@@ -372,12 +366,58 @@ class EchelleModel(nn.Module):
     A Model for Echelle Spectra based on the SparseEmulator
 
     wl_bin_edges (float vector): The input wavelength
-    device (Torch Device or str): GPU or CPU?
     pretrained_emulator (SparsePhoenixEmulator): A pretrained emulator to use for modeling data
+    device (Torch Device or str): GPU or CPU?
+    wing_cut_pixels (int): number of pixels for the wingcut
     """
 
-    def __init__(self, wl_bin_edges, wl_native, device=None):
+    def __init__(
+        self, wl_bin_edges, pretrained_emulator, device=None, wing_cut_pixels=None
+    ):
         super().__init__()
+
+        if device is None:
+            if torch.cuda.is_available():
+                device = "cuda"
+            else:
+                device = "cpu"
+
+        device = torch.device(device)
+
+        ## Define the wing cut
+        # Currently defined in *pixels*
+        if wing_cut_pixels is None:
+            wing_cut_pixels = 1000
+        else:
+            wing_cut_pixels = int(wing_cut_pixels)
+
+        lines = pretrained_emulator.lam_centers.detach().cpu().numpy()
+        wl_native = pretrained_emulator.wl_native.cpu().numpy()
+        print("Initializing a sparse model with {:} spectral lines".format(len(lines)))
+
+        # Find the index position of each spectral line
[…]

### 2022-01-20 experiments/02_synthetic_injection/02_synthetic_injection.py
A-ha! The injected data used wingcut of 1000 pixels, and yes, the emulator is training!

@@ -73,7 +73,7 @@ flux_native = spectrum.flux.value
 # Create the emulator and load a pretrained model
 prominence = 0.01
 emulator = SparsePhoenixEmulator(
-    wl_native, flux_native, prominence=prominence, wing_cut_pixels=100
+    wl_native, flux_native, prominence=prominence, wing_cut_pixels=1000
 )
 emulator.to(device)
 
@@ -115,7 +115,7 @@ loss_fn = nn.MSELoss(reduction="mean")
 optimizer = optim.Adam(
     list(filter(lambda p: p.requires_grad, model.parameters()))
     + list(filter(lambda p: p.requires_grad, emulator.parameters())),
-    0.01,
+    0.05,
     amsgrad=True,
 )
 n_epochs = 5000
@@ -147,6 +147,9 @@ for epoch in t_iter:
     )
     writer.add_scalar("vsini", 0.9 + np.exp(model.ln_vsini.item()), global_step=epoch)
     writer.add_scalar("RV", emulator.radial_velocity.item(), global_step=epoch)
+    writer.add_scalar(
+        "Amplitude2308", emulator.amplitudes[2308].item(), global_step=epoch
+    )
     if (epoch % plot_every_N_steps) == 0:
         # torch.save(model.state_dict(), "model_coeffs.pt")
         wl_plot = data_wavelength.cpu()
@@ -160,4 +163,6 @@ for epoch in t_iter:
             "predictions vs. actuals", plot_spectrum(to_plot), global_step=epoch,
         )
 
-torch.save(model.state_dict(), "model_T4100g3p5_prom0p01_HPF_recovery.pt")
+        torch.save(model.state_dict(), "model_T4100g3p5_prom0p01_HPF_recovery.pt")
+        torch.save(emulator.state_dict(), "emulator_T4100g3p5_prom0p01_HPF_recovery.pt")
+

### 2022-01-21 experiments/02_synthetic_injection/02_synthetic_injection.py
You can spot-check the training on individual line parameters--it works!

@@ -118,12 +118,12 @@ optimizer = optim.Adam(
     0.05,
     amsgrad=True,
 )
-n_epochs = 5000
+n_epochs = 1000
 losses = []
 
 # high_res_model = emulator.flux_native.clone().detach().to(device)
 
-plot_every_N_steps = 100
+plot_every_N_steps = 50
 t_iter = trange(n_epochs, desc="Training", leave=True)
 for epoch in t_iter:
     model.train()
@@ -148,7 +148,7 @@ for epoch in t_iter:
     writer.add_scalar("vsini", 0.9 + np.exp(model.ln_vsini.item()), global_step=epoch)
     writer.add_scalar("RV", emulator.radial_velocity.item(), global_step=epoch)
     writer.add_scalar(
-        "Amplitude2308", emulator.amplitudes[2308].item(), global_step=epoch
+        "Amplitude2085", emulator.amplitudes[2085].item(), global_step=epoch
     )
     if (epoch % plot_every_N_steps) == 0:
         # torch.save(model.state_dict(), "model_coeffs.pt")

### 2022-01-21 paper/paper1/ms.tex
Revise the text based on sparse matrix breakthrough

@@
 There are a few standardization strategies for applying the wing cuts.
-Here we coerce all wing cuts to be exactly 6000 pixels, with the middle pixel being at the line center position, and 3000 pixels to the red and blue side of the line.
+Here we coerce all wing cuts to be exactly the same number of pixels, typically 6000 pixels for \texttt{PHOENIX}, with the middle pixel being at the line center position, and about 3000 pixels to the red and blue side of the line.
 Most spectral lines only appreciably affect the signal in the central $\sim5-50$ pixels, so such a large wing cut may seem over-zealous.
 We emphasize that the line cut serves as the fixed pixel locations where the line will be evaluated with post-processing steps such as radial velocity shifts and line broadening.
-So these 6000 pixels must be capable of handling the fastest-moving and fast-spinning stars you expect to see in your observed spectra of interest.
+So these 6000 pixels must be capable of handling the fastest-moving and fastest-spinning stars you expect to see in your observed spectra of interest.
 
@@
 Most efficient CPU and GPU algorithms are not equipped to deal with ragged arrays, therefore increasing the computation time.
-Instead, the choice of exactly 6000 pixels means we can use standard matrix or tensor arithmetic, which somewhat counterintuitively is faster and handles more pixels for all spectral lines: 6000 pixels everywhere is faster than 6000 pixels for some and 2999 for other lines.
+Instead, the standardization choice of exactly 6000 pixels means we can use standard matrix or tensor arithmetic, which somewhat counterintuitively is faster while handling more pixels for all spectral lines: 6000 pixels everywhere is faster than 6000 pixels for some and 2999 for other lines.
 
-Locating the indicies of these 6000 ``nearest-neighbor'' pixels involves a sorting algorithm that takes about 0.01 seconds per spectral line, so for 6500 spectral lines the preprocessing step takes about 90 seconds.
-This initialization only occurs once and then is fixed, so we should not worry too much that it is slow.
-Heuristics for finding these indices faster are also feasible.
+Locating the indices of these 6000 ``nearest-neighbor'' pixels is very fast, only occurs once, and then is fixed.
 
@@
 
+The argument of the natural logarithm approaches 1 far from the line core, making the terms outside of the matrix approximately zero, a sparse matrix representing the log of the flux.
 Sparse matrix methods generally support an operation known as \emph{coalescing}, which sums values with repeated indices.
-Each pixel gets computed on average about $\sim100$ times in this sparse implementation, which is about $50\times$ better than each pixel getting computed $N_{lines}\sim6500$ times in the dense approach.
+The coalescing operation is extremely fast and efficient on either CPU or GPU hardware.
+We sum the logs and then exponentiate the result before comparing to the native linear flux.
+We enforce that the argument of the natural log must be greater than zero through a simple truncation, which may approximately resemble the effect of saturation in line cores.
+This truncation would only be relevant for saturated lines.
+Each pixel may get computed about $\sim100$ times in this sparse implementation, which is about $50\times$ better than each pixel getting computed $N_{lines}\sim6500$ times in the dense approach.
 
@@
 
-We use the Adam optimizer\authorcomment1{citation} with a learning rate $LR=0.004$ and all the defaults for \texttt{PyTorch v1.11}.
-We set the minibatch fraction to typically $f\sim1-5\%$.
-We defined the number of training epochs $N_{epoch}=100-1000$.
-The optimizer steps $N_{batch}$ times in each epoch, so that on average each epoch sees the entire dataset.
-The minibatching indices may be chosen with replacement so that some pixels are repeated and some are not seen in a given epoch.
+We use the Adam optimizer\authorcomment1{citation} with a typical learning rate $LR\in (0.005, 0.05)$ and all the defaults for \texttt{PyTorch v1.11}.
+Minitbatch testing proceeded with batch fractions of typically $f\sim1-5\%$ of the available pixels, but was rapidly discontinued in favor of the much more performant and convenient sparse matrix method.
+We defined the number of training epochs $N_{epoch}=100-10,000$ depending on the application.
 The user can optionally monitor a live view of the training progress with Tensorboard\authorcomment1{citation} to gain an intuition for the training efficiency.
@@
 
-We save the model parameters with \texttt{torch.save()} and refer to this collection of parameters as the \emph{pre-trained model}.
+We save the model parameters with \texttt{torch.save()} and refer to the entire collection of parameters as the \emph{pre-trained model}.
 

### 2022-01-22 experiments/03_injection_recovery_contours/03_A_run_injection_step.py
Add experiment 03

@@ -1,0 +1,99 @@
+import os
+import torch
+from blase.emulator import SparsePhoenixEmulator
+from gollum.phoenix import PHOENIXSpectrum
+import numpy as np
+from muler.hpf import HPFSpectrum
+import copy
+
+if torch.cuda.is_available():
+    device = "cuda"
+else:
+    device = "cpu"
+
+
+# Pre-process the model as described in the paper
+spectrum = PHOENIXSpectrum(teff=4100, logg=3.5)
+spectrum = spectrum.divide_by_blackbody()
+spectrum = spectrum.normalize()
+continuum_fit = spectrum.fit_continuum(polyorder=5)
+spectrum = spectrum.divide(continuum_fit, handle_meta="ff")
+
+# Fetch the real HPF data
+# For now we will use some data from another project
+fn = "Goldilocks_20210919T084302_v1.0_0063.spectra.fits"
+dir = "~/GitHub/star-witness/data/HPF/goldilocks/UT21-3-015/"
+dir = os.path.expanduser(dir)
+raw_data = HPFSpectrum(file=dir + fn, order=5)  # Just one order for now
+data = raw_data.sky_subtract().deblaze().normalize().trim_edges((8, 2040))
+
+# Numpy arrays: 1 x N_pix
+wl_native = spectrum.wavelength.value
+flux_native = spectrum.flux.value
+
+# Create the emulator and load a pretrained model
+prominence = 0.01
+emulator = SparsePhoenixEmulator(
+    wl_native, flux_native, prominence=prominence, wing_cut_pixels=1000
+)
+emulator.to(device)
+
+wl_native = emulator.wl_native.clone().detach().to(device)
+wl_active = wl_native.to("cpu")[emulator.active_mask.to("cpu").numpy()]
+target = (
+    emulator.flux_native.clone().detach().to(device)[emulator.active_mask.cpu().numpy()]
+)
+
+state_dict_init = torch.load("emulator_T4100g3p5_prom0p01_HPF.pt")
+emulator.load_state_dict(state_dict_init)
+
+emulator.radial_velocity.requires_grad = False
+emulator.lam_centers.requires_grad = False
+emulator.amplitudes.requires_grad = True
+emulator.sigma_widths.requires_grad = False
+emulator.gamma_widths.requires_grad = False
+
+# ---------------------------------------------------------
+# Injection step: Change ~20% of the amplitudes within a factor of ~2.7
[…]

### 2022-01-22 experiments/03_injection_recovery_contours/03_B_run_recovery_step.py

@@ -1,0 +1,126 @@
+import os
+import torch
+from torch import nn
+from tqdm import trange
+import torch.optim as optim
+from blase.emulator import SparsePhoenixEmulator
+import matplotlib.pyplot as plt
+from gollum.phoenix import PHOENIXSpectrum
+import numpy as np
+from muler.hpf import HPFSpectrum
+import copy
+
+if torch.cuda.is_available():
+    device = "cuda"
+else:
+    device = "cpu"
+
+# Pre-process the model as described in the paper
+spectrum = PHOENIXSpectrum(teff=4100, logg=3.5)
+spectrum = spectrum.divide_by_blackbody()
+spectrum = spectrum.normalize()
+continuum_fit = spectrum.fit_continuum(polyorder=5)
+spectrum = spectrum.divide(continuum_fit, handle_meta="ff")
+
+# Fetch the real HPF data
+# For now we will use some data from another project
+fn = "Goldilocks_20210919T084302_v1.0_0063.spectra.fits"
+dir = "~/GitHub/star-witness/data/HPF/goldilocks/UT21-3-015/"
+dir = os.path.expanduser(dir)
+raw_data = HPFSpectrum(file=dir + fn, order=5)  # Just one order for now
+data = raw_data.sky_subtract().deblaze().normalize().trim_edges((8, 2040))
+
+# Numpy arrays: 1 x N_pix
+wl_native = spectrum.wavelength.value
+flux_native = spectrum.flux.value
+
+# Create the emulator and load a pretrained model
+prominence = 0.01
+emulator = SparsePhoenixEmulator(
+    wl_native, flux_native, prominence=prominence, wing_cut_pixels=1000
+)
+emulator.to(device)
+
+wl_native = emulator.wl_native.clone().detach().to(device)
+wl_active = wl_native.to("cpu")[emulator.active_mask.to("cpu").numpy()]
+target = (
+    emulator.flux_native.clone().detach().to(device)[emulator.active_mask.cpu().numpy()]
+)
+
+state_dict_post = torch.load("emulator_T4100g3p5_prom0p01_HPF.pt")
+emulator.load_state_dict(state_dict_post)
+
+emulator.radial_velocity.requires_grad = False
+emulator.lam_centers.requires_grad = False
+emulator.amplitudes.requires_grad = True
+emulator.sigma_widths.requires_grad = False
+emulator.gamma_widths.requires_grad = False
+
+# ---------------------------------------------------------
+# Synthetic Data training
[…]

### 2022-01-22 experiments/03_injection_recovery_contours/03_C_make_result_plot.py

@@ -1,0 +1,61 @@
+from mimetypes import init
+from os import stat
+import torch
+import numpy as np
+import matplotlib.pyplot as plt
+
+state_dict_init = torch.load("emulator_T4100g3p5_prom0p01_HPF.pt")
+state_dict_injected = torch.load("./emulator_T4100g3p5_prom0p01_HPF_injection.pt")
+dict_of_recovered = torch.load("./emulator_T4100g3p5_prom0p01_HPF_recovery.pt")
+
+# Remove the lines outside the region-of-interest
+roi_mask = (state_dict_init["lam_centers"].cpu() > 8656) & (
+    state_dict_init["lam_centers"].cpu() < 8767
+)
+injected = np.exp(state_dict_injected["amplitudes"].cpu())
+initialization = np.exp(state_dict_init["amplitudes"].cpu())
+
+changed = injected != initialization
+
+mask = roi_mask & changed
+injected = injected[mask]
+initialization = initialization[mask]
+
+# Figure 1: The initialization
+plt.figure(figsize=(6, 6))
+plt.plot(injected, initialization, "o", alpha=0.2)
+plt.plot([1.0e-4, 1.0], [1.0e-4, 1.0], color="k", linestyle="dashed")
+plt.yscale("log")
+plt.xscale("log")
+plt.ylim(5e-4, 1)
+plt.xlim(5e-4, 1)
+plt.xlabel("Injected Amplitude")
+plt.ylabel("Initialized Amplitude")
+
+plt.savefig("03_initialization_injection.png", dpi=300, bbox_inches="tight")
+
+
+# Figure 2: The recovery
+plt.figure(figsize=(6, 6))
+
+n_iterations = 5
+recovered_array = np.zeros((len(injected), n_iterations))
+for i in range(5):
+    recovered_array[:, i] = np.exp(dict_of_recovered[i]["amplitudes"].cpu()[mask])
+
+mean_recovered = np.mean(recovered_array, axis=1)
+std_recovered = np.std(recovered_array, axis=1)
+
+plt.errorbar(injected, mean_recovered, yerr=std_recovered, alpha=0.2, linestyle="none")
+plt.plot(injected, mean_recovered, "o", alpha=0.2)
+plt.plot([1.0e-4, 1.0], [1.0e-4, 1.0], color="k", linestyle="dashed")
+plt.yscale("log")
+plt.xscale("log")
+plt.ylim(5e-4, 1)
[…]

### 2022-01-22 experiments/03_injection_recovery_contours/03_B_run_recovery_step.py
Making progress on Issue #14

@@ -81,7 +81,7 @@ optimizer = optim.Adam(
     0.05,
     amsgrad=True,
 )
-n_epochs = 100
+n_epochs = 1000
 losses = []
 
 # high_res_model = emulator.flux_native.clone().detach().to(device)

### 2022-01-22 experiments/03_injection_recovery_contours/03_A_run_injection_step.py
Fixed a bug and we have a working experiment 03

@@ -77,16 +77,6 @@ fn_out = "emulator_T4100g3p5_prom0p01_HPF_injection.pt"
 fake_sigma_angs = torch.tensor(0.07)
 fake_vsini = torch.tensor(4.1)
 
-with torch.no_grad():
-    rotationally_broadened = model.rotational_broaden(emulator.forward(), fake_vsini)
-    instrumentally_broadened = model.instrumental_broaden(
-        rotationally_broadened, fake_sigma_angs
-    )
-    noisefree_signal_perturbed = model.resample_to_data(instrumentally_broadened)
-
-torch.save(noisefree_signal_perturbed, "noisefree_signal_perturbed.pt")
-
-
 print("-" * 80)
 print("Injecting a known perturbation into {:0.1%} of lines".format(frac_perturbed))
 print("We are only perturbing amplitude in this experiment.")
@@ -96,4 +86,14 @@ perturbed_state_dict["amplitudes"] = perturbed_state_dict[
     "amplitudes"
 ] + offsets_perturbed.to(device)
 
+emulator.load_state_dict(perturbed_state_dict)
+
+with torch.no_grad():
+    rotationally_broadened = model.rotational_broaden(emulator.forward(), fake_vsini)
+    instrumentally_broadened = model.instrumental_broaden(
+        rotationally_broadened, fake_sigma_angs
+    )
+    noisefree_signal_perturbed = model.resample_to_data(instrumentally_broadened)
+
+torch.save(noisefree_signal_perturbed, "noisefree_signal_perturbed.pt")
 torch.save(perturbed_state_dict, fn_out)

### 2022-01-22 experiments/03_injection_recovery_contours/03_B_run_recovery_step.py

@@ -65,6 +65,10 @@ from blase.emulator import EchelleModel
 
 model = EchelleModel(data.spectral_axis.bin_edges.value, wl_native.cpu())
 model_init = copy.deepcopy(model.state_dict())
+model_injected = copy.deepcopy(model_init)
+model_injected["ln_vsini"] = torch.tensor(1.163, device=device)
+model_injected["ln_sigma_angs"] = torch.tensor(-2.8134, device=device)
+model.load_state_dict(model_injected)
 
 # Do not train on real data.  Train on Synthetic data with known properties!
 ## data_target = torch.tensor(data.flux.value, device=device, dtype=torch.float64)
@@ -78,10 +82,10 @@ loss_fn = nn.MSELoss(reduction="mean")
 optimizer = optim.Adam(
     list(filter(lambda p: p.requires_grad, model.parameters()))
     + list(filter(lambda p: p.requires_grad, emulator.parameters())),
-    0.05,
+    0.02,
     amsgrad=True,
 )
-n_epochs = 1000
+n_epochs = 200
 losses = []
 
 # high_res_model = emulator.flux_native.clone().detach().to(device)
@@ -92,12 +96,18 @@ dict_out2 = {}
 n_draws = 5
 for i in range(n_draws):
     # Draw a new noise vector each iteration
-    noise_draw = np.random.normal(loc=0, scale=0.01, size=len(noisefree_signal))
+    noise_draw = np.random.normal(loc=0, scale=0.004, size=len(noisefree_signal))
     data_target = noisefree_signal.to(device) + torch.tensor(noise_draw).to(device)
 
     # Reset the models and optimizer back to their starting place.
     emulator.load_state_dict(state_dict_post)
-    model.load_state_dict(model_init)
+    model.load_state_dict(model_injected)
+
+    emulator.radial_velocity.requires_grad = False
+    emulator.lam_centers.requires_grad = False
+    emulator.amplitudes.requires_grad = True
+    emulator.sigma_widths.requires_grad = False
+    emulator.gamma_widths.requires_grad = False
 
     optimizer = optim.Adam(
         list(filter(lambda p: p.requires_grad, model.parameters()))

### 2022-01-22 experiments/03_injection_recovery_contours/03_C_make_result_plot.py

@@ -1,5 +1,3 @@
-from mimetypes import init
-from os import stat
 import torch
 import numpy as np
 import matplotlib.pyplot as plt
@@ -23,7 +21,13 @@ initialization = initialization[mask]
 
 # Figure 1: The initialization
 plt.figure(figsize=(6, 6))
-plt.plot(injected, initialization, "o", alpha=0.2)
+plt.plot(
+    np.exp(state_dict_injected["amplitudes"].cpu()[roi_mask]),
+    np.exp(state_dict_init["amplitudes"].cpu()[roi_mask]),
+    ".",
+    alpha=0.2,
+)
+plt.plot(injected, initialization, ".", alpha=0.5)
 plt.plot([1.0e-4, 1.0], [1.0e-4, 1.0], color="k", linestyle="dashed")
 plt.yscale("log")
 plt.xscale("log")
@@ -46,9 +50,26 @@ for i in range(5):
 mean_recovered = np.mean(recovered_array, axis=1)
 std_recovered = np.std(recovered_array, axis=1)
 
-plt.errorbar(injected, mean_recovered, yerr=std_recovered, alpha=0.2, linestyle="none")
-plt.plot(injected, mean_recovered, "o", alpha=0.2)
-plt.plot([1.0e-4, 1.0], [1.0e-4, 1.0], color="k", linestyle="dashed")
+plt.plot(
+    np.exp(state_dict_injected["amplitudes"].cpu()[roi_mask]),
+    np.exp(dict_of_recovered[0]["amplitudes"].cpu()[roi_mask]),
+    ".",
+    alpha=0.2,
+)
+
+plt.errorbar(
+    injected,
+    np.exp(dict_of_recovered[0]["amplitudes"].cpu()[mask]),
+    yerr=std_recovered,
+    alpha=0.5,
+    color="k",
+    linestyle="none",
+)
+plt.plot(
+    injected, np.exp(dict_of_recovered[0]["amplitudes"].cpu()[mask]), ".", alpha=0.5
+)
+
+plt.plot([1.0e-4, 1.0], [1.0e-4, 1.0], color="k", linestyle="dashed", alpha=0.5)
 plt.yscale("log")
 plt.xscale("log")
 plt.ylim(5e-4, 1)

### 2022-01-22 src/blase/emulator.py
Start with a resolving power closer to HPF

@@ -34,11 +34,9 @@ class PhoenixEmulator(nn.Module):
         self.wl_native = torch.tensor(wl_native)
         self.flux_native = torch.tensor(flux_native)
 
-        (
-            lam_centers,
-            amplitudes,
-            widths_angstroms,
-        ) = self.detect_lines(self.wl_native, self.flux_native, prominence=prominence)
+        (lam_centers, amplitudes, widths_angstroms,) = self.detect_lines(
+            self.wl_native, self.flux_native, prominence=prominence
+        )
 
         self.n_pix = len(wl_native)
 
@@ -394,7 +392,7 @@ class EchelleModel(nn.Module):
         #    torch.tensor(45_000.0, requires_grad=True, dtype=torch.float64)
         # )
         self.ln_sigma_angs = nn.Parameter(
-            torch.tensor(-2.3, requires_grad=True, dtype=torch.float64)
+            torch.tensor(-2.8134, requires_grad=True, dtype=torch.float64)
         )
 
         self.ln_vsini = nn.Parameter(

### 2022-01-22 experiments/04_full_bandwidth_HPF/04_A_pretrain_emulator.py
Woohoo, experiment 4 can work, need to tweak continuum fitting

@@ -1,0 +1,67 @@
+import os
+import torch
+from torch import nn
+from tqdm import trange
+import torch.optim as optim
+from blase.emulator import SparsePhoenixEmulator
+import matplotlib.pyplot as plt
+from gollum.phoenix import PHOENIXSpectrum
+import numpy as np
+from muler.hpf import HPFSpectrum
+import copy
+
+if torch.cuda.is_available():
+    device = "cuda"
+else:
+    device = "cpu"
+
+# Pre-process the model as described in the paper
+spectrum = PHOENIXSpectrum(teff=4700, logg=4.5)
+spectrum = spectrum.divide_by_blackbody()
+spectrum = spectrum.normalize()
+continuum_fit = spectrum.fit_continuum(polyorder=5)
+spectrum = spectrum.divide(continuum_fit, handle_meta="ff")
+
+# Numpy arrays: 1 x N_pix
+wl_native = spectrum.wavelength.value
+flux_native = spectrum.flux.value
+
+# Create the emulator and load a pretrained model
+prominence = 0.01
+emulator = SparsePhoenixEmulator(
+    wl_native, flux_native, prominence=prominence, wing_cut_pixels=1000
+)
+emulator.to(device)
+
+wl_native = emulator.wl_native.clone().detach().to(device)
+wl_active = wl_native.to("cpu")[emulator.active_mask.to("cpu").numpy()]
+target = (
+    emulator.flux_native.clone().detach().to(device)[emulator.active_mask.cpu().numpy()]
+)
+
+emulator.radial_velocity.requires_grad = False
+emulator.lam_centers.requires_grad = False
+emulator.amplitudes.requires_grad = True
+emulator.sigma_widths.requires_grad = True
+emulator.gamma_widths.requires_grad = True
+
+loss_fn = nn.MSELoss(reduction="mean")
+
+optimizer = optim.Adam(
+    list(filter(lambda p: p.requires_grad, emulator.parameters())), 0.01, amsgrad=True,
+)
+n_epochs = 1000
+losses = []
+
+# high_res_model = emulator.flux_native.clone().detach().to(device)
+t_iter = trange(n_epochs, desc="Training", leave=True)
+for epoch in t_iter:
+    emulator.train()
+    yhat = emulator.forward()[emulator.active_mask]
+    loss = loss_fn(yhat, target)
+    loss.backward()
+    optimizer.step()
+    optimizer.zero_grad()
[…]

### 2022-01-22 experiments/04_full_bandwidth_HPF/04_B_train_on_data.py

@@ -1,0 +1,96 @@
+import os
+import torch
+from torch import nn
+from tqdm import trange
+import torch.optim as optim
+from blase.emulator import SparsePhoenixEmulator
+import matplotlib.pyplot as plt
+from gollum.phoenix import PHOENIXSpectrum
+import numpy as np
+from muler.hpf import HPFSpectrum
+import copy
+
+if torch.cuda.is_available():
+    device = "cuda"
+else:
+    device = "cpu"
+
+# Pre-process the model as described in the paper
+spectrum = PHOENIXSpectrum(teff=4700, logg=4.5)
+spectrum = spectrum.divide_by_blackbody()
+spectrum = spectrum.normalize()
+continuum_fit = spectrum.fit_continuum(polyorder=5)
+spectrum = spectrum.divide(continuum_fit, handle_meta="ff")
+
+# Fetch the real HPF data
+# For now we will use some data from another project
+fn = "Goldilocks_20210919T084302_v1.0_0063.spectra.fits"
+dir = "~/GitHub/star-witness/data/HPF/goldilocks/UT21-3-015/"
+dir = os.path.expanduser(dir)
+raw_data = HPFSpectrum(file=dir + fn, order=5)  # Just one order for now
+data = raw_data.sky_subtract().deblaze().normalize().trim_edges((8, 2040))
+
+# Numpy arrays: 1 x N_pix
+wl_native = spectrum.wavelength.value
+flux_native = spectrum.flux.value
+
+# Create the emulator and load a pretrained model
+prominence = 0.01
+emulator = SparsePhoenixEmulator(
+    wl_native, flux_native, prominence=prominence, wing_cut_pixels=1000
+)
+emulator.to(device)
+
+wl_native = emulator.wl_native.clone().detach().to(device)
+wl_active = wl_native.to("cpu")[emulator.active_mask.to("cpu").numpy()]
+target = (
+    emulator.flux_native.clone().detach().to(device)[emulator.active_mask.cpu().numpy()]
+)
+
+state_dict_post = torch.load("emulator_T4700g4p5_prom0p01_HPF.pt")
+emulator.load_state_dict(state_dict_post)
+
+emulator.radial_velocity = nn.Parameter(torch.tensor(27.6, device=device))
+emulator.radial_velocity.requires_grad = True
+emulator.lam_centers.requires_grad = False
+emulator.amplitudes.requires_grad = True
+emulator.sigma_widths.requires_grad = False
[…]

### 2022-02-22 paper/paper1/ms.tex
grammarly to the rescue

@@
 \subsection{Seeking physical self-consistency and high predictive accuracy}
-Ideally we seek models that are both physically self-consistent and highly accurate when compared to data.
+Ideally, we seek models that are both physically self-consistent and highly accurate when compared to data.
 We want tunable models that can be evaluated quickly for use in likelihood calculations.
@@
 We want high out-of-band prediction accuracy.
-Ideally we also want precision and accuracy in the separation of affiliated telluric and instrumental signals.
+Ideally, we also want precision and accuracy in the separation of affiliated telluric and instrumental signals.
 
@@
  \item We use the autodiff and GPU acceleration, but not the neural architectures
-  \item We do not use the neural architectures because we have good understand for how a spectrum is generated
+ \item We do not use the neural architectures because we have a good understand for how a spectrum is generated
  \item Instead of tuning weights of a neural network, we tune the atomic and molecular properties of a spectrum
@@
 For the purposes of this paper, we will showcase examples from two well-known families of precomputed synthetic spectra: \texttt{PHOENIX} \citep{husser13} for stellar spectra $(T_{\mathrm{eff}}\in [2300, 10000]\;K)$ and \texttt{Sonora} \citep{2021ApJ...920...85M} for brown dwarf spectra $(T_{\mathrm{eff}}\in [1000, 2300]\;K)$.
-We place the following demands on the the precomputed spectra.
+We place the following demands on the precomputed spectra.
 They should have sporadic regions of discernable continuum devoid of lines, and the continuum must vary smoothly in wavelength.
 The spectral lines or pseudo-lines should be resolved, and not sub-sampled.
-We suspect most stellar spectra meet these criteria, with the exception of the coolest M dwarfs and brown dwarfs.
+We suspect most stellar spectra meet these criteria, except for the coolest M dwarfs and brown dwarfs.
 The method can hypothetically handle spectra without discernable continua, but we suspect some modifications to the preprocessing steps would be needed.
@@
 The choice of limiting the bandwidth to a region-of-interest around a single echelle spectrograph bandwidth stems from computational constraints.
-In principle there is no fundamental limit to the bandwidth one could clone with the method presented here, up to and including the entire precomputed synthetic spectral model bandwidth.
+In principle, there is no fundamental limit to the bandwidth one could clone with the method presented here, up to and including the entire precomputed synthetic spectral model bandwidth.
 We adopt the exact native wavelength sampling with no smoothing or interpolation, yielding a wavelength vector $\bm{\lambda}_s$ with length $N_\mathrm{s}$ equal to the number of pixels within the extents of our region of interest including the buffers.
 
-At this stage we have the choice of whether to work in linear or log scale flux units.
-Adopting the log of the flux would ensure that the cloned model possesses only positive flux values, a desireable trait of any physical spectral model.
+At this stage, we have the choice of whether to work in linear or log scale flux units.
+Adopting the log of the flux would ensure that the cloned model possesses only positive flux values, a desirable trait of any physical spectral model.
 We chose instead to stick with native linear flux scaling because of the availability of convenient approximations to astrophysically relevant line profiles in the linear scaling domain.
@@
 We found a pixel distance of 50,000 and $3^{rd}$ order polynomial worked across the range of effective temperatures for HPF's bandwidth.
-The IGRINS spectra had more significant continuum opacity and bandheads to accomodate and therefore needed to capture finer scale variation: 5,000 pixels and a $9^{th}$ order polynomial appeared adequate over most of the effective temperature range.
+The IGRINS spectra had more significant continuum opacity and band heads to accommodate and therefore needed to capture finer-scale variation: 5,000 pixels and a $9^{th}$ order polynomial appeared adequate over most of the effective temperature range.
 
 This high-pass filtering step should be set to capture the genuine spectral shape, without over-fitting broad line-wings such as those in deep Hydrogen and sodium lines.
-We have experimentally found it helpful to hand-tune models of exceptionally broad and non-standard line-wings with an optional pre-processing step that depends on the wavelength range of interest.
+We have experimentally found it helpful to hand-tune models of exceptionally broad and non-standard line wings with an optional pre-processing step that depends on the wavelength range of interest.
 Here we sketch the approach for common broad lines found in the long-wavelength visible and near-infrared wavelength regions.
@@
 We emphasize that a recreation of the unvarnished input spectrum can be obtained by multiplying the continuum-flattened signal by the ``perturbed black body'', $\mathsf{B}(\bm{\lambda}_s)\cdot \mathsf{P}(\bm{\lambda}_s)$, that symbolizes the black body modulated by continuum opacity or broad-band radiative transfer effects.
-This smooth spectrum may be useful for applications that need to keep track of broad-band flux, such as low resolution spectra, or regions with molecular bandheads.
+This smooth spectrum may be useful for applications that need to keep track of broad-band flux, such as low-resolution spectra, or regions with molecular band heads.
 The ``perturbed black body'' continuum model contains $n_{\mathrm{poly}}+1$ fixed-but-possibly tunable lookup parameters, plus the fixed input $T_{\mathrm{eff}}$.
 
-Next we identify the spectral lines.
+Next, we identify the spectral lines.
 We apply a local-minimum-finding algorithm on the spectrum by defining a prominence threshold $P_{rom} \in (0.005, 0.02)$.
 This threshold dictates the number of lines that will be modeled: a lower prominence finds more, weaker lines, and a larger prominence finds fewer, deeper lines.
-The prominence algorithm successfully finds lines that reside on top of broad line wings, or unresolved bandheads provided that the individual lines exceed the prominence threshold in their local region.
+The prominence algorithm successfully finds lines that reside on top of broad line wings, or unresolved band heads provided that the individual lines exceed the prominence threshold in their local region.
 The number of lines $N_{\mathrm{lines}}$ depends on the bandwidth, prominence, and the intrinsic properties of the input spectrum, principally effective temperature and metallicity.
@@
 currently on the Gemini South Telescope on Cerro Pach\'on in Chile.
-The $R=55,000$ HPF has a native bandwidth of $8079-12785$~\AA, that we expand to $8049-12815$~\AA~including the edge buffers.
+The $R=55,000$ HPF has a native bandwidth of $8079-12785$~\AA, which we expand to $8049-12815$~\AA~including the edge buffers.
 IGRINS has two cameras for $H$ and $K$ band, with the combined spectrum spanning $14267-25217\;$\AA~ including the edge buffers and the region in-between the two cameras, all at a resolving power of $(R=45,000)$.
@@
 We see between about 2,000 and 20,000 lines depending on the $T_{\mathrm{eff}}$ and $P_{rom}$.
-HPF and IGRINS have comparable number of lines, and halving the prominence increases the number of lines by about $20-30\%$ in these ranges.
+HPF and IGRINS have a comparable number of lines, and halving the prominence increases the number of lines by about $20-30\%$ in these ranges.
 The number of lines monotonically increases towards cooler effective temperatures.
@@
 There does not exist a general-purpose, single-shot algorithm for obtaining the lineshape in the presence of overlapping spectral lines: where do the wings of one line begin and the wings of another adjacent line end?
-We therefore do not attempt to determine anything about the lineshape at this stage, and instead assume that the lines resemble a Voigt profile, with a guess width about equally split between Lorentzian and Gaussian.
+We, therefore, do not attempt to determine anything about the lineshape at this stage and instead assume that the lines resemble a Voigt profile, with a guess width about equally split between Lorentzian and Gaussian.
 
@@
 where $\mathsf{V}(\lambda, \sigma_j, \gamma_j)$ is the normalized Voigt profile with Gaussian standard deviation $\sigma$, Lorentzian half-width $\gamma$, at line center position $\lambda_c$, for the $j^{th}$ spectral line.
-The amplitude $a$ is always expected to be positive for absorption lines, and is equal to the area under the curve of the line since the Voigt profile is normalized.
-The optional Prefactor term $\mathsf{P}$ is fixed at $1$ by default, but can be allowed to vary to fine-adjust imperfections left over from the polynomial-based continuum flattening procedure.
+The amplitude $a$ is always expected to be positive for absorption lines and is equal to the area under the curve of the line since the Voigt profile is normalized.
+The optional Prefactor term $\mathsf{P}$ is fixed at $1$ by default but can be allowed to vary to fine-adjust imperfections leftover from the polynomial-based continuum flattening procedure.
 
 \subsection{Voigt profile evaluation}
-The Voigt profile $\mathsf{V}(\lambda, \sigma_j, \gamma_j)$ can be computed in exact closed form using the Fadeeva function\authorcomment1{citation}.
+The Voigt profile $\mathsf{V}(\lambda, \sigma_j, \gamma_j)$ can be computed in exact closed-form using the Fadeeva function\authorcomment1{citation}.
 Evaluation of the Fadeeva function can be computationally costly, and so approximate forms may be desirable.
@@
 Direct convolution of a sampled Gaussian and Lorentzian profile may be too costly and inexact if pixels are coarsely sampled.
-Analogously, the convolution can be achieved in the Fourier domain through multiplication of the Fast Fourier Transforms of the Gaussian and Lorentzian profiles, followed by the subsequent Inverse Fourier Transforms.
+Analogously, the convolution can be achieved in the Fourier domain through the multiplication of the Fast Fourier Transforms of the Gaussian and Lorentzian profiles, followed by the subsequent Inverse Fourier Transforms.
 The main rationale for either of these two choices would be to anticipate and align with planned convolutions such as rotational broadening and instrumental kernels.
@@
 This model has $N_{\mathrm{lines}}\times 3$ free parameters, where the center wavelength is held fixed and the amplitude, width, and lineshape are allowed to vary.
-We minimize a scalar ``goodness-of-fit'' metric, \emph{aka} loss scalar $\mathcal{L}$, chosen as the mean squared error (MSE), which is proportional to $\chi^2$, the sum of the squares of the residual vector $\mathsf{R} \equiv \mathsf{S}-\mathsf{M}$ but has no notion of per-pixel noise, since the precomputed synthetic spectrum has no uncertainty:
+We minimize a scalar ``goodness-of-fit'' metric, \emph{aka} loss scalar $\mathcal{L}$, chosen as the mean squared error (MSE), which is proportional to $\chi^2$, the sum of the squares of the residual vector $\mathsf{R} \equiv \mathsf{S}-\mathsf{M}$ but has no notion of per-pixel noise since the precomputed synthetic spectrum has no uncertainty:
 
@@
 The optimizer updates the $a_j, \sigma_j, \gamma_j$ parameters by a small fraction of the Jacobian---called the learning rate LR---towards the direction that would improve the fit, for all $N_{\mathrm{lines}} \times 3$ parameters simultaneously.
-The Jacobian is calculated behind-the-scenes with automatic differentiation\authorcomment1{Citation} implemented as the so-called backwards-propagation algorithm or simply ``back-prop''.
-We choose the \texttt{PyTorch} framework that computes these Jacobians efficiently for all of the mathematical primatives in our \texttt{blas\'e} implementation.
+The Jacobian is calculated behind the scenes with automatic differentiation\authorcomment1{Citation} implemented as the so-called backpropagation algorithm or simply ``backprop''.
+We choose the \texttt{PyTorch} framework that computes these Jacobians efficiently for all of the mathematical primitives in our \texttt{blas\'e} implementation.
 
-In principle we could compute the Jacobians analytically---since the partial derivatives of the Voigt function are known in closed-form---circumventing the vendor lock-in and memory overhead of an autodiff implementation with PyTorch.
+In principle, we could compute the Jacobians analytically---since the partial derivatives of the Voigt function are known in closed-form---circumventing the vendor lock-in and memory overhead of an autodiff implementation with PyTorch.
 Further, an autodiff-aware implementation of the \emph{celerit\'e} algorithm\authorcomment1{add DFM citation} does not yet exist in PyTorch, constricting our choices for simultaneous continuum fitting.
@@
 First, we want all the parameters to be positive, forbidding negative amplitudes and negative widths.
-We therefore tune the natural log of the parameters, and exponentiate them before inclusion in Equation \ref{equation1}.
+We, therefore, tune the natural log of the parameters and exponentiate them before inclusion in Equation \ref{equation1}.
 Second, we found through iterative experimentation that the initialization amplitudes and widths were systematically shifted from the optimized values.
@@
 Modern NVIDIA GPUs have $8-40$ GB of RAM.
-The memory bottleneck is even more pernicious than mere storage of the $\bm{\bar{V}}$ matrix, since the CPU/GPU also has to store the computation graph for autodiff.
+The memory bottleneck is even more pernicious than mere storage of the $\bm{\bar{V}}$ matrix since the CPU/GPU also has to store the computation graph for autodiff.
 The computation graph takes up more memory than the matrix itself, so it is generally not possible to evaluate Equation \ref{equation1} in its entirety in one-fell-swoop.
@@
 Minibatches act as a form of regularization, the principal source of stochasticity in the Stochastic Gradient Descent algorithm, which tends to have better convergence than full-batch Gradient Descent\authorcomment1{citation}.
-The Machine Learning and Artificial Intelligence (ML/AI) community employs minibatches to handle similar situations, in which the dataset volume exceeds the available GPU RAM.
-Our configuration departs slightly from the typical ML/AI situation, since their models are usually ``small'' relative to their datasets.
+The Machine Learning/Artificial Intelligence (ML/AI) community employs minibatches to handle similar situations, in which the dataset volume exceeds the available GPU RAM.
+Our configuration departs slightly from the typical ML/AI situation since their models are usually ``small'' relative to their datasets.
 Here our model is ``big'' relative to our dataset.
@@
 The choice to evaluate only a portion of lines at a time would mean the model is inaccurately evaluated at all wavelength pixels.
-Instead we choose to evaluate all lines, but only on a subset $N_{\mathrm{batch}}$ of the total pixels $N_s$, so that the model can eventually converge to exact at those points.
+Instead, we choose to evaluate all lines, but only on a subset $N_{\mathrm{batch}}$ of the total pixels $N_s$, so that the model can eventually converge to exact at those points.
 All lines are allowed to update at each glimpse of a minibatch, but many lines with cores far from minibatch pixels will provide only weak information about how the loss scalar changes for their parameters.
@@
 Another solution to the computation problem is to take advantage of the sparsity of the $\bm{\bar{V}}$ matrix: most of the Voigt profile's entries far from the line center are vanishingly close to zero.
-Ideally we would only evaluate the model close to the line center position, some number of line cuts $N_{\mathrm{cut}}=\alpha \cdot \mathrm{FWHM}$ away.
+Ideally, we would only evaluate the model close to the line center position, some number of line cuts $N_{\mathrm{cut}}=\alpha \cdot \mathrm{FWHM}$ away.
 These line cuts produce a speedup by a factor of $\frac{N_s}{N_{\mathrm{cut}}}$, which can exceed $100\times$ for wide bandwidth spectra.
@@
 The $\alpha$ should be preselected before the training process.
-If the FWHM initialization was poor enough, the question may arise how autodiff training will cope with the poor performance of a model, leading to the prospect of strange and difficult to diagnose artifacts near spurious line-cut truncation artifacts.
+If the FWHM initialization was poor enough, the question may arise how autodiff training will cope with the poor performance of a model, leading to the prospect of strange and difficult-to-diagnose artifacts near spurious line-cut truncation artifacts.
 
@@
 
-There are a few standardization strategies for applying the wing cuts.
-Here we coerce all wing cuts to be exactly the same number of pixels, typically 6000 pixels for \texttt{PHOENIX}, with the middle pixel being at the line center position, and about 3000 pixels to the red and blue side of the line.
+There are a few standardization strategies for applying wing cuts.
+Here we coerce all wing cuts to be the same number of pixels, typically 6000 pixels for \texttt{PHOENIX}, with the middle pixel being at the line center position, and about 3000 pixels to the red and blue side of the line.
 Most spectral lines only appreciably affect the signal in the central $\sim5-50$ pixels, so such a large wing cut may seem over-zealous.
@@
 The PHOENIX spectra double their pixel sampling shortward of $\lambda = 1 \mu$m, so 6000 pixels in the visible is about 30 Angstroms, but 6000 pixels in the infrared is about 60 Angstroms.
-We \emph{could have} demanded that the wingcut be exactly 30 Å, or some other heuristic, but that choice would result in so-called \emph{jagged} or \emph{ragged} arrays: collections of 1D arrays that differ in their lengths.
+We \emph{could have} demanded that the wing cut be exactly 30 $\AA$, or some other heuristic, but that choice would result in so-called \emph{jagged} or \emph{ragged} arrays: collections of 1D arrays that differ in their lengths.
 Some lines would have 5999 pixels, some would have 6000, and others would have 2999 or 3000 pixels, depending whether they occur to the red or blue side of the $1 \;\mu$m kink in sampling.
@@
 The coalescing operation is extremely fast and efficient on either CPU or GPU hardware.
-We sum the logs and then exponentiate the result before comparing to the native linear flux.
+We sum the logs and then exponentiate the result before comparing it to the native linear flux.
 We enforce that the argument of the natural log must be greater than zero through a simple truncation, which may approximately resemble the effect of saturation in line cores.
@@
 
-Extremely broad lines will exhibit truncation effects if the sparse window is small compared to the linewing size.
+Extremely broad lines will exhibit truncation effects if the sparse window is small compared to the line wing size.
 The truncation effects will look like tophat functions severing the asymptotic wings, imbuing artificial step function kinks in the emulated spectrum.
@@
 We then construct and evaluate the entire dense matrix for those lines: $\sim 330\;000 \times 20$.
-The number of FLOPS in each category scales as about 6 Million for the 20 broad lines versus about 36 Million for the sea of about 6500 narrow lines, depending on the exact choices for wingcuts and number of lines.
+The number of FLOPS in each category scales as about 6 Million for the 20 broad lines versus about 36 Million for the sea of about 6500 narrow lines, depending on the exact choices for wing cuts and number of lines.
 
@@
 We use the Adam optimizer\authorcomment1{citation} with a typical learning rate $LR\in (0.005, 0.05)$ and all the defaults for \texttt{PyTorch v1.11}.
-Minitbatch testing proceeded with batch fractions of typically $f\sim1-5\%$ of the available pixels, but was rapidly discontinued in favor of the much more performant and convenient sparse matrix method.
+Minitbatch testing proceeded with batch fractions of typically $f\sim1-5\%$ of the available pixels but was rapidly discontinued in favor of the much more performant and convenient sparse matrix method.
 We defined the number of training epochs $N_{epoch}=100-10,000$ depending on the application.
@@
 \subsection{Line data clustering and conspicuous flaws}
-We see some narrow lines devolve into missing continuum opacity, the favored tradeoff when the continuum estimate's poor performance outweighs the the pain of a narrow-but-tolerably-small spike.
+We see some narrow lines devolve into missing continuum opacity, the favored tradeoff when the continuum estimate's poor performance outweighs the pain of a narrow-but-tolerably-small spike.
 This problem can be seen in Figure \ref{fig_poor_cloning_performance} where a line initialized at $\censor{10210}\;$\AA~ with an amplitude of ZZ and width of YY bloats to \censor{some other value}.
-These devolved lines tend to appear at poorly immitated broad-and-deep line wings.
+These devolved lines tend to appear at poorly imitated broad-and-deep line wings.
 One can mitigate these devolved lines by adding more flexibility to the continuum or supporting more advanced line shapes.
@@
  \1 How to re-sample into the data space?
-  \1 Option 1: Convolve the Pseudo-Voigt with a Instrumenal Gaussian
+ \1 Option 1: Convolve the Pseudo-Voigt with a Instrumental Gaussian
  \2 The convolution of a Gaussian with a Pseudo-Voigt is simply a new Pseudo-Voigt!
@@
  \1 Both options should give the same answer (right?)
-  \2 Need to double-check commutivity of convolution and element-wise product...
+ \2 Need to double-check commutativity of convolution and element-wise product...
  \1 Ideally you need wavelength-dependent resolution, not just a single stock resolving power
@@
  \2 Plot of cloned FWHM versus FWHM transferred
-  \2 The pseudo-voigt line properties can be analytically integrated to give an equivalent width
+ \2 The pseudo-Voigt line properties can be analytically integrated to give an equivalent width
  \2 Plot of cloned EW (before) versus transfer EW (after)
@@
  \2 To First order the lines are Gaussian or Lorentzian
-  \2 To Second order the lines are Voigt Profiles
+ \2 To Second-order the lines are Voigt Profiles
  \2 In detail the lines are the result of radiative transfer that can smear the Voigt profile

### 2022-02-22 paper/paper1/ms.tex
write the paper introduction

@@
 \shortauthors{Gully-Santiago}
-\title{Transfer learning for echelle spectroscopy}
+\title{Transfer learning for \'echelle spectroscopy with \texttt{blas\'e}}
 
 \author{Michael Gully-Santiago}
+\affiliation{University of Texas at Austin Department of Astronomy}
+
+\author{Caroline V. Morley}
 \affiliation{University of Texas at Austin Department of Astronomy}
@@
 
-Below is a table exploring recent techniques for spectral fitting.
+Tens of thousands or more individial spectral lines give rise to a sea of undulations that imbue each stellar spectrum with its characteristic appearance.
+The identification and understanding of these lines has consumed a chief discipline of astrophysics over the last century.
+The field grew from by-eye catalogs of stellar templates (cite \censor{XX} Cecilia Payne, Cannon) to modern pre-computed synthetic forward models including millions or billions of lines (cite \censor{XX}, Husser, Marley, HITRAN, HITEMP, Exomol, radis, Hargraves).
+But as technology has improved, our data and models have become more voluminous, more precise, and more complicated.
+The mere act of comparing models to data has become a computational challenge as much as a scientific one.
+Here we introduce a new state-of-the-art framework \texttt{blas\'e} aimed at solving computational, statistical, and scientific challenges associated with data-model comparisons for modern astronomical spectroscopy.
+
+The \texttt{blas\'e} framework aims to take in an observed astronomical spectrum and report back a semi-empirical noise-free template spectrum, replete with the postition, amplitude, width, and shape of all of its spectral lines, automatically.
+Solving this problem is hard, for at least five reasons.
+First, the spectral lines may overlap, and so the assignment of one line may be partially degenerate with the assignment of some other adjacent line.
+Second, the lines exhibit a large dynamic range in their allowed extents, with extremely wide line wings such as Sodium or Hydrogen lines blending into the continuum, such that the placement of the continuum level may become ill-defined.
+Third, the extent of line blending and realized line shape depends strongly on the spectral resolution of the spectrograph, the rotational broadening of the star, and possibly the instrumental configuration at the time of observation.
+Fourth, telluric absorption lines commingle with the astronomical spectral lines of interest, censoring some spectral regions entirely, or partially confounding other lines with chance allignments.
+And finally, our prediction for where lines should be located may be incomplete, and the decision of where to place lines becomes an uncertain choice that depends on the conditions and assumptions of the surrounding lines.
+
+Luckily, many scientific applications in astrophysics do not need the technically demanding noise-free template, nor catalog-of-all-spectral-lines.
+A few lines suffice.
+For those applications, human inspection of isolated lines and semi-automated equivalent width determination has been and will remain adequate.
+But many new and important questions in the fields of stars and exoplanets aspire to reach the margins of what the data can inform.
+For example, exoplanet cross-correlation spectroscopy (cite \censor{xx} Jayne Birkby, etc.) hinges on accurate templates to detect and characterize the atmospheres of exoplanets.
+Imperfections in these templates can mute the perceived signal strength of these atmospheric features.
+In precision-RV applications, the joint modeling of microtellurics may require an accurate semi-empirical model of both the star and the Earth's atmospheric absorption at for that exact moment in time.
+In the case of Doppler imaging, an accurate underlying template is needed to detect longitudinally symmetric structures such as polar spots (cite \censor{XX} Vogt, Roettenbacher) or zonal bands (cite \censor{XX}, Crossfield, Apai).
+The \texttt{blas\'e} framework is aimed at these and many other applications for which an accurate noise-free template is desired.
 
@@
 
-The PHOENIX spectra double their pixel sampling shortward of $\lambda = 1 \mu$m, so 6000 pixels in the visible is about 30 Angstroms, but 6000 pixels in the infrared is about 60 Angstroms.
-We \emph{could have} demanded that the wing cut be exactly 30 $\AA$, or some other heuristic, but that choice would result in so-called \emph{jagged} or \emph{ragged} arrays: collections of 1D arrays that differ in their lengths.
+The PHOENIX spectra double their pixel sampling shortward of $\lambda = 1 \mu$m, so 6000 pixels in the visible is about 30 Angstroms, but 6000 pixels in the infrared is about 60 \AA.
+We \emph{could have} demanded that the wing cut be exactly 30 \AA, or some other heuristic, but that choice would result in so-called \emph{jagged} or \emph{ragged} arrays: collections of 1D arrays that differ in their lengths.
 Some lines would have 5999 pixels, some would have 6000, and others would have 2999 or 3000 pixels, depending whether they occur to the red or blue side of the $1 \;\mu$m kink in sampling.

### 2022-02-22 paper/paper1/ms.tex
Add comparison to previous frameworks in the introductions, maybe a bit too verbose at the moment

@@
  We introduce blas\'e, a framework for transfer learning for high-grasp echelle spectroscopy.
+\texttt{blas\'e} produces extensible models that are both physically informed and highly accurate when compared to data.
+These models can be evaluated quickly for use in likelihood calculations.
+It works on data with large spectral grasp: simultaneously wide bandwidth and high spectral resolution.
 
@@
 
-The \texttt{blas\'e} framework aims to take in an observed astronomical spectrum and report back a semi-empirical noise-free template spectrum, replete with the postition, amplitude, width, and shape of all of its spectral lines, automatically.
+The \texttt{blas\'e} framework aims to take in an observed astronomical spectrum and report back an extensible semi-empirical spectral forward model, replete with the postition, amplitude, width, and shape of all of its spectral lines, automatically.
 Solving this problem is hard, for at least five reasons.
@@
 
+Addressing these challenges forms the backbone of spectral calibration, an increasingly valuable specialty as the deficits in our models become intolerable with greater data quantity and quality.
 Luckily, many scientific applications in astrophysics do not need the technically demanding noise-free template, nor catalog-of-all-spectral-lines.
@@
 For those applications, human inspection of isolated lines and semi-automated equivalent width determination has been and will remain adequate.
+
 But many new and important questions in the fields of stars and exoplanets aspire to reach the margins of what the data can inform.
@@
 Imperfections in these templates can mute the perceived signal strength of these atmospheric features.
-In precision-RV applications, the joint modeling of microtellurics may require an accurate semi-empirical model of both the star and the Earth's atmospheric absorption at for that exact moment in time.
+In extreme-precision-RV (EPRV) applications, the joint modeling of microtellurics may require an accurate semi-empirical model of both the star and the Earth's atmospheric absorption at that fleeting moment in time.
 In the case of Doppler imaging, an accurate underlying template is needed to detect longitudinally symmetric structures such as polar spots (cite \censor{XX} Vogt, Roettenbacher) or zonal bands (cite \censor{XX}, Crossfield, Apai).
-The \texttt{blas\'e} framework is aimed at these and many other applications for which an accurate noise-free template is desired.
+The \texttt{blas\'e} framework is suitable for these and many other applications for which an accurate noise-free template is desired, or for which the properties of a large number of individual spectral lines is sought.
 
-\movetabledown=13mm
-\begin{rotatetable}
- \begin{deluxetable}{chc}
- \tablecaption{Annotated bibliography for intro\label{table1}}
- \tablehead{
- \colhead{Reference} & \nocolhead{two} & \colhead{Key idea}
- }
- \startdata
- \citet{2015PhDT........82P} & - & \texttt{specmatch-syn} Synthetic spectral template matching \\
- \citet{czekala15} & - & \texttt{starfish} robust spectral inference on precomputed models \\
- \citet{2016ApJS..223....8C} & - & \texttt{sick} Spectral Inference Crank, similar to \texttt{Starfish}\\
- \citet{2017ApJ...836..200G} & - & Modular extensions to \texttt{starfish}: starspots \\
- \citet{2018ApJ...862...85G} & - & Modular extensions to \texttt{starfish}: veiling \\
- \citet{2017ApJ...836...77Y} & - & \texttt{specmatch-emp} Empirical template matching for visible\\
- \citet{exoplanet:astropy18} & - & \texttt{astropy} General purpose tools for spectral manipulation\\
- \citet{2019AJ....158..164B} & - & \texttt{wobble}: data driven stellar and telluric models
- \enddata
- \end{deluxetable}
-
-\end{rotatetable}
-
-\subsection{Seeking physical self-consistency and high predictive accuracy}
-Ideally, we seek models that are both physically self-consistent and highly accurate when compared to data.
-We want tunable models that can be evaluated quickly for use in likelihood calculations.
-We want posterior inference to accurately explore correlations among parameters.
-We want unbiased, noise-free templates for use in cross-correlation spectroscopy.
-We want accuracy over a large spectral grasp: simultaneously wide bandwidth and high spectral resolution.
-We want high out-of-band prediction accuracy.
-Ideally, we also want precision and accuracy in the separation of affiliated telluric and instrumental signals.
-
-These goals form the backbone of spectral calibration, an increasingly valuable specialty as the deficits in our models become intolerable with greater data quantity and quality.
-
-\subsection{Current limitations to our spectral models}
-
-Several practical barriers limit the accuracy of spectral models.
+Existing frameworks have already overcome aspects of these challenges, or have been purpose-built for specific applications.
+The \texttt{specmatch} template matching tool produces noise-free nearest neighbor templates given an input spectrum \citep{2015PhDT........82P}.
+Several practical barriers limit the accuracy of using precomputed synthetic spectral models.
 First and foremost, real stars are usually more complicated than our simplified models of them.
@@
 Conspicuous examples of these hidden variables can be found in protostars: starspots, accretion veiling, dust extinction, and magnetic Zeeman splitting.
-Jointly modeling all of these phenomena alongside the intrinsic stellar photosphere and High latent dimensionality: Many effects matter at the precision level of the data...
+Jointly modeling all of these phenomena alongside the intrinsic stellar photosphere is challenging.
 
-Second, Degeneracy among parameters
+The empirical version, \texttt{specmatch-emp} \citep{2017ApJ...836...77Y} matched spectra better than the synthetic templates, but was still too rigid for some applications and required the assembly of hundreds of standardized high signal-to-noise-ratio templates, ideally with low intrinsic rotational broadening.
+Such a large number of high-quality templates has not yet been established in the near infrared.
 
-\begin{itemize}
- \item Underlying atomic and molecular data may be wrong or approximate or missing
- \item Underlying radiative transfer may be inaccurate or approximate (T-P profile / T-Tau)
- \item Time is a dimension: Star is changing
- \item Stellar surface inhomogeneities become important: Doppler Imaging and limb darkening
- \item We do not know how to accurately parameterize some dimensions
-\end{itemize}
+The \texttt{wobble} framework \citep{2019AJ....158..164B}modernized the construction of high-SNR templates to account for temporally variable telluric lines.
+The tool requires dozens of high-SNR spectra acquired at a range of Barycentric Earth Radial Velocities (BERVs).
+The final telluric-free combined spectrum would still have to be compared to models for absolute calibration, or can be used out-of-the-box for precision relative RVs.
+The \texttt{wobble} framework also pioneered the off-label application of automatic differentiation frameworks---in this case \texttt{TensorFlow}---towards their physically-motivated use in stellar spectra.
+\texttt{blas\'e} can be viewed as an evaluable and interpretable super-resolution version of \texttt{wobble}, that accepts more bias in the bias-variance tradeoff.
 
-\subsection{High dynamic range is both a blessing and a curse:}
-\begin{itemize}
- \item Feeding back locations of model imperfections becomes unwieldy for many lines
- \item Many lines essentially becomes a book-keeping and continuum assignment problem
- \item Computational problem
- \item Some lines have large line wings, blurring continuum and line
- \item Line blanketing has almost no continuum
-\end{itemize}
+The \texttt{starfish} framework \citep{czekala15} provides a robust likelihood function for data-model comparisons, and retires many of the problems in this domain.
+\texttt{starfish} pioneered the use of whole-spectrum fitting with imperfect models.
+In particular, it addressed the problem of what to do when the underlying atomic and molecular data was wrong or approximate or missing.
+It has been extended to inferring starspot physical properties \citep{2017ApJ...836..200G}, measuring veiling in Class 0 protostars \citep{2018ApJ...862...85G}, and quantifying imperfections in brown dwarf models (cite \censor{ZJ BD paper}).
+The Spectral Inference Crank, \texttt{sick}, shares similar aims as \texttt{starfish}, and provides additional useful grid search capabilities.
+
+For very large bandwidths and very many spectral lines, the problem of identifying and cataloging line imperfections essentially becomes a book-keeping and continuum assignment problem.
+\texttt{blas\'e} and \texttt{starfish} provide different strategies for orchestrating the line-mismatch identification procedure, with each route having tradeoffs depending on the application.
 

### 2022-02-22 paper/paper1/ms.tex
missing space

@@
 
-The \texttt{wobble} framework \citep{2019AJ....158..164B}modernized the construction of high-SNR templates to account for temporally variable telluric lines.
+The \texttt{wobble} framework \citep{2019AJ....158..164B} modernized the construction of high-SNR templates to account for temporally variable telluric lines.
 The tool requires dozens of high-SNR spectra acquired at a range of Barycentric Earth Radial Velocities (BERVs).

### 2022-02-23 paper/paper1/ms.tex
Add the astronomy unified thesaurus UAT keywords, probably too many
fill in the references and edit the introduction from yesterday

@@
 
-\keywords{High resolution spectroscopy (2096)}
+\keywords{High resolution spectroscopy (2096), Stellar spectral lines (1630), Astronomy data modeling(1859), GPU Computing (1969), Calibration (2179), Radial Velocity (1332), Maximum likelihood estimation (1901), Deconvolution (1910), Atomic spectroscopy (2099), Stellar photospheres (1237)},
 
@@
 
-Tens of thousands or more individial spectral lines give rise to a sea of undulations that imbue each stellar spectrum with its characteristic appearance.
-The identification and understanding of these lines has consumed a chief discipline of astrophysics over the last century.
-The field grew from by-eye catalogs of stellar templates (cite \censor{XX} Cecilia Payne, Cannon) to modern pre-computed synthetic forward models including millions or billions of lines (cite \censor{XX}, Husser, Marley, HITRAN, HITEMP, Exomol, radis, Hargraves).
-But as technology has improved, our data and models have become more voluminous, more precise, and more complicated.
-The mere act of comparing models to data has become a computational challenge as much as a scientific one.
-Here we introduce a new state-of-the-art framework \texttt{blas\'e} aimed at solving computational, statistical, and scientific challenges associated with data-model comparisons for modern astronomical spectroscopy.
+Tens of thousands or more individual spectral lines give rise to a sea of undulations that imbue each stellar spectrum with its characteristic appearance.
+The identification and understanding of these lines has defined a large category of astrophysics over the last century.
+The field grew from by-eye catalogs of stellar templates \citep{1901AnHar..28..129C} to quantifying the role of atomic ionization balance \citep{1925PhDT.........1P}, to modern synthetic forward models including millions or billions of lines \citep{husser13, 2021ApJ...920...85M}.
+As technology has improved, our data and models have become more voluminous, more precise, and more complicated.
+The mere act of comparing models to observed spectra can resemble a computational statistics challenge as much as a scientific one.
+Here we introduce a machine-learning-based framework \texttt{blas\'e} aimed at solving computational, statistical, and scientific challenges associated with data-model comparisons for modern astronomical spectroscopy.
 
 The \texttt{blas\'e} framework aims to take in an observed astronomical spectrum and report back an extensible semi-empirical spectral forward model, replete with the postition, amplitude, width, and shape of all of its spectral lines, automatically.
-Solving this problem is hard, for at least five reasons.
+Solving this problem is hard, for at least four reasons.
 First, the spectral lines may overlap, and so the assignment of one line may be partially degenerate with the assignment of some other adjacent line.
-Second, the lines exhibit a large dynamic range in their allowed extents, with extremely wide line wings such as Sodium or Hydrogen lines blending into the continuum, such that the placement of the continuum level may become ill-defined.
+Second, extremely wide line wings blend into the continuum, such that the placement of the continuum level may become ill-defined.
 Third, the extent of line blending and realized line shape depends strongly on the spectral resolution of the spectrograph, the rotational broadening of the star, and possibly the instrumental configuration at the time of observation.
-Fourth, telluric absorption lines commingle with the astronomical spectral lines of interest, censoring some spectral regions entirely, or partially confounding other lines with chance allignments.
-And finally, our prediction for where lines should be located may be incomplete, and the decision of where to place lines becomes an uncertain choice that depends on the conditions and assumptions of the surrounding lines.
+Finally, telluric absorption lines commingle with the astronomical spectral lines of interest, censoring some spectral regions entirely, or partially confounding other lines with chance allignments.
 
-Addressing these challenges forms the backbone of spectral calibration, an increasingly valuable specialty as the deficits in our models become intolerable with greater data quantity and quality.
+Addressing these and other challenges forms the backbone of spectral calibration, an increasingly valuable specialty as the deficits in our models become intolerable with greater data quantity and quality.
 Luckily, many scientific applications in astrophysics do not need the technically demanding noise-free template, nor catalog-of-all-spectral-lines.
@@
 But many new and important questions in the fields of stars and exoplanets aspire to reach the margins of what the data can inform.
-For example, exoplanet cross-correlation spectroscopy (cite \censor{xx} Jayne Birkby, etc.) hinges on accurate templates to detect and characterize the atmospheres of exoplanets.
-Imperfections in these templates can mute the perceived signal strength of these atmospheric features.
-In extreme-precision-RV (EPRV) applications, the joint modeling of microtellurics may require an accurate semi-empirical model of both the star and the Earth's atmospheric absorption at that fleeting moment in time.
-In the case of Doppler imaging, an accurate underlying template is needed to detect longitudinally symmetric structures such as polar spots (cite \censor{XX} Vogt, Roettenbacher) or zonal bands (cite \censor{XX}, Crossfield, Apai).
-The \texttt{blas\'e} framework is suitable for these and many other applications for which an accurate noise-free template is desired, or for which the properties of a large number of individual spectral lines is sought.
+For example, exoplanet cross-correlation spectroscopy \citep{2013MNRAS.436L..35B} hinges on accurate molecular spectral templates to detect and characterize the atmospheres of exoplanets.
+Imperfections in these templates can mute the perceived signal strength of these atmospheric features \citep{2015A&A...575A..20H}.
+In extreme precision radial velocity (EPRV) applications, addressing telluric absorption at the $\mathrm{cm/s}$ level may require joint modeling of the star and the Earth's atmospheric absorption before convolution with an instrumental kernel\footnote{\url{https://speakerdeck.com/dwhgg/telluric-line-hack-week-wrap-up}}.
+In the case of Doppler imaging, an accurate underlying template is needed to detect longitudinally symmetric structures such as polar spots \citep{1983PASP...95..565V, roettenbacher16} or zonal bands \citep{Crossfield14,2021ApJ...906...64A}.
+The \texttt{blas\'e} framework is suitable for these and many other applications for which an accurate \emph{evaluable} noise-free template is desired, or for which the properties of a large number of individual spectral lines is sought.
 
 Existing frameworks have already overcome aspects of these challenges, or have been purpose-built for specific applications.
-The \texttt{specmatch} template matching tool produces noise-free nearest neighbor templates given an input spectrum \citep{2015PhDT........82P}.
-Several practical barriers limit the accuracy of using precomputed synthetic spectral models.
+The \texttt{specmatch} synthetic template matching tool produces noise-free nearest neighbor templates given an input spectrum \citep{2015PhDT........82P}.
+Several practical barriers limit the accuracy of using precomputed synthetic spectral models alone.
 First and foremost, real stars are usually more complicated than our simplified models of them.
-Real spectra often vary over more dimensions than specified.
+Real spectra often vary over more dimensions that our models do.
 Conspicuous examples of these hidden variables can be found in protostars: starspots, accretion veiling, dust extinction, and magnetic Zeeman splitting.
@@
 
-The empirical version, \texttt{specmatch-emp} \citep{2017ApJ...836...77Y} matched spectra better than the synthetic templates, but was still too rigid for some applications and required the assembly of hundreds of standardized high signal-to-noise-ratio templates, ideally with low intrinsic rotational broadening.
-Such a large number of high-quality templates has not yet been established in the near infrared.
+The empirical version, \texttt{specmatch-emp} \citep{2017ApJ...836...77Y} matched spectra better than the synthetic templates, but is still too rigid for some applications and requires the assembly of hundreds of standardized high signal-to-noise-ratio templates, ideally with low intrinsic rotational broadening.
+Such a large number of high-quality templates has not yet been established in the near infrared, and may not be feasible for brown dwarfs.
 
@@
 The \texttt{starfish} framework \citep{czekala15} provides a robust likelihood function for data-model comparisons, and retires many of the problems in this domain.
-\texttt{starfish} pioneered the use of whole-spectrum fitting with imperfect models.
-In particular, it addressed the problem of what to do when the underlying atomic and molecular data was wrong or approximate or missing.
-It has been extended to inferring starspot physical properties \citep{2017ApJ...836..200G}, measuring veiling in Class 0 protostars \citep{2018ApJ...862...85G}, and quantifying imperfections in brown dwarf models (cite \censor{ZJ BD paper}).
-The Spectral Inference Crank, \texttt{sick}, shares similar aims as \texttt{starfish}, and provides additional useful grid search capabilities.
+\texttt{starfish} pioneered the use of whole-spectrum fitting with resilience to model imperfections by addressing the problem of what to do when the underlying atomic and molecular data was wrong or approximate or missing.
+It has been extended to inferring starspot physical properties \citep{2017ApJ...836..200G}, measuring veiling in Class 0 protostars \citep{2018ApJ...862...85G}, and quantifying imperfections in brown dwarf models \citep{2021ApJ...921...95Z}.
+The Spectral Inference Crank \citep[\texttt{sick},][]{2016ApJS..223....8C} shares similar aims as \texttt{starfish}, and provides additional useful grid search capabilities.
 

### 2022-02-23 paper/paper1/ms.tex
Refining the introduction

@@
 
-Addressing these and other challenges forms the backbone of spectral calibration, an increasingly valuable specialty as the deficits in our models become intolerable with greater data quantity and quality.
+Addressing these and other challenges forms the backbone of \emph{spectral calibration}, an increasingly valuable specialty as the deficits in our models become intolerable with greater data quantity and quality.
 Luckily, many scientific applications in astrophysics do not need the technically demanding noise-free template, nor catalog-of-all-spectral-lines.
 A few lines suffice.
-For those applications, human inspection of isolated lines and semi-automated equivalent width determination has been and will remain adequate.
+For those applications, human inspection of isolated lines and semi-automated equivalent width determination have been and will remain adequate.
 
-But many new and important questions in the fields of stars and exoplanets aspire to reach the margins of what the data can inform.
+But many new and important questions in the fields of stars and exoplanets aspire to reach the margins of what the entire dataset can inform.
+In particular, data from \emph{high-grasp} \'echelle spectrographs possess simultaneously high spectral resolving power and high bandwidth, yielding tens of thousands or possibly millions of independent spectral data points for each star, substar, or exoplanet.
+Those applications that seek to gain signal by ``stacking'' or cross-correlating spectral lines can hypothetically gain huge boosts in the accessible signal-to-noise ratio compared to a single or few lines.
+Most manual and semi-automated methods cannot take advantage of the entire spectral bandwidth, and may fail to achieve the hypothetical promise of these high-bandwidth spectrographs.
+
 For example, exoplanet cross-correlation spectroscopy \citep{2013MNRAS.436L..35B} hinges on accurate molecular spectral templates to detect and characterize the atmospheres of exoplanets.
 Imperfections in these templates can mute the perceived signal strength of these atmospheric features \citep{2015A&A...575A..20H}.
-In extreme precision radial velocity (EPRV) applications, addressing telluric absorption at the $\mathrm{cm/s}$ level may require joint modeling of the star and the Earth's atmospheric absorption before convolution with an instrumental kernel\footnote{\url{https://speakerdeck.com/dwhgg/telluric-line-hack-week-wrap-up}}.
+In extreme precision radial velocity (EPRV) applications, cross-correlation methods work, but addressing telluric absorption at the $\mathrm{cm/s}$ level may require joint modeling of the star and the Earth's atmospheric absorption before convolution with an instrumental kernel\footnote{\url{https://speakerdeck.com/dwhgg/telluric-line-hack-week-wrap-up}}.
 In the case of Doppler imaging, an accurate underlying template is needed to detect longitudinally symmetric structures such as polar spots \citep{1983PASP...95..565V, roettenbacher16} or zonal bands \citep{Crossfield14,2021ApJ...906...64A}.
-The \texttt{blas\'e} framework is suitable for these and many other applications for which an accurate \emph{evaluable} noise-free template is desired, or for which the properties of a large number of individual spectral lines is sought.
+The \texttt{blas\'e} framework was developed as an enabling technology for these and many other applications for which an accurate \emph{evaluable} noise-free template is desired, or for which the properties of a large number of individual spectral lines is sought.
 

### 2022-02-23 paper/paper1/ms.tex
The introduction is clean now

@@
 \usepackage{amsmath,bm}
+\usepackage{hyperref}
+\Urlmuskip=0mu  plus 10mu
 
@@
 
-The \texttt{blas\'e} framework aims to take in an observed astronomical spectrum and report back an extensible semi-empirical spectral forward model, replete with the postition, amplitude, width, and shape of all of its spectral lines, automatically.
+The metaphorical holy grail of astronomical spectroscopy is a function that takes in an observed astronomical spectrum and reports back the postition, amplitude, width, and shape of all of its spectral lines, automatically, accurately, and precisely.
+The function would go further, to report back the systemic radial velocity (RV) and rotational broadening ($v\sin{i}$).
+Finally the function would---in an \emph{Inception}-like manner---provide what it believes to be the interpretable generating function that produced the data, so that we may gain insights on future examples of this or other stars.
 Solving this problem is hard, for at least four reasons.
@@
 But many new and important questions in the fields of stars and exoplanets aspire to reach the margins of what the entire dataset can inform.
-In particular, data from \emph{high-grasp} \'echelle spectrographs possess simultaneously high spectral resolving power and high bandwidth, yielding tens of thousands or possibly millions of independent spectral data points for each star, substar, or exoplanet.
-Those applications that seek to gain signal by ``stacking'' or cross-correlating spectral lines can hypothetically gain huge boosts in the accessible signal-to-noise ratio compared to a single or few lines.
-Most manual and semi-automated methods cannot take advantage of the entire spectral bandwidth, and may fail to achieve the hypothetical promise of these high-bandwidth spectrographs.
+In particular, data from high-\emph{grasp} \'echelle spectrographs possess simultaneously high spectral resolving power and high bandwidth, yielding tens of thousands or possibly millions of independent spectral resolution elements for each star, substar, or exoplanet.
+Those applications that seek to gain signal by ``stacking'' spectral lines or cross-correlating templates can hypothetically gain huge boosts in the accessible signal-to-noise ratio compared to a single or few lines.
+Most manual and semi-automated methods cannot take advantage of the entire spectral bandwidth, or rely on exact knowledge of the underlying templates and may fail to achieve the hypothetical promise of these high-bandwidth spectrographs \citep{2020AJ....160..198H}.
 
@@
 In the case of Doppler imaging, an accurate underlying template is needed to detect longitudinally symmetric structures such as polar spots \citep{1983PASP...95..565V, roettenbacher16} or zonal bands \citep{Crossfield14,2021ApJ...906...64A}.
-The \texttt{blas\'e} framework was developed as an enabling technology for these and many other applications for which an accurate \emph{evaluable} noise-free template is desired, or for which the properties of a large number of individual spectral lines is sought.
 
-Existing frameworks have already overcome aspects of these challenges, or have been purpose-built for specific applications.
-The \texttt{specmatch} synthetic template matching tool produces noise-free nearest neighbor templates given an input spectrum \citep{2015PhDT........82P}.
-Several practical barriers limit the accuracy of using precomputed synthetic spectral models alone.
-First and foremost, real stars are usually more complicated than our simplified models of them.
-Real spectra often vary over more dimensions that our models do.
-Conspicuous examples of these hidden variables can be found in protostars: starspots, accretion veiling, dust extinction, and magnetic Zeeman splitting.
-Jointly modeling all of these phenomena alongside the intrinsic stellar photosphere is challenging.
+Existing open-source frameworks have overcome some of these challenges, or have been purpose-built for specialized applications. 
+These frameworks include \texttt{ROBOSPECT} \citep{2013PASP..125.1164W}, \texttt{specmatch} \citep{2015PhDT........82P}, \texttt{specmatch-emp} \citep{2017ApJ...836...77Y}, \texttt{wobble} \citep{2019AJ....158..164B}, \texttt{starfish} \citep{czekala15}, \texttt{sick} \citep{2016ApJS..223....8C}, \texttt{psoap} \citep{2017ApJ...840...49C}, \texttt{FAL}\footnote{\url{https://github.com/pacargile/FAL}}, CHIMERA \citep{2015ApJ...807..183L}, the \texttt{Cannon} \citep{2017ApJ...836....5H},  \texttt{MINESweeper} \citep{2020ApJ...900...28C}, and recently \texttt{ExoJAX} \citep{2022ApJS..258...31K}.
+The design of these frameworks necessarily have to make a choice in the bias-variance tradeoff: is the tool more \emph{data}-driven or more \emph{model}-driven?
+The statistical tradeoff can be viewed as a concession in physical self-consistency for model flexibility: more or fewer parameters.
 
-The empirical version, \texttt{specmatch-emp} \citep{2017ApJ...836...77Y} matched spectra better than the synthetic templates, but is still too rigid for some applications and requires the assembly of hundreds of standardized high signal-to-noise-ratio templates, ideally with low intrinsic rotational broadening.
-Such a large number of high-quality templates has not yet been established in the near infrared, and may not be feasible for brown dwarfs.
+A key new enabling technology breaks the historic limits of data-model comparisons for astronomical spectroscopy.
+Automatic differentiation \citep[``autodiff'',][]{2015arXiv150205767G} and the backpropagation algorithm has revolutionized machine learning and neural network architecture design, and is increasingly applied in astrophysical data analysis contexts, \emph{e.g.} kernel phase coronography with \texttt{poppy} \citep{2021ApJ...907...40P}, and exoplanet orbit fitting with \texttt{exoplanet} \citep{2021JOSS....6.3285F}.
+Of the spectroscopy frameworks mentioned above, \texttt{wobble} \citep{2019AJ....158..164B} and \texttt{ExoJAX} \citep{2022ApJS..258...31K} employ autodiff technology.
+\texttt{wobble} treats each pixel as a tunable control point, producing $\sim10^5$ parameters for a modern stellar spectrum.
+The \texttt{ExoJAX} framework has only $\sim$dozens of tunable parameters that describe the fundamental physical properties controlling a brown dwarf atmosphere.
+These two autodiff-aware frameworks span the extreme ends of non-parametric and parametric modeling for spectroscopy.
 
-The \texttt{wobble} framework \citep{2019AJ....158..164B} modernized the construction of high-SNR templates to account for temporally variable telluric lines.
-The tool requires dozens of high-SNR spectra acquired at a range of Barycentric Earth Radial Velocities (BERVs).
-The final telluric-free combined spectrum would still have to be compared to models for absolute calibration, or can be used out-of-the-box for precision relative RVs.
-The \texttt{wobble} framework also pioneered the off-label application of automatic differentiation frameworks---in this case \texttt{TensorFlow}---towards their physically-motivated use in stellar spectra.
-\texttt{blas\'e} can be viewed as an evaluable and interpretable super-resolution version of \texttt{wobble}, that accepts more bias in the bias-variance tradeoff.
+In this paper we show that autodiff-aware semi-empirical models offer a cozy middle ground: informed from self-consistent models, but refined with data.
+In Section \ref{methodology} we show how to supplant the computationally expensive radiative transfer, by cloning pre-computed synthetic spectral models.
+In Section \censor{XX} we describe how to fine-adjust the cloned models to data in a transfer-learning step.
+Section \censor{XX} exhibits some injection-recovery type tests on simulated data of known properties.
+In Section \censor{XX} we show real-world data from high-grasp near-IR spectrographs.
+Finally we discuss many conceivable extensions to the framework, and directions for future work in Section \censor{xx}.
 
-The \texttt{starfish} framework \citep{czekala15} provides a robust likelihood function for data-model comparisons, and retires many of the problems in this domain.
-\texttt{starfish} pioneered the use of whole-spectrum fitting with resilience to model imperfections by addressing the problem of what to do when the underlying atomic and molecular data was wrong or approximate or missing.
-It has been extended to inferring starspot physical properties \citep{2017ApJ...836..200G}, measuring veiling in Class 0 protostars \citep{2018ApJ...862...85G}, and quantifying imperfections in brown dwarf models \citep{2021ApJ...921...95Z}.
-The Spectral Inference Crank \citep[\texttt{sick},][]{2016ApJS..223....8C} shares similar aims as \texttt{starfish}, and provides additional useful grid search capabilities.
-
-For very large bandwidths and very many spectral lines, the problem of identifying and cataloging line imperfections essentially becomes a book-keeping and continuum assignment problem.
-\texttt{blas\'e} and \texttt{starfish} provide different strategies for orchestrating the line-mismatch identification procedure, with each route having tradeoffs depending on the application.
-
-\subsection{This paper: towards semi-empirical models}
-\begin{itemize}
- \item So in practice, tradeoffs between model flexibility and physical self-consistency
- \item Semi-empirical models may be a middle ground: informed from models, but revised with data
- \item The radiative transfer step is computationally expensive, so we want to avoid redoing that
- \item Instead focus on pre-computed models that have that wisdom baked-in
- \item We are essentially compressing pre-computed models into evaluable, interpretable models
- \item Transfer learning is one plausible approach: pre-train on pre-computed models, transfer to data
- \item Neural network frameworks make this possible with autodiff
- \item We use the autodiff and GPU acceleration, but not the neural architectures
- \item We do not use the neural architectures because we have a good understand for how a spectrum is generated
- \item Instead of tuning weights of a neural network, we tune the atomic and molecular properties of a spectrum
-\end{itemize}
-
-\section{Methodology}
+\section{Methodology}\label{methodology}
 
@@
 
+\subsection{Comparison to existing frameworks}
+
+The \texttt{specmatch} synthetic template matching tool produces noise-free nearest neighbor templates given an input spectrum \citep{2015PhDT........82P}.
+Several practical barriers limit the accuracy of using precomputed synthetic spectral models alone.
+First and foremost, real stars are usually more complicated than our simplified models of them.
+Real spectra often vary over more dimensions that our models do.
+Conspicuous examples of these hidden variables can be found in protostars: starspots, accretion veiling, dust extinction, and magnetic Zeeman splitting.
+Jointly modeling all of these phenomena alongside the intrinsic stellar photosphere is challenging.
+
+The empirical version, \texttt{specmatch-emp} \citep{2017ApJ...836...77Y} matched spectra better than the synthetic templates, but is still too rigid for some applications and requires the assembly of hundreds of standardized high signal-to-noise-ratio templates, ideally with low intrinsic rotational broadening.
+Such a large number of high-quality templates has not yet been established in the near infrared, and may not be feasible for brown dwarfs.
+
+The \texttt{wobble} framework \citep{2019AJ....158..164B} modernized the construction of high-SNR templates to account for temporally variable telluric lines.
+The tool requires dozens of high-SNR spectra acquired at a range of Barycentric Earth Radial Velocities (BERVs).
+The final telluric-free combined spectrum would still have to be compared to models for absolute calibration, or can be used out-of-the-box for precision relative RVs.
+The \texttt{wobble} framework also pioneered the off-label application of automatic differentiation frameworks---in this case \texttt{TensorFlow}---towards their physically-motivated use in stellar spectra.
+\texttt{blas\'e} can be viewed as an evaluable and interpretable super-resolution version of \texttt{wobble}, that accepts more bias in the bias-variance tradeoff.
+
+The \texttt{starfish} framework \citep{czekala15} provides a robust likelihood function for data-model comparisons, and retires many of the problems in this domain.
+\texttt{starfish} pioneered the use of whole-spectrum fitting with resilience to model imperfections by addressing the problem of what to do when the underlying atomic and molecular data was wrong or approximate or missing.
+It has been extended to inferring starspot physical properties \citep{2017ApJ...836..200G}, measuring veiling in Class 0 protostars \citep{2018ApJ...862...85G}, and quantifying imperfections in brown dwarf models \citep{2021ApJ...921...95Z}.
+The Spectral Inference Crank \citep[\texttt{sick},][]{2016ApJS..223....8C} shares similar aims as \texttt{starfish}, and provides additional useful grid search capabilities.
+
+For very large bandwidths and very many spectral lines, the problem of identifying and cataloging line imperfections essentially becomes a book-keeping and continuum assignment problem.
+\texttt{blas\'e} and \texttt{starfish} provide different strategies for orchestrating the line-mismatch identification procedure, with each route having tradeoffs depending on the application.
+
 \subsection{Lineshapes choice and tradeoffs}

### 2022-03-02 paper/paper1/ms.tex
Get rid of URLs for now because they raise warnings

@@
 \documentclass[modern]{aastex631}
+\usepackage{hyperref}
 \bibliographystyle{aasjournal}
@@
 \usepackage{amsmath,bm}
-\usepackage{hyperref}
-\Urlmuskip=0mu  plus 10mu
 
@@
 Imperfections in these templates can mute the perceived signal strength of these atmospheric features \citep{2015A&A...575A..20H}.
-In extreme precision radial velocity (EPRV) applications, cross-correlation methods work, but addressing telluric absorption at the $\mathrm{cm/s}$ level may require joint modeling of the star and the Earth's atmospheric absorption before convolution with an instrumental kernel\footnote{\url{https://speakerdeck.com/dwhgg/telluric-line-hack-week-wrap-up}}.
+In extreme precision radial velocity (EPRV) applications, cross-correlation methods work, but addressing telluric absorption at the $\mathrm{cm/s}$ level may require joint modeling of the star and the Earth's atmospheric absorption before convolution with an instrumental kernel \authorcomment1{speakerdeck.com/dwhgg/telluric-line-hack-week-wrap-up}).
+
 In the case of Doppler imaging, an accurate underlying template is needed to detect longitudinally symmetric structures such as polar spots \citep{1983PASP...95..565V, roettenbacher16} or zonal bands \citep{Crossfield14,2021ApJ...906...64A}.
@@
 Existing open-source frameworks have overcome some of these challenges, or have been purpose-built for specialized applications. 
-These frameworks include \texttt{ROBOSPECT} \citep{2013PASP..125.1164W}, \texttt{specmatch} \citep{2015PhDT........82P}, \texttt{specmatch-emp} \citep{2017ApJ...836...77Y}, \texttt{wobble} \citep{2019AJ....158..164B}, \texttt{starfish} \citep{czekala15}, \texttt{sick} \citep{2016ApJS..223....8C}, \texttt{psoap} \citep{2017ApJ...840...49C}, \texttt{FAL}\footnote{\url{https://github.com/pacargile/FAL}}, CHIMERA \citep{2015ApJ...807..183L}, the \texttt{Cannon} \citep{2017ApJ...836....5H},  \texttt{MINESweeper} \citep{2020ApJ...900...28C}, and recently \texttt{ExoJAX} \citep{2022ApJS..258...31K}.
+These frameworks include \texttt{ROBOSPECT} \citep{2013PASP..125.1164W}, \texttt{specmatch} \citep{2015PhDT........82P}, \texttt{specmatch-emp} \citep{2017ApJ...836...77Y}, \texttt{wobble} \citep{2019AJ....158..164B}, \texttt{starfish} \citep{czekala15}, \texttt{sick} \citep{2016ApJS..223....8C}, \texttt{psoap} \citep{2017ApJ...840...49C}, \texttt{FAL} \authorcomment1{Cargile et al.
+in prep}, CHIMERA \citep{2015ApJ...807..183L}, the \texttt{Cannon} \citep{2017ApJ...836....5H},  \texttt{MINESweeper} \citep{2020ApJ...900...28C}, and recently \texttt{ExoJAX} \citep{2022ApJS..258...31K}.
 The design of these frameworks necessarily have to make a choice in the bias-variance tradeoff: is the tool more \emph{data}-driven or more \emph{model}-driven?

### 2022-03-02 paper/paper1/ms.tex
Minor tweaks to the introduction

@@
 The metaphorical holy grail of astronomical spectroscopy is a function that takes in an observed astronomical spectrum and reports back the postition, amplitude, width, and shape of all of its spectral lines, automatically, accurately, and precisely.
-The function would go further, to report back the systemic radial velocity (RV) and rotational broadening ($v\sin{i}$).
+The function would go further.
+It would report back the systemic radial velocity (RV) and rotational broadening ($v\sin{i}$).
 Finally the function would---in an \emph{Inception}-like manner---provide what it believes to be the interpretable generating function that produced the data, so that we may gain insights on future examples of this or other stars.
@@
 In particular, data from high-\emph{grasp} \'echelle spectrographs possess simultaneously high spectral resolving power and high bandwidth, yielding tens of thousands or possibly millions of independent spectral resolution elements for each star, substar, or exoplanet.
-Those applications that seek to gain signal by ``stacking'' spectral lines or cross-correlating templates can hypothetically gain huge boosts in the accessible signal-to-noise ratio compared to a single or few lines.
+Those applications that seek to gain signal by ``stacking'' spectral lines or cross-correlating with templates can hypothetically gain huge boosts in the accessible signal-to-noise ratio compared to a single or few lines.
 Most manual and semi-automated methods cannot take advantage of the entire spectral bandwidth, or rely on exact knowledge of the underlying templates and may fail to achieve the hypothetical promise of these high-bandwidth spectrographs \citep{2020AJ....160..198H}.
@@
 Imperfections in these templates can mute the perceived signal strength of these atmospheric features \citep{2015A&A...575A..20H}.
-In extreme precision radial velocity (EPRV) applications, cross-correlation methods work, but addressing telluric absorption at the $\mathrm{cm/s}$ level may require joint modeling of the star and the Earth's atmospheric absorption before convolution with an instrumental kernel \authorcomment1{speakerdeck.com/dwhgg/telluric-line-hack-week-wrap-up}).
 
-In the case of Doppler imaging, an accurate underlying template is needed to detect longitudinally symmetric structures such as polar spots \citep{1983PASP...95..565V, roettenbacher16} or zonal bands \citep{Crossfield14,2021ApJ...906...64A}.
+In extreme precision radial velocity (EPRV) applications, cross-correlation methods work \citep{2018A_A...620A..47D}, but have many limitations \citep{2022arXiv220110639Z}.
+Addressing telluric absorption at the $\mathrm{cm/s}$ level may require joint modeling of the star and the Earth's atmospheric absorption before convolution with an instrumental kernel\authorcomment1{speakerdeck.com/dwhgg/telluric-line-hack-week-wrap-up}.
+This telluric joint modeling capability does not yet exist at a precision that can meet these strenuous demands.
+
+In the case of Doppler imaging, an accurate underlying template is needed to detect longitudinally symmetric structures \citep{1983PASP...95..565V,2021arXiv211006271L} such as polar spots \citep{roettenbacher16} or zonal bands \citep{Crossfield14,2021ApJ...906...64A}.
 
@@
 The design of these frameworks necessarily have to make a choice in the bias-variance tradeoff: is the tool more \emph{data}-driven or more \emph{model}-driven?
-The statistical tradeoff can be viewed as a concession in physical self-consistency for model flexibility: more or fewer parameters.
+The statistical tradeoff can be viewed as a concession in physical self-consistency for model flexibility: more or fewer parameters; more \emph{accurate} or more \emph{precise}.
 
-A key new enabling technology breaks the historic limits of data-model comparisons for astronomical spectroscopy.
-Automatic differentiation \citep[``autodiff'',][]{2015arXiv150205767G} and the backpropagation algorithm has revolutionized machine learning and neural network architecture design, and is increasingly applied in astrophysical data analysis contexts, \emph{e.g.} kernel phase coronography with \texttt{poppy} \citep{2021ApJ...907...40P}, and exoplanet orbit fitting with \texttt{exoplanet} \citep{2021JOSS....6.3285F}.
-Of the spectroscopy frameworks mentioned above, \texttt{wobble} \citep{2019AJ....158..164B} and \texttt{ExoJAX} \citep{2022ApJS..258...31K} employ autodiff technology.
+A key new enabling technology breaks these classical tradeoffs in data-model comparisons for astronomical spectroscopy.
+Automatic differentiation \citep[``autodiff'',][]{2015arXiv150205767G} and its affiliated backpropagation algorithm has revolutionized machine learning and neural network architecture design, and is increasingly applied in astrophysical data analysis contexts, \emph{e.g.} kernel phase coronography with \texttt{poppy} \citep{2021ApJ...907...40P}, and exoplanet orbit fitting with \texttt{exoplanet} \citep{2021JOSS....6.3285F}.
+Of the spectroscopy frameworks mentioned above, the \texttt{TensorFlow}-based \citep{tensorflow2015-whitepaper} \texttt{wobble} and the \texttt{JAX}-based \citep{jax2018github} \texttt{ExoJAX} employ autodiff technology.
 \texttt{wobble} treats each pixel as a tunable control point, producing $\sim10^5$ parameters for a modern stellar spectrum.
@@
 
-In this paper we show that autodiff-aware semi-empirical models offer a cozy middle ground: informed from self-consistent models, but refined with data.
-In Section \ref{methodology} we show how to supplant the computationally expensive radiative transfer, by cloning pre-computed synthetic spectral models.
-In Section \censor{XX} we describe how to fine-adjust the cloned models to data in a transfer-learning step.
-Section \censor{XX} exhibits some injection-recovery type tests on simulated data of known properties.
-In Section \censor{XX} we show real-world data from high-grasp near-IR spectrographs.
-Finally we discuss many conceivable extensions to the framework, and directions for future work in Section \censor{xx}.
+In this paper we show that autodiff-aware semi-empirical models offer an appealing middle ground: informed from self-consistent models, but refined with data.
+In Section \ref{methodology} we introduce the \emph{blas\'e} forward-model design and its \texttt{PyTorch}-based \cite{2019arXiv191201703P} implementation.
+We show how to supplant the computationally expensive radiative transfer by cloning pre-computed synthetic spectral models.
+In Section \censorbox{XX} we describe how to fine-adjust the cloned models to data in a transfer-learning step.
+Section \censorbox{XX} exhibits some injection-recovery type tests on simulated data of known properties.
+In Section \censorbox{XX} we show real-world data from high-grasp near-IR spectrographs.
+Finally we discuss many conceivable extensions to the framework, and directions for future work in Section \censorbox{XX}.
 

### 2022-03-02 paper/paper1/ms.tex
Add Adam reference, 22000 citations holy moly

@@
 
-We use the Adam optimizer\authorcomment1{citation} with a typical learning rate $LR\in (0.005, 0.05)$ and all the defaults for \texttt{PyTorch v1.11}.
+We use the Adam optimizer \citep{2014arXiv1412.6980K} with a typical learning rate $LR\in (0.005, 0.05)$ and all the defaults for \texttt{PyTorch v1.11}.
 Minitbatch testing proceeded with batch fractions of typically $f\sim1-5\%$ of the available pixels but was rapidly discontinued in favor of the much more performant and convenient sparse matrix method.

### 2022-03-02 paper/paper1/ms.tex
Add figure captions for Figures 2 and 3, to be revised

@@
  \centering
- \includegraphics[width=0.6\textwidth]{example-image-golden}
- \caption{PHOENIX spectrum cloned with blas\'e.}
+ \includegraphics[width=0.8\textwidth]{figures/blase_clone_demo.png}
+ \caption{PHOENIX spectrum cloned with blas\'e.
+This $T_{\mathrm{eff}}=4100\;$K,  $\log{g}=3.5$ solar metallicity model has over 5,000 individual cloned spectral lines.
+The figure zooms in on a $0.7\%$ portion of the large spectral bandwidth that was cloned.
+Some conspicuous flaws can be seen near the wings of broad lines, or surrounding blended lines.
+Some initialization effects can be seen in this example, suggesting that the cloning procedure requires additional training iterations to achieve the best performance.}
  \label{fig_cloned_spectrum_demo}
@@
  \centering
- \includegraphics[width=0.3\textwidth]{example-grid-100x100pt}
- \caption{A \censor{Hydrogen} line with broad line wings exhibiting poor cloning performance.}
+ \includegraphics[width=0.8\textwidth]{figures/blase_clone_demo_zoom.png}
+ \caption{A zoom in of the region from Figure \ref{fig_cloned_spectrum_demo}.}
  \label{fig_poor_cloning_performance}

### 2022-03-03 paper/paper1/ms.tex
Re-ran the cloning with way more training steps, fixed RV offset.  Looks great!

@@
     \caption{PHOENIX spectrum cloned with blas\'e.
-This $T_{\mathrm{eff}}=4100\;$K,  $\log{g}=3.5$ solar metallicity model has over 5,000 individual cloned spectral lines.
-The figure zooms in on a $0.7\%$ portion of the large spectral bandwidth that was cloned.
-Some conspicuous flaws can be seen near the wings of broad lines, or surrounding blended lines.
-Some initialization effects can be seen in this example, suggesting that the cloning procedure requires additional training iterations to achieve the best performance.}
+This $T_{\mathrm{eff}}=4100\;$K,  $\log{g}=3.5$ solar metallicity model has 11,759 individual cloned spectral lines, each with 3 tuned parameters.
+The pictured 25 \AA chunk contains 73 spectral lines and represents a mere $<0.7\%$ portion of the entire spectral bandwidth that was cloned.
+Some flaws can be seen near the cores of deep lines, or wings of broad lines.}
     \label{fig_cloned_spectrum_demo}

### 2022-03-03 src/blase/emulator.py

@@ -282,7 +282,7 @@ class SparsePhoenixEmulator(PhoenixEmulator):
         self.active_mask = self.active_mask.to(device)
 
         self.radial_velocity = nn.Parameter(
-            torch.tensor(10.0, requires_grad=True, dtype=torch.float64)
+            torch.tensor(0.0, requires_grad=True, dtype=torch.float64)
         )
 
     def forward(self):

### 2022-03-03 paper/paper1/ms.tex
Looks good!

@@
 Automatic differentiation \citep[``autodiff'',][]{2015arXiv150205767G} and its affiliated backpropagation algorithm has revolutionized machine learning and neural network architecture design, and is increasingly applied in astrophysical data analysis contexts, \emph{e.g.} kernel phase coronography with \texttt{poppy} \citep{2021ApJ...907...40P}, and exoplanet orbit fitting with \texttt{exoplanet} \citep{2021JOSS....6.3285F}.
-Of the spectroscopy frameworks mentioned above, the \texttt{TensorFlow}-based \citep{tensorflow2015-whitepaper} \texttt{wobble} and the \texttt{JAX}-based \citep{jax2018github} \texttt{ExoJAX} employ autodiff technology.
+Of the spectroscopy frameworks mentioned above, the \texttt{TensorFlow}-based \citep{tensorflow2015-whitepaper} \texttt{wobble} and the \texttt{JAX}-based \censorbox{jax2018github} \texttt{ExoJAX} employ autodiff technology.
 \texttt{wobble} treats each pixel as a tunable control point, producing $\sim10^5$ parameters for a modern stellar spectrum.
@@
 This $T_{\mathrm{eff}}=4100\;$K,  $\log{g}=3.5$ solar metallicity model has 11,759 individual cloned spectral lines, each with 3 tuned parameters.
-The pictured 25 \AA chunk contains 73 spectral lines and represents a mere $<0.7\%$ portion of the entire spectral bandwidth that was cloned.
+The pictured 25 \AA~ chunk contains 73 spectral lines and represents a mere $<0.7\%$ portion of the entire spectral bandwidth that was cloned.
 Some flaws can be seen near the cores of deep lines, or wings of broad lines.}
@@
     \includegraphics[width=0.8\textwidth]{figures/blase_clone_demo_zoom.png}
-    \caption{A zoom in of the region from Figure \ref{fig_cloned_spectrum_demo}.}
+    \caption{A zoom in of the region between the vertical gray bars in Figure \ref{fig_cloned_spectrum_demo}. There are 9 spectral lines in this view.}
     \label{fig_poor_cloning_performance}

### 2022-03-03 experiments/05_regularization_priors/05_A_pretrain_emulator.py
Uncommitted notebooks for the regularization approach

@@ -1,0 +1,67 @@
+import os
+import torch
+from torch import nn
+from tqdm import trange
+import torch.optim as optim
+from blase.emulator import SparsePhoenixEmulator
+import matplotlib.pyplot as plt
+from gollum.phoenix import PHOENIXSpectrum
+import numpy as np
+from muler.hpf import HPFSpectrum
+import copy
+
+if torch.cuda.is_available():
+    device = "cuda"
+else:
+    device = "cpu"
+
+# Pre-process the model as described in the paper
+spectrum = PHOENIXSpectrum(teff=4700, logg=4.5)
+spectrum = spectrum.divide_by_blackbody()
+spectrum = spectrum.normalize()
+continuum_fit = spectrum.fit_continuum(polyorder=5)
+spectrum = spectrum.divide(continuum_fit, handle_meta="ff")
+
+# Numpy arrays: 1 x N_pix
+wl_native = spectrum.wavelength.value
+flux_native = spectrum.flux.value
+
+# Create the emulator and load a pretrained model
+prominence = 0.01
+emulator = SparsePhoenixEmulator(
+    wl_native, flux_native, prominence=prominence, wing_cut_pixels=1000
+)
+emulator.to(device)
+
+wl_native = emulator.wl_native.clone().detach().to(device)
+wl_active = wl_native.to("cpu")[emulator.active_mask.to("cpu").numpy()]
+target = (
+    emulator.flux_native.clone().detach().to(device)[emulator.active_mask.cpu().numpy()]
+)
+
+emulator.radial_velocity.requires_grad = False
+emulator.lam_centers.requires_grad = True
+emulator.amplitudes.requires_grad = True
+emulator.sigma_widths.requires_grad = True
+emulator.gamma_widths.requires_grad = True
+
+loss_fn = nn.MSELoss(reduction="mean")
+
+optimizer = optim.Adam(
+    list(filter(lambda p: p.requires_grad, emulator.parameters())), 0.01, amsgrad=True,
+)
+n_epochs = 1000
+losses = []
+
+# high_res_model = emulator.flux_native.clone().detach().to(device)
+t_iter = trange(n_epochs, desc="Training", leave=True)
+for epoch in t_iter:
+    emulator.train()
+    yhat = emulator.forward()[emulator.active_mask]
+    loss = loss_fn(yhat, target)
+    loss.backward()
+    optimizer.step()
+    optimizer.zero_grad()
[…]

### 2022-03-03 experiments/05_regularization_priors/05_B_inject_fake_signals.py

@@ -1,0 +1,103 @@
+import os
+import torch
+from torch import nn
+from tqdm import trange
+import torch.optim as optim
+from blase.emulator import SparsePhoenixEmulator
+import matplotlib.pyplot as plt
+from gollum.phoenix import PHOENIXSpectrum
+import numpy as np
+from muler.hpf import HPFSpectrumList
+import copy
+
+if torch.cuda.is_available():
+    device = "cuda"
+else:
+    device = "cpu"
+
+# Pre-process the model as described in the paper
+spectrum = PHOENIXSpectrum(teff=4700, logg=4.5)
+spectrum = spectrum.divide_by_blackbody()
+spectrum = spectrum.normalize()
+continuum_fit = spectrum.fit_continuum(polyorder=5)
+spectrum = spectrum.divide(continuum_fit, handle_meta="ff")
+
+# Fetch the real HPF data
+# For now we will use some data from another project
+fn = "Goldilocks_20191022T014138_v1.0_0004.spectra.fits"
+dir = "/home/gully/GitHub/blase/test/data/"
+dir = os.path.expanduser(dir)
+raw_data = HPFSpectrumList.read(file=dir + fn)  # Just one order for now
+raw_data = raw_data.sky_subtract().deblaze().normalize().trim_edges((8, 2040))
+raw_data = raw_data.stitch().mask_tellurics().normalize()
+
+mask = (raw_data.wavelength.value > 8450) & (raw_data.wavelength.value < 8900)
+data = raw_data.apply_boolean_mask(mask).normalize()
+
+
+# Numpy arrays: 1 x N_pix
+wl_native = spectrum.wavelength.value
+flux_native = spectrum.flux.value
+
+# Create the emulator and load a pretrained model
+prominence = 0.01
+emulator = SparsePhoenixEmulator(
+    wl_native, flux_native, prominence=prominence, wing_cut_pixels=1000
+)
+emulator.to(device)
+
+wl_native = emulator.wl_native.clone().detach().to(device)
+wl_active = wl_native.to("cpu")[emulator.active_mask.to("cpu").numpy()]
+target = (
+    emulator.flux_native.clone().detach().to(device)[emulator.active_mask.cpu().numpy()]
+)
+
+state_dict_post = torch.load("emulator_T4700g4p5_prom0p01_HPF.pt")
+emulator.load_state_dict(state_dict_post)
+
+
+# Make synthetic data
+# ---------------------------------------------------------
[…]

### 2022-03-03 experiments/05_regularization_priors/05_C_recover_with_priors.py

@@ -1,0 +1,119 @@
+import os
+import torch
+from torch import nn
+from tqdm import trange
+import torch.optim as optim
+from blase.emulator import SparsePhoenixEmulator
+import matplotlib.pyplot as plt
+from gollum.phoenix import PHOENIXSpectrum
+import numpy as np
+from muler.hpf import HPFSpectrumList
+import copy
+
+if torch.cuda.is_available():
+    device = "cuda"
+else:
+    device = "cpu"
+
+# Pre-process the model as described in the paper
+spectrum = PHOENIXSpectrum(teff=4700, logg=4.5)
+spectrum = spectrum.divide_by_blackbody()
+spectrum = spectrum.normalize()
+continuum_fit = spectrum.fit_continuum(polyorder=5)
+spectrum = spectrum.divide(continuum_fit, handle_meta="ff")
+
+# Fetch the real HPF data
+# For now we will use some data from another project
+fn = "Goldilocks_20191022T014138_v1.0_0004.spectra.fits"
+dir = "/home/gully/GitHub/blase/test/data/"
+dir = os.path.expanduser(dir)
+raw_data = HPFSpectrumList.read(file=dir + fn)  # Just one order for now
+raw_data = raw_data.sky_subtract().deblaze().normalize().trim_edges((8, 2040))
+raw_data = raw_data.stitch().mask_tellurics().normalize()
+
+mask = (raw_data.wavelength.value > 8450) & (raw_data.wavelength.value < 8900)
+data = raw_data.apply_boolean_mask(mask).normalize()
+
+
+# Numpy arrays: 1 x N_pix
+wl_native = spectrum.wavelength.value
+flux_native = spectrum.flux.value
+
+# Create the emulator and load a pretrained model
+prominence = 0.01
+emulator = SparsePhoenixEmulator(
+    wl_native, flux_native, prominence=prominence, wing_cut_pixels=1000
+)
+emulator.to(device)
+
+wl_native = emulator.wl_native.clone().detach().to(device)
+wl_active = wl_native.to("cpu")[emulator.active_mask.to("cpu").numpy()]
+target = (
+    emulator.flux_native.clone().detach().to(device)[emulator.active_mask.cpu().numpy()]
+)
+
+state_dict_post = torch.load("emulator_T4700g4p5_prom0p01_HPF.pt")
+emulator.load_state_dict(state_dict_post)
+
+emulator.radial_velocity = nn.Parameter(torch.tensor(0.0, device=device))
[…]

### 2022-03-03 experiments/05_regularization_priors/05_D_plot_results.py

@@ -1,0 +1,81 @@
+import os
+import torch
+from blase.emulator import SparsePhoenixEmulator
+import matplotlib.pyplot as plt
+from gollum.phoenix import PHOENIXSpectrum
+import numpy as np
+from torch import nn
+from muler.hpf import HPFSpectrum, HPFSpectrumList
+import copy
+
+if torch.cuda.is_available():
+    device = "cuda"
+else:
+    device = "cpu"
+
+# Pre-process the model as described in the paper
+spectrum = PHOENIXSpectrum(teff=4700, logg=4.5)
+spectrum = spectrum.divide_by_blackbody()
+spectrum = spectrum.normalize()
+continuum_fit = spectrum.fit_continuum(polyorder=5)
+spectrum = spectrum.divide(continuum_fit, handle_meta="ff")
+
+# Fetch the real HPF data
+# For now we will use some data from another project
+fn = "Goldilocks_20191022T014138_v1.0_0004.spectra.fits"
+dir = "/home/gully/GitHub/blase/test/data/"
+dir = os.path.expanduser(dir)
+raw_data = HPFSpectrumList.read(file=dir + fn)  # Just one order for now
+raw_data = raw_data.sky_subtract().deblaze().normalize().trim_edges((8, 2040))
+raw_data = raw_data.stitch().mask_tellurics().normalize()
+
+mask = (raw_data.wavelength.value > 8450) & (raw_data.wavelength.value < 8900)
+data = raw_data.apply_boolean_mask(mask).normalize()
+
+
+# Numpy arrays: 1 x N_pix
+wl_native = spectrum.wavelength.value
+flux_native = spectrum.flux.value
+
+# Create the emulator and load a pretrained model
+prominence = 0.01
+emulator = SparsePhoenixEmulator(
+    wl_native, flux_native, prominence=prominence, wing_cut_pixels=1000
+)
+emulator.to(device)
+
+wl_native = emulator.wl_native.clone().detach().to(device)
+wl_active = wl_native.to("cpu")[emulator.active_mask.to("cpu").numpy()]
+target = (
+    emulator.flux_native.clone().detach().to(device)[emulator.active_mask.cpu().numpy()]
+)
+
+state_dict_post = torch.load("./emulator_T4700g4p5_prom0p01_HPF_recovery.pt")
+emulator.load_state_dict(state_dict_post)
+
+from blase.emulator import EchelleModel
+
[…]

### 2022-03-03 experiments/06_bare_emulator_clone/06_bare_emulator_clone.py
experiment 6 is just simple cloning

@@ -1,0 +1,126 @@
+import os
+import torch
+from torch import nn
+from tqdm import trange
+import torch.optim as optim
+import torch.profiler
+from blase.emulator import SparsePhoenixEmulator
+import matplotlib.pyplot as plt
+from gollum.phoenix import PHOENIXSpectrum
+from torch.utils.tensorboard import SummaryWriter
+import webbrowser
+import numpy as np
+from muler.hpf import HPFSpectrum
+
+if torch.cuda.is_available():
+    device = "cuda"
+else:
+    device = "cpu"
+
+
+def plot_spectrum(spectra):
+    """
+    Generates matplotlib Figure using a trained network, along with images
+    and labels from a batch
+    """
+    fig, axes = plt.subplots(2, figsize=(8, 3))
+    axes[0].step(
+        spectra[0]["wl"].numpy(),
+        spectra[0]["flux"].numpy(),
+        label="Cloned Spectrum",
+        lw=1,
+    )
+    axes[0].step(
+        spectra[1]["wl"].numpy(),
+        spectra[1]["flux"].numpy(),
+        label="Native Spectrum",
+        lw=1,
+    )
+    axes[0].set_ylim(0, 1.5)
+    axes[0].legend(loc="upper right", ncol=2)
+    axes[1].step(
+        spectra[1]["wl"].numpy(),
+        spectra[0]["flux"].numpy() - spectra[1]["flux"].numpy(),
+    )
+    axes[1].set_ylim(-0.3, 0.3)
+    return fig
+
+
+log_dir = "runs/synthetic1"
+writer = SummaryWriter(log_dir=log_dir)
+webbrowser.open("http://localhost:6006/", new=2)
+
+
+# Pre-process the model as described in the paper
+spectrum = PHOENIXSpectrum(teff=4100, logg=3.5)
+spectrum = spectrum.divide_by_blackbody()
+spectrum = spectrum.normalize()
+continuum_fit = spectrum.fit_continuum(polyorder=5)
+spectrum = spectrum.divide(continuum_fit, handle_meta="ff")
+
+# Numpy arrays: 1 x N_pix
+wl_native = spectrum.wavelength.value
+flux_native = spectrum.flux.value
+
+# Create the emulator and load a pretrained model
+prominence = 0.01
+emulator = SparsePhoenixEmulator(
+    wl_native, flux_native, prominence=prominence, wing_cut_pixels=1000
+)
+emulator.to(device)
+
+wl_native = emulator.wl_native.clone().detach().to(device)
[…]

### 2022-03-03 paper/paper1/ms.tex
Figures are looking good now
Ok, finally settled on using WASP 69 as the demo source.

@@
     \caption{PHOENIX spectrum cloned with blas\'e.
-This $T_{\mathrm{eff}}=4100\;$K,  $\log{g}=3.5$ solar metallicity model has 11,759 individual cloned spectral lines, each with 3 tuned parameters.
-The pictured 25 \AA~ chunk contains 73 spectral lines and represents a mere $<0.7\%$ portion of the entire spectral bandwidth that was cloned.
+This $T_{\mathrm{eff}}=4700\;$K,  $\log{g}=4.5$ solar metallicity model has 9,028 individual cloned spectral lines, each with 3 tuned parameters.
+The pictured 50 \AA~ chunk contains 121 spectral lines and represents about 1$\%$ of the entire spectral bandwidth that was cloned.
 Some flaws can be seen near the cores of deep lines, or wings of broad lines.}
@@
 The greatest residuals surround the deepest and broadest lines---such as Hydrogen and neutral alkali metal lines---for which the non-Voigt lineshapes stand out.
-Figure \ref{fig_poor_cloning_performance}\authorcomment1{make this figure} zooms in on the \censor{XX} line at $\censor{10200}\;$\AA.
+Figure \ref{fig_zoom_cloning_performance}\authorcomment1{make this figure} zooms in on the \censor{XX} line at $\censor{10200}\;$\AA.
 
@@
     \includegraphics[width=0.8\textwidth]{figures/blase_clone_demo_zoom.png}
-    \caption{A zoom in of the region between the vertical gray bars in Figure \ref{fig_cloned_spectrum_demo}. There are 9 spectral lines in this view.}
-    \label{fig_poor_cloning_performance}
+    \caption{Zoom in of the region between the vertical gray bars in Figure \ref{fig_cloned_spectrum_demo}.
+The cloned model has 7 spectral lines describing the 400 pixels in this 4 \AA~ chunk.
+The native PHOENIX pixel sampling can be see as the boxy steps in both the native and cloned model.
+}
+    \label{fig_zoom_cloning_performance}
 \end{figure}
@@
 We see some narrow lines devolve into missing continuum opacity, the favored tradeoff when the continuum estimate's poor performance outweighs the pain of a narrow-but-tolerably-small spike.
-This problem can be seen in Figure \ref{fig_poor_cloning_performance} where a line initialized at $\censor{10210}\;$\AA~ with an amplitude of ZZ and width of YY bloats to \censor{some other value}.
+This problem can be seen in Figure \ref{fig_zoom_cloning_performance} where a line initialized at $\censor{10210}\;$\AA~ with an amplitude of ZZ and width of YY bloats to \censor{some other value}.
 These devolved lines tend to appear at poorly imitated broad-and-deep line wings.

### 2022-03-03 experiments/06_bare_emulator_clone/06_bare_emulator_clone.py

@@ -52,7 +52,7 @@ webbrowser.open("http://localhost:6006/", new=2)
 
 
 # Pre-process the model as described in the paper
-spectrum = PHOENIXSpectrum(teff=4100, logg=3.5)
+spectrum = PHOENIXSpectrum(teff=4700, logg=4.5)
 spectrum = spectrum.divide_by_blackbody()
 spectrum = spectrum.normalize()
 continuum_fit = spectrum.fit_continuum(polyorder=5)
@@ -122,5 +122,5 @@ for epoch in t_iter:
             "predictions vs. actuals", plot_spectrum(to_plot), global_step=epoch,
         )
 
-        torch.save(emulator.state_dict(), "emulator_T4100g3p5_prom0p01_HPF_recovery.pt")
+        torch.save(emulator.state_dict(), "emulator_T4700g4p5_prom0p01_HPF_recovery.pt")
 

### 2022-03-03 paper/paper1/ms.tex
Woohoo, we made 4 figures today

@@
 \section{Refining the model from data with transfer learning}
+
+\begin{figure}[hbt!]
+    \centering
+    \includegraphics[width=0.8\textwidth]{figures/blase_WASP69_demo.png}
+    \caption{WASP 69 observed with HPF, compared to a $T_{\mathrm{eff}}=4700\;$K,  $\log{g}=4.5$ solar metallicity PHOENIX model warped to $v\sin{i}=2.2$~km/s, $RV=-9.6$~km/s, and HPF resolving power.}
+    \label{fig_WASP69_demo}
+\end{figure}
+
+\begin{figure}[hbt!]
+    \centering
+    \includegraphics[width=0.8\textwidth]{figures/blase_WASP69_transfer_learned.png}
+    \caption{Semi-empirical model of WASP 69 transfer-learned with \texttt{blas\'e}. }
+    \label{fig_WASP69_transferred}
+\end{figure}
+
 \begin{outline}

### 2022-03-03 experiments/04_full_bandwidth_HPF/04_B_train_on_data.py

@@ -7,13 +7,28 @@ from blase.emulator import SparsePhoenixEmulator
 import matplotlib.pyplot as plt
 from gollum.phoenix import PHOENIXSpectrum
 import numpy as np
-from muler.hpf import HPFSpectrumList
+from muler.hpf import HPFSpectrumList, HPFSpectrum
 import copy
+import astropy.units as u
 
 if torch.cuda.is_available():
     device = "cuda"
 else:
     device = "cpu"
+
+from astropy.io import fits
+
+# Fetch the real HPF data
+# We will demo on WASP 69
+hdus = fits.open("../../data/WASP_69_hpf_stack.fits")
+
+# Numpy arrays: 1 x N_pix
+data = HPFSpectrum(
+    flux=hdus[1].data["flux"] * u.dimensionless_unscaled,
+    spectral_axis=hdus[1].data["wavelength"] * u.Angstrom,
+)
+data = data.remove_nans()
+
 
 # Pre-process the model as described in the paper
 spectrum = PHOENIXSpectrum(teff=4700, logg=4.5)
@@ -22,22 +37,9 @@ spectrum = spectrum.normalize()
 continuum_fit = spectrum.fit_continuum(polyorder=5)
 spectrum = spectrum.divide(continuum_fit, handle_meta="ff")
 
-# Fetch the real HPF data
-# For now we will use some data from another project
-fn = "Goldilocks_20191022T014138_v1.0_0004.spectra.fits"
-dir = "/home/gully/GitHub/blase/test/data/"
-dir = os.path.expanduser(dir)
-raw_data = HPFSpectrumList.read(file=dir + fn)  # Just one order for now
-raw_data = raw_data.sky_subtract().deblaze().normalize().trim_edges((8, 2040))
-raw_data = raw_data.stitch().mask_tellurics().normalize()
-
-mask = (raw_data.wavelength.value > 8450) & (raw_data.wavelength.value < 8900)
-data = raw_data.apply_boolean_mask(mask).normalize()
-
-
-# Numpy arrays: 1 x N_pix
 wl_native = spectrum.wavelength.value
 flux_native = spectrum.flux.value
+
 
 # Create the emulator and load a pretrained model
 prominence = 0.01
@@ -55,7 +57,7 @@ target = (
 state_dict_post = torch.load("emulator_T4700g4p5_prom0p01_HPF.pt")
 emulator.load_state_dict(state_dict_post)
 
[…]

### 2022-03-04 experiments/07_regularize_real_data/07_B_regularize_real_data.py
Regularize the emulator amplitudes to avoid overfitting

@@ -1,0 +1,121 @@
+import os
+import torch
+from torch import nn
+from tqdm import trange
+import torch.optim as optim
+from blase.emulator import SparsePhoenixEmulator
+import matplotlib.pyplot as plt
+from gollum.phoenix import PHOENIXSpectrum
+import numpy as np
+from muler.hpf import HPFSpectrumList, HPFSpectrum
+import copy
+import astropy.units as u
+
+if torch.cuda.is_available():
+    device = "cuda"
+else:
+    device = "cpu"
+
+from astropy.io import fits
+
+# Fetch the real HPF data
+# We will demo on WASP 69
+hdus = fits.open("../../data/WASP_69_hpf_stack.fits")
+
+# Numpy arrays: 1 x N_pix
+data = HPFSpectrum(
+    flux=hdus[1].data["flux"] * u.dimensionless_unscaled,
+    spectral_axis=hdus[1].data["wavelength"] * u.Angstrom,
+)
+data = data.remove_nans()
+
+
+# Pre-process the model as described in the paper
+spectrum = PHOENIXSpectrum(teff=4700, logg=4.5)
+spectrum = spectrum.divide_by_blackbody()
+spectrum = spectrum.normalize()
+continuum_fit = spectrum.fit_continuum(polyorder=5)
+spectrum = spectrum.divide(continuum_fit, handle_meta="ff")
+
+wl_native = spectrum.wavelength.value
+flux_native = spectrum.flux.value
+
+
+# Create the emulator and load a pretrained model
+prominence = 0.01
+emulator = SparsePhoenixEmulator(
+    wl_native, flux_native, prominence=prominence, wing_cut_pixels=1000
+)
+emulator.to(device)
+
+wl_native = emulator.wl_native.clone().detach().to(device)
+wl_active = wl_native.to("cpu")[emulator.active_mask.to("cpu").numpy()]
+target = (
+    emulator.flux_native.clone().detach().to(device)[emulator.active_mask.cpu().numpy()]
+)
+
+state_dict_post = torch.load("emulator_T4700g4p5_prom0p01_HPF.pt")
+emulator.load_state_dict(state_dict_post)
+
+emulator.radial_velocity = nn.Parameter(torch.tensor(-9.628, device=device))
+emulator.radial_velocity.requires_grad = True
+emulator.lam_centers.requires_grad = False
+emulator.amplitudes.requires_grad = True
+emulator.sigma_widths.requires_grad = False
[…]

### 2022-03-04 paper/paper1/ms.tex
Ok, figures are looking good!

@@
     \centering
-    \includegraphics[width=0.8\textwidth]{figures/blase_WASP69_transfer_learned.png}
-    \caption{Semi-empirical model of WASP 69 transfer-learned with \texttt{blas\'e}. }
+    \includegraphics[width=0.8\textwidth]{figures/blase_WASP69_regularized.png}
+    \caption{Semi-empirical model of WASP 69 transfer-learned with \texttt{blas\'e}, employing a regularization prior on the learned amplitudes.}
     \label{fig_WASP69_transferred}
+\end{figure}
+
+\begin{figure}[hbt!]
+    \centering
+    \includegraphics[width=0.8\textwidth]{figures/blase_super_resolution_template.png}
+    \caption{The native-resolution semi-empirical model transfer-learned from WASP 69 HPF data. The revised model can be viewed as a super-resolution deconvolution of the HPF spectrum.}
+    \label{fig_WASP69_regularized}
 \end{figure}

### 2022-03-04 experiments/07_regularize_real_data/07_B_regularize_real_data.py

@@ -89,7 +89,7 @@ optimizer = optim.Adam(
     0.01,
     amsgrad=True,
 )
-n_epochs = 1000
+n_epochs = 300
 losses = []
 
 with torch.no_grad():
@@ -104,13 +104,17 @@ def ln_prior(amplitude_vector):
     return 0.5 * torch.sum((amplitude_difference ** 2) / (0.01 ** 2))
 
 
+# We need uncertainty to be able to compute the posterior
+# Assert fixed per-pixel uncertainty for now
+per_pixel_uncertainty = torch.tensor(0.004, device=device, dtype=torch.float64)
+
 t_iter = trange(n_epochs, desc="Training", leave=True)
 for epoch in t_iter:
     model.train()
     emulator.train()
     high_res_model = emulator.forward()
     yhat = model.forward(high_res_model)
-    loss = loss_fn(yhat, data_target)
+    loss = loss_fn(yhat / per_pixel_uncertainty, data_target / per_pixel_uncertainty)
     loss += ln_prior(emulator.amplitudes)
     loss.backward()
     optimizer.step()

### 2022-03-04 paper/paper1/ms.tex
The paper is coming along nicely!

@@
 We use the Adam optimizer \citep{2014arXiv1412.6980K} with a typical learning rate $LR\in (0.005, 0.05)$ and all the defaults for \texttt{PyTorch v1.11}.
-Minitbatch testing proceeded with batch fractions of typically $f\sim1-5\%$ of the available pixels but was rapidly discontinued in favor of the much more performant and convenient sparse matrix method.
+Minibatch testing proceeded with batch fractions of typically $f\sim1-5\%$ of the available pixels but was rapidly discontinued in favor of the much more performant and convenient sparse matrix method.
 We defined the number of training epochs $N_{epoch}=100-10,000$ depending on the application.
@@
 Figure \ref{fig_cloned_spectrum_demo} shows a portion of a PHOENIX spectrum cloned with \emph{blas\'e}.
-The $NN$ epochs of training took X minutes on an RTX2070 GPU with PyTorch v1.11.x, CUDA v11.x, and Intel X CPU, with all tensors as FP64.
-That benchmark works out to X seconds per epoch, or about Z seconds per minibatch.
+The $1000$ epochs of training took 56 seconds on an NVIDIA\textsuperscript{\tiny\textregistered} RTX2070 GPU with \texttt{PyTorch v1.11}, \texttt{CUDA v11.1}, and Intel\textsuperscript{\tiny\textregistered} Core\textsuperscript{\tiny TM} i7-9750H CPUs at 2.60GHz, with all tensors as FP64.
+The same computation on a 2020 M1 Macbook Air took 1$^h$25$^m$ with \texttt{PyTorch v1.9}, $90\times$ slower than the GPU counterpart.
 
-We save the model parameters with \texttt{torch.save()} and refer to the entire collection of parameters as the \emph{pre-trained model}.
+We store the model parameters to disk and refer to the entire collection of parameters as the \emph{pre-trained model}.
 
@@
 
-\subsection{Results and scaling}
+We compute the residual $\mathsf{R}(\bm{\lambda}_s)$ of native PHOENIX minus cloned model, illustrated in the bottom panel of Figure \ref{fig_cloned_spectrum_demo}.
+We see an RMS residuals of 1.2\%/pixel at native resolution.
+We see three main categories of cloning flaws.
 
-We compute the residual $\mathsf{R}(\bm{\lambda}_s)$ of native PHOENIX minus cloned model, illustrated in the bottom panel of Figure \ref{fig_cloned_spectrum_demo}\authorcomment1{make this figure!}.
-We see an RMS of $\blackout{1}\%$ residuals at native resolution, with a long tail of residuals up to $\blackout{20}\%$.
-The greatest residuals surround the deepest and broadest lines---such as Hydrogen and neutral alkali metal lines---for which the non-Voigt lineshapes stand out.
-Figure \ref{fig_zoom_cloning_performance}\authorcomment1{make this figure} zooms in on the \censor{XX} line at $\censor{10200}\;$\AA.
+The first, and expected, source of large residuals is simply missing line opacity due to our finite prominence threshold.
+Lines with prominence less than $P_{rom}$ yield residual notches with strengths comparable to $P_{rom}$.
+Including smaller prominence lines by lowering $P_{rom}$ produces smaller residuals, at the tradeoff of more computing more lines and yielding higher computational cost.
+But at some point, turning down $P_{rom}$ yields diminishing returns, as other imperfections provide a noise floor.
+We have experimentally determined this noise floor to occur near $P_{rom}=0.01$.
+
+Second, another anticipated flaw occurs in the line cores of relatively narrow lines, where the pseudo-Voigt profile becomes a poor approximation of the exact Voigt profile.
+The cloned model tends to overestimate the flux at the core and underestimate the flux alone the slopes of the lines.
+
+Finally, and most perniciously, a large category of residuals stems from conspicuous residuals appear near the wings of the deepest and broadest lines---such as Hydrogen and neutral alkali metal lines.
+The true lines exhibit advanced lineshapes, such as non-Lorentzian line-wings that are not captured with the overly simplistic Voigt line profile.
+Figure \ref{fig_zoom_cloning_performance} highlights super-Lorentzian line wings around a line at 8691~\AA.
+Narrow lines devolve into missing linewing opacity, the favored tradeoff when the continuum estimate's poor performance outweighs the pain of a narrow-but-tolerably-small spike.
+This flaw can be seen  where a line initialized at 8692.5~\AA and another pair of lines at 8690.0 all melt into linewings.
 
@@
 \end{figure}
-
-The other main source of large residuals is simply missing line opacity due to our finite prominence threshold.
-Lines with prominence less than $P_{rom}$ yield residuals that scale with $P_{rom}$.
-Including smaller prominence lines means smaller residuals, but more lines and higher computational cost.
-
-\subsection{Line data clustering and conspicuous flaws}
-We see some narrow lines devolve into missing continuum opacity, the favored tradeoff when the continuum estimate's poor performance outweighs the pain of a narrow-but-tolerably-small spike.
-This problem can be seen in Figure \ref{fig_zoom_cloning_performance} where a line initialized at $\censor{10210}\;$\AA~ with an amplitude of ZZ and width of YY bloats to \censor{some other value}.
-These devolved lines tend to appear at poorly imitated broad-and-deep line wings.
-One can mitigate these devolved lines by adding more flexibility to the continuum or supporting more advanced line shapes.
-
-\begin{outline}
-    \1 We see some lines that devolve into missing continuum opacity
-    \1 How do the lineshapes cluster in $\sigma_G$ vs $FWHM_L$?
-    \1 How do the lineshapes cluster in Amplitude vs FWHM? Can some lines be rejected?
-    \1 Does the pre-trained model round-trip effectively (model saved at end of training = model from loaded params)?
-\end{outline}
 
@@
 
+\begin{outline}
+    \1 We see some lines that devolve into missing continuum opacity
+    \1 How do the lineshapes cluster in $\sigma_G$ vs $FWHM_L$?
+    \1 How do the lineshapes cluster in Amplitude vs FWHM? Can some lines be rejected?
+\end{outline}
+
 \subsection{Lineshapes choice and tradeoffs}

### 2022-03-07 paper/paper1/ms.tex
writing prose: semi empirical model and extrinsic parameters

@@
 Lines with prominence less than $P_{rom}$ yield residual notches with strengths comparable to $P_{rom}$.
-Including smaller prominence lines by lowering $P_{rom}$ produces smaller residuals, at the tradeoff of more computing more lines and yielding higher computational cost.
+Including smaller prominence lines by lowering $P_{rom}$ produces smaller residuals, at the tradeoff of computing more lines and yielding higher computational cost.
 But at some point, turning down $P_{rom}$ yields diminishing returns, as other imperfections provide a noise floor.
@@
 Second, another anticipated flaw occurs in the line cores of relatively narrow lines, where the pseudo-Voigt profile becomes a poor approximation of the exact Voigt profile.
-The cloned model tends to overestimate the flux at the core and underestimate the flux alone the slopes of the lines.
+The cloned model tends to overestimate the flux at the core and underestimate the flux along the slopes of the lines.
 
-Finally, and most perniciously, a large category of residuals stems from conspicuous residuals appear near the wings of the deepest and broadest lines---such as Hydrogen and neutral alkali metal lines.
+Finally, and most perniciously, a large category of residuals appear near the wings of the deepest and broadest lines---such as Hydrogen and neutral alkali metal lines.
 The true lines exhibit advanced lineshapes, such as non-Lorentzian line-wings that are not captured with the overly simplistic Voigt line profile.
@@
 Narrow lines devolve into missing linewing opacity, the favored tradeoff when the continuum estimate's poor performance outweighs the pain of a narrow-but-tolerably-small spike.
-This flaw can be seen  where a line initialized at 8692.5~\AA and another pair of lines at 8690.0 all melt into linewings.
+This flaw can be seen  where a line initialized at 8692.5~\AA~ and another pair of lines at 8690.0 all melt into linewings.
 
@@
 
-\section{Refining the model from data with transfer learning}
+\section{Semi-empirical models with transfer learning techniques}
 
@@
 \end{figure}
+
+The cloned model already represents a useful intermediate product: the distillation of $N_s=335,849\times2$ pixel flux values and coordinates into a more compact quartet of properties for a list of 9,028 spectral lines, a dimensionality reduction of $18\times$ for the cost of 1.2\%/pixel in accuracy.
+But the cloned model serves as a mere stepping stone in our principal quest: the comparison of models to real data.
+
+\subsection{Augmenting the cloned model to match data}
+Real stars possess two key extrinsic properties.
+Rotational broadening $v\sin{i}$ and radial velocity $RV$ depend on the observer's viewing location.
+The qualifier \emph{extrinsic} distiguishes from stellar \emph{intrinsic} properties, such as $T_{\mathrm{eff}}, \log{g}, and [\mathrm{Fe}/\mathrm{H}]$, which appear the same from any viewing location, at least for stars with isotropic surfaces.
+The distinction is important because the extrinsic terms act as simple convolutions and translations to the cloned spectrum.
+We therefore define an augmented model, which we designate the ``extrinsic'' model, $\mathsf{F}_{ex}$:
+
+\begin{eqnarray}
+    \mathsf{F}_{ex}(\bm{\lambda}) = \mathsf{F}(\bm{v}-RV) * \zeta \left(\frac{\bm{v}}{v\sin{i}}\right) \label{eqn_convolution}
+\end{eqnarray}
+
+where $\zeta$ is the convolution kernel for rigid body rotation \censorbox{cite ExoJAX} and $\bm{v}$ is the spectral axis represented as relative velocity coordinates.
+
+Operationally, the Doppler shift gets applied to the line center positions $\tilde \lambda_{c,i} = (\lambda_{c,i}-\delta \lambda)$, and not the individual wavelength grid point coordinates, $\bm{lambda}$.
+This choice yields a key benefit: it makes the $RV$ autodiff-aware, meaning that an infinitesimal change to the RV value can easily be sensed through backpropagation.
+The strategy allows relatively low computational and memory cost since it comports with the sparse matrix implementation.
+
+The instrumental resolving power, $R$, acts as a convolution with a Gaussian line profile of width $\sigma=$\censorbox{todo}.
+Real astronomical instruments usually have wavelength-dependent resolving power, which complicates the implementation.
+The extent to which this effect matters will depend on the science case.
+For now, \emph{blas\'e} simply assumes a fixed resolving power.
+
+The pixel resampling step could be achieved with two choices.
+One would be to simply evaluate the closed-form model at the coordinates of the data pixels $\bm{\lambda_D}$.
+This choice would require a closed-form solution to Equation \ref{eqn_convolution}, which has recently been reported \citep{2021arXiv211006271L}.
+Instead, we choose the conventional sampled approach, in which we evaluate the extrinsic model at all of the native resolution wavelength coordinates, and then compute the mean value of those pixels within the bounds of each coarse data pixel.
+The resampling procedure is autodiff-aware: the same clusters of native resolution coordinates map to the same data pixel coordinates.
 

### 2022-03-07 paper/paper1/ms.tex
woohoo, the transfer learning section is close to complete

@@
 \usepackage{amsmath,bm}
+
+\DeclareMathOperator{\resample}{resample}
 
@@
 
-\begin{figure}[hbt!]
-    \centering
-    \includegraphics[width=0.8\textwidth]{figures/blase_WASP69_demo.png}
-    \caption{WASP 69 observed with HPF, compared to a $T_{\mathrm{eff}}=4700\;$K,  $\log{g}=4.5$ solar metallicity PHOENIX model warped to $v\sin{i}=2.2$~km/s, $RV=-9.6$~km/s, and HPF resolving power.}
-    \label{fig_WASP69_demo}
-\end{figure}
-
-\begin{figure}[hbt!]
-    \centering
-    \includegraphics[width=0.8\textwidth]{figures/blase_WASP69_regularized.png}
-    \caption{Semi-empirical model of WASP 69 transfer-learned with \texttt{blas\'e}, employing a regularization prior on the learned amplitudes.}
-    \label{fig_WASP69_transferred}
-\end{figure}
-
-\begin{figure}[hbt!]
-    \centering
-    \includegraphics[width=0.8\textwidth]{figures/blase_super_resolution_template.png}
-    \caption{The native-resolution semi-empirical model transfer-learned from WASP 69 HPF data. The revised model can be viewed as a super-resolution deconvolution of the HPF spectrum.}
-    \label{fig_WASP69_regularized}
-\end{figure}
-
 The cloned model already represents a useful intermediate product: the distillation of $N_s=335,849\times2$ pixel flux values and coordinates into a more compact quartet of properties for a list of 9,028 spectral lines, a dimensionality reduction of $18\times$ for the cost of 1.2\%/pixel in accuracy.
@@
 Instead, we choose the conventional sampled approach, in which we evaluate the extrinsic model at all of the native resolution wavelength coordinates, and then compute the mean value of those pixels within the bounds of each coarse data pixel.
-The resampling procedure is autodiff-aware: the same clusters of native resolution coordinates map to the same data pixel coordinates.
+The resampling procedure is autodiff-aware: the same clusters of native resolution coordinates map to the same data pixel coordinates, no matter what the $RV$ is.
+The $RV$ only dictates what flux values are realized within those pixels.
+The final forward model for \emph{blas\'e} is designated the ``pixel-level'' model to emphasize that we have achieved a model for each datum in the 1D observed spectrum:
 
-\begin{outline}
-    \1 Augment the pre-trained model with extrinsic parameters:
-    \2 vsini and RV
-    \1 How to re-sample into the data space?
-    \1 Option 1: Convolve the Pseudo-Voigt with a Instrumental Gaussian
-    \2 The convolution of a Gaussian with a Pseudo-Voigt is simply a new Pseudo-Voigt!
-    \2 Simply evaluate on a new coarse wavelength grid! Saves computation!
-    \1 Option 2: Numerically convolve output on a fine grid and sum pixels
-    \2 Computationally wasteful
-    \2 Requires generating full bandwidth spectrum at native resolution
-    \1 Both options should give the same answer (right?)
-    \2 Need to double-check commutativity of convolution and element-wise product...
-    \1 Ideally you need wavelength-dependent resolution, not just a single stock resolving power
-    \1 Need accurate per-pixel uncertainties
-    \1 We have to apply some amount of regularization to the pre-trained model
-    \2 Regularization tunes how much to trust the data versus the model in the semi-empirical model
-    \2 No Regularization means likely to overfit noise in the data
-    \2 Too much regularization resembles a static (fixed) PHOENIX model
-\end{outline}
+\begin{eqnarray}
+    \mathsf{F}_{p}(\bm{\lambda}) = \resample{\Big[\mathsf{F}_{ex} * g(R) \Big]} \label{eqn_final_model}
+\end{eqnarray}
+
+where $g$ is the Gaussian instrumental convolution kernel and the \texttt{resample[]} operation indicates the average of pixels within the data boundaries.
+
+\subsection{Regularization}
+
+Equation \ref{eqn_final_model} has all the same conceivably tunable parameters as the cloned model, \emph{plus} $v\sin{i}$ and $RV$.
+The resolving power may also be treated as tunable if it is not known or varies with \emph{e.g.} seeing or instrumental configuration.
+
+It may appear desirable to simply optimize all of these parameters in a \emph{laissez-faire} manner, allowing them all to take on whatever value the data dictates.
+Such a stratagem would overfit the data, resulting in unphysically perverse lineshapes that do not reflect the air of reality we attempt to impose on our synthetic spectral models.
+Lines would haphazardly fit noise-spikes, and conspire together to warp spectral shapes in unexpected ways.
+This overfit model may suit some rare purposes.
+But most of the time, we prefer to strike a better balance in the bias-variance tradeoff.
+
+So instead, we have to apply some amount of \emph{regularization}, a restiction on the allowed values the model parameters can take on.
+Fortunately we have a firm theoretical basis to justify this regularization.
+We believe our precomputed synthetic spectral models are \emph{quasi-statically correct}: the predicted spectra resemble the unobserved ``latent spectrum'' with lines in the correct place, but just with the wrong amplitudes.
+This statement may stem from the fact that it is easier to predict the mere existence of some energy transition of atoms and molecules than it is to predict their transition rates, abundances, temperature and pressure effects, and all the other line strength effects that flow down to how much light a line ultimately absorbs in a stellar atmosphere.
+
+Even still, the regularization step is in some ways the most subjective, since the degree of regularization will control the degree of overfitting or underfitting.
+The most extreme regularization---the antithesis of the \emph{laissez-faire} scenario---would yield a model too rigid to respond to the data at all, yielding a model entirely unchanged from the cloned PHOENIX model.
+So setting the regularization parameter constitutes the only hyperparameter worthy of tuning in \texttt{blas\'e}.
+The choice of how to set the regularization is problem-specific.
+We default to the following choice.
+We fix all line parameters except for amplitude, which receives a regularization of 0.01, meaning that its ampltidue should be within about 1\% of its cloned value.
+The extrinsic $v\sin{i}$ and $RV$ have no regularization, but in practice they need to be initialized close to their plausible values.
+
+Figure \ref{fig_WASP69_demo} shows an example of comparing HPF data to the closest PHOENIX template, post-processed to the HPF resolution and sampling.
+Figure \ref{fig_WASP69_transferred} shows the same data spectrum compared to a pixel-level \emph{blas\'e} model regularized to the clone in Figure \ref{fig_cloned_spectrum_demo}.
+The latent high-resolution model for WASP 69 is therefore the transfer-learned model, unadjusted for the extrinsic and instrumental properties.
+This semi-empirical model is shown in Figure \ref{fig_WASP69_regularized}.
+
+\begin{figure}[hbt!]
+    \centering
+    \includegraphics[width=0.8\textwidth]{figures/blase_WASP69_demo.png}
+    \caption{WASP 69 observed with HPF, compared to a $T_{\mathrm{eff}}=4700\;$K,  $\log{g}=4.5$ solar metallicity PHOENIX model warped to $v\sin{i}=2.2$~km/s, $RV=-9.6$~km/s, and HPF resolving power.}
+    \label{fig_WASP69_demo}
+\end{figure}
+
+\begin{figure}[hbt!]
+    \centering
+    \includegraphics[width=0.8\textwidth]{figures/blase_WASP69_regularized.png}
+    \caption{Semi-empirical model of WASP 69 transfer-learned with \texttt{blas\'e}, employing a regularization prior on the learned amplitudes.}
+    \label{fig_WASP69_transferred}
+\end{figure}
+
+\begin{figure}[hbt!]
+    \centering
+    \includegraphics[width=0.8\textwidth]{figures/blase_super_resolution_template.png}
+    \caption{The native-resolution semi-empirical model transfer-learned from WASP 69 HPF data. The revised model can be viewed as a super-resolution deconvolution of the HPF spectrum.}
+    \label{fig_WASP69_regularized}
+\end{figure}
 

### 2022-03-08 experiments/06_bare_emulator_clone/06_bare_emulator_clone_M_dwarf.py
a deprecated M-dwarf experiment

@@ -1,0 +1,126 @@
+import os
+import torch
+from torch import nn
+from tqdm import trange
+import torch.optim as optim
+import torch.profiler
+from blase.emulator import SparsePhoenixEmulator
+import matplotlib.pyplot as plt
+from gollum.phoenix import PHOENIXSpectrum
+from torch.utils.tensorboard import SummaryWriter
+import webbrowser
+import numpy as np
+from muler.hpf import HPFSpectrum
+
+if torch.cuda.is_available():
+    device = "cuda"
+else:
+    device = "cpu"
+
+
+def plot_spectrum(spectra):
+    """
+    Generates matplotlib Figure using a trained network, along with images
+    and labels from a batch
+    """
+    fig, axes = plt.subplots(2, figsize=(8, 3))
+    axes[0].step(
+        spectra[0]["wl"].numpy(),
+        spectra[0]["flux"].numpy(),
+        label="Cloned Spectrum",
+        lw=1,
+    )
+    axes[0].step(
+        spectra[1]["wl"].numpy(),
+        spectra[1]["flux"].numpy(),
+        label="Native Spectrum",
+        lw=1,
+    )
+    axes[0].set_ylim(0, 1.5)
+    axes[0].legend(loc="upper right", ncol=2)
+    axes[1].step(
+        spectra[1]["wl"].numpy(),
+        spectra[0]["flux"].numpy() - spectra[1]["flux"].numpy(),
+    )
+    axes[1].set_ylim(-0.3, 0.3)
+    return fig
+
+
+log_dir = "runs/synthetic1"
+writer = SummaryWriter(log_dir=log_dir)
+webbrowser.open("http://localhost:6006/", new=2)
+
+
+# Pre-process the model as described in the paper
+spectrum = PHOENIXSpectrum(teff=2800, logg=5.0)
+spectrum = spectrum.divide_by_blackbody()
+spectrum = spectrum.normalize()
+continuum_fit = spectrum.fit_continuum(polyorder=5)
+spectrum = spectrum.divide(continuum_fit, handle_meta="ff")
+
+# Numpy arrays: 1 x N_pix
+wl_native = spectrum.wavelength.value
+flux_native = spectrum.flux.value
+
+# Create the emulator and load a pretrained model
+prominence = 0.01
+emulator = SparsePhoenixEmulator(
+    wl_native, flux_native, prominence=prominence, wing_cut_pixels=1000
+)
+emulator.to(device)
+
+wl_native = emulator.wl_native.clone().detach().to(device)
[…]

### 2022-03-08 experiments/06_bare_emulator_clone/06_bare_emulator_clone_M_dwarf.py
instantly delete it

@@ -1,126 +1,0 @@
-import os
-import torch
-from torch import nn
-from tqdm import trange
-import torch.optim as optim
-import torch.profiler
-from blase.emulator import SparsePhoenixEmulator
-import matplotlib.pyplot as plt
-from gollum.phoenix import PHOENIXSpectrum
-from torch.utils.tensorboard import SummaryWriter
-import webbrowser
-import numpy as np
-from muler.hpf import HPFSpectrum
-
-if torch.cuda.is_available():
-    device = "cuda"
-else:
-    device = "cpu"
-
-
-def plot_spectrum(spectra):
-    """
-    Generates matplotlib Figure using a trained network, along with images
-    and labels from a batch
-    """
-    fig, axes = plt.subplots(2, figsize=(8, 3))
-    axes[0].step(
-        spectra[0]["wl"].numpy(),
-        spectra[0]["flux"].numpy(),
-        label="Cloned Spectrum",
-        lw=1,
-    )
-    axes[0].step(
-        spectra[1]["wl"].numpy(),
-        spectra[1]["flux"].numpy(),
-        label="Native Spectrum",
-        lw=1,
-    )
-    axes[0].set_ylim(0, 1.5)
-    axes[0].legend(loc="upper right", ncol=2)
-    axes[1].step(
-        spectra[1]["wl"].numpy(),
-        spectra[0]["flux"].numpy() - spectra[1]["flux"].numpy(),
-    )
-    axes[1].set_ylim(-0.3, 0.3)
-    return fig
-
-
-log_dir = "runs/synthetic1"
-writer = SummaryWriter(log_dir=log_dir)
-webbrowser.open("http://localhost:6006/", new=2)
-
-
-# Pre-process the model as described in the paper
-spectrum = PHOENIXSpectrum(teff=2800, logg=5.0)
-spectrum = spectrum.divide_by_blackbody()
-spectrum = spectrum.normalize()
-continuum_fit = spectrum.fit_continuum(polyorder=5)
-spectrum = spectrum.divide(continuum_fit, handle_meta="ff")
-
-# Numpy arrays: 1 x N_pix
-wl_native = spectrum.wavelength.value
-flux_native = spectrum.flux.value
-
-# Create the emulator and load a pretrained model
-prominence = 0.01
-emulator = SparsePhoenixEmulator(
-    wl_native, flux_native, prominence=prominence, wing_cut_pixels=1000
-)
-emulator.to(device)
-
-wl_native = emulator.wl_native.clone().detach().to(device)
[…]

### 2022-03-08 paper/paper1/ms.tex
Make the performance a subsection, not section
Replace placeholder text with actual links and citations

@@
 Automatic differentiation \citep[``autodiff'',][]{2015arXiv150205767G} and its affiliated backpropagation algorithm has revolutionized machine learning and neural network architecture design, and is increasingly applied in astrophysical data analysis contexts, \emph{e.g.} kernel phase coronography with \texttt{poppy} \citep{2021ApJ...907...40P}, and exoplanet orbit fitting with \texttt{exoplanet} \citep{2021JOSS....6.3285F}.
-Of the spectroscopy frameworks mentioned above, the \texttt{TensorFlow}-based \citep{tensorflow2015-whitepaper} \texttt{wobble} and the \texttt{JAX}-based \censorbox{jax2018github} \texttt{ExoJAX} employ autodiff technology.
+Of the spectroscopy frameworks mentioned above, the \texttt{TensorFlow}-based \citep{tensorflow2015-whitepaper} \texttt{wobble} and the \texttt{JAX}-based \citep{jax2018github} \texttt{ExoJAX} employ autodiff technology.
 \texttt{wobble} treats each pixel as a tunable control point, producing $\sim10^5$ parameters for a modern stellar spectrum.
@@
 We show how to supplant the computationally expensive radiative transfer by cloning pre-computed synthetic spectral models.
-In Section \censorbox{XX} we describe how to fine-adjust the cloned models to data in a transfer-learning step.
-Section \censorbox{XX} exhibits some injection-recovery type tests on simulated data of known properties.
-In Section \censorbox{XX} we show real-world data from high-grasp near-IR spectrographs.
-Finally we discuss many conceivable extensions to the framework, and directions for future work in Section \censorbox{XX}.
+In Section \ref{transferLearn} we describe how to fine-adjust the cloned models to real-world data from high-grasp near-IR spectrographs, using a transfer-learning step.
+Section \ref{secInjRec} exhibits injection-recovery tests on simulated data of known properties, revealing the scalability and fundamental limits of the technique.
+Finally we discuss many conceivable extensions to the framework, and directions for future work in Section \ref{secDiscuss}.
 
@@
 Evaluation of the Fadeeva function can be computationally costly, and so approximate forms may be desirable.
-Here we adopt the pseudo-Voigt approximation described in \censor{some paper}\authorcomment1{look up reference to PseudoVoigt}.
+Here we adopt the pseudo-Voigt approximation \citep{Ida:nt0146}.
 
@@
 
-\section{Semi-empirical models with transfer learning techniques}
+\section{Semi-empirical models with transfer learning techniques}\label{transferLearn}
 
@@
 
-where $\zeta$ is the convolution kernel for rigid body rotation \censorbox{cite ExoJAX} and $\bm{v}$ is the spectral axis represented as relative velocity coordinates.
+where $\zeta$ is the convolution kernel for rigid body rotation \citep{2022ApJS..258...31K} and $\bm{v}$ is the spectral axis represented as relative velocity coordinates.
 
@@
 
-The instrumental resolving power, $R$, acts as a convolution with a Gaussian line profile of width $\sigma=$\censorbox{todo}.
+The instrumental resolving power, $R$, acts as a convolution with a Gaussian line profile of width $\sigma=\frac{c}{2.355 R}$.
 Real astronomical instruments usually have wavelength-dependent resolving power, which complicates the implementation.
@@
 
-\section{Performance of transfer learning}
+\subsection{Performance of transfer learning}
+
 \begin{outline}
@@
 
-\section{Discussion}
+\section{Injection/recovery tests}\label{secInjRec}
+Here we conduct injection/recovery tests.
+
+\section{Discussion}\label{secDiscuss}
 
@@
     \1 What about PRV applications
-    \1 What about the choice of minibatch size?
-    \1 How would a sparse tensor change the computational performance?
     \1 Are we using the GPU effectively? Are we memory bandwidth limited?
@@
 \begin{outline}
-    \1 What to do when continuum is absent even in at the native resolution of precomputed synthetic model?
+    \1 What to do when continuum is absent even at the native resolution of precomputed synthetic model?
     \1 Relevant to brown dwarfs and possibly M dwarfs molecular bands

### 2022-03-08 paper/paper1/ms.tex
Add some grant acknowledgements, tweak formatting

@@
 \begin{document}
-\shorttitle{blas\'e}
-\shortauthors{Gully-Santiago}
+\shorttitle{Blas\'e}
+\shortauthors{Gully \& Morley}
 \title{Transfer learning for \'echelle spectroscopy with \texttt{blas\'e}}
@@
 
-Equation \ref{equation1} performs a type of tensor contraction, turning a $N_{\mathrm{lines}}\times N_{s}$ matrix into a $1\times N_{s}$ vector.
+Equation \ref{equation1} performs a matrix contraction, turning a $N_{\mathrm{lines}}\times N_{s}$ matrix into a $1\times N_{s}$ vector.
 The number of Floating Point Operations (FLOPS) scales with the number of entries in this matrix, with a scalar prefactor for the cost of evaluating a single line profile at a single wavelength point.
@@
 Rotational broadening $v\sin{i}$ and radial velocity $RV$ depend on the observer's viewing location.
-The qualifier \emph{extrinsic} distiguishes from stellar \emph{intrinsic} properties, such as $T_{\mathrm{eff}}, \log{g}, and [\mathrm{Fe}/\mathrm{H}]$, which appear the same from any viewing location, at least for stars with isotropic surfaces.
+The qualifier \emph{extrinsic} distiguishes from stellar \emph{intrinsic} properties, such as $T_{\mathrm{eff}}, \log{g},\mathrm{and\,} [\mathrm{Fe}/\mathrm{H}]$, which appear the same from any viewing location, at least for stars with isotropic surfaces.
 The distinction is important because the extrinsic terms act as simple convolutions and translations to the cloned spectrum.
@@
 
-Operationally, the Doppler shift gets applied to the line center positions $\tilde \lambda_{c,i} = (\lambda_{c,i}-\delta \lambda)$, and not the individual wavelength grid point coordinates, $\bm{lambda}$.
+Operationally, the Doppler shift gets applied to the line center positions $\tilde \lambda_{c,i} = (\lambda_{c,i}-\delta \lambda)$, and not the individual wavelength grid point coordinates, $\bm{\lambda}$.
 This choice yields a key benefit: it makes the $RV$ autodiff-aware, meaning that an infinitesimal change to the RV value can easily be sensed through backpropagation.
@@
 \begin{acknowledgements}
-    The author acknowledges the Texas Advanced Computing Center (TACC, \url{http://www.tacc.utexas.edu}) at The University of Texas at Austin for providing HPC resources that have contributed to the research results reported within this paper.
+    This material is based upon work supported by the National Aeronautics and Space Administration under Grant Numbers 80NSSC21K0650 for the NNH20ZDA001N-ADAP:D.2 program, and 80NSSC20K0257 for the XRP program issued through the Science Mission Directorate.
+We acknowledge the National Science Foundation, which supported the work presented here under Grant No.
+1910969.
+This research has made use of NASA's Astrophysics Data System Bibliographic Services.
 \end{acknowledgements}

### 2022-03-10 paper/paper1/ms.tex
Add a flowchart schematic, fill in placeholders, and fill in the table of symbols

@@
 \bibliographystyle{aasjournal}
-\turnoffedit
+%\turnoffedit
 \usepackage[caption=false]{subfig}
@@
 Existing open-source frameworks have overcome some of these challenges, or have been purpose-built for specialized applications.
-These frameworks include \texttt{ROBOSPECT} \citep{2013PASP..125.1164W}, \texttt{specmatch} \citep{2015PhDT........82P}, \texttt{specmatch-emp} \citep{2017ApJ...836...77Y}, \texttt{wobble} \citep{2019AJ....158..164B}, \texttt{starfish} \citep{czekala15}, \texttt{sick} \citep{2016ApJS..223....8C}, \texttt{psoap} \citep{2017ApJ...840...49C}, \texttt{FAL} \authorcomment1{Cargile et al.
-in prep}, CHIMERA \citep{2015ApJ...807..183L}, the \texttt{Cannon} \citep{2017ApJ...836....5H},  \texttt{MINESweeper} \citep{2020ApJ...900...28C}, and recently \texttt{ExoJAX} \citep{2022ApJS..258...31K}.
+These frameworks include \texttt{ROBOSPECT} \citep{2013PASP..125.1164W}, \texttt{specmatch} \citep{2015PhDT........82P}, \texttt{specmatch-emp} \citep{2017ApJ...836...77Y}, \texttt{wobble} \citep{2019AJ....158..164B}, \texttt{starfish} \citep{czekala15}, \texttt{sick} \citep{2016ApJS..223....8C}, \texttt{psoap} \citep{2017ApJ...840...49C}, \texttt{FAL} (Cargile et al.
+\emph{in prep}), CHIMERA \citep{2015ApJ...807..183L}, the \texttt{Cannon} \citep{2017ApJ...836....5H},  \texttt{MINESweeper} \citep{2020ApJ...900...28C}, and recently \texttt{ExoJAX} \citep{2022ApJS..258...31K}.
 The design of these frameworks necessarily have to make a choice in the bias-variance tradeoff: is the tool more \emph{data}-driven or more \emph{model}-driven?
@@
 \subsection{Voigt profile evaluation}
-The Voigt profile $\mathsf{V}(\lambda, \sigma_j, \gamma_j)$ can be computed in exact closed-form using the Fadeeva function\authorcomment1{citation}.
+The Voigt profile $\mathsf{V}(\lambda, \sigma_j, \gamma_j)$ can be computed in exact closed-form using the Voigt-Hjerting function \citep{1938ApJ....88..508H} as the real part of the complex Fadeeva function \citep[\emph{e.g.}][]{2011arXiv1106.0151Z}.
 Evaluation of the Fadeeva function can be computationally costly, and so approximate forms may be desirable.
@@
 We do not currently implement these sampled convolutions in \emph{blas\'e} due to their prohibitive computational cost.
+
+\begin{figure}[hbt!]
+    \centering
+    \includegraphics[width=0.98\textwidth]{figures/blase_flowchart01.png}
+    \caption{Visual flowchart of \texttt{blas\'e}.
+The technique starts with a precomputed synthetic stellar spectrum and a precomputed synthetic telluric spectrum with properties close to those previously estimated for target.
+Both the stellar and telluric spectra get cloned.
+The stellar model is warped to its extrinsic properties, and then the stellar and telluric models get multiplied together.
+Next this joint model is convolved with an instrumental kernel and resampled to the data spectrum.
+Finally the joint model parameters are optimized to match the data, subject to regularization (not shown).
+The final products are semi-empirical super-resolution stellar and telluric templates, the $RV$ and $v\sin{i}$ of the star, and the line amplitude, location, width, and shape of all 10,000+ stellar and telluric lines.
+\authorcomment1{Some of the spectra are placeholders until we re-run the end-to-end-fitting.
+Feedback on this figure layout is sought before investing fine-adjusting the spectra cutouts.} }
+    \label{blase_flowchart}
+\end{figure}
 
@@
 The optimizer updates the $a_j, \sigma_j, \gamma_j$ parameters by a small fraction of the Jacobian---called the learning rate LR---towards the direction that would improve the fit, for all $N_{\mathrm{lines}} \times 3$ parameters simultaneously.
-The Jacobian is calculated behind the scenes with automatic differentiation\authorcomment1{Citation} implemented as the so-called backpropagation algorithm or simply ``backprop''.
+The Jacobian is calculated behind the scenes with automatic differentiation implemented as the so-called backpropagation algorithm or simply ``backprop'' \citep{2015arXiv150205767G}.
 We choose the \texttt{PyTorch} framework that computes these Jacobians efficiently for all of the mathematical primitives in our \texttt{blas\'e} implementation.
@@
 In principle, we could compute the Jacobians analytically---since the partial derivatives of the Voigt function are known in closed-form---circumventing the vendor lock-in and memory overhead of an autodiff implementation with PyTorch.
-Further, an autodiff-aware implementation of the \emph{celerit\'e} algorithm\authorcomment1{add DFM citation} does not yet exist in PyTorch, constricting our choices for simultaneous continuum fitting.
-In practice, PyTorch also offers hardware acceleration (on GPU or TPU)\authorcomment1{citation needed}, mature support for a range of optimizers, and other perks that motivated its choice over analytic implementations or other machine learning frameworks.
-The GPyTorch framework\authorcomment1{add citation} may offer a workaround to the absence of a \emph{celerit\'e} implementation.
+Further, an autodiff-aware implementation of the \emph{celerit\'e} algorithm \citep{2017AJ....154..220F} does not yet exist in PyTorch, constricting our choices for simultaneous continuum fitting.
+In practice, PyTorch also offers hardware acceleration on GPU or TPU \citep{10.1145/3079856.3080246}, mature support for a range of optimizers, and other perks that motivated its choice over analytic implementations or other machine learning frameworks.
+The GPyTorch framework \citep{2018arXiv180911165G} may offer a workaround to the absence of a \emph{celerit\'e} implementation.
 Still, the performance of an analytic Jacobian implementation over the autodiff route presented here offers an interesting avenue of future research.
@@
 Minibatches offer a natural solution to the line-by-line computational bottleneck while also providing optimization performance improvements.
-Minibatches act as a form of regularization, the principal source of stochasticity in the Stochastic Gradient Descent algorithm, which tends to have better convergence than full-batch Gradient Descent\authorcomment1{citation}.
+Minibatches act as a form of regularization, the principal source of stochasticity in the Stochastic Gradient Descent algorithm, which tends to have better convergence than full-batch Gradient Descent \citep{2016arXiv160904747R}.
 The Machine Learning/Artificial Intelligence (ML/AI) community employs minibatches to handle similar situations, in which the dataset volume exceeds the available GPU RAM.
@@
 % Table
-\begin{deluxetable}{cc}
+\begin{deluxetable}{cl}
     \tablecaption{Notation used in this paper\label{table2}}
     \tablehead{
-        \colhead{Symbol} & \nocolhead{Meaning}
+        \colhead{Symbol} & \colhead{Meaning}
     }
     \startdata
-    $N_{pix}$ & Number of pixels in the precomputed spectrum \\
-    $N_{lines}$ & Number of spectral lines \\
-    $\sigma_G$ & Gaussian line profile scale and standard deviation \\
-    $\gamma_L$ & Lorentzian line profile half width\\
-    $\pm \Delta \lambda_{\mathrm{buffer}}$ & Buffer exceeding the red and blue limits of data spectrum
+    \hline
+    \multicolumn{2}{c}{Spectra}\\
+    \hline
+    $\bm{\lambda}_S$ & Native wavelength coordinates of the precomputed stellar spectrum\\
+    $\bm{\lambda}_T$ & Native wavelength coordinates of the telluric spectrum\\
+    $\bm{\lambda}_D$ & Native wavelength coordinates of the data spectrum\\
+    $\mathsf{S}$ & Flux values of the precomputed synthetic stellar spectral model $\bm{\lambda}_S$\\
+    $\mathsf{T}$ & Transmission values of the precomputed synthetic telluric model \\
+    $\mathsf{D}$ & The observed data spectrum flux values\\
+    $\mathsf{B}$ & Blackbody of temperature $T_{\mathrm{eff}}$ to coarsely normalize $\mathsf{S}$\\
+    $\mathsf{P}$ & Smooth polynomial to refine continuum-normalization of $\mathsf{S/B}$\\
+    $\mathsf{S}_{\rm clone}$ & Evaluable and tunable cloned flux model of $\mathsf{S}$\\
+    $\mathsf{T}_{\rm clone}$ & Evaluable and tunable cloned transmission model of $\mathsf{T}$\\
+    $\mathsf{T}_{\rm resamp}$ & The cloned telluric model resampled from $\bm{\lambda}_T$ to $\bm{\lambda}_S$\\
+    $\mathsf{S}_{\rm ext}$ & An augmentation of $\mathsf{S}_{\rm clone}$ with $v\sin{i}$ convolution and $RV$ translation\\
+    $\mathsf{M}_{\rm joint}$ & The joint stellar and telluric model: $\mathsf{S}_{\rm ext} \cdot \mathsf{T}_{\rm resamp}$  \\
+    $\mathsf{M}$ & Joint model convolved with instrument resolving power and resampled to $\bm{\lambda}_D$\\
+    \hline
+    \multicolumn{2}{c}{Line properties}\\
+    \hline
+    $\lambda_{c,i}$ & Line center position of the $i^{th}$ spectral line\\
+    $a_i$ & Gaussian line profile amplitude of the $i^{th}$ spectral line \\
+    $\sigma_i$ & Gaussian line profile scale of the $i^{th}$ spectral line\\
+    $\gamma_i$ & Lorentzian line profile half width of the $i^{th}$ spectral line\\
+    $\mathsf{V}_i$ & The Voigt profile of the $i^{th}$ spectral line \\
+    $\bar{\bm{V}}$ & The dense $(N_{\rm lines} \times N_{\rm pix})$ matrix of all Voigt profiles stacked vertically \\
+    $\hat{\bm{V}}$ & The sparse $(N_{\rm lines} \times N_{\rm sparse})$ matrix of all Voigt profiles stacked vertically \\
+    \hline
+    \multicolumn{2}{c}{Scalars}\\
+    \hline
+    $N_{\rm lines}$ & Number of spectral lines \\
+    $N_{\rm pix}$ & Number of pixel coordinates in the precomputed spectrum \\
+    $N_{\rm sparse}$ & Number of non-zero pixels computed in the sparse implementation\\
+    $\pm \Delta \lambda_{\mathrm{buffer}}$ & Buffer exceeding the red and blue limits of the data spectrum\\
+    $P_{\rm rom}$ & The prominence threshold of spectral lines to include in cloning \\
+    $v\sin{i}$ & Stellar projected rotational broadening at inclination $i$ and equatorial velocity $v$\\
+    $RV$ & Radial velocity of the star\\
     \enddata

### 2022-03-10 paper/paper1/ms.tex
Paper is looking tidy, nice table and figure

@@
 In extreme precision radial velocity (EPRV) applications, cross-correlation methods work \citep{2018A_A...620A..47D}, but have many limitations \citep{2022arXiv220110639Z}.
-Addressing telluric absorption at the $\mathrm{cm/s}$ level may require joint modeling of the star and the Earth's atmospheric absorption before convolution with an instrumental kernel\authorcomment1{speakerdeck.com/dwhgg/telluric-line-hack-week-wrap-up}.
+Addressing telluric absorption at the $\mathrm{cm/s}$ level may require joint modeling of the star and the Earth's atmospheric absorption before convolution with an instrumental kernel.
 This telluric joint modeling capability does not yet exist at a precision that can meet these strenuous demands.
@@
 \authorcomment1{Some of the spectra are placeholders until we re-run the end-to-end-fitting.
-Feedback on this figure layout is sought before investing fine-adjusting the spectra cutouts.} }
+Feedback on this figure layout is sought.} }
     \label{blase_flowchart}
@@
 % Table
-\begin{deluxetable}{cl}
+\begin{deluxetable}{cp{12cm}}
     \tablecaption{Notation used in this paper\label{table2}}
@@
     $\bm{\lambda}_D$ & Native wavelength coordinates of the data spectrum\\
-    $\mathsf{S}$ & Flux values of the precomputed synthetic stellar spectral model $\bm{\lambda}_S$\\
+    $\mathsf{S}_{\rm abs}$ & Flux values of the precomputed synthetic stellar spectral model $\bm{\lambda}_S$\\
+    $\mathsf{B}$ & Blackbody of temperature $T_{\mathrm{eff}}$ to coarsely normalize $\mathsf{S}_{\rm native}$\\
+    $\mathsf{P}$ & Smooth polynomial to refine continuum-normalization\\
+    $\mathsf{S}$ & Continuum normalized augmentation of $\mathsf{S}_{\rm native}$\\ 
     $\mathsf{T}$ & Transmission values of the precomputed synthetic telluric model \\
     $\mathsf{D}$ & The observed data spectrum flux values\\
-    $\mathsf{B}$ & Blackbody of temperature $T_{\mathrm{eff}}$ to coarsely normalize $\mathsf{S}$\\
-    $\mathsf{P}$ & Smooth polynomial to refine continuum-normalization of $\mathsf{S/B}$\\
+    $\bm{\epsilon}$ & The estimated uncertainties in the data spectrum\\
     $\mathsf{S}_{\rm clone}$ & Evaluable and tunable cloned flux model of $\mathsf{S}$\\
     $\mathsf{T}_{\rm clone}$ & Evaluable and tunable cloned transmission model of $\mathsf{T}$\\
-    $\mathsf{T}_{\rm resamp}$ & The cloned telluric model resampled from $\bm{\lambda}_T$ to $\bm{\lambda}_S$\\
+    %$\mathsf{T}_{\rm resamp}$ & The cloned telluric model resampled from $\bm{\lambda}_T$ to $\bm{\lambda}_S$\\
     $\mathsf{S}_{\rm ext}$ & An augmentation of $\mathsf{S}_{\rm clone}$ with $v\sin{i}$ convolution and $RV$ translation\\
-    $\mathsf{M}_{\rm joint}$ & The joint stellar and telluric model: $\mathsf{S}_{\rm ext} \cdot \mathsf{T}_{\rm resamp}$  \\
-    $\mathsf{M}$ & Joint model convolved with instrument resolving power and resampled to $\bm{\lambda}_D$\\
+    $\mathsf{M}_{\rm joint}$ & The joint stellar and telluric model: $\mathsf{S}_{\rm ext} \odot \mathsf{T}_{\rm clone}(\bm{\lambda}_S)$  \\
+    $\mathsf{M}$ & Joint model convolved 
+    with instrumental kernel and resampled to $\bm{\lambda}_D$\\
+    $\mathsf{R}$ & The residual spectrum between a pair of inputs, \emph{e.g.} $\mathsf{D} - \mathsf{M}$\\
+    $\bm{v}$ & The spectral coordinate axis $\bm{\lambda}$ expressed as a velocity difference\\
     \hline
@@
     \hline
-    $\lambda_{c,i}$ & Line center position of the $i^{th}$ spectral line\\
-    $a_i$ & Gaussian line profile amplitude of the $i^{th}$ spectral line \\
-    $\sigma_i$ & Gaussian line profile scale of the $i^{th}$ spectral line\\
-    $\gamma_i$ & Lorentzian line profile half width of the $i^{th}$ spectral line\\
-    $\mathsf{V}_i$ & The Voigt profile of the $i^{th}$ spectral line \\
-    $\bar{\bm{V}}$ & The dense $(N_{\rm lines} \times N_{\rm pix})$ matrix of all Voigt profiles stacked vertically \\
+    $\lambda_{c,j}$ & Line center position of the $j^{th}$ spectral line\\
+    $a_j$ & Gaussian line profile amplitude of the $j^{th}$ spectral line \\
+    $\sigma_j$ & Gaussian line profile scale of the $j^{th}$ spectral line\\
+    $\gamma_j$ & Lorentzian line profile half width of the $j^{th}$ spectral line\\
+    $\mathsf{V}_j$ & The Voigt profile of the $j^{th}$ spectral line \\
+    $\bar{\bm{V}}$ & The dense $(N_{\rm lines} \times N_{x})$ matrix of all Voigt profiles stacked vertically \\
     $\hat{\bm{V}}$ & The sparse $(N_{\rm lines} \times N_{\rm sparse})$ matrix of all Voigt profiles stacked vertically \\
+    $\zeta$ & The rotational broadening convolution kernel\\
+    $g$ & The instrumental broadening convolution kernel, typically a Gaussian\\
     \hline
@@
     $N_{\rm lines}$ & Number of spectral lines \\
-    $N_{\rm pix}$ & Number of pixel coordinates in the precomputed spectrum \\
+    $N_{x}$ & Number of pixel coordinates in the precomputed spectrum $\bm{\lambda}_x$\\
     $N_{\rm sparse}$ & Number of non-zero pixels computed in the sparse implementation\\
@@
     $P_{\rm rom}$ & The prominence threshold of spectral lines to include in cloning \\
-    $v\sin{i}$ & Stellar projected rotational broadening at inclination $i$ and equatorial velocity $v$\\
+    $v\sin{i}$ & Rotational broadening for stellar inclination $i$ and equatorial velocity $v$\\
     $RV$ & Radial velocity of the star\\
+    $R$ & Spectrograph resolving power $\lambda/\delta\lambda$\\
+    $\mathcal{L}$ & The loss scalar, usually the sum of the squares of the residuals\\
+    \hline
+    \multicolumn{2}{c}{Operators}\\
+    \hline
+    $\resample \big[ \mathsf{F}_x \big]_z$ & The resample operator, takes in a flux spectrum $\mathsf{F}$ evaluated at $\bm{\lambda}_x$ coordinates and returns the mean flux within the pixel boundaries of coordinate $\bm{\lambda}_z$\\
+    $*$& The convolution operator\\
+    $\odot$& \emph{Hadamard product}, an elementwise product of two same-length vectors\\
     \enddata

### 2022-03-11 paper/paper1/ms.tex
Major surgery on the paper, closes #20, for now

@@
 
-We start with a high resolution pre-computed synthetic stellar model spectrum, $\mathsf{S}(\bm{\lambda})$.
-The procedure is largely agnostic to the exact details of how this spectrum was made.
-For the purposes of this paper, we will showcase examples from two well-known families of precomputed synthetic spectra: \texttt{PHOENIX} \citep{husser13} for stellar spectra $(T_{\mathrm{eff}}\in [2300, 10000]\;K)$ and \texttt{Sonora} \citep{2021ApJ...920...85M} for brown dwarf spectra $(T_{\mathrm{eff}}\in [1000, 2300]\;K)$.
+We start with a high resolution pre-computed synthetic stellar model spectrum, $\mathsf{S}_{\rm abs}(\bm{\lambda})$ at its native resolution sampling and with its original absolute physical flux units.
+The procedure that follows is largely agnostic to the exact details of how this spectrum was made, or what physics or chemistry it may represent.
+For the purposes of this paper, we will showcase examples from two well-known families of precomputed synthetic astronomical spectra: \texttt{PHOENIX} \citep{husser13} for stellar spectra $(T_{\mathrm{eff}}\in [2300, 10000]\;K)$ and \texttt{Sonora} \citep{2021ApJ...920...85M} for brown dwarf spectra $(T_{\mathrm{eff}}\in [1000, 2300]\;K)$.
+The framework may also work for precomputed synthetic spectra of exoplanets, supernovae, galaxies, or even further afield such as laboratory physical chemistry.
 We place the following demands on the precomputed spectra.
@@
 In principle, there is no fundamental limit to the bandwidth one could clone with the method presented here, up to and including the entire precomputed synthetic spectral model bandwidth.
-We adopt the exact native wavelength sampling with no smoothing or interpolation, yielding a wavelength vector $\bm{\lambda}_s$ with length $N_\mathrm{s}$ equal to the number of pixels within the extents of our region of interest including the buffers.
+We adopt the exact native wavelength sampling with no smoothing or interpolation, yielding a wavelength vector $\bm{\lambda}_S$ with length $N_\mathrm{S}$ equal to the number of pixels within the extents of our region of interest including the buffers.
 
@@
 We initialize the cloned model with a series of preprocessing steps.
-We divide the entire spectrum by a black body of the same effective temperature $T_{\mathrm{eff}}$ as the model template.
+We divide the entire spectrum by a black body $\mathsf{B}(\bm{\lambda}_S)$ of the same effective temperature $T_{\mathrm{eff}}$ as the model template.
 The resulting signal typically possesses a smooth continuum variation.
@@
 Here, we identify the continuum regions by the most prominent peaks in the specturm separated by a distance of at least 5,000$-$90,000 pixels from adjacent peaks, yielding between 60 and 5 control points.
-We then fit an $n_{\mathrm{poly}} = 3^{rd}-15^{th}$ order polynomial to these peaks, evaluate the polynomial on the entire bandwidth $\mathsf{P}(\bm{\lambda}_s)$, and divide the spectrum by this trend.
+We then fit an $n_{\mathrm{poly}} = 3^{rd}-15^{th}$ order polynomial to these peaks, evaluate the polynomial on the entire bandwidth $\mathsf{P}(\bm{\lambda}_S)$, and divide the spectrum by this trend.
+The result should be a flattened ``continuum-normalized'' spectrum familiar to practitioners in high resolution spectroscopy, with the continuum level close to unity.
+It is this spectrum that will serve as the centerpiece of subsequent training steps.
+We therefore drop any subscript and simply refer to this flattened spectrum as $\mathsf{S}$:
+
+\begin{eqnarray}
+    %\mathsf{S} = \frac{\mathsf{S}_{\rm abs}}{\mathsf{B} \; \mathsf{P}}
+    \mathsf{S} = \mathsf{S}_{\rm abs}/\mathsf{B} / \mathsf{P}
+\end{eqnarray}
+
+where the division indicates element-wise division of these arrays or ``vectors'' of flux values.
 Visual inspection of this continuum-flattened spectrum typically possesses variations at the $<1\%$ level.
@@
 
-We emphasize that a recreation of the unvarnished input spectrum can be obtained by multiplying the continuum-flattened signal by the ``perturbed black body'', $\mathsf{B}(\bm{\lambda}_s)\cdot \mathsf{P}(\bm{\lambda}_s)$, that symbolizes the black body modulated by continuum opacity or broad-band radiative transfer effects.
+We emphasize that a recreation of the unvarnished input spectrum can be obtained by multiplying the continuum-flattened signal by the ``perturbed black body'', $\mathsf{B}(\bm{\lambda}_S)\odot \mathsf{P}(\bm{\lambda}_S)$, that symbolizes the black body modulated by continuum opacity or broad-band radiative transfer effects.
 This smooth spectrum may be useful for applications that need to keep track of broad-band flux, such as low-resolution spectra, or regions with molecular band heads.
@@
 Next, we identify the spectral lines.
-We apply a local-minimum-finding algorithm on the spectrum by defining a prominence threshold $P_{rom} \in (0.005, 0.02)$.
+We apply a local-minimum-finding algorithm on the spectrum by defining a prominence threshold $P_{\rm rom} \in (0.005, 0.02)$.
 This threshold dictates the number of lines that will be modeled: a lower prominence finds more, weaker lines, and a larger prominence finds fewer, deeper lines.
@@
 IGRINS has two cameras for $H$ and $K$ band, with the combined spectrum spanning $14267-25217\;$\AA~ including the edge buffers and the region in-between the two cameras, all at a resolving power of $(R=45,000)$.
-The spectrograph acquisition, reduction, and post-processing steps yield data $\mathsf{D}(\bm{\lambda}_{d})$, where $\bm{\lambda}_{d}$ is the wavelength vector at the instrumental resolution and sampling of each instrument, generally much coarser than the resolution and sampling grid of the precomputed synthetic spectra.
+The spectrograph acquisition, reduction, and post-processing steps yield data $\mathsf{D}(\bm{\lambda}_{D})$, where $\bm{\lambda}_{D}$ is the wavelength vector at the instrumental resolution and sampling of each instrument, generally much coarser than the resolution and sampling grid of the precomputed synthetic spectra.
 The data wavelength vector may also contain gaps between \'echelle orders, whereas the precomputed wavelength coordinates are usually contiguous.
@@
 
-Figure \ref{fig_Nlines_vs_teff} shows how the number of detected lines $N_{\mathrm{lines}}$ scales with effective temperature and prominence threshold $P_{rom}$ for the \texttt{PHOENIX} grid, truncated to the bandwidths-plus-buffers for HPF and IGRINS.
-We see between about 2,000 and 20,000 lines depending on the $T_{\mathrm{eff}}$ and $P_{rom}$.
+Figure \ref{fig_Nlines_vs_teff} shows how the number of detected lines $N_{\mathrm{lines}}$ scales with effective temperature and prominence threshold $P_{\rm rom}$ for the \texttt{PHOENIX} grid, truncated to the bandwidths-plus-buffers for HPF and IGRINS.
+We see between about 2,000 and 20,000 lines depending on the $T_{\mathrm{eff}}$ and $P_{\rm rom}$.
 HPF and IGRINS have a comparable number of lines, and halving the prominence increases the number of lines by about $20-30\%$ in these ranges.
@@
 
-We have now arrived at the \emph{blas\'e} clone model $\mathsf{M}(\bf{\lambda}_s)$ for a flattened synthetic spectrum $\mathsf{S}$: it is the cumulative product of transmission through the sea of all overlapping spectral lines:
+We have now arrived at the \texttt{blas\'e} clone model $\mathsf{S}_{\rm clone}(\bf{\lambda}_s)$ for a flattened synthetic spectrum $\mathsf{S}$: it is the cumulative product of transmission through the sea of all overlapping spectral lines:
 
 \begin{eqnarray}
-    \mathsf{M}(\bm{\lambda}_s) = {\displaystyle \mathsf{P} \prod_{j=1}^{N_{\mathrm{lines}}} (1-a_j \mathsf{V}(\bm{\lambda}_s-\lambda_{c,j}, \sigma_j, \gamma_j) )} \label{equation1}
+    \mathsf{S}_{\rm clone} = {\displaystyle \prod_{j=1}^{N_{\mathrm{lines}}} 1-a_j \mathsf{V}_j } \label{equation1}
 \end{eqnarray}
 
-where $\mathsf{V}(\lambda, \sigma_j, \gamma_j)$ is the normalized Voigt profile with Gaussian standard deviation $\sigma$, Lorentzian half-width $\gamma$, at line center position $\lambda_c$, for the $j^{th}$ spectral line.
-The amplitude $a$ is always expected to be positive for absorption lines and is equal to the area under the curve of the line since the Voigt profile is normalized.
-The optional Prefactor term $\mathsf{P}$ is fixed at $1$ by default but can be allowed to vary to fine-adjust imperfections leftover from the polynomial-based continuum flattening procedure.
+where $\mathsf{V}_j$ is the Voigt profile $\mathsf{V}(\bm{\lambda}_S-\lambda_{\mathrm{c},j}, \sigma_j, \gamma_j)$ with Gaussian standard deviation $\sigma$, Lorentzian half-width $\gamma$, at line center position $\lambda_c$, for the $j^{th}$ spectral line.
+The amplitude $a$ is always expected to be positive for absorption lines.
 
-\subsection{Voigt profile evaluation}
 The Voigt profile $\mathsf{V}(\lambda, \sigma_j, \gamma_j)$ can be computed in exact closed-form using the Voigt-Hjerting function \citep{1938ApJ....88..508H} as the real part of the complex Fadeeva function \citep[\emph{e.g.}][]{2011arXiv1106.0151Z}.
@@
 Here we adopt the pseudo-Voigt approximation \citep{Ida:nt0146}.
-
-Two sampled Voigt approaches could be feasible.
-Direct convolution of a sampled Gaussian and Lorentzian profile may be too costly and inexact if pixels are coarsely sampled.
-Analogously, the convolution can be achieved in the Fourier domain through the multiplication of the Fast Fourier Transforms of the Gaussian and Lorentzian profiles, followed by the subsequent Inverse Fourier Transforms.
-The main rationale for either of these two choices would be to anticipate and align with planned convolutions such as rotational broadening and instrumental kernels.
-We do not currently implement these sampled convolutions in \emph{blas\'e} due to their prohibitive computational cost.
 
@@
 This model has $N_{\mathrm{lines}}\times 3$ free parameters, where the center wavelength is held fixed and the amplitude, width, and lineshape are allowed to vary.
-We minimize a scalar ``goodness-of-fit'' metric, \emph{aka} loss scalar $\mathcal{L}$, chosen as the mean squared error (MSE), which is proportional to $\chi^2$, the sum of the squares of the residual vector $\mathsf{R} \equiv \mathsf{S}-\mathsf{M}$ but has no notion of per-pixel noise since the precomputed synthetic spectrum has no uncertainty:
+We minimize a scalar ``goodness-of-fit'' metric, \emph{aka} loss scalar $\mathcal{L}$, chosen as the mean squared error (MSE), which is proportional to $\chi^2$, the sum of the squares of the residual vector $\mathsf{R} \equiv \mathsf{S}-\mathsf{S}_{\rm clone}$ but has no notion of per-pixel noise since the precomputed synthetic spectrum has no uncertainty:
 
 \begin{eqnarray}
-    \mathcal{L} = \sum_i^{N_s} (S_i - M_i)^2 = \mathsf{R^\intercal}\cdot \mathsf{R}
+    \mathcal{L} = \sum_i^{N_S} (S_i - S_{\mathrm{clone},i})^2 = \mathsf{R^\intercal}\cdot \mathsf{R}
 \end{eqnarray}
@@
 As seen in Figure \ref{fig_Nlines_vs_teff}, the number of lines can exceed 7,000, meaning the model has over 7,000 $\times 3 =$ 21,000 free parameters.
-Fitting that large number of parameters is difficult with conventional optimizers, which struggle to converge after hundreds of parameters\authorcomment1{Citation to limits of optimizers}.
-
-Here we employ a variant of Stochastic Gradient Descent (SGD), an optimization technique that can scale to a virtually unlimited number of parameters\authorcomment1{provide a citation}.
-This technique computes the derivative of the loss scalar with respect to each of the parameters, the Jacobian: $(\frac{\partial \mathcal{L}}{\partial a_j}, \frac{\partial \mathcal{L}}{\partial \sigma_j}, \frac{\partial \mathcal{L}}{\partial \gamma_j})$.
+Fitting that large number of parameters is difficult with conventional optimizers.
+Here we employ a variant of Stochastic Gradient Descent (SGD), an optimization technique that can scale to a virtually unlimited number of parameters \citep{2016arXiv160904747R}.
+This technique computes the derivative of the loss scalar with respect to each of the parameters, the so-called \emph{Jacobian}: $(\frac{\partial \mathcal{L}}{\partial a_j}, \frac{\partial \mathcal{L}}{\partial \sigma_j}, \frac{\partial \mathcal{L}}{\partial \gamma_j})$.
 The Jacobian indicates how the MSE would decrease with a change in the parameter-of-interest, or put simply which-way and by-how-much you have to change the line properties to get a better fit.
@@
 We choose the \texttt{PyTorch} framework that computes these Jacobians efficiently for all of the mathematical primitives in our \texttt{blas\'e} implementation.
-
-In principle, we could compute the Jacobians analytically---since the partial derivatives of the Voigt function are known in closed-form---circumventing the vendor lock-in and memory overhead of an autodiff implementation with PyTorch.
-Further, an autodiff-aware implementation of the \emph{celerit\'e} algorithm \citep{2017AJ....154..220F} does not yet exist in PyTorch, constricting our choices for simultaneous continuum fitting.
-In practice, PyTorch also offers hardware acceleration on GPU or TPU \citep{10.1145/3079856.3080246}, mature support for a range of optimizers, and other perks that motivated its choice over analytic implementations or other machine learning frameworks.
-The GPyTorch framework \citep{2018arXiv180911165G} may offer a workaround to the absence of a \emph{celerit\'e} implementation.
-Still, the performance of an analytic Jacobian implementation over the autodiff route presented here offers an interesting avenue of future research.
 
@@
 
+The autodiff machinery has a convenient way to set which parameters are held fixed and which are iteratively fine-tuned.
+One simply disables the Jacobian computations for the fixed parameters.
 We set the \texttt{requires\_grad=True} property for any Torch tensor that we want to vary.
-This allows us to easily explore whether, say, allowing the $\lambda_c$ parameter to vary significantly improves the fit.
-It also allows us to tune or freeze an experimental wavelength-dependent pre-factor to Equation \ref{equation1} to correct for lingering imperfections in the otherwise-fixed continuum flattening procedure.
+This allows us to easily explore whether, say, allowing the $\lambda_\mathrm{c}$ parameter to vary significantly improves the fit.
 
-The computational bottleneck occurs at the evaluation of Equation \ref{equation1}, which includes the assembly of a $N_{\mathrm{lines}}\times N_{s}$ matrix $\bm{\bar{V}}$ assembled by stacking each Voigt profile $\mathsf{V}_j(\bm{\lambda}_s)$ on top of each other:
+The computational bottleneck occurs at the evaluation of Equation \ref{equation1}, which can be viewed as having a $N_{\mathrm{lines}}\times N_{S}$ matrix $\bm{\bar{V}}$ assembled by stacking each Voigt absoption profile $\mathsf{V}_j(\bm{\lambda}_s)$ on top of each other:
 
@@
     \begin{pmatrix}
-        \mathsf{V}_1(\bm{\lambda}_s)                    & \\
-        \mathsf{V}_2(\bm{\lambda}_s)                    & \\
+        1 -\mathsf{V}_1(\bm{\lambda}_s)                     & \\
+        1 - \mathsf{V}_2(\bm{\lambda}_s)                    & \\
         \vdots                                              & \\
-        \mathsf{V}_{N_{\mathrm{lines}}}(\bm{\lambda}_s) &
+        1 - \mathsf{V}_{N_{\mathrm{lines}}}(\bm{\lambda}_s) &
     \end{pmatrix}
@@
 
-Equation \ref{equation1} performs a matrix contraction, turning a $N_{\mathrm{lines}}\times N_{s}$ matrix into a $1\times N_{s}$ vector.
-The number of Floating Point Operations (FLOPS) scales with the number of entries in this matrix, with a scalar prefactor for the cost of evaluating a single line profile at a single wavelength point.
-Efficient GPU algorithms exist for tensor contractions such as this product along an axis of a matrix, allowing this computation to proceed quickly on modern machines\authorcomment1{add citation?}.
-In particular, the proprietary CUDA architecture for NVIDIA\textsuperscript{\tiny\textregistered} GPUs contains Tensor cores with specialized matrix math, although the algorithms place restrictions on floating precision depending on the particular hardware and software available.
+Equation \ref{equation1} performs a type of matrix contraction, turning a $N_{\mathrm{lines}}\times N_{S}$ matrix into a $1\times N_{S}$ vector.
+The number of Floating Point Operations (FLOPS) scales with the number of entries in this matrix.
+Efficient GPU algorithms exist for matrix manipulations such as this one, allowing this computation to proceed quickly on modern machines.
+In particular, the proprietary CUDA architecture for NVIDIA\textsuperscript{\tiny\textregistered} GPUs contains Tensor cores with specialized matrix math.
 The chief bottleneck occurs when the storage of the $\bm{\bar{V}}$ matrix exceeds the available RAM of a GPU or CPU: the computation will fail with an ``Out of Memory'' exception.
@@
 The memory bottleneck is even more pernicious than mere storage of the $\bm{\bar{V}}$ matrix since the CPU/GPU also has to store the computation graph for autodiff.
-The computation graph takes up more memory than the matrix itself, so it is generally not possible to evaluate Equation \ref{equation1} in its entirety in one-fell-swoop.
+It is generally not possible to evaluate Equation \ref{equation1} in its entirety in one-fell-swoop.
 A remedy is needed.
-
-\subsection{Minibatches}
-
-Minibatches offer a natural solution to the line-by-line computational bottleneck while also providing optimization performance improvements.
-Minibatches act as a form of regularization, the principal source of stochasticity in the Stochastic Gradient Descent algorithm, which tends to have better convergence than full-batch Gradient Descent \citep{2016arXiv160904747R}.
-The Machine Learning/Artificial Intelligence (ML/AI) community employs minibatches to handle similar situations, in which the dataset volume exceeds the available GPU RAM.
-Our configuration departs slightly from the typical ML/AI situation since their models are usually ``small'' relative to their datasets.
-Here our model is ``big'' relative to our dataset.
-
-Here we choose to assemble and evaluate only a portion of the matrix at a time, in minibatches.
-The choice to evaluate only a portion of lines at a time would mean the model is inaccurately evaluated at all wavelength pixels.
-Instead, we choose to evaluate all lines, but only on a subset $N_{\mathrm{batch}}$ of the total pixels $N_s$, so that the model can eventually converge to exact at those points.
-All lines are allowed to update at each glimpse of a minibatch, but many lines with cores far from minibatch pixels will provide only weak information about how the loss scalar changes for their parameters.
-The value of $N_{\mathrm{batch}}$ should be set just below the threshold at which the computation runs out of memory.
-At present, this threshold is determined experimentally.
-The indices of the wavelength points can either be set to be contiguous in blocks, or random with possibly large gaps between indices, or else with some other scheme tailored to the most informative wavelength regions.
-We experimented with several techniques and settled on random sampling with replacement, meaning that some pixels get revisited more often than others, with some minibatches tailored only to the line cores.
 
@@
 The first, and expected, source of large residuals is simply missing line opacity due to our finite prominence threshold.
-Lines with prominence less than $P_{rom}$ yield residual notches with strengths comparable to $P_{rom}$.
-Including smaller prominence lines by lowering $P_{rom}$ produces smaller residuals, at the tradeoff of computing more lines and yielding higher computational cost.
-But at some point, turning down $P_{rom}$ yields diminishing returns, as other imperfections provide a noise floor.
-We have experimentally determined this noise floor to occur near $P_{rom}=0.01$.
+Lines with prominence less than $P_{\rm rom}$ yield residual notches with strengths comparable to $P_{\rm rom}$.
+Including smaller prominence lines by lowering $P_{\rm rom}$ produces smaller residuals, at the tradeoff of computing more lines and yielding higher computational cost.
+But at some point, turning down $P_{\rm rom}$ yields diminishing returns, as other imperfections provide a noise floor.
+We have experimentally determined this noise floor to occur near $P_{\rm rom}=0.01$.
 
@@
 
+\subsection{Alternatives to the PseudoVoigt Profile}
+Two sampled Voigt approaches could be feasible.
+Direct convolution of a sampled Gaussian and Lorentzian profile may be too costly and inexact if pixels are coarsely sampled.
+Analogously, the convolution can be achieved in the Fourier domain through the multiplication of the Fast Fourier Transforms of the Gaussian and Lorentzian profiles, followed by the subsequent Inverse Fourier Transforms.
+The main rationale for either of these two choices would be to anticipate and align with planned convolutions such as rotational broadening and instrumental kernels.
+We do not currently implement these sampled convolutions in \emph{blas\'e} due to their prohibitive computational cost.
+
+\subsection{Prospects for closed-form computation of Jacobians}
+In principle, we could compute the Jacobians analytically---since the partial derivatives of the Voigt function are known in closed-form---circumventing the vendor lock-in and memory overhead of an autodiff implementation with PyTorch.
+Further, an autodiff-aware implementation of the \emph{celerit\'e} algorithm \citep{2017AJ....154..220F} does not yet exist in PyTorch, constricting our choices for simultaneous continuum fitting.
+In practice, PyTorch also offers hardware acceleration on GPU or TPU \citep{10.1145/3079856.3080246}, mature support for a range of optimizers, and other perks that motivated its choice over analytic implementations or other machine learning frameworks.
+The GPyTorch framework \citep{2018arXiv180911165G} may offer a workaround to the absence of a \emph{celerit\'e} implementation.
+Still, the performance of an analytic Jacobian implementation over the autodiff route presented here offers an interesting avenue of future research.
+
+\subsection{Tuning the continuum fit}
+We could also tune or freeze an experimental wavelength-dependent pre-factor to Equation \ref{equation1} to correct for lingering imperfections in the otherwise-fixed continuum flattening procedure.
+
+\subsection{What about minibatches?}
+
+Minibatches offer a natural solution to the line-by-line computational bottleneck while also providing optimization performance improvements.
+Minibatches act as a form of regularization, the principal source of stochasticity in the Stochastic Gradient Descent algorithm, which tends to have better convergence than full-batch Gradient Descent \citep{2016arXiv160904747R}.
+The Machine Learning/Artificial Intelligence (ML/AI) community employs minibatches to handle similar situations, in which the dataset volume exceeds the available GPU RAM.
+Our configuration departs slightly from the typical ML/AI situation since their models are usually ``small'' relative to their datasets.
+Here our model is ``big'' relative to our dataset.
+
+Here we choose to assemble and evaluate only a portion of the matrix at a time, in minibatches.
+The choice to evaluate only a portion of lines at a time would mean the model is inaccurately evaluated at all wavelength pixels.
+Instead, we choose to evaluate all lines, but only on a subset $N_{\mathrm{batch}}$ of the total pixels $N_s$, so that the model can eventually converge to exact at those points.
+All lines are allowed to update at each glimpse of a minibatch, but many lines with cores far from minibatch pixels will provide only weak information about how the loss scalar changes for their parameters.
+The value of $N_{\mathrm{batch}}$ should be set just below the threshold at which the computation runs out of memory.
+At present, this threshold is determined experimentally.
+The indices of the wavelength points can either be set to be contiguous in blocks, or random with possibly large gaps between indices, or else with some other scheme tailored to the most informative wavelength regions.
+We experimented with several techniques and settled on random sampling with replacement, meaning that some pixels get revisited more often than others, with some minibatches tailored only to the line cores.
+
 \subsection{Lineshapes choice and tradeoffs}

### 2022-03-11 paper/paper1/ms.tex
More surgery and editting to the methods section, much more coherent

@@
 
-The computational bottleneck occurs at the evaluation of Equation \ref{equation1}, which can be viewed as having a $N_{\mathrm{lines}}\times N_{S}$ matrix $\bm{\bar{V}}$ assembled by stacking each Voigt absoption profile $\mathsf{V}_j(\bm{\lambda}_s)$ on top of each other:
+The computational bottleneck occurs at the evaluation of Equation \ref{equation1}, which can be viewed as having a $N_{\mathrm{lines}}\times N_{S}$ matrix $\bm{\bar{F}}$ assembled by stacking each Voigt absoption profile $\mathsf{V}_j(\bm{\lambda}_s)$ on top of each other:
 
@@
 
-Equation \ref{equation1} performs a type of matrix contraction, turning a $N_{\mathrm{lines}}\times N_{S}$ matrix into a $1\times N_{S}$ vector.
+An element of this matrix, $F_{j\lambda}$, will have the flux value for a given line at a given wavelength coordinate.
+Equation \ref{equation1} performs a type of matrix contraction, turning a $N_{\mathrm{lines}}\times N_{S}$ matrix into a length $N_{S}$ row vector.
 The number of Floating Point Operations (FLOPS) scales with the number of entries in this matrix.
+We can rewrite Equation \ref{equation1} as a sum by taking the log of both sides and dropping in this $\bm{\bar{F}}$ matrix:
+
+\begin{eqnarray}
+    \ln{\mathsf{S}_{\rm clone}} = \sum_{j=1}^{N_{lines}} \ln{F_{j\lambda}} = \mathbf{1} \cdot \ln{\bm{\bar{F}}}
+\end{eqnarray}
+
+where $\mathbf{1}$ is a $1\times N_{\rm lines}$ row vector.
 Efficient GPU algorithms exist for matrix manipulations such as this one, allowing this computation to proceed quickly on modern machines.
@@
 
-Another solution to the computation problem is to take advantage of the sparsity of the $\bm{\bar{V}}$ matrix: most of the Voigt profile's entries far from the line center are vanishingly close to zero.
-Ideally, we would only evaluate the model close to the line center position, some number of line cuts $N_{\mathrm{cut}}=\alpha \cdot \mathrm{FWHM}$ away.
-These line cuts produce a speedup by a factor of $\frac{N_s}{N_{\mathrm{cut}}}$, which can exceed $100\times$ for wide bandwidth spectra.
-Furthermore, efficient algorithms for assembling and coalescing sparse matrices exist in \texttt{PyTorch}.
-A few practical considerations make sparse matrices peskier to work with than their dense counterparts.
-First, the $\alpha$ hyperparameter should be set large enough that the truncation effect is not seen for the deepest lines, or in post-processing steps such as line broadening.
-The $\alpha$ should be preselected before the training process.
-If the FWHM initialization was poor enough, the question may arise how autodiff training will cope with the poor performance of a model, leading to the prospect of strange and difficult-to-diagnose artifacts near spurious line-cut truncation artifacts.
+The $\ln{\bm{\bar{F}}}$ matrix is sparsely populated: most of the entries far from the line center are vanishingly close to zero.
+Here we take advantage of that mostly empty matrix using the mathematics of sparse matrices \citep{saad03:IMS}.
 
-We implemented sparse matrices in \emph{blas\'e} using the \texttt{sparse\_coo\_tensor} function in \emph{PyTorch}.
-
-There are a few standardization strategies for applying wing cuts.
-Here we coerce all wing cuts to be the same number of pixels, typically 6000 pixels for \texttt{PHOENIX}, with the middle pixel being at the line center position, and about 3000 pixels to the red and blue side of the line.
-Most spectral lines only appreciably affect the signal in the central $\sim5-50$ pixels, so such a large wing cut may seem over-zealous.
-We emphasize that the line cut serves as the fixed pixel locations where the line will be evaluated with post-processing steps such as radial velocity shifts and line broadening.
-So these 6000 pixels must be capable of handling the fastest-moving and fastest-spinning stars you expect to see in your observed spectra of interest.
+We retain a relatively small number of pixels $N_{\rm cut}$ adjacent to the line center.
+Setting this wing cut produces a speedup by a factor of $\frac{N_S}{N_{\mathrm{cut}}}$, which can exceed $100\times$ for wide bandwidth spectra.The choice of $N_{\rm cut}$ is nuanced.
+It should be set large enough that truncation effects are not seen for the broadest lines.
+But even more, $N_{\rm cut}$ has to be future-proofed for Doppler-shifting.
+Extreme Doppler shifts could hypothetically send line cores entirely outside the extents of $N_{\rm cut}$.
+We therefore typically set wingcuts comparable to the buffer size $\pm \Delta \lambda_{\mathrm{buffer}}$, even though most weak lines only perceptibly affect $<1\;$\AA.
+We coerce all wing cuts to be the same number of pixels, typically 6000 pixels for \texttt{PHOENIX}, with the middle pixel being at the line center position, and about 3000 pixels to the red and blue side of the line.
+We refer to this approximate sparse matrix as $\ln{\bm{\hat{F}}}$.
 
@@
 Instead, the standardization choice of exactly 6000 pixels means we can use standard matrix or tensor arithmetic, which somewhat counterintuitively is faster while handling more pixels for all spectral lines: 6000 pixels everywhere is faster than 6000 pixels for some and 2999 for other lines.
-
 Locating the indices of these 6000 ``nearest-neighbor'' pixels is very fast, only occurs once, and then is fixed.
 
-We rewrite Equation \ref{equation1} as a sum by taking the log of both sides
-
-\begin{eqnarray}
-    \ln{\mathsf{F}(\bm{\lambda})} = \sum_{i=1}^{N_{lines}} \ln{\left(1-V(\bm{\lambda};A_i, \lambda_{c,i}, \sigma_i) \right)}
-\end{eqnarray}
-
-The argument of the natural logarithm approaches 1 far from the line core, making the terms outside of the matrix approximately zero, a sparse matrix representing the log of the flux.
 Sparse matrix methods generally support an operation known as \emph{coalescing}, which sums values with repeated indices.
-The coalescing operation is extremely fast and efficient on either CPU or GPU hardware.
 We sum the logs and then exponentiate the result before comparing it to the native linear flux.
@@
 Each pixel may get computed about $\sim100$ times in this sparse implementation, which is about $50\times$ better than each pixel getting computed $N_{lines}\sim6500$ times in the dense approach.
-
-\subsection{Broad lines and advanced lineshapes}
-
-Some lines, such as Hydrogen, Sodium, potassium, neutral metal lines, and others have extremely broad line wings, approaching larger than the $\sim6000$ pixels we allocate for the sparse implementation.
-These special lines need to be handled separately from the weak lines, both from a computational performance perspective and an accuracy perspective.
-
-Extremely broad lines will exhibit truncation effects if the sparse window is small compared to the line wing size.
-The truncation effects will look like tophat functions severing the asymptotic wings, imbuing artificial step function kinks in the emulated spectrum.
-We can afford to increase the sparse window on a few, say $N_{broad}\sim20$ of the broadest spectral lines.
-We then construct and evaluate the entire dense matrix for those lines: $\sim 330\;000 \times 20$.
-The number of FLOPS in each category scales as about 6 Million for the 20 broad lines versus about 36 Million for the sea of about 6500 narrow lines, depending on the exact choices for wing cuts and number of lines.
-
-We introduce advanced lineshapes for these $\sim20$ broad lines, perturbing the Voigt line-wings with a smooth wavelength-dependent correction term $\mathsf{G}$:
-
-\begin{eqnarray}
-    \mathsf{\tilde{V}(\bm{\lambda})} &=& \mathsf{V} \odot \mathsf{G}\\
-    \mathsf{G} &=& 1 + (e^{a_j} - 1) \cdot \mathcal{S}\left(\frac{|\bm{\lambda}-\lambda_{c,j}| - \lambda_{t, j}}{b_j}\right)
-\end{eqnarray}
-
-where $\odot$ is the element-wise product (\emph{a.k.a} Hadamard product), $\mathcal{S}$ is the sigmoid function, and we have introduced three new tunable parameters for each of the $j$ broad lines; $\lambda_t$ is the truncation wavelength, $b$ is a scale parameter for how slowly or how rapidly in wavelength-space the transition from non-Lorentzian proceeds, and $a$ is a possibly negative stretch parameter that controls whether the line wing is sub- or super- Lorentzian.
-
-This functional form has a few advantages.
-It is smooth.
-The smoothness of the transition is controlled by a tunable parameter, $b$.
-It can handle either sub- or super-Lorentzian shapes.
-It is roughly based on the $\chi$-factor\authorcomment1{citation to Hartmann et al.
-1989, etc.}.
-In the limit $\lim_{a\to0} \mathsf{G}$, the lineshape becomes exactly Lorentzian.
-The sigmoid is efficiently implemented in PyTorch.
-It enforces that the perturbation only produces absorption and not emission profiles.
+Efficient algorithms for assembling and coalescing sparse matrices exist in \texttt{PyTorch}
 
@@
 
+\subsection{Broad lines and advanced lineshapes}
+
+Some lines, such as Hydrogen, Sodium, potassium, neutral metal lines, and others have extremely broad line wings, approaching larger than the $\sim6000$ pixels we allocate for the sparse implementation.
+These special lines should be handled separately from the weak lines, both from a computational performance perspective and an accuracy perspective.
+
+Extremely broad lines will exhibit truncation effects if the sparse window is small compared to the line wing size.
+The truncation effects will look like tophat functions severing the asymptotic wings, imbuing artificial step function kinks in the emulated spectrum.
+We can afford to increase the sparse window on a few, say $N_{broad}\sim20$ of the broadest spectral lines.
+We then construct and evaluate the entire dense matrix for those lines: $\sim 330\;000 \times 20$.
+The number of FLOPS in each category scales as about 6 Million for the 20 broad lines versus about 36 Million for the sea of about 6500 narrow lines, depending on the exact choices for wing cuts and number of lines.
+
+We introduce advanced lineshapes for these $\sim20$ broad lines, perturbing the Voigt line-wings with a smooth wavelength-dependent correction term $\mathsf{G}$:
+
+\begin{eqnarray}
+    \mathsf{\tilde{V}(\bm{\lambda})} &=& \mathsf{V} \odot \mathsf{G}\\
+    \mathsf{G} &=& 1 + (e^{a_j} - 1) \cdot \mathcal{S}\left(\frac{|\bm{\lambda}-\lambda_{c,j}| - \lambda_{t, j}}{b_j}\right)
+\end{eqnarray}
+
+where $\odot$ is the element-wise product (\emph{a.k.a} Hadamard product), $\mathcal{S}$ is the sigmoid function, and we have introduced three new tunable parameters for each of the $j$ broad lines; $\lambda_t$ is the truncation wavelength, $b$ is a scale parameter for how slowly or how rapidly in wavelength-space the transition from non-Lorentzian proceeds, and $a$ is a possibly negative stretch parameter that controls whether the line wing is sub- or super- Lorentzian.
+
+This functional form has a few advantages.
+It is smooth.
+The smoothness of the transition is controlled by a tunable parameter, $b$.
+It can handle either sub- or super-Lorentzian shapes.
+It is roughly based on the $\chi$-factor\authorcomment1{citation to Hartmann et al.
+1989, etc.}.
+In the limit $\lim_{a\to0} \mathsf{G}$, the lineshape becomes exactly Lorentzian.
+The sigmoid is efficiently implemented in PyTorch.
+It enforces that the perturbation only produces absorption and not emission profiles.
+
 \subsection{Lineshapes choice and tradeoffs}

### 2022-03-11 paper/paper1/ms.tex
Finished writing for the week!

@@
 We use the Adam optimizer \citep{2014arXiv1412.6980K} with a typical learning rate $LR\in (0.005, 0.05)$ and all the defaults for \texttt{PyTorch v1.11}.
-Minibatch testing proceeded with batch fractions of typically $f\sim1-5\%$ of the available pixels but was rapidly discontinued in favor of the much more performant and convenient sparse matrix method.
 We defined the number of training epochs $N_{epoch}=100-10,000$ depending on the application.
@@
 
-We introduce advanced lineshapes for these $\sim20$ broad lines, perturbing the Voigt line-wings with a smooth wavelength-dependent correction term $\mathsf{G}$:
+One could introduce advanced lineshapes for these $\sim20$ broad lines, perturbing the Voigt line-wings with a smooth wavelength-dependent correction term $\mathsf{G}$:
 

### 2022-03-14 paper/paper1/ms.tex
Make a Matrix equation out of Voigt Profile figures, start telluric section

@@
 
-\section{Methodology}\label{methodology}
+\section{Methodology I: Cloning stellar spectra}\label{methodology}
 
@@
 
-\subsection{The \emph{blas\'e} clone model}
+\subsection{The \texttt{blas\'e} Stellar Clone Model}
 
@@
     \begin{pmatrix}
-        1 -\mathsf{V}_1(\bm{\lambda}_s)                     & \\
-        1 - \mathsf{V}_2(\bm{\lambda}_s)                    & \\
+        1 - a_1 \mathsf{V}_1(\bm{\lambda}_s)                     & \\
+        1 - a_2 \mathsf{V}_2(\bm{\lambda}_s)                    & \\
         \vdots                                              & \\
-        1 - \mathsf{V}_{N_{\mathrm{lines}}}(\bm{\lambda}_s) &
+        1 - a_{N_{\mathrm{lines}}} \mathsf{V}_{N_{\mathrm{lines}}}(\bm{\lambda}_s) &
     \end{pmatrix}
@@
 
-An element of this matrix, $F_{j\lambda}$, will have the flux value for a given line at a given wavelength coordinate.
+An element of this matrix, $F_{ji}$, will have the flux value for a given line at a given wavelength coordinate.
 Equation \ref{equation1} performs a type of matrix contraction, turning a $N_{\mathrm{lines}}\times N_{S}$ matrix into a length $N_{S}$ row vector.
@@
 \begin{eqnarray}
-    \ln{\mathsf{S}_{\rm clone}} = \sum_{j=1}^{N_{lines}} \ln{F_{j\lambda}} = \mathbf{1} \cdot \ln{\bm{\bar{F}}}
+    \ln{\mathsf{S}_{\rm clone}} = \sum_{j=1}^{N_{lines}} \ln{F_{ji}} = \mathbf{1} \cdot \ln{\bm{\bar{F}}}
 \end{eqnarray}
 
-where $\mathbf{1}$ is a $1\times N_{\rm lines}$ row vector.
-Efficient GPU algorithms exist for matrix manipulations such as this one, allowing this computation to proceed quickly on modern machines.
+where $\mathbf{1}$ is a $1\times N_{\rm lines}$ row vector of all-ones.
+We re\"emphasize that each spectral line has to be painstakingly evaluated across the entire spectral bandwidth.
+Efficient GPU algorithms exist for voluminous matrix manipulations such as this one, so this computation can proceed as quickly as possible on modern machines.
 In particular, the proprietary CUDA architecture for NVIDIA\textsuperscript{\tiny\textregistered} GPUs contains Tensor cores with specialized matrix math.
-The chief bottleneck occurs when the storage of the $\bm{\bar{V}}$ matrix exceeds the available RAM of a GPU or CPU: the computation will fail with an ``Out of Memory'' exception.
-Modern NVIDIA GPUs have $8-40$ GB of RAM.
-The memory bottleneck is even more pernicious than mere storage of the $\bm{\bar{V}}$ matrix since the CPU/GPU also has to store the computation graph for autodiff.
+The chief bottleneck occurs when the storage of the $\bm{\bar{F}}$ matrix exceeds the available RAM of a GPU or CPU: the computation will fail with an ``Out of Memory'' exception.
+Modern NVIDIA GPUs have $8-40$ GB of RAM, which translates roughly to a few thousand spectral lines across $\sim$300,000 pixels.
+The memory bottleneck is even more pernicious than mere storage of the $\bm{\bar{F}}$ matrix since the CPU/GPU also has to store the computation graph for autodiff.
 It is generally not possible to evaluate Equation \ref{equation1} in its entirety in one-fell-swoop.
@@
 But even more, $N_{\rm cut}$ has to be future-proofed for Doppler-shifting.
-Extreme Doppler shifts could hypothetically send line cores entirely outside the extents of $N_{\rm cut}$.
+Extreme Doppler shifts could hypothetically send line cores entirely outside the extents of $N_{\rm cut}$ if set too low.
 We therefore typically set wingcuts comparable to the buffer size $\pm \Delta \lambda_{\mathrm{buffer}}$, even though most weak lines only perceptibly affect $<1\;$\AA.
 We coerce all wing cuts to be the same number of pixels, typically 6000 pixels for \texttt{PHOENIX}, with the middle pixel being at the line center position, and about 3000 pixels to the red and blue side of the line.
-We refer to this approximate sparse matrix as $\ln{\bm{\hat{F}}}$.
-
-The PHOENIX spectra double their pixel sampling shortward of $\lambda = 1 \mu$m, so 6000 pixels in the visible is about 30 Angstroms, but 6000 pixels in the infrared is about 60 \AA.
+We populate a new approximate sparse matrix $\ln{\bm{\hat{F}}}$ with only these $~6000$ pixels per line and assume zeros everywhere else.
+%\footnote{The PHOENIX spectra double their pixel sampling shortward of $\lambda = 1 \mu$m, so 6000 pixels in the visible is about 30 Angstroms, but 6000 pixels in the infrared is about 60 \AA.
 We \emph{could have} demanded that the wing cut be exactly 30 \AA, or some other heuristic, but that choice would result in so-called \emph{jagged} or \emph{ragged} arrays: collections of 1D arrays that differ in their lengths.
@@
 Instead, the standardization choice of exactly 6000 pixels means we can use standard matrix or tensor arithmetic, which somewhat counterintuitively is faster while handling more pixels for all spectral lines: 6000 pixels everywhere is faster than 6000 pixels for some and 2999 for other lines.
-Locating the indices of these 6000 ``nearest-neighbor'' pixels is very fast, only occurs once, and then is fixed.
+Locating the indices of these 6000 ``nearest-neighbor'' pixels is very fast, only occurs once, and then is fixed.}
+
+The remapping of the sparse matrix can be visualized as the shifting of all lines to the center of this new matrix $\bm{\hat{F}}$.
+The algorithmic machinery keeps track of each $(i, j, F_{ji})$ trio of coordinates and flux values.
+
+\begin{equation} \bm{\bar{F}} = 
+    \begin{pmatrix}
+         \cdots \includegraphics[height=1cm]{figures/voigt_01.pdf}  \cdots & \\
+         \cdots \includegraphics[height=1cm]{figures/voigt_02.pdf}  \cdots & \\
+        \vdots                                              & \\
+        \cdots \includegraphics[height=1cm]{figures/voigt_03.pdf} \cdots &
+    \end{pmatrix} \mapsto 
+    \begin{pmatrix}
+        \includegraphics[height=1cm]{figures/voigt_01b.pdf}  & \\
+        \includegraphics[height=1cm]{figures/voigt_02b.pdf}  & \\
+       \vdots                                              & \\
+        \includegraphics[height=1cm]{figures/voigt_03b.pdf} &
+   \end{pmatrix} = \bm{\hat{F}}
+\end{equation}
 
@@
 We defined the number of training epochs $N_{epoch}=100-10,000$ depending on the application.
-The user can optionally monitor a live view of the training progress with Tensorboard\authorcomment1{citation} to gain an intuition for the training efficiency.
+The user can optionally monitor a live view of the training progress with Tensorboard \citep{tensorflow2015-whitepaper} to gain an intuition for the training efficiency.
 
@@
 
-Figure \ref{fig_cloned_spectrum_demo} shows a portion of a PHOENIX spectrum cloned with \emph{blas\'e}.
+Figure \ref{fig_cloned_spectrum_demo} shows a portion of a PHOENIX spectrum cloned with \texttt{blas\'e}.
 The $1000$ epochs of training took 56 seconds on an NVIDIA\textsuperscript{\tiny\textregistered} RTX2070 GPU with \texttt{PyTorch v1.11}, \texttt{CUDA v11.1}, and Intel\textsuperscript{\tiny\textregistered} Core\textsuperscript{\tiny TM} i7-9750H CPUs at 2.60GHz, with all tensors as FP64.
@@
 
-We store the model parameters to disk and refer to the entire collection of parameters as the \emph{pre-trained model}.
+We store the model parameters to disk and refer to the entire collection of parameters as a \emph{pre-trained model}.
+More specifically this fine-tuned model represents an evaluable and interpretable clone of the original static pixel values flux values.
+
+\section{Methodology II: Cloning Telluric Spectra}
+Ground-based near-IR \'echelle spectra possess countless depressions attributable to molecular line absorption in Earth's atmosphere.
+These telluric lines hamper the unbiased interpretation of \'echelle spectra, so some treatment plan is needed.
+Often the regions of known, deep tellurics are simply discarded.
+In other cases, the lines are modeled with first principles line-by-line radiative transfer \citep[\emph{e.g.} \texttt{TelFit,}][]{2014AJ....148...53G} or through data-driven means \citep[\emph{e.g.} \texttt{wobble,}][]{2019AJ....158..164B}.
+The most demanding EPRV applications require a precision characterization of telluric lines that the astronomical community has not yet been able to achieve, and that may rival even the abilities of Earth Science practitioners.
+A hybrid data-/model- driven approach was the chief recommendation of the \emph{Telluric Hack Week} workshop aimed at improving our mitigation of the atmosphere's deleterious effects.
+The \texttt{blas\'e} framework achieves a key milestone by introducing this \emph{hybrid} data- and model-driven approach.
+
+\subsection{Adjustments from the stellar model}
+The Telluric clone model proceeds almost identically to the stellar model, with the following modifications.
 
@@
 
-We compute the residual $\mathsf{R}(\bm{\lambda}_s)$ of native PHOENIX minus cloned model, illustrated in the bottom panel of Figure \ref{fig_cloned_spectrum_demo}.
+We compute the residual $\mathsf{R}(\bm{\lambda}_S)$ of native PHOENIX minus cloned model, illustrated in the bottom panel of Figure \ref{fig_cloned_spectrum_demo}.
 We see an RMS residuals of 1.2\%/pixel at native resolution.
-We see three main categories of cloning flaws.
+We identify three main categories of cloning flaws.
 
@@
 \begin{eqnarray}
-    \mathsf{F}_{ex}(\bm{\lambda}) = \mathsf{F}(\bm{v}-RV) * \zeta \left(\frac{\bm{v}}{v\sin{i}}\right) \label{eqn_convolution}
+    \mathsf{S}_{\rm ext}(\bm{\lambda}_S) = \mathsf{S}_{\rm clone}(\bm{v}-RV) * \zeta \left(\frac{\bm{v}}{v\sin{i}}\right) \label{eqn_convolution}
 \end{eqnarray}
@@
 
-Operationally, the Doppler shift gets applied to the line center positions $\tilde \lambda_{c,i} = (\lambda_{c,i}-\delta \lambda)$, and not the individual wavelength grid point coordinates, $\bm{\lambda}$.
-This choice yields a key benefit: it makes the $RV$ autodiff-aware, meaning that an infinitesimal change to the RV value can easily be sensed through backpropagation.
+Operationally, the Doppler shift gets applied to the line center positions $\tilde \lambda_{c,j} = (\lambda_{c,j}-\delta \lambda)$, and not the individual wavelength grid point coordinates, $\bm{\lambda}_S$.
+This choice yields a key benefit: it makes the $RV$ autodiff-aware, meaning that an infinitesimal change to the RV value can easily be sensed through backpropagation by affecting only the line center positions.
 The strategy allows relatively low computational and memory cost since it comports with the sparse matrix implementation.
@@
 The extent to which this effect matters will depend on the science case.
-For now, \emph{blas\'e} simply assumes a fixed resolving power.
+For now, \texttt{blas\'e} simply assumes a fixed resolving power.
 

### 2022-03-14 paper/paper1/ms.tex
Revise the telluric prose

@@
     \mathsf{S} = \mathsf{S}_{\rm abs}/\mathsf{B} / \mathsf{P}
+    \label{eqnFlattening}
 \end{eqnarray}
@@
 A hybrid data-/model- driven approach was the chief recommendation of the \emph{Telluric Hack Week} workshop aimed at improving our mitigation of the atmosphere's deleterious effects.
-The \texttt{blas\'e} framework achieves a key milestone by introducing this \emph{hybrid} data- and model-driven approach.
+The \texttt{blas\'e} framework achieves a key milestone by introducing such a \emph{hybrid} approach.
 
-\subsection{Adjustments from the stellar model}
-The Telluric clone model proceeds almost identically to the stellar model, with the following modifications.
+\subsection{The \texttt{blas\'e} Telluric Clone Model}
+We start with a precomputed synthetic telluric model, $\mathsf{T}$ with associated wavelength coordinates $\bm{\lambda}_T$.
+We employ a \texttt{TelFit} model, though any precomputed synthetic telluric model will work.
+The \texttt{TelFit} model does not contain any continuum sources of opacity, and so we can skip the flattening procedure described in Equation \ref{eqnFlattening}.
+We orchestrate the same initialization and line finding as in the stellar models, and obtain a coarse clone.
+
+The number of pixels in the telluric model can be chosen at the time of running a \texttt{TelFit} model.
+Here we choose a spectral resolution \censorbox{$R\sim1\times10^6$}, adequate for resolving telluric lines, and yielding \censorbox{xx} pixels across the HPF bandwidth.
+This pixel sampling is slightly \censorbox{finer} than the native PHOENIX pixel sampling.
+The number of telluric lines depends on the atmospheric properties, in particular the Earth surface temperature and relative humidity $RH$.
+For a \censorbox{T=x}, \censorbox{RH=y}, we detect \censorbox{zz} telluric lines across the entire HPF bandwidth.
+
+One guiding principle departs from the stellar case: telluric lines do not require future-proofing for large radial velocity shifts, hypothetically allowing us to reduce the number of pixels needed for a wingcut.
+Small radial velocity shifts are possible due to bulk motions in the Earth's atmosphere, but those bulk motions should be much smaller than the speeds of stars towards and away from Earth.
+So we can hypothetically tolerate a much smaller wing cut for telluric lines.
+In practice, telluric lines can be extremely broad, and accurately cloning these broad telluric lines requires about $N_{\rm cut}\sim6000$ pixels, comparable to the stellar case.
+
+We optimize the sparse telluric clone, achieving comparable computational speed as the PHOENIX cloning task.
+We are left with $\mathsf{T}_{\rm clone}(\bm{\lambda}_T)$, the tunable telluric clone model evaluated at its original native coordinates.
 
@@
 \end{figure}
+
+Cloning telluric lines suffers from one additional problem.
+Telluric lines can be extremely deep, exhibiting almost vanishing transmission with saturated line cores common in-between the atmospheric windows that define the $I$, $J$, $H$, and $K$- bands.
+The \texttt{blas\'e} method cannot cope with these saturated lines, and often it treats nearby blended saturated lines as one single line.
+The underfit model flounders.
+
+But overall, the telluric model cloning performs adequately, with \censorbox{X percentage} error-per-pixel overall, and only \censorbox{zz} error-per-pixel when restricted to the atmospheric windows.
+We therefore urge some caution when applying \texttt{blas\'e} to telluric modeling in the most saturated regions.
+\authorcomment1{An obvious and good solution is to intialize the model in log flux.
+We may revise the scheme to do so.}
 

### 2022-03-14 paper/paper1/ms.tex
writing more prose and editting equations.  looks good

@@
 
-\subsection{Augmenting the cloned model to match data}
+\subsection{Augmenting the Stellar Clone Model for $RV$ and $v\sin{i}$ }
 Real stars possess two key extrinsic properties.
@@
 The distinction is important because the extrinsic terms act as simple convolutions and translations to the cloned spectrum.
-We therefore define an augmented model, which we designate the ``extrinsic'' model, $\mathsf{F}_{ex}$:
+We therefore define an augmented model, which we designate the ``extrinsic'' model, $\mathsf{S}_{\rm ext}$:
 
 \begin{eqnarray}
-    \mathsf{S}_{\rm ext}(\bm{\lambda}_S) = \mathsf{S}_{\rm clone}(\bm{v}-RV) * \zeta \left(\frac{\bm{v}}{v\sin{i}}\right) \label{eqn_convolution}
+    \mathsf{S}_{\rm ext}(\bm{\lambda}_S) = \mathsf{S}_{\rm clone}(\lambda_\mathrm{c} - \frac{RV}{c}\lambda_\mathrm{c}) * \zeta \left(\frac{\bm{v}}{v\sin{i}}\right) \label{eqn_convolution}
 \end{eqnarray}
 
-where $\zeta$ is the convolution kernel for rigid body rotation \citep{2022ApJS..258...31K} and $\bm{v}$ is the spectral axis represented as relative velocity coordinates.
+where $\zeta$ is the convolution kernel for rigid body rotation \citep{2022ApJS..258...31K}, $\bm{v}$ is the spectral axis represented as relative velocity coordinates, and $*$ denotes the convolution operator.
 
-Operationally, the Doppler shift gets applied to the line center positions $\tilde \lambda_{c,j} = (\lambda_{c,j}-\delta \lambda)$, and not the individual wavelength grid point coordinates, $\bm{\lambda}_S$.
+Operationally, the Doppler shift gets applied to the line center positions and not the individual wavelength grid point coordinates, $\bm{\lambda}_S$.
 This choice yields a key benefit: it makes the $RV$ autodiff-aware, meaning that an infinitesimal change to the RV value can easily be sensed through backpropagation by affecting only the line center positions.
@@
 
+\subsection{Joint Stellar and Telluric Model}
+The telluric model and stellar models possess distinct pixel sampling and coordinates.
+The design of \texttt{blas\'e} permits the telluric model to be re\"evaluated at any wavelength coordinate vector, provided that it adequately samples the underlying lines.
+We choose to evaluate the telluric model at the native resolution of the stellar model.
+We then simply multiply the stellar model by the tunable telluric transmission:
+
+\begin{eqnarray}
+    \mathsf{M}_{\rm joint} = \mathsf{S}_{\rm ext}(\bm{\lambda}_S) \odot \mathsf{T}_{\rm clone}(\bm{\lambda}_S)
+\end{eqnarray}
+
+It is only at this stage that we may apply the instrumental broadening kernel.
 The instrumental resolving power, $R$, acts as a convolution with a Gaussian line profile of width $\sigma=\frac{c}{2.355 R}$.
@@
 
-The pixel resampling step could be achieved with two choices.
+The data-pixel sampling is much coarser than the model pixel sampling.
+This resampling step could be achieved with two choices.
 One would be to simply evaluate the closed-form model at the coordinates of the data pixels $\bm{\lambda_D}$.
@@
 The $RV$ only dictates what flux values are realized within those pixels.
-The final forward model for \emph{blas\'e} is designated the ``pixel-level'' model to emphasize that we have achieved a model for each datum in the 1D observed spectrum:
+The final forward model for \emph{blas\'e} is designated simply as $\mathsf{M}$ without subscripts to emphasize that we have achieved the desired goal of a plausible model for each datum in the 1D observed spectrum:
 
 \begin{eqnarray}
-    \mathsf{F}_{p}(\bm{\lambda}) = \resample{\Big[\mathsf{F}_{ex} * g(R) \Big]} \label{eqn_final_model}
+    \mathsf{M}(\bm{\lambda}_D) = \resample{\Big[\mathsf{M}_{\rm joint}(\bm{\lambda}_S) * g(R) \Big]} \label{eqn_final_model}
 \end{eqnarray}
 
-where $g$ is the Gaussian instrumental convolution kernel and the \texttt{resample[]} operation indicates the average of pixels within the data boundaries.
+where $g$ is the Gaussian instrumental convolution kernel and the \texttt{resample[]} operation indicates the average of model pixels that fall within each data pixel's red and blue boundaries.
 
@@
 
-Equation \ref{eqn_final_model} has all the same conceivably tunable parameters as the cloned model, \emph{plus} $v\sin{i}$ and $RV$.
-The resolving power may also be treated as tunable if it is not known or varies with \emph{e.g.} seeing or instrumental configuration.
+Equation \ref{eqn_final_model} has up to $N_{\mathrm{lines}}\times3 \sim20,000$ tunabale parameters from the star, $N_{\mathrm{lines}}\times3 \sim15,000$ tunabale parameters from the Earth's atmosphere, plus $v\sin{i}$ and $RV$.
+That adds up to 35,002 model parameters.
+The resolving power may also be treated as tunable if it is not known or varies slightly with \emph{e.g.} seeing, slit-or-fiber illumination, or instrumental configuration.
 
@@
 
-So instead, we have to apply some amount of \emph{regularization}, a restiction on the allowed values the model parameters can take on.
+We apply some amount of \emph{regularization}, a restiction on the allowed values the model parameters can take on.
 Fortunately we have a firm theoretical basis to justify this regularization.
@@
 Even still, the regularization step is in some ways the most subjective, since the degree of regularization will control the degree of overfitting or underfitting.
-The most extreme regularization---the antithesis of the \emph{laissez-faire} scenario---would yield a model too rigid to respond to the data at all, yielding a model entirely unchanged from the cloned PHOENIX model.
-So setting the regularization parameter constitutes the only hyperparameter worthy of tuning in \texttt{blas\'e}.
+The most extreme regularization---the antithesis of the \emph{laissez-faire} scenario---would yield a model too rigid to respond to the data at all, yielding a model entirely unchanged from the cloned PHOENIX and \texttt{TelFit} models.
+So regularization constitutes the only hyperparameters worthy of tuning in \texttt{blas\'e}.
 The choice of how to set the regularization is problem-specific.

### 2022-03-14 paper/paper1/ms.tex
Make the table match the revised equation notation conventions

@@
     $\mathsf{P}$ & Smooth polynomial to refine continuum-normalization\\
-    $\mathsf{S}$ & Continuum normalized augmentation of $\mathsf{S}_{\rm native}$\\
+    $\mathsf{S}$ & Continuum normalized augmentation of $\mathsf{S}_{\rm abs}$\\
     $\mathsf{T}$ & Transmission values of the precomputed synthetic telluric model \\
@@
     $\mathsf{T}_{\rm clone}$ & Evaluable and tunable cloned transmission model of $\mathsf{T}$\\
-    %$\mathsf{T}_{\rm resamp}$ & The cloned telluric model resampled from $\bm{\lambda}_T$ to $\bm{\lambda}_S$\\
     $\mathsf{S}_{\rm ext}$ & An augmentation of $\mathsf{S}_{\rm clone}$ with $v\sin{i}$ convolution and $RV$ translation\\
@@
     \hline
-    $\lambda_{c,j}$ & Line center position of the $j^{th}$ spectral line\\
+    $\lambda_{\mathrm{c},j}$ & Line center position of the $j^{th}$ spectral line\\
     $a_j$ & Gaussian line profile amplitude of the $j^{th}$ spectral line \\
@@
     $\mathsf{V}_j$ & The Voigt profile of the $j^{th}$ spectral line \\
-    $\bar{\bm{V}}$ & The dense $(N_{\rm lines} \times N_{x})$ matrix of all Voigt profiles stacked vertically \\
-    $\hat{\bm{V}}$ & The sparse $(N_{\rm lines} \times N_{\rm sparse})$ matrix of all Voigt profiles stacked vertically \\
+    $\bar{\bm{F}}$ & The dense $(N_{\rm lines} \times N_{x})$ matrix of all line fluxes stacked vertically \\
+    $\hat{\bm{F}}$ & The sparse $(N_{\rm lines} \times N_{\rm sparse})$ matrix of all line fluxes stacked vertically \\
     $\zeta$ & The rotational broadening convolution kernel\\
@@
     \hline
-    $\resample \big[ \mathsf{F}_x \big]_z$ & The resample operator, takes in a flux spectrum $\mathsf{F}$ evaluated at $\bm{\lambda}_x$ coordinates and returns the mean flux within the pixel boundaries of coordinate $\bm{\lambda}_z$\\
+    $\resample \big[ \mathsf{F(\bm{\lambda}_x)} \big]$ & The resample operator, takes in a flux spectrum $\mathsf{F}$ evaluated at $\bm{\lambda}_x$ coordinates and returns the mean flux within the pixel boundaries of coordinate $\bm{\lambda}_z$\\
     $*$& The convolution operator\\

### 2022-03-15 paper/paper1/ms.tex
Add appendix describing logarithmic alternative

@@
 Adopting the log of the flux would ensure that the cloned model possesses only positive flux values, a desirable trait of any physical spectral model.
-We chose instead to stick with native linear flux scaling because of the availability of convenient approximations to astrophysically relevant line profiles in the linear scaling domain.
-We normalize the spectrum by its median value since we do not intend to apply these methods to absolutely flux calibrated spectra.
+We have implemented both modes in \texttt{blas\'e}, allowing users to choose their preference.
+We only narrate the linear flux unit description in the main text of this document, for the sake of clarity, and since most practitioners tend to think of flux in terms of linear flux units.
+The data-model comparison step will always take place in linear flux units, so the only operational difference is the behavior for deep and saturated lines.
+Appendix \censorbox{XX} lists the equations adjusted to log flux units.
 
@@
 
-\section{Autodiff themes} \label{appendix:tools}
+\section{Log flux scaling mode} \label{appendixLogScale}
 
-Here are some more details about autodiff
+Here we illustrate how \texttt{blas'e} gets altered when applying the logarithmic flux pre-processing step.
+First, we compute the natural log of the flux directly on the precomputed synthetic spectrum in its absolute flux scaling and native pixel sampling:
+
+\begin{eqnarray}
+    \ln{\mathsf{S}} = \ln{\mathsf{S}_{\rm abs}} - \ln{\mathsf{B}} - \mathsf{P}
+    \label{eqnlogFlat}
+\end{eqnarray}
+
+We simply ``rebrand'' $\mathsf{P}$ as residing in logarithmic flux units, and disregard it since it is largely a nuisance parameter anyways.
+We then treat the \texttt{blas\'e} clone model as a sum of opacities, retaining the Voigt profile:
+
+\begin{eqnarray}
+    \ln{\mathsf{S}_{\rm clone}} = -\sum_{j=1}^{N_{\mathrm{lines}}} a_j \mathsf{V}_j \label{equationOpacitySum}
+\end{eqnarray}
+
+Here, the $a_j$'s have also been slightly rebranded from their meaning in Equation \ref{equation1}.
+We still want to enforce only absorption lines---and not spurious emission lines---so we use the sample trick of sampling the $a_j$'s in log and then exponentiating them to get guaranteed positive values.
+Note that Equations \ref{equation1} and \ref{equationOpacitySum} carry modified meanings for the Voigt profile.
+Specifically, Equation \ref{equation1} can be viewed as the Taylor Series expansion for \ref{equationOpacitySum} in the limit of small opacities:
+
+\begin{eqnarray}
+    e^{-a_j \mathsf{V}_j} \approx (1-a_j\mathsf{V}_j) \label{eqnTaylor}
+\end{eqnarray}
+
+Both equations are approximate.
+A real stellar atmosphere's lineshape arises from a sum of disparate Voigt profiles weighted along a nonuniform column of gas, whereas here we have assumed the column of gas is approximated as a single uniform isothermal backlit layer.
+A sum of unlike-Voigt profiles is not exactly equal to any single Voigt profile.
+Theoreticians may resonate with this more ``first principles'' representation, while data practitioners may find Equation \ref{equation1} more natural, so to some extent the choice is a matter of taste.
+
+The sparse matrix gets rebranded as filled with opacity values, instead of log-fluxes, but operationally remains the same.
+All subsequent steps operate on the summed-and-exponentiated opacities, behaving identically to their linear counterparts.
+For example, we exponentiate before computing the residuals and data-model comparison, $\mathsf{R} = e^{\ln{\mathsf{S}}} - e^{\ln{\mathsf{S}_{\rm clone}}}$.
 

### 2022-03-15 setup.py
Tagging version 0.2!!

@@ -3,7 +3,7 @@ import setuptools
 
 setuptools.setup(
     name="blase",
-    version="0.1",
+    version="0.2",
     author="gully",
     author_email="igully@gmail.com",
     description="Forward Modeling echelle spectra with PyTorch",

### 2022-03-15 src/blase/emulator.py
Start of a major overhaul of the code!

@@ -2,13 +2,8 @@ r"""
 Emulator
 --------------
 
-Precomputed synthetic spectral models are awesome but imperfect and rigid.  Here we clone the most prominent spectral lines and continuum appearance of synthetic spectral models to turn them into tunable, flexible, semi-empirical models.  We can ultimately learn the properties of the pre-computed models with a neural network training loop, and then transfer those weights to real data, where a second transfer-learning training step can take place. The spectrum has :math:`N_{pix} \sim 300,000` pixels and :math:`N_{lines} \sim 5000` spectral lines.  The number of lines is set by the `prominence=` kwarg: lower produces more lines and higher (up to about 0.3) produces fewer lines.  
-
-
-PhoenixEmulator
-###############
+Precomputed synthetic spectral models are awesome but imperfect and rigid.  Here we clone the most prominent spectral lines and continuum appearance of synthetic spectral models to turn them into tunable, flexible, semi-empirical models.  We can ultimately learn the properties of the pre-computed models with a neural network training loop, and then transfer those weights to real data, where a second transfer-learning training step can take place. The spectrum has :math:`N_{\rm pix} \sim 300,000` pixels and :math:`N_{\rm lines} \sim 5000` spectral lines.  The number of lines is set by the `prominence=` kwarg: lower produces more lines and higher (up to about 0.3) produces fewer lines.  
 """
-from importlib.util import resolve_name
 import math
 import torch
 from torch import nn
@@ -17,14 +12,14 @@ from scipy.signal import find_peaks, peak_prominences, peak_widths
 from tqdm import tqdm
 
 
-class PhoenixEmulator(nn.Module):
+class LinearEmulator(nn.Module):
     r"""
     A PyTorch layer that clones precomputed synthetic spectra
 
-    wl_native (int): The input wavelength
-    flux_native (float): The output wavelength
-
-    Currently hardcoded to assume your PHOENIX grid is stored at: ~/libraries/raw/PHOENIX/
+    Parameters
+    ----------
+    wl_native (float vector): The input wavelength at native resolution and sampling 
+    flux_native (float vector): The input flux
     """
 
     def __init__(self, wl_native, flux_native, prominence=0.03):
@@ -107,31 +102,13 @@ class PhoenixEmulator(nn.Module):
         )
 
     def forward(self, wl):
[…]

### 2022-03-15 docs/conf.py
Migrate to sphinx book theme because it colorizes the API much better--more readible

@@ -22,7 +22,7 @@ copyright = "2020, gully"
 author = "gully"
 
 # The full version, including alpha/beta/rc tags
-release = "0.1"
+release = "0.3"
 
 
 # -- General configuration ---------------------------------------------------
@@ -57,7 +57,7 @@ exclude_patterns = ["_build", "Thumbs.db", ".DS_Store"]
 # The theme to use for HTML and HTML Help pages.  See the documentation for
 # a list of builtin themes.
 #
-html_theme = "pyramid"
+html_theme = "sphinx_book_theme"
 
 # Add any paths that contain custom static files (such as style sheets) here,
 # relative to this directory. They are copied after the builtin static files,

### 2022-03-15 src/blase/emulator.py

@@ -14,7 +14,7 @@ from tqdm import tqdm
 
 class LinearEmulator(nn.Module):
     r"""
-    A PyTorch layer that clones precomputed synthetic spectra
+    A PyTorch layer that clones a precomputed synthetic spectrum :math:`\mathsf{S} \mapsto \mathsf{S}_{\rm clone}`
 
     Parameters
     ----------
@@ -123,12 +123,14 @@ class LinearEmulator(nn.Module):
     def detect_lines(self, wl_native, flux_native, prominence=0.03):
         """Identify the spectral lines in the native model
 
-        Args:
+        Parameters
+        ----------
             wl_native (torch.tensor vector): The 1D vector of native model wavelengths (Angstroms)
             flux_native (torch.tensor vector): The 1D vector of native model fluxes (Normalized)
-        Returns:
+
+        Returns
+        -------
             (tuple of tensors): The wavelength centers, prominences, and widths for all ID'ed spectral lines
-        -----
         """
         peaks, _ = find_peaks(-flux_native, distance=4, prominence=prominence)
         prominence_data = peak_prominences(-flux_native, peaks)
@@ -209,8 +211,7 @@ class SparsePhoenixEmulator(PhoenixEmulator):
     flux_native (float vector): The native flux
     prominence (int scalar): The threshold for detecting lines
     device (Torch Device or str): GPU or CPU?
-    wing_cut_pixels (int scalar): the number of pixels centered on the line center
-        to evaluate in the sparse implementation, default: 1000 pixels
+    wing_cut_pixels (int scalar): the number of pixels centered on the line center to evaluate in the sparse implementation, default: 1000 pixels
     """
 
     def __init__(

### 2022-03-15 src/blase/emulator.py
Improving the docstrings with real math

@@ -102,16 +102,34 @@ class LinearEmulator(nn.Module):
         )
 
     def forward(self, wl):
-        r"""The forward pass of the spectral model: 
-            :math:`\mathsf{S}_{\rm clone} = \prod_{j=1}^{N_{\mathrm{lines}}} 1-a_j \mathsf{V}_j`
+        r"""The forward pass of the `blase` clone model
 
-        Returns:
-            (torch.tensor): the 1D generative spectral model destined for backpropagation parameter tuning
+        Conducts the product of PseudoVoigt profiles for each line, with
+        between one and three tunable parameters:
+
+        .. math:: 
+            
+            \mathsf{S}_{\rm clone} = \prod_{j=1}^{N_{\mathrm{lines}}} 1-a_j \mathsf{V}_j(\lambda_S)
+
+        Parameters
+        ----------
+        wl : torch.tensor
+            The input wavelength :math:`\mathbf{\lambda}_S` at which to 
+            evaluate the model
+
+        Returns
+        -------
+        torch.tensor
+            The 1D generative spectral model clone :math:`\mathsf{S}_{\rm clone}` destined for backpropagation parameter tuning 
         """
         return self.product_of_pseudovoigt_model(wl)
 
     def product_of_pseudovoigt_model(self, wl):
-        """Return the PseudoVoight forward model"""
+        """Return the PseudoVoigt forward model
+        
+        
+        
+        """
         net_spectrum = (1 - self.pseudo_voigt_profiles(wl)).prod(0)
 
         wl_normed = (wl - 10_500.0) / 2500.0
@@ -125,12 +143,16 @@ class LinearEmulator(nn.Module):
 
         Parameters
         ----------
-            wl_native (torch.tensor vector): The 1D vector of native model wavelengths (Angstroms)
-            flux_native (torch.tensor vector): The 1D vector of native model fluxes (Normalized)
+        wl_native : torch.tensor
+            The 1D vector of native model wavelengths (Angstroms)
+        flux_native: torch.tensor
+            The 1D vector of continuum-flattened model fluxes
 
         Returns
         -------
-            (tuple of tensors): The wavelength centers, prominences, and widths for all ID'ed spectral lines
+        tuple of tensors 
+            The wavelength centers, prominences, and widths for all ID'ed 
+            spectral lines
         """
         peaks, _ = find_peaks(-flux_native, distance=4, prominence=prominence)
         prominence_data = peak_prominences(-flux_native, peaks)

### 2022-03-16 src/blase/emulator.py
Improve docstrings with equations and references

@@ -14,12 +14,18 @@ from tqdm import tqdm
 
 class LinearEmulator(nn.Module):
     r"""
-    A PyTorch layer that clones a precomputed synthetic spectrum :math:`\mathsf{S} \mapsto \mathsf{S}_{\rm clone}`
+    A PyTorch layer for cloneing a precomputed synthetic spectrum 
+    
+    :math:`\mathsf{S} \mapsto \mathsf{S}_{\rm clone}`
 
     Parameters
     ----------
-    wl_native (float vector): The input wavelength at native resolution and sampling 
-    flux_native (float vector): The input flux
+    wl_native :  torch.tensor
+        The vector of input wavelengths at native resolution and sampling 
+    flux_native : torch.tensor
+        The vector continuum-flattened input fluxes
+    prominence : int
+        The threshold prominence for peak finding
     """
 
     def __init__(self, wl_native, flux_native, prominence=0.03):
@@ -101,12 +107,42 @@ class LinearEmulator(nn.Module):
             torch.tensor(0.0, requires_grad=False, dtype=torch.float64)
         )
 
+        self.wl_normed = (self.wl_native - 10_500.0) / 2500.0
+
     def forward(self, wl):
         r"""The forward pass of the `blase` clone model
 
         Conducts the product of PseudoVoigt profiles for each line, with
-        between one and three tunable parameters:
+        between one and three tunable parameters.  The entire spectrum can
+        optionally be modulated by a tunable continuum polynomial.
 
+        .. math:: 
+            
+            \mathsf{P}(\lambda_S) \prod_{j=1}^{N_{\mathrm{lines}}} 1-a_j \mathsf{V}_j(\lambda_S)
+
+        Parameters
+        ----------
+        wl : torch.tensor
+            The input wavelength :math:`\mathbf{\lambda}_S` at which to 
+            evaluate the model
+
+        Returns
+        -------
+        torch.tensor
+            The 1D generative spectral model clone :math:`\mathsf{S}_{\rm clone}` destined for backpropagation parameter tuning 
+        """
+
+        polynomial_term = (
+            self.a_coeff
+            + self.b_coeff * self.wl_normed
+            + self.c_coeff * self.wl_normed ** 2
+        )
+
+        return self.product_of_pseudovoigt_model(wl) * polynomial_term
+
[…]

### 2022-03-16 src/blase/emulator.py
docstring standardizing and refactoring

@@ -9,12 +9,11 @@ import torch
 from torch import nn
 import numpy as np
 from scipy.signal import find_peaks, peak_prominences, peak_widths
-from tqdm import tqdm
 
 
 class LinearEmulator(nn.Module):
     r"""
-    A PyTorch layer for cloneing a precomputed synthetic spectrum 
+    Model for cloning a precomputed synthetic spectrum in linear flux.
     
     :math:`\mathsf{S} \mapsto \mathsf{S}_{\rm clone}`
 
@@ -141,11 +140,11 @@ class LinearEmulator(nn.Module):
         return self.product_of_pseudovoigt_model(wl) * polynomial_term
 
     def product_of_pseudovoigt_model(self, wl):
-        """Return the Product of pseudo-Voigt forward model
+        r"""Return the Product of pseudo-Voigt profiles
         
-        .. math:: 
-            
-            \mathsf{S}_{\rm clone} = \prod_{j=1}^{N_{\mathrm{lines}}} 1-a_j \mathsf{V}_j(\lambda_S)
+        The product acts like a matrix contraction:
+
+        .. math:: \mathsf{S}_{\rm clone} = \prod_{j=1}^{N_{\mathrm{lines}}} 1-a_j \mathsf{V}_j(\lambda_S)
 
         Parameters
         ----------
@@ -159,8 +158,7 @@ class LinearEmulator(nn.Module):
             The 1D generative spectral model clone :math:`\mathsf{S}_{\rm clone}` 
             destined for backpropagation parameter tuning     
         """
-        net_spectrum = (1 - self.pseudo_voigt_profiles(wl)).prod(0)
-        return net_spectrum
+        return (1 - self.pseudo_voigt_profiles(wl)).prod(0)
 
     def detect_lines(self, wl_native, flux_native, prominence=0.03):
         """Identify the spectral lines in the native model
@@ -196,7 +194,6 @@ class LinearEmulator(nn.Module):
 
     def _gaussian_line(self, lam_center, width, wavelengths):
         """Return a normalized Gaussian line, given properties"""
-
         return (
             1.0
             / (width * 2.5066)
@@ -204,14 +201,12 @@ class LinearEmulator(nn.Module):
         )
 
     def _compute_eta(self, fwhm_L, fwhm):
-        """Compute the eta parameter for pseudo Voigt"""
+        """Compute the eta mixture ratio for pseudo-Voigt weighting"""
         f_ratio = fwhm_L / fwhm
[…]

### 2022-03-16 src/blase/emulator.py
Docstrings overhaul--- noticing similar code between base class and sub class

@@ -117,7 +117,7 @@ class LinearEmulator(nn.Module):
 
         .. math:: 
             
-            \mathsf{P}(\lambda_S) \prod_{j=1}^{N_{\mathrm{lines}}} 1-a_j \mathsf{V}_j(\lambda_S)
+            \mathsf{S}_{\rm clone} = \mathsf{P}(\lambda_S) \prod_{j=1}^{N_{\mathrm{lines}}} 1-a_j \mathsf{V}_j(\lambda_S)
 
         Parameters
         ----------
@@ -144,7 +144,7 @@ class LinearEmulator(nn.Module):
         
         The product acts like a matrix contraction:
 
-        .. math:: \mathsf{S}_{\rm clone} = \prod_{j=1}^{N_{\mathrm{lines}}} 1-a_j \mathsf{V}_j(\lambda_S)
+        .. math:: \prod_{j=1}^{N_{\mathrm{lines}}} 1-a_j \mathsf{V}_j(\lambda_S)
 
         Parameters
         ----------
@@ -431,13 +431,19 @@ class SparseLinearEmulator(LinearEmulator):
         return torch.exp(result_1D)
 
 
-class EchelleModel(nn.Module):
+class ExtrinsicModel(nn.Module):
     r"""
     A Model for Echelle Spectra based on the SparseEmulator
 
-    wl_bin_edges (float vector): The input wavelength
-    device (Torch Device or str): GPU or CPU?
-    pretrained_emulator (SparsePhoenixEmulator): A pretrained emulator to use for modeling data
+
+    Parameters
+    ----------
+    wl_bin_edges : float vector
+        The edges of the wavelength bins
+    wl_native : float vector
+        The native wavelength coordinates
+    device : Torch Device or str
+        GPU or CPU?
     """
 
     def __init__(self, wl_bin_edges, wl_native, device=None):
@@ -507,10 +513,24 @@ class EchelleModel(nn.Module):
         )
 
     def forward(self, high_res_model):
-        """The forward pass of the data-based echelle model implementation--- no wavelengths needed!
+        r"""The forward pass of the data-based echelle model implementation
+        
+        Computes the RV and vsini modulations of the native model:
 
-        Returns:
-            (torch.tensor): the 1D generative spectral model destined for backpropagation parameter tuning
+        .. math::
+
[…]

### 2022-03-16 src/blase/emulator.py
Unit tests to make sure the emulators can forward model

@@ -130,11 +130,10 @@ class LinearEmulator(nn.Module):
         torch.tensor
             The 1D generative spectral model clone :math:`\mathsf{S}_{\rm clone}` destined for backpropagation parameter tuning 
         """
+        wl_normed = (wl - 10_500.0) / 2500.0
 
         polynomial_term = (
-            self.a_coeff
-            + self.b_coeff * self.wl_normed
-            + self.c_coeff * self.wl_normed ** 2
+            self.a_coeff + self.b_coeff * wl_normed + self.c_coeff * wl_normed ** 2
         )
 
         return self.product_of_pseudovoigt_model(wl) * polynomial_term
@@ -403,13 +402,13 @@ class SparseLinearEmulator(LinearEmulator):
 
         flux_2D = torch.exp(self.amplitudes).unsqueeze(1) * (
             eta
-            * self.lorentzian_line(
+            * self._lorentzian_line(
                 rv_shifted_centers.unsqueeze(1),
                 torch.exp(self.gamma_widths).unsqueeze(1),
                 self.wl_2D,
             )
             + (1 - eta)
-            * self.gaussian_line(
+            * self._gaussian_line(
                 rv_shifted_centers.unsqueeze(1),
                 torch.exp(self.sigma_widths).unsqueeze(1),
                 self.wl_2D,

### 2022-03-16 test/test_emulator.py

@@ -1,14 +1,41 @@
 import torch
-from blase.emulator import PhoenixEmulator
+from blase.emulator import LinearEmulator, SparseLinearEmulator
+from gollum.phoenix import PHOENIXSpectrum
+
+spectrum = PHOENIXSpectrum(teff=4700, logg=4.5)
+spectrum = spectrum.divide_by_blackbody()
+spectrum = spectrum.normalize()
+continuum_fit = spectrum.fit_continuum(polyorder=5)
+spectrum = spectrum.divide(continuum_fit, handle_meta="ff")
+spec_native = spectrum
 
 
-def test_forward():
+def test_dense_forward():
     """Can we clone a model?"""
 
-    emulator = PhoenixEmulator(4700, 4.5, prominence=0.1)
+    wl_native = spec_native.wavelength.value
+    flux_native = spec_native.flux.value
+
+    emulator = LinearEmulator(wl_native, flux_native, prominence=0.02)
     wl = torch.arange(10500, 10600, 0.1)
     spectrum = emulator.forward(wl)
 
     assert spectrum is not None
     assert len(spectrum) == len(wl)
+    assert spectrum.dtype == torch.float64
+
+
+def test_sparse_forward():
+    """Can we clone a model?"""
+
+    wl_native = spec_native.wavelength.value
+    flux_native = spec_native.flux.value
+
+    emulator = SparseLinearEmulator(
+        wl_native, flux_native, prominence=0.02, wing_cut_pixels=1000
+    )
+    spectrum = emulator.forward()
+
+    assert spectrum is not None
+    assert len(spectrum) > 0
     assert spectrum.dtype == torch.float64

### 2022-03-24 paper/paper1/ms.tex
Add description of exponentially spaced pixel samples

@@
 But overall, the telluric model cloning performs adequately, with \censorbox{X percentage} error-per-pixel overall, and only \censorbox{zz} error-per-pixel when restricted to the atmospheric windows.
-We therefore urge some caution when applying \texttt{blas\'e} to telluric modeling in the most saturated regions.
-\authorcomment1{An obvious and good solution is to intialize the model in log flux.
-We may revise the scheme to do so.}
 
@@
 Rotational broadening $v\sin{i}$ and radial velocity $RV$ depend on the observer's viewing location.
-The qualifier \emph{extrinsic} distiguishes from stellar \emph{intrinsic} properties, such as $T_{\mathrm{eff}}, \log{g},\mathrm{and\,} [\mathrm{Fe}/\mathrm{H}]$, which appear the same from any viewing location, at least for stars with isotropic surfaces.
+We follow \citet{czekala15} by emphasizing the qualifier \emph{extrinsic}, to distiguish between stellar \emph{intrinsic} properties, such as $T_{\mathrm{eff}}, \log{g},\mathrm{and\,} [\mathrm{Fe}/\mathrm{H}]$.
+Intrinsic properties appear the same from any viewing location---at least for stars with isotropic surfaces---while extrinsic properties do not.
 The distinction is important because the extrinsic terms act as simple convolutions and translations to the cloned spectrum.
@@
 \begin{eqnarray}
-    \mathsf{S}_{\rm ext}(\bm{\lambda}_S) = \mathsf{S}_{\rm clone}(\lambda_\mathrm{c} - \frac{RV}{c}\lambda_\mathrm{c}) * \zeta \left(\frac{\bm{v}}{v\sin{i}}\right) \label{eqn_convolution}
+    \mathsf{S}_{\rm ext}(\bm{\lambda}_z) = \mathsf{S}_{\rm clone}(\bm{\lambda}_\mathrm{z} - \frac{RV}{c}\lambda_\mathrm{c}) * \zeta \left(\frac{\bm{v}}{v\sin{i}}\right) \label{eqn_convolution}
 \end{eqnarray}
@@
 
-Operationally, the Doppler shift gets applied to the line center positions and not the individual wavelength grid point coordinates, $\bm{\lambda}_S$.
-This choice yields a key benefit: it makes the $RV$ autodiff-aware, meaning that an infinitesimal change to the RV value can easily be sensed through backpropagation by affecting only the line center positions.
-The strategy allows relatively low computational and memory cost since it comports with the sparse matrix implementation.
+Most autodiff-aware \texttt{convolve} operators act in pixel-space, approximating kernels as numerically sampled functions.
+There exists a special exponential spectral sample spacing that allows the convolution operators to work out-of-the-box for rotational broadening.
+We therefore change the sampling from the native stellar and telluric wavelength coordinate grids, $\bm{\lambda}_S$ and $\bm{\lambda}_T$, to a new exponentially sampled wavelength grid, denoted with the subscript $Z$:
+
+\begin{eqnarray}
+    \bm{\lambda}_Z = \lambda_0  \exp{\frac{\bm{v}-v_0}{c}}
+\end{eqnarray}
+
+where $\bm{v}-v_0$ is the velocity vector going from zero to the velocity associated with the largest wavelength, with \emph{linear} spaced velocity samples.
+We choose a sampling in velocity space of 0.5 km/s, which corresponds to about $10\times$ finer than the instrumental resolving power of HPF, and delivers a minimum and maximum wavelength spacing of 0.013 and 0.024 \AA$/\text{pixel}$ respectively for the HPF bandwidth.
+
+Operationally, the radial velocity shift gets applied to the line center positions rather than scaling the entire wavelength grid point coordinates, $\bm{\lambda}_Z$.
+This approximate choice ignores relativistic effects, but yields a convenience: it cleanly makes the $RV$ autodiff-aware, meaning that an infinitesimal change to the RV value can be sensed through backpropagation by affecting only the line center positions.
 
@@
 \begin{eqnarray}
-    \mathsf{M}_{\rm joint} = \mathsf{S}_{\rm ext}(\bm{\lambda}_S) \odot \mathsf{T}_{\rm clone}(\bm{\lambda}_S)
+    \mathsf{M}_{\rm joint} = \mathsf{S}_{\rm ext}(\bm{\lambda}_Z) \odot \mathsf{T}_{\rm clone}(\bm{\lambda}_Z)
 \end{eqnarray}

### 2022-03-24 paper/paper1/ms.tex
Refinements to the existing prose

@@
 \begin{eqnarray}
-    \mathsf{S}_{\rm ext}(\bm{\lambda}_z) = \mathsf{S}_{\rm clone}(\bm{\lambda}_\mathrm{z} - \frac{RV}{c}\lambda_\mathrm{c}) * \zeta \left(\frac{\bm{v}}{v\sin{i}}\right) \label{eqn_convolution}
+    \mathsf{S}_{\rm ext}(\bm{\lambda}_Z) = \mathsf{S}_{\rm clone}(\bm{\lambda}_Z - \frac{RV}{c}\lambda_\mathrm{c}) * \zeta \left(\frac{\bm{v}}{v\sin{i}}\right) \label{eqn_convolution}
 \end{eqnarray}
@@
 There exists a special exponential spectral sample spacing that allows the convolution operators to work out-of-the-box for rotational broadening.
+The design of \texttt{blas\'e} permits the telluric model to be re\"evaluated at any wavelength coordinate vector, provided that it adequately samples the underlying lines.
 We therefore change the sampling from the native stellar and telluric wavelength coordinate grids, $\bm{\lambda}_S$ and $\bm{\lambda}_T$, to a new exponentially sampled wavelength grid, denoted with the subscript $Z$:
@@
 
-Operationally, the radial velocity shift gets applied to the line center positions rather than scaling the entire wavelength grid point coordinates, $\bm{\lambda}_Z$.
+Operationally, the radial velocity shift $RV$ gets applied to the line center positions rather than scaling the entire wavelength grid point coordinates, $\bm{\lambda}_Z$.
 This approximate choice ignores relativistic effects, but yields a convenience: it cleanly makes the $RV$ autodiff-aware, meaning that an infinitesimal change to the RV value can be sensed through backpropagation by affecting only the line center positions.
@@
 \subsection{Joint Stellar and Telluric Model}
-The telluric model and stellar models possess distinct pixel sampling and coordinates.
-The design of \texttt{blas\'e} permits the telluric model to be re\"evaluated at any wavelength coordinate vector, provided that it adequately samples the underlying lines.
-We choose to evaluate the telluric model at the native resolution of the stellar model.
-We then simply multiply the stellar model by the tunable telluric transmission:
+We then simply multiply the stellar model by the telluric transmission:
 
@@
 Instead, we choose the conventional sampled approach, in which we evaluate the extrinsic model at all of the native resolution wavelength coordinates, and then compute the mean value of those pixels within the bounds of each coarse data pixel.
-The resampling procedure is autodiff-aware: the same clusters of native resolution coordinates map to the same data pixel coordinates, no matter what the $RV$ is.
+The resampling procedure is autodiff-aware: the same clusters of high resolution coordinates map to the same data pixel coordinates, no matter what the $RV$ is.
 The $RV$ only dictates what flux values are realized within those pixels.
@@
 \begin{eqnarray}
-    \mathsf{M}(\bm{\lambda}_D) = \resample{\Big[\mathsf{M}_{\rm joint}(\bm{\lambda}_S) * g(R) \Big]} \label{eqn_final_model}
+    \mathsf{M}(\bm{\lambda}_D) = \resample{\Big[\mathsf{M}_{\rm joint}(\bm{\lambda}_Z) * g(R) \Big]} \label{eqn_final_model}
 \end{eqnarray}

### 2022-03-24 src/blase/emulator.py
GPU-specific changes--- CUDA

@@ -338,7 +338,7 @@ class SparseLinearEmulator(LinearEmulator):
 
         device = torch.device(device)
 
-        self.target = self.flux_active
+        self.target = torch.tensor(self.flux_active, dtype=torch.float64, device=device)
 
         ## Define the wing cut
         # Currently defined in *pixels*
@@ -784,10 +784,18 @@ class SparseLogEmulator(SparseLinearEmulator):
             init_state_dict=init_state_dict,
         )
 
+        if device is None:
+            if torch.cuda.is_available():
+                device = "cuda"
+            else:
+                device = "cpu"
+
+        device = torch.device(device)
+
         # The clone-native comparison is done in linear space:
         self.target = torch.exp(torch.tensor(lnflux_native, device=device))[
-            self.active_mask
-        ]
+            self.active_mask.to(device)
+        ].to(device)
 
     def forward(self):
         """The forward pass of the sparse implementation--- no wavelengths needed!

### 2022-04-01 paper/paper1/ms.tex
Writing prose in the discussion sections

@@
 As technology has improved, our data and models have become more voluminous, more precise, and more complicated.
-The mere act of comparing models to observed spectra can resemble a computational statistics challenge as much as a scientific one.
-Here we introduce a machine-learning-based framework \texttt{blas\'e} aimed at solving computational, statistical, and scientific challenges associated with data-model comparisons for modern astronomical spectroscopy.
+The mere act of comparing models to observed spectra can now resemble a computational statistics challenge as much as a scientific one.
+Here we introduce a new machine-learning-based framework \texttt{blas\'e} aimed at solving computational, statistical, and scientific challenges associated with data-model comparisons for modern astronomical spectroscopy.
 
@@
 We show how to supplant the computationally expensive radiative transfer by cloning pre-computed synthetic spectral models.
+We show how to combine both stellar and telluric models in a joint modelling approach.
 In Section \ref{transferLearn} we describe how to fine-adjust the cloned models to real-world data from high-grasp near-IR spectrographs, using a transfer-learning step.
-Section \ref{secInjRec} exhibits injection-recovery tests on simulated data of known properties, revealing the scalability and fundamental limits of the technique.
+%Section \ref{secInjRec} exhibits injection-recovery tests on simulated data of known properties, revealing the scalability and fundamental limits of the technique.  
 Finally we discuss many conceivable extensions to the framework, and directions for future work in Section \ref{secDiscuss}.
@@
 \begin{eqnarray}
-    \mathcal{L} = \sum_i^{N_S} (S_i - S_{\mathrm{clone},i})^2 = \mathsf{R^\intercal}\cdot \mathsf{R}
+    \mathcal{L} = \sum_i^{N_S} (S_i - S_{\mathrm{clone},i})^2 = \mathsf{R^\intercal}\cdot \mathsf{R} \label{simpleLikelihood}
 \end{eqnarray}
@@
 
-\subsection{Performance of transfer learning}
+% \begin{outline}
+%     \1 What is the residual level with bare PHOENIX (i.e. tuning continuum only, not lines)?
+%     \1 What is the residual level with no regularization (i.e. tune all lines and continuum)?
+%     \2 Should be near-zero except for missing lines
+%     \2 What should we do about missing lines?
+%     \1 What is the residual level with modest regularization? What is the typical change to the line?
+%     \2 Plot of cloned FWHM versus FWHM transferred
+%     \2 The pseudo-Voigt line properties can be analytically integrated to give an equivalent width
+%     \2 Plot of cloned EW (before) versus transfer EW (after)
+% \end{outline}
 
-\begin{outline}
-    \1 What is the residual level with bare PHOENIX (i.e. tuning continuum only, not lines)?
-    \1 What is the residual level with no regularization (i.e. tune all lines and continuum)?
-    \2 Should be near-zero except for missing lines
-    \2 What should we do about missing lines?
-    \1 What is the residual level with modest regularization? What is the typical change to the line?
-    \2 Plot of cloned FWHM versus FWHM transferred
-    \2 The pseudo-Voigt line properties can be analytically integrated to give an equivalent width
-    \2 Plot of cloned EW (before) versus transfer EW (after)
-\end{outline}
-
-\section{Injection/recovery tests}\label{secInjRec}
-Here we conduct injection/recovery tests.
+%\section{Injection/recovery tests}\label{secInjRec}
+%Here we conduct injection/recovery tests.
 
@@
 \subsection{Comparison to existing frameworks}
+
+Several astronomical spectral frameworks share similar aims as \texttt{blas\'e}.
+These existing frameworks will have enduring value for the wide range of problems in the field of stellar spectroscopy.
+Here we scrutinize the differences among some of these approaches to clarify how This Work fits in.
 
@@
 The empirical version, \texttt{specmatch-emp} \citep{2017ApJ...836...77Y} matched spectra better than the synthetic templates, but is still too rigid for some applications and requires the assembly of hundreds of standardized high signal-to-noise-ratio templates, ideally with low intrinsic rotational broadening.
-Such a large number of high-quality templates has not yet been established in the near infrared, and may not be feasible for brown dwarfs.
+Such a large number of high-quality templates with high resolving power and low $v\sin{i}$ has not yet been established in the near infrared.
 
@@
 
-\begin{outline}
-    \1 We see some lines that devolve into missing continuum opacity
-    \1 How do the lineshapes cluster in $\sigma_G$ vs $FWHM_L$?
-    \1 How do the lineshapes cluster in Amplitude vs FWHM? Can some lines be rejected?
-\end{outline}
+\section{Conceivable Improvements, Extensions, and Limitations}
 
-\subsection{Alternatives to the PseudoVoigt Profile}
-Two sampled Voigt approaches could be feasible.
-Direct convolution of a sampled Gaussian and Lorentzian profile may be too costly and inexact if pixels are coarsely sampled.
-Analogously, the convolution can be achieved in the Fourier domain through the multiplication of the Fast Fourier Transforms of the Gaussian and Lorentzian profiles, followed by the subsequent Inverse Fourier Transforms.
-The main rationale for either of these two choices would be to anticipate and align with planned convolutions such as rotational broadening and instrumental kernels.
-We do not currently implement these sampled convolutions in \emph{blas\'e} due to their prohibitive computational cost.
+As currently implemented, \texttt{blas\'e} represents a minimum viable product: it out-performs some existing approaches for some science cases.
+The scaffolding of the framework is designed in a way to promote extensibility, so the future of the framework is bright.
+Here we enumerate some planned or conceivable extensions to the framework that would unlock new science arenas.
+We order the presentation of these extensions from our perceived ease-of-implementation, from easiest to hardest.
+
+\subsection{Exact instead of pseudo Voigt Profile}
+We currently employ the pseudo-Voigt profile for its low computational cost.
+We have a prototype exact-Voigt-Hjertingimplementation following \citet{2022ApJS..258...31K}.
+We coarsely estimate that moving to this exact-Voigt implementation wculd decrease some residual regions by $\sim 30\%$, while increasing the computational cost by more than $10\times$ over the existing pseudo-Voigt approximation.
+The exact-Voigt-Hjerting implementation still outbids the higher cost of a direct numerical convolution of a Gaussian and Lorentzian profile.
 
 \subsection{Prospects for closed-form computation of Jacobians}
-In principle, we could compute the Jacobians analytically---since the partial derivatives of the Voigt function are known in closed-form---circumventing the vendor lock-in and memory overhead of an autodiff implementation with PyTorch.
-Further, an autodiff-aware implementation of the \emph{celerit\'e} algorithm \citep{2017AJ....154..220F} does not yet exist in PyTorch, constricting our choices for simultaneous continuum fitting.
-In practice, PyTorch also offers hardware acceleration on GPU or TPU \citep{10.1145/3079856.3080246}, mature support for a range of optimizers, and other perks that motivated its choice over analytic implementations or other machine learning frameworks.
-The GPyTorch framework \citep{2018arXiv180911165G} may offer a workaround to the absence of a \emph{celerit\'e} implementation.
-Still, the performance of an analytic Jacobian implementation over the autodiff route presented here offers an interesting avenue of future research.
+In principle, one could compute the Jacobians analytically---since the partial derivatives of the Voigt function are known in closed-form.
+This analytical approach would cut-out the autodiff approach, using the exact Jacobians to compute the Gradient Descent step.
+The analytical approach may be faster and/or possess lower memory overhead than the autodiff one.
+However, working out the chain-rule for the closed-form analytical Jacobians may become difficult or impossible once combined with the many augmentations get applied at the joint telluric-and-stellar modeling step.
+So analytic Jacobians could principally accelerate the cloning step, but would add significant implementation complexity at the important data-model comparison step.
+We therefore deem this extension somewhat diminishing returns on scarce developer time.
 
-\subsection{Tuning the continuum fit}
-We could also tune or freeze an experimental wavelength-dependent pre-factor to Equation \ref{equation1} to correct for lingering imperfections in the otherwise-fixed continuum flattening procedure.
+\subsection{Addressing the pseudo-continuum with Gaussian Process regression}
+
+We currently assume the input spectra are perfectly normalized to the continuum.
+This approximation impacts all training stages, including the stellar cloning, telluric cloning, and data-model comparison.
+Gaussian Processes (GPs) could and should be brought to bear at each of these optimization stages, as described in the ``global kernels'' in \texttt{Starfish} \citet{czekala15}.
+In short, a GP-likelihood relaxes the assumption that the continuum has been perfectly normalized, in favor of the more realistic statement ``the continuum has been coarsely normalized, with some characteristic-but-as-yet-unknown correlation and scale length and amplitude of the imperfections''.
+That statement translated to the following modification to Equation \ref{simpleLikelihood}:
+
+\begin{eqnarray}
+    \mathcal{L} =  \frac{1}{2}\mathsf{R^\intercal} \mathsf{C}^{-1} \mathsf{R} +\frac{1}{2}\ln{\det{\mathsf{C}}} \label{GPLikelihood}
+\end{eqnarray}
+
+where we introduce the covariance matrix $\mathsf{C}$, with its concommittant kernel and collection of typically 2-3 parameters.
+We anticipate that this GP likelihood would have the greatest impact on stellar stars with significant band-heads and line-blanketing: spectra with a so-called ``pseudo-continuum''.
+M-dwarfs and brown dwarfs fall into this fear-inducing category.
+
+The main demerit of moving to a GP-likelihood is computational cost.
+Fortunately a few efficient autodiff-aware implementations of GPs exist.
+The \texttt{celerite} algorithm \emph{celerit\'e} \citep{2017AJ....154..220F} has an exact backpropation implementation \citep{2018RNAAS...2...31F} that scales linearly with the number of data points.
+The celerite algorithm does not currently have a PyTorch implementation.
+The \texttt{GPyTorch} framework \citep{2018arXiv180911165G} has a large category of approximate and exact GPs that could be straight-forwardly dropped into \texttt{blas\'e}.
+Even still these GPs could increase the computation cost by of order $10\times$.
+
+An alternative to GPs could be to simply tune the $\mathsf{P}$ term that represents the wavelength-dependent pre-factor to Equation \ref{equation1}.
+Tuning $\mathsf{P}$ would correct for only large-scale imperfections in the otherwise-fixed continuum flattening procedure.
 

### 2022-04-01 paper/paper1/ms.tex
more prose on doppler imaging, RV jitter, and more

@@
 
-\subsection{Prospects for closed-form computation of Jacobians}
-In principle, one could compute the Jacobians analytically---since the partial derivatives of the Voigt function are known in closed-form.
-This analytical approach would cut-out the autodiff approach, using the exact Jacobians to compute the Gradient Descent step.
-The analytical approach may be faster and/or possess lower memory overhead than the autodiff one.
-However, working out the chain-rule for the closed-form analytical Jacobians may become difficult or impossible once combined with the many augmentations get applied at the joint telluric-and-stellar modeling step.
-So analytic Jacobians could principally accelerate the cloning step, but would add significant implementation complexity at the important data-model comparison step.
-We therefore deem this extension somewhat diminishing returns on scarce developer time.
-
 \subsection{Addressing the pseudo-continuum with Gaussian Process regression}
@@
 
-\subsection{What about minibatches?}
+\subsection{Minibatches and Stochastic Gradient Descent}
 
-Minibatches offer a natural solution to the line-by-line computational bottleneck while also providing optimization performance improvements.
-Minibatches act as a form of regularization, the principal source of stochasticity in the Stochastic Gradient Descent algorithm, which tends to have better convergence than full-batch Gradient Descent \citep{2016arXiv160904747R}.
-The Machine Learning/Artificial Intelligence (ML/AI) community employs minibatches to handle similar situations, in which the dataset volume exceeds the available GPU RAM.
-Our configuration departs slightly from the typical ML/AI situation since their models are usually ``small'' relative to their datasets.
-Here our model is ``big'' relative to our dataset.
+Currently each training epoch ``sees'' the entire dataset, a setup dubbed ``full-batch'' gradient descent.
+An alternative scheme allow training with only a portion of the entire dataset at a time in ``minibatches''.
+The massive data volumes in modern Neural Network applications cannot fit into the GPU memory, and so minibatches are a necessity.
+Our meager $1$ MB dataset can easily fit into the GPU memory, but our model can be large if we have a large number of pixels or lines or both.
+So while minibatches may not be required due to data size limitations they may be useful for particularly large models.
+Minibatches also act as a form of regularization, the principal source of stochasticity in the Stochastic Gradient Descent algorithm, which tends to have better convergence than full-batch Gradient Descent \citep{2016arXiv160904747R}.
 
-Here we choose to assemble and evaluate only a portion of the matrix at a time, in minibatches.
+We experimented with minibatches by assembling and evaluate only a portion of the dense $\bar{\bm{F}}$ matrix at a time, in minibatches.
 The choice to evaluate only a portion of lines at a time would mean the model is inaccurately evaluated at all wavelength pixels.
-Instead, we choose to evaluate all lines, but only on a subset $N_{\mathrm{batch}}$ of the total pixels $N_s$, so that the model can eventually converge to exact at those points.
+Instead, we choose to evaluate all lines, but only on a random subset $N_{\mathrm{batch}}$ of the total pixels $N_s$, so that the model can eventually converge to exact at those points.
 All lines are allowed to update at each glimpse of a minibatch, but many lines with cores far from minibatch pixels will provide only weak information about how the loss scalar changes for their parameters.
@@
 At present, this threshold is determined experimentally.
-The indices of the wavelength points can either be set to be contiguous in blocks, or random with possibly large gaps between indices, or else with some other scheme tailored to the most informative wavelength regions.
+The indices of the wavelength points can either be set to be contiguous in blocks (duplicating the sparse matrix approach), or random with possibly large gaps between indices, or else with some other scheme tailored to the most informative wavelength regions.
 We experimented with several techniques and settled on random sampling with replacement, meaning that some pixels get revisited more often than others, with some minibatches tailored only to the line cores.
+Overall minibatches as implemented above performed worse than the sparse implementation, with both lower accuracy and slower computation time.
 
@@
 
-Some lines, such as Hydrogen, Sodium, potassium, neutral metal lines, and others have extremely broad line wings, approaching larger than the $\sim6000$ pixels we allocate for the sparse implementation.
+Some lines, such as Hydrogen, sodium, potassium, and others have extremely broad line wings, approaching larger than the $\sim6000$ pixels we allocate for the sparse implementation.
 These special lines should be handled separately from the weak lines, both from a computational performance perspective and an accuracy perspective.
@@
 We then construct and evaluate the entire dense matrix for those lines: $\sim 330\;000 \times 20$.
-The number of FLOPS in each category scales as about 6 Million for the 20 broad lines versus about 36 Million for the sea of about 6500 narrow lines, depending on the exact choices for wing cuts and number of lines.
+The number of FLOPS in each category scales as about 6 Million for the 20 broad and dense lines versus about 36 Million for the sea of about 6500 narrow and sparse lines, depending on the exact choices for wing cuts and number of lines.
 
@@
 
-where $\odot$ is the element-wise product (\emph{a.k.a} Hadamard product), $\mathcal{S}$ is the sigmoid function, and we have introduced three new tunable parameters for each of the $j$ broad lines; $\lambda_t$ is the truncation wavelength, $b$ is a scale parameter for how slowly or how rapidly in wavelength-space the transition from non-Lorentzian proceeds, and $a$ is a possibly negative stretch parameter that controls whether the line wing is sub- or super- Lorentzian.
+where $\odot$ is again the element-wise product (\emph{a.k.a} Hadamard product), $\mathcal{S}$ is the sigmoid function, and we have introduced three new tunable parameters for each of the $j$ broad lines; $\lambda_t$ is the truncation wavelength, $b$ is a scale parameter for how slowly or how rapidly in wavelength-space the transition from non-Lorentzian proceeds, and $a$ is a possibly negative stretch parameter that controls whether the line wing is sub- or super- Lorentzian.
 
@@
 It enforces that the perturbation only produces absorption and not emission profiles.
+
+\subsection{Wavelength dependent limb darkening}
+Currently the extrinsic model step possesses up to four parameters: the $v\sin{i}$ and $RV$, and 2 optional parameters for limb-darkening.
+These four parameters may adequately parameterize a star with a uniform stellar disk.
+Extremely high signal-to-noise-ratio spectra of rapidly rotating stars may require additional flexibility.
+The limb darkening is generally wavelength dependent, and so a pan-chomatic spectrum may require a different limb-darkening from the blue end to the red end.
+The limb darkening may instead depend on physical properties of the spectral line formation, such as physical depth of formation, and so the extent of limb-darkening may jump haphazardly from line-to-line-to-line, rather than as a predictably smooth function across wavelength.
+\texttt{blas\'e} could be built to handle such a seemingly pathological scenario by adding a vector of limb darkening parameters, one for each line.
+One would have to regularize the fits with some typical limb darkening and a heuristic penalty for departures from this mean.
+
+\subsection{Doppler Imaging}
+The fixed $v\sin{i}$ approximation breaks down for stars with large-scale surface features.
+Doppler imaging \authorcomment1{xx, yy, zz} attempts to reverse engineer the surface map from the extent to which observed line profiles depart from a pristine rotational broadening kernel.
+This reverse engineering step suffers from a vast number of geometrical degeneracies, but still provides useful constraints on stellar surfaces.
+Following LugerXX, we emphasize a distinction between A) longitudinally symmetric surface features, and B) longitudinally asymmetric surface features.
+Most radial velocity practitioners think about the latter, since longitudinally asymmetric surface features imbue skewness to the line profiles, causing radial velocity perturbations easily detectable in radial velocity time series.
+These confound exoplanet searches.
+
+Longitudinally symmetric surface features to not change as the star rotates on its axis.
+The existence of these features change the \emph{kurtosis} of the spectral line.
+For example a hypothetical non-emitting (black) polar starspot exhibits a deficit in flux at the line core, resulting in less zero-velocity flux than its homogeneous counterpart.
+A dark zonal band results in equal-sized bites out of the red- and blue- sides of the line.
+
+It is much easier to reverse-engineer longitudinally asymmetric features than longitudinally symmetric ones.
+The latter requires exact knowledge of the underlying spectral template.
+Isolated deep spectral lines constitute the only practical scenario where exact knowledge can be claimed.
+Isolated spectral lines may be scarce or absent for M-dwarfs and brown dwarfs where lines blend together ostensibly in an inseparable way, counfounding Doppler imaging.
+
+\texttt{blas\'e} offer a new approach to Doppler imaging that may overcome these historical limitations.  In \texttt{blas\'e} can...
+
+\subsection{RV jitter and microturbulence}
+
+Already \texttt{blas\'e} is equipped to fit every single line with its own systemic $RV$, by tuning the line center position $\lambda_\mathrm{c}$ at each training epoch.
+There exists both empirical evidence and some theoretical motivation that $RV$ jitter varies from line-to-line \authorcomment1{Xavier Dumusque 2018}.
+Importantly, the extent of this line-by-line RV jitter could be predicted in part by the depth of line formation.
+A future extreme-precision-RV (EPRV) version of \texttt{blase} could leverage this information in a few ways.
+As an example, the depth-of-formation for all or a subset of lines could be obtained and associated with each spectral line.
+A regulariztion scaling term could be introduced to allow the line positions to vary, but only in proportion to their depth of formation.
+
+\subsection{Prospects for closed-form computation of Jacobians}
+In principle, one could compute the Jacobians analytically---since the partial derivatives of the Voigt function are known in closed-form.
+This analytical approach would cut-out the autodiff approach, using the exact Jacobians to compute the Gradient Descent step.
+The analytical approach may be faster and/or possess lower memory overhead than the autodiff one.
+However, working out the chain-rule for the closed-form analytical Jacobians may become difficult or impossible once combined with the many augmentations get applied at the joint telluric-and-stellar modeling step.
+So analytic Jacobians could principally accelerate the cloning step, but would add significant implementation complexity at the important data-model comparison step.
+We therefore deem this extension somewhat diminishing returns on scarce developer time.
 

### 2022-06-10 paper/paper1/ms.tex
Delete placeholder text, tbd return back to Doppler Imaginge section later
Overhaul paper abstract based on CoolStars21, fix LaTeX warnings

@@
 
-    We introduce blas\'e, a framework for transfer learning for high-grasp echelle spectroscopy.
-\texttt{blas\'e} produces extensible models that are both physically informed and highly accurate when compared to data.
-These models can be evaluated quickly for use in likelihood calculations.
-It works on data with large spectral grasp: simultaneously wide bandwidth and high spectral resolution.
+Comparison of \'echelle spectra to synthetic models has become a computational statistics challenge, with over ten thousand individual spectral lines affecting a typical cool star \'echelle spectrum.
+Telluric artifacts, imperfect line lists, inexact continuum placement, and inflexible models frustrate the scientific promise of these information-rich datasets.
+Here we debut an interpretable machine-learning framework \emph{blas\'e} that addresses these challenges.
+The semi-empirical approach can be viewed as ``transfer learning''--first pre-training models on noise-free precomputed synthetic spectral models, then learning the corrections to line depths and widths from whole-spectrum fitting to an observed spectrum.
+The auto-differentiable model employs back-propagation, the fundamental algorithm empowering modern Neural Networks.
+Here, however, the 40,000+ parameters symbolize physically interpretable line profile properties such as amplitude, width, location, and shape, plus radial velocity and rotational broadening.
+This hybrid data-/model- driven framework allows joint modeling of stellar and telluric lines simultaneously, a potentially transformative step forwards for mitigating the deleterious telluric contamination in the near-infrared.
+The \emph{blas\'e} approach also acts as a deconvolution tool.
+It is suitable for Doppler Imaging scenarios with longitudinally symmetric surface features like bands or polar spots, which evade detection in differential techniques.
+Blas\'e can also create super-resolution semi-empirical templates useful for critically evaluating atomic and molecular line lists.
+Its sparse-matrix architecture and GPU-acceleration make \emph{blas\'e} fast, with end-to-end training of an octave of bandwidth (50+ \'echlle orders) in under 1 minute.
+The open-source PyTorch-based code \texttt{blase} includes tutorials and API documentation.
+We show how the tool fits into the existing Python spectroscopy ecosystem, demonstrate a range of astrophysical applications, and discuss limitations and future extensions.
 
@@
 
-\keywords{High resolution spectroscopy (2096), Stellar spectral lines (1630), Astronomy data modeling(1859), GPU Computing (1969), Calibration (2179), Radial Velocity (1332), Maximum likelihood estimation (1901), Deconvolution (1910), Atomic spectroscopy (2099), Stellar photospheres (1237)},
+\keywords{High resolution spectroscopy (2096), Stellar spectral lines (1630), Astronomy data modeling(1859), GPU Computing (1969), Calibration (2179), Radial Velocity (1332), Maximum likelihood estimation (1901), Deconvolution (1910), Atomic spectroscopy (2099), Stellar photospheres (1237)}
 
@@
 
-\subsection{Augmenting the Stellar Clone Model for $RV$ and $v\sin{i}$ }
+\subsection{Augmenting the Stellar Clone with radial velocity and rotational broadening }
 Real stars possess two key extrinsic properties.
@@
 
-\texttt{blas\'e} offer a new approach to Doppler imaging that may overcome these historical limitations.  In \texttt{blas\'e} can...
+\texttt{blas\'e} offer a new approach to Doppler imaging that may overcome these historical limitations.
 
@@
 % Table
-\begin{deluxetable}{cp{12cm}}
+\begin{deluxetable}{cp{10cm}}
+    \tabletypesize{\scriptsize}
     \tablecaption{Notation used in this paper\label{table2}}
@@
     \enddata
+
 \end{deluxetable}

### 2022-06-10 docs/conf.py
Add frequently asked questions... and Log/Linear comparison

@@ -61,6 +61,7 @@ nbsphinx_thumbnails = {
     "tutorials/demo3": "_static/397px-PyTorch_logo_icon.svg.png",
     "tutorials/demo4": "_static/397px-PyTorch_logo_icon.svg.png",
     "tutorials/demo5": "_static/397px-PyTorch_logo_icon.svg.png",
+    "tutorials/demo6": "_static/397px-PyTorch_logo_icon.svg.png",
 }
 
 # -- Options for HTML output -------------------------------------------------

### 2022-06-10 docs/conf.py
Make a standalone FAQ page rather than a subsection in Tutorials

@@ -61,7 +61,7 @@ nbsphinx_thumbnails = {
     "tutorials/demo3": "_static/397px-PyTorch_logo_icon.svg.png",
     "tutorials/demo4": "_static/397px-PyTorch_logo_icon.svg.png",
     "tutorials/demo5": "_static/397px-PyTorch_logo_icon.svg.png",
-    "tutorials/demo6": "_static/397px-PyTorch_logo_icon.svg.png",
+    "FAQ/demo6": "_static/397px-PyTorch_logo_icon.svg.png",
 }
 
 # -- Options for HTML output -------------------------------------------------

### 2022-06-10 docs/conf.py
Split tutorial steps 4 and 5 into standalone notebooks

@@ -61,6 +61,7 @@ nbsphinx_thumbnails = {
     "tutorials/demo3": "_static/397px-PyTorch_logo_icon.svg.png",
     "tutorials/demo4": "_static/397px-PyTorch_logo_icon.svg.png",
     "tutorials/demo5": "_static/397px-PyTorch_logo_icon.svg.png",
+    "tutorials/demo6": "_static/397px-PyTorch_logo_icon.svg.png",
     "FAQ/demo6": "_static/397px-PyTorch_logo_icon.svg.png",
 }
 

### 2022-06-22 paper/paper1/ms.tex
re-enable format-on-save with latexindent, flaky on linux tho...
One less placeholder
Minor paper edit and nice formatting

@@
 \usepackage{amsmath,bm}
+\definecolor{belize}{RGB}{41, 128, 185}
+\definecolor{peter}{RGB}{52, 152, 219}
+\definecolor{nephritis}{RGB}{39, 174, 96}
+\definecolor{asbestos}{RGB}{127, 140, 141}
+\hypersetup{linkcolor=nephritis,citecolor=belize,filecolor=asbestos,urlcolor=peter}
 
@@
 In extreme precision radial velocity (EPRV) applications, cross-correlation methods work \citep{2018A_A...620A..47D}, but have many limitations \citep{2022arXiv220110639Z}.
-Addressing telluric absorption at the $\mathrm{cm/s}$ level may require joint modeling of the star and the Earth's atmospheric absorption before convolution with an instrumental kernel.
+Among the many such limitations, one pernicious noise floor stands out as enigmatic: telluric mitigation.
+Many practitioners today simply mask these telluric regions, yet micro-tellurics still inject variance into the spectrum that cannot be easily accounted for with existing methods.
+A robust accounting of telluric absorption at the $\mathrm{cm/s}$ level may require joint modeling of the star and the Earth's atmospheric absorption \emph{before} convolution with an instrumental kernel.
 This telluric joint modeling capability does not yet exist at a precision that can meet these strenuous demands.
@@
 In this paper we show that autodiff-aware semi-empirical models offer an appealing middle ground: informed from self-consistent models, but refined with data.
-In Section \ref{methodology} we introduce the \emph{blas\'e} forward-model design and its \texttt{PyTorch}-based \cite{2019arXiv191201703P} implementation.
+In Section \ref{methodology} we introduce the \emph{blas\'e} forward-model design and its \texttt{PyTorch}-based \citep{2019arXiv191201703P} implementation.
 We show how to supplant the computationally expensive radiative transfer by cloning pre-computed synthetic spectral models.
@@
 The data-model comparison step will always take place in linear flux units, so the only operational difference is the behavior for deep and saturated lines.
-Appendix \censorbox{XX} lists the equations adjusted to log flux units.
+Appendix \ref{appendixLogScale} lists the equations adjusted to log flux units.
 
@@
     \enddata
-
 \end{deluxetable}

### 2022-06-22 paper/paper1/ms.tex
Fill in placeholder text with some numbers on telluric lines.

@@
 The number of pixels in the telluric model can be chosen at the time of running a \texttt{TelFit} model.
-Here we choose a spectral resolution \censorbox{$R\sim1\times10^6$}, adequate for resolving telluric lines, and yielding \censorbox{xx} pixels across the HPF bandwidth.
-This pixel sampling is slightly \censorbox{finer} than the native PHOENIX pixel sampling.
+Here we choose a spectral resolution $R\sim10^6$, adequate for resolving narrow telluric lines, and yielding about 2 million pixels across the entire HPF bandwidth.
+This pixel sampling is about $6\times$ finer than the native PHOENIX pixel sampling.
 The number of telluric lines depends on the atmospheric properties, in particular the Earth surface temperature and relative humidity $RH$.
-For a \censorbox{T=x}, \censorbox{RH=y}, we detect \censorbox{zz} telluric lines across the entire HPF bandwidth.
+For a surface temperature of $T=290 K\; (62^\circ~F)$, a relative humidity of 40\%, and typical conditions for McDonald Observatory, we anticipate 3615 telluric lines across the entire HPF bandwidth.
 

### 2022-06-22 paper/paper1/ms.tex
Dump the paper into Grammarly for high-quality spelling and grammar checking

@@
 \author{Michael Gully-Santiago}
-\affiliation{University of Texas at Austin Department of Astronomy}
+\affiliation{The University of Texas at Austin Department of Astronomy}
 
 \author{Caroline V. Morley}
-\affiliation{University of Texas at Austin Department of Astronomy}
+\affiliation{The University of Texas at Austin Department of Astronomy}
 
@@
 Blas\'e can also create super-resolution semi-empirical templates useful for critically evaluating atomic and molecular line lists.
-Its sparse-matrix architecture and GPU-acceleration make \emph{blas\'e} fast, with end-to-end training of an octave of bandwidth (50+ \'echlle orders) in under 1 minute.
+Its sparse-matrix architecture and GPU-acceleration make \emph{blas\'e} fast, with end-to-end training of an octave of bandwidth (50+ \'echelle orders) in under 1 minute.
 The open-source PyTorch-based code \texttt{blase} includes tutorials and API documentation.
@@
 Tens of thousands or more individual spectral lines give rise to a sea of undulations that imbue each stellar spectrum with its characteristic appearance.
-The identification and understanding of these lines has defined a large category of astrophysics over the last century.
+The identification and understanding of these lines have defined a large category of astrophysics over the last century.
 The field grew from by-eye catalogs of stellar templates \citep{1901AnHar..28..129C} to quantifying the role of atomic ionization balance \citep{1925PhDT.........1P}, to modern synthetic forward models including millions or billions of lines \citep{husser13, 2021ApJ...920...85M}.
@@
 
-The metaphorical holy grail of astronomical spectroscopy is a function that takes in an observed astronomical spectrum and reports back the postition, amplitude, width, and shape of all of its spectral lines, automatically, accurately, and precisely.
+The metaphorical holy grail of astronomical spectroscopy is a function that takes in an observed astronomical spectrum and reports back the position, amplitude, width, and shape of all of its spectral lines, automatically, accurately, and precisely.
 The function would go further.
 It would report back the systemic radial velocity (RV) and rotational broadening ($v\sin{i}$).
-Finally the function would---in an \emph{Inception}-like manner---provide what it believes to be the interpretable generating function that produced the data, so that we may gain insights on future examples of this or other stars.
+Finally, the function would---in an \emph{Inception}-like manner---provide what it believes to be the interpretable generating function that produced the data, so that we may gain insights on future examples of this or other stars.
 Solving this problem is hard, for at least four reasons.
@@
 Third, the extent of line blending and realized line shape depends strongly on the spectral resolution of the spectrograph, the rotational broadening of the star, and possibly the instrumental configuration at the time of observation.
-Finally, telluric absorption lines commingle with the astronomical spectral lines of interest, censoring some spectral regions entirely, or partially confounding other lines with chance allignments.
+Finally, telluric absorption lines commingle with the astronomical spectral lines of interest, censoring some spectral regions entirely, or partially confounding other lines with chance alignments.
 
@@
 \emph{in prep}), CHIMERA \citep{2015ApJ...807..183L}, the \texttt{Cannon} \citep{2017ApJ...836....5H},  \texttt{MINESweeper} \citep{2020ApJ...900...28C}, and recently \texttt{ExoJAX} \citep{2022ApJS..258...31K}.
-The design of these frameworks necessarily have to make a choice in the bias-variance tradeoff: is the tool more \emph{data}-driven or more \emph{model}-driven?
+The designs of these frameworks necessarily have to make a choice in the bias-variance tradeoff: is the tool more \emph{data}-driven or more \emph{model}-driven?
 The statistical tradeoff can be viewed as a concession in physical self-consistency for model flexibility: more or fewer parameters; more \emph{accurate} or more \emph{precise}.
@@
 
-In this paper we show that autodiff-aware semi-empirical models offer an appealing middle ground: informed from self-consistent models, but refined with data.
+In this paper, we show that autodiff-aware semi-empirical models offer an appealing middle ground: informed by self-consistent models but refined with data.
 In Section \ref{methodology} we introduce the \emph{blas\'e} forward-model design and its \texttt{PyTorch}-based \citep{2019arXiv191201703P} implementation.
 We show how to supplant the computationally expensive radiative transfer by cloning pre-computed synthetic spectral models.
-We show how to combine both stellar and telluric models in a joint modelling approach.
+We show how to combine both stellar and telluric models in a joint modeling approach.
 In Section \ref{transferLearn} we describe how to fine-adjust the cloned models to real-world data from high-grasp near-IR spectrographs, using a transfer-learning step.
 %Section \ref{secInjRec} exhibits injection-recovery tests on simulated data of known properties, revealing the scalability and fundamental limits of the technique.  
-Finally we discuss many conceivable extensions to the framework, and directions for future work in Section \ref{secDiscuss}.
+Finally, we discuss many conceivable extensions to the framework and directions for future work in Section \ref{secDiscuss}.
 
@@
 
-The choice of limiting the bandwidth to a region-of-interest around a single echelle spectrograph bandwidth stems from computational constraints.
+The choice of limiting the bandwidth to a region of interest around a single echelle spectrograph bandwidth stems from computational constraints.
 In principle, there is no fundamental limit to the bandwidth one could clone with the method presented here, up to and including the entire precomputed synthetic spectral model bandwidth.
@@
 An additional continuum flattening step ensures that subsequent spectral line finding steps get applied uniformly.
-Here, we identify the continuum regions by the most prominent peaks in the specturm separated by a distance of at least 5,000$-$90,000 pixels from adjacent peaks, yielding between 60 and 5 control points.
+Here, we identify the continuum regions by the most prominent peaks in the spectrum separated by a distance of at least 5,000$-$90,000 pixels from adjacent peaks, yielding between 60 and 5 control points.
 We then fit an $n_{\mathrm{poly}} = 3^{rd}-15^{th}$ order polynomial to these peaks, evaluate the polynomial on the entire bandwidth $\mathsf{P}(\bm{\lambda}_S)$, and divide the spectrum by this trend.
-The result should be a flattened ``continuum-normalized'' spectrum familiar to practitioners in high resolution spectroscopy, with the continuum level close to unity.
+The result should be a flattened ``continuum-normalized'' spectrum familiar to practitioners in high-resolution spectroscopy, with the continuum level close to unity.
 It is this spectrum that will serve as the centerpiece of subsequent training steps.
-We therefore drop any subscript and simply refer to this flattened spectrum as $\mathsf{S}$:
+We, therefore, drop any subscript and simply refer to this flattened spectrum as $\mathsf{S}$:
 
@@
 
-This high-pass filtering step should be set to capture the genuine spectral shape, without over-fitting broad line-wings such as those in deep Hydrogen and sodium lines.
+This high-pass filtering step should be set to capture the genuine spectral shape, without over-fitting broad line wings such as those in deep Hydrogen and sodium lines.
 We have experimentally found it helpful to hand-tune models of exceptionally broad and non-standard line wings with an optional pre-processing step that depends on the wavelength range of interest.
@@
     \caption{Visual flowchart of \texttt{blas\'e}.
-The technique starts with a precomputed synthetic stellar spectrum and a precomputed synthetic telluric spectrum with properties close to those previously estimated for target.
+The technique starts with a precomputed synthetic stellar spectrum and a precomputed synthetic telluric spectrum with properties close to those previously estimated for the target.
 Both the stellar and telluric spectra get cloned.
 The stellar model is warped to its extrinsic properties, and then the stellar and telluric models get multiplied together.
-Next this joint model is convolved with an instrumental kernel and resampled to the data spectrum.
-Finally the joint model parameters are optimized to match the data, subject to regularization (not shown).
+Next, this joint model is convolved with an instrumental kernel and resampled to the data spectrum.
+Finally, the joint model parameters are optimized to match the data, subject to regularization (not shown).
 The final products are semi-empirical super-resolution stellar and telluric templates, the $RV$ and $v\sin{i}$ of the star, and the line amplitude, location, width, and shape of all 10,000+ stellar and telluric lines.
@@
 As seen in Figure \ref{fig_Nlines_vs_teff}, the number of lines can exceed 7,000, meaning the model has over 7,000 $\times 3 =$ 21,000 free parameters.
-Fitting that large number of parameters is difficult with conventional optimizers.
+Fitting more than about 300 parameters is difficult with conventional optimizers-- a $70\times$ shortfall in computation ability.
 Here we employ a variant of Stochastic Gradient Descent (SGD), an optimization technique that can scale to a virtually unlimited number of parameters \citep{2016arXiv160904747R}.
 This technique computes the derivative of the loss scalar with respect to each of the parameters, the so-called \emph{Jacobian}: $(\frac{\partial \mathcal{L}}{\partial a_j}, \frac{\partial \mathcal{L}}{\partial \sigma_j}, \frac{\partial \mathcal{L}}{\partial \gamma_j})$.
-The Jacobian indicates how the MSE would decrease with a change in the parameter-of-interest, or put simply which-way and by-how-much you have to change the line properties to get a better fit.
+The Jacobian indicates how the MSE would decrease with a change in the parameter-of-interest, or put simply ``which way and by how much'' you have to change the line properties to get a better fit.
 
@@
 We retain a relatively small number of pixels $N_{\rm cut}$ adjacent to the line center.
-Setting this wing cut produces a speedup by a factor of $\frac{N_S}{N_{\mathrm{cut}}}$, which can exceed $100\times$ for wide bandwidth spectra.The choice of $N_{\rm cut}$ is nuanced.
+Setting this wing cut produces a speedup by a factor of $\frac{N_S}{N_{\mathrm{cut}}}$, which can exceed $100\times$ for wide bandwidth spectra.
+The choice of $N_{\rm cut}$ is nuanced.
 It should be set large enough that truncation effects are not seen for the broadest lines.
@@
 We \emph{could have} demanded that the wing cut be exactly 30 \AA, or some other heuristic, but that choice would result in so-called \emph{jagged} or \emph{ragged} arrays: collections of 1D arrays that differ in their lengths.
-Some lines would have 5999 pixels, some would have 6000, and others would have 2999 or 3000 pixels, depending whether they occur to the red or blue side of the $1 \;\mu$m kink in sampling.
+Some lines would have 5999 pixels, some would have 6000, and others would have 2999 or 3000 pixels, depending on whether they occur to the red or blue side of the $1 \;\mu$m kink in sampling.
 Most efficient CPU and GPU algorithms are not equipped to deal with ragged arrays, therefore increasing the computation time.
@@
 We employ a \texttt{TelFit} model, though any precomputed synthetic telluric model will work.
-The \texttt{TelFit} model does not contain any continuum sources of opacity, and so we can skip the flattening procedure described in Equation \ref{eqnFlattening}.
-We orchestrate the same initialization and line finding as in the stellar models, and obtain a coarse clone.
+The \texttt{TelFit} model does not contain any continuum sources of opacity, so we can skip the flattening procedure described in Equation \ref{eqnFlattening}.
+We orchestrate the same initialization and line finding as in the stellar models and obtain a coarse clone.
 
@@
 This pixel sampling is about $6\times$ finer than the native PHOENIX pixel sampling.
-The number of telluric lines depends on the atmospheric properties, in particular the Earth surface temperature and relative humidity $RH$.
-For a surface temperature of $T=290 K\; (62^\circ~F)$, a relative humidity of 40\%, and typical conditions for McDonald Observatory, we anticipate 3615 telluric lines across the entire HPF bandwidth.
+The number of telluric lines depends on the atmospheric properties, in particular, the local surface temperature and relative humidity $RH$.
+For a surface temperature of $T=290 \mathrm{K}\; (62^\circ~\mathrm{F})$, relative humidity of 40\%, and typical conditions for McDonald Observatory, we anticipate 3615 telluric lines across the entire HPF bandwidth.
 
@@
 So we can hypothetically tolerate a much smaller wing cut for telluric lines.
-In practice, telluric lines can be extremely broad, and accurately cloning these broad telluric lines requires about $N_{\rm cut}\sim6000$ pixels, comparable to the stellar case.
+In practice, telluric lines can be extremely broad, and accurately cloning these broad telluric lines requires about $N_{\rm cut}\sim6000$ pixels, comparable to the stellar scenario.
 
@@
 We compute the residual $\mathsf{R}(\bm{\lambda}_S)$ of native PHOENIX minus cloned model, illustrated in the bottom panel of Figure \ref{fig_cloned_spectrum_demo}.
-We see an RMS residuals of 1.2\%/pixel at native resolution.
+We see an RMS residual of 1.2\%/pixel at native resolution.
 We identify three main categories of cloning flaws.
 
-The first, and expected, source of large residuals is simply missing line opacity due to our finite prominence threshold.
+The first---and expected---source of large residuals is simply missing line opacity due to our finite prominence threshold.
 Lines with prominence less than $P_{\rm rom}$ yield residual notches with strengths comparable to $P_{\rm rom}$.
@@
 Finally, and most perniciously, a large category of residuals appear near the wings of the deepest and broadest lines---such as Hydrogen and neutral alkali metal lines.
-The true lines exhibit advanced lineshapes, such as non-Lorentzian line-wings that are not captured with the overly simplistic Voigt line profile.
+The true lines exhibit advanced lineshapes, such as non-Lorentzian line wings that are not captured with the overly simplistic Voigt line profile.
 Figure \ref{fig_zoom_cloning_performance} highlights super-Lorentzian line wings around a line at 8691~\AA.
-Narrow lines devolve into missing linewing opacity, the favored tradeoff when the continuum estimate's poor performance outweighs the pain of a narrow-but-tolerably-small spike.
-This flaw can be seen  where a line initialized at 8692.5~\AA~ and another pair of lines at 8690.0 all melt into linewings.
+Narrow lines devolve into missing line wing opacity, the favored tradeoff when the continuum estimate's poor performance outweighs the pain of a narrow-but-tolerably-small spike.
+This flaw can be seen where a line initialized at 8692.5~\AA~ and another pair of lines at 8690.0 all melt into line wings.
 
@@
 
-But overall, the telluric model cloning performs adequately, with \censorbox{X percentage} error-per-pixel overall, and only \censorbox{zz} error-per-pixel when restricted to the atmospheric windows.
+But overall, the telluric model cloning performs adequately, with \censorbox{X percentage} error-per-pixel, and only \censorbox{zz} error-per-pixel when restricted to the atmospheric windows.
 
@@
 Most autodiff-aware \texttt{convolve} operators act in pixel-space, approximating kernels as numerically sampled functions.
-There exists a special exponential spectral sample spacing that allows the convolution operators to work out-of-the-box for rotational broadening.
+There exists a special exponential spectral sample spacing that allows the convolution operators to work out of the box for rotational broadening.
 The design of \texttt{blas\'e} permits the telluric model to be re\"evaluated at any wavelength coordinate vector, provided that it adequately samples the underlying lines.
-We therefore change the sampling from the native stellar and telluric wavelength coordinate grids, $\bm{\lambda}_S$ and $\bm{\lambda}_T$, to a new exponentially sampled wavelength grid, denoted with the subscript $Z$:
+We, therefore, change the sampling from the native stellar and telluric wavelength coordinate grids, $\bm{\lambda}_S$ and $\bm{\lambda}_T$, to a new exponentially sampled wavelength grid, denoted with the subscript $Z$:
 
@@
 Real astronomical instruments usually have wavelength-dependent resolving power, which complicates the implementation.
-The extent to which this effect matters will depend on the science case.
+The extent to which this effect matters will depend on the science application.
 For now, \texttt{blas\'e} simply assumes a fixed resolving power.
@@
 
-We apply some amount of \emph{regularization}, a restiction on the allowed values the model parameters can take on.
-Fortunately we have a firm theoretical basis to justify this regularization.
+We apply some amount of \emph{regularization}, a restriction on the allowed values the model parameters can take on.
+Fortunately, we have a firm theoretical basis to justify this regularization.
 We believe our precomputed synthetic spectral models are \emph{quasi-statically correct}: the predicted spectra resemble the unobserved ``latent spectrum'' with lines in the correct place, but just with the wrong amplitudes.
@@
 We default to the following choice.
-We fix all line parameters except for amplitude, which receives a regularization of 0.01, meaning that its ampltidue should be within about 1\% of its cloned value.
-The extrinsic $v\sin{i}$ and $RV$ have no regularization, but in practice they need to be initialized close to their plausible values.
+We fix all line parameters except for amplitude, which receives a regularization of 0.01, meaning that its amplitude should be within about 1\% of its cloned value.
+The extrinsic $v\sin{i}$ and $RV$ have no regularization, but in practice, they need to be initialized close to their plausible values.
 
@@
 The empirical version, \texttt{specmatch-emp} \citep{2017ApJ...836...77Y} matched spectra better than the synthetic templates, but is still too rigid for some applications and requires the assembly of hundreds of standardized high signal-to-noise-ratio templates, ideally with low intrinsic rotational broadening.
-Such a large number of high-quality templates with high resolving power and low $v\sin{i}$ has not yet been established in the near infrared.
+Such a large number of high-quality templates with high resolving power and low $v\sin{i}$ has not yet been established in the near-infrared.
 
@@
 The tool requires dozens of high-SNR spectra acquired at a range of Barycentric Earth Radial Velocities (BERVs).
-The final telluric-free combined spectrum would still have to be compared to models for absolute calibration, or can be used out-of-the-box for precision relative RVs.
+The final telluric-free combined spectrum would still have to be compared to models for absolute calibration or can be used out-of-the-box for precision relative RVs.
 The \texttt{wobble} framework also pioneered the off-label application of automatic differentiation frameworks---in this case \texttt{TensorFlow}---towards their physically-motivated use in stellar spectra.
@@
 
-The \texttt{starfish} framework \citep{czekala15} provides a robust likelihood function for data-model comparisons, and retires many of the problems in this domain.
+The \texttt{starfish} framework \citep{czekala15} provides a robust likelihood function for data-model comparisons and retires many of the problems in this domain.
 \texttt{starfish} pioneered the use of whole-spectrum fitting with resilience to model imperfections by addressing the problem of what to do when the underlying atomic and molecular data was wrong or approximate or missing.
@@
 Here we enumerate some planned or conceivable extensions to the framework that would unlock new science arenas.
-We order the presentation of these extensions from our perceived ease-of-implementation, from easiest to hardest.
+We order the presentation of these extensions from our perceived ease of implementation, from easiest to hardest.
 
@@
 We have a prototype exact-Voigt-Hjertingimplementation following \citet{2022ApJS..258...31K}.
-We coarsely estimate that moving to this exact-Voigt implementation wculd decrease some residual regions by $\sim 30\%$, while increasing the computational cost by more than $10\times$ over the existing pseudo-Voigt approximation.
+We coarsely estimate that moving to this exact-Voigt implementation could decrease some residual regions by $\sim 30\%$, while increasing the computational cost by more than $10\times$ over the existing pseudo-Voigt approximation.
 The exact-Voigt-Hjerting implementation still outbids the higher cost of a direct numerical convolution of a Gaussian and Lorentzian profile.
@@
 We currently assume the input spectra are perfectly normalized to the continuum.
-This approximation impacts all training stages, including the stellar cloning, telluric cloning, and data-model comparison.
+This approximation impacts all training stages, including stellar cloning, telluric cloning, and data-model comparison.
 Gaussian Processes (GPs) could and should be brought to bear at each of these optimization stages, as described in the ``global kernels'' in \texttt{Starfish} \citet{czekala15}.
@@
 
-where we introduce the covariance matrix $\mathsf{C}$, with its concommittant kernel and collection of typically 2-3 parameters.
+where we introduce the covariance matrix $\mathsf{C}$, with its concomitant kernel and collection of typically 2-3 parameters.
 We anticipate that this GP likelihood would have the greatest impact on stellar stars with significant band-heads and line-blanketing: spectra with a so-called ``pseudo-continuum''.
@@
 The main demerit of moving to a GP-likelihood is computational cost.
-Fortunately a few efficient autodiff-aware implementations of GPs exist.
+Fortunately, a few efficient autodiff-aware implementations of GPs exist.
 The \texttt{celerite} algorithm \emph{celerit\'e} \citep{2017AJ....154..220F} has an exact backpropation implementation \citep{2018RNAAS...2...31F} that scales linearly with the number of data points.
-The celerite algorithm does not currently have a PyTorch implementation.
+The \emph{celerit\'e} algorithm does not currently have a PyTorch implementation.
 The \texttt{GPyTorch} framework \citep{2018arXiv180911165G} has a large category of approximate and exact GPs that could be straight-forwardly dropped into \texttt{blas\'e}.
@@
 
-Currently each training epoch ``sees'' the entire dataset, a setup dubbed ``full-batch'' gradient descent.
-An alternative scheme allow training with only a portion of the entire dataset at a time in ``minibatches''.
-The massive data volumes in modern Neural Network applications cannot fit into the GPU memory, and so minibatches are a necessity.
+Currently, each training epoch ``sees'' the entire dataset, a setup dubbed ``full-batch'' gradient descent.
+An alternative scheme allows training with only a portion of the entire dataset at a time in ``minibatches''.
+The massive data volumes in modern Neural Network applications cannot fit into the GPU memory, so minibatches are a necessity.
 Our meager $1$ MB dataset can easily fit into the GPU memory, but our model can be large if we have a large number of pixels or lines or both.
@@
 
-We experimented with minibatches by assembling and evaluate only a portion of the dense $\bar{\bm{F}}$ matrix at a time, in minibatches.
+We experimented with minibatches by assembling and evaluating only a portion of the dense $\bar{\bm{F}}$ matrix at a time, in minibatches.
 The choice to evaluate only a portion of lines at a time would mean the model is inaccurately evaluated at all wavelength pixels.
@@
 We then construct and evaluate the entire dense matrix for those lines: $\sim 330\;000 \times 20$.
-The number of FLOPS in each category scales as about 6 Million for the 20 broad and dense lines versus about 36 Million for the sea of about 6500 narrow and sparse lines, depending on the exact choices for wing cuts and number of lines.
+The number of FLOPS in each category scales as about 6 Million for the 20 broad and dense lines versus about 36 Million for the sea of about 6500 narrow and sparse lines, depending on the exact choices for wing cuts and the number of lines.
 
@@
 Extremely high signal-to-noise-ratio spectra of rapidly rotating stars may require additional flexibility.
-The limb darkening is generally wavelength dependent, and so a pan-chomatic spectrum may require a different limb-darkening from the blue end to the red end.
+The limb darkening is generally wavelength-dependent, and so a pan-chromatic spectrum may require a different limb darkening from the blue end to the red end.
 The limb darkening may instead depend on physical properties of the spectral line formation, such as physical depth of formation, and so the extent of limb-darkening may jump haphazardly from line-to-line-to-line, rather than as a predictably smooth function across wavelength.
@@
 
-Longitudinally symmetric surface features to not change as the star rotates on its axis.
-The existence of these features change the \emph{kurtosis} of the spectral line.
-For example a hypothetical non-emitting (black) polar starspot exhibits a deficit in flux at the line core, resulting in less zero-velocity flux than its homogeneous counterpart.
+Longitudinally symmetric surface features do not change as the star rotates on its axis.
+The existence of these features changes the \emph{kurtosis} of the spectral line.
+For example, a hypothetical non-emitting (black) polar starspot exhibits a deficit in flux at the line core, resulting in less zero-velocity flux than its homogeneous counterpart.
 A dark zonal band results in equal-sized bites out of the red- and blue- sides of the line.
@@
 Isolated deep spectral lines constitute the only practical scenario where exact knowledge can be claimed.
-Isolated spectral lines may be scarce or absent for M-dwarfs and brown dwarfs where lines blend together ostensibly in an inseparable way, counfounding Doppler imaging.
+Isolated spectral lines may be scarce or absent for M-dwarfs and brown dwarfs where lines blend ostensibly in an inseparable way, counfounding Doppler imaging.
 
@@
 As an example, the depth-of-formation for all or a subset of lines could be obtained and associated with each spectral line.
-A regulariztion scaling term could be introduced to allow the line positions to vary, but only in proportion to their depth of formation.
+A regularization scaling term could be introduced to allow the line positions to vary, but only in proportion to their depth of formation.
 
@@
 In principle, one could compute the Jacobians analytically---since the partial derivatives of the Voigt function are known in closed-form.
-This analytical approach would cut-out the autodiff approach, using the exact Jacobians to compute the Gradient Descent step.
+This analytical approach would cut out the autodiff approach, using the exact Jacobians to compute the Gradient Descent step.
 The analytical approach may be faster and/or possess lower memory overhead than the autodiff one.
-However, working out the chain-rule for the closed-form analytical Jacobians may become difficult or impossible once combined with the many augmentations get applied at the joint telluric-and-stellar modeling step.
-So analytic Jacobians could principally accelerate the cloning step, but would add significant implementation complexity at the important data-model comparison step.
-We therefore deem this extension somewhat diminishing returns on scarce developer time.
+However, working out the chain-rule for the closed-form analytical Jacobians may become difficult or impossible once combined with the many augmentations that get applied at the joint telluric-and-stellar modeling step.
+So analytic Jacobians could principally accelerate the cloning step but would add significant implementation complexity at the important data-model comparison step.
+We, therefore, deem this extension somewhat diminishing returns on scarce developer time.
 

### 2022-06-22 paper/paper1/ms.tex
Add a second figure of line density to Figure 1, might delete

@@
 \definecolor{asbestos}{RGB}{127, 140, 141}
-\hypersetup{linkcolor=nephritis,citecolor=belize,filecolor=asbestos,urlcolor=peter}
+\hypersetup{linkcolor=asbestos, 
+citecolor=belize,filecolor=asbestos,urlcolor=peter}
 
@@
     \centering
-    \includegraphics[width=0.5\textwidth]{figures/N_lines_vs_Teff_prom.png}
-    \caption{Number of prominent spectral lines versus effective temperature for PHOENIX models truncated to IGRINS (\emph{blue, connected} points) and HPF (\emph{orange, free-standing} points) bandwidths, for different prominence thresholds of $0.02$, $0.01$, and $0.005$.}
+    \includegraphics[width=0.45\textwidth]{figures/N_lines_vs_Teff_prom.png}
+    \includegraphics[width=0.45\textwidth]{figures/line_density_demo.png}
+    \caption{Scaling of prominence and density of spectral lines.  
+    \emph{Left:} Number of lines versus effective temperature for PHOENIX models truncated to IGRINS (\textbf{blue, connected} points) and HPF (\emph{red, free-standing} points) bandwidths, for different prominence thresholds of $0.02$, $0.01$, and $0.005$.
+    \textbf{Right:} Number of lines per 100 \AA\ wavelength bin for stellar (\emph{blue, upper envelope of steps}) and telluric (\emph{orange, lower envelope of steps}), illustrated for a $T_\mathrm{eff}=4700$~K, $\log{g}=4.5$ PHOENIX model and a $T=290 \mathrm{K}$, relative humidity of 40\% TelFit model.}
     \label{fig_Nlines_vs_teff}
@@
 We see an RMS residual of 1.2\%/pixel at native resolution.
+\authorcomment1{Need more prose here.}
+
 We identify three main categories of cloning flaws.
@@
 Cloning telluric lines suffers from one additional problem.
-Telluric lines can be extremely deep, exhibiting almost vanishing transmission with saturated line cores common in-between the atmospheric windows that define the $I$, $J$, $H$, and $K$- bands.
-The \texttt{blas\'e} method cannot cope with these saturated lines, and often it treats nearby blended saturated lines as one single line.
-The underfit model flounders.
+Telluric lines can be extremely deep, exhibiting almost vanishing transmission with saturated line cores common in-between the atmospheric windows that define the $I-$, $J-$, $H-$, and $K-$ bands.
+The \texttt{blas\'e} method can cope with these saturated lines, but often it treats nearby-and-blended saturated lines as one single line.
+This glomming together of lines has little practical effect since few astronomical practitioners can make use of such profoundly saturated data.
 
@@
 
-Here we illustrate how \texttt{blas'e} gets altered when applying the logarithmic flux pre-processing step.
+Here we illustrate how \texttt{blas\'e} gets altered when applying the logarithmic flux pre-processing step.
 First, we compute the natural log of the flux directly on the precomputed synthetic spectrum in its absolute flux scaling and native pixel sampling:

### 2022-06-22 paper/paper1/ms.tex
Add regularization matrh

@@
     \includegraphics[width=0.45\textwidth]{figures/N_lines_vs_Teff_prom.png}
-    \includegraphics[width=0.45\textwidth]{figures/line_density_demo.png}
+    \includegraphics[width=0.45\textwidth]{figures/N_lines_vs_Teff_prom.png}
+    %\includegraphics[width=0.45\textwidth]{figures/line_density_demo.png}
     \caption{Scaling of prominence and density of spectral lines.
@@
 
-Equation \ref{eqn_final_model} has up to $N_{\mathrm{lines}}\times3 \sim20,000$ tunabale parameters from the star, $N_{\mathrm{lines}}\times3 \sim15,000$ tunabale parameters from the Earth's atmosphere, plus $v\sin{i}$ and $RV$.
+Equation \ref{eqn_final_model} has up to $N_{\mathrm{lines}}\times3 \sim20,000$ tunable parameters from the star, $N_{\mathrm{lines}}\times3 \sim15,000$ tunable parameters from the Earth's atmosphere, plus $v\sin{i}$ and $RV$.
 That adds up to 35,002 model parameters.
@@
 We default to the following choice.
-We fix all line parameters except for amplitude, which receives a regularization of 0.01, meaning that its amplitude should be within about 1\% of its cloned value.
+We usually fix all line parameters except for amplitude, which receives a log-normal distribution:
+
+\begin{eqnarray}
+    \mathrm{regularization} \equiv \ln{\mathrm{prior}}  = \frac{1}{2} \sum_{j=1}^{N_\mathrm{lines}} \Big(\frac{\ln{a_j} - \ln{a^\prime_j }}{\Lambda}\Big)^2
+\end{eqnarray}
+
+where the prime notation demarcates the amplitudes obtained from cloning, or the ``initial amplitude''.
+We assign $\Lambda_{\ln{a}}\sim5$.
+This weak degree of regularization has the effect of permitting refinement of only the most conspicuous data-model mismatches; the weakest lines don't both to move from their initial state because doing so would penalize the regularization penalty without enough reduction of the overall loss.
 The extrinsic $v\sin{i}$ and $RV$ have no regularization, but in practice, they need to be initialized close to their plausible values.

### 2022-06-23 paper/paper1/ms.tex
Some formatting-- figure placement in two column mode

@@
-\documentclass[modern]{aastex631}
+\documentclass[twocolumn]{aastex631}
 \usepackage{hyperref}
@@
 citecolor=belize,filecolor=asbestos,urlcolor=peter}
+
+\hyphenation{long-it-udin-ally hy-po-thet-ical de-mon-stra-tion}
 
@@
 
+\begin{figure}[hbt!]
+    \centering
+    \includegraphics[width=0.45\textwidth]{figures/N_lines_vs_Teff_prom.png}
+    %\includegraphics[width=0.45\textwidth]{figures/N_lines_vs_Teff_prom.png}
+    \includegraphics[width=0.45\textwidth]{figures/line_density_demo.png}
+    \caption{Scaling of prominence and density of spectral lines.
+        \emph{Left:} Number of lines versus effective temperature for PHOENIX models truncated to IGRINS (\textbf{blue, connected} points) and HPF (\emph{red, free-standing} points) bandwidths, for different prominence thresholds of $0.02$, $0.01$, and $0.005$.
+        \textbf{Right:} Number of lines per 100 \AA\ wavelength bin for stellar (\emph{blue, upper envelope of steps}) and telluric (\emph{orange, lower envelope of steps}), illustrated for a $T_\mathrm{eff}=4700$~K, $\log{g}=4.5$ PHOENIX model and a $T=290 \mathrm{K}$, relative humidity of 40\% TelFit model.}
+    \label{fig_Nlines_vs_teff}
+\end{figure}
+
 This high-pass filtering step should be set to capture the genuine spectral shape, without over-fitting broad line wings such as those in deep Hydrogen and sodium lines.
@@
 HPF may have up to $2048\times28=$57,344 pixels, and IGRINS has typically about 75,000 pixels, after common trimming of noisy edge pixels and unusable telluric regions.
-
-\begin{figure}[hbt!]
-    \centering
-    \includegraphics[width=0.45\textwidth]{figures/N_lines_vs_Teff_prom.png}
-    \includegraphics[width=0.45\textwidth]{figures/N_lines_vs_Teff_prom.png}
-    %\includegraphics[width=0.45\textwidth]{figures/line_density_demo.png}
-    \caption{Scaling of prominence and density of spectral lines.
-        \emph{Left:} Number of lines versus effective temperature for PHOENIX models truncated to IGRINS (\textbf{blue, connected} points) and HPF (\emph{red, free-standing} points) bandwidths, for different prominence thresholds of $0.02$, $0.01$, and $0.005$.
-        \textbf{Right:} Number of lines per 100 \AA\ wavelength bin for stellar (\emph{blue, upper envelope of steps}) and telluric (\emph{orange, lower envelope of steps}), illustrated for a $T_\mathrm{eff}=4700$~K, $\log{g}=4.5$ PHOENIX model and a $T=290 \mathrm{K}$, relative humidity of 40\% TelFit model.}
-    \label{fig_Nlines_vs_teff}
-\end{figure}
 
@@
 
-\begin{figure}[hbt!]
+\begin{figure*}[hbt!]
     \centering
@@
     \label{blase_flowchart}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{equation} \bm{\bar{F}} =
+\begin{equation} 
+    \begin{aligned}
+    \bm{\bar{F}} &=
     \begin{pmatrix}
@@
         \cdots \includegraphics[height=1cm]{figures/voigt_03.pdf} \cdots  &
-    \end{pmatrix} \mapsto
-    \begin{pmatrix}
+    \end{pmatrix} \\
+    \bm{\bar{F}}&\mapsto \bm{\hat{F}} =\begin{pmatrix}
         \includegraphics[height=1cm]{figures/voigt_01b.pdf} & \\
@@
         \includegraphics[height=1cm]{figures/voigt_03b.pdf} &
-    \end{pmatrix} = \bm{\hat{F}}
+    \end{pmatrix}
+\end{aligned}
 \end{equation}
@@
     \label{fig_cloned_spectrum_demo}
-\end{figure}
+\end{figure*}
 
@@
     \label{fig_zoom_cloning_performance}
-\end{figure}
+\end{figure*}
 
@@
 \begin{eqnarray}
-    \mathrm{regularization} \equiv \ln{\mathrm{prior}}  = \frac{1}{2} \sum_{j=1}^{N_\mathrm{lines}} \Big(\frac{\ln{a_j} - \ln{a^\prime_j }}{\Lambda}\Big)^2
+    \mathrm{regularization} \equiv -\ln{\mathrm{prior}}  = \frac{1}{2} \sum_{j=1}^{N_\mathrm{lines}} \Big(\frac{\ln{a_j} - \ln{a^\prime_j }}{\Lambda}\Big)^2
 \end{eqnarray}
@@
     \label{fig_WASP69_demo}
-\end{figure}
+\end{figure*}
 
-\begin{figure}[hbt!]
+\begin{figure*}[hbt!]
     \centering
@@
     \label{fig_WASP69_transferred}
-\end{figure}
+\end{figure*}
 
-\begin{figure}[hbt!]
+\begin{figure*}[hbt!]
     \centering
@@
     \label{fig_WASP69_regularized}
-\end{figure}
+\end{figure*}
 
@@
 
-Currently, each training epoch ``sees'' the entire dataset, a setup dubbed ``full-batch'' gradient descent.
-An alternative scheme allows training with only a portion of the entire dataset at a time in ``minibatches''.
+Currently, each training epoch sees the entire dataset, a setup dubbed \emph{full-batch} gradient descent.  
+An alternative scheme allows training with only a portion of the entire dataset at a time in \emph{minibatches}.
 The massive data volumes in modern Neural Network applications cannot fit into the GPU memory, so minibatches are a necessity.
-Our meager $1$ MB dataset can easily fit into the GPU memory, but our model can be large if we have a large number of pixels or lines or both.
+Our meager 1 MB dataset can easily fit into the GPU memory, but our model can be large if we have a large number of pixels or lines or both.
 So while minibatches may not be required due to data size limitations they may be useful for particularly large models.  
@@
 
-It is much easier to reverse-engineer longitudinally asymmetric features than longitudinally symmetric ones.
+It is easier to reverse-engineer longitudinally asymmetric features than longitudinally symmetric ones.  
 The latter requires exact knowledge of the underlying spectral template.
@@
     \1 Computational tradeoff of Lorentzian/Gaussian versus Pseudo-Voigt versus Voigt
-    \2 Additional Implementation challenge: Fadeeva function not implemented in PyTorch
+    \2 Additional challenge: Fadeeva function not in PyTorch
     \1 We choose a pseudo-Voigt as a balance between computation and adequate accuracy

### 2022-06-23 paper/paper1/ms.tex
Looking nice, about to print for some revisions

@@
 \usepackage{amsmath,bm}
+
 \definecolor{belize}{RGB}{41, 128, 185}
@@
 
-\hyphenation{long-it-udin-ally hy-po-thet-ical de-mon-stra-tion}
+\hyphenation{long-it-udin-ally hy-po-thet-ical de-mon-stra-tion Ad-min-is-tra-tion}
 
@@
     %\includegraphics[width=0.45\textwidth]{figures/N_lines_vs_Teff_prom.png}
-    \includegraphics[width=0.45\textwidth]{figures/line_density_demo.png}
+    \includegraphics[width=0.41\textwidth]{figures/line_density_demo.png}
     \caption{Scaling of prominence and density of spectral lines.
@@
     \centering
-    \includegraphics[width=0.8\textwidth]{figures/blase_clone_demo.png}
+    \includegraphics[width=0.6\textwidth]{figures/blase_clone_demo.png}
     \caption{PHOENIX spectrum cloned with blas\'e.
@@
 
-\begin{figure*}[hbt!]
+\begin{figure}[hbt!]
     \centering
-    \includegraphics[width=0.8\textwidth]{figures/blase_clone_demo_zoom.png}
+    \includegraphics[width=0.48\textwidth]{figures/blase_clone_demo_zoom.png}
     \caption{Zoom in of the region between the vertical gray bars in Figure \ref{fig_cloned_spectrum_demo}.
@@
     \label{fig_zoom_cloning_performance}
-\end{figure*}
+\end{figure}
 
@@
 \begin{eqnarray}
-    \mathrm{regularization} \equiv -\ln{\mathrm{prior}}  = \frac{1}{2} \sum_{j=1}^{N_\mathrm{lines}} \Big(\frac{\ln{a_j} - \ln{a^\prime_j }}{\Lambda}\Big)^2
+    \mathrm{regularization} &\equiv& -\ln{\mathrm{prior}} \nonumber \\ 
+    &=& \frac{1}{2} \sum_{j=1}^{N_\mathrm{lines}} \Big(\frac{\ln{a_j} - \ln{a^\prime_j }}{\Lambda}\Big)^2
 \end{eqnarray}
@@
     \centering
-    \includegraphics[width=0.8\textwidth]{figures/blase_WASP69_demo.png}
+    \includegraphics[width=0.6\textwidth]{figures/blase_WASP69_demo.png}
     \caption{WASP 69 observed with HPF, compared to a $T_{\mathrm{eff}}=4700\;$K,  $\log{g}=4.5$ solar metallicity PHOENIX model warped to $v\sin{i}=2.2$~km/s, $RV=-9.6$~km/s, and HPF resolving power.}
@@
     \centering
-    \includegraphics[width=0.8\textwidth]{figures/blase_WASP69_regularized.png}
+    \includegraphics[width=0.6\textwidth]{figures/blase_WASP69_regularized.png}
     \caption{Semi-empirical model of WASP 69 transfer-learned with \texttt{blas\'e}, employing a regularization prior on the learned amplitudes.}
@@
     \centering
-    \includegraphics[width=0.8\textwidth]{figures/blase_super_resolution_template.png}
+    \includegraphics[width=0.6\textwidth]{figures/blase_super_resolution_template.png}
     \caption{The native-resolution semi-empirical model transfer-learned from WASP 69 HPF data. The revised model can be viewed as a super-resolution deconvolution of the HPF spectrum.}
@@
 \subsection{Prospects for closed-form computation of Jacobians}
-In principle, one could compute the Jacobians analytically---since the partial derivatives of the Voigt function are known in closed-form.
+In principle, one could compute the Jacobians analytically since the partial derivatives of the Voigt function are known in closed-form.
 This analytical approach would cut out the autodiff approach, using the exact Jacobians to compute the Gradient Descent step.
@@
 
-\
+\begin{acknowledgments}
 
-\begin{acknowledgements}
-    This material is based upon work supported by the National Aeronautics and Space Administration under Grant Numbers 80NSSC21K0650 for the NNH20ZDA001N-ADAP:D.2 program, and 80NSSC20K0257 for the XRP program issued through the Science Mission Directorate.
-We acknowledge the National Science Foundation, which supported the work presented here under Grant No.
-1910969.
+These results are based on observations obtained with the Habitable-zone Planet Finder Spectrograph on the HET.
+The HPF team was supported by NSF grants AST-1006676, AST-1126413, AST-1310885, AST-1517592, AST-1310875, AST-1910954, AST-1907622, AST-1909506, ATI 2009889, ATI-2009982, and the NASA Astrobiology Institute (NNA09DA76A) in the pursuit of precision radial velocities in the NIR.
+The HPF team was also supported by the Heising-Simons Foundation via grant 2017-0494.
+
+The Hobby-Eberly Telescope (HET) is a joint project of the University of Texas at Austin, the Pennsylvania State University, Ludwig-Maximilians-Universit\"at M\"unchen, and Georg-August-Universit\"at G\"ottingen.
+The HET is named in honor of its principal benefactors, William P.
+Hobby and Robert E.
+Eberly.
+
+This material is based upon work supported by the National Aeronautics and Space Administration under Grant Numbers 80NSSC21K0650 for the NNH20ZDA001N-ADAP:D.2 program,
+and 80NSSC20K0257 for the XRP program issued through the Science Mission Directorate.
+
+We acknowledge the National Science Foundation, which supported the work presented here under Grant No. 1910969.
+
 This research has made use of NASA's Astrophysics Data System Bibliographic Services.
-\end{acknowledgements}
 
-\clearpage
+\end{acknowledgments}
 

### 2022-06-23 setup.py
bump version to v0.3, yay!

@@ -3,7 +3,7 @@ import setuptools
 
 setuptools.setup(
     name="blase",
-    version="0.2",
+    version="0.3",
     author="gully",
     author_email="igully@gmail.com",
     description="Forward Modeling echelle spectra with PyTorch",

### 2022-06-23 setup.py
Add some ostensible dependencies, may want to revisit though

@@ -10,7 +10,7 @@ setuptools.setup(
     long_description="Hybrid data/model-driven forward modeling of echelle spectra",
     long_description_content_type="text/markdown",
     url="https://github.com/gully/blase",
-    install_requires=["numpy", "scipy", "torch"],
+    install_requires=["numpy", "scipy", "torch", "tqdm", "kornia"],
     packages=setuptools.find_packages(where="src"),
     package_dir={"": "src"},
     classifiers=[

### 2022-08-10 paper/paper1/ms.tex
Focus on science in the discussion section, technical stuff to appendix

@@
 \usepackage{censor}
+\usepackage[arrowdel]{physics}
 \usepackage{mathtools}
 \usepackage{outlines}
+\usepackage{lipsum}  
 \usepackage{amsmath,bm}
@@
 \definecolor{asbestos}{RGB}{127, 140, 141}
-\hypersetup{linkcolor=asbestos, 
-citecolor=belize,filecolor=asbestos,urlcolor=peter}
+\definecolor{clouds}{RGB}{236, 240, 241}
+
+\hypersetup{linkcolor=belize, citecolor=belize,filecolor=asbestos,urlcolor=peter}
+
+%\setlipsum{
+%  par-before = \begingroup\color{clouds},
+%  par-after = \endgroup
+%}
+%\SetLipsumParListSurrounders{\begingroup\color{gray}}{\endgroup}
 
@@
 \begin{document}
-\shorttitle{Blas\'e}
+\shorttitle{Transfer learning with blas\'e}
 \shortauthors{Gully \& Morley}
-\title{Transfer learning for \'echelle spectroscopy with \texttt{blas\'e}}
+\title{Automatic line-by-line semi-empirical models for high-grasp stellar and telluric spectra}
 
@@
 Telluric artifacts, imperfect line lists, inexact continuum placement, and inflexible models frustrate the scientific promise of these information-rich datasets.
-Here we debut an interpretable machine-learning framework \emph{blas\'e} that addresses these challenges.
+Here we debut an interpretable machine-learning framework \emph{blas\'e} that addresses these and other challenges.
 The semi-empirical approach can be viewed as ``transfer learning''--first pre-training models on noise-free precomputed synthetic spectral models, then learning the corrections to line depths and widths from whole-spectrum fitting to an observed spectrum.
-The auto-differentiable model employs back-propagation, the fundamental algorithm empowering modern Neural Networks.
+The auto-differentiable model employs back-propagation, the fundamental algorithm empowering modern Deep Learning and Neural Networks.
 Here, however, the 40,000+ parameters symbolize physically interpretable line profile properties such as amplitude, width, location, and shape, plus radial velocity and rotational broadening.
@@
 Its sparse-matrix architecture and GPU-acceleration make \emph{blas\'e} fast, with end-to-end training of an octave of bandwidth (50+ \'echelle orders) in under 1 minute.
-The open-source PyTorch-based code \texttt{blase} includes tutorials and API documentation.
+The open-source PyTorch-based code \texttt{blase} includes tutorials, deep dives, reproducible Python environments, and API documentation.
 We show how the tool fits into the existing Python spectroscopy ecosystem, demonstrate a range of astrophysical applications, and discuss limitations and future extensions.
@@
 The identification and understanding of these lines have defined a large category of astrophysics over the last century.
-The field grew from by-eye catalogs of stellar templates \citep{1901AnHar..28..129C} to quantifying the role of atomic ionization balance \citep{1925PhDT.........1P}, to modern synthetic forward models including millions or billions of lines \citep{husser13, 2021ApJ...920...85M}.
+The field grew from by-eye catalogs of stellar templates \citep{1901AnHar..28..129C} to quantifying the role of atomic ionization balance \citep{1925PhDT.........1P}, to modern synthetic forward models including millions or billions of lines \citep[\emph{e.g.}][]{husser13, 2021ApJ...920...85M}.
 As technology has improved, our data and models have become more voluminous, more precise, and more complicated.
@@
 
-The metaphorical holy grail of astronomical spectroscopy is a function that takes in an observed astronomical spectrum and reports back the position, amplitude, width, and shape of all of its spectral lines, automatically, accurately, and precisely.
+The metaphorical holy grail of astronomical spectroscopy is a function that takes in an observed stellar spectrum and reports back the position, amplitude, width, and shape of all of its spectral lines, automatically, accurately, and precisely.
 The function would go further.
-It would report back the systemic radial velocity (RV) and rotational broadening ($v\sin{i}$).
-Finally, the function would---in an \emph{Inception}-like manner---provide what it believes to be the interpretable generating function that produced the data, so that we may gain insights on future examples of this or other stars.
+It would report back the systemic radial velocity ($RV$) and rotational broadening ($v\sin{i}$)and optionally fundamental stellar properties, including $T_{\mathrm{eff}}$, $\log{g}$, and $[\mathrm{Fe}/\mathrm{H}]$.
+Finally, the function would---in a feat of artificial intelligence---provide what it believes to be the interpretable generating function that produced this data in the first place, so that we may gain insights on future examples of this or other stars.
 Solving this problem is hard, for at least four reasons.
@@
 Many practitioners today simply mask these telluric regions, yet micro-tellurics still inject variance into the spectrum that cannot be easily accounted for with existing methods.
-A robust accounting of telluric absorption at the $\mathrm{cm/s}$ level may require joint modeling of the star and the Earth's atmospheric absorption \emph{before} convolution with an instrumental kernel.
+Instead, a robust accounting of telluric absorption at the $\mathrm{cm/s}$ level may require joint modeling of the star and the Earth's atmospheric absorption \emph{before} convolution with an instrumental kernel.
 This telluric joint modeling capability does not yet exist at a precision that can meet these strenuous demands.
 
-In the case of Doppler imaging, an accurate underlying template is needed to detect longitudinally symmetric structures \citep{1983PASP...95..565V,2021arXiv211006271L} such as polar spots \citep{roettenbacher16} or zonal bands \citep{Crossfield14,2021ApJ...906...64A}.
+In the case of Doppler imaging, an accurate underlying spectral template is needed to detect longitudinally symmetric structures \citep{1983PASP...95..565V,2021arXiv211006271L} such as polar spots \citep{roettenbacher16} or zonal bands \citep{Crossfield14,2021ApJ...906...64A}.
+There exists a nearly circular reasoning: we need to know the extent of line profile perturbations to reveal the underlying spectral template, but we need the underlying spectral template to estimate the line profile perturbations.
+The approach introduced here offers a path forward.
+
+\subsection{Automatic differentiation technology}
 
@@
 These frameworks include \texttt{ROBOSPECT} \citep{2013PASP..125.1164W}, \texttt{specmatch} \citep{2015PhDT........82P}, \texttt{specmatch-emp} \citep{2017ApJ...836...77Y}, \texttt{wobble} \citep{2019AJ....158..164B}, \texttt{starfish} \citep{czekala15}, \texttt{sick} \citep{2016ApJS..223....8C}, \texttt{psoap} \citep{2017ApJ...840...49C}, \texttt{FAL} (Cargile et al.
-\emph{in prep}), CHIMERA \citep{2015ApJ...807..183L}, the \texttt{Cannon} \citep{2017ApJ...836....5H},  \texttt{MINESweeper} \citep{2020ApJ...900...28C}, and recently \texttt{ExoJAX} \citep{2022ApJS..258...31K}.
+\emph{in prep}), CHIMERA \citep{2015ApJ...807..183L}, the \texttt{Cannon} \citep{2017ApJ...836....5H},  \texttt{MOOG} \citep{2012ascl.soft02009S}, \texttt{MOOGStokes} \citep{2013AJ....146...51D}, \texttt{MINESweeper} \citep{2020ApJ...900...28C}, and recently \texttt{ExoJAX} \citep{2022ApJS..258...31K}.
 The designs of these frameworks necessarily have to make a choice in the bias-variance tradeoff: is the tool more \emph{data}-driven or more \emph{model}-driven?
@@
 A key new enabling technology breaks these classical tradeoffs in data-model comparisons for astronomical spectroscopy.
-Automatic differentiation \citep[``autodiff'',][]{2015arXiv150205767G} and its affiliated backpropagation algorithm has revolutionized machine learning and neural network architecture design, and is increasingly applied in astrophysical data analysis contexts, \emph{e.g.} kernel phase coronography with \texttt{poppy} \citep{2021ApJ...907...40P}, and exoplanet orbit fitting with \texttt{exoplanet} \citep{2021JOSS....6.3285F}.
+Automatic differentiation \citep[``autodiff'' or ``autograd'',][]{2015arXiv150205767G, 2016PhDT.......317M} and its affiliated backpropagation algorithm \citep{kelley1960,Linnainmaa1976TaylorEO, 1986Natur.323..533R, 1990JGCD...13..926D} has revolutionized machine learning and neural network architecture design, and is increasingly applied in astrophysical data analysis contexts, \emph{e.g.} kernel phase coronography with \texttt{poppy} \citep{2021ApJ...907...40P}, and exoplanet orbit fitting with \texttt{exoplanet} \citep{2021JOSS....6.3285F}.
 Of the spectroscopy frameworks mentioned above, the \texttt{TensorFlow}-based \citep{tensorflow2015-whitepaper} \texttt{wobble} and the \texttt{JAX}-based \citep{jax2018github} \texttt{ExoJAX} employ autodiff technology.
@@
 In this paper, we show that autodiff-aware semi-empirical models offer an appealing middle ground: informed by self-consistent models but refined with data.
-In Section \ref{methodology} we introduce the \emph{blas\'e} forward-model design and its \texttt{PyTorch}-based \citep{2019arXiv191201703P} implementation.
-We show how to supplant the computationally expensive radiative transfer by cloning pre-computed synthetic spectral models.
-We show how to combine both stellar and telluric models in a joint modeling approach.
-In Section \ref{transferLearn} we describe how to fine-adjust the cloned models to real-world data from high-grasp near-IR spectrographs, using a transfer-learning step.
-%Section \ref{secInjRec} exhibits injection-recovery tests on simulated data of known properties, revealing the scalability and fundamental limits of the technique.  
-Finally, we discuss many conceivable extensions to the framework and directions for future work in Section \ref{secDiscuss}.
+This sweet spot in the bias-variance tradoff can be thought of as a hybrid data-and-model driven approach.
+The algorithm presented here focuses on modeling the spectra of stars and brown dwarfs.
+In Section \ref{methodology} we introduce the interpretable forward-model design and its \texttt{PyTorch}-based \citep{2019arXiv191201703P} implementation, \texttt{blas\'e}.
+Existing stellar models \citep[\emph{e.g.}][]{husser13} and substellar models \citep[\emph{e.g.}][]{2021ApJ...920...85M} laboriously solve for a self consistent thermal structure in the atmosphere given the copious opacity sources that themselves depend on temperture and pressure.
+Here we build upon that hard work by cloning pre-existing synthetic stellar or substellar models (Section \ref{methodology}), and optionally by cloning models of Earth's atmospheric ``telluric'' absorption (Section \ref{sectionTelluric}).
+In Section \ref{transferLearn}  we describe how to adapt both stellar and telluric cloned models simultaneously, using a transfer-learning step.
+We obtain semi-empirical models by comparing to real-world \'echelle data in Section \ref{secResults}.
+Finally, we compare the framework to alternatives (Section \ref{secDiscuss}) and chronicle many conceivable extensions for future work (Section \ref{secFutureWork}).
 
@@
 
-We start with a high resolution pre-computed synthetic stellar model spectrum, $\mathsf{S}_{\rm abs}(\bm{\lambda})$ at its native resolution sampling and with its original absolute physical flux units.
+We start with a high resolution pre-computed synthetic stellar or substellar model spectrum, $\mathsf{S}_{\rm abs}(\bm{\lambda})$ at its native resolution sampling and with its original absolute physical flux units.
 The procedure that follows is largely agnostic to the exact details of how this spectrum was made, or what physics or chemistry it may represent.
-For the purposes of this paper, we will showcase examples from two well-known families of precomputed synthetic astronomical spectra: \texttt{PHOENIX} \citep{husser13} for stellar spectra $(T_{\mathrm{eff}}\in [2300, 10000]\;K)$ and \texttt{Sonora} \citep{2021ApJ...920...85M} for brown dwarf spectra $(T_{\mathrm{eff}}\in [1000, 2300]\;K)$.
-The framework may also work for precomputed synthetic spectra of exoplanets, supernovae, galaxies, or even further afield such as laboratory physical chemistry.
-We place the following demands on the precomputed spectra.
-They should have sporadic regions of discernable continuum devoid of lines, and the continuum must vary smoothly in wavelength.
-The spectral lines or pseudo-lines should be resolved, and not sub-sampled.
+For the purposes of this paper, we will showcase examples from two well-known families of precomputed synthetic astronomical spectra: \texttt{PHOENIX} \citep{husser13} for stellar spectra $(T_{\mathrm{eff}}\in [2300, 10000]\;K)$ and \texttt{Sonora} \citep{2021ApJ...920...85M} for substellar spectra $(T_{\mathrm{eff}}\in [200, 2300]\;K)$.
+The algorithms in the framework may also work for precomputed synthetic spectra of reflected light exoplanets, supernovae, galaxies, or even further afield such as laboratory physical chemistry, plasma physics, or materials science.
+
+We place mild demands on the precomputed spectra.
+They should have sporadic regions of discernable continuum devoid of lines, and the continuum should vary smoothly in wavelength.
+The spectral lines or pseudo-lines should be resolved (and not sub-sampled).
 We suspect most stellar spectra meet these criteria, except for the coolest M dwarfs and brown dwarfs.
-The method can hypothetically handle spectra without discernable continua, but we suspect some modifications to the preprocessing steps would be needed.
+The method can handle spectra without discernable continua, with some fine-adjustments to the preprocessing steps.
 We truncate the red and blue limits of the precomputed synthetic spectrum to match a high-bandwidth echelle spectrograph, extended with a buffer at the edges of size $\pm \Delta \lambda_{\mathrm{buffer}}$, chosen to account for plausible radial velocity and rotational broadening of real stars.
@@
 We have implemented both modes in \texttt{blas\'e}, allowing users to choose their preference.
-We only narrate the linear flux unit description in the main text of this document, for the sake of clarity, and since most practitioners tend to think of flux in terms of linear flux units.
+We only narrate the linear flux unit description in the main text of this document, for the sake of clarity, and since most practitioners may tend to think of flux in terms of linear flux units.
 The data-model comparison step will always take place in linear flux units, so the only operational difference is the behavior for deep and saturated lines.
@@
 
-\subsection{Initialization}
+\subsection{Initialization}\label{subsecInit}
 
@@
 We divide the entire spectrum by a black body $\mathsf{B}(\bm{\lambda}_S)$ of the same effective temperature $T_{\mathrm{eff}}$ as the model template.
-The resulting signal typically possesses a smooth continuum variation.
-An additional continuum flattening step ensures that subsequent spectral line finding steps get applied uniformly.
-Here, we identify the continuum regions by the most prominent peaks in the spectrum separated by a distance of at least 5,000$-$90,000 pixels from adjacent peaks, yielding between 60 and 5 control points.
-We then fit an $n_{\mathrm{poly}} = 3^{rd}-15^{th}$ order polynomial to these peaks, evaluate the polynomial on the entire bandwidth $\mathsf{P}(\bm{\lambda}_S)$, and divide the spectrum by this trend.
+The resulting signal usually still has smooth wiggles around the continuum.
+An optional continuum flattening step ensures that subsequent spectral line finding steps get applied uniformly.
+This high-pass filtering step should be set to capture the genuine spectral shape, without over-fitting broad line wings such as those in deep Hydrogen and sodium lines.
+Any high-pass filter will work, a Gaussian Process approach would be ideal.
+Instead we apply a simple and familiar heuristic: fit a polynomial $\mathsf{P}(\bm{\lambda}_S)$ to a few continuum peaks and divide out the trend.
+
 The result should be a flattened ``continuum-normalized'' spectrum familiar to practitioners in high-resolution spectroscopy, with the continuum level close to unity.
@@
 where the division indicates element-wise division of these arrays or ``vectors'' of flux values.
-Visual inspection of this continuum-flattened spectrum typically possesses variations at the $<1\%$ level.
-Tuning the pixel distance and polynomial order can lower the undulations to $<0.05\%$ level for some input spectra.
-We found a pixel distance of 50,000 and $3^{rd}$ order polynomial worked across the range of effective temperatures for HPF's bandwidth.
-The IGRINS spectra had more significant continuum opacity and band heads to accommodate and therefore needed to capture finer-scale variation: 5,000 pixels and a $9^{th}$ order polynomial appeared adequate over most of the effective temperature range.
 
@@
     \caption{Scaling of prominence and density of spectral lines.
-        \emph{Left:} Number of lines versus effective temperature for PHOENIX models truncated to IGRINS (\textbf{blue, connected} points) and HPF (\emph{red, free-standing} points) bandwidths, for different prominence thresholds of $0.02$, $0.01$, and $0.005$.
-        \textbf{Right:} Number of lines per 100 \AA\ wavelength bin for stellar (\emph{blue, upper envelope of steps}) and telluric (\emph{orange, lower envelope of steps}), illustrated for a $T_\mathrm{eff}=4700$~K, $\log{g}=4.5$ PHOENIX model and a $T=290 \mathrm{K}$, relative humidity of 40\% TelFit model.}
+        \emph{Top:} Number of lines versus effective temperature for PHOENIX models truncated to IGRINS (\textbf{blue, connected} points) and HPF (\emph{red, free-standing} points) bandwidths, for different prominence thresholds of $0.02$, $0.01$, and $0.005$.
+        \emph{Bottom:} Number of lines per 100 \AA\ wavelength bin for stellar (\emph{blue, upper envelope of steps}) and telluric (\emph{orange, lower envelope of steps}), illustrated for a $T_\mathrm{eff}=4700$~K, $\log{g}=4.5$ PHOENIX model and a $T=290 \mathrm{K}$, relative humidity of 40\% TelFit model.}
     \label{fig_Nlines_vs_teff}
@@
 
-This high-pass filtering step should be set to capture the genuine spectral shape, without over-fitting broad line wings such as those in deep Hydrogen and sodium lines.
-We have experimentally found it helpful to hand-tune models of exceptionally broad and non-standard line wings with an optional pre-processing step that depends on the wavelength range of interest.
-Here we sketch the approach for common broad lines found in the long-wavelength visible and near-infrared wavelength regions.
-
 % TODO We pre-assign spectral lines for the conspicuous Hydrogen lines and neutral metal lines. Table
 
-We emphasize that a recreation of the unvarnished input spectrum can be obtained by multiplying the continuum-flattened signal by the ``perturbed black body'', $\mathsf{B}(\bm{\lambda}_S)\odot \mathsf{P}(\bm{\lambda}_S)$, that symbolizes the black body modulated by continuum opacity or broad-band radiative transfer effects.
+We emphasize that a recreation of the unvarnished input spectrum---if desired---can be obtained by multiplying the continuum-flattened signal by the ``perturbed black body'', $\mathsf{B}(\bm{\lambda}_S)\odot \mathsf{P}(\bm{\lambda}_S)$, that symbolizes the black body modulated by continuum opacity or broad-band radiative transfer effects.
 This smooth spectrum may be useful for applications that need to keep track of broad-band flux, such as low-resolution spectra, or regions with molecular band heads.
 The ``perturbed black body'' continuum model contains $n_{\mathrm{poly}}+1$ fixed-but-possibly tunable lookup parameters, plus the fixed input $T_{\mathrm{eff}}$.
+For most practitioners these terms serve as nuisance parameters and are perfunctorily discarded.
 
 Next, we identify the spectral lines.
-We apply a local-minimum-finding algorithm on the spectrum by defining a prominence threshold $P_{\rm rom} \in (0.005, 0.02)$.
+We apply a local-minimum-finding algorithm \citep{2020SciPy-NMeth} on the spectrum by defining a prominence threshold $P_{\rm rom} \in (0.005, 0.02)$.
 This threshold dictates the number of lines that will be modeled: a lower prominence finds more, weaker lines, and a larger prominence finds fewer, deeper lines.
@@
 
-We have now arrived at the \texttt{blas\'e} clone model $\mathsf{S}_{\rm clone}(\bf{\lambda}_s)$ for a flattened synthetic spectrum $\mathsf{S}$: it is the cumulative product of transmission through the sea of all overlapping spectral lines:
+We have now arrived at the \texttt{blas\'e} clone model $\mathsf{S}_{\rm clone}(\bm{\lambda}_S)$ for a flattened synthetic spectrum $\mathsf{S}$: it is the cumulative product of transmission through the sea of all overlapping spectral lines:
 
@@
     \centering
-    \includegraphics[width=0.98\textwidth]{figures/blase_flowchart01.png}
-    \caption{Visual flowchart of \texttt{blas\'e}.
-The technique starts with a precomputed synthetic stellar spectrum and a precomputed synthetic telluric spectrum with properties close to those previously estimated for the target.
-Both the stellar and telluric spectra get cloned.
-The stellar model is warped to its extrinsic properties, and then the stellar and telluric models get multiplied together.
-Next, this joint model is convolved with an instrumental kernel and resampled to the data spectrum.
-Finally, the joint model parameters are optimized to match the data, subject to regularization (not shown).
-The final products are semi-empirical super-resolution stellar and telluric templates, the $RV$ and $v\sin{i}$ of the star, and the line amplitude, location, width, and shape of all 10,000+ stellar and telluric lines.
-\authorcomment1{Some of the spectra are placeholders until we re-run the end-to-end-fitting.
-Feedback on this figure layout is sought.} }
+    \includegraphics[width=1.0\textwidth]{figures/blase_flowchart03.pdf}
+    \caption{Visual flowchart of the \texttt{blas\'e} forward model.
+\textbf{Step 0} (not shown) is to choose a precomputed synthetic stellar spectrum---and, optionally, a precomputed synthetic telluric spectrum---with physical properties close to the target and observing conditions.
+Both the stellar and telluric spectra get cloned (\textbf{Step 1}).
+The stellar model is warped to its extrinsic properties (\textbf{Step 2}), and then the stellar and telluric models get multiplied together (\textbf{Step 3}).
+Next, this joint model is convolved with an instrumental kernel and resampled to the wavelength coordinates of the data spectrum (\textbf{Step 4}).
+It is this forward model that gets directly compared to the observed spectrum (not shown).
+    }
     \label{blase_flowchart}
@@
 Instead, we tune the parameters of the model, starting from these coarse initial values.
-This model has $N_{\mathrm{lines}}\times 3$ free parameters, where the center wavelength is held fixed and the amplitude, width, and lineshape are allowed to vary.
+This model has between $N_{\mathrm{lines}}\times 1$ and $N_{\mathrm{lines}}\times 4$ free parameters, depending on how many of the 4 physical line properties you wish to tune.
+We default to fitting three parameters per line, where the center wavelength is held fixed and the amplitude, width, and lineshape are allowed to vary.
 We minimize a scalar ``goodness-of-fit'' metric, \emph{aka} loss scalar $\mathcal{L}$, chosen as the mean squared error (MSE), which is proportional to $\chi^2$, the sum of the squares of the residual vector $\mathsf{R} \equiv \mathsf{S}-\mathsf{S}_{\rm clone}$ but has no notion of per-pixel noise since the precomputed synthetic spectrum has no uncertainty:
@@
 
-As seen in Figure \ref{fig_Nlines_vs_teff}, the number of lines can exceed 7,000, meaning the model has over 7,000 $\times 3 =$ 21,000 free parameters.
-Fitting more than about 300 parameters is difficult with conventional optimizers-- a $70\times$ shortfall in computation ability.
-Here we employ a variant of Stochastic Gradient Descent (SGD), an optimization technique that can scale to a virtually unlimited number of parameters \citep{2016arXiv160904747R}.
-This technique computes the derivative of the loss scalar with respect to each of the parameters, the so-called \emph{Jacobian}: $(\frac{\partial \mathcal{L}}{\partial a_j}, \frac{\partial \mathcal{L}}{\partial \sigma_j}, \frac{\partial \mathcal{L}}{\partial \gamma_j})$.
-The Jacobian indicates how the MSE would decrease with a change in the parameter-of-interest, or put simply ``which way and by how much'' you have to change the line properties to get a better fit.
+As seen in Figure \ref{fig_Nlines_vs_teff}, the number of lines can exceed 7,000, meaning the clone model has over 21,000 free parameters.
+Fitting more than about 300 parameters is difficult with conventional optimizers that rely on minmizing a single scalar loss: the search space becomes too large.
+Here we employ a variant of Stochastic Gradient Descent (SGD), a transformative optimization technique that can scale to a virtually unlimited number of parameters \citep{2016arXiv160904747R}.
+This technique computes the derivative of the loss scalar with respect to each of the parameters, the so-called \emph{Jacobian}: $\grad \mathcal{L} = (\frac{\partial \mathcal{L}}{\partial a_j}, \frac{\partial \mathcal{L}}{\partial \sigma_j}, \cdots, \frac{\partial \mathcal{L}}{\partial \gamma_j})$.
+The Jacobian indicates how the MSE would decrease with a change in the parameter-of-interest, or put simply ``which way and by how much'' you have to change each individual line property to get a better fit.
 
-The optimizer updates the $a_j, \sigma_j, \gamma_j$ parameters by a small fraction of the Jacobian---called the learning rate LR---towards the direction that would improve the fit, for all $N_{\mathrm{lines}} \times 3$ parameters simultaneously.
+The optimizer updates the $a_j, \sigma_j, \gamma_j$ parameters by a small fraction of the Jacobian---called the learning rate LR---towards the direction that would improve the fit, for all parameters simultaneously.
 The Jacobian is calculated behind the scenes with automatic differentiation implemented as the so-called backpropagation algorithm or simply ``backprop'' \citep{2015arXiv150205767G}.
-We choose the \texttt{PyTorch} framework that computes these Jacobians efficiently for all of the mathematical primitives in our \texttt{blas\'e} implementation.
+We choose the \texttt{PyTorch} framework that computes these Jacobians efficiently for all of the mathematical primitives in our \texttt{blas\'e} implementation \citep{2019arXiv191201703P}.
+
+It is this ability to automatically compute Jacobians that sets \texttt{PyTorch} (and \texttt{JAX} and \texttt{TensorFlow}) apart.
+These frameworks give exact Jacobians, instantaneously, for free (or cheap).
+Without exact Jacobians a conventional optimization step only obtains \emph{one} piece of information: how much the overall loss changed.
+With exact Jacobians, we obtain $N_\mathrm{params}$ pieces of information for each evaluation of the forward model.
+The power of gradient descent becomes transformative as the number of parameters grows into the tens of thousands, as is the case for \emph{blas\'e}.
 
 \subsection{GPU and Autodiff specific considerations}
-We make a few tweaks to the implementation for numerical purposes.
-First, we want all the parameters to be positive, forbidding negative amplitudes and negative widths.
-We, therefore, tune the natural log of the parameters and exponentiate them before inclusion in Equation \ref{equation1}.
-Second, we found through iterative experimentation that the initialization amplitudes and widths were systematically shifted from the optimized values.
-We built-in scalar tweaks to the initialization amplitudes, Gaussian width, and Lorentzian width, which dramatically accelerated the optimization step.
+
+Forward modeling with tens of thousands of physics-informed parameters may seem like such a significant paradigm shift that it can feel too good to be true.
+In this section we introduce the non-negligible architectural design tradeoffs that arise when adopting an autodiff framework---such as \texttt{PyTorch}---for physics based forward modeling.
+
+First, we make a few tweaks to the implementation for numerical purposes.
+We enforce that all Gaussian and Lorenzian widths are positive by tuning the natural log of the widths, and exponentiate them before inclusion in Equation \ref{equation1}.
+Based on the minus sign in Equation \ref{equation1}, an amplitude $a_j$ could hypothetically take on either positive values (flux loss, absorption) or negative values (flux gain, emission).
+For now we focus on photospheric absorption lines---as opposed to, say, chromospheric emission lines---and therefore enforce all the amplitudes to be positive by tuning $\ln{a_j}$ and then exponentiating in the same strategy as above.
+Emission lines---if desired---could be included with a mere sign flip to Equation \ref{equation1}.
+Lines that can manifest either in absorption or emission could hypothetically relax the natural log pre-processing step for isolated lines.
+We narate only the absorption scenario moving forward.
 
 The autodiff machinery has a convenient way to set which parameters are held fixed and which are iteratively fine-tuned.
-One simply disables the Jacobian computations for the fixed parameters.
-We set the \texttt{requires\_grad=True} property for any Torch tensor that we want to vary.
-This allows us to easily explore whether, say, allowing the $\lambda_\mathrm{c}$ parameter to vary significantly improves the fit.
+One simply disables the autodiff flag for the fixed parameters: we set the \texttt{requires\_grad=True} property for any \texttt{PyTorch} tensor that we want to vary.
+This design allows us to easily explore whether, say, allowing the $\lambda_\mathrm{c}$ parameter to vary significantly improves the fit.
 
@@
 
-An element of this matrix, $F_{ji}$, will have the flux value for a given line at a given wavelength coordinate.
+An element of this matrix, $F_{ji}$, will have the flux value for a given $j^{th}$ line at a given $i^{th}$ wavelength coordinate.
 Equation \ref{equation1} performs a type of matrix contraction, turning a $N_{\mathrm{lines}}\times N_{S}$ matrix into a length $N_{S}$ row vector.
 The number of Floating Point Operations (FLOPS) scales with the number of entries in this matrix.
+So we face a tradeoff of wanting to make the matrix large for accuracy and small for computational expedience.
+
 We can rewrite Equation \ref{equation1} as a sum by taking the log of both sides and dropping in this $\bm{\bar{F}}$ matrix:
@@
 \begin{eqnarray}
-    \ln{\mathsf{S}_{\rm clone}} = \sum_{j=1}^{N_{lines}} \ln{F_{ji}} = \mathbf{1} \cdot \ln{\bm{\bar{F}}}
+    \ln{\mathsf{S}_{\rm clone}} = \sum_{j=1}^{N_{lines}} \ln{F_{ji}} = \mathbf{1} \cdot \ln{\bm{\bar{F}}}  \label{eqnFbar}
 \end{eqnarray}
@@
 where $\mathbf{1}$ is a $1\times N_{\rm lines}$ row vector of all-ones.
-We re\"emphasize that each spectral line has to be painstakingly evaluated across the entire spectral bandwidth.
-Efficient GPU algorithms exist for voluminous matrix manipulations such as this one, so this computation can proceed as quickly as possible on modern machines.
+We re\"emphasize that---in its current form---each spectral line has to be painstakingly evaluated across the entire spectral bandwidth.
+Efficient GPU algorithms exist for voluminous matrix manipulations such as this one, so this voluminous computation will proceed as quickly as possible on modern machines.
 In particular, the proprietary CUDA architecture for NVIDIA\textsuperscript{\tiny\textregistered} GPUs contains Tensor cores with specialized matrix math.
@@
 Modern NVIDIA GPUs have $8-40$ GB of RAM, which translates roughly to a few thousand spectral lines across $\sim$300,000 pixels.
-The memory bottleneck is even more pernicious than mere storage of the $\bm{\bar{F}}$ matrix since the CPU/GPU also has to store the computation graph for autodiff.
-It is generally not possible to evaluate Equation \ref{equation1} in its entirety in one-fell-swoop.
+It is generally not possible to construct Equation \ref{eqnFbar} in its entirety in one-fell-swoop, even on a GPU.
 A remedy is needed.
@@
 Extreme Doppler shifts could hypothetically send line cores entirely outside the extents of $N_{\rm cut}$ if set too low.
-We therefore typically set wingcuts comparable to the buffer size $\pm \Delta \lambda_{\mathrm{buffer}}$, even though most weak lines only perceptibly affect $<1\;$\AA.
-We coerce all wing cuts to be the same number of pixels, typically 6000 pixels for \texttt{PHOENIX}, with the middle pixel being at the line center position, and about 3000 pixels to the red and blue side of the line.
-We populate a new approximate sparse matrix $\ln{\bm{\hat{F}}}$ with only these $~6000$ pixels per line and assume zeros everywhere else.
-%\footnote{The PHOENIX spectra double their pixel sampling shortward of $\lambda = 1 \mu$m, so 6000 pixels in the visible is about 30 Angstroms, but 6000 pixels in the infrared is about 60 \AA.
-We \emph{could have} demanded that the wing cut be exactly 30 \AA, or some other heuristic, but that choice would result in so-called \emph{jagged} or \emph{ragged} arrays: collections of 1D arrays that differ in their lengths.
-Some lines would have 5999 pixels, some would have 6000, and others would have 2999 or 3000 pixels, depending on whether they occur to the red or blue side of the $1 \;\mu$m kink in sampling.
-Most efficient CPU and GPU algorithms are not equipped to deal with ragged arrays, therefore increasing the computation time.
-Instead, the standardization choice of exactly 6000 pixels means we can use standard matrix or tensor arithmetic, which somewhat counterintuitively is faster while handling more pixels for all spectral lines: 6000 pixels everywhere is faster than 6000 pixels for some and 2999 for other lines.
-Locating the indices of these 6000 ``nearest-neighbor'' pixels is very fast, only occurs once, and then is fixed.}
+We therefore typically set wingcuts comparable to the buffer size $2 \Delta \lambda_{\mathrm{buffer}}$, even though most weak lines only perceptibly affect $<1\;$\AA.
+We coerce all wing cuts to be the same number of pixels, typically 6000 pixels, $\sim30-60~$\AA~ for \texttt{PHOENIX}, with the middle pixel being at the line center position, and about 3000 pixels to the red and blue side of the line.
+We populate a new approximate sparse matrix $\ln{\bm{\hat{F}}}$ with only these $6000$ pixels per line and assume zeros everywhere else.
 
-The remapping of the sparse matrix can be visualized as the shifting of all lines to the center of this new matrix $\bm{\hat{F}}$.
+The remapping of the sparse matrix can be pictured as having shifted all lines to the center of this new matrix $\bm{\hat{F}}$, visualized pictorially in Equation \ref{eqnPictograph}.
 The algorithmic machinery keeps track of each $(i, j, F_{ji})$ trio of coordinates and flux values.
 
-\begin{equation} 
+\begin{equation} \label{eqnPictograph}
     \begin{aligned}
-    \bm{\bar{F}} &=
+        \bm{\bar{F}} & =
         \begin{pmatrix}
@@
         \end{pmatrix}                        \\
-    \bm{\bar{F}}&\mapsto \bm{\hat{F}} =\begin{pmatrix}
+        \bm{\bar{F}} & \mapsto \bm{\hat{F}} =\begin{pmatrix}
                                                  \includegraphics[height=1cm]{figures/voigt_01b.pdf} & \\
@@
 Sparse matrix methods generally support an operation known as \emph{coalescing}, which sums values with repeated indices.
-We sum the logs and then exponentiate the result before comparing it to the native linear flux.
-We enforce that the argument of the natural log must be greater than zero through a simple truncation, which may approximately resemble the effect of saturation in line cores.
-This truncation would only be relevant for saturated lines.
-Each pixel may get computed about $\sim100$ times in this sparse implementation, which is about $50\times$ better than each pixel getting computed $N_{lines}\sim6500$ times in the dense approach.
-Efficient algorithms for assembling and coalescing sparse matrices exist in \texttt{PyTorch}
+Each pixel may get computed about $\sim100$ times in this sparse implementation, which is about $70\times$ faster than each pixel getting computed $N_\mathrm{lines}\sim7000$ times in the dense approach.
+Efficient algorithms for assembling and coalescing sparse matrices exist in \texttt{PyTorch}.
+Some GPUs now support additional hardware-acceleration of sparse matrices, providing even greater speedups.
 
@@
 
-We use the Adam optimizer \citep{2014arXiv1412.6980K} with a typical learning rate $LR\in (0.005, 0.05)$ and all the defaults for \texttt{PyTorch v1.11}.
+We use the Adam optimizer \citep{2014arXiv1412.6980K} with a typical learning rate $LR\in (0.005, 0.1)$ and all the defaults for \texttt{PyTorch v1.11}.
 We defined the number of training epochs $N_{epoch}=100-10,000$ depending on the application.
@@
 
-\begin{figure*}[hbt!]
+\begin{figure}[hbt!]
     \centering
-    \includegraphics[width=0.6\textwidth]{figures/blase_clone_demo.png}
+    \includegraphics[width=1.0\columnwidth]{figures/blase_clone_demo.png}
     \caption{PHOENIX spectrum cloned with blas\'e.
@@
     \label{fig_cloned_spectrum_demo}
-\end{figure*}
+\end{figure}
 
@@
 We store the model parameters to disk and refer to the entire collection of parameters as a \emph{pre-trained model}.
-More specifically this fine-tuned model represents an evaluable and interpretable clone of the original static pixel values flux values.
+More specifically this fine-tuned model represents an evaluable and interpretable clone of the original static pixel-by-pixel flux values.
 
-\section{Methodology II: Cloning Telluric Spectra}
-Ground-based near-IR \'echelle spectra possess countless depressions attributable to molecular line absorption in Earth's atmosphere.
+\section{Methodology II: Cloning Telluric Spectra} \label{sectionTelluric}
+Ground-based near-IR \'echelle spectra possess thousands of depressions attributable to molecular line absorption in Earth's atmosphere.
 These telluric lines hamper the unbiased interpretation of \'echelle spectra, so some treatment plan is needed.
@@
 The most demanding EPRV applications require a precision characterization of telluric lines that the astronomical community has not yet been able to achieve, and that may rival even the abilities of Earth Science practitioners.
-A hybrid data-/model- driven approach was the chief recommendation of the \emph{Telluric Hack Week} workshop aimed at improving our mitigation of the atmosphere's deleterious effects.
-The \texttt{blas\'e} framework achieves a key milestone by introducing such a \emph{hybrid} approach.
+A hybrid data-/model- driven approach was among the chief recommendations of the \emph{Telluric Hack Week} Workshop\footnote{\url{https://speakerdeck.com/dwhgg/telluric-line-hack-week-wrap-up}} aimed at improving mitigation of the atmosphere's deleterious effects (David W.
+Hogg, \emph{priv.
+comm.}).
+The \texttt{blas\'e} framework achieves a key milestone by introducing a \emph{hybrid} approach to tellurics.
 
@@
 This pixel sampling is about $6\times$ finer than the native PHOENIX pixel sampling.
-The number of telluric lines depends on the atmospheric properties, in particular, the local surface temperature and relative humidity $RH$.
-For a surface temperature of $T=290 \mathrm{K}\; (62^\circ~\mathrm{F})$, relative humidity of 40\%, and typical conditions for McDonald Observatory, we anticipate 3615 telluric lines across the entire HPF bandwidth.
+The number of telluric lines depends on the atmospheric properties, in particular, the local surface temperature $T_\oplus$ and relative humidity $RH$.
+For a surface temperature of $T=290 \mathrm{K}\; (62^\circ~\mathrm{F})$, relative humidity of 40\%, and typical conditions for McDonald Observatory, we anticipate 3615 telluric lines across the entire HPF bandwidth, distributed as shown in the bottom panel of Figure \ref{fig_Nlines_vs_teff}.
 
@@
 So we can hypothetically tolerate a much smaller wing cut for telluric lines.
-In practice, telluric lines can be extremely broad, and accurately cloning these broad telluric lines requires about $N_{\rm cut}\sim6000$ pixels, comparable to the stellar scenario.
+In practice, telluric lines can be saturated, and accurately cloning the resulting broad telluric wings still benefits from $N_{\rm cut}\sim6000$ pixels, comparable to the stellar scenario.
+
+\begin{figure*}[hbt!]
+    \centering
+    \includegraphics[width=0.98\textwidth]{figures/TelFit_clone_2panel.pdf}
+    \caption{Pre-computed TelFit model cloned with \texttt{blase}.
+\textbf{Before:} The top panel shows a forward model evaluated with coarsely initialized line-by-line properties, based on simple threshold-based peak-finding and crude estimation of the shape of the line.
+\textbf{After:} The bottom panel shows the \emph{blas\'e} fine-tuned clone, with nearly indentical spectral structure.}
+    \label{fig_telluric_clone}
+\end{figure*}
 
@@
 We are left with $\mathsf{T}_{\rm clone}(\bm{\lambda}_T)$, the tunable telluric clone model evaluated at its original native coordinates.
+Figure \ref{fig_telluric_clone} shows a before-and-after view of the blas\'e cloning procedure for a TelFit model zoomed in on a $J-$band region with dense Telluric lines.
+We see that the blas\'e fine-tuning appears nearly indistinguishable from the native resolution TelFit precomputed model.
 
@@
 
-We compute the residual $\mathsf{R}(\bm{\lambda}_S)$ of native PHOENIX minus cloned model, illustrated in the bottom panel of Figure \ref{fig_cloned_spectrum_demo}.
+We compute the residual $\mathsf{R}(\bm{\lambda}_S)$ of native-minus-cloned PHOENIX model, illustrated in the bottom panel of Figure \ref{fig_cloned_spectrum_demo}.
 We see an RMS residual of 1.2\%/pixel at native resolution.
-\authorcomment1{Need more prose here.}
-
-We identify three main categories of cloning flaws.
+The telluric clone shows a comparable level of performance.
+These residuals tend to pile up in local symmetrically balanced clusters that get cancelled out once convolved with coarser instrumental line profiles, so their overall effect at instrumental resolution is typically negligible.
+We identify three main categories of cloning flaws that may not be negligible depending on the science application.
 
@@
 Cloning telluric lines suffers from one additional problem.
-Telluric lines can be extremely deep, exhibiting almost vanishing transmission with saturated line cores common in-between the atmospheric windows that define the $I-$, $J-$, $H-$, and $K-$ bands.
+Telluric lines can be extremely deep, exhibiting almost vanishing transmission with saturated line cores common in-between the atmospheric windows that define the $I$, $J$, $H$, and $K$ bands.
 The \texttt{blas\'e} method can cope with these saturated lines, but often it treats nearby-and-blended saturated lines as one single line.
@@
 
-But overall, the telluric model cloning performs adequately, with \censorbox{X percentage} error-per-pixel, and only \censorbox{zz} error-per-pixel when restricted to the atmospheric windows.
-
 \section{Semi-empirical models with transfer learning techniques}\label{transferLearn}
 
-The cloned model already represents a useful intermediate product: the distillation of $N_s=335,849\times2$ pixel flux values and coordinates into a more compact quartet of properties for a list of 9,028 spectral lines, a dimensionality reduction of $18\times$ for the cost of 1.2\%/pixel in accuracy.
+The cloned models already represents a useful intermediate product: the distillation of $N_s=335,849\times2$ pixel flux values and their wavelength coordinates into a more compact quartet of properties for a list of $<10,000$ spectral lines, a dimensionality reduction of $18\times$ for the cost of 1.2\%/pixel in accuracy.
+So the process so far can be myopically viewed as a physics-informed compression algorithm.
 But the cloned model serves as a mere stepping stone in our principal quest: the comparison of models to real data.
 
-\subsection{Augmenting the Stellar Clone with radial velocity and rotational broadening }
+\subsection{Augmenting the stellar clone with radial velocity and rotational broadening }
 Real stars possess two key extrinsic properties.
@@
 Intrinsic properties appear the same from any viewing location---at least for stars with isotropic surfaces---while extrinsic properties do not.
-The distinction is important because the extrinsic terms act as simple convolutions and translations to the cloned spectrum.
+The distinction is important because the extrinsic terms act as simple convolutions and translations to the cloned spectrum, and can be treated after the cloning procedure.
 We therefore define an augmented model, which we designate the ``extrinsic'' model, $\mathsf{S}_{\rm ext}$:
@@
 
-where $\zeta$ is the convolution kernel for rigid body rotation \citep{2022ApJS..258...31K}, $\bm{v}$ is the spectral axis represented as relative velocity coordinates, and $*$ denotes the convolution operator.
+where $\zeta$ is the convolution kernel for rigid body rotation \citep[\emph{e.g.}][]{2022ApJS..258...31K}, $\bm{v}$ is the spectral axis represented as relative velocity coordinates, and $*$ denotes the convolution operator.
 
@@
 There exists a special exponential spectral sample spacing that allows the convolution operators to work out of the box for rotational broadening.
-The design of \texttt{blas\'e} permits the telluric model to be re\"evaluated at any wavelength coordinate vector, provided that it adequately samples the underlying lines.
-We, therefore, change the sampling from the native stellar and telluric wavelength coordinate grids, $\bm{\lambda}_S$ and $\bm{\lambda}_T$, to a new exponentially sampled wavelength grid, denoted with the subscript $Z$:
+The design of \texttt{blas\'e} permits the stellar and telluric models to be re\"evaluated at any wavelength coordinate vector, provided that it adequately samples the underlying lines.
+We, therefore, change the sampling from the native stellar and telluric wavelength coordinate grids, $\bm{\lambda}_S$ and $\bm{\lambda}_T$, to this special exponentially sampled wavelength grid, denoted with the subscript $Z$:
 
@@
 Operationally, the radial velocity shift $RV$ gets applied to the line center positions rather than scaling the entire wavelength grid point coordinates, $\bm{\lambda}_Z$.
-This approximate choice ignores relativistic effects, but yields a convenience: it cleanly makes the $RV$ autodiff-aware, meaning that an infinitesimal change to the RV value can be sensed through backpropagation by affecting only the line center positions.
+This choice yields a convenience: it cleanly makes the $RV$ autodiff-aware, meaning that an infinitesimal change to the RV value can be sensed through backpropagation by affecting only the line center positions.
+
+The $\mathsf{S}_{\rm ext}$ spectrum is shown in Step 2 of Figure \ref{blase_flowchart}.
+There is currently no equivalent post-processing of the telluric spectrum $\mathsf{T}$.
+As mentioned previously, we assume motions of the Earth's atmosphere are much less than the desired stellar radial velocity precisions.
+However, demanding Extreme Precision Radial Velocity (EPRV) exoplanet searches may need to consider minuscule systematic $RV$ shifts and broadening of the telluric templates, arising from the turbulent and bulk motions of the Earth's atmosphere.
+A $\mathsf{T}_{\rm ext}$ could hypothetically be implemented in \emph{blas\'e} to achieve these strenuous precision demands.
 
 \subsection{Joint Stellar and Telluric Model}
-We then simply multiply the stellar model by the telluric transmission:
+Figure \ref{blase_flowchart} shows a visual guide to all the steps in \texttt{blase}.
+We have arrived at what may be the most intriguing-and-yet-simple of these steps: we simply multiply the rotationally-broadened-and-$RV$-shifted stellar model by the telluric transmission:
 
@@
 The instrumental resolving power, $R$, acts as a convolution with a Gaussian line profile of width $\sigma=\frac{c}{2.355 R}$.
-Real astronomical instruments usually have wavelength-dependent resolving power, which complicates the implementation.
+Real astronomical instruments usually have wavelength-dependent resolving power, which complicates the implementation for high-grasp spectra.
 The extent to which this effect matters will depend on the science application.
@@
 
-The data-pixel sampling is much coarser than the model pixel sampling.
-This resampling step could be achieved with two choices.
-One would be to simply evaluate the closed-form model at the coordinates of the data pixels $\bm{\lambda_D}$.
-This choice would require a closed-form solution to Equation \ref{eqn_convolution}, which has recently been reported \citep{2021arXiv211006271L}.
-Instead, we choose the conventional sampled approach, in which we evaluate the extrinsic model at all of the native resolution wavelength coordinates, and then compute the mean value of those pixels within the bounds of each coarse data pixel.
+Notice that the order of Steps 3 and 4 in Figure \ref{blase_flowchart} cannot be swapped.
+Mathematically speaking, multiplication and convolution do not commute.
+While the distinction may seem negligible, it matters at the level of precisions sought in EPRV applications (Suvrath Mahadevan, \emph{priv.
+comm.}).
+Water vapor lines in our own atmosphere can ``beat'' with water vapor in the spectrum of, say, an M-dwarf atmosphere.
+The systematic telluric mis-cancelation would imbue a Moir\'e pattern of residuals that is most acute for sources with sharp lines, namely low projected rotational broadening ($v\sin{i}\sim\frac{c}{2R}$).
+The approach in \emph{blas\'e} may therefore unlock a level of telluric calibration that has evaded previous efforts.
+
+As noted, typical data-pixel sampling $\bm{\lambda}_D$ is much coarser than the model pixel sampling $\bm{\lambda}_Z$.
+We therefore resample the model to the data spectrum in the following way.
+We evaluate the joint model $\mathsf{M}_{\rm joint}$ at all of the super-resolution wavelength coordinates, and then compute the mean value of those pixels within the bounds of each coarse data pixel.
 The resampling procedure is autodiff-aware: the same clusters of high resolution coordinates map to the same data pixel coordinates, no matter what the $RV$ is.
-The $RV$ only dictates what flux values are realized within those pixels.
-The final forward model for \emph{blas\'e} is designated simply as $\mathsf{M}$ without subscripts to emphasize that we have achieved the desired goal of a plausible model for each datum in the 1D observed spectrum:
+The $RV$ only dictates what flux values are realized within those pixel bounds.
+
+The final forward model for \emph{blas\'e} is designated simply as $\mathsf{M}$ without subscripts to emphasize that we have achieved the desired goal of a plausible end-to-end physics-informed yet highly flexible forward model for each datum in the 1D observed spectrum:
 
@@
 
-Equation \ref{eqn_final_model} has up to $N_{\mathrm{lines}}\times3 \sim20,000$ tunable parameters from the star, $N_{\mathrm{lines}}\times3 \sim15,000$ tunable parameters from the Earth's atmosphere, plus $v\sin{i}$ and $RV$.
-That adds up to 35,002 model parameters.
-The resolving power may also be treated as tunable if it is not known or varies slightly with \emph{e.g.} seeing, slit-or-fiber illumination, or instrumental configuration.
+Equation \ref{eqn_final_model} has $\sim21,000$ tunable parameters from the star, $\sim9,000$ tunable parameters from the Earth's atmosphere, plus $v\sin{i}$ and $RV$.
+That adds up to about $30,002$ model parameters.
+The resolving power may also be treated as tunable if it is not known or varies slightly with \emph{e.g.} seeing, slit-or-fiber illumination, or instrumental configuration: $30,003$.
 
 It may appear desirable to simply optimize all of these parameters in a \emph{laissez-faire} manner, allowing them all to take on whatever value the data dictates.
-Such a stratagem would overfit the data, resulting in unphysically perverse lineshapes that do not reflect the air of reality we attempt to impose on our synthetic spectral models.
+Such a stratagem would overfit the data, resulting in unphysically perverse lineshapes that do not reflect the air of reality we aspire to impose on our synthetic spectral models.
 Lines would haphazardly fit noise-spikes, and conspire together to warp spectral shapes in unexpected ways.
@@
 Fortunately, we have a firm theoretical basis to justify this regularization.
-We believe our precomputed synthetic spectral models are \emph{quasi-statically correct}: the predicted spectra resemble the unobserved ``latent spectrum'' with lines in the correct place, but just with the wrong amplitudes.
+We believe our precomputed synthetic spectral models are \emph{quasi-statically correct}: the predicted spectra resemble the unobserved ``True spectrum'' with lines in the correct place, but just with the wrong area under the curve.
 This statement may stem from the fact that it is easier to predict the mere existence of some energy transition of atoms and molecules than it is to predict their transition rates, abundances, temperature and pressure effects, and all the other line strength effects that flow down to how much light a line ultimately absorbs in a stellar atmosphere.
 
-Even still, the regularization step is in some ways the most subjective, since the degree of regularization will control the degree of overfitting or underfitting.
-The most extreme regularization---the antithesis of the \emph{laissez-faire} scenario---would yield a model too rigid to respond to the data at all, yielding a model entirely unchanged from the cloned PHOENIX and \texttt{TelFit} models.
+The degree of regularization will control the extent of overfitting or underfitting.
+The most extreme regularization---the antithesis of the \emph{laissez-faire} scenario---would yield a model too rigid to respond to the data at all, yielding a model entirely unchanged from the cloned PHOENIX and \texttt{TelFit} models, the extreme end of underfitting.
 So regularization constitutes the only hyperparameters worthy of tuning in \texttt{blas\'e}.
@@
 We default to the following choice.
-We usually fix all line parameters except for amplitude, which receives a log-normal distribution:
+We fix all line parameters except for amplitude, which receives an L1 loss-- namely we penalize the absolute value of departures from an amplitude's starting place:
 
 \begin{eqnarray}
-    \mathrm{regularization} &\equiv& -\ln{\mathrm{prior}} \nonumber \\ 
-    &=& \frac{1}{2} \sum_{j=1}^{N_\mathrm{lines}} \Big(\frac{\ln{a_j} - \ln{a^\prime_j }}{\Lambda}\Big)^2
+    \mathcal{L}_{reg} &\equiv& \sum_{j=1}^{N_\mathrm{lines}} \Big|\frac{\ln{a_j} - \ln{\hat{a}_j}}{\Lambda}\Big|
 \end{eqnarray}
 
-where the prime notation demarcates the amplitudes obtained from cloning, or the ``initial amplitude''.
-We assign $\Lambda_{\ln{a}}\sim5$.
-This weak degree of regularization has the effect of permitting refinement of only the most conspicuous data-model mismatches; the weakest lines don't both to move from their initial state because doing so would penalize the regularization penalty without enough reduction of the overall loss.
+where the hat notation demarcates the amplitudes obtained from cloning, \emph{i.e.} the ``initial, theory-inspired amplitude''.  We assign $\Lambda\sim5$.  The total, overall loss then becomes:
+
+\begin{eqnarray}
+    \mathcal{L}_{tot} = \mathcal{L}_{MSE} + \mathcal{L}_{reg}
+\end{eqnarray}
+
+This weak degree of regularization has the effect of permitting refinement of only the most conspicuous data-model mismatches; the weakest lines do not bother to move from their initial state because doing so would penalize the regularization without enough reduction of the overall loss.
 The extrinsic $v\sin{i}$ and $RV$ have no regularization, but in practice, they need to be initialized close to their plausible values.
 
-Figure \ref{fig_WASP69_demo} shows an example of comparing HPF data to the closest PHOENIX template, post-processed to the HPF resolution and sampling.
-Figure \ref{fig_WASP69_transferred} shows the same data spectrum compared to a pixel-level \emph{blas\'e} model regularized to the clone in Figure \ref{fig_cloned_spectrum_demo}.
-The latent high-resolution model for WASP 69 is therefore the transfer-learned model, unadjusted for the extrinsic and instrumental properties.
+\section{Results: Comparison to data}\label{secResults}
+\subsection{WASP 69 with HPF}
+
+The planet-host star WASP 69 makes a great benchmark because it has a low $v\sin{i}=2.2\pm0.4~\text{km/s}$ \citep{2017A&A...608A.135C}, making its lines sharp and easy to perceive.
+The K5 dwarf has an effective temperature of about 4700 K, $\log{g}=4.535\pm0.023$, and slightly super-solar metallicity \citep{2014MNRAS.445.1114A}.
+
+Figure \ref{fig_WASP69_demo} shows a portion of an HPF spectrum of WASP 69 centered on a region devoid of telluric lines.
+This figure highlights a baseline case with no line-by-line fitting, simply conventional template matching: comparing HPF data of WASP 69 to the closest PHOENIX template, convolved and resampled to the HPF resolution and sampling.
+Conspicuous residuals of $\pm10\%$ appear throughout the spectrum, with lines in the correct place, but with the amplitudes systematically biased.
+These lines residuals arise from bona-fide imperfections in the PHOENIX spectrum, with a minor contribution from the coarse sampling of the PHOENIX grid.
+
+Figure \ref{fig_WASP69_transferred} shows the same data spectrum compared to a pixel-level model trained using the \emph{blas\'e} technique described in Section \ref{transferLearn}.
+The model fit appears much better, with typical residuals approaching the photon noise of the data themselves.
+The model is not perfect however, especially around line cores.
+These residiuals stem from a combination of causes.
+First, our finite regularization restricts line amplitudes and widths from straying too far from their values.
+This computational tug-of-war makes line cores land just-short of the values the otherwise would have obtained in the absence of regularization.
+Second, the spectral resolution kernel may get biased from the domineering model-misspecification of broad lines, setting up a slightly sub-par performance for all the other lines.
+
+The inferred, semi-empirical high-resolution model for WASP 69 is therefore the transfer-learned model $\mathsf{\hat{S}}_\mathrm{clone}$, unadorned with the extrinsic and instrumental properties.
 This semi-empirical model is shown in Figure \ref{fig_WASP69_regularized}.
+The departures from the PHOENIX model appear dramatic at this native resolution.
+The transfer-learned semi-empirical template exhibits both deeper and shallower lines than the native PHOENIX model.
+
+\begin{figure}[hbt!]
+    \centering
+    \includegraphics[width=0.98\columnwidth]{figures/blase_WASP69_demo.png}
+    \caption{WASP 69 observed with HPF, compared to a $T_{\mathrm{eff}}=4700\;$K,  $\log{g}=4.5$ solar metallicity PHOENIX model warped to $v\sin{i}=2.2$~km/s, $RV=-9.6$~km/s, and HPF resolving power.}
+    \label{fig_WASP69_demo}
+\end{figure}
+
+\begin{figure}[hbt!]
+    \centering
+    \includegraphics[width=0.98\columnwidth]{figures/blase_WASP69_regularized.png}
+    \caption{Semi-empirical model of WASP 69 transfer-learned with \texttt{blas\'e}, employing a regularization prior on the learned amplitudes.}
+    \label{fig_WASP69_transferred}
+\end{figure}
+
+\begin{figure}[hbt!]
+    \centering
+    \includegraphics[width=0.98\columnwidth]{figures/blase_super_resolution_template.png}
+    \caption{The native-resolution semi-empirical model transfer-learned from WASP 69 HPF data. The revised model can be viewed as a super-resolution deconvolution of the HPF spectrum.}
+    \label{fig_WASP69_regularized}
+\end{figure}
+
+The low $v\sin{i}$ of WASP 69 means that the line profile broadening arises principally from the finite instrumental resolution.
+The interplay of telluric lines and stellar lines within an instrumental resolution element is exactly one of the challenges blas\'e was designed to solve, as discussed in Section \ref{sectionTelluric}.
+In Figure \ref{fig_multi_panel_WASP69} we show a multi-panel dissection of the HPF spectrum in a wavelength region in which conspicuous stellar and telluric lines coexist.
+The blas\'e end-to-end model exhibits excellent agreement with the data.
+The trustworthiness of the line properties inferred in locations where stellar and telluric lines exactly overlap remains an open research question: a tie-breaker is needed to distinguish these overlapping line inferrences.
+We discuss conceivable tie-brakers later in the paper.
 
@@
     \centering
-    \includegraphics[width=0.6\textwidth]{figures/blase_WASP69_demo.png}
-    \caption{WASP 69 observed with HPF, compared to a $T_{\mathrm{eff}}=4700\;$K,  $\log{g}=4.5$ solar metallicity PHOENIX model warped to $v\sin{i}=2.2$~km/s, $RV=-9.6$~km/s, and HPF resolving power.}
-    \label{fig_WASP69_demo}
+    \includegraphics[width=0.95\textwidth]{figures/WASP_69_multi.pdf}
+    \caption{End-to-end training on a portion of HPF spectrum of WASP-69.
+The top two panels show the cloning in from Step 1 of the visual guide.
+The middle two panels depict the before-and-after of the instrument-convolved and resampled joint model illustrated in Step 4 of the visual guide.
+The final two panels show the underlying semi-empirical models learned in the process.}
+    \label{fig_multi_panel_WASP69}
 \end{figure*}
+
+\subsection{IGRINS spectrum of a T6 Ultracool Dwarf}
+
+The T6 ultracool dwarf 2MASS J08173001$-$6155158 was recently observed with IGRINS, revealing a rich spectroscopic atlas of molecules in its $T_\mathrm{eff} = 1060 \pm 50$ cloud-free atmosphere \citep{2022MNRAS.tmp.1421T}.
+The sea of moleculaar lines are so rich as to blur entirely the notion of isolated lines.
+Virtually all lines should instead be considered pseudo lines, computed using the opacity agglomeration technique known as ``correlated-$k$''.
+The boundaries of lines for 2MASS J08173001$-$6155158 become even more amorphous in the presence of its moderately high $v\sin{i} = 22.5\pm 0.5 ~ \text{km/s}$ \citep{2022MNRAS.tmp.1421T} rotational broadening.
+This T6 IGRINS spectrum therefore represents an extreme end of a test-case, in which the underlying notions of \emph{blas\'e} go outside the comparatively safe stellar spectra.
+The application of blas\'e to 2MASS J08173001$-$6155158 can be viewed as conducting a deconvolution step simultaneously paired with its line-by-line inference capability in one-fell-swoop.
+
+We initialize \texttt{blas\'e} with the nearest Sonora template with $T_\mathrm{eff} = 1100~\text{K}$, $\log{g}=5.0$, and solar metallicity.
+Figure \ref{fig_IGRINS_sonora_demo} shows the clone of this native spectrum near the peak of the $H-$band.
+The cloning performance appears adequate to capture most of the molecular pseudo-lines, with a few notable shortcomings.
+The high signal-to-noise ratio of this IGRINS spectrum lays bare the departures from the broadened Sonora-Bobcat model, albeit with surprisingly close agreement given the challenge of substellar atmosphere modeling.
+The line-by-line departures from Sonora exceed any minor flaws from the cloning process, indicating genuine opacity differences between the data and Sonora template.
+Our technique automatically detects and reports the underlying structure of data-model mismatches, a key milestone for the assembly of refined opacity tables in this cool dwarf temperature regime.
 
@@
     \centering
-    \includegraphics[width=0.6\textwidth]{figures/blase_WASP69_regularized.png}
-    \caption{Semi-empirical model of WASP 69 transfer-learned with \texttt{blas\'e}, employing a regularization prior on the learned amplitudes.}
-    \label{fig_WASP69_transferred}
+    \includegraphics[width=0.98\textwidth]{figures/IGRINS_T6_sonora_demo.png}
+    \caption{Line-by-line dissection of the 2MASS J08173001$-$6155158 IGRINS spectrum.
+The \emph{top} panel compares the native Sonora-Bobcat spectrum to the blas\'e clone.
+We match the RV shift of the observed spectrum (thick black line) and convolve the initial and learned Sonora template to the large $v\sin{i}$.
+The bottom panel shows the locations and extents of line mismatches at the native resolution of Sonora.}
+    \label{fig_IGRINS_sonora_demo}
 \end{figure*}
-
-\begin{figure*}[hbt!]
-    \centering
-    \includegraphics[width=0.6\textwidth]{figures/blase_super_resolution_template.png}
-    \caption{The native-resolution semi-empirical model transfer-learned from WASP 69 HPF data. The revised model can be viewed as a super-resolution deconvolution of the HPF spectrum.}
-    \label{fig_WASP69_regularized}
-\end{figure*}
-
-% \begin{outline}
-%     \1 What is the residual level with bare PHOENIX (i.e. tuning continuum only, not lines)?
-%     \1 What is the residual level with no regularization (i.e. tune all lines and continuum)?
-%     \2 Should be near-zero except for missing lines
-%     \2 What should we do about missing lines?
-%     \1 What is the residual level with modest regularization? What is the typical change to the line?
-%     \2 Plot of cloned FWHM versus FWHM transferred
-%     \2 The pseudo-Voigt line properties can be analytically integrated to give an equivalent width
-%     \2 Plot of cloned EW (before) versus transfer EW (after)
-% \end{outline}
-
-%\section{Injection/recovery tests}\label{secInjRec}
-%Here we conduct injection/recovery tests.
 
@@
 
-\subsection{Comparison to existing frameworks}
+\subsection{How to interpret line-by-line properties}
 
-Several astronomical spectral frameworks share similar aims as \texttt{blas\'e}.
-These existing frameworks will have enduring value for the wide range of problems in the field of stellar spectroscopy.
-Here we scrutinize the differences among some of these approaches to clarify how This Work fits in.
+Up to this point we have presented the mechanics of blas\'e, and shown that it applies to a wide range of applications including stars, substars, and Earth's atmosphere.
+The interpretation of the approach and its line-by-line outputs may carry different meaning to different practitioners.
+Here we caricature some of these perspectives:
 
-The \texttt{specmatch} synthetic template matching tool produces noise-free nearest neighbor templates given an input spectrum \citep{2015PhDT........82P}.
-Several practical barriers limit the accuracy of using precomputed synthetic spectral models alone.
-First and foremost, real stars are usually more complicated than our simplified models of them.
-Real spectra often vary over more dimensions that our models do.
-Conspicuous examples of these hidden variables can be found in protostars: starspots, accretion veiling, dust extinction, and magnetic Zeeman splitting.
-Jointly modeling all of these phenomena alongside the intrinsic stellar photosphere is challenging.
+\emph{``Reverse-engineering the line lists''}-- Equation \ref{equation1} (and its exponential twin Equation \ref{equationOpacitySum}) could have been populated with initial line-by-line properties from atomic and molecular line lists, instead of the coarse detection-threshold approach we presented in Section \ref{subsecInit}.
+So our cloning procedure can be viewed as reverse-engineering the quantum mechanical properties housed in line lists such as HITRAN, HITEMP, EXOMOL, VALD, or countless other primary source documents.
+This interpretation is only partly accurate.
+The line lists generally store the temperature and pressure scaling terms that get included in 1-D thermal structure calculations with distinct Voigt coefficient terms computed for each layer in the stellar or substellar or Earth atmosphere.
+The cloned properties therefore represent the flux-weighted average line property, having undergone an integral through the radiative transfer in the atmosphere.
 
-The empirical version, \texttt{specmatch-emp} \citep{2017ApJ...836...77Y} matched spectra better than the synthetic templates, but is still too rigid for some applications and requires the assembly of hundreds of standardized high signal-to-noise-ratio templates, ideally with low intrinsic rotational broadening.
-Such a large number of high-quality templates with high resolving power and low $v\sin{i}$ has not yet been established in the near-infrared.
+Our deliberate choice not to initialize from the atomic and molecular lines means we do not have to compute an expensive multi-layer atmosphere at each inference step.
+The precomputed models have aleady taken care of that step, and so our cloning procedure encodes all the quantum mechanics and stellar atmospheres knowledge written down at the time of the computing of PHOENIX, Sonora, or LBLRTM.
 
-The \texttt{wobble} framework \citep{2019AJ....158..164B} modernized the construction of high-SNR templates to account for temporally variable telluric lines.
-The tool requires dozens of high-SNR spectra acquired at a range of Barycentric Earth Radial Velocities (BERVs).
-The final telluric-free combined spectrum would still have to be compared to models for absolute calibration or can be used out-of-the-box for precision relative RVs.
-The \texttt{wobble} framework also pioneered the off-label application of automatic differentiation frameworks---in this case \texttt{TensorFlow}---towards their physically-motivated use in stellar spectra.
-\texttt{blas\'e} can be viewed as an evaluable and interpretable super-resolution version of \texttt{wobble}, that accepts more bias in the bias-variance tradeoff.
+\emph{``Equivalent Width machine''}-- The area under the curve of each $j^{th}$ blas\'e line represents its (neighbor-weighted) Equivalent Width (EW).
+That means that blas\'e can automatically and quickly distill a massive high-banwidth spectrum into a catalog of line positions and EWs.
+Even more, the catalog would have some resilience to rotaional broadening, line blending, and other friction points that hamper such an industrial-scale approach with existing tooling.
+Current applications of Equivalent Widths rely on either isolated lines or---for multi-epoch applications---the assumption of non-variable neighboring lines.
+Blas\'e does not have these limitations.
+The interpretations of these EWs may have to be treated with care if intercompared with EWs obtained in the conventional way.
+This approach may suit stellar abundance applications, such as Galactic archaeology and near-field cosmology.
 
-The \texttt{starfish} framework \citep{czekala15} provides a robust likelihood function for data-model comparisons and retires many of the problems in this domain.
-\texttt{starfish} pioneered the use of whole-spectrum fitting with resilience to model imperfections by addressing the problem of what to do when the underlying atomic and molecular data was wrong or approximate or missing.
-It has been extended to inferring starspot physical properties \citep{2017ApJ...836..200G}, measuring veiling in Class 0 protostars \citep{2018ApJ...862...85G}, and quantifying imperfections in brown dwarf models \citep{2021ApJ...921...95Z}.
-The Spectral Inference Crank \citep[\texttt{sick},][]{2016ApJS..223....8C} shares similar aims as \texttt{starfish}, and provides additional useful grid search capabilities.
+\emph{``Fancy deconvolution''}-- The blas\'e approach can be viewed as a fancy deconvolution procedure, in which the convolution kernels gets inferred alongside the imperfections of the model template.
+This perspective may suit practitioners who wish to obtain a robust rotational broadening estimate in the presence of imperfect templates, and vice versa: robust templates in the presence of imperfect rotational broadening.
 
-For very large bandwidths and very many spectral lines, the problem of identifying and cataloging line imperfections essentially becomes a book-keeping and continuum assignment problem.
-\texttt{blas\'e} and \texttt{starfish} provide different strategies for orchestrating the line-mismatch identification procedure, with each route having tradeoffs depending on the application.
+\emph{``A surrogate model, emulator, or discrepancy modeling''}-- The machine learning community is increasingly incorporating physics-based emulation into machine learning models.
+Blas\'e can be viewed as such a technique, often labeled as ``surrogate models''.
 
-\section{Conceivable Improvements, Extensions, and Limitations}
+\emph{``A good enough substrate''}-- The imperfections in pre-computed stellar and substellar models have hampered their use in high-fidelity applications, where model flaws overwhelm the sigal-to-noise in the data.
+The adaptability of blas\'e can be viewed as increasing the number of applications that could conceivably benefit from the rich-albeit-imperfect information enoded in these model grid spectra.
+In the next section we illustrate how these adaptable models can serve as a subtrate for applications in which the imperfections in precomputed models would have otherwise precluded their use.
 
-As currently implemented, \texttt{blas\'e} represents a minimum viable product: it out-performs some existing approaches for some science cases.
-The scaffolding of the framework is designed in a way to promote extensibility, so the future of the framework is bright.
-Here we enumerate some planned or conceivable extensions to the framework that would unlock new science arenas.
-We order the presentation of these extensions from our perceived ease of implementation, from easiest to hardest.
+\section{Unlocking new science}\label{secFutureWork}
 
-\subsection{Exact instead of pseudo Voigt Profile}
-We currently employ the pseudo-Voigt profile for its low computational cost.
-We have a prototype exact-Voigt-Hjertingimplementation following \citet{2022ApJS..258...31K}.
-We coarsely estimate that moving to this exact-Voigt implementation could decrease some residual regions by $\sim 30\%$, while increasing the computational cost by more than $10\times$ over the existing pseudo-Voigt approximation.
-The exact-Voigt-Hjerting implementation still outbids the higher cost of a direct numerical convolution of a Gaussian and Lorentzian profile.
+The scaffolding of \texttt{blase} is designed in a way to promote extensibility, so many scientific questions can be written down into the flexible spectral modeling approach presented in this paper.
+In some cases the flexibility, precision, speed, and ease-of-use may unlock new approaches to long-standing astrophysical questions and practical challenges.
+Here we enumerate some planned or conceivable extensions.
+We break up the themes into two categories: scientific extensions (this Section) and technical innovations (Section \ref{technicalInnovations}).
+We order the presentation of these extensions by perceived scientific impact:
 
-\subsection{Addressing the pseudo-continuum with Gaussian Process regression}
-
-We currently assume the input spectra are perfectly normalized to the continuum.
-This approximation impacts all training stages, including stellar cloning, telluric cloning, and data-model comparison.
-Gaussian Processes (GPs) could and should be brought to bear at each of these optimization stages, as described in the ``global kernels'' in \texttt{Starfish} \citet{czekala15}.
-In short, a GP-likelihood relaxes the assumption that the continuum has been perfectly normalized, in favor of the more realistic statement ``the continuum has been coarsely normalized, with some characteristic-but-as-yet-unknown correlation and scale length and amplitude of the imperfections''.
-That statement translated to the following modification to Equation \ref{simpleLikelihood}:
-
-\begin{eqnarray}
-    \mathcal{L} =  \frac{1}{2}\mathsf{R^\intercal} \mathsf{C}^{-1} \mathsf{R} +\frac{1}{2}\ln{\det{\mathsf{C}}} \label{GPLikelihood}
-\end{eqnarray}
-
-where we introduce the covariance matrix $\mathsf{C}$, with its concomitant kernel and collection of typically 2-3 parameters.
-We anticipate that this GP likelihood would have the greatest impact on stellar stars with significant band-heads and line-blanketing: spectra with a so-called ``pseudo-continuum''.
-M-dwarfs and brown dwarfs fall into this fear-inducing category.
-
-The main demerit of moving to a GP-likelihood is computational cost.
-Fortunately, a few efficient autodiff-aware implementations of GPs exist.
-The \texttt{celerite} algorithm \emph{celerit\'e} \citep{2017AJ....154..220F} has an exact backpropation implementation \citep{2018RNAAS...2...31F} that scales linearly with the number of data points.
-The \emph{celerit\'e} algorithm does not currently have a PyTorch implementation.
-The \texttt{GPyTorch} framework \citep{2018arXiv180911165G} has a large category of approximate and exact GPs that could be straight-forwardly dropped into \texttt{blas\'e}.
-Even still these GPs could increase the computation cost by of order $10\times$.
-
-An alternative to GPs could be to simply tune the $\mathsf{P}$ term that represents the wavelength-dependent pre-factor to Equation \ref{equation1}.
-Tuning $\mathsf{P}$ would correct for only large-scale imperfections in the otherwise-fixed continuum flattening procedure.
-
-\subsection{Minibatches and Stochastic Gradient Descent}
-
-Currently, each training epoch sees the entire dataset, a setup dubbed \emph{full-batch} gradient descent.  
-An alternative scheme allows training with only a portion of the entire dataset at a time in \emph{minibatches}.
-The massive data volumes in modern Neural Network applications cannot fit into the GPU memory, so minibatches are a necessity.
-Our meager 1 MB dataset can easily fit into the GPU memory, but our model can be large if we have a large number of pixels or lines or both.
-So while minibatches may not be required due to data size limitations they may be useful for particularly large models.  
-Minibatches also act as a form of regularization, the principal source of stochasticity in the Stochastic Gradient Descent algorithm, which tends to have better convergence than full-batch Gradient Descent \citep{2016arXiv160904747R}.
-
-We experimented with minibatches by assembling and evaluating only a portion of the dense $\bar{\bm{F}}$ matrix at a time, in minibatches.
-The choice to evaluate only a portion of lines at a time would mean the model is inaccurately evaluated at all wavelength pixels.
-Instead, we choose to evaluate all lines, but only on a random subset $N_{\mathrm{batch}}$ of the total pixels $N_s$, so that the model can eventually converge to exact at those points.
-All lines are allowed to update at each glimpse of a minibatch, but many lines with cores far from minibatch pixels will provide only weak information about how the loss scalar changes for their parameters.
-The value of $N_{\mathrm{batch}}$ should be set just below the threshold at which the computation runs out of memory.
-At present, this threshold is determined experimentally.
-The indices of the wavelength points can either be set to be contiguous in blocks (duplicating the sparse matrix approach), or random with possibly large gaps between indices, or else with some other scheme tailored to the most informative wavelength regions.
-We experimented with several techniques and settled on random sampling with replacement, meaning that some pixels get revisited more often than others, with some minibatches tailored only to the line cores.
-Overall minibatches as implemented above performed worse than the sparse implementation, with both lower accuracy and slower computation time.
-
-\subsection{Broad lines and advanced lineshapes}
-
-Some lines, such as Hydrogen, sodium, potassium, and others have extremely broad line wings, approaching larger than the $\sim6000$ pixels we allocate for the sparse implementation.
-These special lines should be handled separately from the weak lines, both from a computational performance perspective and an accuracy perspective.
-
-Extremely broad lines will exhibit truncation effects if the sparse window is small compared to the line wing size.
-The truncation effects will look like tophat functions severing the asymptotic wings, imbuing artificial step function kinks in the emulated spectrum.
-We can afford to increase the sparse window on a few, say $N_{broad}\sim20$ of the broadest spectral lines.
-We then construct and evaluate the entire dense matrix for those lines: $\sim 330\;000 \times 20$.
-The number of FLOPS in each category scales as about 6 Million for the 20 broad and dense lines versus about 36 Million for the sea of about 6500 narrow and sparse lines, depending on the exact choices for wing cuts and the number of lines.
-
-One could introduce advanced lineshapes for these $\sim20$ broad lines, perturbing the Voigt line-wings with a smooth wavelength-dependent correction term $\mathsf{G}$:
-
-\begin{eqnarray}
-    \mathsf{\tilde{V}(\bm{\lambda})} &=& \mathsf{V} \odot \mathsf{G}\\
-    \mathsf{G} &=& 1 + (e^{a_j} - 1) \cdot \mathcal{S}\left(\frac{|\bm{\lambda}-\lambda_{c,j}| - \lambda_{t, j}}{b_j}\right)
-\end{eqnarray}
-
-where $\odot$ is again the element-wise product (\emph{a.k.a} Hadamard product), $\mathcal{S}$ is the sigmoid function, and we have introduced three new tunable parameters for each of the $j$ broad lines; $\lambda_t$ is the truncation wavelength, $b$ is a scale parameter for how slowly or how rapidly in wavelength-space the transition from non-Lorentzian proceeds, and $a$ is a possibly negative stretch parameter that controls whether the line wing is sub- or super- Lorentzian.
-
-This functional form has a few advantages.
-It is smooth.
-The smoothness of the transition is controlled by a tunable parameter, $b$.
-It can handle either sub- or super-Lorentzian shapes.
-It is roughly based on the $\chi$-factor\authorcomment1{citation to Hartmann et al.
-1989, etc.}.
-In the limit $\lim_{a\to0} \mathsf{G}$, the lineshape becomes exactly Lorentzian.
-The sigmoid is efficiently implemented in PyTorch.
-It enforces that the perturbation only produces absorption and not emission profiles.
-
-\subsection{Wavelength dependent limb darkening}
-Currently the extrinsic model step possesses up to four parameters: the $v\sin{i}$ and $RV$, and 2 optional parameters for limb-darkening.
-These four parameters may adequately parameterize a star with a uniform stellar disk.
-Extremely high signal-to-noise-ratio spectra of rapidly rotating stars may require additional flexibility.
-The limb darkening is generally wavelength-dependent, and so a pan-chromatic spectrum may require a different limb darkening from the blue end to the red end.
-The limb darkening may instead depend on physical properties of the spectral line formation, such as physical depth of formation, and so the extent of limb-darkening may jump haphazardly from line-to-line-to-line, rather than as a predictably smooth function across wavelength.
-\texttt{blas\'e} could be built to handle such a seemingly pathological scenario by adding a vector of limb darkening parameters, one for each line.
-One would have to regularize the fits with some typical limb darkening and a heuristic penalty for departures from this mean.
+\subsection{Line-by-line fundamental parameter estimates}
+to-be-completed...
 
@@
 The fixed $v\sin{i}$ approximation breaks down for stars with large-scale surface features.
-Doppler imaging \authorcomment1{xx, yy, zz} attempts to reverse engineer the surface map from the extent to which observed line profiles depart from a pristine rotational broadening kernel.
+Doppler imaging attempts to reverse engineer the surface map from the extent to which observed line profiles depart from a pristine rotational broadening kernel.
 This reverse engineering step suffers from a vast number of geometrical degeneracies, but still provides useful constraints on stellar surfaces.
-Following LugerXX, we emphasize a distinction between A) longitudinally symmetric surface features, and B) longitudinally asymmetric surface features.
-Most radial velocity practitioners think about the latter, since longitudinally asymmetric surface features imbue skewness to the line profiles, causing radial velocity perturbations easily detectable in radial velocity time series.
+We emphasize a distinction between A) longitudinally symmetric surface features, and B) longitudinally asymmetric surface features.
+Most radial velocity practitioners think about the latter, since longitudinally asymmetric surface features imbue changing-in-time skewness to the line profiles, causing radial velocity perturbations easily detectable in radial velocity time series.
 These confound exoplanet searches.
@@
 Longitudinally symmetric surface features do not change as the star rotates on its axis.
-The existence of these features changes the \emph{kurtosis} of the spectral line.
+The existence of these features manifest as static-in-time \emph{kurtosis} of the spectral line.
 For example, a hypothetical non-emitting (black) polar starspot exhibits a deficit in flux at the line core, resulting in less zero-velocity flux than its homogeneous counterpart.
@@
 
-It is easier to reverse-engineer longitudinally asymmetric features than longitudinally symmetric ones.  
+It is easier to reverse-engineer longitudinally asymmetric features than longitudinally symmetric ones, since we assume that we occasionally catch a glimpse of the spot-free limbs and their pristine line profiles.
 The latter requires exact knowledge of the underlying spectral template.
-Isolated deep spectral lines constitute the only practical scenario where exact knowledge can be claimed.
+Isolated, deep, well-calibrated spectral lines constitute the only practical scenario where exact knowledge can plausibly be claimed.
 Isolated spectral lines may be scarce or absent for M-dwarfs and brown dwarfs where lines blend ostensibly in an inseparable way, counfounding Doppler imaging.
 
-\texttt{blas\'e} offer a new approach to Doppler imaging that may overcome these historical limitations.
+\texttt{blas\'e} offer a new approach to Doppler imaging that may overcome these historical limitations by simultaneously fitting both the imperfections in the underlying spectrum and its line profile perturbations.
+This approach is analogous to the linearized model in \citet{2021arXiv211006271L}, but with the benefit of also handling non-linear properties of the spectrum such as line widths, shapes, and locations, while also handling telluric contamination.
+Hypothetically the \citet{2021arXiv211006271L} approach could be partially absorbed into \texttt{blase}, or vice versa, though such a merger may be complicated to implement.
 
-\subsection{RV jitter and microturbulence}
+\subsection{Extreme Precision Radial Velocity}
 
 Already \texttt{blas\'e} is equipped to fit every single line with its own systemic $RV$, by tuning the line center position $\lambda_\mathrm{c}$ at each training epoch.
-There exists both empirical evidence and some theoretical motivation that $RV$ jitter varies from line-to-line \authorcomment1{Xavier Dumusque 2018}.
+There exists both empirical evidence and some theoretical motivation that $RV$ jitter varies from line-to-line \citep{2018A&A...620A..47D}.
 Importantly, the extent of this line-by-line RV jitter could be predicted in part by the depth of line formation.
-A future extreme-precision-RV (EPRV) version of \texttt{blase} could leverage this information in a few ways.
-As an example, the depth-of-formation for all or a subset of lines could be obtained and associated with each spectral line.
+A future extreme-precision-RV (EPRV) version of \texttt{blase} could leverage this information.
+As an example, the depth-of-formation for all (or a subset) of lines could be obtained and associated with each spectral line.
 A regularization scaling term could be introduced to allow the line positions to vary, but only in proportion to their depth of formation.
 
-\subsection{Prospects for closed-form computation of Jacobians}
-In principle, one could compute the Jacobians analytically since the partial derivatives of the Voigt function are known in closed-form.
-This analytical approach would cut out the autodiff approach, using the exact Jacobians to compute the Gradient Descent step.
-The analytical approach may be faster and/or possess lower memory overhead than the autodiff one.
-However, working out the chain-rule for the closed-form analytical Jacobians may become difficult or impossible once combined with the many augmentations that get applied at the joint telluric-and-stellar modeling step.
-So analytic Jacobians could principally accelerate the cloning step but would add significant implementation complexity at the important data-model comparison step.
-We, therefore, deem this extension somewhat diminishing returns on scarce developer time.
+\subsection{Circumstellar disk veiling}
+To-be-completed
 
-\subsection{Lineshapes choice and tradeoffs}
-\begin{outline}
-    \1 Faced with a question: what lineshape to use
-    \2 To First order the lines are Gaussian or Lorentzian
-    \2 To Second-order the lines are Voigt Profiles
-    \2 In detail the lines are the result of radiative transfer that can smear the Voigt profile
-    \2 In practice the high-res lines get convolved with instrumental profiles, so details don't matter too much
-    \1 Computational tradeoff of Lorentzian/Gaussian versus Pseudo-Voigt versus Voigt
-    \2 Additional challenge: Fadeeva function not in PyTorch
-    \1 We choose a pseudo-Voigt as a balance between computation and adequate accuracy
-\end{outline}
-
-\subsection{Questioning the assumptions of the method}
-
-\begin{outline}
-    \1 Is PseudoVoigt Adequate?
-    \1 What performance level is adequate? (depends on your application)
-    \1 What about PRV applications
-    \1 Are we using the GPU effectively? Are we memory bandwidth limited?
-    \1 Should we update the line center positions?
-\end{outline}
-
-\subsection{Limitations}
-\begin{outline}
-    \1 What to do when continuum is absent even at the native resolution of precomputed synthetic model?
-    \1 Relevant to brown dwarfs and possibly M dwarfs molecular bands
-    \1 Bandwidth limitations
-    \1 Interplay with tellurics
-\end{outline}
+\subsection{Accretion Veiling}
 
 \section{Conclusions}
-More placeholder text...
+to-be-completed...
 
@@
 
+\section{Comparison to existing spectroscopy frameworks}
+Several astronomical spectral frameworks share similar aims as \texttt{blas\'e}.
+These existing frameworks will have enduring value for the wide range of problems in the field of stellar spectroscopy.
+Here we scrutinize the differences among some of these approaches to clarify how this work fits in.
+
+The \texttt{specmatch} synthetic template matching tool produces noise-free nearest neighbor templates given an input spectrum \citep{2015PhDT........82P}.
+Several practical barriers limit the accuracy of using precomputed synthetic spectral models alone.
+First and foremost, real stars are usually more complicated than our simplified models of them.
+Real spectra often vary over more dimensions that our models do.
+Conspicuous examples of these hidden variables can be found in protostars: starspots, accretion veiling, dust extinction, and magnetic Zeeman splitting.
+Jointly modeling all of these phenomena alongside the intrinsic stellar photosphere is challenging.
+
+The empirical version, \texttt{specmatch-emp} \citep{2017ApJ...836...77Y} matched spectra better than the synthetic templates, but is still too rigid for some applications and requires the assembly of hundreds of standardized high signal-to-noise-ratio templates, ideally with low intrinsic rotational broadening.
+Such a large number of high-quality templates with high resolving power and low $v\sin{i}$ has not yet been established in the near-infrared.
+
+The \texttt{wobble} framework \citep{2019AJ....158..164B} modernized the construction of high-SNR templates to account for temporally variable telluric lines.
+The tool requires dozens of high-SNR spectra acquired at a range of Barycentric Earth Radial Velocities (BERVs).
+The final telluric-free combined spectrum would still have to be compared to models for absolute calibration or can be used out-of-the-box for precision relative RVs.
+The \texttt{wobble} framework also pioneered the off-label application of automatic differentiation frameworks---in this case \texttt{TensorFlow}---towards their physically-motivated use in stellar spectra.
+\texttt{blas\'e} can be viewed as an evaluable and interpretable super-resolution version of \texttt{wobble}, that accepts more bias in the bias-variance tradeoff.
+
+The \texttt{starfish} framework \citep{czekala15} provides a robust likelihood function for data-model comparisons and retires many of the problems in this domain.
+\texttt{starfish} pioneered the use of whole-spectrum fitting with resilience to model imperfections by addressing the problem of what to do when the underlying atomic and molecular data was wrong or approximate or missing.
+It has been extended to inferring starspot physical properties \citep{2017ApJ...836..200G}, measuring veiling in Class 0 protostars \citep{2018ApJ...862...85G}, and quantifying imperfections in brown dwarf models \citep{2021ApJ...921...95Z}.
+The Spectral Inference Crank \citep[\texttt{sick},][]{2016ApJS..223....8C} shares similar aims as \texttt{starfish}, and provides additional useful grid search capabilities.
+
+For very large bandwidths and very many spectral lines, the problem of identifying and cataloging line imperfections essentially becomes a book-keeping and continuum assignment problem.
+\texttt{blas\'e} and \texttt{starfish} provide different strategies for orchestrating the line-mismatch identification procedure, with each route having tradeoffs depending on the application.
+
+\section{Conceivable Technical Improvements}\label{technicalInnovations}
+
+\subsection{Exact instead of pseudo Voigt Profile}
+We currently employ the pseudo-Voigt profile for its low computational cost.
+We have a prototype exact-Voigt-Hjerting implementation following \citet{2022ApJS..258...31K}.
+We coarsely estimate that moving to this exact-Voigt implementation could decrease some residual regions by $\sim 30\%$, while increasing the computational cost by more than $10\times$ over the existing pseudo-Voigt approximation.
+The exact-Voigt-Hjerting implementation still outbids the higher cost of a direct numerical convolution of a Gaussian and Lorentzian profile.
+
+\subsection{Addressing the pseudo-continuum with Gaussian Process regression}
+
+We currently assume the input spectra are adequately normalized to the continuum.
+We have a few options to relax this assumption.
+We could  simply tune the $\mathsf{P}$ term that represents the wavelength-dependent pre-factor to Equation \ref{equation1}.
+Tuning $\mathsf{P}$ would correct for large-scale imperfections in the otherwise-fixed continuum flattening procedure.
+This change would be easy and effective, but has some challenge with model selection and flexibility: how to set the polynomial order to avoid over- and under-fitting.
+Gaussian Processes (GPs) offer many advantages for continuum fitting \citet{czekala15}.
+In short, a GP-likelihood relaxes the assumption that the continuum has been perfectly normalized, in favor of the more realistic statement ``the continuum has been coarsely normalized, with some characteristic-but-as-yet-unknown correlation and scale length and amplitude of the imperfections''.
+That statement translates to the following modification to Equation \ref{simpleLikelihood}:
+
+\begin{eqnarray}
+    \mathcal{L} =  \frac{1}{2}\mathsf{R^\intercal} \mathsf{C}^{-1} \mathsf{R} +\frac{1}{2}\ln{\det{\mathsf{C}}} \label{GPLikelihood}
+\end{eqnarray}
+
+where we introduce the covariance matrix $\mathsf{C}$, with its associated kernel and collection of typically 2-3 parameters.
+We anticipate that this GP likelihood would have the greatest impact on stellar stars with significant band-heads and line-blanketing: spectra with a so-called ``pseudo-continuum''.
+M-dwarfs and brown dwarfs fall into this challenging category.
+
+The main demerit of moving to a GP-likelihood is computational cost.
+Fortunately, a few efficient autodiff-aware implementations of GPs exist.
+The \texttt{celerite} algorithm \emph{celerit\'e} \citep{2017AJ....154..220F} has an exact backpropation implementation \citep{2018RNAAS...2...31F} that scales linearly with the number of data points.
+The \emph{celerit\'e} algorithm does not currently have a PyTorch implementation.
+The \texttt{GPyTorch} framework \citep{2018arXiv180911165G} has a large category of approximate and exact GPs that could be straight-forwardly dropped into \texttt{blas\'e}.
+Even still these GPs could increase the computation cost by of order $10\times$.
+
+\subsection{Minibatches and Stochastic Gradient Descent}
+
+Currently, each training epoch sees the entire dataset, a setup dubbed \emph{full-batch} gradient descent.
+An alternative scheme allows training with only a portion of the entire dataset at a time in \emph{minibatches}.
+The massive data volumes in modern Neural Network applications cannot fit into the GPU memory, so minibatches are a necessity.
+Our meager 1 MB dataset can easily fit into the GPU memory, but our model can be large if we have a large number of pixels or lines or both.
+So while minibatches may not be required due to data size limitations they may be useful for particularly large models.
+Minibatches also act as a form of regularization, the principal source of stochasticity in the Stochastic Gradient Descent algorithm, which tends to have better convergence than full-batch Gradient Descent \citep{2016arXiv160904747R}.
+
+We experimented with minibatches by assembling and evaluating only a portion of the dense $\bar{\bm{F}}$ matrix at a time, in minibatches.
+The choice to evaluate only a portion of lines at a time would mean the model is inaccurately evaluated at all wavelength pixels.
+Instead, we choose to evaluate all lines, but only on a random subset $N_{\mathrm{batch}}$ of the total pixels $N_s$, so that the model can eventually converge to exact at those points.
+All lines are allowed to update at each glimpse of a minibatch, but many lines with cores far from minibatch pixels will provide only weak information about how the loss scalar changes for their parameters.
+
+Overall minibatches as implemented above performed worse than the sparse implementation, with both lower accuracy and slower computation time.
+
+\subsection{Broad lines and advanced lineshapes}
+
+Some lines---such as those arising from hydrogen, sodium, potassium, and others---have extremely broad line wings, approaching larger than the $\sim6000$ pixels we allocate for the sparse implementation.
+These special lines should be handled separately from the weak lines, both from a computational performance perspective and an accuracy perspective.
+
+Extremely broad lines will exhibit truncation effects if the sparse window is small compared to the line wing size.
+The truncation effects will look like tophat functions severing the asymptotic wings, imbuing artificial step function kinks in the emulated spectrum.
+We can afford to increase the sparse window on a few, say $N_{broad}\sim20$ of the broadest spectral lines.
+We then construct and evaluate the entire dense matrix for those lines: $\sim 330\;000 \times 20$.
+The number of FLOPS in each category scales as about 6 Million for the 20 broad and dense lines versus about 36 Million for the sea of about 7000 narrow and sparse lines, depending on the exact choices for wing cuts and the number of lines.
+
+One could introduce advanced lineshapes for these $\sim20$ broad lines, perturbing the Voigt line-wings with a smooth wavelength-dependent correction term $\mathsf{G}$:
+
+\begin{eqnarray}
+    \mathsf{\tilde{V}(\bm{\lambda})} &=& \mathsf{V} \odot \mathsf{G}\\
+    \mathsf{G} &=& 1 + (e^{\alpha_j} - 1) \cdot \mathcal{S}\left(\frac{|\bm{\lambda}-\lambda_{c,j}| - \lambda_{t, j}}{b_j}\right)
+\end{eqnarray}
+
+where $\odot$ is again the element-wise product (\emph{a.k.a} Hadamard product), $\mathcal{S}$ is the sigmoid function, and we have introduced three new tunable parameters for each of the $j$ broad lines; $\lambda_t$ is the truncation wavelength, $b$ is a scale parameter for how slowly or how rapidly in wavelength-space the transition from non-Lorentzian proceeds, and $\alpha$ is a possibly negative stretch parameter that controls whether the line wing is sub- or super- Lorentzian.
+
+This functional form has a few advantages.
+It is smooth.
+The smoothness of the transition is controlled by a tunable parameter, $b$.
+It can handle either sub- or super-Lorentzian shapes.
+It is roughly based on the $\chi$-factor\authorcomment1{citation to Hartmann et al.
+1989, etc.}.
+In the limit $\lim_{a\to0} \mathsf{G}$, the lineshape becomes exactly Lorentzian.
+The sigmoid is efficiently implemented in PyTorch.
+Finally, it enforces that the perturbation only produces absorption and not emission profiles.
+
+\subsection{Using native line lists rather than clones}
+For many practitioners, the choice to clone precomputed synthetic models in the first place may seem roundabout: ``Why not just use the line lists?''.
+Adopting the line lists would have many advantages: it would provide chemical and molecular provenance tags.
+Metadata associated with the quality of the atomic and molecular data could be used to assign physics-informed regularization.
+Many other benefits would effortlessly accrue from adopting the native line lists.
+The \texttt{FAL} project (Cargile et al.
+\emph{in prep}) follows such a principled prescription.
+
+There exists at least a few demerits of adopting the line lists, and therefore supporting the \emph{blase} strategy.
+First, these line lists need to undergo expensive multi-level radiative transfer calculations in order to obtain their amplitudes, so adopting the line lists would mean a laborious and computationally expensive pursuit simply to get close to what has already been computed.
+Second, as the effective temperature scales to ultracool dwarfs (Figure \ref{fig_Nlines_vs_teff}) the number of lines sky-rockets, tending towards the billions for T-dwarfs.
+The methane line list alone \censorbox{Hargraves et al 2022} represents a prohibitive data volume.
+The ExoJAX and Radis \censorbox{Radis paper} libraries offer a breakthrough solution to the voluminous line list problem.
+Even still, \emph{blase} deals with the less pure but more practical ``pseudo line'' that gets closer to the astronomical observables anyways, and offers a middle ground between the extremes of interpretability and performance.
+
+\subsection{Wavelength dependent limb darkening}
+Currently the extrinsic model step possesses up to four parameters: the $v\sin{i}$ and $RV$, and 2 optional parameters for limb-darkening.
+These four parameters may adequately parameterize a star with a uniform stellar disk.
+Extremely high signal-to-noise-ratio spectra of rapidly rotating stars may require additional flexibility.
+The limb darkening is generally wavelength-dependent, and so a pan-chromatic spectrum may require a different limb darkening from the blue end to the red end.
+The limb darkening may instead depend on physical properties of the spectral line formation, such as physical depth of formation, and so the extent of limb-darkening may jump haphazardly from line-to-line-to-line, rather than as a predictably smooth function across wavelength.
+\texttt{blas\'e} could be built to handle such a seemingly pathological scenario by adding a vector of limb darkening parameters, one for each line.
+One would have to regularize the fits with some typical limb darkening and a heuristic penalty for departures from this mean.
+
+\emph{``Transfer-transfer learning''}
+To-be-completed
+
 \end{document}

### 2022-08-10 paper/paper1/ms.tex
The paper is beginning to look near complete, yay

@@
 \subsection{Accretion Veiling}
+To-be-completed
 
@@
 
+Blas\'e already performs very well under a wide range of cloning and transfer-learning tasks.
+However, some precision applications may demand even-more-strenuous performance than what the current implementation can accomodate.
+Here we describe some of these technical improvements, their design, and/or mention some science case they may unlock.
+
 \subsection{Exact instead of pseudo Voigt Profile}

### 2022-08-11 paper/paper1/ms.tex
Change name of the paper for now
add section on how to get to fundamental properties

@@
 \shortauthors{Gully \& Morley}
-\title{Automatic line-by-line semi-empirical models for high-grasp stellar and telluric spectra}
+\title{Precision line-by-line forward models for \'echelle spectra}
 
@@
 \subsection{Line-by-line fundamental parameter estimates}
-to-be-completed...
+The conversion of spectral line depths to spectral type\citep{1901AnHar..28..129C}---and by extension $T_\mathrm{eff}$ \citep{1925PhDT.........1P}---has occupied a large chunk of stellar spectroscopy in the last century, and the measurement of fundamental stellar properties remains one of the principal applications of astronomical spectroscopy.
+Blas\'e as it currently stands does not output fundamental properties, and in fact it takes them \emph{for granted}: you must specify \emph{e.g.} the $T_\mathrm{eff}$, $\log{g}$, [Fe/H], and [$\alpha$/Fe] of your PHOENIX template as Step Zero before that template gets cloned and warped to match data.
+So obtaining a $T_\mathrm{eff}$ estimate is impossible.
+
+Simply put, we seek a function, $f$:
+$$f_j(a_j, \sigma_j, \gamma_j) \to (T_\mathrm{eff}, \log{g}, [\mathrm{Fe}/\mathrm{H}])$$
+
+There are $N_{\mathrm{lines}}$ such functions---one for each spectral line---because each line has its own temperature, gravity, and metallicity dependence.
+For visualization simplicity purposes, we have assumed the PHOENIX grid is merely 3-dimensional ($T_\mathrm{eff}$, $\log{g}$, [Fe/H]), but the same logic applies to 4-D and higher, and it works for sparse and irregularly sampled grids too.
+
+Fascinatingly, there are two related ways forward towards this function.
+They both involved ensembles of spectra, such as a spectral atlas or library, and they both involve first obtaining the inverse function:
+
+\begin{eqnarray}
+    g_j(T_\mathrm{eff}, \log{g}, [\mathrm{Fe}/\mathrm{H}]) &\to& (a_j, \sigma_j, \gamma_j)
+\end{eqnarray}
+
+A purely model-based approach would start by cloning every single PHOENIX spectrum in the grid dimensions-of-interest, over the wavelength range-of-interest.
+The resulting product would be a line-by-line catalog of cloned spectral properties for each 3D coordinate on this grid.
+We could then assemble a heatmap of for how each $j^{th}$ line properties changes across this heatmap, $g_j$.
+In this way, we are reverse-engineering the \emph{e.g.} temperature dependence of each spectral line, as encoded by the PHOENIX atmosphere models.
+The final step would involve finding the neareast neighbor of each $j^{th}$ observed spectral line to each $j^{th}$ grid heatmap point; that is, finding a way to invert the function $g_j$ to get $f_j$.
+The information-weighted mean, median, or mode of these nearest neighbors could then be reported as a revised ``best-fit $T_\mathrm{eff}$'', for instance.
+
+A semi-empirical approach could improve on this purely model-based approach.
+We know that the PHOENIX line depths, widths, and shapes are imperfect, and so this ($T_\mathrm{eff}$, $\log{g}$, [Fe/H]) heatmap will have large flaws in the line-by-line properties: the contours are systematically too bright or dark.
+There are many conceivable ways to quantify these heatmap flaws.
+Blas\'e provides an expedient route.
+One can pull the heatmap towards the locus of points established by running blas\'e on benchmark systems.
+The known $T_\mathrm{eff}$ of such systems would anchor the trend.
+We suspect the directionality of the purely model based heatmap must be correct, simply the slope, offset, and concavity of the trends may be wrong.
+With new information from each benchmark, the heatmap would get lifted like a central tentpole propping up an under-supported tent.
+
+The construction and calibration of these ensembles of spectra represents a tremendous amount of work beyond the scope of this paper.
+But its inevitable creation represents a major breakthrough for the field of stellar astrophysics.
+It would be an extremely precise, accurate, fast, and interpretable way to measure the properties of stars based solely on their high bandwidth \'echelle spectrum.
+
+\subsection{Abundances}
 
@@
 It can handle either sub- or super-Lorentzian shapes.
-It is roughly based on the $\chi$-factor\authorcomment1{citation to Hartmann et al.
-1989, etc.}.
 In the limit $\lim_{a\to0} \mathsf{G}$, the lineshape becomes exactly Lorentzian.
@@
 Second, as the effective temperature scales to ultracool dwarfs (Figure \ref{fig_Nlines_vs_teff}) the number of lines sky-rockets, tending towards the billions for T-dwarfs.
-The methane line list alone \censorbox{Hargraves et al 2022} represents a prohibitive data volume.
-The ExoJAX and Radis \censorbox{Radis paper} libraries offer a breakthrough solution to the voluminous line list problem.
+The methane line list alone \citep{2020ApJS..247...55H} represents a prohibitive data volume.
+The ExoJAX and Radis \cite{2019JQSRT.222...12P,2021JQSRT.26107476V} libraries offer a breakthrough solution to the voluminous line list problem.
 Even still, \emph{blase} deals with the less pure but more practical ``pseudo line'' that gets closer to the astronomical observables anyways, and offers a middle ground between the extremes of interpretability and performance.

### 2022-08-11 paper/paper1/ms.tex
Add a bunch of text, probably too much, on conceivable science

@@
 
-\subsection{GPU and Autodiff specific considerations}
+\subsection{GPU and Autodiff specific considerations}\label{gpuConsiderations}
 
@@
 
+\subsection{Extreme Precision Radial Velocity} \label{eprv}
+
+Already \texttt{blas\'e} is equipped to fit every single line with its own systemic $RV$, by tuning the line center position $\lambda_\mathrm{c}$ at each training epoch.
+There exists both empirical evidence and some theoretical motivation that $RV$ jitter varies from line-to-line \citep{2018A&A...620A..47D}.
+Importantly, the extent of this line-by-line RV jitter could be predicted in part by the depth of line formation \citep{2021A&A...653A..43C, 2022A&A...664A..34A}.
+A future extreme-precision-RV (EPRV) version of \texttt{blase} could leverage this information.
+As an example, the depth-of-formation for all (or a subset) of lines could be obtained and associated with each spectral line.
+A regularization scaling term could be introduced to allow the line positions to vary, but only in proportion to their depth of formation.
+
 \subsection{Line-by-line fundamental parameter estimates}
-The conversion of spectral line depths to spectral type\citep{1901AnHar..28..129C}---and by extension $T_\mathrm{eff}$ \citep{1925PhDT.........1P}---has occupied a large chunk of stellar spectroscopy in the last century, and the measurement of fundamental stellar properties remains one of the principal applications of astronomical spectroscopy.
+The conversion of spectral line strengths to spectral type \citep{1901AnHar..28..129C}---and by extension line strengths to $T_\mathrm{eff}$ \citep{1925PhDT.........1P}---has occupied a large chunk of stellar spectroscopy in the last century.
+The measurement of fundamental stellar properties remains one of the principal applications of stellar spectroscopy.
 Blas\'e as it currently stands does not output fundamental properties, and in fact it takes them \emph{for granted}: you must specify \emph{e.g.} the $T_\mathrm{eff}$, $\log{g}$, [Fe/H], and [$\alpha$/Fe] of your PHOENIX template as Step Zero before that template gets cloned and warped to match data.
-So obtaining a $T_\mathrm{eff}$ estimate is impossible.
-
-Simply put, we seek a function, $f$:
+So obtaining a $T_\mathrm{eff}$ estimate from blas\'e is impossible, at least as disscussed so far.
+
+The clearest way forward would be to calibrate the blas\'e-derived line properties.  Simply put, we seek a set of functions, $f$, that relate the line properties back to fundamental properties:
 $$f_j(a_j, \sigma_j, \gamma_j) \to (T_\mathrm{eff}, \log{g}, [\mathrm{Fe}/\mathrm{H}])$$
@@
 Fascinatingly, there are two related ways forward towards this function.
-They both involved ensembles of spectra, such as a spectral atlas or library, and they both involve first obtaining the inverse function:
-
-\begin{eqnarray}
+They both involve ensembles of spectra---such as a spectral atlas, spectral sequence, or library---and they both involve first obtaining the inverse function:
+
+\begin{eqnarray}
+    g_j &\equiv& f^{-1}_j\\
     g_j(T_\mathrm{eff}, \log{g}, [\mathrm{Fe}/\mathrm{H}]) &\to& (a_j, \sigma_j, \gamma_j)
@@
 
-A purely model-based approach would start by cloning every single PHOENIX spectrum in the grid dimensions-of-interest, over the wavelength range-of-interest.
+A purely model-based approach would start by blas\'e-cloning every single PHOENIX spectrum in the grid dimensions-of-interest, over the wavelength range-of-interest.
 The resulting product would be a line-by-line catalog of cloned spectral properties for each 3D coordinate on this grid.
@@
 The construction and calibration of these ensembles of spectra represents a tremendous amount of work beyond the scope of this paper.
-But its inevitable creation represents a major breakthrough for the field of stellar astrophysics.
-It would be an extremely precise, accurate, fast, and interpretable way to measure the properties of stars based solely on their high bandwidth \'echelle spectrum.
+But its creation could an extremely precise, fast, interpretable and reasonably accurate way to measure the properties of stars based solely on their high bandwidth \'echelle spectra, albeit with the overall accuracy hinging on the accuracy of the stellar benchmark labels.
+The mechanics of this approach could be adapted to fit within existing pipelines such as SAPP \citep{2022A&A...658A.147G}.
+
+There exists a corrolary from the method described here to the approach of \citet{czekala15}.
+There, a 3D heatmap was created for relating the eigenweights of a PCA basis back to the stellar fundamental properties, for spectrum emulation purposes.
+Here, the 3D heatmap is created on a line-by-line basis, and is therefore interpretable.
+The PCA eigenspectra were generally uninterpretable, at least not easily.
+So in the limiting case of obtaining a densely calibrated semi-empirical heatmap for all spectral lines across all grid dimensions, we will have achieved a spectral emulator that would obviate the need to clone spectra in the first place.
+We would simply start with this powerful line-by-line emulator as a forward model, and go directly to the extrinsic warping and data-model comparison.
 
 \subsection{Abundances}
+
+Stellar abundance work involves precisely measuring line strengths (often reported as EWs) of different chemical constituents evinced in a star or substar's photosphere.
+Relating those EWs back to physics can be done in a few ways.
+Most easily, trends and patterns in the EWs can be assembled, and metal-rich and metal-poor clusters can be identified.
+There exists a precision/accuracy tradeoff in the measurement process, usually stemming from the placement of the continuum, or assumptions about line blending.
+Blas\'e offers an immediate solution to these challenges, since it starts from our best-guess for how nearby lines may be shaping the continuum.
+Blas\'e appears to be a gateway to extremely fast and intervention-free way to obtain abundances, potentially useful for large surveys like APOGEE \citep{2017AJ....154...94M}, Gaia-ESO \citep{2012Msngr.147...25G}, RAVE \citep{2006AJ....132.1645S,}, and more.
+
+This spirit resembles \texttt{ROBOSPECT} \citep{2013PASP..125.1164W}, and other approaches aimed at automatically measuring equivalent widths for a large number of stars.
+
+One missing piece is either to post-process the blas\'e-output EW catalog with chemical labels (C, N, O, Si, Ba, Eu, etc.), or incorporate that information directly into the structure of blas\'e.
+The latter case is intriguing.
+For example, PHOENIX may not yet be aware of the existence of some obscure Lanthanum (La) line, possibly because the line was not yet understood and incorporated at the time of PHOENIX computation.
+Rather than wait for PHOENIX to receive the necessary update, blas\'e could instantiate one or more such lines in the locations guided by the atomic line list.
+This line could get initialized with some small amplitude, and would get added onto the underlying PHOENIX source as yet another freely tunable line.
+In this way, all the other nearby lines would be used for assigning continuum and blending, but this free-standing Lanthanum line would acquire a value properly taking into account the impact of its neighbors.
 
@@
 
-\texttt{blas\'e} offer a new approach to Doppler imaging that may overcome these historical limitations by simultaneously fitting both the imperfections in the underlying spectrum and its line profile perturbations.
+\texttt{blas\'e} offers a new approach to Doppler imaging that may overcome these historical limitations by simultaneously fitting both the imperfections in the underlying spectrum and its line profile perturbations.
 This approach is analogous to the linearized model in \citet{2021arXiv211006271L}, but with the benefit of also handling non-linear properties of the spectrum such as line widths, shapes, and locations, while also handling telluric contamination.
@@
 
-\subsection{Extreme Precision Radial Velocity}
-
-Already \texttt{blas\'e} is equipped to fit every single line with its own systemic $RV$, by tuning the line center position $\lambda_\mathrm{c}$ at each training epoch.
-There exists both empirical evidence and some theoretical motivation that $RV$ jitter varies from line-to-line \citep{2018A&A...620A..47D}.
-Importantly, the extent of this line-by-line RV jitter could be predicted in part by the depth of line formation.
-A future extreme-precision-RV (EPRV) version of \texttt{blase} could leverage this information.
-As an example, the depth-of-formation for all (or a subset) of lines could be obtained and associated with each spectral line.
-A regularization scaling term could be introduced to allow the line positions to vary, but only in proportion to their depth of formation.
-
-\subsection{Circumstellar disk veiling}
-To-be-completed
-
-\subsection{Accretion Veiling}
-To-be-completed
+\subsection{Starspots and magnetic fields}
+One current assumption of starspot spectral decomposition is that the starspot spectrum itself resembles the stellar photosphere of a cooler star.
+This assumption appears adequate for detecting starspot spectra and measuring their physical properties \citep{gullysantiago17}.
+But to second order starspots should exhibit some spectral peculiarities that make them depart from a ``normal'' stellar photosphere.
+We may be probing deeper into the photosphere, and so the lines may experiencing higher pressure, with slightly different line widths.
+Or maybe the finite convective velocity shift can be directly seen as systematic shifts of the spectral lines (as stated in Section \ref{eprv}).
+
+Rather than applying fixed a mixture model of PHOENIX templates, one could adopt a mixture model of pre-cloned blas\'e models.
+Then, the imperfections in the starspot spectrum can be learned alongside the filling factor of the starspots.
+
+Often the most spotted stars are also rapidly rotating, and so the large $v\sin{i}$ blurs the blurs the starspot spectrum into the ambient photosphere.
+One experiment could measure the starspot spectral peculiarities more easily: obtain multi-epoch high resolution spectra of a pole-on (or nearly pole-on) young star, and monitor it for secular variations attributable to stellar activity cycles.
+Those cycles should confirm the ebb-and-flow of starspot area confirming the interpretation of two components as spots, and not binarity or some other anomaly.
+
+Relatedly, magnetic fields information could be incorporated into blas\'e as Land\'e $g-$factors, to allow magnetic sensitive lines to exhibit either Zeeman broadening or Zeeman splitting.
+This labelling could be conducted in the same way as described in the Abundances subsection.
+
+\subsection{Circumstellar disk and accretion veiling}
+Circumstellar disk veiling suppresses the strengths of all spectral lines \emph{en masse}, as the stellar photosphere gets outshined by a hot disk and/or envelope.
+
+The bulk properties of that disk/envelope could be incorporated with a simple physical model: a black body of temperature $T_\mathrm{disk}$ and solid angle $\Omega_{disk}$ \citep{2018ApJ...862...85G}.
+The spectral shape of the veiling would require sufficiently high bandwidth spectra, such as Xshooter \citep{2011A&A...536A.105V} or possibly IGRINS.
+The extent of veiling would be fit alongside all the other stellar spectral lines.
+An MCMC approach may outperform blas\'e in accuracy, since the choice of picking an underlying spectral template is degenerate with the derived veiling.
+
+Accretion veiling can be treated in an analogous way, but with some more complications.
+Emission lines can be easily incorporated as described in Section \ref{gpuConsiderations}.
+These emission lines could be initialized with a line list of Hydrogen lines, forbidden lines, and other conspicuous features.
+The shapes of those lines can be affected by winds and other physical phenomena, yielding a variety of physically interpretable line shapes \citep{2022arXiv220802940E}.
+Blas\'e could incorporate those lineshapes as templates shared among several lines originating from similar physical enviroments, but scaled and shifted based on the details of each individual line's radiative transfer properties.
 

### 2022-08-11 paper/paper1/ms.tex
add conclusions, almost there

@@
 Blas\'e offers an immediate solution to these challenges, since it starts from our best-guess for how nearby lines may be shaping the continuum.
-Blas\'e appears to be a gateway to extremely fast and intervention-free way to obtain abundances, potentially useful for large surveys like APOGEE \citep{2017AJ....154...94M}, Gaia-ESO \citep{2012Msngr.147...25G}, RAVE \citep{2006AJ....132.1645S,}, and more.
+Blas\'e appears to be a gateway to extremely fast and intervention-free way to obtain abundances, potentially useful for large surveys like APOGEE \citep{2017AJ....154...94M}, Gaia-ESO \citep{2012Msngr.147...25G}, RAVE \citep{2006AJ....132.1645S}, and more.
 
 This spirit resembles \texttt{ROBOSPECT} \citep{2013PASP..125.1164W}, and other approaches aimed at automatically measuring equivalent widths for a large number of stars.
-
-One missing piece is either to post-process the blas\'e-output EW catalog with chemical labels (C, N, O, Si, Ba, Eu, etc.), or incorporate that information directly into the structure of blas\'e.
+One missing piece here is either to post-process the blas\'e-output EW catalog with chemical labels (C, N, O, Si, Ba, Eu, etc.), or incorporate that information directly into the structure of blas\'e.
 The latter case is intriguing.
@@
 
-Accretion veiling can be treated in an analogous way, but with some more complications.
-Emission lines can be easily incorporated as described in Section \ref{gpuConsiderations}.
+Accretion veiling can be treated in an analogous way, but with some more complications.  Emission lines can be easily incorporated as described in Section \ref{gpuConsiderations}.  
 These emission lines could be initialized with a line list of Hydrogen lines, forbidden lines, and other conspicuous features.
@@
 \section{Conclusions}
-to-be-completed...
+We have introduced an line-by-line approach to forward modelling stellar, substellar, and telluric spectroscopic data.
+The approach relies on a key enabling technology, automatic differentiation, that enables nearly unlimited number of spectral lines to be fit simultaneously.
+We initialize these lines to match precomputed synthetic stellar spectra, achieving excellent performance, and lending some confidence that the approach has a capacity to capture a tremendous amount of information at once.
 

### 2022-08-12 paper/paper1/ms.tex
A complete read-through of the paper

@@
 The \emph{blas\'e} approach also acts as a deconvolution tool.
-It is suitable for Doppler Imaging scenarios with longitudinally symmetric surface features like bands or polar spots, which evade detection in differential techniques.
-Blas\'e can also create super-resolution semi-empirical templates useful for critically evaluating atomic and molecular line lists.
+The general purpose scaffolding may be extensible to many scientific applications, including Precision Radial Velocities, Doppler Imaging, chemical abundances for Galactic archaeology, line veiling, magnetic fields, and remote sensing.
+Blas\'e can create super-resolution semi-empirical templates useful for critically evaluating imperfections in atomic and molecular line lists.
 Its sparse-matrix architecture and GPU-acceleration make \emph{blas\'e} fast, with end-to-end training of an octave of bandwidth (50+ \'echelle orders) in under 1 minute.
@@
 The identification and understanding of these lines have defined a large category of astrophysics over the last century.
-The field grew from by-eye catalogs of stellar templates \citep{1901AnHar..28..129C} to quantifying the role of atomic ionization balance \citep{1925PhDT.........1P}, to modern synthetic forward models including millions or billions of lines \citep[\emph{e.g.}][]{husser13, 2021ApJ...920...85M}.
+The field grew from by-eye catalogs of stellar templates \citep{1901AnHar..28..129C} to quantifying the role of atomic ionization balance \citep{1925PhDT.........1P}, to modern synthetic forward models including millions or billions of lines \citep[\emph{e.g.}][]{husser13, 2021ApJ...920...85M, 2021JQSRT.26107476V}.
 As technology has improved, our data and models have become more voluminous, more precise, and more complicated.
@@
 
-The metaphorical holy grail of astronomical spectroscopy is a function that takes in an observed stellar spectrum and reports back the position, amplitude, width, and shape of all of its spectral lines, automatically, accurately, and precisely.
+The metaphorical holy grail of astronomical spectroscopy would be a function that takes in an observed stellar spectrum and reports back the position, amplitude, width, and shape of all of its spectral lines, automatically, accurately, and precisely.
 The function would go further.
-It would report back the systemic radial velocity ($RV$) and rotational broadening ($v\sin{i}$)and optionally fundamental stellar properties, including $T_{\mathrm{eff}}$, $\log{g}$, and $[\mathrm{Fe}/\mathrm{H}]$.
+It would report back the systemic radial velocity ($RV$) and rotational broadening ($v\sin{i}$) and fundamental stellar properties, including $T_{\mathrm{eff}}$, $\log{g}$, and $[\mathrm{Fe}/\mathrm{H}]$.
 Finally, the function would---in a feat of artificial intelligence---provide what it believes to be the interpretable generating function that produced this data in the first place, so that we may gain insights on future examples of this or other stars.
@@
 A few lines suffice.
-For those applications, human inspection of isolated lines and semi-automated equivalent width determination have been and will remain adequate.
+For those applications, human inspection of isolated lines and semi-automated equivalent width determination have been---and will remain---adequate.
 
@@
 In particular, data from high-\emph{grasp} \'echelle spectrographs possess simultaneously high spectral resolving power and high bandwidth, yielding tens of thousands or possibly millions of independent spectral resolution elements for each star, substar, or exoplanet.
-Those applications that seek to gain signal by ``stacking'' spectral lines or cross-correlating with templates can hypothetically gain huge boosts in the accessible signal-to-noise ratio compared to a single or few lines.
+Scientific applications that seek to gain signal by ``stacking'' spectral lines or cross-correlating with templates can hypothetically gain huge boosts in the accessible signal-to-noise ratio compared to a single or few lines.
 Most manual and semi-automated methods cannot take advantage of the entire spectral bandwidth, or rely on exact knowledge of the underlying templates and may fail to achieve the hypothetical promise of these high-bandwidth spectrographs \citep{2020AJ....160..198H}.
@@
 There exists a nearly circular reasoning: we need to know the extent of line profile perturbations to reveal the underlying spectral template, but we need the underlying spectral template to estimate the line profile perturbations.
-The approach introduced here offers a path forward.
+The approaches introduced here offer a path forward on these long-standing friction points.
 
@@
 The algorithm presented here focuses on modeling the spectra of stars and brown dwarfs.
-In Section \ref{methodology} we introduce the interpretable forward-model design and its \texttt{PyTorch}-based \citep{2019arXiv191201703P} implementation, \texttt{blas\'e}.
 Existing stellar models \citep[\emph{e.g.}][]{husser13} and substellar models \citep[\emph{e.g.}][]{2021ApJ...920...85M} laboriously solve for a self consistent thermal structure in the atmosphere given the copious opacity sources that themselves depend on temperture and pressure.
 Here we build upon that hard work by cloning pre-existing synthetic stellar or substellar models (Section \ref{methodology}), and optionally by cloning models of Earth's atmospheric ``telluric'' absorption (Section \ref{sectionTelluric}).
+We introduce the interpretable forward-model design and its \texttt{PyTorch}-based \citep{2019arXiv191201703P} implementation, \texttt{blas\'e}.
 In Section \ref{transferLearn}  we describe how to adapt both stellar and telluric cloned models simultaneously, using a transfer-learning step.
 We obtain semi-empirical models by comparing to real-world \'echelle data in Section \ref{secResults}.
-Finally, we compare the framework to alternatives (Section \ref{secDiscuss}) and chronicle many conceivable extensions for future work (Section \ref{secFutureWork}).
+Finally, we discuss perspectives on how to think of blas\'e (Section \ref{secDiscuss}) and chronicle many conceivable extensions for unlocking new science (Section \ref{secFutureWork}).
 
@@
 For the purposes of this paper, we will showcase examples from two well-known families of precomputed synthetic astronomical spectra: \texttt{PHOENIX} \citep{husser13} for stellar spectra $(T_{\mathrm{eff}}\in [2300, 10000]\;K)$ and \texttt{Sonora} \citep{2021ApJ...920...85M} for substellar spectra $(T_{\mathrm{eff}}\in [200, 2300]\;K)$.
-The algorithms in the framework may also work for precomputed synthetic spectra of reflected light exoplanets, supernovae, galaxies, or even further afield such as laboratory physical chemistry, plasma physics, or materials science.
+The algorithms in the framework may also work for precomputed synthetic spectra of reflected light exoplanets, supernovae, galaxies, or even further afield such as laboratory physical chemistry, plasma physics, materials science, or remote sensing.
 
 We place mild demands on the precomputed spectra.
-They should have sporadic regions of discernable continuum devoid of lines, and the continuum should vary smoothly in wavelength.
+They should have sporadic regions of discernable continuum devoid of lines, and the continuum shape should vary smoothly in wavelength.
 The spectral lines or pseudo-lines should be resolved (and not sub-sampled).
@@
 We have implemented both modes in \texttt{blas\'e}, allowing users to choose their preference.
-We only narrate the linear flux unit description in the main text of this document, for the sake of clarity, and since most practitioners may tend to think of flux in terms of linear flux units.
+We only narrate the linear flux unit description in the main text of this document for the sake of clarity, and since most practitioners may tend to think of flux in terms of linear flux units.
 The data-model comparison step will always take place in linear flux units, so the only operational difference is the behavior for deep and saturated lines.
@@
 IGRINS has two cameras for $H$ and $K$ band, with the combined spectrum spanning $14267-25217\;$\AA~ including the edge buffers and the region in-between the two cameras, all at a resolving power of $(R=45,000)$.
-The spectrograph acquisition, reduction, and post-processing steps yield data $\mathsf{D}(\bm{\lambda}_{D})$, where $\bm{\lambda}_{D}$ is the wavelength vector at the instrumental resolution and sampling of each instrument, generally much coarser than the resolution and sampling grid of the precomputed synthetic spectra.
+The spectrograph acquisition, reduction, and post-processing steps yield data $\mathsf{D}(\bm{\lambda}_{D})$, where $\bm{\lambda}_{D}$ is the wavelength vector at the instrumental resolution and sampling of each instrument, generally much coarser than the resolution and sampling of the precomputed synthetic spectra.
 The data wavelength vector may also contain gaps between \'echelle orders, whereas the precomputed wavelength coordinates are usually contiguous.
 HPF may have up to $2048\times28=$57,344 pixels, and IGRINS has typically about 75,000 pixels, after common trimming of noisy edge pixels and unusable telluric regions.
+Meanwhile, the HPF-truncated model spectra have $N_s=$335,849 native resolution samples, comparable to the IGRINS-truncated model spectra, $N_s=$330,052.
 
@@
 The number of lines monotonically increases towards cooler effective temperatures.
-The HPF-truncated spectra have $N_s=$335,849 native resolution samples, comparable to the IGRINS-truncated spectra, $N_s=$330,052.
-
-So far we have only one piece of information about the peaks: their location.
+
+So far we have only one piece of information about the spectral lines: their location.
 Next, we derive coarse properties about each detected peak: its amplitude and width, again using the prominence algorithms implemented in \texttt{scipy} \citep{2020SciPy-NMeth}.
@@
 
-The optimizer updates the $a_j, \sigma_j, \gamma_j$ parameters by a small fraction of the Jacobian---called the learning rate LR---towards the direction that would improve the fit, for all parameters simultaneously.
+The optimizer updates the $a_j, \sigma_j, \gamma_j$ parameters by a small fraction of the Jacobian---called the learning rate (LR)---towards the direction that would improve the fit, for all parameters \emph{simultaneously}.
 The Jacobian is calculated behind the scenes with automatic differentiation implemented as the so-called backpropagation algorithm or simply ``backprop'' \citep{2015arXiv150205767G}.
@@
 First, we make a few tweaks to the implementation for numerical purposes.
-We enforce that all Gaussian and Lorenzian widths are positive by tuning the natural log of the widths, and exponentiate them before inclusion in Equation \ref{equation1}.
+We enforce that all Gaussian and Lorenzian widths are positive by tuning the natural log of the widths, then exponentiating them before inclusion in Equation \ref{equation1}.
 Based on the minus sign in Equation \ref{equation1}, an amplitude $a_j$ could hypothetically take on either positive values (flux loss, absorption) or negative values (flux gain, emission).
@@
 
-The computational bottleneck occurs at the evaluation of Equation \ref{equation1}, which can be viewed as having a $N_{\mathrm{lines}}\times N_{S}$ matrix $\bm{\bar{F}}$ assembled by stacking each Voigt absoption profile $\mathsf{V}_j(\bm{\lambda}_s)$ on top of each other:
+The computational bottleneck occurs at the evaluation of Equation \ref{equation1}, which can be viewed as having an $N_{\mathrm{lines}}\times N_{S}$ matrix $\bm{\bar{F}}$ assembled by stacking each Voigt absoption profile $\mathsf{V}_j(\bm{\lambda}_s)$ on top of each other:
 
@@
 An element of this matrix, $F_{ji}$, will have the flux value for a given $j^{th}$ line at a given $i^{th}$ wavelength coordinate.
-Equation \ref{equation1} performs a type of matrix contraction, turning a $N_{\mathrm{lines}}\times N_{S}$ matrix into a length $N_{S}$ row vector.
+Equation \ref{equation1} performs a type of matrix contraction, turning an $N_{\mathrm{lines}}\times N_{S}$ matrix into a length $N_{S}$ row vector.
 The number of Floating Point Operations (FLOPS) scales with the number of entries in this matrix.
@@
 Often the regions of known, deep tellurics are simply discarded.
-In other cases, the lines are modeled with first principles line-by-line radiative transfer \citep[\emph{e.g.} \texttt{TelFit,}][]{2014AJ....148...53G} or through data-driven means \citep[\emph{e.g.} \texttt{wobble,}][]{2019AJ....158..164B}.
+In other cases, the lines are modeled with first principles line-by-line radiative transfer \citep[\emph{e.g.} \texttt{TelFit,}][]{2014AJ....148...53G, 2005JQSRT..91..233C} or through data-driven means \citep[\emph{e.g.} \texttt{wobble,}][]{2019AJ....158..164B}.
 The most demanding EPRV applications require a precision characterization of telluric lines that the astronomical community has not yet been able to achieve, and that may rival even the abilities of Earth Science practitioners.
@@
 We start with a precomputed synthetic telluric model, $\mathsf{T}$ with associated wavelength coordinates $\bm{\lambda}_T$.
-We employ a \texttt{TelFit} model, though any precomputed synthetic telluric model will work.
+We employ a \texttt{TelFit} model, though any precomputed synthetic telluric model will work, such as \texttt{MOLECFIT} \citep{2015A&A...576A..77}.
 The \texttt{TelFit} model does not contain any continuum sources of opacity, so we can skip the flattening procedure described in Equation \ref{eqnFlattening}.
@@
 
-The number of pixels in the telluric model can be chosen at the time of running a \texttt{TelFit} model.
+The number of pixels in the telluric model can be chosen at the time of running \texttt{TelFit}.
 Here we choose a spectral resolution $R\sim10^6$, adequate for resolving narrow telluric lines, and yielding about 2 million pixels across the entire HPF bandwidth.
@@
 The telluric clone shows a comparable level of performance.
-These residuals tend to pile up in local symmetrically balanced clusters that get cancelled out once convolved with coarser instrumental line profiles, so their overall effect at instrumental resolution is typically negligible.
+These residuals tend to pile up in local symmetrically balanced clusters that get cancelled out once convolved with coarser instrumental line profiles, so their overall effect at instrumental resolution is typically negligible: the clones are almost perfect.
 We identify three main categories of cloning flaws that may not be negligible depending on the science application.
@@
 Mathematically speaking, multiplication and convolution do not commute.
-While the distinction may seem negligible, it matters at the level of precisions sought in EPRV applications (Suvrath Mahadevan, \emph{priv.
+While the distinction may seem negligible, it matters at the level of precisions sought in EPRV applications (Suvrath Mahadevan, Arpita Roy, Sharon Xuesong Wang \emph{priv.
 comm.}).
@@
 The model is not perfect however, especially around line cores.
-These residiuals stem from a combination of causes.
+These residuals stem from a combination of causes.
 First, our finite regularization restricts line amplitudes and widths from straying too far from their values.
@@
 Virtually all lines should instead be considered pseudo lines, computed using the opacity agglomeration technique known as ``correlated-$k$''.
-The boundaries of lines for 2MASS J08173001$-$6155158 become even more amorphous in the presence of its moderately high $v\sin{i} = 22.5\pm 0.5 ~ \text{km/s}$ \citep{2022MNRAS.tmp.1421T} rotational broadening.
-This T6 IGRINS spectrum therefore represents an extreme end of a test-case, in which the underlying notions of \emph{blas\'e} go outside the comparatively safe stellar spectra.
-The application of blas\'e to 2MASS J08173001$-$6155158 can be viewed as conducting a deconvolution step simultaneously paired with its line-by-line inference capability in one-fell-swoop.
-
-We initialize \texttt{blas\'e} with the nearest Sonora template with $T_\mathrm{eff} = 1100~\text{K}$, $\log{g}=5.0$, and solar metallicity.
+The boundaries of lines for 2MASS J08173001$-$6155158 become even more amorphous in the presence of its moderately high $v\sin{i} = 22.5\pm 0.5 ~ \text{km/s}$ rotational broadening \citep{2022MNRAS.tmp.1421T}.
+This T6 IGRINS spectrum therefore represents an extreme end of a test-case, in which the underlying notions of \emph{blas\'e} go outside the comparatively safe assumptions in stellar spectra.
+The application of blas\'e to 2MASS J08173001$-$6155158 can be viewed as conducting a deconvolution step simultaneously paired with line-by-line inference.
+
+We initialize \texttt{blas\'e} with the nearest Sonora template with $T_\mathrm{eff} = 1100~\text{K}$, $\log{g}=5.0$, and solar metallicity \citep{2022MNRAS.tmp.1421T}.
 Figure \ref{fig_IGRINS_sonora_demo} shows the clone of this native spectrum near the peak of the $H-$band.
@@
 
-\emph{``Fancy deconvolution''}-- The blas\'e approach can be viewed as a fancy deconvolution procedure, in which the convolution kernels gets inferred alongside the imperfections of the model template.
+\emph{``Fancy deconvolution''}-- The blas\'e approach can be viewed as a fancy deconvolution procedure, in which the convolution kernels get inferred alongside the imperfections of the model template.
 This perspective may suit practitioners who wish to obtain a robust rotational broadening estimate in the presence of imperfect templates, and vice versa: robust templates in the presence of imperfect rotational broadening.
@@
 
-The scaffolding of \texttt{blase} is designed in a way to promote extensibility, so many scientific questions can be written down into the flexible spectral modeling approach presented in this paper.
+The scaffolding of \texttt{blas\'e} is designed in a way to promote extensibility, so many scientific questions can be written down in the flexible language presented in this paper.
 In some cases the flexibility, precision, speed, and ease-of-use may unlock new approaches to long-standing astrophysical questions and practical challenges.
@@
 The resulting product would be a line-by-line catalog of cloned spectral properties for each 3D coordinate on this grid.
-We could then assemble a heatmap of for how each $j^{th}$ line properties changes across this heatmap, $g_j$.
+We could then assemble a heatmap of how each $j^{th}$ line property changes across this heatmap, $g_j$.
 In this way, we are reverse-engineering the \emph{e.g.} temperature dependence of each spectral line, as encoded by the PHOENIX atmosphere models.
-The final step would involve finding the neareast neighbor of each $j^{th}$ observed spectral line to each $j^{th}$ grid heatmap point; that is, finding a way to invert the function $g_j$ to get $f_j$.
+The final step could involve finding the neareast neighbor of each $j^{th}$ observed spectral line to each $j^{th}$ grid heatmap point; that is, finding a way to invert the function $g_j$ to get $f_j$.
 The information-weighted mean, median, or mode of these nearest neighbors could then be reported as a revised ``best-fit $T_\mathrm{eff}$'', for instance.
@@
 Blas\'e provides an expedient route.
-One can pull the heatmap towards the locus of points established by running blas\'e on benchmark systems.
+One can pull the heatmap towards the locus of points established by running blas\'e on benchmark stars.
 The known $T_\mathrm{eff}$ of such systems would anchor the trend.
@@
 The construction and calibration of these ensembles of spectra represents a tremendous amount of work beyond the scope of this paper.
-But its creation could an extremely precise, fast, interpretable and reasonably accurate way to measure the properties of stars based solely on their high bandwidth \'echelle spectra, albeit with the overall accuracy hinging on the accuracy of the stellar benchmark labels.
+But its creation could yield an extremely precise, fast, interpretable and reasonably accurate way to measure the properties of stars based solely on their high bandwidth \'echelle spectra.
+The overall accuracy hinges on the accuracy of the stellar benchmark labels.
 The mechanics of this approach could be adapted to fit within existing pipelines such as SAPP \citep{2022A&A...658A.147G}.
@@
 
-\subsection{Abundances}
+\subsection{Abundances}\label{secAbundance}
 
@@
 Blas\'e offers an immediate solution to these challenges, since it starts from our best-guess for how nearby lines may be shaping the continuum.
-Blas\'e appears to be a gateway to extremely fast and intervention-free way to obtain abundances, potentially useful for large surveys like APOGEE \citep{2017AJ....154...94M}, Gaia-ESO \citep{2012Msngr.147...25G}, RAVE \citep{2006AJ....132.1645S}, and more.
+Blas\'e appears to be a gateway to extremely fast and intervention-free industrial-scale abundances, potentially useful for large surveys like APOGEE \citep{2017AJ....154...94M}, Gaia-ESO \citep{2012Msngr.147...25G}, RAVE \citep{2006AJ....132.1645S}, and more.
 
@@
 Doppler imaging attempts to reverse engineer the surface map from the extent to which observed line profiles depart from a pristine rotational broadening kernel.
-This reverse engineering step suffers from a vast number of geometrical degeneracies, but still provides useful constraints on stellar surfaces.
+This reverse engineering step suffers from a vast number of geometrical degeneracies, but still provides useful constraints on stellar surfaces \citet{2021arXiv211006271L}.
 We emphasize a distinction between A) longitudinally symmetric surface features, and B) longitudinally asymmetric surface features.
@@
 
-Longitudinally symmetric surface features do not change as the star rotates on its axis.
+Longitudinally symmetric surface features---on the other hand---do not change as the star rotates on its axis.
 The existence of these features manifest as static-in-time \emph{kurtosis} of the spectral line.
@@
 But to second order starspots should exhibit some spectral peculiarities that make them depart from a ``normal'' stellar photosphere.
-We may be probing deeper into the photosphere, and so the lines may experiencing higher pressure, with slightly different line widths.
+We may be probing deeper into the photosphere, and so the lines may experience higher pressure, with slightly different line widths.
 Or maybe the finite convective velocity shift can be directly seen as systematic shifts of the spectral lines (as stated in Section \ref{eprv}).
 
-Rather than applying fixed a mixture model of PHOENIX templates, one could adopt a mixture model of pre-cloned blas\'e models.
+Rather than applying a mixture model of fixed PHOENIX templates, one could adopt a mixture model of pre-cloned blas\'e models.
 Then, the imperfections in the starspot spectrum can be learned alongside the filling factor of the starspots.
 
-Often the most spotted stars are also rapidly rotating, and so the large $v\sin{i}$ blurs the blurs the starspot spectrum into the ambient photosphere.
-One experiment could measure the starspot spectral peculiarities more easily: obtain multi-epoch high resolution spectra of a pole-on (or nearly pole-on) young star, and monitor it for secular variations attributable to stellar activity cycles.
-Those cycles should confirm the ebb-and-flow of starspot area confirming the interpretation of two components as spots, and not binarity or some other anomaly.
-
-Relatedly, magnetic fields information could be incorporated into blas\'e as Land\'e $g-$factors, to allow magnetic sensitive lines to exhibit either Zeeman broadening or Zeeman splitting.
-This labelling could be conducted in the same way as described in the Abundances subsection.
+Often the most spotted stars are also rapidly rotating, and so the large $v\sin{i}$ blurs the starspot spectrum into the spot-free component of the photosphere.
+One experiment could measure the starspot spectral peculiarities more easily: obtainhigh resolution spectra of a pole-on (or nearly pole-on) young star where the lines will be sharp and the two components of the spectra should be more easily discernable.
+
+Relatedly, magnetic fields information could be incorporated into blas\'e as line-by-line Land\'e $g-$factors, to allow magnetic sensitive lines to exhibit either Zeeman broadening or Zeeman splitting.
+This labelling could be conducted in the same way as described in the Abundances subsection \citep{secAbundance}.
 
@@
 The bulk properties of that disk/envelope could be incorporated with a simple physical model: a black body of temperature $T_\mathrm{disk}$ and solid angle $\Omega_{disk}$ \citep{2018ApJ...862...85G}.
-The spectral shape of the veiling would require sufficiently high bandwidth spectra, such as Xshooter \citep{2011A&A...536A.105V} or possibly IGRINS.
+Revealing the spectral shape of the veiling would require sufficiently high bandwidth spectra, such as Xshooter \citep{2011A&A...536A.105V} or possibly IGRINS.
 The extent of veiling would be fit alongside all the other stellar spectral lines.
-An MCMC approach may outperform blas\'e in accuracy, since the choice of picking an underlying spectral template is degenerate with the derived veiling.
+However, an MCMC approach may outperform blas\'e in accuracy, since the choice of picking an underlying spectral template is partially degenerate with the derived veiling.
 
@@
 \section{Conclusions}
-We have introduced an line-by-line approach to forward modelling stellar, substellar, and telluric spectroscopic data.
-The approach relies on a key enabling technology, automatic differentiation, that enables nearly unlimited number of spectral lines to be fit simultaneously.
+We have introduced a line-by-line approach to forward modelling stellar, substellar, and telluric spectroscopic data.
+The approach relies on a key enabling technology, automatic differentiation, that allows nearly unlimited number of spectral lines to be forward modelled simultaneously.
 We initialize these lines to match precomputed synthetic stellar spectra, achieving excellent performance, and lending some confidence that the approach has a capacity to capture a tremendous amount of information at once.
@@
 \begin{acknowledgments}
+
+    MGS thanks Ian Czekala, Phill Cargile, Rodrigo Luger, Greg Herczeg, Will Best, Dan Foreman-Mackey, Meg Bedell, Mark Marley, Erwan Pannier, Dirk C.M.
+van den Bekerom, Dan Clemens, Dan Jaffe, Tom Greene, Adam Kraus, David Hogg, Greg Zeimann, Ben Montet, Christina Hedges, Brittany Miles, Arpita Roy, and Kevin Gullikson for conversations and resources that shaped his thinking on spectral calibration and telluric mitigation.
 
@@
     seaborn \citep{Waskom2021},
-    pytorch \citep{2019arXiv191201703P}}
+    pytorch \citep{2019arXiv191201703P},
+    muler \citep{2022JOSS....7.4302G}}
 

### 2022-08-12 paper/paper1/ms.tex
Tiny fixes to appendix

@@
 This change would be easy and effective, but has some challenge with model selection and flexibility: how to set the polynomial order to avoid over- and under-fitting.
-Gaussian Processes (GPs) offer many advantages for continuum fitting \citet{czekala15}.
+Gaussian Processes (GPs) offer many advantages for continuum fitting \citep{czekala15}.
 In short, a GP-likelihood relaxes the assumption that the continuum has been perfectly normalized, in favor of the more realistic statement ``the continuum has been coarsely normalized, with some characteristic-but-as-yet-unknown correlation and scale length and amplitude of the imperfections''.
@@
 where we introduce the covariance matrix $\mathsf{C}$, with its associated kernel and collection of typically 2-3 parameters.
-We anticipate that this GP likelihood would have the greatest impact on stellar stars with significant band-heads and line-blanketing: spectra with a so-called ``pseudo-continuum''.
+We anticipate that this GP likelihood would have the greatest impact on stars with significant band-heads and line-blanketing: spectra with a so-called ``pseudo-continuum''.
 M-dwarfs and brown dwarfs fall into this challenging category.
@@
 
-There exists at least a few demerits of adopting the line lists, and therefore supporting the \emph{blase} strategy.
+As already emphasized, there exist at least a few demerits of adopting the line lists, and therefore supporting the \emph{blase} strategy.
 First, these line lists need to undergo expensive multi-level radiative transfer calculations in order to obtain their amplitudes, so adopting the line lists would mean a laborious and computationally expensive pursuit simply to get close to what has already been computed.
@@
 The methane line list alone \citep{2020ApJS..247...55H} represents a prohibitive data volume.
-The ExoJAX and Radis \cite{2019JQSRT.222...12P,2021JQSRT.26107476V} libraries offer a breakthrough solution to the voluminous line list problem.
+The ExoJAX and Radis \citep{2019JQSRT.222...12P,2021JQSRT.26107476V} libraries offer a breakthrough solution to the voluminous line list problem.
 Even still, \emph{blase} deals with the less pure but more practical ``pseudo line'' that gets closer to the astronomical observables anyways, and offers a middle ground between the extremes of interpretability and performance.
@@
 
-\emph{``Transfer-transfer learning''}
-To-be-completed
-
 \end{document}

### 2022-08-12 paper/paper1/ms.tex
Edits from @astrocaroline printed paper draft on Aug 12, 2022

@@
 \shortauthors{Gully \& Morley}
-\title{Precision line-by-line forward models for \'echelle spectra}
+\title{Precision line-by-line forward models for \'echelle spectra\footnote{Open source code at \url{https://github.com/gully/blase}}}
 
@@
 The T6 ultracool dwarf 2MASS J08173001$-$6155158 was recently observed with IGRINS, revealing a rich spectroscopic atlas of molecules in its $T_\mathrm{eff} = 1060 \pm 50$ cloud-free atmosphere \citep{2022MNRAS.tmp.1421T}.
-The sea of moleculaar lines are so rich as to blur entirely the notion of isolated lines.
-Virtually all lines should instead be considered pseudo lines, computed using the opacity agglomeration technique known as ``correlated-$k$''.
+The sea of molecular lines are so rich as to blur entirely the notion of isolated lines.
+Many lines should instead be considered pseudo lines.
 The boundaries of lines for 2MASS J08173001$-$6155158 become even more amorphous in the presence of its moderately high $v\sin{i} = 22.5\pm 0.5 ~ \text{km/s}$ rotational broadening \citep{2022MNRAS.tmp.1421T}.
@@
 \emph{``Equivalent Width machine''}-- The area under the curve of each $j^{th}$ blas\'e line represents its (neighbor-weighted) Equivalent Width (EW).
-That means that blas\'e can automatically and quickly distill a massive high-banwidth spectrum into a catalog of line positions and EWs.
-Even more, the catalog would have some resilience to rotaional broadening, line blending, and other friction points that hamper such an industrial-scale approach with existing tooling.
+That means that blas\'e can automatically and quickly distill a massive high-bandwidth spectrum into a catalog of line positions and EWs.
+Even more, the catalog would have some resilience to rotational broadening, line blending, and other friction points that hamper such an industrial-scale approach with existing tooling.
 Current applications of Equivalent Widths rely on either isolated lines or---for multi-epoch applications---the assumption of non-variable neighboring lines.
@@
 We break up the themes into two categories: scientific extensions (this Section) and technical innovations (Section \ref{technicalInnovations}).
-We order the presentation of these extensions by perceived scientific impact:
 

### 2022-08-12 paper/paper1/ms.tex
Title brainstorm, conclusions, feedback by @astrocaroline, yay!

@@
 \shortauthors{Gully \& Morley}
-\title{Precision line-by-line forward models for \'echelle spectra\footnote{Open source code at \url{https://github.com/gully/blase}}}
+\title{An Interpretable Machine Learning Framework for Modeling High-Resolution Spectroscopic Data\footnote{Open source code at \url{https://github.com/gully/blase}}}
 
@@
 \emph{``A good enough substrate''}-- The imperfections in pre-computed stellar and substellar models have hampered their use in high-fidelity applications, where model flaws overwhelm the sigal-to-noise in the data.
-The adaptability of blas\'e can be viewed as increasing the number of applications that could conceivably benefit from the rich-albeit-imperfect information enoded in these model grid spectra.
+The adaptability of blas\'e can be viewed as increasing the number of applications that could conceivably benefit from the rich-albeit-imperfect information encoded in these model grid spectra.
 In the next section we illustrate how these adaptable models can serve as a subtrate for applications in which the imperfections in precomputed models would have otherwise precluded their use.
@@
 
-This spirit resembles \texttt{ROBOSPECT} \citep{2013PASP..125.1164W}, and other approaches aimed at automatically measuring equivalent widths for a large number of stars.
+\subsection{Identifying missing lines}\label{secMissingLines}
+The automatic abundance determination spirit of blas\'e resembles \texttt{ROBOSPECT} \citep{2013PASP..125.1164W}, and other approaches aimed at automatically measuring equivalent widths for a large number of stars.
 One missing piece here is either to post-process the blas\'e-output EW catalog with chemical labels (C, N, O, Si, Ba, Eu, etc.), or incorporate that information directly into the structure of blas\'e.
@@
 Doppler imaging attempts to reverse engineer the surface map from the extent to which observed line profiles depart from a pristine rotational broadening kernel.
-This reverse engineering step suffers from a vast number of geometrical degeneracies, but still provides useful constraints on stellar surfaces \citet{2021arXiv211006271L}.
+This reverse engineering step suffers from a vast number of geometrical degeneracies, but still provides useful constraints on stellar surfaces \citep{2021arXiv211006271L}.
 We emphasize a distinction between A) longitudinally symmetric surface features, and B) longitudinally asymmetric surface features.
@@
 Relatedly, magnetic fields information could be incorporated into blas\'e as line-by-line Land\'e $g-$factors, to allow magnetic sensitive lines to exhibit either Zeeman broadening or Zeeman splitting.
-This labelling could be conducted in the same way as described in the Abundances subsection \citep{secAbundance}.
+This labelling could be conducted in the same way as described in Section \citep{secMissingLines}.
 
@@
 \section{Conclusions}
-We have introduced a line-by-line approach to forward modelling stellar, substellar, and telluric spectroscopic data.
-The approach relies on a key enabling technology, automatic differentiation, that allows nearly unlimited number of spectral lines to be forward modelled simultaneously.
+We have introduced an interpretable machine learning approach to forward modelling stellar, substellar, and telluric spectroscopic data.
+The line-by-line approach approach relies on a key enabling technology, automatic differentiation, that allows a nearly unlimited number of spectral lines to be forward modelled simultaneously.
 We initialize these lines to match precomputed synthetic stellar spectra, achieving excellent performance, and lending some confidence that the approach has a capacity to capture a tremendous amount of information at once.
+
+We demo the framework on two sources: the K5 exoplanet host star WASP 69 using a precomputed PHOENIX model, and the T6 ultracool dwarf 2MASS J08173001$-$6155158 using a precomputed Sonora-Bobcat model.
+We discuss how blas\'e can be used to measure Equivalent Widths for thousands of lines automatically, understand line lists, measure rotation rates, generate surrogate models, and construct semi-empirical models.
+This tool could readily have applications across stellar and substellar astronomy, including for PRV work, stellar compositions, Doppler imaging, and stellar activity.
 

### 2022-08-15 paper/paper1/ms.tex
change to flat directory structure, all files in same directory

@@
     \centering
-    \includegraphics[width=0.45\textwidth]{figures/N_lines_vs_Teff_prom.png}
-    %\includegraphics[width=0.45\textwidth]{figures/N_lines_vs_Teff_prom.png}
-    \includegraphics[width=0.41\textwidth]{figures/line_density_demo.png}
+    \includegraphics[width=0.45\textwidth]{N_lines_vs_Teff_prom.png}
+    %\includegraphics[width=0.45\textwidth]{N_lines_vs_Teff_prom.png}
+    \includegraphics[width=0.41\textwidth]{line_density_demo.png}
     \caption{Scaling of prominence and density of spectral lines.
@@
     \centering
-    \includegraphics[width=1.0\textwidth]{figures/blase_flowchart03.pdf}
+    \includegraphics[width=1.0\textwidth]{blase_flowchart03.pdf}
     \caption{Visual flowchart of the \texttt{blas\'e} forward model.
@@
         \begin{pmatrix}
-            \cdots \includegraphics[height=1cm]{figures/voigt_01.pdf}  \cdots & \\
-            \cdots \includegraphics[height=1cm]{figures/voigt_02.pdf}  \cdots & \\
+            \cdots \includegraphics[height=1cm]{voigt_01.pdf}  \cdots & \\
+            \cdots \includegraphics[height=1cm]{voigt_02.pdf}  \cdots & \\
             \vdots                                                    & \\
-            \cdots \includegraphics[height=1cm]{figures/voigt_03.pdf} \cdots  &
+            \cdots \includegraphics[height=1cm]{voigt_03.pdf} \cdots  &
         \end{pmatrix}                        \\
         \bm{\bar{F}} & \mapsto \bm{\hat{F}} =\begin{pmatrix}
-                                                 \includegraphics[height=1cm]{figures/voigt_01b.pdf} & \\
-                                                 \includegraphics[height=1cm]{figures/voigt_02b.pdf} & \\
+                                                 \includegraphics[height=1cm]{voigt_01b.pdf} & \\
+                                                 \includegraphics[height=1cm]{voigt_02b.pdf} & \\
                                                  \vdots                                      & \\
-                                                 \includegraphics[height=1cm]{figures/voigt_03b.pdf} &
+                                                 \includegraphics[height=1cm]{voigt_03b.pdf} &
                                              \end{pmatrix}
@@
     \centering
-    \includegraphics[width=1.0\columnwidth]{figures/blase_clone_demo.png}
+    \includegraphics[width=1.0\columnwidth]{blase_clone_demo.png}
     \caption{PHOENIX spectrum cloned with blas\'e.
@@
     \centering
-    \includegraphics[width=0.98\textwidth]{figures/TelFit_clone_2panel.pdf}
+    \includegraphics[width=0.98\textwidth]{TelFit_clone_2panel.pdf}
     \caption{Pre-computed TelFit model cloned with \texttt{blase}.
@@
     \centering
-    \includegraphics[width=0.48\textwidth]{figures/blase_clone_demo_zoom.png}
+    \includegraphics[width=0.48\textwidth]{blase_clone_demo_zoom.png}
     \caption{Zoom in of the region between the vertical gray bars in Figure \ref{fig_cloned_spectrum_demo}.
@@
     \centering
-    \includegraphics[width=0.98\columnwidth]{figures/blase_WASP69_demo.png}
+    \includegraphics[width=0.98\columnwidth]{blase_WASP69_demo.png}
     \caption{WASP 69 observed with HPF, compared to a $T_{\mathrm{eff}}=4700\;$K,  $\log{g}=4.5$ solar metallicity PHOENIX model warped to $v\sin{i}=2.2$~km/s, $RV=-9.6$~km/s, and HPF resolving power.}
@@
     \centering
-    \includegraphics[width=0.98\columnwidth]{figures/blase_WASP69_regularized.png}
+    \includegraphics[width=0.98\columnwidth]{blase_WASP69_regularized.png}
     \caption{Semi-empirical model of WASP 69 transfer-learned with \texttt{blas\'e}, employing a regularization prior on the learned amplitudes.}
@@
     \centering
-    \includegraphics[width=0.98\columnwidth]{figures/blase_super_resolution_template.png}
+    \includegraphics[width=0.98\columnwidth]{blase_super_resolution_template.png}
     \caption{The native-resolution semi-empirical model transfer-learned from WASP 69 HPF data. The revised model can be viewed as a super-resolution deconvolution of the HPF spectrum.}
@@
     \centering
-    \includegraphics[width=0.95\textwidth]{figures/WASP_69_multi.pdf}
+    \includegraphics[width=0.95\textwidth]{WASP_69_multi.pdf}
     \caption{End-to-end training on a portion of HPF spectrum of WASP-69.
@@
     \centering
-    \includegraphics[width=0.98\textwidth]{figures/IGRINS_T6_sonora_demo.png}
+    \includegraphics[width=0.98\textwidth]{IGRINS_T6_sonora_demo.png}
     \caption{Line-by-line dissection of the 2MASS J08173001$-$6155158 IGRINS spectrum.

### 2022-08-15 paper/paper1/ms.tex
Shorten the abstract to meet 250 words, on CCA rooftop :)

@@
 This hybrid data-/model- driven framework allows joint modeling of stellar and telluric lines simultaneously, a potentially transformative step forwards for mitigating the deleterious telluric contamination in the near-infrared.
-The \emph{blas\'e} approach also acts as a deconvolution tool.
+The \emph{blas\'e} approach acts as both a deconvolution tool and semi-empirical model.
 The general purpose scaffolding may be extensible to many scientific applications, including Precision Radial Velocities, Doppler Imaging, chemical abundances for Galactic archaeology, line veiling, magnetic fields, and remote sensing.
-Blas\'e can create super-resolution semi-empirical templates useful for critically evaluating imperfections in atomic and molecular line lists.
-Its sparse-matrix architecture and GPU-acceleration make \emph{blas\'e} fast, with end-to-end training of an octave of bandwidth (50+ \'echelle orders) in under 1 minute.
-The open-source PyTorch-based code \texttt{blase} includes tutorials, deep dives, reproducible Python environments, and API documentation.
+Its sparse-matrix architecture and GPU-acceleration make \emph{blas\'e} fast.
+The open-source PyTorch-based code \texttt{blase} includes tutorials, Application Programming Interface (API) documentation, and more.
 We show how the tool fits into the existing Python spectroscopy ecosystem, demonstrate a range of astrophysical applications, and discuss limitations and future extensions.

### 2022-08-17 paper/paper1/ms.tex
fix one missing citation letter
Close Issue #50

@@
-\documentclass[twocolumn]{aastex631}
+\documentclass[linenumbers]{aastex631}
 \usepackage{hyperref}
@@
 We start with a precomputed synthetic telluric model, $\mathsf{T}$ with associated wavelength coordinates $\bm{\lambda}_T$.
-We employ a \texttt{TelFit} model, though any precomputed synthetic telluric model will work, such as \texttt{MOLECFIT} \citep{2015A&A...576A..77}.
+We employ a \texttt{TelFit} model, though any precomputed synthetic telluric model will work, such as \texttt{MOLECFIT} \citep{2015A&A...576A..77S}.
 The \texttt{TelFit} model does not contain any continuum sources of opacity, so we can skip the flattening procedure described in Equation \ref{eqnFlattening}.
@@
 Relatedly, magnetic fields information could be incorporated into blas\'e as line-by-line Land\'e $g-$factors, to allow magnetic sensitive lines to exhibit either Zeeman broadening or Zeeman splitting.
-This labelling could be conducted in the same way as described in Section \citep{secMissingLines}.
+This labelling could be conducted in the same way as described in Section \ref{secMissingLines}.
 

### 2022-08-17 src/blase/emulator.py
Dan @dfm was wondering what happens here

@@ -485,7 +485,7 @@ class SparseLinearEmulator(LinearEmulator):
         sparse_matrix = torch.sparse_coo_tensor(
             self.indices, ln_term, size=(self.n_pix,), requires_grad=True
         )
-
+        # The operation below does a COALESCE operation on the sparse matrix by default!
         result_1D = sparse_matrix.to_dense()
 
         return torch.exp(result_1D)

### 2022-08-17 src/blase/jaxemulator.py
Start of a JAX version of blasé

@@ -1,0 +1,437 @@
+r"""
+emulator
+--------------
+
+Precomputed synthetic spectral models are awesome but imperfect and rigid.  Here we clone the most prominent spectral lines and continuum appearance of synthetic spectral models to turn them into tunable, flexible, semi-empirical models.  We can ultimately learn the properties of the pre-computed models with a neural network training loop, and then transfer those weights to real data, where a second transfer-learning training step can take place. The spectrum has :math:`N_{\rm pix} \sim 300,000` pixels and :math:`N_{\rm lines} \sim 5000` spectral lines.  The number of lines is set by the `prominence=` kwarg: lower produces more lines and higher (up to about 0.3) produces fewer lines.  
+"""
+import math
+import jax
+import jax.numpy as jnp
+import numpy as np
+from scipy.signal import find_peaks, peak_prominences, peak_widths
+from tqdm import trange
+
+
+class LinearEmulator(object):
+    r"""
+    Model for cloning a precomputed synthetic spectrum in linear flux.
+
+    :math:`\mathsf{S} \mapsto \mathsf{S}_{\rm clone}`
+
+    Parameters
+    ----------
+    wl_native :  torch.tensor
+        The vector of input wavelengths at native resolution and sampling
+    flux_native : torch.tensor or None
+        The vector of continuum-flattened input fluxes.  If None, line-finding is skipped, init_state_dict is required, and the
+        optimize method does not work.
+    prominence : int or None
+        The threshold prominence for peak finding, defaults to 0.03.  Ignored if init_state_dict is provided.
+    init_state_dict : dict
+        A dictionary of model parameters to initialize the model with
+    """
+
+    def __init__(
+        self, wl_native, flux_native=None, prominence=None, init_state_dict=None
+    ):
+        super().__init__()
+
+        # Read in the synthetic spectra at native resolution
+        self.wl_native = jnp.array(wl_native)
+        self.wl_min = wl_native.min()
+        self.wl_max = wl_native.max()
+        self.n_pix = len(wl_native)
+
+        ## Set up "active area", where the region-of-interest is:
+        ## Restrict the lines to the active region plus 30 A buffer region
+        ## These are hardcoded, and care should be taken if changing them
+        line_buffer = 30  # Angstroms
[…]

### 2022-08-17 src/blase/jaxemulator.py
JAX specific tweaks

@@ -304,7 +304,7 @@ class SparseLinearEmulator(LinearEmulator):
         )
 
         if self.flux_native is not None:
-            self.target = jax.array(self.flux_active)
+            self.target = jnp.array(self.flux_active)
         else:
             self.target = None
 
@@ -325,24 +325,32 @@ class SparseLinearEmulator(LinearEmulator):
         # From that, determine the beginning and ending indices
         zero_indices = center_indices - (wing_cut_pixels // 2)
         too_low = zero_indices < 0
-        zero_indices[too_low] = 0
+
+        ## The next lines are a JAX workaround for item assignment:
+        # zero_indices[too_low] = 0 # can't do this in JAX
+        zero_indices = zero_indices.at[too_low].set(0)
+
         end_indices = zero_indices + wing_cut_pixels
         too_high = end_indices > self.n_pix
-        zero_indices[too_high] = self.n_pix - wing_cut_pixels - 1
-        end_indices[too_high] = self.n_pix - 1
+
+        ## The next lines are a JAX workaround for item assignment:
+        # zero_indices[too_low] = 0 # can't do this in JAX
+        zero_indices = zero_indices.at[too_high].set(self.n_pix - wing_cut_pixels - 1)
+        end_indices = end_indices.at[too_high].set(self.n_pix - 1)
 
         # Make a 2D array of the indices
         indices_2D = np.linspace(
             zero_indices, end_indices, num=wing_cut_pixels, endpoint=True
         )
 
-        self.indices_2D = jnp.array(indices_2D.T, dtype=jnp.long)
+        self.indices_2D = jnp.array(indices_2D.T, dtype=jnp.int32)
         self.indices_1D = self.indices_2D.reshape(-1)
-        self.indices = self.indices_1D.unsqueeze(0)
-        self.wl_2D = self.wl_native.to(device)[self.indices_2D]
-        self.wl_1D = self.wl_2D.reshape(-1)
-        self.active_mask = self.active_mask.to(device)
+        self.indices = np.expand_dims(self.indices_1D, axis=0)
 
+        ## TODO ... keep unsqueeze should be expand dims
+        self.wl_2D = self.wl_native[self.indices_2D]
+        self.wl_1D = self.wl_2D.reshape(-1)
+        self.active_mask = self.active_mask
         self.radial_velocity = jnp.array(0.0)
 
     def forward(self):

### 2022-08-18 src/blase/jaxemulator.py
holy moly it can make a forward model with JAX.  nice!

@@ -241,23 +241,23 @@ class LinearEmulator(object):
         ----------
         .. [1] https://en.wikipedia.org/wiki/Pseudo-Voigt_profile
         """
-        fwhm_G = 2.3548 * jnp.exp(self.sigma_widths).unsqueeze(1)
-        fwhm_L = 2.0 * jnp.exp(self.gamma_widths).unsqueeze(1)
+        fwhm_G = 2.3548 * jnp.expand_dims(jnp.exp(self.sigma_widths), 1)
+        fwhm_L = 2.0 * jnp.expand_dims(jnp.exp(self.gamma_widths), 1)
         fwhm = self._compute_fwhm(fwhm_L, fwhm_G)
         eta = self._compute_eta(fwhm_L, fwhm)
 
-        return jnp.exp(self.amplitudes).unsqueeze(1) * (
+        return jnp.exp(self.amplitudes)[:, None] * (
             eta
             * self._lorentzian_line(
-                self.lam_centers.unsqueeze(1),
-                jnp.exp(self.gamma_widths).unsqueeze(1),
-                wavelengths.unsqueeze(0),
+                self.lam_centers[:, None],
+                jnp.exp(self.gamma_widths)[:, None],
+                wavelengths[:, None],
             )
             + (1 - eta)
             * self._gaussian_line(
-                self.lam_centers.unsqueeze(1),
-                jnp.exp(self.sigma_widths).unsqueeze(1),
-                wavelengths.unsqueeze(0),
+                self.lam_centers[:, None],
+                jnp.exp(self.sigma_widths)[:, None],
+                wavelengths[:, None],
             )
         )
 
@@ -369,9 +369,9 @@ class SparseLinearEmulator(LinearEmulator):
         Returns:
             (torch.tensor): the 1D generative spectral model destined for backpropagation parameter tuning
         """
-        flux_2D = jnp.exp(self.amplitudes).unsqueeze(1) * self.gaussian_line(
-            self.lam_centers.unsqueeze(1),
-            jnp.exp(self.sigma_widths).unsqueeze(1),
+        flux_2D = jnp.exp(self.amplitudes)[:, None] * self.gaussian_line(
+            self.lam_centers[:, None],
+            jnp.exp(self.sigma_widths)[:, None],
             self.wl_2D,
         )
 
@@ -405,8 +405,8 @@ class SparseLinearEmulator(LinearEmulator):
         torch.tensor
[…]

### 2022-08-18 src/blase/jaxemulator.py
The JAX optax-based optimization works!

@@ -11,6 +11,8 @@ import numpy as np
 from scipy.signal import find_peaks, peak_prominences, peak_widths
 from tqdm import trange
 
+# jax.config.update("jax_enable_x64", True)
+
 
 class LinearEmulator(object):
     r"""
@@ -353,7 +355,7 @@ class SparseLinearEmulator(LinearEmulator):
         self.active_mask = self.active_mask
         self.radial_velocity = jnp.array(0.0)
 
-    def forward(self):
+    def forward(self, ln_amplitudes):
         r"""The forward pass of the sparse implementation--- no wavelengths needed!
 
         Returns
@@ -361,34 +363,9 @@ class SparseLinearEmulator(LinearEmulator):
         torch.tensor
             The 1D generative spectral model destined for backpropagation
         """
-        return self.sparse_pseudo_Voigt_model()
+        return self.sparse_pseudo_Voigt_model(ln_amplitudes)
 
-    def sparse_gaussian_model(self):
-        """A sparse Gaussian-only model
-
-        Returns:
-            (torch.tensor): the 1D generative spectral model destined for backpropagation parameter tuning
-        """
-        flux_2D = jnp.exp(self.amplitudes)[:, None] * self.gaussian_line(
-            self.lam_centers[:, None],
-            jnp.exp(self.sigma_widths)[:, None],
-            self.wl_2D,
-        )
-
-        flux_1D = flux_2D.reshape(-1)
-        ln_term = jnp.log(1 - flux_1D)
-
-        ## TODO fix this part!
-
-        # sparse_matrix = jnp.sparse_coo_tensor(
-        #    self.indices, ln_term, size=(self.n_pix,), requires_grad=True
-        # )
-
-        # result_1D = sparse_matrix.to_dense()
-
-        return ln_term
-
-    def sparse_pseudo_Voigt_model(self):
+    def sparse_pseudo_Voigt_model(self, ln_amplitudes):
         r"""A sparse pseudo-Voigt model
 
         The sparse matrix :math:`\hat{F}` is composed of the log flux
@@ -414,7 +391,7 @@ class SparseLinearEmulator(LinearEmulator):
             1.0 + self.radial_velocity / 299_792.458
         )
 
-        flux_2D = jnp.exp(self.amplitudes)[:, None] * (
+        flux_2D = jnp.exp(ln_amplitudes)[:, None] * (
             eta
             * self._lorentzian_line(
[…]

### 2022-08-19 src/blase/jaxemulator.py
work from on Thursday at CCA

@@ -124,27 +124,6 @@ class LinearEmulator(object):
 
         self.wl_normed = (self.wl_native - 10_500.0) / 2500.0
 
-    def product_of_pseudovoigt_model(self, wl):
-        r"""Return the Product of pseudo-Voigt profiles
-
-        The product acts like a matrix contraction:
-
-        .. math:: \prod_{j=1}^{N_{\mathrm{lines}}} 1-a_j \mathsf{V}_j(\lambda_S)
-
-        Parameters
-        ----------
-        wl : torch.tensor
-            The input wavelength :math:`\mathbf{\lambda}_S` at which to
-            evaluate the model
-
-        Returns
-        -------
-        torch.tensor
-            The 1D generative spectral model clone :math:`\mathsf{S}_{\rm clone}`
-            destined for backpropagation parameter tuning
-        """
-        return (1 - self.pseudo_voigt_profiles(wl)).prod(0)
-
     def detect_lines(self, wl_native, flux_native, prominence=0.03):
         """Identify the spectral lines in the native model
 
@@ -201,71 +180,6 @@ class LinearEmulator(object):
             + 0.07842 * fwhm_G**1 * fwhm_L**4
             + fwhm_L**5
         ) ** (1 / 5)
-
-    def pseudo_voigt_profiles(self, wavelengths):
-        r"""Compute the pseudo-Voigt Profile for a collection of lines
-
-        Much faster than the exact Voigt profile, but not as accurate:
-
-        .. math::
-
-            \mathsf{V}(\lambda_S-\lambda_{\mathrm{c},j}, \sigma_j, \gamma_j)
-
-
-        Parameters
-        ----------
-        wavelengths : torch.tensor
-            The 1D vector of wavelengths :math:`\mathbf{\lambda}_S` at which to
-            evaluate the model
-
-        Returns
-        -------
-        torch.tensor
-            The 1D pseudo-Voigt profiles
-
-        Notes
-        -----
-        The pseudo-Voigt [1]_ is an approximation to the convolution of a
-        Lorentzian profile :math:`L(\lambda,f)` and Gaussian profile :math:`G(\lambda,f)`
-
-        .. math::  V_p(\lambda,f) = \eta \cdot L(\lambda, f) + (1 - \eta) \cdot G(\lambda,f)
-
-        with mixture pre-factor:
-
[…]

### 2022-08-19 src/blase/jaxemulator.py
Deprecate linear emulator in favor of Sparse-only

@@ -14,30 +14,36 @@ from tqdm import trange
 # jax.config.update("jax_enable_x64", True)
 
 
-class LinearEmulator(object):
+class SparseLinearEmulator(object):
     r"""
-    Model for cloning a precomputed synthetic spectrum in linear flux.
-
-    :math:`\mathsf{S} \mapsto \mathsf{S}_{\rm clone}`
+    A sparse implementation of the LinearEmulator
 
     Parameters
     ----------
-    wl_native :  torch.tensor
-        The vector of input wavelengths at native resolution and sampling
-    flux_native : torch.tensor or None
-        The vector of continuum-flattened input fluxes.  If None, line-finding is skipped, init_state_dict is required, and the
-        optimize method does not work.
-    prominence : int or None
-        The threshold prominence for peak finding, defaults to 0.03.  Ignored if init_state_dict is provided.
+    wl_native : float vector
+        The input wavelength at native sampling
+    flux_native : float vector
+        The continuum-flattened flux at native sampling
+    prominence : int
+        The threshold for detecting lines
+    device : Torch Device or str
+        GPU or CPU?
+    wing_cut_pixels : int
+        The number of pixels centered on the line center to evaluate in the
+        sparse implementation, default: 1000 pixels
     init_state_dict : dict
         A dictionary of model parameters to initialize the model with
     """
 
     def __init__(
-        self, wl_native, flux_native=None, prominence=None, init_state_dict=None
+        self,
+        wl_native,
+        flux_native=None,
+        prominence=None,
+        device=None,
+        wing_cut_pixels=None,
+        init_state_dict=None,
     ):
-        super().__init__()
-
         # Read in the synthetic spectra at native resolution
         self.wl_native = jnp.array(wl_native)
         self.wl_min = wl_native.min()
@@ -124,6 +130,56 @@ class LinearEmulator(object):
 
         self.wl_normed = (self.wl_native - 10_500.0) / 2500.0
 
+        if self.flux_native is not None:
+            self.target = jnp.array(self.flux_active)
+        else:
+            self.target = None
+
+        ## Define the wing cut
+        # Currently defined in *pixels*
+        if wing_cut_pixels is None:
+            wing_cut_pixels = 1000
+        else:
+            wing_cut_pixels = int(wing_cut_pixels)
+
+        lines = self.lam_centers
[…]

### 2022-08-19 src/blase/jaxemulator.py
remove deprecated code from jax version

@@ -4,6 +4,7 @@ emulator
 
 Precomputed synthetic spectral models are awesome but imperfect and rigid.  Here we clone the most prominent spectral lines and continuum appearance of synthetic spectral models to turn them into tunable, flexible, semi-empirical models.  We can ultimately learn the properties of the pre-computed models with a neural network training loop, and then transfer those weights to real data, where a second transfer-learning training step can take place. The spectrum has :math:`N_{\rm pix} \sim 300,000` pixels and :math:`N_{\rm lines} \sim 5000` spectral lines.  The number of lines is set by the `prominence=` kwarg: lower produces more lines and higher (up to about 0.3) produces fewer lines.  
 """
+from dataclasses import dataclass
 import math
 import jax
 import jax.numpy as jnp
@@ -26,8 +27,6 @@ class SparseLinearEmulator(object):
         The continuum-flattened flux at native sampling
     prominence : int
         The threshold for detecting lines
-    device : Torch Device or str
-        GPU or CPU?
     wing_cut_pixels : int
         The number of pixels centered on the line center to evaluate in the
         sparse implementation, default: 1000 pixels
@@ -40,7 +39,6 @@ class SparseLinearEmulator(object):
         wl_native,
         flux_native=None,
         prominence=None,
-        device=None,
         wing_cut_pixels=None,
         init_state_dict=None,
     ):
@@ -174,7 +172,6 @@ class SparseLinearEmulator(object):
         self.indices_1D = self.indices_2D.reshape(-1)
         self.indices = np.expand_dims(self.indices_1D, axis=0)
 
-        ## TODO ... keep unsqueeze should be expand dims
         self.wl_2D = self.wl_native[self.indices_2D]
         self.wl_1D = self.wl_2D.reshape(-1)
         self.active_mask = self.active_mask

### 2022-09-20 paper/paper1/ms.tex
Start of adressing the referee report

@@
 
-where the division indicates element-wise division of these arrays or ``vectors'' of flux values.
+\noindent where the division indicates element-wise division of these arrays or ``vectors'' of flux values.
 
@@
 
-where $\mathsf{V}_j$ is the Voigt profile $\mathsf{V}(\bm{\lambda}_S-\lambda_{\mathrm{c},j}, \sigma_j, \gamma_j)$ with Gaussian standard deviation $\sigma$, Lorentzian half-width $\gamma$, at line center position $\lambda_c$, for the $j^{th}$ spectral line.
+\noindent where $\mathsf{V}_j$ is the Voigt profile $\mathsf{V}(\bm{\lambda}_S-\lambda_{\mathrm{c},j}, \sigma_j, \gamma_j)$ with Gaussian standard deviation $\sigma$, Lorentzian half-width $\gamma$, at line center position $\lambda_c$, for the $j^{th}$ spectral line.
 The amplitude $a$ is always expected to be positive for absorption lines.
@@
 
-where $\mathbf{1}$ is a $1\times N_{\rm lines}$ row vector of all-ones.
+\noindent where $\mathbf{1}$ is a $1\times N_{\rm lines}$ row vector of all-ones.
 We re\"emphasize that---in its current form---each spectral line has to be painstakingly evaluated across the entire spectral bandwidth.
@@
 
-where $\zeta$ is the convolution kernel for rigid body rotation \citep[\emph{e.g.}][]{2022ApJS..258...31K}, $\bm{v}$ is the spectral axis represented as relative velocity coordinates, and $*$ denotes the convolution operator.
+\noindent where $\zeta$ is the convolution kernel for rigid body rotation \citep[\emph{e.g.}][]{2022ApJS..258...31K}, $\bm{v}$ is the spectral axis represented as relative velocity coordinates, and $*$ denotes the convolution operator.
 
@@
 
-where $\bm{v}-v_0$ is the velocity vector going from zero to the velocity associated with the largest wavelength, with \emph{linear} spaced velocity samples.
+\noindent where $\bm{v}-v_0$ is the velocity vector going from zero to the velocity associated with the largest wavelength, with \emph{linear} spaced velocity samples.
 We choose a sampling in velocity space of 0.5 km/s, which corresponds to about $10\times$ finer than the instrumental resolving power of HPF, and delivers a minimum and maximum wavelength spacing of 0.013 and 0.024 \AA$/\text{pixel}$ respectively for the HPF bandwidth.
@@
 
-where $g$ is the Gaussian instrumental convolution kernel and the \texttt{resample[]} operation indicates the average of model pixels that fall within each data pixel's red and blue boundaries.
+\noindent where $g$ is the Gaussian instrumental convolution kernel and the \texttt{resample[]} operation indicates the average of model pixels that fall within each data pixel's red and blue boundaries.
 
@@
 
-where the hat notation demarcates the amplitudes obtained from cloning, \emph{i.e.} the ``initial, theory-inspired amplitude''.  We assign $\Lambda\sim5$.  The total, overall loss then becomes:
+\noindent where the hat notation demarcates the amplitudes obtained from cloning, \emph{i.e.} the ``initial, theory-inspired amplitude''.
+We assign $\Lambda\sim5$.
+The total, overall loss then becomes:
 
@@
 
-There exists a corrolary from the method described here to the approach of \citet{czekala15}.
+There exists a corollary from the method described here to the approach of \citet{czekala15}.
 There, a 3D heatmap was created for relating the eigenweights of a PCA basis back to the stellar fundamental properties, for spectrum emulation purposes.
@@
 The fixed $v\sin{i}$ approximation breaks down for stars with large-scale surface features.
-Doppler imaging attempts to reverse engineer the surface map from the extent to which observed line profiles depart from a pristine rotational broadening kernel.
-This reverse engineering step suffers from a vast number of geometrical degeneracies, but still provides useful constraints on stellar surfaces \citep{2021arXiv211006271L}.
+Doppler imaging attempts to infer the surface map from the extent to which observed line profiles depart from a pristine rotational broadening kernel.
+This inference procedure suffers from a vast number of geometrical degeneracies, but still provides useful constraints on stellar surfaces \citep{2021arXiv211006271L}.
 We emphasize a distinction between A) longitudinally symmetric surface features, and B) longitudinally asymmetric surface features.
@@
 
-where we introduce the covariance matrix $\mathsf{C}$, with its associated kernel and collection of typically 2-3 parameters.
+\noindent where we introduce the covariance matrix $\mathsf{C}$, with its associated kernel and collection of typically 2-3 parameters.
 We anticipate that this GP likelihood would have the greatest impact on stars with significant band-heads and line-blanketing: spectra with a so-called ``pseudo-continuum''.
@@
 
-where $\odot$ is again the element-wise product (\emph{a.k.a} Hadamard product), $\mathcal{S}$ is the sigmoid function, and we have introduced three new tunable parameters for each of the $j$ broad lines; $\lambda_t$ is the truncation wavelength, $b$ is a scale parameter for how slowly or how rapidly in wavelength-space the transition from non-Lorentzian proceeds, and $\alpha$ is a possibly negative stretch parameter that controls whether the line wing is sub- or super- Lorentzian.
+\noindent where $\odot$ is again the element-wise product (\emph{a.k.a} Hadamard product), $\mathcal{S}$ is the sigmoid function, and we have introduced three new tunable parameters for each of the $j$ broad lines; $\lambda_t$ is the truncation wavelength, $b$ is a scale parameter for how slowly or how rapidly in wavelength-space the transition from non-Lorentzian proceeds, and $\alpha$ is a possibly negative stretch parameter that controls whether the line wing is sub- or super- Lorentzian.
 

### 2022-09-20 paper/paper1/ms.tex
Address the SGD and optimization verbiage feedback

@@
-\documentclass[linenumbers]{aastex631}
+\documentclass[trackchanges]{aastex631}
 \usepackage{hyperref}
@@
     \includegraphics[width=0.45\textwidth]{N_lines_vs_Teff_prom.png}
-    %\includegraphics[width=0.45\textwidth]{N_lines_vs_Teff_prom.png}
     \includegraphics[width=0.41\textwidth]{line_density_demo.png}
@@
 
-\subsection{Goodness of fit metric}
+\subsection{Goodness of fit metric and gradients}
 The model evaluated with its coarse initial values would have terrible performance: it would only vaguely resemble the synthetic spectral model, with up to $\pm 50\%$ undulations from the inexact assignment of widths, lineshapes, and amplitudes.
@@
 As seen in Figure \ref{fig_Nlines_vs_teff}, the number of lines can exceed 7,000, meaning the clone model has over 21,000 free parameters.
-Fitting more than about 300 parameters is difficult with conventional optimizers that rely on minmizing a single scalar loss: the search space becomes too large.
-Here we employ a variant of Stochastic Gradient Descent (SGD), a transformative optimization technique that can scale to a virtually unlimited number of parameters \citep{2016arXiv160904747R}.
-This technique computes the derivative of the loss scalar with respect to each of the parameters, the so-called \emph{Jacobian}: $\grad \mathcal{L} = (\frac{\partial \mathcal{L}}{\partial a_j}, \frac{\partial \mathcal{L}}{\partial \sigma_j}, \cdots, \frac{\partial \mathcal{L}}{\partial \gamma_j})$.
-The Jacobian indicates how the MSE would decrease with a change in the parameter-of-interest, or put simply ``which way and by how much'' you have to change each individual line property to get a better fit.
+Fitting more than about 300 parameters is difficult with conventional optimizers\edit1{, and so historically the practice of finding the best settings for all 21,000 parameters would have been considered hopeless, a fool's errand.
+A new category of ``gradient-based'' optimizers offers a breakthrough.
+Gradient-based optimizers have virtually no restriction on the number of parameters, scaling to functions of possibly millions or billions of tuning parameters.
+Gradient-based optimizers get vastly more information per likelihood evaluation: $N_\mathrm{params}+1$ pieces of information-- the scalar value of the loss evaluated for some setting, \emph{and} the derivative of the loss scalar with respect to each of the parameters, the so-called \emph{Jacobian}: $\grad \mathcal{L} = (\frac{\partial \mathcal{L}}{\partial a_j}, \frac{\partial \mathcal{L}}{\partial \sigma_j}, \cdots, \frac{\partial \mathcal{L}}{\partial \gamma_j})$.}  The Jacobian indicates how the MSE would decrease with a change in the parameter-of-interest, or put simply ``which way and by how much'' you have to change each individual line property to get a better fit.
+
+\edit1{We adopt ``Full-batch'' Gradient Descent optimization, as opposed to the more popular Stochastic Gradient Descent \citep[SGD,][]{2016arXiv160904747R}.
+In Full-Batch Gradient Descent, the entire dataset is evaluated in Equation \ref{simpleLikelihood}, as opposed to only a portion of the dataset in the Stochastic counterpart.
+Appendix \ref{minibatches} discusses the tradeoffs and rationale for ``Full-Batch''.}
 
@@
 Without exact Jacobians a conventional optimization step only obtains \emph{one} piece of information: how much the overall loss changed.
-With exact Jacobians, we obtain $N_\mathrm{params}$ pieces of information for each evaluation of the forward model.
-The power of gradient descent becomes transformative as the number of parameters grows into the tens of thousands, as is the case for \emph{blas\'e}.
+With exact Jacobians, we obtain $N_\mathrm{params}+1$ pieces of information for each evaluation of the forward model.
+\edit1{ Conventional gradient-free optimizers evaluate the Jacobian through finite differences, which typically results in numerical instabilities.
+Those instabilities accrue and break down at around the 300 dimension mark.}  The power of gradient descent becomes transformative as the number of parameters grows into the tens of thousands, as is the case for \emph{blas\'e}.
 
@@
 
-\subsection{Minibatches and Stochastic Gradient Descent}
+\subsection{Minibatches and Stochastic Gradient Descent}\label{minibatches}
 

### 2022-09-20 paper/paper1/ms.tex
Referee report response ongoing...

@@
 An optional continuum flattening step ensures that subsequent spectral line finding steps get applied uniformly.
-This high-pass filtering step should be set to capture the genuine spectral shape, without over-fitting broad line wings such as those in deep Hydrogen and sodium lines.
+This high-pass filtering step should be set to capture the genuine spectral shape, without over-fitting broad line wings such as those in deep hydrogen and sodium lines.
 Any high-pass filter will work, a Gaussian Process approach would be ideal.
@@
 \end{figure}
-
-% TODO We pre-assign spectral lines for the conspicuous Hydrogen lines and neutral metal lines. Table
 
@@
 
+\edit1{Geometrically, the $\grad \mathcal{L}$ Jacobian resembles a compass needle guiding the direction to the best fit.
+This compass needle resides in, say, 21,000 dimensional space, instead of our mere 3 dimensions of physical space.
+Conventional gradient-free optimizers evaluate the Jacobian through approximate finite differences, which result in numerical instabilities.
+Those instabilities accrue and break down at around the 300 dimension mark.
+So non-gradient-based optimization can be pictured as having only a coarse needle---assembled from approximate finite differences---that begins to spin aimlessly after too many dimensions.}
+
 \edit1{We adopt ``Full-batch'' Gradient Descent optimization, as opposed to the more popular Stochastic Gradient Descent \citep[SGD,][]{2016arXiv160904747R}.
@@
 With exact Jacobians, we obtain $N_\mathrm{params}+1$ pieces of information for each evaluation of the forward model.
-\edit1{ Conventional gradient-free optimizers evaluate the Jacobian through finite differences, which typically results in numerical instabilities.
-Those instabilities accrue and break down at around the 300 dimension mark.}  The power of gradient descent becomes transformative as the number of parameters grows into the tens of thousands, as is the case for \emph{blas\'e}.
+The power of gradient descent becomes transformative as the number of parameters grows into the tens of thousands, as is the case for \emph{blas\'e}.
 
@@
 The most demanding EPRV applications require a precision characterization of telluric lines that the astronomical community has not yet been able to achieve, and that may rival even the abilities of Earth Science practitioners.
-A hybrid data-/model- driven approach was among the chief recommendations of the \emph{Telluric Hack Week} Workshop\footnote{\url{https://speakerdeck.com/dwhgg/telluric-line-hack-week-wrap-up}} aimed at improving mitigation of the atmosphere's deleterious effects (David W.
-Hogg, \emph{priv.
-comm.}).
+A hybrid data-/model- driven approach was among the chief recommendations of the \emph{Telluric Hack Week} Workshop\footnote{\url{https://speakerdeck.com/dwhgg/telluric-line-hack-week-wrap-up}} aimed at improving mitigation of the atmosphere's deleterious effects.
 The \texttt{blas\'e} framework achieves a key milestone by introducing a \emph{hybrid} approach to tellurics.
@@
 
-Finally, and most perniciously, a large category of residuals appear near the wings of the deepest and broadest lines---such as Hydrogen and neutral alkali metal lines.
+Finally, and most perniciously, a large category of residuals appear near the wings of the deepest and broadest lines---such as hydrogen and neutral alkali metal lines.
 The true lines exhibit advanced lineshapes, such as non-Lorentzian line wings that are not captured with the overly simplistic Voigt line profile.
@@
 The latter case is intriguing.
-For example, PHOENIX may not yet be aware of the existence of some obscure Lanthanum (La) line, possibly because the line was not yet understood and incorporated at the time of PHOENIX computation.
+For example, PHOENIX may not yet be aware of the existence of some obscure lanthanum (La) line, possibly because the line was not yet understood and incorporated at the time of PHOENIX computation.
 Rather than wait for PHOENIX to receive the necessary update, blas\'e could instantiate one or more such lines in the locations guided by the atomic line list.
 This line could get initialized with some small amplitude, and would get added onto the underlying PHOENIX source as yet another freely tunable line.
-In this way, all the other nearby lines would be used for assigning continuum and blending, but this free-standing Lanthanum line would acquire a value properly taking into account the impact of its neighbors.
+In this way, all the other nearby lines would be used for assigning continuum and blending, but this free-standing lanthanum line would acquire a value properly taking into account the impact of its neighbors.
 
@@
 Accretion veiling can be treated in an analogous way, but with some more complications.  Emission lines can be easily incorporated as described in Section \ref{gpuConsiderations}.
-These emission lines could be initialized with a line list of Hydrogen lines, forbidden lines, and other conspicuous features.
+These emission lines could be initialized with a line list of hydrogen lines, forbidden lines, and other conspicuous features.
 The shapes of those lines can be affected by winds and other physical phenomena, yielding a variety of physically interpretable line shapes \citep{2022arXiv220802940E}.
@@
 
-As already emphasized, there exist at least a few demerits of adopting the line lists, and therefore supporting the \emph{blase} strategy.
+As already emphasized, there exist at least a few demerits of adopting the line lists, and therefore supporting the \emph{blas\'e} strategy.
 First, these line lists need to undergo expensive multi-level radiative transfer calculations in order to obtain their amplitudes, so adopting the line lists would mean a laborious and computationally expensive pursuit simply to get close to what has already been computed.

### 2022-09-20 paper/paper1/ms.tex
standardize the use of blase italics or texttt

@@
 The mere act of comparing models to observed spectra can now resemble a computational statistics challenge as much as a scientific one.
-Here we introduce a new machine-learning-based framework \texttt{blas\'e} aimed at solving computational, statistical, and scientific challenges associated with data-model comparisons for modern astronomical spectroscopy.
+Here we introduce a new machine-learning-based framework \emph{blas\'e} aimed at solving computational, statistical, and scientific challenges associated with data-model comparisons for modern astronomical spectroscopy.
 
@@
 Here we build upon that hard work by cloning pre-existing synthetic stellar or substellar models (Section \ref{methodology}), and optionally by cloning models of Earth's atmospheric ``telluric'' absorption (Section \ref{sectionTelluric}).
-We introduce the interpretable forward-model design and its \texttt{PyTorch}-based \citep{2019arXiv191201703P} implementation, \texttt{blas\'e}.
+We introduce the interpretable forward-model design and its \texttt{PyTorch}-based \citep{2019arXiv191201703P} implementation, \texttt{blase}.
 In Section \ref{transferLearn}  we describe how to adapt both stellar and telluric cloned models simultaneously, using a transfer-learning step.
 We obtain semi-empirical models by comparing to real-world \'echelle data in Section \ref{secResults}.
-Finally, we discuss perspectives on how to think of blas\'e (Section \ref{secDiscuss}) and chronicle many conceivable extensions for unlocking new science (Section \ref{secFutureWork}).
+Finally, we discuss perspectives on how to think of \emph{blas\'e} (Section \ref{secDiscuss}) and chronicle many conceivable extensions for unlocking new science (Section \ref{secFutureWork}).
 
@@
 Adopting the log of the flux would ensure that the cloned model possesses only positive flux values, a desirable trait of any physical spectral model.
-We have implemented both modes in \texttt{blas\'e}, allowing users to choose their preference.
+We have implemented both modes in \texttt{blase}, allowing users to choose their preference.
 We only narrate the linear flux unit description in the main text of this document for the sake of clarity, and since most practitioners may tend to think of flux in terms of linear flux units.
@@
 
-\subsection{The \texttt{blas\'e} Stellar Clone Model}
-
-We have now arrived at the \texttt{blas\'e} clone model $\mathsf{S}_{\rm clone}(\bm{\lambda}_S)$ for a flattened synthetic spectrum $\mathsf{S}$: it is the cumulative product of transmission through the sea of all overlapping spectral lines:
+\subsection{The \emph{blas\'e} Stellar Clone Model}
+
+We have now arrived at the \emph{blas\'e} clone model $\mathsf{S}_{\rm clone}(\bm{\lambda}_S)$ for a flattened synthetic spectrum $\mathsf{S}$: it is the cumulative product of transmission through the sea of all overlapping spectral lines:
 
@@
     \includegraphics[width=1.0\textwidth]{blase_flowchart03.pdf}
-    \caption{Visual flowchart of the \texttt{blas\'e} forward model.
+    \caption{Visual flowchart of the  \emph{blas\'e} forward model.
 \textbf{Step 0} (not shown) is to choose a precomputed synthetic stellar spectrum---and, optionally, a precomputed synthetic telluric spectrum---with physical properties close to the target and observing conditions.
@@
 The Jacobian is calculated behind the scenes with automatic differentiation implemented as the so-called backpropagation algorithm or simply ``backprop'' \citep{2015arXiv150205767G}.
-We choose the \texttt{PyTorch} framework that computes these Jacobians efficiently for all of the mathematical primitives in our \texttt{blas\'e} implementation \citep{2019arXiv191201703P}.
+We choose the \texttt{PyTorch} framework that computes these Jacobians efficiently for all of the mathematical primitives in our \texttt{blase} implementation \citep{2019arXiv191201703P}.
 
@@
     \includegraphics[width=1.0\columnwidth]{blase_clone_demo.png}
-    \caption{PHOENIX spectrum cloned with blas\'e.
+    \caption{PHOENIX spectrum cloned with \emph{blas\'e}.
 This $T_{\mathrm{eff}}=4700\;$K,  $\log{g}=4.5$ solar metallicity model has 9,028 individual cloned spectral lines, each with 3 tuned parameters.
@@
 
-Figure \ref{fig_cloned_spectrum_demo} shows a portion of a PHOENIX spectrum cloned with \texttt{blas\'e}.
+Figure \ref{fig_cloned_spectrum_demo} shows a portion of a PHOENIX spectrum cloned with \texttt{blase}.
 The $1000$ epochs of training took 56 seconds on an NVIDIA\textsuperscript{\tiny\textregistered} RTX2070 GPU with \texttt{PyTorch v1.11}, \texttt{CUDA v11.1}, and Intel\textsuperscript{\tiny\textregistered} Core\textsuperscript{\tiny TM} i7-9750H CPUs at 2.60GHz, with all tensors as FP64.
@@
 A hybrid data-/model- driven approach was among the chief recommendations of the \emph{Telluric Hack Week} Workshop\footnote{\url{https://speakerdeck.com/dwhgg/telluric-line-hack-week-wrap-up}} aimed at improving mitigation of the atmosphere's deleterious effects.
-The \texttt{blas\'e} framework achieves a key milestone by introducing a \emph{hybrid} approach to tellurics.
-
-\subsection{The \texttt{blas\'e} Telluric Clone Model}
+The  \emph{blas\'e} framework achieves a key milestone by introducing a \emph{hybrid} approach to tellurics.
+
+\subsection{The  \emph{blas\'e} Telluric Clone Model}
 We start with a precomputed synthetic telluric model, $\mathsf{T}$ with associated wavelength coordinates $\bm{\lambda}_T$.
@@
 \textbf{Before:} The top panel shows a forward model evaluated with coarsely initialized line-by-line properties, based on simple threshold-based peak-finding and crude estimation of the shape of the line.
-\textbf{After:} The bottom panel shows the \emph{blas\'e} fine-tuned clone, with nearly indentical spectral structure.}
+\textbf{After:} The bottom panel shows the \texttt{blase} fine-tuned clone, with nearly indentical spectral structure.}
     \label{fig_telluric_clone}
@@
 We are left with $\mathsf{T}_{\rm clone}(\bm{\lambda}_T)$, the tunable telluric clone model evaluated at its original native coordinates.
-Figure \ref{fig_telluric_clone} shows a before-and-after view of the blas\'e cloning procedure for a TelFit model zoomed in on a $J-$band region with dense Telluric lines.
+Figure \ref{fig_telluric_clone} shows a before-and-after view of the  \emph{blase} cloning procedure for a TelFit model zoomed in on a $J-$band region with dense Telluric lines.
 We see that the blas\'e fine-tuning appears nearly indistinguishable from the native resolution TelFit precomputed model.
@@
 There exists a special exponential spectral sample spacing that allows the convolution operators to work out of the box for rotational broadening.
-The design of \texttt{blas\'e} permits the stellar and telluric models to be re\"evaluated at any wavelength coordinate vector, provided that it adequately samples the underlying lines.
+The design of \emph{blas\'e} permits the stellar and telluric models to be re\"evaluated at any wavelength coordinate vector, provided that it adequately samples the underlying lines.
 We, therefore, change the sampling from the native stellar and telluric wavelength coordinate grids, $\bm{\lambda}_S$ and $\bm{\lambda}_T$, to this special exponentially sampled wavelength grid, denoted with the subscript $Z$:
@@
 However, demanding Extreme Precision Radial Velocity (EPRV) exoplanet searches may need to consider minuscule systematic $RV$ shifts and broadening of the telluric templates, arising from the turbulent and bulk motions of the Earth's atmosphere.
-A $\mathsf{T}_{\rm ext}$ could hypothetically be implemented in \emph{blas\'e} to achieve these strenuous precision demands.
+A $\mathsf{T}_{\rm ext}$ could hypothetically be included in  \emph{blas\'e} to achieve these strenuous precision demands.
 
 \subsection{Joint Stellar and Telluric Model}
-Figure \ref{blase_flowchart} shows a visual guide to all the steps in \texttt{blase}.
+Figure \ref{blase_flowchart} shows a visual guide to all the steps in  \emph{blas\'e}.
 We have arrived at what may be the most intriguing-and-yet-simple of these steps: we simply multiply the rotationally-broadened-and-$RV$-shifted stellar model by the telluric transmission:
@@
 The extent to which this effect matters will depend on the science application.
-For now, \texttt{blas\'e} simply assumes a fixed resolving power.
+For now, \texttt{blase} simply assumes a fixed resolving power.
 
@@
 The most extreme regularization---the antithesis of the \emph{laissez-faire} scenario---would yield a model too rigid to respond to the data at all, yielding a model entirely unchanged from the cloned PHOENIX and \texttt{TelFit} models, the extreme end of underfitting.
-So regularization constitutes the only hyperparameters worthy of tuning in \texttt{blas\'e}.
+So regularization constitutes the only hyperparameters worthy of tuning in \emph{blas\'e}.
 The choice of how to set the regularization is problem-specific.
@@
 The low $v\sin{i}$ of WASP 69 means that the line profile broadening arises principally from the finite instrumental resolution.
-The interplay of telluric lines and stellar lines within an instrumental resolution element is exactly one of the challenges blas\'e was designed to solve, as discussed in Section \ref{sectionTelluric}.
+The interplay of telluric lines and stellar lines within an instrumental resolution element is exactly one of the challenges  \emph{blas\'e} was designed to solve, as discussed in Section \ref{sectionTelluric}.
 In Figure \ref{fig_multi_panel_WASP69} we show a multi-panel dissection of the HPF spectrum in a wavelength region in which conspicuous stellar and telluric lines coexist.
-The blas\'e end-to-end model exhibits excellent agreement with the data.
+The  \texttt{blase} end-to-end model exhibits excellent agreement with the data.
 The trustworthiness of the line properties inferred in locations where stellar and telluric lines exactly overlap remains an open research question: a tie-breaker is needed to distinguish these overlapping line inferrences.
@@
 This T6 IGRINS spectrum therefore represents an extreme end of a test-case, in which the underlying notions of \emph{blas\'e} go outside the comparatively safe assumptions in stellar spectra.
-The application of blas\'e to 2MASS J08173001$-$6155158 can be viewed as conducting a deconvolution step simultaneously paired with line-by-line inference.
-
-We initialize \texttt{blas\'e} with the nearest Sonora template with $T_\mathrm{eff} = 1100~\text{K}$, $\log{g}=5.0$, and solar metallicity \citep{2022MNRAS.tmp.1421T}.
+The application of  \emph{blas\'e} to 2MASS J08173001$-$6155158 can be viewed as conducting a deconvolution step simultaneously paired with line-by-line inference.
+
+We initialize \texttt{blase} with the nearest Sonora template with $T_\mathrm{eff} = 1100~\text{K}$, $\log{g}=5.0$, and solar metallicity \citep{2022MNRAS.tmp.1421T}.
 Figure \ref{fig_IGRINS_sonora_demo} shows the clone of this native spectrum near the peak of the $H-$band.
@@
 
-Up to this point we have presented the mechanics of blas\'e, and shown that it applies to a wide range of applications including stars, substars, and Earth's atmosphere.
+Up to this point we have presented the mechanics of  \emph{blas\'e}, and have shown that it applies to a wide range of applications including stars, substars, and Earth's atmosphere.
 The interpretation of the approach and its line-by-line outputs may carry different meaning to different practitioners.
@@
 
-\emph{``Equivalent Width machine''}-- The area under the curve of each $j^{th}$ blas\'e line represents its (neighbor-weighted) Equivalent Width (EW).
-That means that blas\'e can automatically and quickly distill a massive high-bandwidth spectrum into a catalog of line positions and EWs.
+\emph{``Equivalent Width machine''}-- The area under the curve of each $j^{th}$  \emph{blas\'e} line represents its (neighbor-weighted) Equivalent Width (EW).
+That means that  \emph{blas\'e} can automatically and quickly distill a massive high-bandwidth spectrum into a catalog of line positions and EWs.
 Even more, the catalog would have some resilience to rotational broadening, line blending, and other friction points that hamper such an industrial-scale approach with existing tooling.
 Current applications of Equivalent Widths rely on either isolated lines or---for multi-epoch applications---the assumption of non-variable neighboring lines.
-Blas\'e does not have these limitations.
+\emph{Blas\'e} does not have these limitations.
 The interpretations of these EWs may have to be treated with care if intercompared with EWs obtained in the conventional way.
@@
 
-\emph{``Fancy deconvolution''}-- The blas\'e approach can be viewed as a fancy deconvolution procedure, in which the convolution kernels get inferred alongside the imperfections of the model template.
+\emph{``Fancy deconvolution''}-- The  \emph{blas\'e} approach can be viewed as a fancy deconvolution procedure, in which the convolution kernels get inferred alongside the imperfections of the model template.
 This perspective may suit practitioners who wish to obtain a robust rotational broadening estimate in the presence of imperfect templates, and vice versa: robust templates in the presence of imperfect rotational broadening.
@@
 \emph{``A surrogate model, emulator, or discrepancy modeling''}-- The machine learning community is increasingly incorporating physics-based emulation into machine learning models.
-Blas\'e can be viewed as such a technique, often labeled as ``surrogate models''.
+\emph{Blas\'e} can be viewed as such a technique, often labeled as ``surrogate models''.
 
 \emph{``A good enough substrate''}-- The imperfections in pre-computed stellar and substellar models have hampered their use in high-fidelity applications, where model flaws overwhelm the sigal-to-noise in the data.
-The adaptability of blas\'e can be viewed as increasing the number of applications that could conceivably benefit from the rich-albeit-imperfect information encoded in these model grid spectra.
+The adaptability of  \emph{blas\'e} can be viewed as increasing the number of applications that could conceivably benefit from the rich-albeit-imperfect information encoded in these model grid spectra.
 In the next section we illustrate how these adaptable models can serve as a subtrate for applications in which the imperfections in precomputed models would have otherwise precluded their use.
@@
 
-The scaffolding of \texttt{blas\'e} is designed in a way to promote extensibility, so many scientific questions can be written down in the flexible language presented in this paper.
+The scaffolding of \texttt{blase} is designed in a way to promote extensibility, so many scientific questions can be written down in the flexible language presented in this paper.
 In some cases the flexibility, precision, speed, and ease-of-use may unlock new approaches to long-standing astrophysical questions and practical challenges.
@@
 
-Already \texttt{blas\'e} is equipped to fit every single line with its own systemic $RV$, by tuning the line center position $\lambda_\mathrm{c}$ at each training epoch.
+Already \texttt{blase} is equipped to fit every single line with its own systemic $RV$, by tuning the line center position $\lambda_\mathrm{c}$ at each training epoch.
 There exists both empirical evidence and some theoretical motivation that $RV$ jitter varies from line-to-line \citep{2018A&A...620A..47D}.
@@
 The measurement of fundamental stellar properties remains one of the principal applications of stellar spectroscopy.
-Blas\'e as it currently stands does not output fundamental properties, and in fact it takes them \emph{for granted}: you must specify \emph{e.g.} the $T_\mathrm{eff}$, $\log{g}$, [Fe/H], and [$\alpha$/Fe] of your PHOENIX template as Step Zero before that template gets cloned and warped to match data.
-So obtaining a $T_\mathrm{eff}$ estimate from blas\'e is impossible, at least as disscussed so far.
+\emph{Blas\'e} as it currently stands does not output fundamental properties, and in fact it takes them \emph{for granted}: you must specify \emph{e.g.} the $T_\mathrm{eff}$, $\log{g}$, [Fe/H], and [$\alpha$/Fe] of your PHOENIX template as Step Zero before that template gets cloned and warped to match data.
+So obtaining a $T_\mathrm{eff}$ estimate from  \emph{blas\'e} is impossible, at least as disscussed so far.
 
@@
 
-A purely model-based approach would start by blas\'e-cloning every single PHOENIX spectrum in the grid dimensions-of-interest, over the wavelength range-of-interest.
+A purely model-based approach would start by  \emph{blas\'e}-cloning every single PHOENIX spectrum in the grid dimensions-of-interest, over the wavelength range-of-interest.
 The resulting product would be a line-by-line catalog of cloned spectral properties for each 3D coordinate on this grid.
@@
 There are many conceivable ways to quantify these heatmap flaws.
-Blas\'e provides an expedient route.
-One can pull the heatmap towards the locus of points established by running blas\'e on benchmark stars.
+\emph{Blas\'e} provides an expedient route.
+One can pull the heatmap towards the locus of points established by running  \emph{blas\'e} on benchmark stars.
 The known $T_\mathrm{eff}$ of such systems would anchor the trend.
@@
 There exists a precision/accuracy tradeoff in the measurement process, usually stemming from the placement of the continuum, or assumptions about line blending.
-Blas\'e offers an immediate solution to these challenges, since it starts from our best-guess for how nearby lines may be shaping the continuum.
-Blas\'e appears to be a gateway to extremely fast and intervention-free industrial-scale abundances, potentially useful for large surveys like APOGEE \citep{2017AJ....154...94M}, Gaia-ESO \citep{2012Msngr.147...25G}, RAVE \citep{2006AJ....132.1645S}, and more.
+\emph{Blas\'e} offers an immediate solution to these challenges, since it starts from our best-guess for how nearby lines may be shaping the continuum.
+\emph{Blas\'e} appears to be a gateway to extremely fast and intervention-free industrial-scale abundances, potentially useful for large surveys like APOGEE \citep{2017AJ....154...94M}, Gaia-ESO \citep{2012Msngr.147...25G}, RAVE \citep{2006AJ....132.1645S}, and more.
 
 \subsection{Identifying missing lines}\label{secMissingLines}
-The automatic abundance determination spirit of blas\'e resembles \texttt{ROBOSPECT} \citep{2013PASP..125.1164W}, and other approaches aimed at automatically measuring equivalent widths for a large number of stars.
-One missing piece here is either to post-process the blas\'e-output EW catalog with chemical labels (C, N, O, Si, Ba, Eu, etc.), or incorporate that information directly into the structure of blas\'e.
+The automatic abundance determination spirit of  \emph{blas\'e} resembles \texttt{ROBOSPECT} \citep{2013PASP..125.1164W}, and other approaches aimed at automatically measuring equivalent widths for a large number of stars.
+One missing piece here is either to post-process the  \emph{blas\'e}-output EW catalog with chemical labels (C, N, O, Si, Ba, Eu, etc.), or incorporate that information directly into the structure of  \emph{blas\'e}.
 The latter case is intriguing.
 For example, PHOENIX may not yet be aware of the existence of some obscure lanthanum (La) line, possibly because the line was not yet understood and incorporated at the time of PHOENIX computation.
-Rather than wait for PHOENIX to receive the necessary update, blas\'e could instantiate one or more such lines in the locations guided by the atomic line list.
+Rather than wait for PHOENIX to receive the necessary update,  \emph{blas\'e} could instantiate one or more such lines in the locations guided by the atomic line list.
 This line could get initialized with some small amplitude, and would get added onto the underlying PHOENIX source as yet another freely tunable line.
@@
 
-\texttt{blas\'e} offers a new approach to Doppler imaging that may overcome these historical limitations by simultaneously fitting both the imperfections in the underlying spectrum and its line profile perturbations.
+\emph{Blas\'e} offers a new approach to Doppler imaging that may overcome these historical limitations by simultaneously fitting both the imperfections in the underlying spectrum and its line profile perturbations.
 This approach is analogous to the linearized model in \citet{2021arXiv211006271L}, but with the benefit of also handling non-linear properties of the spectrum such as line widths, shapes, and locations, while also handling telluric contamination.
@@
 
-Rather than applying a mixture model of fixed PHOENIX templates, one could adopt a mixture model of pre-cloned blas\'e models.
+Rather than applying a mixture model of fixed PHOENIX templates, one could adopt a mixture model of pre-cloned  \emph{blas\'e} models.
 Then, the imperfections in the starspot spectrum can be learned alongside the filling factor of the starspots.
@@
 
-Relatedly, magnetic fields information could be incorporated into blas\'e as line-by-line Land\'e $g-$factors, to allow magnetic sensitive lines to exhibit either Zeeman broadening or Zeeman splitting.
+Relatedly, magnetic fields information could be incorporated into  \emph{blas\'e} as line-by-line Land\'e $g-$factors, to allow magnetic sensitive lines to exhibit either Zeeman broadening or Zeeman splitting.
 This labelling could be conducted in the same way as described in Section \ref{secMissingLines}.
@@
 The extent of veiling would be fit alongside all the other stellar spectral lines.
-However, an MCMC approach may outperform blas\'e in accuracy, since the choice of picking an underlying spectral template is partially degenerate with the derived veiling.
+However, an MCMC approach may outperform  \emph{blas\'e} in accuracy, since the choice of picking an underlying spectral template is partially degenerate with the derived veiling.
 
@@
 The shapes of those lines can be affected by winds and other physical phenomena, yielding a variety of physically interpretable line shapes \citep{2022arXiv220802940E}.
-Blas\'e could incorporate those lineshapes as templates shared among several lines originating from similar physical enviroments, but scaled and shifted based on the details of each individual line's radiative transfer properties.
+\emph{Blas\'e} could incorporate those lineshapes as templates shared among several lines originating from similar physical enviroments, but scaled and shifted based on the details of each individual line's radiative transfer properties.
 
@@
 We demo the framework on two sources: the K5 exoplanet host star WASP 69 using a precomputed PHOENIX model, and the T6 ultracool dwarf 2MASS J08173001$-$6155158 using a precomputed Sonora-Bobcat model.
-We discuss how blas\'e can be used to measure Equivalent Widths for thousands of lines automatically, understand line lists, measure rotation rates, generate surrogate models, and construct semi-empirical models.
+We discuss how  \emph{blas\'e} can be used to measure Equivalent Widths for thousands of lines automatically, understand line lists, measure rotation rates, generate surrogate models, and construct semi-empirical models.
 This tool could readily have applications across stellar and substellar astronomy, including for PRV work, stellar compositions, Doppler imaging, and stellar activity.
@@
 
-Here we illustrate how \texttt{blas\'e} gets altered when applying the logarithmic flux pre-processing step.
+Here we illustrate how  \emph{blas\'e} gets altered when applying the logarithmic flux pre-processing step.
 First, we compute the natural log of the flux directly on the precomputed synthetic spectrum in its absolute flux scaling and native pixel sampling:
@@
 We simply ``rebrand'' $\mathsf{P}$ as residing in logarithmic flux units, and disregard it since it is largely a nuisance parameter anyways.
-We then treat the \texttt{blas\'e} clone model as a sum of opacities, retaining the Voigt profile:
+We then treat the  \emph{blas\'e} clone model as a sum of opacities, retaining the Voigt profile:
 
@@
 \section{Comparison to existing spectroscopy frameworks}
-Several astronomical spectral frameworks share similar aims as \texttt{blas\'e}.
+Several astronomical spectral frameworks share similar aims as  \emph{blas\'e}.
 These existing frameworks will have enduring value for the wide range of problems in the field of stellar spectroscopy.
@@
 The \texttt{wobble} framework also pioneered the off-label application of automatic differentiation frameworks---in this case \texttt{TensorFlow}---towards their physically-motivated use in stellar spectra.
-\texttt{blas\'e} can be viewed as an evaluable and interpretable super-resolution version of \texttt{wobble}, that accepts more bias in the bias-variance tradeoff.
+\texttt{blase} can be viewed as an evaluable and interpretable super-resolution version of \texttt{wobble}, that accepts more bias in the bias-variance tradeoff.
 
@@
 For very large bandwidths and very many spectral lines, the problem of identifying and cataloging line imperfections essentially becomes a book-keeping and continuum assignment problem.
-\texttt{blas\'e} and \texttt{starfish} provide different strategies for orchestrating the line-mismatch identification procedure, with each route having tradeoffs depending on the application.
+\texttt{blase} and \texttt{starfish} provide different strategies for orchestrating the line-mismatch identification procedure, with each route having tradeoffs depending on the application.
 
@@
 The \emph{celerit\'e} algorithm does not currently have a PyTorch implementation.
-The \texttt{GPyTorch} framework \citep{2018arXiv180911165G} has a large category of approximate and exact GPs that could be straight-forwardly dropped into \texttt{blas\'e}.
+The \texttt{GPyTorch} framework \citep{2018arXiv180911165G} has a large category of approximate and exact GPs that could be straight-forwardly dropped into \texttt{blase}.
 Even still these GPs could increase the computation cost by of order $10\times$.
@@
 The ExoJAX and Radis \citep{2019JQSRT.222...12P,2021JQSRT.26107476V} libraries offer a breakthrough solution to the voluminous line list problem.
-Even still, \emph{blase} deals with the less pure but more practical ``pseudo line'' that gets closer to the astronomical observables anyways, and offers a middle ground between the extremes of interpretability and performance.
+Even still,  \emph{blas\'e} deals with the less pure but more practical ``pseudo line'' that gets closer to the astronomical observables anyways, and offers a middle ground between the extremes of interpretability and performance.
 
@@
 The limb darkening may instead depend on physical properties of the spectral line formation, such as physical depth of formation, and so the extent of limb-darkening may jump haphazardly from line-to-line-to-line, rather than as a predictably smooth function across wavelength.
-\texttt{blas\'e} could be built to handle such a seemingly pathological scenario by adding a vector of limb darkening parameters, one for each line.
+\texttt{Blas\'e} could be built to handle such a seemingly pathological scenario by adding a vector of limb darkening parameters, one for each line.
 One would have to regularize the fits with some typical limb darkening and a heuristic penalty for departures from this mean.

### 2022-09-20 paper/paper1/ms.tex
More referee report progress...

@@
 A key new enabling technology breaks these classical tradeoffs in data-model comparisons for astronomical spectroscopy.
-Automatic differentiation \citep[``autodiff'' or ``autograd'',][]{2015arXiv150205767G, 2016PhDT.......317M} and its affiliated backpropagation algorithm \citep{kelley1960,Linnainmaa1976TaylorEO, 1986Natur.323..533R, 1990JGCD...13..926D} has revolutionized machine learning and neural network architecture design, and is increasingly applied in astrophysical data analysis contexts, \emph{e.g.} kernel phase coronography with \texttt{poppy} \citep{2021ApJ...907...40P}, and exoplanet orbit fitting with \texttt{exoplanet} \citep{2021JOSS....6.3285F}.
+Automatic differentiation \citep[``autodiff'' or ``autograd'',][]{2015arXiv150205767G, 2016PhDT.......317M} and its affiliated backpropagation algorithm \citep{kelley1960,Linnainmaa1976TaylorEO, 1986Natur.323..533R, 1990JGCD...13..926D} has revolutionized machine learning and neural network architecture design, and is increasingly applied in astrophysical data analysis contexts, \emph{e.g.} kernel phase coronography with \texttt{morphine} \citep{2021ApJ...907...40P,2021JOSAB..38.2465W} \edit1{and $\partial$\texttt{lux} \citep{10.1117/12.2629774}}, and exoplanet orbit fitting with \texttt{exoplanet} \citep{2021JOSS....6.3285F}.
 Of the spectroscopy frameworks mentioned above, the \texttt{TensorFlow}-based \citep{tensorflow2015-whitepaper} \texttt{wobble} and the \texttt{JAX}-based \citep{jax2018github} \texttt{ExoJAX} employ autodiff technology.
@@
 \noindent where the division indicates element-wise division of these arrays or ``vectors'' of flux values.
+\edit1{For elementwise-multiplication, we will adopt the $\odot$ operator symbol to distinguish that the multiplication occurs elementwise between two ``vectors'' of spectral fluxes, as opposed to a dot product (scalar ouput) or matrix multiplication (matrix output).
+This style of vector multiplication and division is standard in numerical numerical array manipulations such as those in \texttt{numpy}, \texttt{IDL}, \texttt{matlab}, \emph{etc.}}
 
@@
 
-We emphasize that a recreation of the unvarnished input spectrum---if desired---can be obtained by multiplying the continuum-flattened signal by the ``perturbed black body'', $\mathsf{B}(\bm{\lambda}_S)\odot \mathsf{P}(\bm{\lambda}_S)$, that symbolizes the black body modulated by continuum opacity or broad-band radiative transfer effects.
+A recreation of the unvarnished input spectrum---if desired---can be obtained by multiplying the continuum-flattened signal by the ``perturbed black body'', $\mathsf{B}(\bm{\lambda}_S)\odot \mathsf{P}(\bm{\lambda}_S)$, that symbolizes the black body modulated by continuum opacity or broad-band radiative transfer effects.
 This smooth spectrum may be useful for applications that need to keep track of broad-band flux, such as low-resolution spectra, or regions with molecular band heads.
@@
 Notice that the order of Steps 3 and 4 in Figure \ref{blase_flowchart} cannot be swapped.
-Mathematically speaking, multiplication and convolution do not commute.
+Mathematically speaking, elementwise-multiplication and convolution do not commute.
 While the distinction may seem negligible, it matters at the level of precisions sought in EPRV applications (Suvrath Mahadevan, Arpita Roy, Sharon Xuesong Wang \emph{priv.

### 2022-09-23 paper/paper1/ms.tex
What would a Machine Learning practitioner call this?

@@
 \begin{document}
-\shorttitle{Transfer learning with blas\'e}
-\shortauthors{Gully \& Morley}
+\shorttitle{\emph{blas\'e}: Interpretable Machine Learning for Spectroscopy}
+\shortauthors{Gully-Santiago \& Morley}
 \title{An Interpretable Machine Learning Framework for Modeling High-Resolution Spectroscopic Data\footnote{Open source code at \url{https://github.com/gully/blase}}}
@@
 
+\edit1{Our first stage is to \emph{clone} a precomputed synthetic spectrum.
+Cloning means to mimic, emulate, or otherwise approximate the appearance of a discretely sampled spectrum with some function or combination of functions.
+The subject of the cloning shall be a noise-free spectrum generated from physics-based atmosphere models.
+The purpose of cloning may appear superficial at first glance, but it serves as a necessary gateway for our ultimate data-model comparison goals.}
+
 \subsection{Overall Architecture and Design Choices}
@@
 \edit1{For elementwise-multiplication, we will adopt the $\odot$ operator symbol to distinguish that the multiplication occurs elementwise between two ``vectors'' of spectral fluxes, as opposed to a dot product (scalar ouput) or matrix multiplication (matrix output).
-This style of vector multiplication and division is standard in numerical numerical array manipulations such as those in \texttt{numpy}, \texttt{IDL}, \texttt{matlab}, \emph{etc.}}
+This style of vector multiplication and division is standard in numerical array manipulations such as those in \texttt{numpy}, \texttt{IDL}, \texttt{matlab}, \emph{etc.}}
 
@@
 
-\emph{``A surrogate model, emulator, or discrepancy modeling''}-- The machine learning community is increasingly incorporating physics-based emulation into machine learning models.
-\emph{Blas\'e} can be viewed as such a technique, often labeled as ``surrogate models''.
+\emph{``Surrogate modeling / Simulation Intelligence''}-- \edit1{Our slightly off-label use of machine learning straddles a boundary of machine learning and conventional physics-based models.
+The rapidly evolving field of machine learning may therefore classify the approach here using a few related terms.
+The phrases surrogate modeling, emulation, discrepancy modeling, transfer learning, simulation intelligence, and physics-informed machine learning could describe some aspects of \emph{blas\'e}, depending on the ultimate application by the end-user.
+For example, machine learning has been applied to physical chemistry models using a neural network as a cheaper-but-still-accurate ``surrogate'' for the real theory-based simulations.
+The realized speedups in quantum chemistry can approach 300,000$\times$ \citep{pmlr-v70-gilmer17a}.
+The key idea is that information from expensive physics-based simulations has been distilled by machine learning and can be applied to new examples.}
+
+\emph{``Narrowing the Synthetic Gap'}-- \edit1{Callendar et al. describe the notion of a Synthetic Gap: the fact that real astronomical spectra to not resemble their machine learning clones...}
 

### 2022-09-23 paper/paper1/ms.tex
neat discussion

@@
 
+\edit1{\emph{``Fancy interpolator''}--  A linear interpolator has to store the slope, offsets, and start and end points of all the line segments from pixel-to-pixel, merely to recreate a spectrum losslessly at some arbitrary coordinate or coordinates.
+\emph{Blas\'e} lossily recreates a spectrum, while having to store less information.
+In this way, it is fair to think of \emph{Blas\'e} as an interpolator that happens to have an intuitive interpretation.}
+
 \emph{``Surrogate modeling / Simulation Intelligence''}-- \edit1{Our slightly off-label use of machine learning straddles a boundary of machine learning and conventional physics-based models.
@@
 \subsection{Identifying missing lines}\label{secMissingLines}
-The automatic abundance determination spirit of  \emph{blas\'e} resembles \texttt{ROBOSPECT} \citep{2013PASP..125.1164W}, and other approaches aimed at automatically measuring equivalent widths for a large number of stars.
+\edit1{So far we have not addressed the inevitable prospect that some spectral line is present in the data, but was not predicted at the time of the atmosphere simulation.
+This ``missing lines'' scenario arises from our incomplete and evolving knowledge of quantum mechanics and molecular chemistry.
+\texttt{ROBOSPECT} \citep{2013PASP..125.1164W} and the \emph{Cannon} \citep{2017ApJ...836....5H} have data-driven ways to identify these lines.
+The \texttt{Cannon} can go further to indicate how these previously unknown lines correlate with the stellar labels, providing some physical undersanding for their cause.
+\emph{Blas\'e} as it currently stands has no mechanism for identifying these lines.
+Here we consider two scenarios: \emph{I)} information about the lines now exists, but they were not included in the spectrum forward modeling process, or \emph{II)} the lines are truly anomalous and unknown.
+For the first scenario, a Voigt profile or profiles could be instantiated on top of the line center location(s) of the previously unknown line, and it would be fitted alongside all its other fellow lines:}
+
+\begin{eqnarray}
+    \mathsf{S}_{\rm clone} = \Big( \prod_{m=1}^{N_{\mathrm{missing}}} 1-a_m \mathsf{V}_m \Big) \Big(\prod_{j=1}^{N_{\mathrm{lines}}} 1-a_j \mathsf{V}_j \Big) \label{missing}
+\end{eqnarray}
+
+for a large number of stars.
 One missing piece here is either to post-process the  \emph{blas\'e}-output EW catalog with chemical labels (C, N, O, Si, Ba, Eu, etc.), or incorporate that information directly into the structure of  \emph{blas\'e}.
 The latter case is intriguing.
-For example, PHOENIX may not yet be aware of the existence of some obscure lanthanum (La) line, possibly because the line was not yet understood and incorporated at the time of PHOENIX computation.
 Rather than wait for PHOENIX to receive the necessary update,  \emph{blas\'e} could instantiate one or more such lines in the locations guided by the atomic line list.

### 2022-09-23 paper/paper1/ms.tex
Nice discussion on missing lines!

@@
 \subsection{Identifying missing lines}\label{secMissingLines}
-\edit1{So far we have not addressed the inevitable prospect that some spectral line is present in the data, but was not predicted at the time of the atmosphere simulation.
-This ``missing lines'' scenario arises from our incomplete and evolving knowledge of quantum mechanics and molecular chemistry.
-\texttt{ROBOSPECT} \citep{2013PASP..125.1164W} and the \emph{Cannon} \citep{2017ApJ...836....5H} have data-driven ways to identify these lines.
+\edit1{So far we have not addressed the inevitable prospect that some spectral line or lines reside in the observed data spectrum, but are absent entirely from the precomputed model.
+This ``missing line'' scenario arises from our incomplete and evolving knowledge of quantum mechanics and molecular chemistry.
+\texttt{ROBOSPECT} \citep{2013PASP..125.1164W} and the \texttt{Cannon} \citep{2017ApJ...836....5H} have data-driven ways to identify these lines.
 The \texttt{Cannon} can go further to indicate how these previously unknown lines correlate with the stellar labels, providing some physical undersanding for their cause.
 \emph{Blas\'e} as it currently stands has no mechanism for identifying these lines.
-Here we consider two scenarios: \emph{I)} information about the lines now exists, but they were not included in the spectrum forward modeling process, or \emph{II)} the lines are truly anomalous and unknown.
-For the first scenario, a Voigt profile or profiles could be instantiated on top of the line center location(s) of the previously unknown line, and it would be fitted alongside all its other fellow lines:}
+Here we consider that there are $N_\mathrm{missing}$ lines with two scenarios: \emph{I)} the identification, line center location, and possibly other ancillary information about the lines has become newly available,  or \emph{II)} the lines are truly anomalous and unknown.}
+
+\edit1{For the first scenario, Voigt profiles could be instantiated at the newly-established line center locations $\lambda_{c,m}$ of the previously unknown line.
+So Equation \ref{equation1} would become a product of the new and old physics:}
 
@@
 
-for a large number of stars.
-One missing piece here is either to post-process the  \emph{blas\'e}-output EW catalog with chemical labels (C, N, O, Si, Ba, Eu, etc.), or incorporate that information directly into the structure of  \emph{blas\'e}.
-The latter case is intriguing.
-Rather than wait for PHOENIX to receive the necessary update,  \emph{blas\'e} could instantiate one or more such lines in the locations guided by the atomic line list.
-This line could get initialized with some small amplitude, and would get added onto the underlying PHOENIX source as yet another freely tunable line.
-In this way, all the other nearby lines would be used for assigning continuum and blending, but this free-standing lanthanum line would acquire a value properly taking into account the impact of its neighbors.
+\noindent \edit1{Scenario \emph{II}---truly unknown lines---would not have the luxury of knowing where to initialize the line center location, at least not based on precise theory.
+So instead, an educated guess for $\lambda_\mathrm{c,m}$ could be made based on the data centroid.
+}
+
+\edit1{In either case, the initialization of $A_m, \sigma_m, \gamma_m$ would be a matter of taste, and they would therefore be difficult to regularize, at least in the manner that we have described so far.
+For these reasons, an even better strategy may be to treat the data residuals not as lines, but in some other basis.
+For example it could be possible to model the \emph{blas\'e} residual spectrum with, say, \texttt{wobble}, or with a Gaussian process.
+These questions remain an open research area.}
 

### 2022-09-26 paper/paper1/ms.tex
Address the referee comments on PyTorch ecosystem

@@
 A key new enabling technology breaks these classical tradeoffs in data-model comparisons for astronomical spectroscopy.
-Automatic differentiation \citep[``autodiff'' or ``autograd'',][]{2015arXiv150205767G, 2016PhDT.......317M} and its affiliated backpropagation algorithm \citep{kelley1960,Linnainmaa1976TaylorEO, 1986Natur.323..533R, 1990JGCD...13..926D} has revolutionized machine learning and neural network architecture design, and is increasingly applied in astrophysical data analysis contexts, \emph{e.g.} kernel phase coronography with \texttt{morphine} \citep{2021ApJ...907...40P,2021JOSAB..38.2465W} \edit1{and $\partial$\texttt{lux} \citep{10.1117/12.2629774}}, and exoplanet orbit fitting with \texttt{exoplanet} \citep{2021JOSS....6.3285F}.
+Automatic differentiation \citep[``autodiff'' or ``autograd'',][]{2015arXiv150205767G, 2016PhDT.......317M} and its affiliated backpropagation algorithm \citep{kelley1960,Linnainmaa1976TaylorEO, 1986Natur.323..533R, 1990JGCD...13..926D} has revolutionized machine learning and neural network architecture design, and is increasingly applied in astrophysical data analysis contexts, \emph{e.g.} kernel phase coronography with \texttt{morphine} \citep{2021ApJ...907...40P,2021JOSAB..38.2465W} \edit1{and $\partial$\texttt{lux} \citep{10.1117/12.2629774}}, \edit1{stellar surface modeling with \texttt{starry} \citep{2021AJ....162..123L}}, and exoplanet orbit fitting with \texttt{exoplanet} \citep{2021JOSS....6.3285F}.
 Of the spectroscopy frameworks mentioned above, the \texttt{TensorFlow}-based \citep{tensorflow2015-whitepaper} \texttt{wobble} and the \texttt{JAX}-based \citep{jax2018github} \texttt{ExoJAX} employ autodiff technology.
@@
 This high-pass filtering step should be set to capture the genuine spectral shape, without over-fitting broad line wings such as those in deep hydrogen and sodium lines.
-Any high-pass filter will work, a Gaussian Process approach would be ideal.
+Any high-pass filter will work, a Gaussian Process approach would be ideal \citep{czekala15}.
 Instead we apply a simple and familiar heuristic: fit a polynomial $\mathsf{P}(\bm{\lambda}_S)$ to a few continuum peaks and divide out the trend.
@@
 The realized speedups in quantum chemistry can approach 300,000$\times$ \citep{pmlr-v70-gilmer17a}.
-The key idea is that information from expensive physics-based simulations has been distilled by machine learning and can be applied to new examples.}
+The key idea is that information from expensive physics-based simulations has been distilled by machine learning and can be applied to new examples.
+Scientific domain knowledge has been shown to aid in automatic exoplanet transit classification with Deep Learning \citep{2018ApJ...869L...7A}.}
 
@@
 In the next section we illustrate how these adaptable models can serve as a subtrate for applications in which the imperfections in precomputed models would have otherwise precluded their use.
+
+\subsection{PyTorch and the autodiff ecosystem}
+
+\edit1{One key limitation stems from the need for autodiff to ``know about'' or sense each step of the end-to-end computation.
+Practically speaking, this requirement means \texttt{blase} had to be written in a single differentiable (\emph{i.e.} autodiff-aware) programming framework.
+This requirement is so strong as to forbid the intermingling of frameworks and languages.
+The choice of which machine learning framework to adopt must therefore be made up-front, with some risk of \emph{vendor lock-in} once that choice has been set.
+We considered three such differentiable frameworks: \texttt{TensorFlow}, \texttt{JAX}, and \texttt{PyTorch}.
+We chose \texttt{PyTorch} because of its ease of use and its ability to structure differentiable models in an object-oriented style familiar within the \texttt{Python} community, with robust optimization support.
+PyTorch had already been applied to several astronomy projects by this time, including \texttt{Exonet} \citep{2018ApJ...869L...7A}, \texttt{PySR} \citep{pysr}, and \texttt{MPoL} \citep{mpol}.
+Ultimately, the choice of \texttt{PyTorch} made it easy to employ Sparse Matrices.}
+
+\edit1{This vendor lock-in gives rise to a frustrating incompatibility among exoplanet-relevant autodiff applications.
+For example, no differentiable piece of code in \texttt{blase} can call an \texttt{exoplanet} orbital radial velocity, a \texttt{starry} surface map, an \texttt{astropy} spectral coordinate transformation, a \texttt{wobble} spectral residual model, or an \texttt{ExoJAX} Voigt profile.
+All of those tools use either \texttt{Theano}, \texttt{TensorFlow}, \texttt{Numpy}, or \texttt{JAX}, which are incompatible with \texttt{PyTorch}.
+We use \texttt{SciPy}, but only for preprocessing, before the differentiable model has entered.
+Some heuristics exist for interchanging code among the frameworks, but they pose some implementation complexity and/or some performance hit.}
+
+\edit1{It may appear attractive to coordinate the relatively small autodiff-for-astrophysics community to adopt a single autodiff-able machine learning framework to facilitate compatibility.
+In practice such coordination may be ill-fated, since the rapid pace of machine learning development has not yet settled on a single best way forward.
+Each framework offers some strengths and demerits, and exploring those tradeoffs will continue to yield new experimental approaches.
+For example, \texttt{JAX} may offer speedups attributable to Just In Time (JIT) compilation, at the expense of a more rigid functional programming style.
+Access to scalable Gaussian Process implementations may be another deciding factor \citep{2022arXiv220908940A}.}
 

### 2022-09-26 paper/paper1/ms.tex
Almost finished with the report, two more tasks

@@
 The method can handle spectra without discernable continua, with some fine-adjustments to the preprocessing steps.
-We truncate the red and blue limits of the precomputed synthetic spectrum to match a high-bandwidth echelle spectrograph, extended with a buffer at the edges of size $\pm \Delta \lambda_{\mathrm{buffer}}$, chosen to account for plausible radial velocity and rotational broadening of real stars.
+We truncate the red and blue limits of the precomputed synthetic spectrum to match a high-bandwidth echelle spectrograph, extended with a buffer at the edges of size $\pm \Delta \lambda_{\mathrm{buffer}}$.
+\edit1{The buffers are needed for two reasons: 1) to include the influence of spectra lines whose centers lie just outside the region-of-interest, but whose wings are broad enough to affect the neighboring pixels; and 2) to account for astrophysical radial velocity shifts that may send line centers into- and out-of- the region of interest.}  We set the value of the buffers to account for plausible radial velocity and rotational broadening of real stars.
 A generous buffer of $v \sin{i} < 500 \;\kmps$ and $|RV|<500 \;\kmps$ yields a typical buffer of about 30~\AA.
@@
 Next, we identify the spectral lines.
-We apply a local-minimum-finding algorithm \citep{2020SciPy-NMeth} on the spectrum by defining a prominence threshold $P_{\rm rom} \in (0.005, 0.02)$.
-This threshold dictates the number of lines that will be modeled: a lower prominence finds more, weaker lines, and a larger prominence finds fewer, deeper lines.
+We apply the \edit1{\texttt{find\_peaks} local-maximum-finding algorithm implemented in \texttt{SciPy} \citep{2020SciPy-NMeth} on the negative of the spectrum, in order to find the local minima}.
+We select a ``topographic prominence threshold'' $P_{\rm rom} \in (0.005, 0.02)$, \edit1{defined as the vertical distance between the peak and the local baseline.
+The baseline is defined with topographic heuristics that act as effective pseudo-continuum-detectors.}  This threshold dictates the number of lines that will be modeled: a lower prominence finds more, weaker lines, and a larger prominence finds fewer, deeper lines.
 The prominence algorithm successfully finds lines that reside on top of broad line wings, or unresolved band heads provided that the individual lines exceed the prominence threshold in their local region.
@@
 The interpretations of these EWs may have to be treated with care if intercompared with EWs obtained in the conventional way.
-This approach may suit stellar abundance applications, such as Galactic archaeology and near-field cosmology.
+This approach may suit stellar abundance applications, such as Galactic archaeology and \edit1{chemically peculiar stars \citep[\emph{e.g.
+HgMn stars,}][]{2020MNRAS.496..832C}}.
 

### 2022-09-26 paper/paper1/ms.tex
The biggest thorn in the cloning--- line core residuals

@@
 But even more, $N_{\rm cut}$ has to be future-proofed for Doppler-shifting.
-Extreme Doppler shifts could hypothetically send line cores entirely outside the extents of $N_{\rm cut}$ if set too low.
+Extreme Doppler shifts could hypothetically send line \edit1{centers} entirely outside the extents of $N_{\rm cut}$ if set too low.
 We therefore typically set wingcuts comparable to the buffer size $2 \Delta \lambda_{\mathrm{buffer}}$, even though most weak lines only perceptibly affect $<1\;$\AA.
@@
 
-Second, another anticipated flaw occurs in the line cores of relatively narrow lines, where the pseudo-Voigt profile becomes a poor approximation of the exact Voigt profile.
+Second, a conspicuous flaw occurs in the line cores of relatively narrow lines.
 The cloned model tends to overestimate the flux at the core and underestimate the flux along the slopes of the lines.
-
-Finally, and most perniciously, a large category of residuals appear near the wings of the deepest and broadest lines---such as hydrogen and neutral alkali metal lines.
-The true lines exhibit advanced lineshapes, such as non-Lorentzian line wings that are not captured with the overly simplistic Voigt line profile.
+\edit1{The residuals therefore exhibit a \texttt{W}-shape ringing artifact.
+The scientific impact of this flaw will depend on the precision demands of the application, but it may not be as bad as it seems for many applications.
+Recall that the cloning occurs at native spectral resolution, much greater than typical instrumental resolution.
+The \texttt{W}-shape ringing residuals will tend to cancel each other out once the clone model undergoes convolution with an instrumental profile (Section \ref{jointModel}).
+We ascribe this flaw to the adoption of approximate pseudo-Voigt profiles rather than exact Voigt profiles.
+High-precision applications may wish to employ the exact Voigt profile, which we estimate to cost about 27$\times$ more than this approximate form.}
+
+\edit1{The overprediction of line cores in narrow lines is only partly attributable to the pseudo-Voigt approximation.
+Even exact Voigt profiles inadequately represent true atmospheric opacities.
+Typical stellar and substellar atmosphere models compute opacities for a stratified atmosphere consisting of tens or hundreds of locally isothermal, isobaric atmospheric layers.
+Each of those layers may exhibit a local per-line opacity that matches a Voigt profile.
+But the cumulative product of many slightly-different Voigt profiles is not exactly equal to any single Voigt profile.
+So the \emph{blas\'e} approach of asserting lines as Voigt profiles can be best understood as a phenomenological model-- that lines tend to look Voigt-like, even if they are non-Voigt in detail.
+In principle, \texttt{blas\'e} could be configured in a way to learn the requisite perturbations to the Voigt profile in a systematic manner, pooling information among lines.
+Such an approach is not yet implemented, and may be an enticing avenue of future research.}
+
+Finally, and most perniciously, a large category of residuals appear near the wings of the deepest and broadest lines---such as hydrogen lines and neutral alkali metal lines.
+The true lines exhibit advanced lineshapes, such as non-Lorentzian line wings that are not captured with the overly simplistic Voigt line profile, even an exact one.
 Figure \ref{fig_zoom_cloning_performance} highlights super-Lorentzian line wings around a line at 8691~\AA.
-Narrow lines devolve into missing line wing opacity, the favored tradeoff when the continuum estimate's poor performance outweighs the pain of a narrow-but-tolerably-small spike.
+\edit1{Nearby} narrow lines devolve into missing line wing opacity, the favored tradeoff when the continuum estimate's poor performance outweighs the pain of a narrow-but-tolerably-small spike.
 This flaw can be seen where a line initialized at 8692.5~\AA~ and another pair of lines at 8690.0 all melt into line wings.
@@
 
-\subsection{Joint Stellar and Telluric Model}
+\subsection{Joint Stellar and Telluric Model}\label{jointModel}
 Figure \ref{blase_flowchart} shows a visual guide to all the steps in  \emph{blas\'e}.

### 2022-09-26 paper/paper1/ms.tex
OK, I think I might be done with the referee comments.  Yay!

@@
 One current assumption of starspot spectral decomposition is that the starspot spectrum itself resembles the stellar photosphere of a cooler star.
-This assumption appears adequate for detecting starspot spectra and measuring their physical properties \citep{gullysantiago17}.
-But to second order starspots should exhibit some spectral peculiarities that make them depart from a ``normal'' stellar photosphere.
-We may be probing deeper into the photosphere, and so the lines may experience higher pressure, with slightly different line widths.
+This assumption appears adequate for detecting starspot spectra and measuring their physical properties \citep{2017ApJ...836..200G}.
+But to second order, starspots should exhibit some spectral peculiarities that make them depart from a ``normal'' stellar photosphere.
+We may be probing deeper into the photosphere, and so the lines may experience higher pressure, with slightly different line widths \citep{oneal96}.
 Or maybe the finite convective velocity shift can be directly seen as systematic shifts of the spectral lines (as stated in Section \ref{eprv}).
@@
 
-Often the most spotted stars are also rapidly rotating, and so the large $v\sin{i}$ blurs the starspot spectrum into the spot-free component of the photosphere.
-One experiment could measure the starspot spectral peculiarities more easily: obtainhigh resolution spectra of a pole-on (or nearly pole-on) young star where the lines will be sharp and the two components of the spectra should be more easily discernable.
-
-Relatedly, magnetic fields information could be incorporated into  \emph{blas\'e} as line-by-line Land\'e $g-$factors, to allow magnetic sensitive lines to exhibit either Zeeman broadening or Zeeman splitting.
-This labelling could be conducted in the same way as described in Section \ref{secMissingLines}.
+Relatedly, magnetic field information could be incorporated into \emph{blas\'e} as line-by-line Land\'e $g-$factor ``labels'', when available.
+The use of the labels could take a few different forms.
+\edit1{The labels could simply serve as flags, to indicate that regularization should be weakened to allow these magnetic sensitive lines to wander farther from their naive non-magnetic expectations.
+Or the labels could be used to infer the extent of Zeeman broadening or Zeeman splitting by parameterizing these operations as convolutions.
+The best choice for how to incorporate magnetic sensitivity into the flexibility of \emph{blas\'e} remains an open research question.}
 

### 2022-10-03 paper/paper1/ms.tex
Add some text from comments by Phill Cargile and Hajime Kawahara, thank you!!

@@
 
-For example, exoplanet cross-correlation spectroscopy \citep{2013MNRAS.436L..35B} hinges on accurate molecular spectral templates to detect and characterize the atmospheres of exoplanets.
+For example, exoplanet cross-correlation spectroscopy \citep{2010Natur.465.1049S,2012Natur.486..502B,2013MNRAS.436L..35B} hinges on accurate molecular spectral templates to detect and characterize the atmospheres of exoplanets.
 Imperfections in these templates can mute the perceived signal strength of these atmospheric features \citep{2015A&A...575A..20H}.
@@
 \texttt{wobble} treats each pixel as a tunable control point, producing $\sim10^5$ parameters for a modern stellar spectrum.
-The \texttt{ExoJAX} framework has only $\sim$dozens of tunable parameters that describe the fundamental physical properties controlling a brown dwarf atmosphere.
+The \texttt{ExoJAX} framework has only $\sim$100 tunable parameters that describe the fundamental physical properties controlling a \edit2{stratified} brown dwarf atmosphere.
 These two autodiff-aware frameworks span the extreme ends of non-parametric and parametric modeling for spectroscopy.
@@
 We place mild demands on the precomputed spectra.
-They should have sporadic regions of discernable continuum devoid of lines, and the continuum shape should vary smoothly in wavelength.
+They should have sporadic regions of discernable continuum \edit2{(or pseudocontinuum)} devoid of lines, and the continuum shape should vary smoothly in wavelength.
 The spectral lines or pseudo-lines should be resolved (and not sub-sampled).
-We suspect most stellar spectra meet these criteria, except for the coolest M dwarfs and brown dwarfs.
-The method can handle spectra without discernable continua, with some fine-adjustments to the preprocessing steps.
+Many precomputed synthetic stellar spectra will meet these criteria.
+\edit2{For those spectra that do not meet these criteria, the method should still work with some additional fine-tuning of the pre-processing steps that follow.
+M dwarfs in the red-optical, solar-like stars in the blue visible, and brown dwarfs at virtually any wavelength may fall into this category.}
+
 We truncate the red and blue limits of the precomputed synthetic spectrum to match a high-bandwidth echelle spectrograph, extended with a buffer at the edges of size $\pm \Delta \lambda_{\mathrm{buffer}}$.
-\edit1{The buffers are needed for two reasons: 1) to include the influence of spectra lines whose centers lie just outside the region-of-interest, but whose wings are broad enough to affect the neighboring pixels; and 2) to account for astrophysical radial velocity shifts that may send line centers into- and out-of- the region of interest.}  We set the value of the buffers to account for plausible radial velocity and rotational broadening of real stars.
+\edit1{The buffers are needed for two reasons: 1) to include the influence of spectral lines whose centers lie just outside the region-of-interest, but whose wings are broad enough to affect the neighboring pixels; and 2) to account for astrophysical radial velocity shifts that may send line centers into- and out-of- the region of interest.}  We set the value of the buffers to account for plausible radial velocity and rotational broadening of real stars.
 A generous buffer of $v \sin{i} < 500 \;\kmps$ and $|RV|<500 \;\kmps$ yields a typical buffer of about 30~\AA.
@@
 This model has between $N_{\mathrm{lines}}\times 1$ and $N_{\mathrm{lines}}\times 4$ free parameters, depending on how many of the 4 physical line properties you wish to tune.
-We default to fitting three parameters per line, where the center wavelength is held fixed and the amplitude, width, and lineshape are allowed to vary.
+\edit2{The treshold-based line-finding algorithm performs very well at finding the line center positions, so in practice we can fix the line center wavelengths.}  We therefore default to fitting three parameters per line, where the center wavelength is held fixed and the amplitude, width, and lineshape are allowed to vary.
 We minimize a scalar ``goodness-of-fit'' metric, \emph{aka} loss scalar $\mathcal{L}$, chosen as the mean squared error (MSE), which is proportional to $\chi^2$, the sum of the squares of the residual vector $\mathsf{R} \equiv \mathsf{S}-\mathsf{S}_{\rm clone}$ but has no notion of per-pixel noise since the precomputed synthetic spectrum has no uncertainty:
@@
 There exists a special exponential spectral sample spacing that allows the convolution operators to work out of the box for rotational broadening.
-The design of \emph{blas\'e} permits the stellar and telluric models to be re\"evaluated at any wavelength coordinate vector, provided that it adequately samples the underlying lines.
+The design of \emph{blas\'e} permits the stellar and telluric models to be reevaluated at any wavelength coordinate vector, provided that it adequately samples the underlying lines.
 We, therefore, change the sampling from the native stellar and telluric wavelength coordinate grids, $\bm{\lambda}_S$ and $\bm{\lambda}_T$, to this special exponentially sampled wavelength grid, denoted with the subscript $Z$:
@@
 The resolving power may also be treated as tunable if it is not known or varies slightly with \emph{e.g.} seeing, slit-or-fiber illumination, or instrumental configuration: $30,003$.
+\edit2{The pseudocontinuum polynomial term $\mathsf{P}$ can also be tuned to refine the continuum placement initialized at the pre-processing stage, yielding an additional 5-15 parameters, depending on the choice of polynomial order.}
 
@@
 \emph{Blas\'e} could incorporate those lineshapes as templates shared among several lines originating from similar physical enviroments, but scaled and shifted based on the details of each individual line's radiative transfer properties.
+
+\subsection{Sky emission lines and wavelength calibration}
+\edit2{Night sky emission lines---arising from OH, for example---add emission line spikes on top of the astronomical target spectrum of interest.
+These lines get subtracted in most pipeline packages, but in some cases a more careful treatment may be desirable.
+In HPF, a sky reference fiber points towards blank sky, needing some careful calibration before subtraction from the target fiber \citep{2022JOSS....7.4302G}.
+Remote sensing applications may wish to measure such emission as the primary science of interest.
+It would be straightforward for \emph{blas\'e} to handle emission lines, following the augmentation for emission lines in Section \ref{gpuConsiderations}.
+The main benefit over conventional sky subtraction methods may be the ability to disentangle the underlying spectrum at super-resolution.
+A predictive model for sky emission could be built-up over time, building data-driven corrections to theoretical models.}
+
+\edit2{The well-known center wavelengths of these lines could be used for additional wavelength calibration, as is already done in IGRINS \citep{jaejoonlee16}.
+An imperfect instrumental wavelength calibration could be diagnosed based on departure of the predicted versus realized wavelength center positions, enabling a relatively fast and intervention-free mitigation strategy.
+Hypothetically, a prescription for the wavelength calibration could be built directly into \emph{blas\'e}, with the coefficents for a Polynomial re-routing the realized wavelength center locations from their known wavelength centers.
+}
 

### 2022-10-03 paper/paper1/ms.tex
Grammarly edits and some extra references, etc.

@@
 In this paper, we show that autodiff-aware semi-empirical models offer an appealing middle ground: informed by self-consistent models but refined with data.
-This sweet spot in the bias-variance tradoff can be thought of as a hybrid data-and-model driven approach.
+This sweet spot in the bias-variance tradeoff can be thought of as a hybrid data-and-model driven approach.
 The algorithm presented here focuses on modeling the spectra of stars and brown dwarfs.
-Existing stellar models \citep[\emph{e.g.}][]{husser13} and substellar models \citep[\emph{e.g.}][]{2021ApJ...920...85M} laboriously solve for a self consistent thermal structure in the atmosphere given the copious opacity sources that themselves depend on temperture and pressure.
+Existing stellar models \citep[\emph{e.g.}][]{husser13} and substellar models \citep[\emph{e.g.}][]{2021ApJ...920...85M} laboriously solve for a self-consistent thermal structure in the atmosphere given the copious opacity sources that themselves depend on temperature and pressure.
 Here we build upon that hard work by cloning pre-existing synthetic stellar or substellar models (Section \ref{methodology}), and optionally by cloning models of Earth's atmospheric ``telluric'' absorption (Section \ref{sectionTelluric}).
@@
 We truncate the red and blue limits of the precomputed synthetic spectrum to match a high-bandwidth echelle spectrograph, extended with a buffer at the edges of size $\pm \Delta \lambda_{\mathrm{buffer}}$.
-\edit1{The buffers are needed for two reasons: 1) to include the influence of spectral lines whose centers lie just outside the region-of-interest, but whose wings are broad enough to affect the neighboring pixels; and 2) to account for astrophysical radial velocity shifts that may send line centers into- and out-of- the region of interest.}  We set the value of the buffers to account for plausible radial velocity and rotational broadening of real stars.
+\edit1{The buffers are needed for two reasons: 1) to include the influence of spectral lines whose centers lie just outside the region-of-interest, but whose wings are broad enough to affect the neighboring pixels; and 2) to account for astrophysical radial velocity shifts that may send line centers into- and out of- the region of interest.}  We set the value of the buffers to account for plausible radial velocity and rotational broadening of real stars.
 A generous buffer of $v \sin{i} < 500 \;\kmps$ and $|RV|<500 \;\kmps$ yields a typical buffer of about 30~\AA.
@@
 Any high-pass filter will work, a Gaussian Process approach would be ideal \citep{czekala15}.
-Instead we apply a simple and familiar heuristic: fit a polynomial $\mathsf{P}(\bm{\lambda}_S)$ to a few continuum peaks and divide out the trend.
+Instead, we apply a simple and familiar heuristic: fit a polynomial $\mathsf{P}(\bm{\lambda}_S)$ to a few continuum peaks and divide out the trend.
 
@@
 \noindent where the division indicates element-wise division of these arrays or ``vectors'' of flux values.
-\edit1{For elementwise-multiplication, we will adopt the $\odot$ operator symbol to distinguish that the multiplication occurs elementwise between two ``vectors'' of spectral fluxes, as opposed to a dot product (scalar ouput) or matrix multiplication (matrix output).
+\edit1{For elementwise-multiplication, we will adopt the $\odot$ operator symbol to distinguish that the multiplication occurs elementwise between two ``vectors'' of spectral fluxes, as opposed to a dot product (scalar output) or matrix multiplication (matrix output).
 This style of vector multiplication and division is standard in numerical array manipulations such as those in \texttt{numpy}, \texttt{IDL}, \texttt{matlab}, \emph{etc.}}
@@
 The ``perturbed black body'' continuum model contains $n_{\mathrm{poly}}+1$ fixed-but-possibly tunable lookup parameters, plus the fixed input $T_{\mathrm{eff}}$.
-For most practitioners these terms serve as nuisance parameters and are perfunctorily discarded.
+For most practitioners, these terms serve as nuisance parameters and are perfunctorily discarded.
 
@@
 The Voigt profile $\mathsf{V}(\lambda, \sigma_j, \gamma_j)$ can be computed in exact closed-form using the Voigt-Hjerting function \citep{1938ApJ....88..508H} as the real part of the complex Fadeeva function \citep[\emph{e.g.}][]{2011arXiv1106.0151Z}.
-Evaluation of the Fadeeva function can be computationally costly, and so approximate forms may be desirable.
+Evaluation of the Fadeeva function can be computationally costly, so approximate forms may be desirable.
 Here we adopt the pseudo-Voigt approximation \citep{Ida:nt0146}.
@@
 This model has between $N_{\mathrm{lines}}\times 1$ and $N_{\mathrm{lines}}\times 4$ free parameters, depending on how many of the 4 physical line properties you wish to tune.
-\edit2{The treshold-based line-finding algorithm performs very well at finding the line center positions, so in practice we can fix the line center wavelengths.}  We therefore default to fitting three parameters per line, where the center wavelength is held fixed and the amplitude, width, and lineshape are allowed to vary.
+\edit2{The threshold-based line-finding algorithm performs very well at finding the line center positions, so in practice, we can fix the line center wavelengths.}  We, therefore, default to fitting three parameters per line, where the center wavelength is held fixed and the amplitude, width, and lineshape are allowed to vary.
 We minimize a scalar ``goodness-of-fit'' metric, \emph{aka} loss scalar $\mathcal{L}$, chosen as the mean squared error (MSE), which is proportional to $\chi^2$, the sum of the squares of the residual vector $\mathsf{R} \equiv \mathsf{S}-\mathsf{S}_{\rm clone}$ but has no notion of per-pixel noise since the precomputed synthetic spectrum has no uncertainty:
@@
 \edit1{Geometrically, the $\grad \mathcal{L}$ Jacobian resembles a compass needle guiding the direction to the best fit.
-This compass needle resides in, say, 21,000 dimensional space, instead of our mere 3 dimensions of physical space.
+This compass needle resides in, say, 21,000-dimensional space, instead of our mere 3 dimensions of physical space.
 Conventional gradient-free optimizers evaluate the Jacobian through approximate finite differences, which result in numerical instabilities.
-Those instabilities accrue and break down at around the 300 dimension mark.
+Those instabilities accrue and break down at around the 300-dimension mark.
 So non-gradient-based optimization can be pictured as having only a coarse needle---assembled from approximate finite differences---that begins to spin aimlessly after too many dimensions.}
@@
 These frameworks give exact Jacobians, instantaneously, for free (or cheap).
-Without exact Jacobians a conventional optimization step only obtains \emph{one} piece of information: how much the overall loss changed.
+Without exact Jacobians, a conventional optimization step only obtains \emph{one} piece of information: how much the overall loss changed.
 With exact Jacobians, we obtain $N_\mathrm{params}+1$ pieces of information for each evaluation of the forward model.
@@
 Forward modeling with tens of thousands of physics-informed parameters may seem like such a significant paradigm shift that it can feel too good to be true.
-In this section we introduce the non-negligible architectural design tradeoffs that arise when adopting an autodiff framework---such as \texttt{PyTorch}---for physics based forward modeling.
+In this section, we introduce the non-negligible architectural design tradeoffs that arise when adopting an autodiff framework---such as \texttt{PyTorch}---for physics-based forward modeling.
 
 First, we make a few tweaks to the implementation for numerical purposes.
-We enforce that all Gaussian and Lorenzian widths are positive by tuning the natural log of the widths, then exponentiating them before inclusion in Equation \ref{equation1}.
+We enforce that all Gaussian and Lorentzian widths are positive by tuning the natural log of the widths, then exponentiating them before inclusion in Equation \ref{equation1}.
 Based on the minus sign in Equation \ref{equation1}, an amplitude $a_j$ could hypothetically take on either positive values (flux loss, absorption) or negative values (flux gain, emission).
-For now we focus on photospheric absorption lines---as opposed to, say, chromospheric emission lines---and therefore enforce all the amplitudes to be positive by tuning $\ln{a_j}$ and then exponentiating in the same strategy as above.
+For now, we focus on photospheric absorption lines---as opposed to, say, chromospheric emission lines---and therefore enforce all the amplitudes to be positive by tuning $\ln{a_j}$ and then exponentiating in the same strategy as above.
 Emission lines---if desired---could be included with a mere sign flip to Equation \ref{equation1}.
 Lines that can manifest either in absorption or emission could hypothetically relax the natural log pre-processing step for isolated lines.
-We narate only the absorption scenario moving forward.
+We narrate only the absorption scenario moving forward.
 
@@
 Efficient algorithms for assembling and coalescing sparse matrices exist in \texttt{PyTorch}.
-Some GPUs now support additional hardware-acceleration of sparse matrices, providing even greater speedups.
+Some GPUs now support additional hardware acceleration of sparse matrices, providing even greater speedups.
 
@@
 The pictured 50 \AA~ chunk contains 121 spectral lines and represents about 1$\%$ of the entire spectral bandwidth that was cloned.
-Some flaws can be seen near the cores of deep lines, or wings of broad lines.}
+Some flaws can be seen near the cores of deep lines or wings of broad lines.}
     \label{fig_cloned_spectrum_demo}
@@
 \textbf{Before:} The top panel shows a forward model evaluated with coarsely initialized line-by-line properties, based on simple threshold-based peak-finding and crude estimation of the shape of the line.
-\textbf{After:} The bottom panel shows the \texttt{blase} fine-tuned clone, with nearly indentical spectral structure.}
+\textbf{After:} The bottom panel shows the \texttt{blase} fine-tuned clone, with nearly identical spectral structure.}
     \label{fig_telluric_clone}
@@
 The telluric clone shows a comparable level of performance.
-These residuals tend to pile up in local symmetrically balanced clusters that get cancelled out once convolved with coarser instrumental line profiles, so their overall effect at instrumental resolution is typically negligible: the clones are almost perfect.
+These residuals tend to pile up in local symmetrically balanced clusters that get canceled out once convolved with coarser instrumental line profiles, so their overall effect at instrumental resolution is typically negligible: the clones are almost perfect.
 We identify three main categories of cloning flaws that may not be negligible depending on the science application.
@@
 The cloned model tends to overestimate the flux at the core and underestimate the flux along the slopes of the lines.
-\edit1{The residuals therefore exhibit a \texttt{W}-shape ringing artifact.
+\edit1{The residuals, therefore, exhibit a \texttt{W}-shape ringing artifact.
 The scientific impact of this flaw will depend on the precision demands of the application, but it may not be as bad as it seems for many applications.
@@
 In principle, \texttt{blas\'e} could be configured in a way to learn the requisite perturbations to the Voigt profile in a systematic manner, pooling information among lines.
-Such an approach is not yet implemented, and may be an enticing avenue of future research.}
+Such an approach is not yet implemented and may be an enticing avenue for future research.}
 
@@
 The cloned model has 7 spectral lines describing the 400 pixels in this 4 \AA~ chunk.
-The native PHOENIX pixel sampling can be see as the boxy steps in both the native and cloned model.
+The native PHOENIX pixel sampling can be seen as the boxy steps in both the native and cloned model.
 }
@@
 
-The cloned models already represents a useful intermediate product: the distillation of $N_s=335,849\times2$ pixel flux values and their wavelength coordinates into a more compact quartet of properties for a list of $<10,000$ spectral lines, a dimensionality reduction of $18\times$ for the cost of 1.2\%/pixel in accuracy.
+The cloned model already represents a useful intermediate product: the distillation of $N_s=335,849\times2$ pixel flux values and their wavelength coordinates into a more compact quartet of properties for a list of $<10,000$ spectral lines, a dimensionality reduction of $18\times$ for the cost of 1.2\%/pixel in accuracy.
 So the process so far can be myopically viewed as a physics-informed compression algorithm.
@@
 Rotational broadening $v\sin{i}$ and radial velocity $RV$ depend on the observer's viewing location.
-We follow \citet{czekala15} by emphasizing the qualifier \emph{extrinsic}, to distiguish between stellar \emph{intrinsic} properties, such as $T_{\mathrm{eff}}, \log{g},\mathrm{and\,} [\mathrm{Fe}/\mathrm{H}]$.
+We follow \citet{czekala15} by emphasizing the qualifier \emph{extrinsic}, to distinguish between stellar \emph{intrinsic} properties, such as $T_{\mathrm{eff}}, \log{g},\mathrm{and\,} [\mathrm{Fe}/\mathrm{H}]$.
 Intrinsic properties appear the same from any viewing location---at least for stars with isotropic surfaces---while extrinsic properties do not.
 The distinction is important because the extrinsic terms act as simple convolutions and translations to the cloned spectrum, and can be treated after the cloning procedure.
-We therefore define an augmented model, which we designate the ``extrinsic'' model, $\mathsf{S}_{\rm ext}$:
+We, therefore, define an augmented model, which we designate the ``extrinsic'' model, $\mathsf{S}_{\rm ext}$:
 
@@
 There is currently no equivalent post-processing of the telluric spectrum $\mathsf{T}$.
-As mentioned previously, we assume motions of the Earth's atmosphere are much less than the desired stellar radial velocity precisions.
+As mentioned previously, we assume the motions of Earth's atmosphere are much less than the desired stellar radial velocity precisions.
 However, demanding Extreme Precision Radial Velocity (EPRV) exoplanet searches may need to consider minuscule systematic $RV$ shifts and broadening of the telluric templates, arising from the turbulent and bulk motions of the Earth's atmosphere.
@@
 Notice that the order of Steps 3 and 4 in Figure \ref{blase_flowchart} cannot be swapped.
-Mathematically speaking, elementwise-multiplication and convolution do not commute.
+Mathematically speaking, elementwise multiplication and convolution do not commute.
 While the distinction may seem negligible, it matters at the level of precisions sought in EPRV applications (Suvrath Mahadevan, Arpita Roy, Sharon Xuesong Wang \emph{priv.
@@
 As noted, typical data-pixel sampling $\bm{\lambda}_D$ is much coarser than the model pixel sampling $\bm{\lambda}_Z$.
-We therefore resample the model to the data spectrum in the following way.
-We evaluate the joint model $\mathsf{M}_{\rm joint}$ at all of the super-resolution wavelength coordinates, and then compute the mean value of those pixels within the bounds of each coarse data pixel.
-The resampling procedure is autodiff-aware: the same clusters of high resolution coordinates map to the same data pixel coordinates, no matter what the $RV$ is.
+We, therefore, resample the model to the data spectrum in the following way.
+We evaluate the joint model $\mathsf{M}_{\rm joint}$ at all of the super-resolution wavelength coordinates and then compute the mean value of those pixels within the bounds of each coarse data pixel.
+The resampling procedure is autodiff-aware: the same clusters of high-resolution coordinates map to the same data pixel coordinates, no matter what the $RV$ is.
 The $RV$ only dictates what flux values are realized within those pixel bounds.
@@
 Such a stratagem would overfit the data, resulting in unphysically perverse lineshapes that do not reflect the air of reality we aspire to impose on our synthetic spectral models.
-Lines would haphazardly fit noise-spikes, and conspire together to warp spectral shapes in unexpected ways.
+Lines would haphazardly fit noise spikes, and conspire together to warp spectral shapes in unexpected ways.
 This overfit model may suit some rare purposes.
@@
 We default to the following choice.
-We fix all line parameters except for amplitude, which receives an L1 loss-- namely we penalize the absolute value of departures from an amplitude's starting place:
+We fix all line parameters except for amplitude, which receives an L1 loss-- namely, we penalize the absolute value of departures from an amplitude's starting place:
 
@@
 Conspicuous residuals of $\pm10\%$ appear throughout the spectrum, with lines in the correct place, but with the amplitudes systematically biased.
-These lines residuals arise from bona-fide imperfections in the PHOENIX spectrum, with a minor contribution from the coarse sampling of the PHOENIX grid.
+These line residuals arise from bona-fide imperfections in the PHOENIX spectrum, with a minor contribution from the coarse sampling of the PHOENIX grid.
 
@@
 First, our finite regularization restricts line amplitudes and widths from straying too far from their values.
-This computational tug-of-war makes line cores land just-short of the values the otherwise would have obtained in the absence of regularization.
+This computational tug-of-war makes line cores land just short of the values they otherwise would have obtained in the absence of regularization.
 Second, the spectral resolution kernel may get biased from the domineering model-misspecification of broad lines, setting up a slightly sub-par performance for all the other lines.
@@
 The  \texttt{blase} end-to-end model exhibits excellent agreement with the data.
-The trustworthiness of the line properties inferred in locations where stellar and telluric lines exactly overlap remains an open research question: a tie-breaker is needed to distinguish these overlapping line inferrences.
-We discuss conceivable tie-brakers later in the paper.
+The trustworthiness of the line properties inferred in locations where stellar and telluric lines exactly overlap remains an open research question: a tie-breaker is needed to distinguish these overlapping line inferences.
+We discuss conceivable tie-breakers later in the paper.
 
@@
     \caption{End-to-end training on a portion of HPF spectrum of WASP-69.
-The top two panels show the cloning in from Step 1 of the visual guide.
+The top two panels show the cloning from Step 1 of the visual guide.
 The middle two panels depict the before-and-after of the instrument-convolved and resampled joint model illustrated in Step 4 of the visual guide.
@@
 The T6 ultracool dwarf 2MASS J08173001$-$6155158 was recently observed with IGRINS, revealing a rich spectroscopic atlas of molecules in its $T_\mathrm{eff} = 1060 \pm 50$ cloud-free atmosphere \citep{2022MNRAS.tmp.1421T}.
-The sea of molecular lines are so rich as to blur entirely the notion of isolated lines.
+The sea of molecular lines is so rich as to blur entirely the notion of isolated lines.
 Many lines should instead be considered pseudo lines.
 The boundaries of lines for 2MASS J08173001$-$6155158 become even more amorphous in the presence of its moderately high $v\sin{i} = 22.5\pm 0.5 ~ \text{km/s}$ rotational broadening \citep{2022MNRAS.tmp.1421T}.
-This T6 IGRINS spectrum therefore represents an extreme end of a test-case, in which the underlying notions of \emph{blas\'e} go outside the comparatively safe assumptions in stellar spectra.
+This T6 IGRINS spectrum, therefore, represents an extreme test-case, in which the underlying notions of \emph{blas\'e} go outside the comparatively safe assumptions in stellar spectra.
 The application of  \emph{blas\'e} to 2MASS J08173001$-$6155158 can be viewed as conducting a deconvolution step simultaneously paired with line-by-line inference.
@@
 
-Up to this point we have presented the mechanics of  \emph{blas\'e}, and have shown that it applies to a wide range of applications including stars, substars, and Earth's atmosphere.
+Up to this point, we have presented the mechanics of  \emph{blas\'e}, and have shown that it applies to a wide range of applications including stars, substars, and Earth's atmosphere.
 The interpretation of the approach and its line-by-line outputs may carry different meaning to different practitioners.
@@
 The line lists generally store the temperature and pressure scaling terms that get included in 1-D thermal structure calculations with distinct Voigt coefficient terms computed for each layer in the stellar or substellar or Earth atmosphere.
-The cloned properties therefore represent the flux-weighted average line property, having undergone an integral through the radiative transfer in the atmosphere.
+The cloned properties, therefore, represent the flux-weighted average line property, having undergone an integral through the radiative transfer in the atmosphere.
 
 Our deliberate choice not to initialize from the atomic and molecular lines means we do not have to compute an expensive multi-layer atmosphere at each inference step.
-The precomputed models have aleady taken care of that step, and so our cloning procedure encodes all the quantum mechanics and stellar atmospheres knowledge written down at the time of the computing of PHOENIX, Sonora, or LBLRTM.
+The precomputed models have already taken care of that step, and so our cloning procedure encodes all the quantum mechanics and stellar atmospheres knowledge written down at the time of the computing of PHOENIX, Sonora, or LBLRTM.
 
@@
 \emph{Blas\'e} does not have these limitations.
-The interpretations of these EWs may have to be treated with care if intercompared with EWs obtained in the conventional way.
+The interpretations of these EWs may have to be treated with care if intercompared with EWs obtained conventionally.
 This approach may suit stellar abundance applications, such as Galactic archaeology and \edit1{chemically peculiar stars \citep[\emph{e.g.
@@
 The phrases surrogate modeling, emulation, discrepancy modeling, transfer learning, simulation intelligence, and physics-informed machine learning could describe some aspects of \emph{blas\'e}, depending on the ultimate application by the end-user.
-For example, machine learning has been applied to physical chemistry models using a neural network as a cheaper-but-still-accurate ``surrogate'' for the real theory-based simulations.
+The surrogate modeling approach has been applied with mixed success for M-dwarfs \citep{2020A&A...642A..22P} and brown dwarfs \citep{2020PASP..132d4502J}.
 The realized speedups in quantum chemistry can approach 300,000$\times$ \citep{pmlr-v70-gilmer17a}.
-The key idea is that information from expensive physics-based simulations has been distilled by machine learning and can be applied to new examples.
-Scientific domain knowledge has been shown to aid in automatic exoplanet transit classification with Deep Learning \citep{2018ApJ...869L...7A}.}
-
-\emph{``Narrowing the Synthetic Gap'}-- \edit1{Callendar et al. describe the notion of a Synthetic Gap: the fact that real astronomical spectra to not resemble their machine learning clones...}
-
-\emph{``A good enough substrate''}-- The imperfections in pre-computed stellar and substellar models have hampered their use in high-fidelity applications, where model flaws overwhelm the sigal-to-noise in the data.
+The key idea is that information from expensive physics-based simulations can be distilled by machine learning and then applied cheaply to new examples.
+The performance of these surrogates should improve when more scientific domain knowledge is included at training time \citep{2018ApJ...869L...7A}.}
+
+\emph{``Narrowing the Synthetic Gap'}-- \edit2{Our approach to pretraining on synthetic spectra and comparing to data resembles the technique of \citet{2020A&A...642A..22P}.
+There, a convolutional neural network is pretrained on PHOENIX spectra, and then applied to CARMENES spectra of M dwarfs.
+\citet{2020A&A...642A..22P} designate the difference in appearance between the synthetic spectra and observed spectra as the ``synthetic gap'', showing that this gap leads to inaccurate metallicity inferences.
+\emph{Blas\'e} shows a path towards shrinking the synthetic gap, with the prospect of re-training convolutional neural networks with semi-empirical spectra produced via the method described here, assuming high-fidelity metallicity labels can be obtained through co-natal binary studies.}
+
+\emph{``A good enough substrate''}-- The imperfections in pre-computed stellar and substellar models have hampered their use in high-fidelity applications, where model flaws overwhelm the signal-to-noise in the data.
 The adaptability of  \emph{blas\'e} can be viewed as increasing the number of applications that could conceivably benefit from the rich-albeit-imperfect information encoded in these model grid spectra.
-In the next section we illustrate how these adaptable models can serve as a subtrate for applications in which the imperfections in precomputed models would have otherwise precluded their use.
+In the next Section, we illustrate how these adaptable models can serve as a substrate for applications in which the imperfections in precomputed models would have otherwise precluded their use.
 
@@
 All of those tools use either \texttt{Theano}, \texttt{TensorFlow}, \texttt{Numpy}, or \texttt{JAX}, which are incompatible with \texttt{PyTorch}.
-We use \texttt{SciPy}, but only for preprocessing, before the differentiable model has entered.
+We use \texttt{SciPy}, but only for preprocessing, before the differentiable model has been entered.
 Some heuristics exist for interchanging code among the frameworks, but they pose some implementation complexity and/or some performance hit.}
@@
 \edit1{It may appear attractive to coordinate the relatively small autodiff-for-astrophysics community to adopt a single autodiff-able machine learning framework to facilitate compatibility.
-In practice such coordination may be ill-fated, since the rapid pace of machine learning development has not yet settled on a single best way forward.
+In practice, such coordination may be ill-fated, since the rapid pace of machine learning development has not yet settled on a single best way forward.
 Each framework offers some strengths and demerits, and exploring those tradeoffs will continue to yield new experimental approaches.
@@
 The scaffolding of \texttt{blase} is designed in a way to promote extensibility, so many scientific questions can be written down in the flexible language presented in this paper.
-In some cases the flexibility, precision, speed, and ease-of-use may unlock new approaches to long-standing astrophysical questions and practical challenges.
+In some cases, the flexibility, precision, speed, and ease of use may unlock new approaches to long-standing astrophysical questions and practical challenges.
 Here we enumerate some planned or conceivable extensions.
@@
 \emph{Blas\'e} as it currently stands does not output fundamental properties, and in fact it takes them \emph{for granted}: you must specify \emph{e.g.} the $T_\mathrm{eff}$, $\log{g}$, [Fe/H], and [$\alpha$/Fe] of your PHOENIX template as Step Zero before that template gets cloned and warped to match data.
-So obtaining a $T_\mathrm{eff}$ estimate from  \emph{blas\'e} is impossible, at least as disscussed so far.
-
-The clearest way forward would be to calibrate the blas\'e-derived line properties.  Simply put, we seek a set of functions, $f$, that relate the line properties back to fundamental properties:
+So obtaining a $T_\mathrm{eff}$ estimate from  \emph{blas\'e} is impossible, at least as discussed so far.
+
+The clearest way forward would be to calibrate the blas\'e-derived line properties.  Simply put, we seek a set of functions, $f$, that relate the line properties to fundamental properties:
 $$f_j(a_j, \sigma_j, \gamma_j) \to (T_\mathrm{eff}, \log{g}, [\mathrm{Fe}/\mathrm{H}])$$
@@
 In this way, we are reverse-engineering the \emph{e.g.} temperature dependence of each spectral line, as encoded by the PHOENIX atmosphere models.
-The final step could involve finding the neareast neighbor of each $j^{th}$ observed spectral line to each $j^{th}$ grid heatmap point; that is, finding a way to invert the function $g_j$ to get $f_j$.
+The final step could involve finding the nearest neighbor of each $j^{th}$ observed spectral line to each $j^{th}$ grid heatmap point; that is, finding a way to invert the function $g_j$ to get $f_j$.
 The information-weighted mean, median, or mode of these nearest neighbors could then be reported as a revised ``best-fit $T_\mathrm{eff}$'', for instance.
@@
 The known $T_\mathrm{eff}$ of such systems would anchor the trend.
-We suspect the directionality of the purely model based heatmap must be correct, simply the slope, offset, and concavity of the trends may be wrong.
+We suspect the directionality of the purely model-based heatmap must be correct, simply the slope, offset, and concavity of the trends may be wrong.
 With new information from each benchmark, the heatmap would get lifted like a central tentpole propping up an under-supported tent.
@@
 The construction and calibration of these ensembles of spectra represents a tremendous amount of work beyond the scope of this paper.
-But its creation could yield an extremely precise, fast, interpretable and reasonably accurate way to measure the properties of stars based solely on their high bandwidth \'echelle spectra.
+But its creation could yield an extremely precise, fast, interpretable, and reasonably accurate way to measure the properties of stars based solely on their high bandwidth \'echelle spectra.
 The overall accuracy hinges on the accuracy of the stellar benchmark labels.
@@
 There exists a corollary from the method described here to the approach of \citet{czekala15}.
-There, a 3D heatmap was created for relating the eigenweights of a PCA basis back to the stellar fundamental properties, for spectrum emulation purposes.
-Here, the 3D heatmap is created on a line-by-line basis, and is therefore interpretable.
+There, a 3D heatmap was created for relating the eigenweights of a PCA basis to the stellar fundamental properties, for spectrum emulation purposes.
+Here, the 3D heatmap is created on a line-by-line basis and is therefore interpretable.
 The PCA eigenspectra were generally uninterpretable, at least not easily.
@@
 Stellar abundance work involves precisely measuring line strengths (often reported as EWs) of different chemical constituents evinced in a star or substar's photosphere.
-Relating those EWs back to physics can be done in a few ways.
+Relating those EWs to physics can be done in a few ways.
 Most easily, trends and patterns in the EWs can be assembled, and metal-rich and metal-poor clusters can be identified.
 There exists a precision/accuracy tradeoff in the measurement process, usually stemming from the placement of the continuum, or assumptions about line blending.
-\emph{Blas\'e} offers an immediate solution to these challenges, since it starts from our best-guess for how nearby lines may be shaping the continuum.
+\emph{Blas\'e} offers an immediate solution to these challenges since it starts from our best guess for how nearby lines may be shaping the continuum.
 \emph{Blas\'e} appears to be a gateway to extremely fast and intervention-free industrial-scale abundances, potentially useful for large surveys like APOGEE \citep{2017AJ....154...94M}, Gaia-ESO \citep{2012Msngr.147...25G}, RAVE \citep{2006AJ....132.1645S}, and more.
@@
 \texttt{ROBOSPECT} \citep{2013PASP..125.1164W} and the \texttt{Cannon} \citep{2017ApJ...836....5H} have data-driven ways to identify these lines.
-The \texttt{Cannon} can go further to indicate how these previously unknown lines correlate with the stellar labels, providing some physical undersanding for their cause.
+The \texttt{Cannon} can go further to indicate how these previously unknown lines correlate with the stellar labels, providing some physical understanding of their cause.
 \emph{Blas\'e} as it currently stands has no mechanism for identifying these lines.
-Here we consider that there are $N_\mathrm{missing}$ lines with two scenarios: \emph{I)} the identification, line center location, and possibly other ancillary information about the lines has become newly available,  or \emph{II)} the lines are truly anomalous and unknown.}
+Here we consider that there are $N_\mathrm{missing}$ lines with two scenarios: \emph{I)} the identification, line center location, and possibly other ancillary information about the lines have become newly available,  or \emph{II)} the lines are truly anomalous and unknown.}
 
@@
 For these reasons, an even better strategy may be to treat the data residuals not as lines, but in some other basis.
-For example it could be possible to model the \emph{blas\'e} residual spectrum with, say, \texttt{wobble}, or with a Gaussian process.
+For example, it could be possible to model the \emph{blas\'e} residual spectrum with, say, \texttt{wobble}, or with a Gaussian process.
 These questions remain an open research area.}
@@
 Longitudinally symmetric surface features---on the other hand---do not change as the star rotates on its axis.
-The existence of these features manifest as static-in-time \emph{kurtosis} of the spectral line.
+The existence of these features manifests as static-in-time \emph{kurtosis} of the spectral line.
 For example, a hypothetical non-emitting (black) polar starspot exhibits a deficit in flux at the line core, resulting in less zero-velocity flux than its homogeneous counterpart.
@@
 
-It is easier to reverse-engineer longitudinally asymmetric features than longitudinally symmetric ones, since we assume that we occasionally catch a glimpse of the spot-free limbs and their pristine line profiles.
+It is easier to \edit1{infer} longitudinally asymmetric features than longitudinally symmetric ones, since we assume that we occasionally catch a glimpse of the spot-free limbs and their pristine line profiles.
 The latter requires exact knowledge of the underlying spectral template.
 Isolated, deep, well-calibrated spectral lines constitute the only practical scenario where exact knowledge can plausibly be claimed.
-Isolated spectral lines may be scarce or absent for M-dwarfs and brown dwarfs where lines blend ostensibly in an inseparable way, counfounding Doppler imaging.
+Isolated spectral lines may be scarce or absent for M-dwarfs and brown dwarfs where lines blend ostensibly in an inseparable way, confounding Doppler imaging.
 
@@
 Relatedly, magnetic field information could be incorporated into \emph{blas\'e} as line-by-line Land\'e $g-$factor ``labels'', when available.
-The use of the labels could take a few different forms.
+The use of these labels could take a few different forms.
 \edit1{The labels could simply serve as flags, to indicate that regularization should be weakened to allow these magnetic sensitive lines to wander farther from their naive non-magnetic expectations.
@@
 
-Accretion veiling can be treated in an analogous way, but with some more complications.  Emission lines can be easily incorporated as described in Section \ref{gpuConsiderations}.
+Accretion veiling can be treated analogously, but with some more complications.  Emission lines can be easily incorporated as described in Section \ref{gpuConsiderations}.
 These emission lines could be initialized with a line list of hydrogen lines, forbidden lines, and other conspicuous features.
 The shapes of those lines can be affected by winds and other physical phenomena, yielding a variety of physically interpretable line shapes \citep{2022arXiv220802940E}.
-\emph{Blas\'e} could incorporate those lineshapes as templates shared among several lines originating from similar physical enviroments, but scaled and shifted based on the details of each individual line's radiative transfer properties.
+\emph{Blas\'e} could incorporate those lineshapes as templates shared among several lines originating from similar physical environments, but scaled and shifted based on the details of each line's radiative transfer properties.
 
@@
 \edit2{Night sky emission lines---arising from OH, for example---add emission line spikes on top of the astronomical target spectrum of interest.
-These lines get subtracted in most pipeline packages, but in some cases a more careful treatment may be desirable.
+These lines get subtracted in most pipeline packages, but in some cases, a more careful treatment may be desirable.
 In HPF, a sky reference fiber points towards blank sky, needing some careful calibration before subtraction from the target fiber \citep{2022JOSS....7.4302G}.
-Remote sensing applications may wish to measure such emission as the primary science of interest.
+Remote sensing applications may wish to measure such emission lines as the primary science of interest.
 It would be straightforward for \emph{blas\'e} to handle emission lines, following the augmentation for emission lines in Section \ref{gpuConsiderations}.
@@
 \edit2{The well-known center wavelengths of these lines could be used for additional wavelength calibration, as is already done in IGRINS \citep{jaejoonlee16}.
-An imperfect instrumental wavelength calibration could be diagnosed based on departure of the predicted versus realized wavelength center positions, enabling a relatively fast and intervention-free mitigation strategy.
+An imperfect instrumental wavelength calibration could be diagnosed based on the departure of the predicted versus realized wavelength center positions, enabling a relatively fast and intervention-free correction strategy.
 Hypothetically, a prescription for the wavelength calibration could be built directly into \emph{blas\'e}, with the coefficents for a Polynomial re-routing the realized wavelength center locations from their known wavelength centers.
@@
 \section{Conclusions}
-We have introduced an interpretable machine learning approach to forward modelling stellar, substellar, and telluric spectroscopic data.
-The line-by-line approach approach relies on a key enabling technology, automatic differentiation, that allows a nearly unlimited number of spectral lines to be forward modelled simultaneously.
+We have introduced an interpretable machine learning approach to forward modeling stellar, substellar, and telluric spectroscopic data.
+The line-by-line approach relies on a key enabling technology, automatic differentiation, that allows a nearly unlimited number of spectral lines to be forward modeled simultaneously.
 We initialize these lines to match precomputed synthetic stellar spectra, achieving excellent performance, and lending some confidence that the approach has a capacity to capture a tremendous amount of information at once.
@@
 
+    \edit1{We thank the anonymous referee for comments that improved the paper.}  \edit2{We thank Phill Cargile, Hajime Kawahara, and Dan Jaffe for pre-print comments that were incorporated into the draft text.}
     MGS thanks Ian Czekala, Phill Cargile, Rodrigo Luger, Greg Herczeg, Will Best, Dan Foreman-Mackey, Meg Bedell, Mark Marley, Erwan Pannier, Dirk C.M.
@@
     astropy \citep{exoplanet:astropy13,exoplanet:astropy18},
-    exoplanet \citep{exoplanet:joss}, %celerite?
+    exoplanet \citep{2021JOSS....6.3285F},
     numpy \citep{harris2020array},

### 2022-10-03 paper/paper1/ms.tex
remove all the edit flags, switch to twocolumn

@@
-\documentclass[trackchanges]{aastex631}
+\documentclass[twocolumn]{aastex631}
 \usepackage{hyperref}
 \bibliographystyle{aasjournal}
-%\turnoffedit
 \usepackage[caption=false]{subfig}
@@
 A key new enabling technology breaks these classical tradeoffs in data-model comparisons for astronomical spectroscopy.
-Automatic differentiation \citep[``autodiff'' or ``autograd'',][]{2015arXiv150205767G, 2016PhDT.......317M} and its affiliated backpropagation algorithm \citep{kelley1960,Linnainmaa1976TaylorEO, 1986Natur.323..533R, 1990JGCD...13..926D} has revolutionized machine learning and neural network architecture design, and is increasingly applied in astrophysical data analysis contexts, \emph{e.g.} kernel phase coronography with \texttt{morphine} \citep{2021ApJ...907...40P,2021JOSAB..38.2465W} \edit1{and $\partial$\texttt{lux} \citep{10.1117/12.2629774}}, \edit1{stellar surface modeling with \texttt{starry} \citep{2021AJ....162..123L}}, and exoplanet orbit fitting with \texttt{exoplanet} \citep{2021JOSS....6.3285F}.
+Automatic differentiation \citep[``autodiff'' or ``autograd'',][]{2015arXiv150205767G, 2016PhDT.......317M} and its affiliated backpropagation algorithm \citep{kelley1960,Linnainmaa1976TaylorEO, 1986Natur.323..533R, 1990JGCD...13..926D} has revolutionized machine learning and neural network architecture design, and is increasingly applied in astrophysical data analysis contexts, \emph{e.g.} kernel phase coronography with \texttt{morphine} \citep{2021ApJ...907...40P,2021JOSAB..38.2465W} and $\partial$\texttt{lux} \citep{10.1117/12.2629774}, stellar surface modeling with \texttt{starry} \citep{2021AJ....162..123L}, and exoplanet orbit fitting with \texttt{exoplanet} \citep{2021JOSS....6.3285F}.
 Of the spectroscopy frameworks mentioned above, the \texttt{TensorFlow}-based \citep{tensorflow2015-whitepaper} \texttt{wobble} and the \texttt{JAX}-based \citep{jax2018github} \texttt{ExoJAX} employ autodiff technology.
 \texttt{wobble} treats each pixel as a tunable control point, producing $\sim10^5$ parameters for a modern stellar spectrum.
-The \texttt{ExoJAX} framework has only $\sim$100 tunable parameters that describe the fundamental physical properties controlling a \edit2{stratified} brown dwarf atmosphere.
+The \texttt{ExoJAX} framework has only $\sim$100 tunable parameters that describe the fundamental physical properties controlling a stratified brown dwarf atmosphere.
 These two autodiff-aware frameworks span the extreme ends of non-parametric and parametric modeling for spectroscopy.
@@
 
-\edit1{Our first stage is to \emph{clone} a precomputed synthetic spectrum.
+Our first stage is to \emph{clone} a precomputed synthetic spectrum.
 Cloning means to mimic, emulate, or otherwise approximate the appearance of a discretely sampled spectrum with some function or combination of functions.
 The subject of the cloning shall be a noise-free spectrum generated from physics-based atmosphere models.
-The purpose of cloning may appear superficial at first glance, but it serves as a necessary gateway for our ultimate data-model comparison goals.}
+The purpose of cloning may appear superficial at first glance, but it serves as a necessary gateway for our ultimate data-model comparison goals.
 
@@
 We place mild demands on the precomputed spectra.
-They should have sporadic regions of discernable continuum \edit2{(or pseudocontinuum)} devoid of lines, and the continuum shape should vary smoothly in wavelength.
+They should have sporadic regions of discernable continuum (or pseudocontinuum) devoid of lines, and the continuum shape should vary smoothly in wavelength.
 The spectral lines or pseudo-lines should be resolved (and not sub-sampled).
 Many precomputed synthetic stellar spectra will meet these criteria.
-\edit2{For those spectra that do not meet these criteria, the method should still work with some additional fine-tuning of the pre-processing steps that follow.
-M dwarfs in the red-optical, solar-like stars in the blue visible, and brown dwarfs at virtually any wavelength may fall into this category.}
+For those spectra that do not meet these criteria, the method should still work with some additional fine-tuning of the pre-processing steps that follow.
+M dwarfs in the red-optical, solar-like stars in the blue visible, and brown dwarfs at virtually any wavelength may fall into this category.
 
 We truncate the red and blue limits of the precomputed synthetic spectrum to match a high-bandwidth echelle spectrograph, extended with a buffer at the edges of size $\pm \Delta \lambda_{\mathrm{buffer}}$.
-\edit1{The buffers are needed for two reasons: 1) to include the influence of spectral lines whose centers lie just outside the region-of-interest, but whose wings are broad enough to affect the neighboring pixels; and 2) to account for astrophysical radial velocity shifts that may send line centers into- and out of- the region of interest.}  We set the value of the buffers to account for plausible radial velocity and rotational broadening of real stars.
+The buffers are needed for two reasons: 1) to include the influence of spectral lines whose centers lie just outside the region-of-interest, but whose wings are broad enough to affect the neighboring pixels; and 2) to account for astrophysical radial velocity shifts that may send line centers into- and out of- the region of interest.
+We set the value of the buffers to account for plausible radial velocity and rotational broadening of real stars.
 A generous buffer of $v \sin{i} < 500 \;\kmps$ and $|RV|<500 \;\kmps$ yields a typical buffer of about 30~\AA.
@@
 \noindent where the division indicates element-wise division of these arrays or ``vectors'' of flux values.
-\edit1{For elementwise-multiplication, we will adopt the $\odot$ operator symbol to distinguish that the multiplication occurs elementwise between two ``vectors'' of spectral fluxes, as opposed to a dot product (scalar output) or matrix multiplication (matrix output).
-This style of vector multiplication and division is standard in numerical array manipulations such as those in \texttt{numpy}, \texttt{IDL}, \texttt{matlab}, \emph{etc.}}
+For elementwise-multiplication, we will adopt the $\odot$ operator symbol to distinguish that the multiplication occurs elementwise between two ``vectors'' of spectral fluxes, as opposed to a dot product (scalar output) or matrix multiplication (matrix output).
+This style of vector multiplication and division is standard in numerical array manipulations such as those in \texttt{numpy}, \texttt{IDL}, \texttt{matlab}, \emph{etc.}
 
@@
 Next, we identify the spectral lines.
-We apply the \edit1{\texttt{find\_peaks} local-maximum-finding algorithm implemented in \texttt{SciPy} \citep{2020SciPy-NMeth} on the negative of the spectrum, in order to find the local minima}.
-We select a ``topographic prominence threshold'' $P_{\rm rom} \in (0.005, 0.02)$, \edit1{defined as the vertical distance between the peak and the local baseline.
-The baseline is defined with topographic heuristics that act as effective pseudo-continuum-detectors.}  This threshold dictates the number of lines that will be modeled: a lower prominence finds more, weaker lines, and a larger prominence finds fewer, deeper lines.
+We apply the \texttt{find\_peaks} local-maximum-finding algorithm implemented in \texttt{SciPy} \citep{2020SciPy-NMeth} on the negative of the spectrum, in order to find the local minima.
+We select a ``topographic prominence threshold'' $P_{\rm rom} \in (0.005, 0.02)$, defined as the vertical distance between the peak and the local baseline.
+The baseline is defined with topographic heuristics that act as effective pseudo-continuum-detectors.
+This threshold dictates the number of lines that will be modeled: a lower prominence finds more, weaker lines, and a larger prominence finds fewer, deeper lines.
 The prominence algorithm successfully finds lines that reside on top of broad line wings, or unresolved band heads provided that the individual lines exceed the prominence threshold in their local region.
@@
 This model has between $N_{\mathrm{lines}}\times 1$ and $N_{\mathrm{lines}}\times 4$ free parameters, depending on how many of the 4 physical line properties you wish to tune.
-\edit2{The threshold-based line-finding algorithm performs very well at finding the line center positions, so in practice, we can fix the line center wavelengths.}  We, therefore, default to fitting three parameters per line, where the center wavelength is held fixed and the amplitude, width, and lineshape are allowed to vary.
+The threshold-based line-finding algorithm performs very well at finding the line center positions, so in practice, we can fix the line center wavelengths.
+We, therefore, default to fitting three parameters per line, where the center wavelength is held fixed and the amplitude, width, and lineshape are allowed to vary.
 We minimize a scalar ``goodness-of-fit'' metric, \emph{aka} loss scalar $\mathcal{L}$, chosen as the mean squared error (MSE), which is proportional to $\chi^2$, the sum of the squares of the residual vector $\mathsf{R} \equiv \mathsf{S}-\mathsf{S}_{\rm clone}$ but has no notion of per-pixel noise since the precomputed synthetic spectrum has no uncertainty:
@@
 As seen in Figure \ref{fig_Nlines_vs_teff}, the number of lines can exceed 7,000, meaning the clone model has over 21,000 free parameters.
-Fitting more than about 300 parameters is difficult with conventional optimizers\edit1{, and so historically the practice of finding the best settings for all 21,000 parameters would have been considered hopeless, a fool's errand.
+Fitting more than about 300 parameters is difficult with conventional optimizers, and so historically the practice of finding the best settings for all 21,000 parameters would have been considered hopeless, a fool's errand.
 A new category of ``gradient-based'' optimizers offers a breakthrough.
 Gradient-based optimizers have virtually no restriction on the number of parameters, scaling to functions of possibly millions or billions of tuning parameters.
-Gradient-based optimizers get vastly more information per likelihood evaluation: $N_\mathrm{params}+1$ pieces of information-- the scalar value of the loss evaluated for some setting, \emph{and} the derivative of the loss scalar with respect to each of the parameters, the so-called \emph{Jacobian}: $\grad \mathcal{L} = (\frac{\partial \mathcal{L}}{\partial a_j}, \frac{\partial \mathcal{L}}{\partial \sigma_j}, \cdots, \frac{\partial \mathcal{L}}{\partial \gamma_j})$.}  The Jacobian indicates how the MSE would decrease with a change in the parameter-of-interest, or put simply ``which way and by how much'' you have to change each individual line property to get a better fit.
-
-\edit1{Geometrically, the $\grad \mathcal{L}$ Jacobian resembles a compass needle guiding the direction to the best fit.
+Gradient-based optimizers get vastly more information per likelihood evaluation: $N_\mathrm{params}+1$ pieces of information-- the scalar value of the loss evaluated for some setting, \emph{and} the derivative of the loss scalar with respect to each of the parameters, the so-called \emph{Jacobian}: $\grad \mathcal{L} = (\frac{\partial \mathcal{L}}{\partial a_j}, \frac{\partial \mathcal{L}}{\partial \sigma_j}, \cdots, \frac{\partial \mathcal{L}}{\partial \gamma_j})$.
+The Jacobian indicates how the MSE would decrease with a change in the parameter-of-interest, or put simply ``which way and by how much'' you have to change each individual line property to get a better fit.
+
+Geometrically, the $\grad \mathcal{L}$ Jacobian resembles a compass needle guiding the direction to the best fit.
 This compass needle resides in, say, 21,000-dimensional space, instead of our mere 3 dimensions of physical space.
@@
 Those instabilities accrue and break down at around the 300-dimension mark.
-So non-gradient-based optimization can be pictured as having only a coarse needle---assembled from approximate finite differences---that begins to spin aimlessly after too many dimensions.}
-
-\edit1{We adopt ``Full-batch'' Gradient Descent optimization, as opposed to the more popular Stochastic Gradient Descent \citep[SGD,][]{2016arXiv160904747R}.
+So non-gradient-based optimization can be pictured as having only a coarse needle---assembled from approximate finite differences---that begins to spin aimlessly after too many dimensions.
+
+We adopt ``Full-batch'' Gradient Descent optimization, as opposed to the more popular Stochastic Gradient Descent \citep[SGD,][]{2016arXiv160904747R}.
 In Full-Batch Gradient Descent, the entire dataset is evaluated in Equation \ref{simpleLikelihood}, as opposed to only a portion of the dataset in the Stochastic counterpart.
-Appendix \ref{minibatches} discusses the tradeoffs and rationale for ``Full-Batch''.}
+Appendix \ref{minibatches} discusses the tradeoffs and rationale for ``Full-Batch''.
 
@@
 But even more, $N_{\rm cut}$ has to be future-proofed for Doppler-shifting.
-Extreme Doppler shifts could hypothetically send line \edit1{centers} entirely outside the extents of $N_{\rm cut}$ if set too low.
+Extreme Doppler shifts could hypothetically send line centers entirely outside the extents of $N_{\rm cut}$ if set too low.
 We therefore typically set wingcuts comparable to the buffer size $2 \Delta \lambda_{\mathrm{buffer}}$, even though most weak lines only perceptibly affect $<1\;$\AA.
@@
 The cloned model tends to overestimate the flux at the core and underestimate the flux along the slopes of the lines.
-\edit1{The residuals, therefore, exhibit a \texttt{W}-shape ringing artifact.
+The residuals, therefore, exhibit a \texttt{W}-shape ringing artifact.
 The scientific impact of this flaw will depend on the precision demands of the application, but it may not be as bad as it seems for many applications.
@@
 We ascribe this flaw to the adoption of approximate pseudo-Voigt profiles rather than exact Voigt profiles.
-High-precision applications may wish to employ the exact Voigt profile, which we estimate to cost about 27$\times$ more than this approximate form.}
-
-\edit1{The overprediction of line cores in narrow lines is only partly attributable to the pseudo-Voigt approximation.
+High-precision applications may wish to employ the exact Voigt profile, which we estimate to cost about 27$\times$ more than this approximate form.
+
+The overprediction of line cores in narrow lines is only partly attributable to the pseudo-Voigt approximation.
 Even exact Voigt profiles inadequately represent true atmospheric opacities.
@@
 In principle, \texttt{blas\'e} could be configured in a way to learn the requisite perturbations to the Voigt profile in a systematic manner, pooling information among lines.
-Such an approach is not yet implemented and may be an enticing avenue for future research.}
+Such an approach is not yet implemented and may be an enticing avenue for future research.
 
@@
 Figure \ref{fig_zoom_cloning_performance} highlights super-Lorentzian line wings around a line at 8691~\AA.
-\edit1{Nearby} narrow lines devolve into missing line wing opacity, the favored tradeoff when the continuum estimate's poor performance outweighs the pain of a narrow-but-tolerably-small spike.
+Nearby narrow lines devolve into missing line wing opacity, the favored tradeoff when the continuum estimate's poor performance outweighs the pain of a narrow-but-tolerably-small spike.
 This flaw can be seen where a line initialized at 8692.5~\AA~ and another pair of lines at 8690.0 all melt into line wings.
@@
 The resolving power may also be treated as tunable if it is not known or varies slightly with \emph{e.g.} seeing, slit-or-fiber illumination, or instrumental configuration: $30,003$.
-\edit2{The pseudocontinuum polynomial term $\mathsf{P}$ can also be tuned to refine the continuum placement initialized at the pre-processing stage, yielding an additional 5-15 parameters, depending on the choice of polynomial order.}
+The pseudocontinuum polynomial term $\mathsf{P}$ can also be tuned to refine the continuum placement initialized at the pre-processing stage, yielding an additional 5-15 parameters, depending on the choice of polynomial order.
 
@@
 The interpretations of these EWs may have to be treated with care if intercompared with EWs obtained conventionally.
-This approach may suit stellar abundance applications, such as Galactic archaeology and \edit1{chemically peculiar stars \citep[\emph{e.g.
-HgMn stars,}][]{2020MNRAS.496..832C}}.
+This approach may suit stellar abundance applications, such as Galactic archaeology and chemically peculiar stars \citep[\emph{e.g.
+HgMn stars,}][]{2020MNRAS.496..832C}.
 
@@
 
-\edit1{\emph{``Fancy interpolator''}--  A linear interpolator has to store the slope, offsets, and start and end points of all the line segments from pixel-to-pixel, merely to recreate a spectrum losslessly at some arbitrary coordinate or coordinates.
+\emph{``Fancy interpolator''}--  A linear interpolator has to store the slope, offsets, and start and end points of all the line segments from pixel-to-pixel, merely to recreate a spectrum losslessly at some arbitrary coordinate or coordinates.
 \emph{Blas\'e} lossily recreates a spectrum, while having to store less information.
-In this way, it is fair to think of \emph{Blas\'e} as an interpolator that happens to have an intuitive interpretation.}
-
-\emph{``Surrogate modeling / Simulation Intelligence''}-- \edit1{Our slightly off-label use of machine learning straddles a boundary of machine learning and conventional physics-based models.
+In this way, it is fair to think of \emph{Blas\'e} as an interpolator that happens to have an intuitive interpretation.
+
+\emph{``Surrogate modeling / Simulation Intelligence''}-- Our slightly off-label use of machine learning straddles a boundary of machine learning and conventional physics-based models.
 The rapidly evolving field of machine learning may therefore classify the approach here using a few related terms.
@@
 The key idea is that information from expensive physics-based simulations can be distilled by machine learning and then applied cheaply to new examples.
-The performance of these surrogates should improve when more scientific domain knowledge is included at training time \citep{2018ApJ...869L...7A}.}
-
-\emph{``Narrowing the Synthetic Gap'}-- \edit2{Our approach to pretraining on synthetic spectra and comparing to data resembles the technique of \citet{2020A&A...642A..22P}.
+The performance of these surrogates should improve when more scientific domain knowledge is included at training time \citep{2018ApJ...869L...7A}.
+
+\emph{``Narrowing the Synthetic Gap'}-- Our approach to pretraining on synthetic spectra and comparing to data resembles the technique of \citet{2020A&A...642A..22P}.
 There, a convolutional neural network is pretrained on PHOENIX spectra, and then applied to CARMENES spectra of M dwarfs.
 \citet{2020A&A...642A..22P} designate the difference in appearance between the synthetic spectra and observed spectra as the ``synthetic gap'', showing that this gap leads to inaccurate metallicity inferences.
-\emph{Blas\'e} shows a path towards shrinking the synthetic gap, with the prospect of re-training convolutional neural networks with semi-empirical spectra produced via the method described here, assuming high-fidelity metallicity labels can be obtained through co-natal binary studies.}
+\emph{Blas\'e} shows a path towards shrinking the synthetic gap, with the prospect of re-training convolutional neural networks with semi-empirical spectra produced via the method described here, assuming high-fidelity metallicity labels can be obtained through co-natal binary studies.
 
@@
 
-\edit1{One key limitation stems from the need for autodiff to ``know about'' or sense each step of the end-to-end computation.
+One key limitation stems from the need for autodiff to ``know about'' or sense each step of the end-to-end computation.
 Practically speaking, this requirement means \texttt{blase} had to be written in a single differentiable (\emph{i.e.} autodiff-aware) programming framework.
@@
 PyTorch had already been applied to several astronomy projects by this time, including \texttt{Exonet} \citep{2018ApJ...869L...7A}, \texttt{PySR} \citep{pysr}, and \texttt{MPoL} \citep{mpol}.
-Ultimately, the choice of \texttt{PyTorch} made it easy to employ Sparse Matrices.}
-
-\edit1{This vendor lock-in gives rise to a frustrating incompatibility among exoplanet-relevant autodiff applications.
+Ultimately, the choice of \texttt{PyTorch} made it easy to employ Sparse Matrices.
+
+This vendor lock-in gives rise to a frustrating incompatibility among exoplanet-relevant autodiff applications.
 For example, no differentiable piece of code in \texttt{blase} can call an \texttt{exoplanet} orbital radial velocity, a \texttt{starry} surface map, an \texttt{astropy} spectral coordinate transformation, a \texttt{wobble} spectral residual model, or an \texttt{ExoJAX} Voigt profile.
@@
 We use \texttt{SciPy}, but only for preprocessing, before the differentiable model has been entered.
-Some heuristics exist for interchanging code among the frameworks, but they pose some implementation complexity and/or some performance hit.}
-
-\edit1{It may appear attractive to coordinate the relatively small autodiff-for-astrophysics community to adopt a single autodiff-able machine learning framework to facilitate compatibility.
+Some heuristics exist for interchanging code among the frameworks, but they pose some implementation complexity and/or some performance hit.
+
+It may appear attractive to coordinate the relatively small autodiff-for-astrophysics community to adopt a single autodiff-able machine learning framework to facilitate compatibility.
 In practice, such coordination may be ill-fated, since the rapid pace of machine learning development has not yet settled on a single best way forward.
@@
 For example, \texttt{JAX} may offer speedups attributable to Just In Time (JIT) compilation, at the expense of a more rigid functional programming style.
-Access to scalable Gaussian Process implementations may be another deciding factor \citep{2022arXiv220908940A}.}
+Access to scalable Gaussian Process implementations may be another deciding factor \citep{2022arXiv220908940A}.
 
@@
 \subsection{Identifying missing lines}\label{secMissingLines}
-\edit1{So far we have not addressed the inevitable prospect that some spectral line or lines reside in the observed data spectrum, but are absent entirely from the precomputed model.
+So far we have not addressed the inevitable prospect that some spectral line or lines reside in the observed data spectrum, but are absent entirely from the precomputed model.
 This ``missing line'' scenario arises from our incomplete and evolving knowledge of quantum mechanics and molecular chemistry.
@@
 \emph{Blas\'e} as it currently stands has no mechanism for identifying these lines.
-Here we consider that there are $N_\mathrm{missing}$ lines with two scenarios: \emph{I)} the identification, line center location, and possibly other ancillary information about the lines have become newly available,  or \emph{II)} the lines are truly anomalous and unknown.}
-
-\edit1{For the first scenario, Voigt profiles could be instantiated at the newly-established line center locations $\lambda_{c,m}$ of the previously unknown line.
-So Equation \ref{equation1} would become a product of the new and old physics:}
+Here we consider that there are $N_\mathrm{missing}$ lines with two scenarios: \emph{I)} the identification, line center location, and possibly other ancillary information about the lines have become newly available,  or \emph{II)} the lines are truly anomalous and unknown.
+
+For the first scenario, Voigt profiles could be instantiated at the newly-established line center locations $\lambda_{c,m}$ of the previously unknown line.
+So Equation \ref{equation1} would become a product of the new and old physics:
 
@@
 
-\noindent \edit1{Scenario \emph{II}---truly unknown lines---would not have the luxury of knowing where to initialize the line center location, at least not based on precise theory.
+\noindent Scenario \emph{II}---truly unknown lines---would not have the luxury of knowing where to initialize the line center location, at least not based on precise theory.
 So instead, an educated guess for $\lambda_\mathrm{c,m}$ could be made based on the data centroid.
-}
-
-\edit1{In either case, the initialization of $A_m, \sigma_m, \gamma_m$ would be a matter of taste, and they would therefore be difficult to regularize, at least in the manner that we have described so far.
+
+In either case, the initialization of $A_m, \sigma_m, \gamma_m$ would be a matter of taste, and they would therefore be difficult to regularize, at least in the manner that we have described so far.
 For these reasons, an even better strategy may be to treat the data residuals not as lines, but in some other basis.
 For example, it could be possible to model the \emph{blas\'e} residual spectrum with, say, \texttt{wobble}, or with a Gaussian process.
-These questions remain an open research area.}
+These questions remain an open research area.
 
@@
 
-It is easier to \edit1{infer} longitudinally asymmetric features than longitudinally symmetric ones, since we assume that we occasionally catch a glimpse of the spot-free limbs and their pristine line profiles.
+It is easier to infer longitudinally asymmetric features than longitudinally symmetric ones, since we assume that we occasionally catch a glimpse of the spot-free limbs and their pristine line profiles.
 The latter requires exact knowledge of the underlying spectral template.
@@
 The use of these labels could take a few different forms.
-\edit1{The labels could simply serve as flags, to indicate that regularization should be weakened to allow these magnetic sensitive lines to wander farther from their naive non-magnetic expectations.
+The labels could simply serve as flags, to indicate that regularization should be weakened to allow these magnetic sensitive lines to wander farther from their naive non-magnetic expectations.
 Or the labels could be used to infer the extent of Zeeman broadening or Zeeman splitting by parameterizing these operations as convolutions.
-The best choice for how to incorporate magnetic sensitivity into the flexibility of \emph{blas\'e} remains an open research question.}
+The best choice for how to incorporate magnetic sensitivity into the flexibility of \emph{blas\'e} remains an open research question.
 
@@
 \subsection{Sky emission lines and wavelength calibration}
-\edit2{Night sky emission lines---arising from OH, for example---add emission line spikes on top of the astronomical target spectrum of interest.
+Night sky emission lines---arising from OH, for example---add emission line spikes on top of the astronomical target spectrum of interest.
 These lines get subtracted in most pipeline packages, but in some cases, a more careful treatment may be desirable.
@@
 The main benefit over conventional sky subtraction methods may be the ability to disentangle the underlying spectrum at super-resolution.
-A predictive model for sky emission could be built-up over time, building data-driven corrections to theoretical models.}
-
-\edit2{The well-known center wavelengths of these lines could be used for additional wavelength calibration, as is already done in IGRINS \citep{jaejoonlee16}.
+A predictive model for sky emission could be built-up over time, building data-driven corrections to theoretical models.
+
+The well-known center wavelengths of these lines could be used for additional wavelength calibration, as is already done in IGRINS \citep{jaejoonlee16}.
 An imperfect instrumental wavelength calibration could be diagnosed based on the departure of the predicted versus realized wavelength center positions, enabling a relatively fast and intervention-free correction strategy.
 Hypothetically, a prescription for the wavelength calibration could be built directly into \emph{blas\'e}, with the coefficents for a Polynomial re-routing the realized wavelength center locations from their known wavelength centers.
-}
 
@@
 
-    \edit1{We thank the anonymous referee for comments that improved the paper.}  \edit2{We thank Phill Cargile, Hajime Kawahara, and Dan Jaffe for pre-print comments that were incorporated into the draft text.}
+    We thank the anonymous referee for comments that improved the paper.  We thank Phill Cargile, Hajime Kawahara, and Dan Jaffe for pre-print comments that were incorporated into the draft text.
     MGS thanks Ian Czekala, Phill Cargile, Rodrigo Luger, Greg Herczeg, Will Best, Dan Foreman-Mackey, Meg Bedell, Mark Marley, Erwan Pannier, Dirk C.M.

### 2022-10-03 paper/paper1/ms.tex
Some changes to get the arxiv version to render

@@
 
+\pagebreak
+\newpage
+
 \begin{acknowledgments}
-
-    We thank the anonymous referee for comments that improved the paper.  We thank Phill Cargile, Hajime Kawahara, and Dan Jaffe for pre-print comments that were incorporated into the draft text.
-    MGS thanks Ian Czekala, Phill Cargile, Rodrigo Luger, Greg Herczeg, Will Best, Dan Foreman-Mackey, Meg Bedell, Mark Marley, Erwan Pannier, Dirk C.M.
-van den Bekerom, Dan Clemens, Dan Jaffe, Tom Greene, Adam Kraus, David Hogg, Greg Zeimann, Ben Montet, Christina Hedges, Brittany Miles, Arpita Roy, and Kevin Gullikson for conversations and resources that shaped his thinking on spectral calibration and telluric mitigation.
+    \footnotesize{
+        We thank the anonymous referee for comments that improved the paper.  We thank Phill Cargile, Hajime Kawahara, and Dan Jaffe for preprint comments.
+
+        This material is based upon work supported by the National Aeronautics and Space Administration under Grant Numbers 80NSSC21K0650 for the NNH20ZDA001N-ADAP:D.2 program,
+        and 80NSSC20K0257 for the XRP program issued through the Science Mission Directorate.
+
+        We acknowledge the National Science Foundation, which supported the work presented here under Grant No. 1910969.
 
@@
 
-    This material is based upon work supported by the National Aeronautics and Space Administration under Grant Numbers 80NSSC21K0650 for the NNH20ZDA001N-ADAP:D.2 program,
-    and 80NSSC20K0257 for the XRP program issued through the Science Mission Directorate.
-
-    We acknowledge the National Science Foundation, which supported the work presented here under Grant No. 1910969.
-
-    This research has made use of NASA's Astrophysics Data System Bibliographic Services.
-
+        MGS thanks Ian Czekala, Phill Cargile, Rodrigo Luger, Greg Herczeg, Will Best, Dan Foreman-Mackey, Meg Bedell, Mark Marley, Erwan Pannier, Dirk C.M.
+van den Bekerom, Dan Clemens, Dan Jaffe, Tom Greene, Adam Kraus, David Hogg, Greg Zeimann, Ben Montet, Christina Hedges, Brittany Miles, Arpita Roy, and Kevin Gullikson for conversations and resources that shaped his thinking on spectral calibration and telluric mitigation.
+    }
 \end{acknowledgments}

### 2022-11-07 paper/paper1/ms.tex
Add changes in the proofs stage

@@
 \emph{``Reverse-engineering the line lists''}-- Equation \ref{equation1} (and its exponential twin Equation \ref{equationOpacitySum}) could have been populated with initial line-by-line properties from atomic and molecular line lists, instead of the coarse detection-threshold approach we presented in Section \ref{subsecInit}.
-So our cloning procedure can be viewed as reverse-engineering the quantum mechanical properties housed in line lists such as HITRAN, HITEMP, EXOMOL, VALD, or countless other primary source documents.
+So our cloning procedure can be viewed as reverse-engineering the quantum mechanical properties housed in line lists such as HITRAN \citep{2022JQSRT.27707949G}, HITEMP \citep{2010JQSRT.111.2139R}, EXOMOL \citep{2012MNRAS.425...21T}, VALD \citep{1995A&AS..112..525P}, or countless other primary source documents.
 This interpretation is only partly accurate.
@@
 A sum of unlike-Voigt profiles is not exactly equal to any single Voigt profile.
-Theoreticians may resonate with this more ``first principles'' representation, while data practitioners may find Equation \ref{equation1} more natural, so to some extent the choice is a matter of taste.
+Theoreticians may resonate with this more ``first principles'' representation, while data practitioners may find Equation \ref{equation1} more natural, so to some extent, the choice is a matter of taste.
 
@@
 First and foremost, real stars are usually more complicated than our simplified models of them.
-Real spectra often vary over more dimensions that our models do.
+Real spectra often vary over more dimensions than our models do.
 Conspicuous examples of these hidden variables can be found in protostars: starspots, accretion veiling, dust extinction, and magnetic Zeeman splitting.
@@
 
-For very large bandwidths and very many spectral lines, the problem of identifying and cataloging line imperfections essentially becomes a book-keeping and continuum assignment problem.
+For very large bandwidths and very many spectral lines, the problem of identifying and cataloging line imperfections essentially becomes a bookkeeping and continuum assignment problem.
 \texttt{blase} and \texttt{starfish} provide different strategies for orchestrating the line-mismatch identification procedure, with each route having tradeoffs depending on the application.
@@
 Blas\'e already performs very well under a wide range of cloning and transfer-learning tasks.
-However, some precision applications may demand even-more-strenuous performance than what the current implementation can accomodate.
-Here we describe some of these technical improvements, their design, and/or mention some science case they may unlock.
+However, some precision applications may demand even-more-strenuous performance than what the current implementation can accommodate.
+Here we describe some of these technical improvements, their design, and/or mention some science cases they may unlock.
 
@@
 Tuning $\mathsf{P}$ would correct for large-scale imperfections in the otherwise-fixed continuum flattening procedure.
-This change would be easy and effective, but has some challenge with model selection and flexibility: how to set the polynomial order to avoid over- and under-fitting.
+This change would be easy and effective, but has some challenges with model selection and flexibility: how to set the polynomial order to avoid over- and under-fitting.
 Gaussian Processes (GPs) offer many advantages for continuum fitting \citep{czekala15}.
@@
 Fortunately, a few efficient autodiff-aware implementations of GPs exist.
-The \texttt{celerite} algorithm \emph{celerit\'e} \citep{2017AJ....154..220F} has an exact backpropation implementation \citep{2018RNAAS...2...31F} that scales linearly with the number of data points.
+The \texttt{celerite} algorithm \emph{celerit\'e} \citep{2017AJ....154..220F} has an exact backpropagation implementation \citep{2018RNAAS...2...31F} that scales linearly with the number of data points.
 The \emph{celerit\'e} algorithm does not currently have a PyTorch implementation.
-The \texttt{GPyTorch} framework \citep{2018arXiv180911165G} has a large category of approximate and exact GPs that could be straight-forwardly dropped into \texttt{blase}.
+The \texttt{GPyTorch} framework \citep{2018arXiv180911165G} has a large category of approximate and exact GPs that could be straightforwardly dropped into \texttt{blase}.
 Even still these GPs could increase the computation cost by of order $10\times$.
@@
 \subsection{Wavelength dependent limb darkening}
-Currently the extrinsic model step possesses up to four parameters: the $v\sin{i}$ and $RV$, and 2 optional parameters for limb-darkening.
+Currently, the extrinsic model step possesses up to four parameters: the $v\sin{i}$ and $RV$, and 2 optional parameters for limb-darkening.
 These four parameters may adequately parameterize a star with a uniform stellar disk.
@@
 The limb darkening may instead depend on physical properties of the spectral line formation, such as physical depth of formation, and so the extent of limb-darkening may jump haphazardly from line-to-line-to-line, rather than as a predictably smooth function across wavelength.
-\texttt{Blas\'e} could be built to handle such a seemingly pathological scenario by adding a vector of limb darkening parameters, one for each line.
+\texttt{Blas\'e} could be built to handle such a seemingly pathological scenario by adding a vector of limb-darkening parameters, one for each line.
 One would have to regularize the fits with some typical limb darkening and a heuristic penalty for departures from this mean.

### 2022-11-07 docs/conf.py
the RTD build has been failing for months! whoops.  turn off notebook builds for now...

@@ -72,6 +72,10 @@ nbsphinx_thumbnails = {
 #
 html_theme = "sphinx_book_theme"
 
+nbsphinx_allow_errors = True
+
+nbsphinx_execute = "never"
+
 pygments_style = "vs"
 
 # Add any paths that contain custom static files (such as style sheets) here,
@@ -85,5 +89,5 @@ html_theme_options = {
     "use_repository_button": True,
     "repository_branch": "main",
     "home_page_in_toc": True,
-    "announcement": "🆕 blasé will be featured in a talk at the Machine Learning Splinter Session at the Cool Star 21 Conference in Toulouse, France on July 5th, 2022!",
+    "announcement": "Our paper is now published! https://arxiv.org/abs/2210.01827",
 }

### 2023-10-06 src/blase/emulator.py
uncomitted changes

@@ -9,6 +9,7 @@ import torch
 from torch import nn
 import numpy as np
 from scipy.signal import find_peaks, peak_prominences, peak_widths
+from scipy.special import voigt_profile
 import torch.optim as optim
 from tqdm import trange
 from torch.special import erfc
@@ -329,15 +330,44 @@ class LinearEmulator(nn.Module):
         raise NotImplementedError
     
 
-    def animate(self, old_state_dict):
+    def animate(self, old_state_dict, size=1.5):
         """Animate the model from a previous state to the current state"""
         try: 
             import manim
         except ImportError:
             print("Manim is required for the .animate() feature, but it is not installed.  Please install manim and try again.")
-            return
+            return None
+        
+        ## Manim requires a scene class, so we'll make one here
+        class ArgMinExample(manim.Scene, size=0.9):
+            def construct(self, size):
+                ax = manim.Axes(
+                    x_range=[-10, 10], y_range=[0, 1, 0.25], axis_config={"include_tip": False}
+                )
+                labels = ax.get_axis_labels(x_label=manim.Tex(r"$\lambda$"), 
+                                            y_label=manim.Tex(r"$f(\lambda)$"))
+                
+                graph1 = ax.plot(lambda x: 1 - 2*voigt_profile(x, 1.0, 0), color=manim.MAROON)
+                graph2 = ax.plot(lambda x: 1 - 2*voigt_profile(x, 1.0, size), color=manim.MAROON)
+            
+                # Plot noisy data with manim:
+                x = np.linspace(-10, 10, 100)  
+                y = 1 - 2*voigt_profile(x, 1.0, size)
+                y += np.random.normal(0, 0.03, y.shape)
+                coords = np.vstack((x,y)).T
 
-        raise NotImplementedError
+                dots = manim.VGroup(*[manim.Dot().move_to(ax.c2p(coord[0],coord[1])) for coord in coords])
+                self.add(dots)
+
+            
+                self.add(ax, labels)
+                self.play(manim.Create(graph1))
+                self.play(manim.Transform(graph1, graph2))
+                self.play(manim.FadeOut(graph2))
+
+        scene = ArgMinExample(size=size)
+        scene.render(preview=False) # That's it!
+        return scene.renderer.file_writer.movie_file_path
 
 
 

### 2023-10-09 paper/paper2/ms.tex
Paper changes, torch device util, try .env

\documentclass[twocolumn]{aastex631}
\bibliographystyle{aasjournal}

\usepackage{amsmath}
\usepackage{blindtext}

\usepackage{mdframed}

\begin{document}
\title{Blas\'e3D}
\shorttitle{\emph{blas\'e3D}}

\author[0000-0002-2290-6810]{Sujay Shankar}
\author[0000-0002-4020-3457]{Michael Gully-Santiago}
\author[0000-0002-4404-0456]{Caroline V. Morley}
\affil{Department of Astronomy, The University of Texas at Austin, 2515 Speedway, Austin, TX 78712, USA}
\shortauthors{Shankar \& Gully-Santiago \& Morley}

\begin{abstract}
    \blindtext
\end{abstract}

\keywords{}

\section{Introduction}
\blindtext

\section{Data}
\begin{mdframed}
    \textbf{PHOENIX Model Grid Subset}

    \textcolor{lightgray}{\blindtext}
\end{mdframed}

\begin{mdframed}
    \textbf{Figure: Line Population Heatmap}
\end{mdframed}

\section{Methods}
\begin{mdframed}
    \textbf{Illustration}
\end{mdframed}

\begin{mdframed}
    \textbf{Model Optimization}

    - Per-grid point optimization procedure
    - Optimization outputs
    - Output storage

    \textcolor{lightgray}{\blindtext}
\end{mdframed}

\begin{mdframed}
    \textbf{Line Recognition}

    - Line set
    - Pre-shift centers

    \textcolor{lightgray}{\blindtext}
\end{mdframed}

\begin{mdframed}
    \textbf{Math of the Surface Fit}

    \textcolor{lightgray}{\blindtext}
\end{mdframed}

\section{Discussion}
\begin{mdframed}
    \textbf{Revisiting Model Assumptions}

   \textcolor{lightgray}{\blindtext}
\end{mdframed}

\begin{mdframed}
    \textbf{Limitations}

    - Computational resources
    - Line profile inaccuracy
    - Surface functional form

    \textcolor{lightgray}{\blindtext}
\end{mdframed}

\begin{mdframed}
    \textbf{Conceivable Extensions}

    \textcolor{lightgray}{\blindtext}
\end{mdframed}

\pagebreak
\newpage

\begin{acknowledgments}
    \blindtext
\end{acknowledgments}

\software{}

\bibliography{ms}
\clearpage

\appendix
\section{Linear Regression}
\blindtext

\end{document}

### 2023-10-09 src/blase/utils.py

@@ -4,23 +4,24 @@ utilities
 
 These general purpose methods are used in one or more other modules.
 """
-from contextlib import contextmanager
-import sys, os
 import numpy as np
+import torch
+
 from astropy import constants as const
 from astropy import units as u
+from contextlib import contextmanager
+from os import devnull
+from sys import stdout
 
 
 @contextmanager
 def suppress_stdout():
-    with open(os.devnull, "w") as devnull:
-        old_stdout = sys.stdout
-        sys.stdout = devnull
+    with open(devnull, "w") as devnull:
+        old_stdout, stdout = stdout, devnull
         try:
             yield
         finally:
-            sys.stdout = old_stdout
-
+            stdout = old_stdout
 
 def doppler_grid(lambda_0, lambda_max, velocity_resolution_kmps=0.5):
     """Compute the exponential sampling interval for a given lambda_0 and velocity resolution."""
@@ -28,3 +29,12 @@ def doppler_grid(lambda_0, lambda_max, velocity_resolution_kmps=0.5):
     velocity_max = c_kmps * np.log(lambda_max / lambda_0)
     velocity_vector = np.arange(0, velocity_max, velocity_resolution_kmps)
     return lambda_0 * np.exp(velocity_vector / c_kmps)
+
+def auto_torch_device():
+    if torch.cuda.is_available():
+        return torch.device("cuda")
+    else:
+        try:
+            return torch.device("mps")
+        except RuntimeError:
+            return torch.device("cpu")

### 2023-10-10 paper/paper2/ms.tex
Tiny comment
Paper work
Pair coding on the outline, fun!

@@
 \usepackage{blindtext}
-
+% Temporary packages
 \usepackage{mdframed}
@@
 
-\section{Data}
+\section{Emulating the PHOENIX grid}
+
 \begin{mdframed}
-    \textbf{PHOENIX Model Grid Subset}
+    \textbf{The PHOENIX grid subset}
 
@@
 \begin{mdframed}
-    \textbf{Figure: Line Population Heatmap}
-\end{mdframed}
-
-\section{Methods}
-\begin{mdframed}
-    \textbf{Illustration}
-\end{mdframed}
-
-\begin{mdframed}
-    \textbf{Model Optimization}
+    \textbf{Emulation with blas\'e}
 
     - Per-grid point optimization procedure
-    - Optimization outputs
+    - Pretrained model caching
     - Output storage
@@
 \begin{mdframed}
-    \textbf{Line Recognition}
-
-    - Line set
-    - Pre-shift centers
+    \textbf{Open Source Availability}
 
@@
 \begin{mdframed}
-    \textbf{Math of the Surface Fit}
+    \textbf{Figure: Line density across the grid heatmap}
+\end{mdframed}
+
+\section{Line-by-Line Fundamental Stellar Properties}
+\begin{mdframed}
+    \textbf{Conceptual Illustration: Faceted Plot (Teff, Logg, Z) -> Line Profiles}
+\end{mdframed}
+
+\begin{mdframed}
+    \textbf{Line Recognition}
+    - Need: Identifying unique lines
+    - Anticipated friction points: line centroids drift
+    - Strategy: (Pre-shift centers)
+    \textcolor{lightgray}{\blindtext}
+\end{mdframed}
+
+\begin{mdframed}
+    \textbf{Line property trends across the grid}
+    \textcolor{lightgray}{\blindtext}
+\end{mdframed}
+
+\begin{mdframed}
+    \textbf{Figure: Heatmap $T_\mathrm{eff}$ vs $\log{g}$ for a single line $2\times2$ panels for $\sigma$, $\gamma$, $A$, $\lambda$ }
+\end{mdframed}
+
+\begin{mdframed}
+    \textbf{Line-by-Line Parameter Surfaces}
 

### 2023-10-10 paper/paper2/ms.tex
Pair coding V3
Neat ideas for repittion of a flowchart!
Pair coding V5

@@
 \blindtext
+
+\begin{mdframed}
+    \textbf{Figure: Flowchart-with-words}
+\end{mdframed}
 
@@
 \begin{mdframed}
-    \textbf{Line property trends across the grid}
+    \textbf{Line property bulk trends across the grid}
     \textcolor{lightgray}{\blindtext}
@@
 \begin{mdframed}
-    \textbf{Line-by-Line Parameter Surfaces}
+    \textbf{Anomalies in Heatmaps}
+    \textcolor{lightgray}{\blindtext}
+\end{mdframed}
 
+\section{Mapping Line Parameters to Fundamental Properties}
+\begin{mdframed}
+    \textbf{Statement: A Bidirectonal Relation exists}
+    - Goal: identify a functional form
+    \textcolor{lightgray}{\blindtext}
+\end{mdframed}
+
+\begin{mdframed}
+    \textbf{Figure: Flowchart-with-equations}
+\end{mdframed}
+
+\begin{mdframed}
+    \textbf{Functional Form}
+    - Describe functional form options and refinement procedures
+    - List possibilities, Linreg, GPs, NNs, etc.
+    - We choose LSTSQ
+    - Enumerate functional form
+    - Adapting model complexity
+    \textcolor{lightgray}{\blindtext}
+\end{mdframed}
+
+\begin{mdframed}
+    \textbf{Address what to do with missing lines}
     \textcolor{lightgray}{\blindtext}

### 2023-10-10 paper/paper2/ms.tex
End of pair-coding session on the paper outline, super helpful!

@@
     - Describe functional form options and refinement procedures
-    - List possibilities, Linreg, GPs, NNs, etc.
+    - List possibilities, Linreg, GPs \citep{2023ARA&A..61..329A}, NNs, etc.
     - We choose LSTSQ
     - Enumerate functional form
-    - Adapting model complexity
+    - Adapting model complexity-- AIC
     \textcolor{lightgray}{\blindtext}
@@
 \begin{mdframed}
-    \textbf{Address what to do with missing lines}
+    \textbf{Problem: missing lines}
+    - Conceivable Solution 1: treat as NaNs and deal with sparsity
+    - Conceivable Solution 2: increase model complexity
+    \textcolor{lightgray}{\blindtext}
+\end{mdframed}
+
+\section{Performance evaluation}
+
+\begin{mdframed}
+    \textbf{Typical Line reconstruction performance}
+    \textcolor{lightgray}{\blindtext}
+\end{mdframed}
+
+\begin{mdframed}
+    \emph{stretch goal}\par
+    \textbf{End-to-end PHOENIX grid replication and residual}
+    - State the per-pixel residual
     \textcolor{lightgray}{\blindtext}

### 2023-10-24 experiments/09_surface_fitting/09_surface_fitting.py
Experiment 09 + Schematic playground

@@ -4,7 +4,7 @@ import torch
 
 from collections import defaultdict
 from functools import reduce
-from numpy.polynomial.polynomial import polyvander3d
+from numpy.polynomial.polynomial import polyvander2d, polyvander3d
 from os import listdir
 from re import split
 from scipy.linalg import lstsq
@@ -39,9 +39,12 @@ def main():
     df[fcols] = df[fcols].apply(pd.to_numeric, downcast='float')
     df[icols] = df[icols].apply(pd.to_numeric, downcast='integer')
 
-    current_line = line_set[1]
-    MAX_DEGREE = 2
+    line_counts = df.value_counts('center')
+
+    current_line = line_counts.index[0]
+    MAX_DEGREE = 3
     df_line = df.query('center == @current_line')
+    '''
     print('Creating Design and Observation Matrices...')
     design_matrix = polyvander3d(df_line['amp'], df_line['sigma'], df_line['gamma'], deg=(MAX_DEGREE, MAX_DEGREE, MAX_DEGREE)).astype(float)
     observation_matrix = df_line[['teff', 'logg', 'Z']].to_numpy(dtype=float)
@@ -54,6 +57,25 @@ def main():
     print(f'{MAX_DEGREE}-th Order Raw Model AIC: {AIC}')
     # Loop over design matrix, and test drop columns to improve AIC
     # Refined Model AIC here
+    manifold = design_matrix @ coefficients
+    df_visual = pd.DataFrame({'amp': df_line['amp'], 'sigma': df_line['sigma'], 'gamma': df_line['gamma'], 'teff': manifold[:, 0], 'logg': manifold[:, 1], 'Z': manifold[:, 2]})
+    print('Writing Manifold to Parquet...')
+    df_visual.to_parquet('experiments/09_surface_fitting/surface_visualization_sparse.parquet.gz', compression='gzip')'''
+
+    print('Creating Design and Observation Matrices...')
+    design_matrix = polyvander2d(df_line['teff'], df_line['logg'], deg=(7, 7)).astype(float)
+    observation_matrix = df_line[['amp', 'sigma', 'gamma']].to_numpy(dtype=float)
+    print('Solving for Least Squares Coefficients...')
+    coefficients, residuals = lstsq(design_matrix, observation_matrix)[0:2]
+    print('Evaluating Akaike Information Criterion...')
+    loss_function = np.sum(residuals**2)
[…]

### 2023-10-24 paper/paper2/ms.tex
Add the Paper 1 deluxe table of notation for editing

@@
 \appendix
-\section{Linear Regression}
-\blindtext
+\section{Notation}
+
+\begin{deluxetable}{cp{10cm}}
+    \tabletypesize{\scriptsize}
+    \tablecaption{Notation used in this paper\label{table2}}
+    \tablehead{
+        \colhead{Symbol} & \colhead{Meaning}
+    }
+    \startdata
+    \hline
+    \multicolumn{2}{c}{Spectra}\\
+    \hline
+    $\bm{\lambda}_S$ & Native wavelength coordinates of the precomputed stellar spectrum\\
+    $\bm{\lambda}_T$ & Native wavelength coordinates of the telluric spectrum\\
+    $\bm{\lambda}_D$ & Native wavelength coordinates of the data spectrum\\
+    $\mathsf{S}_{\rm abs}$ & Flux values of the precomputed synthetic stellar spectral model $\bm{\lambda}_S$\\
+    $\mathsf{B}$ & Blackbody of temperature $T_{\mathrm{eff}}$ to coarsely normalize $\mathsf{S}_{\rm native}$\\
+    $\mathsf{P}$ & Smooth polynomial to refine continuum-normalization\\
+    $\mathsf{S}$ & Continuum normalized augmentation of $\mathsf{S}_{\rm abs}$\\
+    $\mathsf{T}$ & Transmission values of the precomputed synthetic telluric model \\
+    $\mathsf{D}$ & The observed data spectrum flux values\\
+    $\bm{\epsilon}$ & The estimated uncertainties in the data spectrum\\
+    $\mathsf{S}_{\rm clone}$ & Evaluable and tunable cloned flux model of $\mathsf{S}$\\
+    $\mathsf{T}_{\rm clone}$ & Evaluable and tunable cloned transmission model of $\mathsf{T}$\\
+    $\mathsf{S}_{\rm ext}$ & An augmentation of $\mathsf{S}_{\rm clone}$ with $v\sin{i}$ convolution and $RV$ translation\\
+    $\mathsf{M}_{\rm joint}$ & The joint stellar and telluric model: $\mathsf{S}_{\rm ext} \odot \mathsf{T}_{\rm clone}(\bm{\lambda}_S)$ \\
+    $\mathsf{M}$ & Joint model convolved
+    with instrumental kernel and resampled to $\bm{\lambda}_D$\\
+    $\mathsf{R}$ & The residual spectrum between a pair of inputs, \emph{e.g.} $\mathsf{D} - \mathsf{M}$\\
+    $\bm{v}$ & The spectral coordinate axis $\bm{\lambda}$ expressed as a velocity difference\\
+    \hline
+    \multicolumn{2}{c}{Line properties}\\
+    \hline
+    $\lambda_{\mathrm{c},j}$ & Line center position of the $j^{th}$ spectral line\\
+    $a_j$ & Gaussian line profile amplitude of the $j^{th}$ spectral line \\
+    $\sigma_j$ & Gaussian line profile scale of the $j^{th}$ spectral line\\
+    $\gamma_j$ & Lorentzian line profile half width of the $j^{th}$ spectral line\\
+    $\mathsf{V}_j$ & The Voigt profile of the $j^{th}$ spectral line \\
+    $\bar{\bm{F}}$ & The dense $(N_{\rm lines} \times N_{x})$ matrix of all line fluxes stacked vertically \\
+    $\hat{\bm{F}}$ & The sparse $(N_{\rm lines} \times N_{\rm sparse})$ matrix of all line fluxes stacked vertically \\
+    $\zeta$ & The rotational broadening convolution kernel\\
+    $g$ & The instrumental broadening convolution kernel, typically a Gaussian\\
+    \hline
+    \multicolumn{2}{c}{Scalars}\\
+    \hline
+    $N_{\rm lines}$ & Number of spectral lines \\
+    $N_{x}$ & Number of pixel coordinates in the precomputed spectrum $\bm{\lambda}_x$\\
+    $N_{\rm sparse}$ & Number of non-zero pixels computed in the sparse implementation\\
+    $\pm \Delta \lambda_{\mathrm{buffer}}$ & Buffer exceeding the red and blue limits of the data spectrum\\
+    $P_{\rm rom}$ & The prominence threshold of spectral lines to include in cloning \\
+    $v\sin{i}$ & Rotational broadening for stellar inclination $i$ and equatorial velocity $v$\\
+    $RV$ & Radial velocity of the star\\
+    $R$ & Spectrograph resolving power $\lambda/\delta\lambda$\\
+    $\mathcal{L}$ & The loss scalar, usually the sum of the squares of the residuals\\
+    \hline
+    \multicolumn{2}{c}{Operators}\\
+    \hline
+    $\resample \big[ \mathsf{F(\bm{\lambda}_x)} \big]$ & The resample operator, takes in a flux spectrum $\mathsf{F}$ evaluated at $\bm{\lambda}_x$ coordinates and returns the mean flux within the pixel boundaries of coordinate $\bm{\lambda}_z$\\
+    $*$& The convolution operator\\
+    $\odot$& \emph{Hadamard product}, an elementwise product of two same-length vectors\\
+    \enddata
+\end{deluxetable}
 

### 2023-11-02 experiments/09_surface_fitting/09_surface_fitting.py
2D Interpolation

@@ -4,13 +4,15 @@ import torch
 
 from collections import defaultdict
 from functools import reduce
+from itertools import product
 from numpy.polynomial.polynomial import polyvander2d, polyvander3d
 from os import listdir
 from re import split
+from scipy.interpolate import RegularGridInterpolator
 from scipy.linalg import lstsq
 from tqdm import tqdm
 
-def main():
+def main(scheme=None):
     path = "/home/sujay/data/10K_12.5K_clones"
 
     line_stats = defaultdict(list)
@@ -42,40 +44,42 @@ def main():
     line_counts = df.value_counts('center')
 
     current_line = line_counts.index[0]
-    MAX_DEGREE = 3
     df_line = df.query('center == @current_line')
-    '''
-    print('Creating Design and Observation Matrices...')
-    design_matrix = polyvander3d(df_line['amp'], df_line['sigma'], df_line['gamma'], deg=(MAX_DEGREE, MAX_DEGREE, MAX_DEGREE)).astype(float)
-    observation_matrix = df_line[['teff', 'logg', 'Z']].to_numpy(dtype=float)
-    print('Solving for Least Squares Coefficients...')
-    coefficients, residuals = lstsq(design_matrix, observation_matrix)[0:2]
-    print('Evaluating Akaike Information Criterion...')
-    loss_function = np.sum(residuals**2)
-    AIC = 2 * (coefficients.size - np.log(loss_function))
-    # Use inflect ordinals in the future, because why not
-    print(f'{MAX_DEGREE}-th Order Raw Model AIC: {AIC}')
-    # Loop over design matrix, and test drop columns to improve AIC
-    # Refined Model AIC here
-    manifold = design_matrix @ coefficients
-    df_visual = pd.DataFrame({'amp': df_line['amp'], 'sigma': df_line['sigma'], 'gamma': df_line['gamma'], 'teff': manifold[:, 0], 'logg': manifold[:, 1], 'Z': manifold[:, 2]})
-    print('Writing Manifold to Parquet...')
-    df_visual.to_parquet('experiments/09_surface_fitting/surface_visualization_sparse.parquet.gz', compression='gzip')'''
 
-    print('Creating Design and Observation Matrices...')
-    design_matrix = polyvander2d(df_line['teff'], df_line['logg'], deg=(7, 7)).astype(float)
-    observation_matrix = df_line[['amp', 'sigma', 'gamma']].to_numpy(dtype=float)
[…]

### 2023-11-05 paper/paper2/ms.tex
Comment out deluxetable for now
Add back in the resample operator as a command
Add interpolated grid

--- 2023-10-10 paper/paper2/ms.tex
@@
 \usepackage{blindtext}
+\usepackage{bm}
 % Temporary packages
 \usepackage{mdframed}
+\usepackage{verbatim}
+
+\DeclareMathOperator{\resample}{resample}
+
+\def\Teff{T_{\rm eff}}
+\def\vsini{v\sin{i}}
+\def\kmps{\mathrm{km}\;\mathrm{s}^{-1}}
 
@@
 \appendix
-\section{Linear Regression}
-\blindtext
+\section{Notation}
+We adopt a mostly similar notation to the blas\'e paper, with some modifications.
+
+\begin{deluxetable}{cp{10cm}}[h]
+    \tabletypesize{\small}
+    \tablecaption{Notation used in this paper\label{table2}}
+    \tablehead{
+        \colhead{Symbol} & \colhead{Meaning}
+    }
+    \startdata
+    \hline
+    \multicolumn{2}{c}{Indexes}\\
+    \hline
+    $i$ & Pixel index of the precomputed spectrum\\
+    $j$ & Spectral line index\\
+    $k$ & Grid point index\\
+    \hline
+    \multicolumn{2}{c}{Spectra}\\
+    \hline
+    $\bm{\lambda}$ & Native wavelength coordinate axis of the PHOENIX model grid\\
+    $\mathsf{S}_k$ & Flux values of the $k^{th}$ precomputed spectrum\\
+    $\mathsf{P}_k$ & Smooth polynomial to refine continuum-normalization\\
+    $\mathsf{S}$ & Continuum normalized augmentation of $\mathsf{S}_{\rm abs}$\\
+    $\mathsf{T}$ & Transmission values of the precomputed synthetic telluric model \\
+    $\mathsf{D}$ & The observed data spectrum flux values\\
+    $\bm{\epsilon}$ & The estimated uncertainties in the data spectrum\\
+    $\mathsf{S}_{\rm clone}$ & Evaluable and tunable cloned flux model of $\mathsf{S}$\\
+    $\mathsf{T}_{\rm clone}$ & Evaluable and tunable cloned transmission model of $\mathsf{T}$\\
+    $\mathsf{S}_{\rm ext}$ & An augmentation of $\mathsf{S}_{\rm clone}$ with $v\sin{i}$ convolution and $RV$ translation\\
+    $\mathsf{M}_{\rm joint}$ & The joint stellar and telluric model: $\mathsf{S}_{\rm ext} \odot \mathsf{T}_{\rm clone}(\bm{\lambda}_S)$ \\
+    $\mathsf{M}$ & Joint model convolved
+    with instrumental kernel and resampled to $\bm{\lambda}_D$\\
+    $\mathsf{R}$ & The residual spectrum between a pair of inputs, \emph{e.g.} $\mathsf{D} - \mathsf{M}$\\
+    $\bm{v}$ & The spectral coordinate axis $\bm{\lambda}$ expressed as a velocity difference\\
+    \hline
+    \multicolumn{2}{c}{Line properties}\\
+    \hline
+    $\lambda_{\mathrm{c},j}$ & Line center position of the $j^{th}$ spectral line\\
+    $a_j$ & Gaussian line profile amplitude of the $j^{th}$ spectral line \\
+    $\sigma_j$ & Gaussian line profile scale of the $j^{th}$ spectral line\\
+    $\gamma_j$ & Lorentzian line profile half width of the $j^{th}$ spectral line\\
+    $\mathsf{V}_j$ & The Voigt profile of the $j^{th}$ spectral line \\
+    $\bar{\bm{F}}\bar{F}$ & The dense $(N_{\rm lines} \times N_{x})$ matrix of all line fluxes stacked vertically \\
+    $\hat{\bm{F}}$ & The sparse $(N_{\rm lines} \times N_{\rm sparse})$ matrix of all line fluxes stacked vertically \\
+    $\zeta$ & The rotational broadening convolution kernel\\
+    $g$ & The instrumental broadening convolution kernel, typically a Gaussian\\
+    \hline
+    \multicolumn{2}{c}{Scalars}\\
+    \hline
+    $N_{\rm lines}$ & Number of spectral lines \\
+    $N_{x}$ & Number of pixel coordinates in the precomputed spectrum $\bm{\lambda}_x$\\
+    $N_{\rm sparse}$ & Number of non-zero pixels computed in the sparse implementation\\
+    $\pm \Delta \lambda_{\mathrm{buffer}}$ & Buffer exceeding the red and blue limits of the data spectrum\\
+    $P_{\rm rom}$ & The prominence threshold of spectral lines to include in cloning \\
+    $v\sin{i}$ & Rotational broadening for stellar inclination $i$ and equatorial velocity $v$\\
+    $RV$ & Radial velocity of the star\\
+    $R$ & Spectrograph resolving power $\lambda/\delta\lambda$\\
+    $\mathcal{L}$ & The loss scalar, usually the sum of the squares of the residuals\\
+    \hline
+    \multicolumn{2}{c}{Operators}\\
+    \hline
+    $\resample \big[ \mathsf{F(\bm{\lambda}_x)} \big]$ & The resample operator, takes in a flux spectrum $\mathsf{F}$ evaluated at $\bm{\lambda}_x$ coordinates and returns the mean flux within the pixel boundaries of coordinate $\bm{\lambda}_z$\\
+    $*$& The convolution operator\\
+    $\odot$& \emph{Hadamard product}, an elementwise product of two same-length vectors\\
+    \enddata
+\end{deluxetable}
 

### 2023-11-05 experiments/09_surface_fitting/09_surface_fitting.py

@@ -64,7 +64,7 @@ def main(scheme=None):
         df_visual.to_parquet('experiments/09_surface_fitting/surface_visualization_grid.parquet.gz', compression='gzip')
 
     elif scheme == 'interpolation':
-        METHOD = 'linear' # linear, slinear, cubic, quintic, pchip
+        METHOD = 'slinear' # linear, slinear, cubic, quintic, pchip
         df_line.sort_values(by=['teff', 'logg'], inplace=True)
         teff_points = np.sort(df.teff.unique()).astype(float)
         logg_points = np.sort(df.logg.unique()).astype(float)

### 2023-11-05 src/blase/optimizer.py

@@ -9,9 +9,9 @@ extract = lambda x: x.detach().cpu().numpy()
 
 
 def default_clean(spec: PHOENIXSpectrum):
-    norm_spec = spec.divide_by_blackbody().normalize(99)
-    continuum = norm_spec.fit_continuum(polyorder=5)
-    return norm_spec.divide(continuum, handle_meta="ff")
+    norm_spec = spec.divide_by_blackbody().normalize(99) # B in paper
+    continuum = norm_spec.fit_continuum(polyorder=5) # P in paper
+    return norm_spec.divide(continuum, handle_meta="ff") # S in paper
 
 
 def run_emulator(

### 2023-11-07 paper/paper2/ms.tex
Paper symbol table

--- 2023-10-10 paper/paper2/ms.tex
@@
 \usepackage{blindtext}
+\usepackage{bm}
 % Temporary packages
 \usepackage{mdframed}
+\usepackage{verbatim}
+
+\def\Teff{T_{\rm eff}}
+\def\vsini{v\sin{i}}
+\def\kmps{\mathrm{km}\;\mathrm{s}^{-1}}
 
@@
 \appendix
-\section{Linear Regression}
-\blindtext
+\section{Notation}
+We adopt similar notation to the blas\'e paper, with some small modifications.
+
+\begin{deluxetable}{cp{10cm}}[h]
+    \tabletypesize{\small}
+    \tablecaption{Notation used in this paper\label{table2}}
+    \tablehead{
+        \colhead{Symbol} & \colhead{Meaning}
+    }
+    \startdata
+    \hline
+    \multicolumn{2}{c}{Indexes}\\
+    \hline
+    $i$ & Pixel index of the precomputed spectrum\\
+    $j$ & Spectral line index\\
+    $k$ & Grid point index\\
+    \hline
+    \multicolumn{2}{c}{Spectra}\\
+    \hline
+    $\bm{\lambda}$ & Spectral axis of the model grid\\
+    $\mathbb{G}$ & The grid of synthetic spectral models\\
+    $\mathsf{B}_k$ & Blackbody spectrum of the $k^{th}$ model\\
+    $\mathsf{P}_k$ & Continuum polynomial of the $k^{th}$ model\\
+    $\mathsf{S}_k$ & Preprocessed flux vector of the $k^{th}$ model\\
+    $\mathsf{\hat{S}}$ & Flux vector of the continuous model reconstruction\\
+    $\mathsf{R}_k$ & Residual between a $\mathsf{S}_k$ and its reconstruction\\
+    \hline
+    \multicolumn{2}{c}{Spectral Line Properties}\\
+    \hline
+    $\mu_{jk}$ & Center position of the $j^{th}$ line in the $k^{th}$ model\\
+    $A_{jk}$ & Amplitude of the $j^{th}$ line in the $k^{th}$ model\\
+    $\sigma_{jk}$ & Gaussian shape parameter of the $j^{th}$ line in the $k^{th}$ model\\
+    $\gamma_{jk}$ & Lorentzian shape parameter of the $j^{th}$ line in the $k^{th}$ model\\
+    \hline
+    \multicolumn{2}{c}{Scalars}\\
+    \hline
+    $I$ & Number of pixels in the model grid's spectral axis\\
+    $J$ & Number of spectral lines detected across the grid \\
+    $K$ & Number of grid points in the model grid\\
+    $p$ & The prominence threshold of spectral lines to identify\\
+    $v\sin(i)$ & Observed stellar rotation for inclination $i$ and true rotation $v$\\
+    $v_r$ & Radial velocity of the star\\
+    $R$ & Spectrograph resolving power $\lambda/\delta\lambda$\\
+    $Q$ & Quality of a spectral reconstruction\\
+    $H$ & The average storage impact of a spectral reconstruction\\
+    $T$ & The average computational impact of a spectral reconstruction\\
+    \enddata
+\end{deluxetable}
 

### 2023-11-07 paper/paper2/ms.tex
Pair writing paper

@@
 \usepackage{bm}
+\usepackage{enumitem}
 % Temporary packages
@@
     - Conceivable Solution 2: increase model complexity
-    \textcolor{lightgray}{\blindtext}
+    \\--------\\
+    When using blas\'e, we aggregate the detected lines from all of the grid points
+    into one superset of lines. However, not all of the lines in this superset are 
+    detected at every grid point. In order to deal with this sparsity, we ignore lines
+    have $xx$ detections or less. This removes the most sparse lines, but some sparsity still
+    remains. The remaining 'missing lines' are dealt with by treating them as Voigt profiles 
+    with amplitudes of 0, and Gaussian and Lorentzian widths equal to $xx$ and $xx$ respectively.
 \end{mdframed}
@@
     \textbf{Conceivable Extensions}
-
-    \textcolor{lightgray}{\blindtext}
+    \\-------------\\
+    Future studies can extend this work in several ways:
+    \begin{enumerate}[label=-]
+        \item Extending the data from just the PHOENIX Grid to additional synthethic spectral
+        model grids, such as the Sonora substellar model grids or the COOLTLUSTY substellar model grid.
+    \end{enumerate}
 \end{mdframed}
@@
     $\bm{\lambda}$ & Spectral axis of the model grid\\
-    $\mathbb{G}$ & The grid of synthetic spectral models\\
+    $\bm{G}$ & The grid of synthetic spectral models\\
     $\mathsf{B}_k$ & Blackbody spectrum of the $k^{th}$ model\\
@@
     $K$ & Number of grid points in the model grid\\
-    $p$ & The prominence threshold of spectral lines to identify\\
-    $v\sin(i)$ & Observed stellar rotation for inclination $i$ and true rotation $v$\\
+    $p$ & The prominence threshold for spectral line identification\\
+    $w$ & Wing cut pixel threshold for spectral line identification\\
     $v_r$ & Radial velocity of the star\\

### 2023-11-07 paper/paper2/ms.tex
Pair writing with sujay

@@
 \section{Introduction}
-\blindtext
+
+\subsection{General spectroscopy ideas}
+
+\textbf{what do we want}\\
+Key goal: semi-empirical models-- a model that can continuously learn from and adapt to data.
+Each new observed data spectrum informs changes to our preconceived theories of how stellar atmospheres produce spectra.
+We seek model flexibility where we expect there to be deficiencies in these theories, and rigidity where we expect our theories to be most robust.
+
+\textbf{Why is that hard}\\
+
+Comparison to \citet{czekala15}.  Describe limitation of pre-designating 1 grid point.
+
+blasePaper1 describes a way to overcome this limitation by building up a so-called ``heatmap'' of line properties.  The realization of such a function enables a key capability: semi-empirical models.
+This paper implements ths
 

### 2023-11-07 paper/paper2/ms.tex
Pair writing: again this time

@@
 \begin{mdframed}
-    \textbf{Emulation with blas\'e}
+    \textbf{Cloning with blas\'e}
 
@@
     - Output storage
-
-    \textcolor{lightgray}{\blindtext}
+    \\---------------\\
+    Blas\'e's algorithm is designed to take a spectrum as input, and outputs two products:
+    1. A list of detected spectral line properties, all assumed to be Voigt profiles, and 2.
+    A cloned spectrum, which is the aggregation of all of those detected lines into a spectrum. We
+    don't use the cloned spectrum from this process, and only focus on the line properties that are output.
+    These two products are stored as a .pt file, which can then be loaded in in order to 
+    analyze all the lines in bulk. Each grid point is processed in this way, and the resulting .pt files'
+    line properties are aggregated into a single DataFrame, which can then be queried by $T_eff$, $\log(g)$, $Z$,
+    in addition to the actual line in question, identified by its center $\mu_j$.
 \end{mdframed}
@@
     \textbf{Open Source Availability}
-
-    \textcolor{lightgray}{\blindtext}
+    \\--------------\\
+    The PHOENIX model grid is available open-source at $xx$, and can be interfaced with gollum,
+    also available open-source at $xx$. Blas\'e is also available open-source at $xx$, with this work
+    existing as a fork named blas\'e3D, available at $xx$.
 \end{mdframed}

### 2023-11-09 paper/paper2/ms.tex
Paper deluxetable tweaks

@@
 \section{Notation}
-We adopt similar notation to the blas\'e paper, with some small modifications.
+We adopt similar notation to the original blas\'e paper, with some small modifications. We make indexes and
+their limits consistent, and remove the use of subscript labeling. We also add new symbols for
+wing cut pixels, and radial velocity, as well as introducing three new metrics for reconstruction quality,
+storage impact, and computational impact.
 
@@
     \tablehead{
-        \colhead{Symbol} & \colhead{Meaning}
+        \colhead{Symbol} & \colhead{Meaning [former symbol, if applicable]}
     }
@@
     \hline
-    $\bm{\lambda}$ & Spectral axis of the model grid\\
+    $\bm{\lambda}$ & Spectral axis of the model grid [$\bm{\lambda}_S$]\\
     $\bm{G}$ & The grid of synthetic spectral models\\
@@
     \hline
-    $\mu_{jk}$ & Center position of the $j^{th}$ line in the $k^{th}$ model\\
-    $A_{jk}$ & Amplitude of the $j^{th}$ line in the $k^{th}$ model\\
+    $\mu_{jk}$ & Center position of the $j^{th}$ line in the $k^{th}$ model [$\lambda_c$]\\
+    $A_{jk}$ & Amplitude of the $j^{th}$ line in the $k^{th}$ model [$a$]\\
     $\sigma_{jk}$ & Gaussian shape parameter of the $j^{th}$ line in the $k^{th}$ model\\
@@
     \hline
-    $I$ & Number of pixels in the model grid's spectral axis\\
-    $J$ & Number of spectral lines detected across the grid \\
+    $I$ & Number of pixels in the model grid's spectral axis [$N_x$]\\
+    $J$ & Number of spectral lines detected across the grid [$N_{lines}$]\\
     $K$ & Number of grid points in the model grid\\
-    $p$ & The prominence threshold for spectral line identification\\
     $w$ & Wing cut pixel threshold for spectral line identification\\
-    $v_r$ & Radial velocity of the star\\
+    $v_r$ & Radial velocity of the star [$RV$]\\
     $R$ & Spectrograph resolving power $\lambda/\delta\lambda$\\
@@
     $H$ & The average storage impact of a spectral reconstruction\\
-    $T$ & The average computational impact of a spectral reconstruction\\
+    $Y$ & The average computational impact of a spectral reconstruction\\
     \enddata

### 2023-11-09 paper/paper2/ms.tex
Table things

@@
 \usepackage{enumitem}
+\usepackage{graphicx}
 % Temporary packages
@@
 
+Here is a table of the performance of different manifold choices:
+
+\begin{deluxetable*}{cccc}[h]
+    \tabletypesize{\small}
+    \tablecaption{Notation used in this paper\label{table2}}
+    \tablehead{
+        \colhead{Model Architecture} & \colhead{Reconstruction RMS} & \colhead{Model Size} & \colhead{Training Time}
+        \\\colhead{} & \colhead{RMS \%} & \colhead{MB} & \colhead{x} 
+        }
+    \startdata
+    Linear Polynomial & a & b & c\\
+    FFT & a & b & c\\
+    Random Forest & a & b & c\\
+    Gaussian Process & a & b & c\\
+    Interpolation & a & b & c\\
+    Interpolation & a & b & c\\ 
+    \enddata
+\end{deluxetable*}
+
 \begin{mdframed}

### 2023-11-14 paper/paper2/ms.tex
Add intro text to the paper

@@
 \subsection{General spectroscopy ideas}
+
+Precomputed synthetic stellar spectral models represent a distillation of all available physics and chemistry theory into effortlessly-reproducible comparisons for spectral observations.
+Practitioners may simply download a file and overplot it against an observed spectrum, and apply human pattern recognition to obtain insights on fundamental stellar properties.
+But such exercises typically fall short.
+Inevitably, individual line depth and shape predictions do not match the observed spectrum.
+\citet{2022ApJ...941..200G} introduced a two-stage interpretable Machine Learning (ML) technique to alleviate some of these challenges.
+First clone the lines in a precomuted spectrum with fine-tuned Voigt profiles, and then warp those line properties in response to data using the power of automatic differentiation.
+This design satisfied a common use case, the desire to obtain a semi-empirical super-resolution stellar template for the observed star: the pristine $R=\frac{\lambda}{\delta \lambda}\sim1,000,000$ emergent spectrum coming from a patch of the star, having accounted for rotational broadening, instrumental resolution, and telluric artifacts.
+We called this algorithm \emph{blas\'e}.
+
+This capability alone was useful for Extreme Precision Radial Velocity (EPRV), but a key limitation of \emph{bla\'e} was its requirement to pre-select a grid point as the initial starting point.
 

### 2023-11-15 experiments/10_end_to_end/10_end_to_end.py
Beginning of production run code

@@ -1,0 +1,57 @@
+import numpy as np
+import pandas as pd
+import torch
+
+from collections import defaultdict
+from functools import partial
+from os import listdir
+from re import split
+from scipy.interpolate import griddata
+from tqdm import tqdm
+
+
+def read_state_dicts(path: str) -> pd.DataFrame:
+    line_stats = defaultdict(list)
+    for f in tqdm(listdir(path)):
+        state_dict = state_dict = torch.load(f'{path}/{f}', map_location='cuda:0')
+        tokens = split('[TGZ]', f[:-3])
+        line_stats['teff'].append(int(tokens[1]))
+        line_stats['logg'].append(float(tokens[2]))
+        line_stats['Z'].append(float(tokens[3]))
+        line_stats['center'].append(state_dict['pre_line_centers'].cpu().numpy())
+        line_stats['shift_center'].append(state_dict['lam_centers'].cpu().numpy())
+        line_stats['amp'].append(state_dict['amplitudes'].cpu().numpy())
+        line_stats['sigma'].append(state_dict['sigma_widths'].cpu().numpy())
+        line_stats['gamma'].append(state_dict['gamma_widths'].cpu().numpy())
+    return pd.DataFrame(line_stats)
+
+def optimize_memory(df: pd.DataFrame):
+    fcols = df.select_dtypes('float').columns
+    icols = df.select_dtypes('integer').columns
+    df[fcols] = df[fcols].apply(pd.to_numeric, downcast='float')
+    df[icols] = df[icols].apply(pd.to_numeric, downcast='integer')
+
+def local_run():
+    path = '/home/sujay/data/10K_12.5K_clones'
+    df = read_state_dicts(path).query('Z == 0').explode(['center', 'amp', 'sigma', 'gamma', 'shift_center']).convert_dtypes(dtype_backend='numpy_nullable')
+    print('DataFrame created')
+    optimize_memory(df)
+    print('DataFrame memory optimized')
+    interpolator_list = []
+    for line in tqdm(df.center.unique()):
+        df_line = df.query('center == @line')
+        interpolator_list.append(partial(griddata, points=(df_line.teff, df_line.logg), values=df_line[['amp', 'sigma', 'gamma']].to_numpy(), method='linear'))
+    print('Interpolator partials created')
+    for interpolator in interpolator_list:
+        print(interpolator(xi=(5777, 4.44))) 
+
+def triton_run():
[…]

### 2023-11-28 paper/paper2/ms.tex
Big paper work

--- 2023-11-09 paper/paper2/ms.tex
@@
 \usepackage{enumitem}
+\usepackage{graphicx}
+\usepackage{hyperref}
 % Temporary packages
@@
 \def\Teff{T_{\rm eff}}
-\def\vsini{v\sin{i}}
+\def\logg{\log(g)}
+\def\Z{\mathrm{[Fe/H]}}
+\def\vsini{v\sin(i)}
 \def\kmps{\mathrm{km}\;\mathrm{s}^{-1}}
@@
 \begin{document}
-\title{Blas\'e3D}
-\shorttitle{\emph{blas\'e3D}}
+\title{Estimating Fundamental Stellar Properties using PHOENIX Grid Interpolators}
 
@@
 \subsection{General spectroscopy ideas}
+
+Precomputed synthetic stellar spectral models represent a distillation of all available physics
+and chemistry theory into effortlessly-reproducible comparisons for spectral observations.  
+Practitioners may simply download a file and overplot it against an observed spectrum, relying on 
+human pattern recognition to obtain insights on fundamental stellar properties. But such exercises 
+typically fall short. Inevitably, individual line depth and shape predictions do not match the 
+observed spectrum. \citet{2022ApJ...941..200G} introduced a two-stage interpretable machine learning (ML) 
+technique to alleviate some of these challenges. First, cloning the lines in a precomputed spectrum 
+with fine-tuned Voigt profiles, and then warping their line properties in response to data using 
+the power of automatic differentiation. This design satisfied a common use case, the desire to 
+obtain a semi-empirical super-resolution stellar template for the observed star: the pristine 
+$R=\frac{\lambda}{\delta \lambda}\sim1,000,000$ emergent spectrum coming from a patch of the star, 
+having accounted for rotational broadening, instrumental resolution, and telluric artifacts. 
+We called this algorithm \emph{blas\'e}, available with the Python package \texttt{blase}.
+
+This capability alone was useful for Extreme Precision Radial Velocity (EPRV), but a key limitation 
+of \emph{blas\'e} was its requirement to pre-select a grid point as the initial starting point. The code
+introduced in this study bypasses this requirement by comparing spectral reconstructions to an input
+spectrum in order to automatically ascertain the fundamental stellar parameters.
 
@@
 
-\begin{mdframed}
-    \textbf{The PHOENIX grid subset}
+For the purposes of this study, we use a slice of the PHOENIX grid where the alpha abundance
+is set to zero. Within this slice, we allow the full range of $\Teff, \logg, \Z$. The specific
+sampling of the grid points is shown in Figure $XX$. Our approach could feasibly be extended to
+include alpha abundances, but given that the \texttt{gollum} interface does not yet support the access
+of spectra with nonzero alpha abundances, as well as bearing a significant increase in computational 
+impact, we leave this for future work.
 
-    \textcolor{lightgray}{\blindtext}
-\end{mdframed}
+The \texttt{blase} code takes in any spectrum as input, and outputs a cloned
+version of that spectrum. However, there is a hidden product to this
+procedure, the state dictionary of the model itself, stored as a \texttt{.pt} file. In this lies metadata about
+where spectral lines were identified by \texttt{blase}, as well as those lines'
+specific properties, namely amplitude, Gaussian shape, and Lorentzian shape.
+In addition, \texttt{blase} also allows the line centers to shift slightly during
+cloning, meaning that in addition to the initial line centers, we 
+also have access to the shifted line centers. This hidden metadata is 
+what formed the backbone for this study and allowed us to retrieve line-by-line data
+that could then be interpolated. We input the PHOENIX grid spectra into \texttt{blase}
+and accessed the state dictionaries of each clone. Once this data was retrieved from the state dictionaries
+across the grid, it was then aggregated into a single \texttt{pandas} DataFrame, enabling
+querying of the data by any property, be it a fundamental stellar parameter or the output
+line properties.
 
-\begin{mdframed}
-    \textbf{Cloning with blas\'e}
-
-    - Per-grid point optimization procedure
-    - Pretrained model caching
-    - Output storage
-    \\---------------\\
-    Blas\'e's algorithm is designed to take a spectrum as input, and outputs two products:
-    1. A list of detected spectral line properties, all assumed to be Voigt profiles, and 2.
-    A cloned spectrum, which is the aggregation of all of those detected lines into a spectrum. We
-    don't use the cloned spectrum from this process, and only focus on the line properties that are output.
-    These two products are stored as a .pt file, which can then be loaded in in order to 
-    analyze all the lines in bulk. Each grid point is processed in this way, and the resulting .pt files'
-    line properties are aggregated into a single DataFrame, which can then be queried by $T_eff$, $\log(g)$, $Z$,
-    in addition to the actual line in question, identified by its center $\mu_j$.
-\end{mdframed}
-
-\begin{mdframed}
-    \textbf{Open Source Availability}
-    \\--------------\\
-    The PHOENIX model grid is available open-source at $xx$, and can be interfaced with gollum,
-    also available open-source at $xx$. Blas\'e is also available open-source at $xx$, with this work
-    existing as a fork named blas\'e3D, available at $xx$.
-\end{mdframed}
+All data and code used in this study is available online and is open source. The PHOENIX grid itself takes quite a while to download
+due to server bandwidth limitations; this is something that \texttt{gollum} aims to
+improve upon by implementing both caching behavior and underlying data structure compression.
+\texttt{gollum} itself is available on GitHub and is pip installable, although this
+method lags in features significantly compared to the bleeding edge code from GitHub.
+\texttt{blase} is the main library of note here, also on GitHub, while the
+materials related to this study are currently residing as a fork named \texttt{blase3D}. When polished and complete,
+the fork is set to merge into \texttt{blase} itself. Links to all of these are provided in the appendix.
 
@@
 
+We have established that the \texttt{blase} code is capable of cloning a spectrum and
+returning a set of line properties for each line. In order to visualize line-by-line
+properties across the grid, we need to be able to identify unique lines. This begs the question:
+what property lets us uniquely identify a spectral line? We have two possible solutions.
+First, we consider the shifted line centers. However, this shifting comes from \texttt{blase}'s
+emulating physics from its training dataset, meaning that the same line at two grid points could
+be shifted differently. Could we then round or truncate the wavelength coordinate of the line centers?
+The answer is no, because these shifts vary in precision, and some lines are close enough to begin with
+that what should have been two different lines get identified as the same line. This, obviously is incorrect;
+one spectral line is just that, one spectral line. So we now consider the original unshifted line centers.
+These are the same across the grid, and there is no issue with unintentionally binning lines together
+thanks to a uniform precision. This means that we now use the unshifted line centers as the unique
+identifier for lines, which vary across the grid in their properties, namely amplitude, Gaussian shape,
+and Lorentzian shape.
+
+The next thing that we want to do is to visualize the trends of line properties over
+the PHOENIX grid. The best way to visualize this is with heatmaps. This means that for
+each line, for each line property, we have one heatmap, where the axes are the fundamental
+parameters, and the `temperature' is the value of the line property. As you can imagine, this
+is a lot of heatmaps, each of which is incredibly hard to visualize due to the dimensionality.
+So, we slice the heatmaps at solar metallicity to reduce them to 2D, and then we can easily 
+visualize the trend of any one line property of any one line across the grid on demand.
+
 \begin{mdframed}
-    \textbf{Line Recognition}
-    - Need: Identifying unique lines
-    - Anticipated friction points: line centroids drift
-    - Strategy: (Pre-shift centers)
-    \textcolor{lightgray}{\blindtext}
+    \textbf{Figure: Heatmap $\Teff$ vs $\logg$ for a single line $2\times2$ panels for $\sigma$, $\gamma$, $A$, $\lambda$ }
 \end{mdframed}
 
-\begin{mdframed}
-    \textbf{Line property bulk trends across the grid}
-    \textcolor{lightgray}{\blindtext}
-\end{mdframed}
-
-\begin{mdframed}
-    \textbf{Figure: Heatmap $T_\mathrm{eff}$ vs $\log{g}$ for a single line $2\times2$ panels for $\sigma$, $\gamma$, $A$, $\lambda$ }
-\end{mdframed}
-
-\begin{mdframed}
-    \textbf{Anomalies in Heatmaps}
-    \textcolor{lightgray}{\blindtext}
-\end{mdframed}
+While we see heatmaps for lines that have values at all grid points, this is only true for
+a select few spectral lines. The vast majority of lines do not seem to show up at certain
+grid points, meaning that our heatmaps are sparse. Initially this seems like it would not be an 
+issue; when reconstructing, simply retain knowledge of which points were missing and override
+the interpolated value with a missing value. Now, in addition to an
+interpolator, a sequence of grid points with missing values must be stored for each line just to
+override the interpolator. This increases the storage cost of the model. In addition to this, we would
+not be able to deal with points lying on the boundary between missing and visible lines. So, we need
+to process these missing lines. Since the line properties are on a logarithmic scale,
+a line with an amplitude of 0 would have a log amplitude of $-\infty$. This means that we should just
+set all missing lines to have log amplitudes of $-\infty$, right? Close, however interpolation
+algorithms absolutely despise infinities. So, we set the log amplitude of missing lines to be
+$-1000$, which is a large enough negative number that it will effectively evaluate to zero without
+causing problems with the interpolation process. We do not really have to worry about the Gaussian
+and Lorentzian shape parameters; no matter how `wrong' their interpolated values are, an amplitude of
+zero will override that and make the line disappear. This way, we have effectively dealt with
+pathological behavior in such a way that interpolation returns a sensible value without sending
+the storage impact through the roof.
 
@@
 
-\begin{mdframed}
-    \textbf{Conceivable Extensions}
-    \\-------------\\
-    Future studies can extend this work in several ways:
+Future studies can expand upon this work in several ways:
 \begin{enumerate}[label=-]
-        \item Extending the data from just the PHOENIX Grid to additional synthethic spectral
+    \item Extending the data from just the PHOENIX grid to additional synthetic spectral
           model grids, such as the Sonora substellar model grids or the COOLTLUSTY substellar model grid.
+    \item Expanding the section of the PHOENIX grid on which \texttt{blase} was run to include all available alpha abundances, encompassing all of PHOENIX.
+    \item Using more advanced regressors with ML techniques to predict line properties instead of using interpolation to brute-force the model predictions.
+    \item Extending support to accommodate large-bandwidth spectra across the full electromagnetic spectrum.
+    \item Identifying spectral lines using species-based line lists rather than fitting Voigt profiles whenever possible.
 \end{enumerate}
-\end{mdframed}
 

### 2023-11-28 experiments/10_end_to_end/10_end_to_end.py

@@ -4,6 +4,7 @@ import torch
 
 from collections import defaultdict
 from functools import partial
+from multiprocessing import Pool
 from os import listdir
 from re import split
 from scipy.interpolate import griddata
@@ -31,19 +32,27 @@ def optimize_memory(df: pd.DataFrame):
     df[fcols] = df[fcols].apply(pd.to_numeric, downcast='float')
     df[icols] = df[icols].apply(pd.to_numeric, downcast='integer')
 
+def interpolate(line: float, df: pd.DataFrame) -> partial:
+    df_line = df.query('center == @line')
+    return partial(griddata, points=(df_line.teff, df_line.logg), values=df_line[['amp', 'sigma', 'gamma']].to_numpy(), method='linear')
+
 def local_run():
     path = '/home/sujay/data/10K_12.5K_clones'
-    df = read_state_dicts(path).query('Z == 0').explode(['center', 'amp', 'sigma', 'gamma', 'shift_center']).convert_dtypes(dtype_backend='numpy_nullable')
+    df_native = read_state_dicts(path).query('Z == 0 & 4000 <= teff <= 7000 & 2 <= logg <= 4')
+    df = df_native.explode(['center', 'amp', 'sigma', 'gamma', 'shift_center']).convert_dtypes(dtype_backend='numpy_nullable')
     print('DataFrame created')
     optimize_memory(df)
     print('DataFrame memory optimized')
     interpolator_list = []
-    for line in tqdm(df.center.unique()):
-        df_line = df.query('center == @line')
-        interpolator_list.append(partial(griddata, points=(df_line.teff, df_line.logg), values=df_line[['amp', 'sigma', 'gamma']].to_numpy(), method='linear'))
-    print('Interpolator partials created')
-    for interpolator in interpolator_list:
-        print(interpolator(xi=(5777, 4.44))) 
+    #for line in tqdm(df.center.unique()):
+    #    df_line = df.query('center == @line')
+    #    interpolator_list.append(partial(griddata, points=(df_line.teff, df_line.logg), values=df_line[['amp', 'sigma', 'gamma']].to_numpy(), method='linear'))
+    df_gp = df_native[['teff', 'logg', 'Z']]
+    df_line = df.query('center == @df.center.unique()[-1]')
+    print(df_line.merge(df_gp, how='right', on=['teff', 'logg', 'Z']).fillna(-1000))
+    #print('Interpolator partials created')
[…]

### 2023-11-30 paper/paper2/ms.tex
Line density figure addition
First draft of the flowcharts, split into steps 1-3 and 4-6
Generate plots

@@
 \begin{document}
-\title{Estimating Fundamental Stellar Properties using PHOENIX Grid Interpolators}
+\title{Precision Fundamental Stellar Properties with Interpretable Machine Learning}
 
@@
 
+\begin{figure*}
+    \includegraphics[width=\textwidth]{images/flowchart_frame1_v0p3.png}
+    \caption{Flowchart of the cloning steps for emulating precomputed spectra.}
+\end{figure*}
+
 \section{Emulating the PHOENIX grid}
@@
 
-\begin{mdframed}
-    \textbf{Figure: Line density across the grid heatmap}
-\end{mdframed}
+\begin{figure*}
+    \includegraphics[width=\textwidth]{images/line_density.png}
+    \caption{Heatmap of the number of spectral lines detected at each grid point.
+    The grid has been sliced at solar metallicity to make the heatmap 2D.}
+\end{figure*}
 
@@
 
-\begin{mdframed}
-    \textbf{Figure: Heatmap $\Teff$ vs $\logg$ for a single line $2\times2$ panels for $\sigma$, $\gamma$, $A$, $\lambda$ }
-\end{mdframed}
+\begin{figure*}
+    \includegraphics[width=\textwidth]{images/heatmap0.png}
+    \caption{Heatmap of the most commonly detected spectral line at 8185.51 Angstroms. Amplitude,
+    Gaussian shape, and Lorentzian shape are on a logarithmic scale. Jitter is a measure of how 
+    much \texttt{blase} shifted the line center during cloning.}
+\end{figure*}
 
@@
     \textbf{Statement: A Bidirectonal Relation exists}
-    - Goal: identify a functional form
-    \textcolor{lightgray}{\blindtext}
+
+    Now that we have our list of line properties for each grid point, we need to find the link 
+    between these line properties and the fundamental stellar parameters. We know that there is
+    some relation between the two, but we do not even know its functional form. While we have a
+    ridiculously large array of methods to choose from, we can start with a simple linear
+    regression with a polynomial basis. We can quickly test the efficacy of this method by 
+    running it on the line we visualized earlier.
 \end{mdframed}
 
+\begin{figure*}
+    \includegraphics[width=\textwidth]{images/heatmap_poly.png}
+\end{figure*}
+
 \begin{mdframed}
-    \textbf{Figure: Flowchart-with-equations}
+    \textbf{Big equation section}
 \end{mdframed}
 
-\begin{mdframed}
-    \textbf{Functional Form}
-    - Describe functional form options and refinement procedures
-    - List possibilities, Linreg, GPs \citep{2023ARA&A..61..329A}, NNs, etc.
-    - We choose LSTSQ
-    - Enumerate functional form
-    - Adapting model complexity-- AIC
-    \textcolor{lightgray}{\blindtext}
-\end{mdframed}
+\begin{figure*}
+    \includegraphics[width=\textwidth]{images/flowchart_frame2_v0p3.png}
+    \caption{Flowchart of the manifold steps for reconstructing a precomputed spectrum.}
+\end{figure*}
 
@@
 
+    \textcolor{lightgray}{\blindtext}
+\end{mdframed}
+
+\begin{mdframed}
+    \textbf{Functional Form}
+    - Describe functional form options and refinement procedures
+    - List possibilities, Linreg, GPs \citep{2023ARA&A..61..329A}, NNs, etc.
+    - We choose LSTSQ
+    - Enumerate functional form
+    - Adapting model complexity-- AIC
     \textcolor{lightgray}{\blindtext}

### 2023-12-05 paper/paper2/ms.tex
Paper push

@@
 \usepackage{hyperref}
+\usepackage{physics}
 % Temporary packages
@@
 For the purposes of this study, we use a slice of the PHOENIX grid where the alpha abundance
-is set to zero. Within this slice, we allow the full range of $\Teff, \logg, \Z$. The specific
-sampling of the grid points is shown in Figure $XX$. Our approach could feasibly be extended to
+is set to zero. Within this slice, we allow the full range of $\Teff, \logg, \Z$, leaving us with 1,959
+grid points, each with its own PHOENIX spectrum. The specific domain and
+sampling of the grid points are shown in Figure $XX$. Our approach could feasibly be extended to
 include alpha abundances, but given that the \texttt{gollum} interface does not yet support the access
@@
 version of that spectrum. However, there is a hidden product to this
-procedure, the state dictionary of the model itself, stored as a \texttt{.pt} file. In this lies metadata about
+procedure, the state dictionary of the model itself, which can be written to a file. In this lies metadata about
 where spectral lines were identified by \texttt{blase}, as well as those lines'
@@
 cloning, meaning that in addition to the initial line centers, we
-also have access to the shifted line centers. This hidden metadata is
-what formed the backbone for this study and allowed us to retrieve line-by-line data
-that could then be interpolated. We input the PHOENIX grid spectra into \texttt{blase}
-and accessed the state dictionaries of each clone. Once this data was retrieved from the state dictionaries
+also have access to the shifted line centers. This hidden metadata is what
+formed the backbone for this study and allowed us to retrieve line-by-line data
+that could then be interpolated.
+
+The ability to store these models on disk in itself is quite novel, as it offers a reduction
+in storage space while also being interpretable. What we effectively do is store the
+generator for the data rather than the data itself, which is a new paradigm in the 
+storage of model grids such as PHOENIX.
+Using this method, we input the PHOENIX grid spectra into \texttt{blase},
+which wrote the models to disk, and then accessed the state dictionaries of each clone. Once this data was retrieved from the state dictionaries
 across the grid, it was then aggregated into a single \texttt{pandas} DataFrame, enabling
@@
 
-All data and code used in this study is available online and is open source. The PHOENIX grid itself takes quite a while to download
+Before we proceed further, it is important to mention that all data and code used in this study is available online and is open source. The PHOENIX grid itself takes quite a while to download
 due to server bandwidth limitations; this is something that \texttt{gollum} aims to
@@
 \texttt{blase} is the main library of note here, also on GitHub, while the
-materials related to this study are currently residing as a fork named \texttt{blase3D}. When polished and complete,
+materials related to this study are currently residing in a fork named \texttt{blase3D}. When polished and complete,
 the fork is set to merge into \texttt{blase} itself. Links to all of these are provided in the appendix.
@@
 \section{Mapping Line Parameters to Fundamental Properties}
-\begin{mdframed}
-    \textbf{Statement: A Bidirectonal Relation exists}
 
 Now that we have our list of line properties for each grid point, we need to find the link 
-    between these line properties and the fundamental stellar parameters. We know that there is
-    some relation between the two, but we do not even know its functional form. While we have a
-    ridiculously large array of methods to choose from, we can start with a simple linear
-    regression with a polynomial basis. We can quickly test the efficacy of this method by 
-    running it on the line we visualized earlier.
-\end{mdframed}
+between these line properties and the fundamental stellar parameters, and we want to obtain
+some relation between the two. For this study, we decided to use linear interpolation to
+supersample the mapping between fundamental stellar parameters and spectral line properties.
 
@@
 
-\begin{mdframed}
-    \textbf{Big equation section}
-\end{mdframed}
+While we have discussed the process so far fairly qualitatively, we now establish the
+mathematical notation for the methods that we employ in this study. Our input
+fundamental line properties can be written as a vector with entries as follows:
+
+\begin{gather}
+    \theta: \mqty[\Teff, \logg, \Z]
+\end{gather}
+
+The PHOENIX grid $\bm{G}$ has $K$ grid points, indexed as $k$, and gollum
+evaluates the following mapping:
+\begin{gather}
+    \theta_k \mapsto \bm{G}_k
+\end{gather}
+
+where $\bm{G}_k$ is the $k^{th}$ PHOENIX spectrum. All PHOENIX spectra share the same
+spectral axis but differ in their flux axis.
 
@@
 \end{figure*}
-
-\begin{mdframed}
-    \textbf{Problem: missing lines}
-    - Conceivable Solution 1: treat as NaNs and deal with sparsity
-    - Conceivable Solution 2: increase model complexity
-    \\--------\\
-    When using blas\'e, we aggregate the detected lines from all of the grid points
-    into one superset of lines. However, not all of the lines in this superset are
-    detected at every grid point. In order to deal with this sparsity, we ignore lines
-    have $xx$ detections or less. This removes the most sparse lines, but some sparsity still
-    remains. The remaining 'missing lines' are dealt with by treating them as Voigt profiles
-    with amplitudes of 0, and Gaussian and Lorentzian widths equal to $xx$ and $xx$ respectively.
-\end{mdframed}
 
@@
 \section{Discussion}
-\begin{mdframed}
-    \textbf{Revisiting Model Assumptions}
 
-    \textcolor{lightgray}{\blindtext}
-\end{mdframed}
-
-\begin{mdframed}
-    \textbf{Functional Form}
-    - Describe functional form options and refinement procedures
-    - List possibilities, Linreg, GPs \citep{2023ARA&A..61..329A}, NNs, etc.
-    - We choose LSTSQ
-    - Enumerate functional form
-    - Adapting model complexity-- AIC
-    \textcolor{lightgray}{\blindtext}
-\end{mdframed}
+While we used interpolation to model the mapping between fundamental stellar parameters
+and spectral line properties, there are also other methods that could feasibly used. Very early in the study, we considered a rather simple
+solution such as polynomial linear regression, but this was quickly discarded as the regressors
+had difficulty retaining detail in the dataset and scaled poorly, with the number
+of terms growing as the cube of the desired degree. The scaling issue would quickly lead to the regression's design matrix
+having more columns than it did rows, which would cause regression to fail due to having more degrees
+of freedom than data points. Interpolation was our better solution, but others exist that we 
+intend to explore, such as random forest regression, Gaussian processes, and neural networks. These methods offer an increase in the generalization of
+PHOENIX's spectral features by not simply memorizing the data as an interpolator would,
+with the drawback that they can be trickier to implement correctly. We also make
+the assumption that every single spectral line follows a Voigt profile, which gets us 
+most of the way there but does admittedly fall flat in some cases.
 

### 2024-03-06 paper/paper2/ms.tex
Complete interpolation code and test dropout residual

@@
 \begin{document}
-\title{Precision Fundamental Stellar Properties with Interpretable Machine Learning}
+\title{Precision Fundamental Stellar Properties with Adaptive Machine Learning}
 
@@
 observed spectrum. \citet{2022ApJ...941..200G} introduced a two-stage interpretable machine learning (ML)
-technique to alleviate some of these challenges. First, cloning the lines in a precomputed spectrum
+technique to alleviate some of these challenges: first, cloning the lines in a precomputed spectrum
 with fine-tuned Voigt profiles, and then warping their line properties in response to data using
@@
 obtain a semi-empirical super-resolution stellar template for the observed star: the pristine
-$R=\frac{\lambda}{\delta \lambda}\sim1,000,000$ emergent spectrum coming from a patch of the star,
+$R\equiv\frac{\lambda}{\delta \lambda}\sim1,000,000$ emergent spectrum coming from a patch of the star,
 having accounted for rotational broadening, instrumental resolution, and telluric artifacts.
@@
 This capability alone was useful for Extreme Precision Radial Velocity (EPRV), but a key limitation
-of \emph{blas\'e} was its requirement to pre-select a grid point as the initial starting point. The code
-introduced in this study bypasses this requirement by comparing spectral reconstructions to an input
-spectrum in order to automatically ascertain the fundamental stellar parameters.
+of \emph{blas\'e} was its requirement to pre-select a single grid point as the initial starting point. The approach
+introduced in this study bypasses this requirement by analyzing line properties of the input spectrum
+in order to automatically ascertain fundamental stellar parameters.
 
@@
 
-For the purposes of this study, we use a slice of the PHOENIX grid where the alpha abundance
-is set to zero. Within this slice, we allow the full range of $\Teff, \logg, \Z$, leaving us with 1,959
+For the purposes of this study, we use a slice of the PHOENIX grid (Husser et al) where the alpha abundance
+is set to zero. Within this slice, we choose the full range of $\Teff, \logg, \Z$, leaving us with 1,959
 grid points, each with its own PHOENIX spectrum. The specific domain and
-sampling of the grid points are shown in Figure $XX$. Our approach could feasibly be extended to
-include alpha abundances, but given that the \texttt{gollum} interface does not yet support the access
-of spectra with nonzero alpha abundances, as well as bearing a significant increase in computational
-impact, we leave this for future work.
+sampling of the grid points are shown in Figure $XX$.
 
 The \texttt{blase} code takes in any spectrum as input, and outputs a cloned
-version of that spectrum. However, there is a hidden product to this
+version of that spectrum. There is a hidden product to this
 procedure, the state dictionary of the model itself, which can be written to a file. In this lies metadata about
 where spectral lines were identified by \texttt{blase}, as well as those lines'
-specific properties, namely amplitude, Gaussian shape, and Lorentzian shape.
+specific properties, namely amplitude, Gaussian shape, and Lorentzian shape. This metadata is 
+all of the information needed to reconstruct the spectrum.
 In addition, \texttt{blase} also allows the line centers to shift slightly during
@@
 storage of model grids such as PHOENIX.
-Using this method, we input the PHOENIX grid spectra into \texttt{blase},
+Using this method, we input 1,959 PHOENIX grid spectra into \texttt{blase},
 which wrote the models to disk, and then accessed the state dictionaries of each clone. Once this data was retrieved from the state dictionaries
-across the grid, it was then aggregated into a single \texttt{pandas} DataFrame, enabling
+across the grid, it was then aggregated into a single table of shape $XX$, enabling
 querying of the data by any property, be it a fundamental stellar parameter or the output

### 2024-03-06 experiments/10_end_to_end/10_end_to_end.py

@@ -1,10 +1,12 @@
+import matplotlib.pyplot as plt
 import numpy as np
 import pandas as pd
 import torch
 
+from blase.emulator import SparseLinearEmulator as SLE
+from blase.utils import doppler_grid
 from collections import defaultdict
 from functools import partial
-from multiprocessing import Pool
 from os import listdir
 from re import split
 from scipy.interpolate import griddata
@@ -14,7 +16,7 @@ from tqdm import tqdm
 def read_state_dicts(path: str) -> pd.DataFrame:
     line_stats = defaultdict(list)
     for f in tqdm(listdir(path)):
-        state_dict = state_dict = torch.load(f'{path}/{f}', map_location='cuda:0')
+        state_dict = torch.load(f'{path}/{f}', map_location='cuda:0')
         tokens = split('[TGZ]', f[:-3])
         line_stats['teff'].append(int(tokens[1]))
         line_stats['logg'].append(float(tokens[2]))
@@ -32,27 +34,57 @@ def optimize_memory(df: pd.DataFrame):
     df[fcols] = df[fcols].apply(pd.to_numeric, downcast='float')
     df[icols] = df[icols].apply(pd.to_numeric, downcast='integer')
 
-def interpolate(line: float, df: pd.DataFrame) -> partial:
-    df_line = df.query('center == @line')
-    return partial(griddata, points=(df_line.teff, df_line.logg), values=df_line[['amp', 'sigma', 'gamma']].to_numpy(), method='linear')
-
-def local_run():
+def local_run(p_teff, p_logg, p_Z):
     path = '/home/sujay/data/10K_12.5K_clones'
-    df_native = read_state_dicts(path).query('Z == 0 & 4000 <= teff <= 7000 & 2 <= logg <= 4')
+    df_native = read_state_dicts(path).query('-0.5 <= Z <= 0.5 & 4000 <= teff <= 7000 & 2 <= logg <= 4')
+    df_gp = df_native[['teff', 'logg', 'Z']]
     df = df_native.explode(['center', 'amp', 'sigma', 'gamma', 'shift_center']).convert_dtypes(dtype_backend='numpy_nullable')
     print('DataFrame created')
     optimize_memory(df)
     print('DataFrame memory optimized')
     interpolator_list = []
-    #for line in tqdm(df.center.unique()):
-    #    df_line = df.query('center == @line')
[…]

### 2024-03-08 paper/paper2_neo/paper.tex
Paper reformulation, Finish triton code

\documentclass[twocolumn]{aastex631}
\usepackage{graphicx}
\usepackage{physics}

\begin{document}
\title{Using Hybrid Machine Learning for Fundamental Stellar Parameter Inference}
\author[0000-0002-2290-6810]{Sujay Shankar}
\author[0000-0002-4020-3457]{Michael Gully-Santiago}
\author[0000-0002-4404-0456]{Caroline V. Morley}
\affil{Department of Astronomy, The University of Texas at Austin, Austin, TX 78712, USA}

\begin{abstract}
    Text
\end{abstract}
\keywords{Text}

\section{Introduction}
Text

\section{Cloning the PHOENIX Model Grid}
\subsection{Preprocessing with \texttt{gollum}}
Text
\subsection{Optimization with \texttt{blase}}
Text
\subsection{Voigt Component Compression}
Text

\section{Interpolating Manifolds}
\subsection{Maintaining Line Persistence}
Text
\subsection{Continuously Evaluable Partials}
Text

\section{The PHOENIX Generator}
\subsection{Generator Design}
Text
\subsection{Measuring Information Loss}
Text

\section{Bayesian Optimization}
Text

\section{Inference Testing}
\subsection{Test Data}
Text
\subsection{Performance Results}
Text

\section{Discussion}
Text

\section*{Acknowledgements}
Text

\end{document}

### 2024-03-08 experiments/10_end_to_end/10_end_to_end.py

@@ -8,6 +8,7 @@ from blase.utils import doppler_grid
 from collections import defaultdict
 from functools import partial
 from os import listdir
+from pickle import dump, load
 from re import split
 from scipy.interpolate import griddata
 from tqdm import tqdm
@@ -33,6 +34,23 @@ def optimize_memory(df: pd.DataFrame):
     icols = df.select_dtypes('integer').columns
     df[fcols] = df[fcols].apply(pd.to_numeric, downcast='float')
     df[icols] = df[icols].apply(pd.to_numeric, downcast='integer')
+
+def create_interpolators(df: pd.DataFrame, df_gp: pd.DataFrame) -> list[partial]:
+    interpolator_list = []
+    for line in tqdm(df.center.unique()):
+        df_line = df.query('center == @line').merge(df_gp, how='right', on=['teff', 'logg', 'Z']).fillna(-1000)
+        interpolator_list.append(partial(griddata, points=(df_line.teff, df_line.logg, df_line.Z), values=df_line[['amp', 'sigma', 'gamma', 'shift_center']].to_numpy()))
+    return interpolator_list
+
+def reconstruct(wl_grid: np.ndarray, interpolators: list[partial], point: tuple[int, float, float]) -> np.ndarray:
+    output = np.vstack([r for interpolator in interpolators if (r := interpolator(xi=(point[0], point[1], point[2])))[0] != -1000])
+    state_dict = {
+        'amplitudes': torch.from_numpy(output[:, 0]),
+        'sigma_widths': torch.from_numpy(output[:, 1]),
+        'gamma_widths': torch.from_numpy(output[:, 2]),
+        'lam_centers': torch.from_numpy(output[:, 3]),
+    }
+    return np.nan_to_num(SLE(wl_native=wl_grid, init_state_dict=state_dict, device="cpu").forward().detach().numpy(), nan=1)
 
 def local_run(p_teff, p_logg, p_Z):
     path = '/home/sujay/data/10K_12.5K_clones'
@@ -88,10 +106,14 @@ def local_run(p_teff, p_logg, p_Z):
 
 def triton_run():
     path = '/home/sujays/github/blase/experiments/08_blase3D_HPC_test/emulator_states'
-    df = read_state_dicts(path).explode(['center', 'amp', 'sigma', 'gamma', 'shift_center']).convert_dtypes(dtype_backend='numpy_nullable')
+    df = read_state_dicts(path)
[…]

### 2024-03-26 experiments/10_end_to_end/10_end_to_end.py
Run the inference function

@@ -135,4 +135,4 @@ def inference_run():
     print(res.func_vals)
 
 if __name__ == '__main__':
-    triton_run()
+    inference_run()

### 2024-03-27 experiments/10_end_to_end/10_end_to_end.py
Debugging

@@ -135,4 +135,6 @@ def inference_run():
     print(res.func_vals)
 
 if __name__ == '__main__':
-    inference_run()
+    spec = PHOENIXSpectrum(teff=5000, logg=4, Z=0, download=True)
+    x = chisq_loss(spec.wavelength.value, spec.flux.value)((5000, 4, 0))
+    print(x)

### 2024-03-27 paper/paper2_neo/paper.tex

@@
 \usepackage{physics}
+\usepackage[super]{nth}
+\usepackage{enumitem}
 
@@
 \section{Cloning the PHOENIX Model Grid}
+\subsection{The PHOENIX Subset}
+For the purposes of this study, we did not consider the full range of the
+PHOENIX synthetic spectral model grid. Instead, we focused on a subset of 
+the grid, whose parameter ranges are given in Table 1.
+
+\begin{table}[h]
+    \hspace*{-1.1cm}\begin{tabular}{|c|c|c|}
+        \hline
+        \multicolumn{3}{|c|}{\textbf{PHOENIX Grid Subset}}\\
+        \hline\hline
+        Parameter & Symbol & Range\\
+        \hline
+        Alpha Element Abundance & $\alpha$ & [0] dex\\
+        Iron Abundance & [Fe/H] & [-0.5, 0.5] dex\\
+        Effective Temperature & $T_{\mathrm{eff}}$ & [2300, 12000] K\\
+        Surface Gravity & $\log(g)$ & [0, 5]\\
+        Wavelength & $\lambda$ & [8050, 12850] \AA\\
+        \hline
+    \end{tabular}
+    \caption{The subset of the PHOENIX grid used in this study. These limits 
+    were imposed to reduce the computational cost of the algorithms. The 
+    wavelength limits in particular roughly line up with that of the 
+    Habitable Zone Planet Finder (HPF) spectrograph.}
+\end{table}
+
 \subsection{Preprocessing with \texttt{gollum}}
-Text
-\subsection{Optimization with \texttt{blase}}
-Text
-\subsection{Voigt Component Compression}
-Text
+First, the PHOENIX subset was accessed directly from the PHOENIX website
+using \texttt{gollum}'s download option, which uses the website's FTP server
+to download spectra without using any disk space. This came at the cost of
+longer load times due to internet connection being significantly more of a
+bottleneck than disk I/O, but was acceptable as this was a one-time operation.
+
+The spectra were then put through a three-step preprocessing pipeline.
+\begin{enumerate}
+    \item Blackbody Division: Since the $T_{\mathrm{eff}}$ of each spectrum 
+    is known, the according blackbody spectrum was divided out.
+    \item Percentile Normalization: The spectra were normalized by dividing
+    them by their 99th percentile.
+    \item Continuum Normalization: The spectra were further normalized by 
+    dividing them by a \nth{5} order polynomial continuum.
+\end{enumerate}
+
+\subsection{Line Identification with \texttt{blase}}
+The next step was to convert the PHOENIX subset into an interpretable format.
+We wanted to represent the spectra as a list of spectral lines rather than a
+list of fluxes. This was done using \texttt{blase}, which detects spectral 
+lines as Voigt profiles and tunes the profiles to mimic the original 
+PHOENIX spectrum with back propagation. Four parameters were optimized: the 
+line center $\mu$, the log-amplitude $\ln(a)$, the Gaussian width $\sigma$, and the 
+Lorentzian width $\gamma$. The optimization used the Adam optimizer with a
+learning rate of 0.05 over 100 epochs. In addition, we limited two custom 
+parameters: wing cut to 6000 and prominence to 0.005. Wing cut is a parameter
+that determines the extent of the Voigt profile to evaluate, saving
+computational resources by not evaluating small numbers. Prominence sets a
+lower limit for the amplitude of detected lines, which also saves resources 
+by disregarding shallow lines. In short, larger values for wing cut and 
+smaller values for prominence both increase the accuracy of using
+\texttt{blase}'s line detection and optimization at the expense of an
+increased computational cost.
+
+Once optimization was complete, the list of identified lines, present in the
+state dictionary of \texttt{blase}'s model, was saved to a \texttt{.pt} file.
 
 \section{Interpolating Manifolds}
-\subsection{Maintaining Line Persistence}
-Text
-\subsection{Continuously Evaluable Partials}
-Text
+\subsection{Cross-Model Line Identification}
+Earlier we mentioned that \texttt{blase} tuned the line centers of
+detected lines. This means that from one PHOENIX spectrum to the next, the
+same line could have a slightly different line center. Since the goal of this
+study is to interpolate the properties of each line, we needed to identify the
+presence of a particular line across the PHOENIX subset. We decided to do
+this by using the line centers of the detected lines pre-optimization. This
+introduced a new variable called the jitter $\delta\mu$, which measures the shift of the
+line center between pre-optimization and post-optimization. Now with each
+spectral line indexed by $\mu$, we had four parameters to interpolate:
+$\delta\mu$, $\ln(a)$, $\sigma$, and $\gamma$.
 
-\section{The PHOENIX Generator}
-\subsection{Generator Design}
-Text
-\subsection{Measuring Information Loss}
-Text
+There was also a second issue, and that is that spectral lines were often
+only detected in some spectra from the PHOENIX subset. This means that in 
+theory, interpolation would be inaccurate in regions where a line does not 
+appear, and also breaks in practice because \texttt{scipy}'s regular grid
+interpolation relies on a rectilinear grid, which we would not have if
+regions of grid points were missing. To solve this, we artificially
+populated missing grid points with log-amplitudes of -1000, which retained
+interpolator stability by not being an infinity, but also essentially 
+nullified the line when evaluated.
+
+\subsection{Continuously Evaluable Manifolds}
+For each line, the inputs to the interpolator were the three input parameters $T_{\mathrm{eff}}$,
+$\log(g)$, and $[Fe/H]$, and the output was a list of four parameters, 
+$\delta\mu$, $\ln(a)$, $\sigma$, and $\gamma$. For each line, one of these 
+interpolator objects was created using linear interpolation, and these 
+interpolators were aggregated into a single list, which was when pickled 
+and written to a \texttt{.pkl} file. These interpolators represent multiple 
+manifolds mapping fundamental stellar parameters to spectral line properties. 
+These interpolators could now be evaluated at any point lying within the 
+domain of the PHOENIX subset, turning a discretely sampled PHOENIX subset 
+into a continuous one. With the given size of the PHOENIX subset, the 
+interpolator list takes up 19 GB of disk space.
 
@@
 Text
+
 \subsection{Performance Results}
@@
 \section{Discussion}
-Text
+It should be noted that this study represents a proof of concept, and that 
+there are numerous design considerations that could be improved upon with 
+future work. These considerations include but are not limited to the 
+following:
+\begin{enumerate}[label=-]
+    \item \textit{Limited PHOENIX Subset}: The PHOENIX subset used in this 
+    study was just that, a subset. The full PHOENIX grid not only expands 
+    the [Fe/H] range to [-4.0, 1.0] dex, but also includes the alpha element 
+    abundance, which we elected to fix at 0 for this study. In addition to 
+    the actual fundamental stellar parameters, we also took a subset of the 
+    PHOENIX wavelength range, with the full [500, 55000] \AA \ wavelength 
+    range also being left to future work.
+    \item \textit{Strict Wavelength Range}: Currently, the generator only 
+    supports inference on spectra whose wavelength limits are either equal 
+    to it, or encompass that of the generator and have been truncated to 
+    match. However, when the spectrum in question has a smaller wavelength 
+    range than the generator, currently there is no functionality to truncate 
+    the generator. This would require externally indexing the generator's 
+    individual interpolators by line center position and selectively
+    evaluating those to eliminate wasteful computation.
+    \item \textit{Single Model Grid}: The PHOENIX grid is not the only model
+    grid of synthetic spectra available, and it does not apply to all types 
+    of stars. Future work would extend the reach of this study's algorithm 
+    to encompass other model grids such as the Sonora series for substellar
+    models. \texttt{blase} should be able to have an option for the user to 
+    input which model grid they would like to base the inference on, and 
+    to get even more advanced, perhaps even have the ability to intelligently
+    determine which model grid to use automatically.
+    \item \textit{Memorization vs.\ Generalization}: The current design of
+    the algorithm constructs manifolds using interpolation. This means that
+    performance is good at points close to PHOENIX subset grid points, but
+    is highly dependent on the type of interpolation used. As interpolators
+    require memorization of the data, advanced interpolation becomes extremely
+    expensive in terms of disk utilization. Future work would involve 
+    constructing manifolds using regression, which would allow for much
+    better generalization and lower disk utilization at the expense of
+    some accuracy.
+    \item \textit{Extrinsic Absence}: The current design of our algorithm
+    does not account for extrinsic parameters that modify the appearance of 
+    spectra such as rotational broadening and doppler shifting. Future work
+    would need to develop ways to tune these extrinsic parameters alongside
+    the fundamental stellar parameters.
+    \item \textit{Framework Overhead}: As this algorithm is currently more 
+    proof of concept than practical, it uses convenience functions from 
+    various libraries, which naturally introduces some level of overhead and 
+    leaves performance on the table. Future work would involve writing 
+    custom functions expressly designed for \texttt{blase}, most likely a 
+    complete rewrite of the library from the ground up.
+    \item \textit{Pseudo-Interpretability}: Our algorithm boasts interpretability
+    by considering spectral lines as the objects of interest as opposed to 
+    the rather uninterpretable flux values of other approaches. However, this
+    is only a step in the direction of interpretability. True interpretability
+    would decompose a spectrum not into a set of spectral lines, but into a
+    set of species component spectra, which requires a much more advanced 
+    understanding of different species and their behavior, as well as
+    direct access to a radiative transfer code as opposed to an off-the-shelf
+    model grid. This approach would also extend the inference from just
+    fundamental stellar parameters defined by a grid to any set of parameters
+    accounted for in the radiative transfer model, down to specific species
+    abundances.
+    \item \textit{The Continuum Black Box}: Continuum normalization is a
+    process that is not yet completely understood, and is currently done as 
+    a preprocessing step with a fairly simple algorithm. Future work would
+    dive deeper into the science of continuums and develop more advanced 
+    methods that can discern continuums with greater accuracy and less
+    modeling restrictions.
+    \item \textit{One Voigt Fits All}: The current assumption of \texttt{blase}
+    is that every spectral line is a Voigt profile. This assumption is largely
+    true, but there are situations where that is simply not enough. Future
+    studies need to account for more advanced spectral line profiles and 
+    procedures to deal with phenomena such as ro-vibrational bands.
+\end{enumerate}
+In short, it is quite clear that \texttt{blase} is a long way from being a
+powerhouse in spectral inference, however it represents a step down a road
+not yet traveled, and the potential for future growth is immense. At the 
+end of the day, we aim to develop a tool that can analyze spectrum based on 
+how nature behaves rather than how data behaves.
 

### 2024-03-27 paper/paper2_old/ms.tex

--- 2024-03-06 paper/paper2/ms.tex
@@
 \begin{document}
-\title{Precision Fundamental Stellar Properties with Adaptive Machine Learning}
+\title{Inferencing Fundamental Stellar Parameters with Interpretable Machine Learning}
 
@@
 \author[0000-0002-4404-0456]{Caroline V. Morley}
-\affil{Department of Astronomy, The University of Texas at Austin, 2515 Speedway, Austin, TX 78712, USA}
+\affil{Department of Astronomy, The University of Texas at Austin, Austin, TX 78712, USA}
 \shortauthors{Shankar \& Gully-Santiago \& Morley}
@@
 \begin{abstract}
-    \blindtext
+    Text
 \end{abstract}
 
-\keywords{}
+\keywords{Text, text, text}
 
 \section{Introduction}
-
 \subsection{General spectroscopy ideas}
@@
 
-\pagebreak
-\newpage
+\begin{acknowledgments}
+    Text
+\end{acknowledgments}
 
-\begin{acknowledgments}
-    \blindtext
-\end{acknowledgments}
+\break
 

### 2024-04-09 experiments/10_end_to_end/10_end_to_end.py
Epic inference fail

@@ -128,13 +128,13 @@ def triton_run():
 
 def inference_run():
     spec = PHOENIXSpectrum(teff=5000, logg=4, Z=0, download=True)
-    res = gp_minimize(chisq_loss(doppler_grid(8038, 12849), spec.flux.value), dimensions=[(2300, 12000), (0, 5), (-0.5, 0.5)], n_calls=100, n_random_starts=10, random_state=0, n_jobs=16)
+    res = gp_minimize(chisq_loss(spec.wavelength.value, spec.flux.value), dimensions=[(2300, 12000), (0, 5), (-0.5, 0.5)], n_calls=100, n_random_starts=10, random_state=0, n_jobs=16)
     print(res.x)
     print(res.x_iters)
     print(res.fun)
     print(res.func_vals)
 
 if __name__ == '__main__':
-    spec = PHOENIXSpectrum(teff=5000, logg=4, Z=0, download=True)
-    x = chisq_loss(spec.wavelength.value, spec.flux.value)((5000, 4, 0))
-    print(x)
+    sys.stderr = open('log.txt', 'w')
+    sys.stdout = open('output.txt', 'w')
+    inference_run()

### 2024-04-09 paper/paper2_neo/paper.tex
Auto stash before merge of "working" and "origin/working"

@@
 \section{Introduction}
-Text
+Stellar spectra are exceedingly rich sources of information about the stars
+which create them. They contain information about the fundamental properties
+of the star such as its temperature and surface gravity, as well as
+countless absorption lines reflecting the star's chemical composition. On 
+top of that, extrinsic factors such as the radial velocity of the star or 
+the rotation of the star itself further transform spectra. When these spectra
+are observed, even the location of the observing instrument and the 
+capabilities of the instrument itself introduce further modifications. 
+In short, stellar spectra represent extremely complex, high-dimensional 
+data that has gone through multiple transformations before reaching our 
+eyes as a deceptively simple graph of flux versus wavelength.
+
+Naturally, scientists want to extract as much information as possible from
+a star's spectra to get back the properties that lie at the root of it all. 
+However, this has proven to be no easy task. Spectroscopic surveys such as 
+APOGEE develop their own in-house pipelines to extract stellar parameters
+from spectra (ASPCAP in the case of APOGEE), and these pipelines are fairly
+effective. However, they all work on the core assumption that the data to be 
+analyzed is a list of pixels. In addition, the pipelines that surveys 
+develop are closed-source and usually limited to the scope of the survey 
+itself, causing a sort of fragmentation. This motivates the development of a 
+more universal, open-source tool that can do what spectral analysis pipelines 
+do, but in an alternative way.
+
+Multiple efforts have been made in this direction, treating spectra in 
+different ways. Some take the wavelength and flux values and construct
+neural networks to create this mapping between stellar parameters and the
+observed spectra, such as [EXAMPLE] and [EXAMPLE]. Others take a more 
+physics-informed approach by decomposing spectra into some eigenbasis and 
+treating it as a regression problem, such as \texttt{starfish}.
+However, \texttt{blase} takes a different approach. It treats spectra not as 
+a set of pixels or a set of eigenbasis coefficients but as a set of 
+spectral lines, specifically Voigt profiles. There is no metric of this 
+approach being better or worse than the others mentioned so far, but it is 
+a different paradigm and is worth exploring as a semi-empirical tool for
+spectral inference. This study aims to take advantage of \texttt{blase}'s
+ability to change a spectrum's basis from flux versus wavelength into
+spectral line versus spectral line properties, and use that to develop a
+robust tool for stellar parameter inference.
+
+However, this is only a proof of concept. Here we develop a tool that 
+represents a very basic realization of the idea of line-by-line inference. 
+We make no claims as to the performance of this tool, but rather aim to show 
+that the idea is sound and able to be implemented, even if rudimentary, and 
+that with further research and development, this tool has the potential to 
+become useful for spectral inference.
+
+We chose the PHOENIX synthetic spectral model grid as the basis for this 
+study, because it is fairly well-known. While we plan to extend this tool to 
+leverage information across multiple model grids in the future, both for 
+accuracy and to increase the scope to include substellar objects with grids
+such as Sonora, for now we limit our scope to a subset of the PHOENIX grid.
 
@@
         Surface Gravity & $\log(g)$ & [0, 5]\\
-        Wavelength & $\lambda$ & [8050, 12850] \AA\\
+        Wavelength & $\lambda$ & [8038, 12849] \AA\\
         \hline
@@
 Once optimization was complete, the list of identified lines, present in the
-state dictionary of \texttt{blase}'s model, was saved to a \texttt{.pt} file.
+state dictionary of \texttt{blase}'s model, was saved to a \texttt{.pt} file
+for each PHOENIX subset grid point.
 
@@
 For each line, the inputs to the interpolator were the three input parameters $T_{\mathrm{eff}}$,
-$\log(g)$, and $[Fe/H]$, and the output was a list of four parameters, 
+$\log(g)$, and [Fe/H], and the output was a list of four parameters, 
 $\delta\mu$, $\ln(a)$, $\sigma$, and $\gamma$. For each line, one of these 
@@
 into a continuous one. With the given size of the PHOENIX subset, the 
-interpolator list takes up 19 GB of disk space.
+interpolator list takes up 19 GB of disk space. Since this is now evaluable,
+we call this the PHOENIX generator.
 
-\section{Bayesian Optimization}
-Text
+\section{Bayesian Inference and Testing}
+\subsection{Inference Algorithm}
+With the PHOENIX generator complete, we used Bayesian optimization as the
+inference algorithm, specifically the \texttt{gp\_minimize} function from
+the \texttt{scikit-optimize} library. This algorithm uses a Gaussian Process
+to model the objective function, which in this case was the $\chi^2$ loss
+function between the interpolated spectrum and the true spectrum. The 
+optimizer was configured to first run 10 random evaluations to seed the
+surrogate model, then run 90 more evaluations now guided by the surrogate
+model. This totals to 100 evaluations, which was deemed sufficient for this
+study.
 
-\section{Inference Testing}
 \subsection{Test Data}
@@
 \section{Discussion}
+The computer used for this study, Triton, has the following specifications:
+\begin{table}[h]
+    \hspace*{0.7cm}\begin{tabular}{|c|c|}
+        \hline
+        \multicolumn{2}{|c|}{\textbf{Triton Specs}}\\
+        \hline\hline
+        CPU & AMD EPYC 7513\\
+        Memory & 256 GB\\
+        GPU & Nvidia A100 40GB ($\times$2)\\
+        \hline
+    \end{tabular}
+    \caption{This computer is owned by Caroline Morley and is available to
+    members of her research group. It was used for all computations, but not 
+    for generating visualizations.}
+\end{table}
+
 It should be noted that this study represents a proof of concept, and that 
@@
     evaluating those to eliminate wasteful computation.
+    \item \textit{Low-Order Interpolation}: The current design of the 
+    algorithm uses linear interpolation to construct manifolds, however this
+    could be made more sophisticated with more complex interpolation schemes.
+    However, the drawback of this is that higher orders of interpolation
+    become increasingly expensive both in terms of disk utilization, and 
+    in terms of the computational cost to generate and evaluate them. Future
+    studies should sidestep this issue by using more generalizable methods
+    that are much more efficient.
     \item \textit{Single Model Grid}: The PHOENIX grid is not the only model

### 2024-04-09 experiments/10_end_to_end/10_end_to_end.py

@@ -135,6 +135,6 @@ def inference_run():
     print(res.func_vals)
 
 if __name__ == '__main__':
-    sys.stderr = open('log.txt', 'w')
-    sys.stdout = open('output.txt', 'w')
-    inference_run()
+    spec = PHOENIXSpectrum(teff=5000, logg=4, Z=0, download=True)
+    x = chisq_loss(spec.wavelength.value, spec.flux.value)((5000, 4, 0))
+    print(x)

### 2024-04-09 experiments/10_end_to_end/10_end_to_end.py
Change interpolator backend to RegularGridInterpolator

@@ -13,7 +13,7 @@ from os import listdir
 from pickle import dump, load
 from re import split
 from skopt import gp_minimize
-from scipy.interpolate import griddata
+from scipy.interpolate import griddata, RegularGridInterpolator
 from tqdm import tqdm
 from typing import Callable
 
@@ -43,7 +43,11 @@ def create_interpolators(df: pd.DataFrame, df_gp: pd.DataFrame) -> list[partial]
     interpolator_list = []
     for line in tqdm(df.center.unique()):
         df_line = df.query('center == @line', engine='python').merge(df_gp, how='right', on=['teff', 'logg', 'Z']).fillna(-1000)
-        interpolator_list.append(partial(griddata, points=(df_line.teff, df_line.logg, df_line.Z), values=df_line[['amp', 'sigma', 'gamma', 'shift_center']].to_numpy()))
+        interpolator_list.append(RegularGridInterpolator(
+            (df_line.teff.unique(), df_line.logg.unique(), df_line.Z.unique()), 
+            df_line[['amp', 'sigma', 'gamma', 'shift_center']].to_numpy().reshape(len(df_line.teff.unique()), len(df_line.logg.unique()), len(df_line.Z.unique()), 4)),
+            method='linear')
+        #interpolator_list.append(partial(griddata, points=(df_line.teff, df_line.logg, df_line.Z), values=df_line[['amp', 'sigma', 'gamma', 'shift_center']].to_numpy()))
     return interpolator_list
 
 def reconstruct(wl_grid: np.ndarray, point: tuple[int, float, float]) -> np.ndarray:
@@ -60,60 +64,8 @@ def reconstruct(wl_grid: np.ndarray, point: tuple[int, float, float]) -> np.ndar
 def chisq_loss(wl_grid: np.ndarray, data: np.ndarray) -> Callable:
     return lambda point: np.sum((reconstruct(wl_grid, point) - data)**2)
 
-def local_run(p_teff, p_logg, p_Z):
-    path = '/home/sujay/data/10K_12.5K_clones'
-    df_native = read_state_dicts(path).query('-0.5 <= Z <= 0.5 & 4000 <= teff <= 7000 & 2 <= logg <= 4')
-    df_gp = df_native[['teff', 'logg', 'Z']]
[…]

### 2024-04-12 experiments/10_end_to_end/10_end_to_end.py
Successfully finish single inference

@@ -43,16 +43,17 @@ def create_interpolators(df: pd.DataFrame, df_gp: pd.DataFrame) -> list[partial]
     interpolator_list = []
     for line in tqdm(df.center.unique()):
         df_line = df.query('center == @line', engine='python').merge(df_gp, how='right', on=['teff', 'logg', 'Z']).fillna(-1000)
+        df_line.sort_values(['teff', 'logg', 'Z'], inplace=True)
         interpolator_list.append(RegularGridInterpolator(
             (df_line.teff.unique(), df_line.logg.unique(), df_line.Z.unique()), 
-            df_line[['amp', 'sigma', 'gamma', 'shift_center']].to_numpy().reshape(len(df_line.teff.unique()), len(df_line.logg.unique()), len(df_line.Z.unique()), 4)),
-            method='linear')
+            df_line[['amp', 'sigma', 'gamma', 'shift_center']].to_numpy().reshape(len(df_line.teff.unique()), len(df_line.logg.unique()), len(df_line.Z.unique()), 4),
+            method='linear'))
         #interpolator_list.append(partial(griddata, points=(df_line.teff, df_line.logg, df_line.Z), values=df_line[['amp', 'sigma', 'gamma', 'shift_center']].to_numpy()))
     return interpolator_list
 
 def reconstruct(wl_grid: np.ndarray, point: tuple[int, float, float]) -> np.ndarray:
-    interpolators = load(open('interpolators.pkl', 'rb'))
-    output = np.vstack([r for interpolator in interpolators if (r := interpolator(xi=(point[0], point[1], point[2])))[0] != -1000])
+    interpolators = load(open('interpolator_list.pkl', 'rb'))
+    output = np.vstack([r for interpolator in interpolators if (r := interpolator([point[0], point[1], point[2]]).squeeze())[0] != -1000])
     state_dict = {
         'amplitudes': torch.from_numpy(output[:, 0]),
         'sigma_widths': torch.from_numpy(output[:, 1]),
@@ -61,8 +62,8 @@ def reconstruct(wl_grid: np.ndarray, point: tuple[int, float, float]) -> np.ndar
     }
     return np.nan_to_num(SLE(wl_native=wl_grid, init_state_dict=state_dict, device="cpu").forward().detach().numpy(), nan=1)
 
[…]

### 2024-04-12 paper/paper2_neo/paper.tex
Update paper test data section

@@
 neural networks to create this mapping between stellar parameters and the
-observed spectra, such as [EXAMPLE] and [EXAMPLE]. Others take a more 
+observed spectra, such as \texttt{IN-SYNC}. Others take a more 
 physics-informed approach by decomposing spectra into some eigenbasis and 
@@
         Alpha Element Abundance & $\alpha$ & [0] dex\\
-        Iron Abundance & [Fe/H] & [-0.5, 0.5] dex\\
+        Iron Abundance & [Fe/H] & [-0.5, 0] dex\\
         Effective Temperature & $T_{\mathrm{eff}}$ & [2300, 12000] K\\
-        Surface Gravity & $\log(g)$ & [0, 5]\\
+        Surface Gravity & $\log(g)$ & [2, 6]\\
         Wavelength & $\lambda$ & [8038, 12849] \AA\\
@@
     \caption{The subset of the PHOENIX grid used in this study. These limits 
-    were imposed to reduce the computational cost of the algorithms. The 
-    wavelength limits in particular roughly line up with that of the 
-    Habitable Zone Planet Finder (HPF) spectrograph.}
+    were imposed to reduce the computational cost of the algorithms and to 
+    ensure a rectilinear parameter space in order to work with the chosen
+    interpolator backend. The wavelength limits in particular roughly line 
+    up with that of the Habitable Zone Planet Finder (HPF) spectrograph.}
 \end{table}
@@
 into a continuous one. With the given size of the PHOENIX subset, the 
-interpolator list takes up 19 GB of disk space. Since this is now evaluable,
+interpolator list takes up 13.2 GB of disk space. Since this is now evaluable,
 we call this the PHOENIX generator.
@@
 the \texttt{scikit-optimize} library. This algorithm uses a Gaussian Process
-to model the objective function, which in this case was the $\chi^2$ loss
-function between the interpolated spectrum and the true spectrum. The 
-optimizer was configured to first run 10 random evaluations to seed the
-surrogate model, then run 90 more evaluations now guided by the surrogate
-model. This totals to 100 evaluations, which was deemed sufficient for this
-study.
+to model the objective function, which in this case was the RMS 
+(Root-Mean-Square) loss between the interpolated spectrum and the true 
+spectrum. The optimizer was configured to first run 30 random evaluations to 
+seed the surrogate model, then run 20 more evaluations now guided by the 
+surrogate model. This totals to 50 evaluations, which was deemed sufficient 
+for this study. We then decided to use ensemble methods, running 32 parallel
+optimizers to get a more statistically sound result. One inference run 
+takes an average of [TIME] to complete on Triton.
 
 \subsection{Test Data}
-Text
+To test the performance of the inference algorithm, we used the PHOENIX subset
+itself. At first glance, this may seem like a useless endeavor, as the 
+PHOENIX generator has indeed memorized the PHOENIX subset, being able to
+reconstruct a PHOENIX spectrum when evaluated at that grid point. However,
+that is precisely what allows us to use the PHOENIX subset as test data.
+The Bayesian optimizer's surrogate model is seeded by random continuously 
+sampled generator evaluations within the search space, \textit{not} grid 
+points of the PHOENIX subset, meaning the surrogate model has no `memorization' 
+to speak of. If the optimizer had been an iterative grid refinement search, 
+this would not have been possible, because grid methods rely on just that, 
+the grid.
 
@@
 \begin{table}[h]
-    \hspace*{0.7cm}\begin{tabular}{|c|c|}
+    \hspace*{0.6cm}\begin{tabular}{|c|c|}
         \hline
@@
     members of her research group. It was used for all computations, but not 
-    for generating visualizations.}
+    for generating visualizations. The EPYC 7513 is a 32c/64t CPU with a 
+    boost clock of 3.65 GHz.}
 \end{table}
@@
     study was just that, a subset. The full PHOENIX grid not only expands 
-    the [Fe/H] range to [-4.0, 1.0] dex, but also includes the alpha element 
-    abundance, which we elected to fix at 0 for this study. In addition to 
-    the actual fundamental stellar parameters, we also took a subset of the 
-    PHOENIX wavelength range, with the full [500, 55000] \AA \ wavelength 
-    range also being left to future work.
+    the [Fe/H] range to [-4.0, 1.0] dex and the $\log(g)$ range to [0, 6], 
+    but also includes the alpha element abundance parameter, which we 
+    elected to fix at 0 for this study. In addition to the actual fundamental 
+    stellar parameters, we also took a subset of the PHOENIX wavelength range, 
+    with the full [500, 55000] \AA \ wavelength range also being left to 
+    future work.
     \item \textit{Strict Wavelength Range}: Currently, the generator only 
@@
 not yet traveled, and the potential for future growth is immense. At the 
-end of the day, we aim to develop a tool that can analyze spectrum based on 
-how nature behaves rather than how data behaves.
+end of the day, we aim to develop a tool that can analyze spectra with the 
+advent of physics-informed machine learning.
 

### 2024-04-12 experiments/10_end_to_end/10_end_to_end.py

@@ -48,7 +48,6 @@ def create_interpolators(df: pd.DataFrame, df_gp: pd.DataFrame) -> list[partial]
             (df_line.teff.unique(), df_line.logg.unique(), df_line.Z.unique()), 
             df_line[['amp', 'sigma', 'gamma', 'shift_center']].to_numpy().reshape(len(df_line.teff.unique()), len(df_line.logg.unique()), len(df_line.Z.unique()), 4),
             method='linear'))
-        #interpolator_list.append(partial(griddata, points=(df_line.teff, df_line.logg, df_line.Z), values=df_line[['amp', 'sigma', 'gamma', 'shift_center']].to_numpy()))
     return interpolator_list
 
 def reconstruct(wl_grid: np.ndarray, point: tuple[int, float, float]) -> np.ndarray:
@@ -88,7 +87,4 @@ def inference_run():
     print(f'Result: {res.x} achieved in {perf_counter() - start} s')
 
 if __name__ == '__main__':
-    #df = pd.DataFrame({'x': [0, 0, 1, 1], 'y': [2, 3, 2, 3], 'u': [4, 5, 6, 7], 'v': [8, 9, 10, 11], 'w': [12, 13, 14, 15]})
-    #interp2 = RegularGridInterpolator((df.x.unique(), df.y.unique()), df[['u', 'v', 'w']].to_numpy().reshape(2, 2, 3))
-    #print(interp2([0, 3]))
     inference_run()

### 2024-04-16 paper/paper2_neo/paper.tex
Write multiprocessing full grid inference

@@
 
-Naturally, scientists want to extract as much information as possible from
+Scientists want to extract as much information as possible from
 a star's spectra to get back the properties that lie at the root of it all. 
@@
 analyzed is a list of pixels. In addition, the pipelines that surveys 
-develop are closed-source and usually limited to the scope of the survey 
-itself, causing a sort of fragmentation. This motivates the development of a 
-more universal, open-source tool that can do what spectral analysis pipelines 
-do, but in an alternative way.
+develop are typically closed-source and usually limited to the scope of the 
+survey itself, causing a sort of fragmentation. This motivates the 
+development of a more universal, open-source tool that can do what spectral 
+analysis pipelines do, but in an alternative way.
 
 Multiple efforts have been made in this direction, treating spectra in 
-different ways. Some take the wavelength and flux values and construct
-neural networks to create this mapping between stellar parameters and the
-observed spectra, such as \texttt{IN-SYNC}. Others take a more 
-physics-informed approach by decomposing spectra into some eigenbasis and 
-treating it as a regression problem, such as \texttt{starfish}.
-However, \texttt{blase} takes a different approach. It treats spectra not as 
-a set of pixels or a set of eigenbasis coefficients but as a set of 
-spectral lines, specifically Voigt profiles. There is no metric of this 
-approach being better or worse than the others mentioned so far, but it is 
-a different paradigm and is worth exploring as a semi-empirical tool for
-spectral inference. This study aims to take advantage of \texttt{blase}'s
-ability to change a spectrum's basis from flux versus wavelength into
-spectral line versus spectral line properties, and use that to develop a
-robust tool for stellar parameter inference.
+different ways. The standard practice is to treat the wavelength and flux as 
+simply two arrays and implement any from a myriad of different algorithms to 
+inference fundamental stellar parameters. Others take a more physics-informed 
+approach by decomposing spectra into some eigenbasis and treating it as a 
+regression problem, such as \texttt{starfish}. However, \texttt{blase} takes 
+a different approach. It treats spectra not as a set of pixels or a set of 
+eigenbasis coefficients but as a set of spectral lines, specifically Voigt 
+profiles. There is no metric of this approach being better or worse than the 
+others mentioned so far, but it is a different paradigm and is worth 
+exploring as a semi-empirical tool. This study aims to take advantage of 
+\texttt{blase}'s ability to change a spectrum's basis from flux versus 
+wavelength into spectral line versus spectral line properties, and use that 
+to develop a robust tool for stellar parameter inference.
 
@@
 \begin{table}[h]
-    \hspace*{-1.1cm}\begin{tabular}{|c|c|c|}
+    \hspace*{-1.35cm}\begin{tabular}{|c|c|c|}
         \hline
@@
     were imposed to reduce the computational cost of the algorithms and to 
-    ensure a rectilinear parameter space in order to work with the chosen
-    interpolator backend. The wavelength limits in particular roughly line 
-    up with that of the Habitable Zone Planet Finder (HPF) spectrograph.}
+    ensure a rectilinear parameter space in order to work with \texttt{scipy}'s
+    \texttt{RegularGridInterpolator}. The wavelength limits in particular 
+    roughly line up with that of the Habitable Zone Planet Finder (HPF) 
+    spectrograph.}
 \end{table}
@@
 First, the PHOENIX subset was accessed directly from the PHOENIX website
-using \texttt{gollum}'s download option, which uses the website's FTP server
-to download spectra without using any disk space. This came at the cost of
-longer load times due to internet connection being significantly more of a
-bottleneck than disk I/O, but was acceptable as this was a one-time operation.
-
-The spectra were then put through a three-step preprocessing pipeline.
+using \texttt{gollum}'s download option. The spectra were then put through a 
+three-step preprocessing pipeline.
 \begin{enumerate}

### 2024-04-16 experiments/10_end_to_end/10_end_to_end.py

@@ -1,19 +1,19 @@
-import matplotlib.pyplot as plt
 import numpy as np
 import pandas as pd
 import sys
 import torch
 
 from blase.emulator import SparseLinearEmulator as SLE
-from blase.utils import doppler_grid
+from blase.optimizer import default_clean
 from collections import defaultdict
-from functools import partial
-from gollum.phoenix import PHOENIXSpectrum
+from gollum.phoenix import PHOENIXSpectrum, PHOENIXGrid
+from itertools import repeat
+from multiprocessing import Pool
 from os import listdir
 from pickle import dump, load
 from re import split
 from skopt import gp_minimize
-from scipy.interpolate import griddata, RegularGridInterpolator
+from scipy.interpolate import RegularGridInterpolator
 from tqdm import tqdm
 from typing import Callable
 
@@ -39,7 +39,7 @@ def optimize_memory(df: pd.DataFrame):
     df[fcols] = df[fcols].apply(pd.to_numeric, downcast='float')
     df[icols] = df[icols].apply(pd.to_numeric, downcast='integer')
 
-def create_interpolators(df: pd.DataFrame, df_gp: pd.DataFrame) -> list[partial]:
+def create_interpolators(df: pd.DataFrame, df_gp: pd.DataFrame) -> list[RegularGridInterpolator]:
     interpolator_list = []
     for line in tqdm(df.center.unique()):
         df_line = df.query('center == @line', engine='python').merge(df_gp, how='right', on=['teff', 'logg', 'Z']).fillna(-1000)
@@ -78,13 +78,28 @@ def triton_run():
     with open('log.txt', 'w') as f:
         f.write('Interpolator partials dumped to pickle.')
 
-def inference_run():
-    from blase.optimizer import default_clean
+def inference_test():
     from time import perf_counter
     start = perf_counter()
     spec = default_clean(PHOENIXSpectrum(teff=5000, logg=4, Z=0, download=True))
     res = gp_minimize(loss_fn(spec.wavelength.value, spec.flux.value), dimensions=[(2300, 12000), (2, 6), (-0.5, 0)], n_calls=50, n_random_starts=30)
     print(f'Result: {res.x} achieved in {perf_counter() - start} s')
 
+def inference(spec: PHOENIXSpectrum):
[…]

### 2024-04-17 experiments/10_end_to_end/10_end_to_end.py
Almost killed Triton, multiprocessing is flaky

@@ -85,21 +85,21 @@ def inference_test():
     res = gp_minimize(loss_fn(spec.wavelength.value, spec.flux.value), dimensions=[(2300, 12000), (2, 6), (-0.5, 0)], n_calls=50, n_random_starts=30)
     print(f'Result: {res.x} achieved in {perf_counter() - start} s')
 
-def inference(spec: PHOENIXSpectrum):
-    return np.array(gp_minimize(loss_fn(spec.wavelength.value, spec.flux.value), dimensions=[(2300, 12000), (2, 6), (-0.5, 0)], n_calls=50, n_random_starts=30).x)
+def inference(spec: tuple[np.ndarray]):
+    return np.array(gp_minimize(loss_fn(spec[0], spec[1]), dimensions=[(2300, 12000), (2, 6), (-0.5, 0)], n_calls=50, n_random_starts=30).x)
 
 def inference_grid():
     sys.stderr = open('log.txt', 'w')
+    N_CORES = 10
     grid = PHOENIXGrid(teff_range=(2300, 12000), logg_range=(2, 6), Z_range=(-0.5, 0), path='/data/libraries/raw/PHOENIX/')
     df_list = []
     for point in grid.grid_points:
-        spec = default_clean(grid[point])
-        with Pool(32) as p:
-            result = p.imap_unordered(inference, repeat(spec, 32))
-        result = np.vstack(list(result))
-        df_list.append(pd.DataFrame({'teff': [point[0]]*32, 'logg': [point[1]]*32, 'Z': [point[2]]*32, 'i_teff': result[:, 0], 'i_logg': result[:, 1], 'i_Z': result[:, 2]}))
+        spec = default_clean(grid[grid.lookup_dict[point]])
+        with Pool(N_CORES) as p:
+            result = np.vstack([*p.imap_unordered(inference, repeat((spec.wavelength.value, spec.flux.value), N_CORES))])
+        df_list.append(pd.DataFrame({'teff': [point[0]]*N_CORES, 'logg': [point[1]]*N_CORES, 'Z': [point[2]]*N_CORES, 'i_teff': result[:, 0], 'i_logg': result[:, 1], 'i_Z': result[:, 2]}))
     pd.concat(df_list).to_parquet('inference_results.parquet.gz', compression='gzip')
 
 
 if __name__ == '__main__':
-    inference_grid()
+    inference_test()

### 2024-04-17 experiments/10_end_to_end/10_end_to_end.py
Try to optimize inference

@@ -7,8 +7,6 @@ from blase.emulator import SparseLinearEmulator as SLE
 from blase.optimizer import default_clean
 from collections import defaultdict
 from gollum.phoenix import PHOENIXSpectrum, PHOENIXGrid
-from itertools import repeat
-from multiprocessing import Pool
 from os import listdir
 from pickle import dump, load
 from re import split
@@ -50,21 +48,20 @@ def create_interpolators(df: pd.DataFrame, df_gp: pd.DataFrame) -> list[RegularG
             method='linear'))
     return interpolator_list
 
-def reconstruct(wl_grid: np.ndarray, point: tuple[int, float, float]) -> np.ndarray:
-    interpolators = load(open('interpolator_list.pkl', 'rb'))
-    output = np.vstack([r for interpolator in interpolators if (r := interpolator([point[0], point[1], point[2]]).squeeze())[0] != -1000])
+def reconstruct(wl_grid: np.ndarray, point: tuple[int, float, float], interpolator_list: list[RegularGridInterpolator]) -> np.ndarray:
+    output = np.vstack([r for interpolator in interpolator_list if (r := interpolator([point[0], point[1], point[2]]).squeeze())[0] != -1000])
     state_dict = {
         'amplitudes': torch.from_numpy(output[:, 0]),
         'sigma_widths': torch.from_numpy(output[:, 1]),
         'gamma_widths': torch.from_numpy(output[:, 2]),
         'lam_centers': torch.from_numpy(output[:, 3]),
     }
-    return np.nan_to_num(SLE(wl_native=wl_grid, init_state_dict=state_dict, device="cpu").forward().detach().numpy(), nan=1)
+    return np.nan_to_num(SLE(wl_native=wl_grid, init_state_dict=state_dict).forward().detach().numpy(), nan=1)
 
-def loss_fn(wl_grid: np.ndarray, data: np.ndarray) -> Callable:
-    return lambda point: ((reconstruct(wl_grid, point) - data)**2).mean()**0.5
+def loss_fn(wl_grid: np.ndarray, data: np.ndarray, interpolator_list: list[RegularGridInterpolator]) -> Callable:
+    return lambda point: ((reconstruct(wl_grid, point, interpolator_list) - data)**2).mean()**0.5
 
-def triton_run():
+def pickling_run():
[…]

### 2024-04-19 paper/paper2_neo/paper.tex
Add figures and math to paper

@@
 simply two arrays and implement any from a myriad of different algorithms to 
-inference fundamental stellar parameters. Others take a more physics-informed 
+inference fundamental stellar parameters. Others take a more mathematical 
 approach by decomposing spectra into some eigenbasis and treating it as a 
@@
 
-\begin{table}[h]
+\begin{table}[h!]
     \hspace*{-1.35cm}\begin{tabular}{|c|c|c|}
@@
     roughly line up with that of the Habitable Zone Planet Finder (HPF) 
-    spectrograph.}
+    spectrograph. This subset is comprised of 1314 individual spectra.}
 \end{table}
@@
 \begin{enumerate}
-    \item Blackbody Division: Since the $T_{\mathrm{eff}}$ of each spectrum 
+    \item \textit{Blackbody Division}: Since the $T_{\mathrm{eff}}$ of each spectrum 
     is known, the according blackbody spectrum was divided out.
-    \item Percentile Normalization: The spectra were normalized by dividing
+    \item \textit{Percentile Normalization}: The spectra were normalized by dividing
     them by their 99th percentile.
-    \item Continuum Normalization: The spectra were further normalized by 
+    \item \textit{Continuum Normalization}: The spectra were further normalized by 
     dividing them by a \nth{5} order polynomial continuum.
 \end{enumerate}
+Mathematically, we can express the preprocessing as follows:
+\begin{gather}
+    \mathsf{\bar{S}} = \dfrac{\mathsf{S}}{\mathsf{B}\mathsf{Q}_5P_{99}}
+\end{gather}
+where $\mathsf{\bar{S}}$ is the preprocessed spectrum, $\mathsf{S}$ is the 
+original spectrum,  $\mathsf{B}$ is the blackbody spectrum, $\mathsf{Q}_n$ 
+is the $n^\mathrm{th}$ order polynomial continuum fit, and $P_n$ is the 
+$n^\mathrm{th}$ percentile function. Arithmetic operations between arrays 
+are assumed to be elementwise in all notation from here on out.
 
@@
 state dictionary of \texttt{blase}'s model, was saved to a \texttt{.pt} file
-for each PHOENIX subset grid point.
+for each PHOENIX subset grid point. The total disk space these files took up 
+is 465 MB.
 
@@
 presence of a particular line across the PHOENIX subset. We decided to do
-this by using the line centers of the detected lines pre-optimization. This
-introduced a new variable called the jitter $\delta\mu$, which measures the shift of the
-line center between pre-optimization and post-optimization. Now with each
-spectral line indexed by $\mu$, we had four parameters to interpolate:
-$\delta\mu$, $\ln(a)$, $\sigma$, and $\gamma$.
+this by using the line centers $\mu$ of the detected lines pre-optimization, 
+while the post-optimization line centers are labeled $\mu'$. Now with 
+each spectral line indexed by $\mu$, we had four parameters to interpolate:
+$\mu'$, $\ln(a)$, $\sigma$, and $\gamma$. Mathematically, the Voigt profile
+$\mathsf{V}_\mu$ in terms of our four parameters is given by:
 
-There was also a second issue, and that is that spectral lines were often
-only detected in some spectra from the PHOENIX subset. This means that in 
-theory, interpolation would be inaccurate in regions where a line does not 
-appear, and also breaks in practice because \texttt{scipy}'s regular grid
-interpolation relies on a rectilinear grid, which we would not have if
-regions of grid points were missing. To solve this, we artificially
+\begin{gather}
+    \mathsf{V}_\mu(\lambda) = e^{\ln(a)}\mathrm{Re}\left[w\left(\dfrac{\lambda - \mu' + i\gamma}{\sigma\sqrt{2}}\right)\right] \bigg/ \left(\sigma\sqrt{2\pi}\right)
+\end{gather}
+where $w$ is the Faddeeva function. Note that since we are dealing with 
+the parameters of these Voigt profiles, even if the interpolation method is 
+linear, the actual result of interpolation varies non-linearly.
+
+There is also a second issue, and that is that spectral lines were often
+only detected in some spectra from the PHOENIX subset. In Figure 1, we show
+that different grid points sport differing counts of detected spectral lines.
+This means that in theory, interpolation would be inaccurate in regions where 
+a line does not appear, and also breaks in practice because \texttt{scipy}'s 
+\texttt{RegularGridInterpolator} relies on a rectilinear grid, which we would not 
+have if regions of grid points were missing.
+
+\begin{figure*}[t!]
+    \centering
+    \includegraphics[width=\textwidth]{images/figure1.png}
+    \caption{The number of detected spectral lines at each grid point of 
+    a slice of the PHOENIX subset at solar metallicity. We can see that
+    the number of detected lines decreases with increasing $T_{\mathrm{eff}}$.
+    Also note that from $T_{\mathrm{eff}} = 7000$ K onward, the spacing 
+    between grid points increases from 100 K to 200 K.}
+\end{figure*}
+
+To solve this, we artificially
 populated missing grid points with log-amplitudes of -1000, which retained
 interpolator stability by not being an infinity, but also essentially 
-nullified the line when evaluated.
+nullified the line when evaluated. An example of the appearance of missing 
+sections in heatmaps where a line does not appear is shown in Figure 2. 
+In total, across the entire PHOENIX subset, \texttt{blase} detected 
+128,723 individual spectral lines, meaning we have that many manifolds to 
+interpolate.
+
+\begin{figure*}[t!]
+    \centering
+    \includegraphics[width=\textwidth]{images/figure2.png}
+    \caption{Heatmap showing how $\ln(a)$ varies over the PHOENIX subset
+    slice at solar metallicity. Notice the missing chunk in the top left
+    of the figure; \texttt{blase} did not detect a spectral line here,
+    but we have to artificially populate those points with lines that 
+    have $\ln(a) = -1000$.}
+\end{figure*}
 
@@
 $\log(g)$, and [Fe/H], and the output was a list of four parameters, 
-$\delta\mu$, $\ln(a)$, $\sigma$, and $\gamma$. For each line, one of these 
+$\mu'$, $\ln(a)$, $\sigma$, and $\gamma$. For each line, one of these 
 interpolator objects was created using linear interpolation, and these 
 interpolators were aggregated into a single list, which was when pickled 
-and written to a \texttt{.pkl} file. These interpolators represent multiple 
-manifolds mapping fundamental stellar parameters to spectral line properties. 
+and written to a \texttt{.pkl} file. These interpolators generate multiple 
+manifolds with the following mapping:
+\begin{gather}
+    \mqty[T_{\mathrm{eff}} \\ \log(g) \\ \mathrm{[Fe/H]}] \rightarrow \mqty[\mu' \\ \ln(a) \\ \sigma \\ \gamma]
+\end{gather}
 These interpolators could now be evaluated at any point lying within the 
@@
 optimizers to get a more statistically sound result. One inference run 
-takes an average of [TIME] to complete on Triton.
+takes an average of 15 minutes to complete on Triton (see Section 5 for
+specifications).
 
@@
 The computer used for this study, Triton, has the following specifications:
-\begin{table}[h]
+\begin{table}[h!]
     \hspace*{0.6cm}\begin{tabular}{|c|c|}

### 2024-04-19 experiments/10_end_to_end/10_end_to_end.py

@@ -48,19 +48,6 @@ def create_interpolators(df: pd.DataFrame, df_gp: pd.DataFrame) -> list[RegularG
             method='linear'))
     return interpolator_list
 
-def reconstruct(wl_grid: np.ndarray, point: tuple[int, float, float], interpolator_list: list[RegularGridInterpolator]) -> np.ndarray:
-    output = np.vstack([r for interpolator in interpolator_list if (r := interpolator([point[0], point[1], point[2]]).squeeze())[0] != -1000])
-    state_dict = {
-        'amplitudes': torch.from_numpy(output[:, 0]),
-        'sigma_widths': torch.from_numpy(output[:, 1]),
-        'gamma_widths': torch.from_numpy(output[:, 2]),
-        'lam_centers': torch.from_numpy(output[:, 3]),
-    }
-    return np.nan_to_num(SLE(wl_native=wl_grid, init_state_dict=state_dict).forward().detach().numpy(), nan=1)
-
-def loss_fn(wl_grid: np.ndarray, data: np.ndarray, interpolator_list: list[RegularGridInterpolator]) -> Callable:
-    return lambda point: ((reconstruct(wl_grid, point, interpolator_list) - data)**2).mean()**0.5
-
 def pickling_run():
     sys.stderr = open('log.txt', 'w')
     path = '/home/sujays/github/blase/experiments/08_blase3D_HPC_test/emulator_states'
@@ -74,6 +61,25 @@ def pickling_run():
     dump(interpolators, open('interpolator_list.pkl', 'wb'))
     with open('log.txt', 'w') as f:
         f.write('Interpolator partials dumped to pickle.')
+
+def reconstruct(wl_grid: np.ndarray, point: tuple[int, float, float], interpolator_list: list[RegularGridInterpolator]) -> np.ndarray:
+    output = np.vstack([r for interpolator in interpolator_list if (r := interpolator([point[0], point[1], point[2]]).squeeze())[0] != -1000])
+    state_dict = {
+        'amplitudes': torch.from_numpy(output[:, 0]),
+        'sigma_widths': torch.from_numpy(output[:, 1]),
+        'gamma_widths': torch.from_numpy(output[:, 2]),
+        'lam_centers': torch.from_numpy(output[:, 3]),
+    }
+    return np.nan_to_num(SLE(wl_native=wl_grid, init_state_dict=state_dict).forward().detach().numpy(), nan=1)
+
[…]

### 2024-04-23 experiments/10_end_to_end/10_end_to_end.py
Generate DataFrame for interpolated heatmap visualization

@@ -7,6 +7,7 @@ from blase.emulator import SparseLinearEmulator as SLE
 from blase.optimizer import default_clean
 from collections import defaultdict
 from gollum.phoenix import PHOENIXSpectrum, PHOENIXGrid
+from itertools import product
 from os import listdir
 from pickle import dump, load
 from re import split
@@ -39,7 +40,7 @@ def optimize_memory(df: pd.DataFrame):
 
 def create_interpolators(df: pd.DataFrame, df_gp: pd.DataFrame) -> list[RegularGridInterpolator]:
     interpolator_list = []
-    for line in tqdm(df.center.unique()):
+    for line in tqdm(df.value_counts('center').index):
         df_line = df.query('center == @line', engine='python').merge(df_gp, how='right', on=['teff', 'logg', 'Z']).fillna(-1000)
         df_line.sort_values(['teff', 'logg', 'Z'], inplace=True)
         interpolator_list.append(RegularGridInterpolator(
@@ -49,7 +50,8 @@ def create_interpolators(df: pd.DataFrame, df_gp: pd.DataFrame) -> list[RegularG
     return interpolator_list
 
 def pickling_run():
-    sys.stderr = open('log.txt', 'w')
+    from time import perf_counter
+    sys.stderr = sys.stdout = open('log.txt', 'w')
     path = '/home/sujays/github/blase/experiments/08_blase3D_HPC_test/emulator_states'
     df = read_state_dicts(path)
     df_gp = df[['teff', 'logg', 'Z']]
@@ -57,10 +59,10 @@ def pickling_run():
     print('DataFrame created')
     optimize_memory(df)
     print('DataFrame memory optimized')
+    start = perf_counter()
     interpolators = create_interpolators(df, df_gp)
     dump(interpolators, open('interpolator_list.pkl', 'wb'))
-    with open('log.txt', 'w') as f:
-        f.write('Interpolator partials dumped to pickle.')
+    print(f'Interpolator partials dumped to pickle ({perf_counter() - start} s).')
 
 def reconstruct(wl_grid: np.ndarray, point: tuple[int, float, float], interpolator_list: list[RegularGridInterpolator]) -> np.ndarray:
     output = np.vstack([r for interpolator in interpolator_list if (r := interpolator([point[0], point[1], point[2]]).squeeze())[0] != -1000])
[…]

### 2024-04-23 paper/paper2_neo/paper.tex
Start adding figures to paper

@@
     \includegraphics[width=\textwidth]{images/figure2.png}
-    \caption{Heatmap showing how $\ln(a)$ varies over the PHOENIX subset
+    \caption{Heatmap showing how $\ln(a)$ and $\sigma$ vary over the PHOENIX subset
     slice at solar metallicity. Notice the missing chunk in the top left
@@
 into a continuous one. With the given size of the PHOENIX subset, the 
-interpolator list takes up 13.2 GB of disk space. Since this is now evaluable,
+interpolator list takes up 13.2 GB of disk space. This evaluation 
+is able to reconstruct an existing PHOENIX spectrum or alternatively 
+interpolate a new spectrum within the domain of the PHOENIX subset,
 we call this the PHOENIX generator.
@@
 
-\subsection{Performance Results}
-Text
+\subsection{Spectral Reconstruction Time}
+Typically a practitioner would wish to use a reconstruction in an iterative 
+way, repeatedly generating spectra from input (teff, logg, FeH) 
+coordinates.  Therefore, there is a strong desire to bring the computational 
+cost of reconstruction to extremely low values close to subsecond.
+We evaluate the computational time of our generator in two ways.  First, 
+for a single input. When evaluating a vector of inputs, we parallelize the interpolator evaluation, 
+however the reconstruction done by \texttt{blase} on each evaluated state 
+dictionary is done serially.
+
+The single input performance 
+takes $\sim$30 seconds, which is much too large for typical Bayesian inference procedures.
+
+I hand in a (teff, logg, FeH) pair, how long does it take to get back a reconstructed spectrum?
+
+\subsection{Bayesian Optimizer Accuracy}
+TACC
 

### 2024-04-25 paper/paper2_neo/paper.tex
Change paper and create multi-reconstructor

--- 2024-04-19 paper/paper2_neo/paper.tex
@@
 \texttt{blase}'s line detection and optimization at the expense of an
-increased computational cost.
-
-Once optimization was complete, the list of identified lines, present in the
-state dictionary of \texttt{blase}'s model, was saved to a \texttt{.pt} file
-for each PHOENIX subset grid point. The total disk space these files took up 
-is 465 MB.
+increased computational cost. In addition, it should be mentioned that
+\texttt{blase} uses the pseudo-Voigt approximation, which saves on 
+computational cost while remaining accurate to about 1\%. It is a 
+weighted average of a Gaussian and Lorentzian as opposed to a convolution.
+\texttt{blase}'s pseudo-Voight profile implementation uses the following:
+\begin{gather}
+    \tilde{\mathsf{V}}_\mu(\lambda) = \eta L(\lambda - \mu'; f) + (1 - \eta)G(\lambda - \mu'; f)\\
+    \nonumber \eta = \sum_{n=1}^3 a_n\left(\dfrac{2\gamma}{f}\right)^n \qquad a = \mqty[1.36603 \\ -0.47719 \\ 0.11116]\\
+    \nonumber f = 32\sum_{n=0}^5 b_n \left(\sqrt{2\ln(2)}\sigma\right)^{5-n}(\gamma)^n \qquad b = \mqty[1 \\ 2.69269 \\ 2.42843 \\ 4.47163 \\ 0.07842 \\ 1]
+\end{gather}
+Notice that we use $\mu'$ instead of $\mu$ in the formula. This is because
+\texttt{blase} allows the line center to shift slightly during optimization, 
+and it is this shifted center which is used in computation. The individual 
+Voigt profiles are still indexed by $\mu$ for cross-model line identification,
+explained in the next section. Once optimization was complete, the list of 
+identified lines, present in the state dictionary of \texttt{blase}'s model, 
+was saved to a \texttt{.pt} file for each PHOENIX subset grid point. The total 
+disk space these files took up is 465 MB.
 
@@
 presence of a particular line across the PHOENIX subset. We decided to do
-this by using the line centers $\mu$ of the detected lines pre-optimization, 
-while the post-optimization line centers are labeled $\mu'$. Now with 
-each spectral line indexed by $\mu$, we had four parameters to interpolate:
-$\mu'$, $\ln(a)$, $\sigma$, and $\gamma$. Mathematically, the Voigt profile
-$\mathsf{V}_\mu$ in terms of our four parameters is given by:
-
-\begin{gather}
-    \mathsf{V}_\mu(\lambda) = e^{\ln(a)}\mathrm{Re}\left[w\left(\dfrac{\lambda - \mu' + i\gamma}{\sigma\sqrt{2}}\right)\right] \bigg/ \left(\sigma\sqrt{2\pi}\right)
-\end{gather}
-where $w$ is the Faddeeva function. Note that since we are dealing with 
-the parameters of these Voigt profiles, even if the interpolation method is 
-linear, the actual result of interpolation varies non-linearly.
+this by using the line centers $\mu$ of the detected lines pre-optimization. 
+Now with each spectral line indexed by $\mu$, we had four parameters to interpolate:
+$\mu'$, $\ln(a)$, $\sigma$, and $\gamma$. Note that since we are dealing with 
+the parameters of Voigt profiles, we can see in Equation 2 that even if the 
+interpolation method is linear, the actual result of interpolation varies 
+non-linearly.
 
@@
     but we have to artificially populate those points with lines that 
-    have $\ln(a) = -1000$.}
+    have $\ln(a) = -1000$. The change in spacing on the x-axis is owing to
+    the step of $T_\mathrm{eff}$ in the PHOENIX grid changing from 100 K to 
+    200 K after $T_\mathrm{eff} = 7000$ K.}
 \end{figure*}
@@
 into a continuous one. With the given size of the PHOENIX subset, the 
-interpolator list takes up 13.2 GB of disk space. Since this is now evaluable,
-we call this the PHOENIX generator.
+interpolator list takes up 13.2 GB of disk space. This evaluation 
+is able to reconstruct an existing PHOENIX spectrum or alternatively 
+interpolate a new spectrum within the domain of the PHOENIX subset,
+we call this the PHOENIX generator. In Figure 3, we show the same 
+spectral line as in Figure 2, but now supersampled using the PHOENIX
+generator evaluated over the same slice.
+
+\begin{figure*}[t!]
+    \centering
+    \includegraphics[width=\textwidth]{images/figure3.png}
+    \caption{Heatmap showing how $\ln(a)$ and $\sigma$ vary over the PHOENIX 
+    subset slice at solar metallicity, now supersampled with the PHOENIX 
+    generator. Notice that the missing chunk in the top left still exists 
+    and does not display any artifacts, as the artificially populated points 
+    are removed after interpolation to retain the model's integrity. Also
+    see that the x-axis spacing is now uniform, as the PHOENIX generator 
+    was evaluated at constant step.}
+\end{figure*}
 
@@
 
-\subsection{Performance Results}
-Text
+\subsection{Spectral Reconstruction Time}
+The typical use case for the PHOENIX generator may be to batch 
+reconstruct spectra from an array of input coordinates. Therefore, there
+is some motivation to attempt to reduce the computational cost of this
+procedure to tractable levels. We evaluate the computational time needed
+to use the PHOENIX generator in two distinct ways. First, for a single
+input, which would be relevant in serial applications. Second, for an array
+of multiple inputs, which would be relevant in parallel applications.
+The \texttt{RegularGridInterpolator} API allows for the passing in of 
+an entire array of input coordinates to be evaluated at once, however 
+using \texttt{blase} to reconstruct the spectrum from our interpolated
+state dictionary is done serially, leading to what is actually more of 
+a psuedo-parallel evaluation.
+
+\subsection{Bayesian Optimizer Accuracy}
+TACC
 

### 2024-04-25 experiments/10_end_to_end/10_end_to_end.py

@@ -7,12 +7,12 @@ from blase.emulator import SparseLinearEmulator as SLE
 from blase.optimizer import default_clean
 from collections import defaultdict
 from gollum.phoenix import PHOENIXSpectrum, PHOENIXGrid
-from itertools import product
 from os import listdir
 from pickle import dump, load
 from re import split
 from skopt import gp_minimize
 from scipy.interpolate import RegularGridInterpolator
+from time import perf_counter
 from tqdm import tqdm
 from typing import Callable
 
@@ -64,8 +64,8 @@ def pickling_run():
     dump(interpolators, open('interpolator_list.pkl', 'wb'))
     print(f'Interpolator partials dumped to pickle ({perf_counter() - start} s).')
 
-def reconstruct(wl_grid: np.ndarray, point: tuple[int, float, float], interpolator_list: list[RegularGridInterpolator]) -> np.ndarray:
-    output = np.vstack([r for interpolator in interpolator_list if (r := interpolator([point[0], point[1], point[2]]).squeeze())[0] != -1000])
+def reconstruct1(wl_grid: np.ndarray, point: np.ndarray, interpolator_list: list[RegularGridInterpolator]) -> np.ndarray:
+    output = np.vstack([r for interpolator in interpolator_list if (r := interpolator(point).squeeze())[0] != -1000])
     state_dict = {
         'amplitudes': torch.from_numpy(output[:, 0]),
         'sigma_widths': torch.from_numpy(output[:, 1]),
@@ -74,44 +74,39 @@ def reconstruct(wl_grid: np.ndarray, point: tuple[int, float, float], interpolat
     }
     return np.nan_to_num(SLE(wl_native=wl_grid, init_state_dict=state_dict).forward().detach().numpy(), nan=1)
 
-def random_starts(wl_grid: np.ndarray, data: np.ndarray, interpolator_list: list[RegularGridInterpolator], n: int) -> tuple[np.ndarray, np.ndarray]:
-    random_teff =  np.random.randint(2300, 12000, n)
-    random_logg = np.random.uniform(2, 6, n)
-    random_Z = np.random.uniform(-0.5, 0, n)
-    
+def reconstructn(wl_grid: np.ndarray, points: np.ndarray, interpolator_list: list[RegularGridInterpolator]) -> np.ndarray:
+    raw_outputs = [[] for _ in points]
[…]

### 2024-04-25 experiments/10_end_to_end/10_end_to_end.py
Add SLE devices, multi-reconstruction successful

@@ -72,7 +72,7 @@ def reconstruct1(wl_grid: np.ndarray, point: np.ndarray, interpolator_list: list
         'gamma_widths': torch.from_numpy(output[:, 2]),
         'lam_centers': torch.from_numpy(output[:, 3]),
     }
-    return np.nan_to_num(SLE(wl_native=wl_grid, init_state_dict=state_dict).forward().detach().numpy(), nan=1)
+    return np.nan_to_num(SLE(wl_native=wl_grid, init_state_dict=state_dict, device='cpu').forward().detach().numpy(), nan=1)
 
 def reconstructn(wl_grid: np.ndarray, points: np.ndarray, interpolator_list: list[RegularGridInterpolator]) -> np.ndarray:
     raw_outputs = [[] for _ in points]
@@ -88,7 +88,7 @@ def reconstructn(wl_grid: np.ndarray, points: np.ndarray, interpolator_list: lis
         'gamma_widths': torch.from_numpy(output[:, 2]),
         'lam_centers': torch.from_numpy(output[:, 3]),
     } for output in outputs]
-    return np.nan_to_num(np.vstack([SLE(wl_native=wl_grid, init_state_dict=state_dict).forward().detach().numpy() for state_dict in state_dicts]), nan=1)
+    return np.nan_to_num(np.vstack([SLE(wl_native=wl_grid, init_state_dict=state_dict, device='cpu').forward().detach().numpy() for state_dict in state_dicts]), nan=1)
 
 def rms_loss(wl_grid: np.ndarray, data: np.ndarray, interpolator_list: list[RegularGridInterpolator]) -> Callable:
     return lambda point: ((reconstruct1(wl_grid, point, interpolator_list) - data)**2).mean()**0.5

### 2024-04-25 experiments/10_end_to_end/10_end_to_end.py
Single vs multi-reconstruction test

@@ -65,7 +65,7 @@ def pickling_run():
     print(f'Interpolator partials dumped to pickle ({perf_counter() - start} s).')
 
 def reconstruct1(wl_grid: np.ndarray, point: np.ndarray, interpolator_list: list[RegularGridInterpolator]) -> np.ndarray:
-    output = np.vstack([r for interpolator in interpolator_list if (r := interpolator(point).squeeze())[0] != -1000])
+    output = np.vstack([r for interpolator in interpolator_list if (r := interpolator(point).squeeze())[0] > -100])
     state_dict = {
         'amplitudes': torch.from_numpy(output[:, 0]),
         'sigma_widths': torch.from_numpy(output[:, 1]),
@@ -108,5 +108,8 @@ if __name__ == '__main__':
     points = np.array([[5000, 2, 0], [6000, 3, -0.5], [7000, 4, 0], [8000, 5, -0.5], [9000, 6, 0]])
     start = perf_counter()
     x = reconstructn(wl, points, interpolator_list)
-    print(f'{len(points)} points reconstructed in {perf_counter() - start} s')
+    print(f'{len(points)} points multi-reconstructed in {perf_counter() - start} s')
+    start = perf_counter()
+    x = [reconstruct1(wl, point, interpolator_list) for point in points]
+    print(f'{len(points)} points single-reconstructed in {perf_counter() - start} s')
 

### 2024-04-25 experiments/10_end_to_end/10_end_to_end.py
Micro-optimizations

@@ -77,18 +77,14 @@ def reconstruct1(wl_grid: np.ndarray, point: np.ndarray, interpolator_list: list
 def reconstructn(wl_grid: np.ndarray, points: np.ndarray, interpolator_list: list[RegularGridInterpolator]) -> np.ndarray:
     raw_outputs = [[] for _ in points]
     for interpolator in interpolator_list:
-        results = interpolator(points).squeeze()
-        for i, r in enumerate(results):
+        for i, r in enumerate(interpolator(points).squeeze()):
             if r[0] > -100:
                 raw_outputs[i].append(r)
-    outputs = [np.vstack(raw_output) for raw_output in raw_outputs]
-    state_dicts = [{
-        'amplitudes': torch.from_numpy(output[:, 0]),
-        'sigma_widths': torch.from_numpy(output[:, 1]),
-        'gamma_widths': torch.from_numpy(output[:, 2]),
-        'lam_centers': torch.from_numpy(output[:, 3]),
-    } for output in outputs]
-    return np.nan_to_num(np.vstack([SLE(wl_native=wl_grid, init_state_dict=state_dict, device='cpu').forward().detach().numpy() for state_dict in state_dicts]), nan=1)
+    return np.nan_to_num(np.vstack([SLE(wl_native=wl_grid, init_state_dict={'amplitudes': torch.from_numpy(output[:, 0]),
+            'sigma_widths': torch.from_numpy(output[:, 1]),
+            'gamma_widths': torch.from_numpy(output[:, 2]),
+            'lam_centers': torch.from_numpy(output[:, 3])}, device='cpu').forward().detach().numpy() for output in (np.vstack(raw_output) for raw_output in raw_outputs)]), nan=1)
+
 
 def rms_loss(wl_grid: np.ndarray, data: np.ndarray, interpolator_list: list[RegularGridInterpolator]) -> Callable:
     return lambda point: ((reconstruct1(wl_grid, point, interpolator_list) - data)**2).mean()**0.5

### 2024-04-26 experiments/10_end_to_end/10_end_to_end.py
Prepare input scaling test

@@ -98,14 +98,16 @@ def inference_test():
 
 
 if __name__ == '__main__':
+    G = np.random.default_rng()
+    point_random = G.uniform([2300, 2, -0.5], [12000, 6, 0], (100, 3))
     sys.stderr = sys.stdout = open('log.txt', 'w')
     interpolator_list = load(open('interpolator_list.pkl', 'rb'))
     wl = PHOENIXSpectrum(teff=5000, logg=4, Z=0, download=True).wavelength.value
-    points = np.array([[5000, 2, 0], [6000, 3, -0.5], [7000, 4, 0], [8000, 5, -0.5], [9000, 6, 0]])
-    start = perf_counter()
-    x = reconstructn(wl, points, interpolator_list)
-    print(f'{len(points)} points multi-reconstructed in {perf_counter() - start} s')
-    start = perf_counter()
-    x = [reconstruct1(wl, point, interpolator_list) for point in points]
-    print(f'{len(points)} points single-reconstructed in {perf_counter() - start} s')
+    for n in [1, 2, 5, 10, 20, 50, 100]:
+        start = perf_counter()
+        x = reconstructn(wl, point_random[:n], interpolator_list)
+        print(f'{n} points multi-reconstructed in {perf_counter() - start} s')
+        start = perf_counter()
+        x = [reconstruct1(wl, point, interpolator_list) for point in point_random[:n]]
+        print(f'{n} points single-reconstructed in {perf_counter() - start} s')
 

### 2024-04-26 experiments/10_end_to_end/10_end_to_end.py
Run reconstructor profiling

@@ -98,12 +98,15 @@ def inference_test():
 
 
 if __name__ == '__main__':
+    #sys.stderr = sys.stdout = open('log.txt', 'w')
     G = np.random.default_rng()
     point_random = G.uniform([2300, 2, -0.5], [12000, 6, 0], (100, 3))
-    sys.stderr = sys.stdout = open('log.txt', 'w')
+    print('100 Random Points Generated')
     interpolator_list = load(open('interpolator_list.pkl', 'rb'))
+    print('PHOENIX Generator Loaded')
     wl = PHOENIXSpectrum(teff=5000, logg=4, Z=0, download=True).wavelength.value
-    for n in [1, 2, 5, 10, 20, 50, 100]:
+    print('Spectral Axis Loaded')
+    for n in [2, 5, 10, 25, 50, 75, 100]:
         start = perf_counter()
         x = reconstructn(wl, point_random[:n], interpolator_list)
         print(f'{n} points multi-reconstructed in {perf_counter() - start} s')

### 2024-05-03 paper/paper2_neo/paper.tex
Update paper
- Add abstract
- Add reconstruction profiling
- Add style thingies

@@
 \usepackage{enumitem}
+\usepackage{natbib}
+
+\graphicspath{{./images/}}
 
@@
 
+\shortauthors{Shankar \& Gully-Santiago \& Morley}
+
 \begin{abstract}
-    Text
+    The advent of machine learning methods has revolutionized 
+    numerous scientific methods, among which is spectroscopic 
+    inference. Studies can and have shown that machine learning 
+    methods prove effective in inferring what we call
+    fundamental stellar parameters. These include $T_{\mathrm{eff}}$,
+    $\log(g)$, and [Fe/H] for the purposes of this study but can 
+    also encompass other parameters such as [$\alpha$/Fe], C/O, and
+    $f_{sed}$. While current methods tend to treat spectra purely
+    from a data perspective, we propose to use a hybrid machine 
+    learning (HML) method that uses the semi-empirical approach 
+    of treating spectra as a set of spectral lines put forth by 
+    \texttt{blase}, combined with traditional linear interpolation 
+    as well as Bayesian optimization. We interpolate the four 
+    line properties ($\mu'$, $\ln(a)$, $\sigma$, $\gamma$) of 
+    128,723 unique spectral lines across a subset of the PHOENIX
+    synthetic model grid to create the PHOENIX generator, allowing
+    us to reconstruct spectra at any continuously valued point 
+    lying within the domain of the PHOENIX subset, and even reconstruct 
+    batches of spectra in pseudo-parallel fashion, with the speedup 
+    exceeding 20 times once more than 60 spectra are reconstructed. 
+    We then combine the PHOENIX generator with Bayesian optimization to 
+    efficiently infer fundamental stellar parameters. This study is 
+    however a proof of concept, as we have not yet included various
+    features that would be necessary for a full-fledged tool for 
+    spectroscopic inference, however we show that semi-empirical 
+    HML methods are a viable, albeit fledgling alternative to 
+    more traditional approaches.
 \end{abstract}
-\keywords{Text}
 
@@
 \begin{table}[h!]
-    \hspace*{-1.35cm}\begin{tabular}{|c|c|c|}
+    \centering
+    \begin{tabular}{lll}
         \hline
-        \multicolumn{3}{|c|}{\textbf{PHOENIX Grid Subset}}\\
-        \hline\hline
-        Parameter & Symbol & Range\\
+        \bf{Parameter} & \bf{Symbol} & \bf{Range}\\
         \hline
@@
 \texttt{blase}'s pseudo-Voight profile implementation uses the following:
+\begin{subequations}
 \begin{gather}
-    \tilde{\mathsf{V}}_\mu(\lambda) = \eta L(\lambda - \mu'; f) + (1 - \eta)G(\lambda - \mu'; f)\\
-    \nonumber \eta = \sum_{n=1}^3 a_n\left(\dfrac{2\gamma}{f}\right)^n \qquad a = \mqty[1.36603 \\ -0.47719 \\ 0.11116]\\
-    \nonumber f = 32\sum_{n=0}^5 b_n \left(\sqrt{2\ln(2)}\sigma\right)^{5-n}(\gamma)^n \qquad b = \mqty[1 \\ 2.69269 \\ 2.42843 \\ 4.47163 \\ 0.07842 \\ 1]
+    \tilde{\mathsf{V}}_\mu(\lambda) = a\left[\eta \mathbf{L}(\lambda - \mu'; f) + (1 - \eta)\mathbf{G}(\lambda - \mu'; f)\right]\\
+    \eta = \sum_{n=1}^3 \mathbf{u}_n\left(\dfrac{2\gamma}{f}\right)^n\\
+    f = 32\sum_{n=0}^5 \mathbf{v}_n \left(\sqrt{2\ln(2)}\sigma\right)^{5-n}(\gamma)^n\\
+    \nonumber \mathbf{u} = \mqty[1.36603 \\ -0.47719 \\ 0.11116] \quad \mathbf{v} = \mqty[1 \\ 2.69269 \\ 2.42843 \\ 4.47163 \\ 0.07842 \\ 1]
 \end{gather}
-Notice that we use $\mu'$ instead of $\mu$ in the formula. This is because
-\texttt{blase} allows the line center to shift slightly during optimization, 
-and it is this shifted center which is used in computation. The individual 
-Voigt profiles are still indexed by $\mu$ for cross-model line identification,
-explained in the next section. Once optimization was complete, the list of 
-identified lines, present in the state dictionary of \texttt{blase}'s model, 
-was saved to a \texttt{.pt} file for each PHOENIX subset grid point. The total 
-disk space these files took up is 465 MB.
+\end{subequations}
+where $\mathbf{L}$ and $\mathbf{G}$ are abbreviations for Lorentzian and 
+Gaussian profiles, respectively. Notice that we use $\mu'$ instead of $\mu$ 
+in the formula. This is because \texttt{blase} allows the line center to shift 
+slightly during optimization, and it is this shifted center which is used in 
+computation. The individual Voigt profiles are still indexed by $\mu$ for 
+cross-model line identification, explained in the next section. Once optimization 
+was complete, the list of identified lines, present in the state dictionary of 
+\texttt{blase}'s model, was saved to a \texttt{.pt} file for each PHOENIX subset 
+grid point. The total disk space these files took up is 465 MB.
 
@@
 
-\begin{figure*}[t!]
+\begin{figure*}
     \centering
-    \includegraphics[width=\textwidth]{images/figure1.png}
+    \includegraphics[width=\textwidth]{figure1}
     \caption{The number of detected spectral lines at each grid point of 
@@
 
-\begin{figure*}[t!]
+\begin{figure*}
     \centering
-    \includegraphics[width=\textwidth]{images/figure2.png}
+    \includegraphics[width=\textwidth]{figure2}
     \caption{Heatmap showing how $\ln(a)$ and $\sigma$ vary over the PHOENIX subset
@@
 
-\begin{figure*}[t!]
+\begin{figure*}
     \centering
-    \includegraphics[width=\textwidth]{images/figure3.png}
+    \includegraphics[width=\textwidth]{figure3}
     \caption{Heatmap showing how $\ln(a)$ and $\sigma$ vary over the PHOENIX 
@@
 \section{Bayesian Inference and Testing}
+\subsection{Spectral Reconstruction Time}
+The typical use case for the PHOENIX generator may be to batch 
+reconstruct spectra from an array of input coordinates. Therefore, there
+is some motivation to attempt to reduce the computational cost of this
+procedure to tractable levels. We evaluate the computational time needed
+to use the PHOENIX generator in two distinct ways. First, for a single
+input, which would be relevant in serial applications. Second, for an array
+of multiple inputs, which would be relevant in parallel applications.
+The \texttt{RegularGridInterpolator} API allows for the passing in of 
+an entire array of input coordinates to be evaluated at once, however 
+using \texttt{blase} to reconstruct the spectrum from our interpolated
+state dictionary is done serially, leading to what is actually more of 
+a psuedo-parallel evaluation. Performance results are shown in Figure 4,
+and we can see that the pseudo-parallel implementation is much faster.
+
 \subsection{Inference Algorithm}
-With the PHOENIX generator complete, we used Bayesian optimization as the
-inference algorithm, specifically the \texttt{gp\_minimize} function from
-the \texttt{scikit-optimize} library. This algorithm uses a Gaussian Process
-to model the objective function, which in this case was the RMS 
-(Root-Mean-Square) loss between the interpolated spectrum and the true 
-spectrum. The optimizer was configured to first run 30 random evaluations to 
-seed the surrogate model, then run 20 more evaluations now guided by the 
-surrogate model. This totals to 50 evaluations, which was deemed sufficient 
-for this study. We then decided to use ensemble methods, running 32 parallel
-optimizers to get a more statistically sound result. One inference run 
-takes an average of 15 minutes to complete on Triton (see Section 5 for
-specifications).
+We elected to use Bayesian optimization as the inference algorithm, specifically 
+the \texttt{gp\_minimize} function from the \texttt{scikit-optimize} library. 
+This algorithm uses a Gaussian Process to model the objective function, which in 
+this case was the RMS (Root-Mean-Square) loss between the interpolated spectrum 
+$\mathsf{M}$ and the true spectrum $\mathsf{D}$, defined as:
+\begin{gather}
+    \mathcal{L} = \ev{(\mathsf{M} - \mathsf{D})^{\circ 2}}^{1/2}
+\end{gather}
+The optimizer was configured to first run 30 random 
+evaluations to seed the surrogate model, then run 20 more evaluations now guided 
+by the surrogate model. This totals to 50 evaluations, which was deemed 
+sufficient for this study. One inference run took us an average of 15 minutes to 
+complete.
 
-\subsection{Test Data}
+\subsection{Bayesian Optimizer Performance}
 To test the performance of the inference algorithm, we used the PHOENIX subset
@@
 
-\subsection{Spectral Reconstruction Time}
-The typical use case for the PHOENIX generator may be to batch 
-reconstruct spectra from an array of input coordinates. Therefore, there
-is some motivation to attempt to reduce the computational cost of this
-procedure to tractable levels. We evaluate the computational time needed
-to use the PHOENIX generator in two distinct ways. First, for a single
-input, which would be relevant in serial applications. Second, for an array
-of multiple inputs, which would be relevant in parallel applications.
-The \texttt{RegularGridInterpolator} API allows for the passing in of 
-an entire array of input coordinates to be evaluated at once, however 
-using \texttt{blase} to reconstruct the spectrum from our interpolated
-state dictionary is done serially, leading to what is actually more of 
-a psuedo-parallel evaluation.
+\begin{figure}
+    \centering
+    \includegraphics[width=\columnwidth]{figure4}
+    \caption{Line chart showing the time taken to reconstruct varying
+    numbers of spectra using the PHOENIX generator (lower is better). 
+    We can see that the time taken per spectra for the serial 
+    implementation hovers around 15 seconds within run-to-run variance, 
+    while the pseudo-parallel implementation continually decreases in time 
+    taken per spectrum as the number of inputs increases. The speedup factor 
+    (higher is better) increases as more spectra are generated, which is 
+    also a desirable outcome.}
+\end{figure}
 
-\subsection{Bayesian Optimizer Accuracy}
-TACC
-
-\section{Discussion}
+\section{Final Considerations}
 The computer used for this study, Triton, has the following specifications:
 \begin{table}[h!]
-    \hspace*{0.6cm}\begin{tabular}{|c|c|}
+    \centering
+    \begin{tabular}{ll}
         \hline
-        \multicolumn{2}{|c|}{\textbf{Triton Specs}}\\
-        \hline\hline
         CPU & AMD EPYC 7513\\
-        Memory & 256 GB\\
+        RAM & 256 GB\\
         GPU & Nvidia A100 40GB ($\times$2)\\
@@
     for generating visualizations. The EPYC 7513 is a 32c/64t CPU with a 
-    boost clock of 3.65 GHz.}
+    boost clock of 3.65 GHz. The A100 has 6912 CUDA cores.}
 \end{table}
 
+This study combines backpropagation, interpolation, and Bayesian optimization
+to develop a HML tool for stellar parameter inference. A summative overview 
+of the end-to-end process is shown in Figure 5.
+\begin{figure*}
+    \centering
+    \includegraphics[width=\textwidth]{figure5}
+    \caption{A simple diagram showing the essence of the process used in 
+    this paper.}
+\end{figure*}
 It should be noted that this study represents a proof of concept, and that 
@@
     accounted for in the radiative transfer model, down to specific species
-    abundances.
+    abundances. So while readers may be tempted to try to identify certain 
+    spectral lines such as the one we use for our figures, this is not 
+    the point. \texttt{blase} is \textit{agnostic} to the identity of the line
+    that it is optimizing. We study these lines as \texttt{blase} sees them
+    (i.e. their four shape parameters), because for the purposes of this study, 
+    that is the only information that is useful.
     \item \textit{The Continuum Black Box}: Continuum normalization is a
@@
 end of the day, we aim to develop a tool that can analyze spectra with the 
-advent of physics-informed machine learning.
+advent of physics-informed machine learning \citep{test} \citep{test2}.
 
-\section*{Acknowledgements}
+\begin{acknowledgements}
     Text
+\end{acknowledgements}
+
+\software{Text}
+
+\begin{thebibliography}{dummy}
+    \bibitem[Test 2024()]{test} Woop
+    \bibitem[Test2 2024()]{test2} Woop woop
+\end{thebibliography}
 

### 2024-05-07 paper/paper2_neo/paper.tex
Add bibliography and citations to paper

@@
 \usepackage{enumitem}
-\usepackage{natbib}
 
 \graphicspath{{./images/}}
+\makeatletter
+\renewcommand\hyper@natlinkbreak[2]{#1}
+\makeatother
 
@@
     of treating spectra as a set of spectral lines put forth by 
-    \texttt{blase}, combined with traditional linear interpolation 
+    \texttt{blase} \citep{blase}, combined with traditional linear interpolation 
     as well as Bayesian optimization. We interpolate the four 
@@
     128,723 unique spectral lines across a subset of the PHOENIX
-    synthetic model grid to create the PHOENIX generator, allowing
+    synthetic model grid \citep{PHOENIX} to create the PHOENIX generator, allowing
     us to reconstruct spectra at any continuously valued point 
@@
 from spectra (ASPCAP in the case of APOGEE), and these pipelines are fairly
-effective. However, they all work on the core assumption that the data to be 
+effective \citep{APOGEE, ASPCAP}. However, they all work on the core assumption that the data to be 
 analyzed is a list of pixels. In addition, the pipelines that surveys 
@@
 approach by decomposing spectra into some eigenbasis and treating it as a 
-regression problem, such as \texttt{starfish}. However, \texttt{blase} takes 
+regression problem, such as \texttt{starfish} \citep{starfish}. However, \texttt{blase} takes 
 a different approach. It treats spectra not as a set of pixels or a set of 
@@
 accuracy and to increase the scope to include substellar objects with grids
-such as Sonora, for now we limit our scope to a subset of the PHOENIX grid.
+such as Sonora \citep{bobcat, cholla, diamondback, elfowl}, for now we limit our scope to a subset of the PHOENIX grid.
 
@@
     ensure a rectilinear parameter space in order to work with \texttt{scipy}'s
-    \texttt{RegularGridInterpolator}. The wavelength limits in particular 
+    \texttt{RegularGridInterpolator} \citep{scipy}. The wavelength limits in particular 
     roughly line up with that of the Habitable Zone Planet Finder (HPF) 
-    spectrograph. This subset is comprised of 1314 individual spectra.}
+    spectrograph \citep{HPF}. This subset is comprised of 1314 individual spectra.}
 \end{table}
@@
 First, the PHOENIX subset was accessed directly from the PHOENIX website
-using \texttt{gollum}'s download option. The spectra were then put through a 
+using \texttt{gollum}'s download option \citep{gollum}. The spectra were then put through a 
 three-step preprocessing pipeline.
@@
 Lorentzian width $\gamma$. The optimization used the Adam optimizer with a
-learning rate of 0.05 over 100 epochs. In addition, we limited two custom 
+learning rate of 0.05 over 100 epochs \citep{adam}. In addition, we limited two custom 
 parameters: wing cut to 6000 and prominence to 0.005. Wing cut is a parameter
@@
 \texttt{blase} uses the pseudo-Voigt approximation, which saves on 
-computational cost while remaining accurate to about 1\%. It is a 
+computational cost while remaining accurate to about 1\% \citep{pseudovoigt1, pseudovoigt2}. It is a 
 weighted average of a Gaussian and Lorentzian as opposed to a convolution.
@@
 \texttt{blase}'s model, was saved to a \texttt{.pt} file for each PHOENIX subset 
-grid point. The total disk space these files took up is 465 MB.
+grid point \citep{pytorch}. The total disk space these files took up is 465 MB.
 
@@
 We elected to use Bayesian optimization as the inference algorithm, specifically 
-the \texttt{gp\_minimize} function from the \texttt{scikit-optimize} library. 
+the \texttt{gp\_minimize} function from the \texttt{scikit-optimize} library \citep{skopt}. 
 This algorithm uses a Gaussian Process to model the objective function, which in 
@@
     evaluating those to eliminate wasteful computation.
-    \item \textit{Low-Order Interpolation}: The current design of the 
-    algorithm uses linear interpolation to construct manifolds, however this
-    could be made more sophisticated with more complex interpolation schemes.
-    However, the drawback of this is that higher orders of interpolation
-    become increasingly expensive both in terms of disk utilization, and 
-    in terms of the computational cost to generate and evaluate them. Future
-    studies should sidestep this issue by using more generalizable methods
-    that are much more efficient.
     \item \textit{Single Model Grid}: The PHOENIX grid is not the only model
@@
 end of the day, we aim to develop a tool that can analyze spectra with the 
-advent of physics-informed machine learning \citep{test} \citep{test2}.
+advent of physics-informed machine learning.
 
@@
 
-\software{Text}
+\software{
+    \texttt{altair} \citep{altair1, altair2},
+    \texttt{astropy} \citep{astropy1, astropy2, astropy3},
+    blas\'e/\texttt{blase} \citep{blase},
+    CUDA \citep{cuda},
+    \texttt{gollum} \citep{gollum},
+    \texttt{matplotlib} \citep{mpl},
+    \texttt{numpy} \citep{numpy},
+    \texttt{pandas} \citep{pandas1, pandas2},
+    Python \citep{python},
+    PyTorch/\texttt{torch} \citep{pytorch},
+    scikit-optimize/\texttt{skopt} \citep{skopt},
+    \texttt{scipy} \citep{scipy},
+    \texttt{tqdm} \citep{tqdm},
+    \texttt{vegafusion} \citep{vegafusion},
+}
 
-\begin{thebibliography}{dummy}
-    \bibitem[Test 2024()]{test} Woop
-    \bibitem[Test2 2024()]{test2} Woop woop
-\end{thebibliography}
+\bibliography{paper}
 

### 2024-05-10 paper/paper2_neo/paper.tex
Test 120-point inference

@@
 
+The spectral reconstruction process is done by iterating over the 
+PHOENIX generator, evaluating each interpolator at the given coordinates, 
+then reshaping the data into the same format that PyTorch uses for 
+state dictionaries. During the iteration, if the interpolated
+log-amplitude of the line is less than -100, the line is excluded from the 
+state dictionary. Why -100 instead of -1000? This is because near the 
+areas that we artificially populated, the manifold experiences artifacts due 
+the sudden decline. We have to anticipate that decline and cut off the 
+behavior in advance. This saves on computational time. Then, the state
+dictionary is fed into \texttt{blase}'s \texttt{SparseLinearEmulator},
+which reconstructs the spectrum by constructing a forward model based on 
+the input state dictionary. Any \texttt{nan} values are set to 1 (which we 
+can do because the spectra are all normalized), and the spectrum is 
+complete.
+
 \begin{figure*}
@@
 end of the day, we aim to develop a tool that can analyze spectra with the 
-advent of physics-informed machine learning.
+growing power of physics-informed machine learning.
 

### 2024-05-10 experiments/10_end_to_end/10_end_to_end.py

@@ -98,19 +98,13 @@ def inference_test():
 
 
 if __name__ == '__main__':
-    #sys.stderr = sys.stdout = open('log.txt', 'w')
-    G = np.random.default_rng()
-    point_random = G.uniform([2300, 2, -0.5], [12000, 6, 0], (100, 3))
-    print('100 Random Points Generated')
+    start = perf_counter()
+    sys.stderr = sys.stdout = open('log.txt', 'w')
     interpolator_list = load(open('interpolator_list.pkl', 'rb'))
-    print('PHOENIX Generator Loaded')
-    wl = PHOENIXSpectrum(teff=5000, logg=4, Z=0, download=True).wavelength.value
-    print('Spectral Axis Loaded')
-    for n in [2, 5, 10, 25, 50, 75, 100]:
-        start = perf_counter()
-        x = reconstructn(wl, point_random[:n], interpolator_list)
-        print(f'{n} points multi-reconstructed in {perf_counter() - start} s')
-        start = perf_counter()
-        x = [reconstruct1(wl, point, interpolator_list) for point in point_random[:n]]
-        print(f'{n} points single-reconstructed in {perf_counter() - start} s')
-
+    G = np.random.default_rng()
+    point_random = G.uniform([5500, 2, -0.5], [5900, 6, 0], (100, 3))
+    spec = default_clean(PHOENIXSpectrum(teff=5000, logg=4.5, Z=0, download=True))
+    reconstructions = reconstructn(spec.wavelength.value, point_random, interpolator_list)
+    rms_array = np.sqrt(((reconstructions - spec.flux.value)**2).mean(axis=1))
+    res = gp_minimize(rms_loss(spec.wavelength.value, spec.flux.value, interpolator_list), dimensions=[(5500, 5900), (2, 6), (-0.5, 0)], n_calls=120, x0=point_random, y0=rms_array, n_initial_points=0)
+    print(f'Result: {res.x} achieved in {perf_counter() - start} s')

### 2024-05-16 experiments/10_end_to_end/10_end_to_end.py
Get single test working

@@ -99,12 +99,13 @@ def inference_test():
 
 if __name__ == '__main__':
     start = perf_counter()
-    sys.stderr = sys.stdout = open('log.txt', 'w')
+    sys.stdout = open('log.txt', 'w')
     interpolator_list = load(open('interpolator_list.pkl', 'rb'))
     G = np.random.default_rng()
-    point_random = G.uniform([5500, 2, -0.5], [5900, 6, 0], (100, 3))
-    spec = default_clean(PHOENIXSpectrum(teff=5000, logg=4.5, Z=0, download=True))
+    point_random = G.uniform([5200, 2, -0.5], [6200, 6, 0], (100, 3))
+    spec = default_clean(PHOENIXSpectrum(teff=5700, logg=4.5, Z=0, download=True))
     reconstructions = reconstructn(spec.wavelength.value, point_random, interpolator_list)
+    print('Random reconstructions complete')
     rms_array = np.sqrt(((reconstructions - spec.flux.value)**2).mean(axis=1))
-    res = gp_minimize(rms_loss(spec.wavelength.value, spec.flux.value, interpolator_list), dimensions=[(5500, 5900), (2, 6), (-0.5, 0)], n_calls=120, x0=point_random, y0=rms_array, n_initial_points=0)
+    res = gp_minimize(rms_loss(spec.wavelength.value, spec.flux.value, interpolator_list), dimensions=[(5200.0, 6200), (2.0, 6), (-0.5, 0)], n_calls=20, x0=[list(array) for array in point_random], y0=list(rms_array), n_initial_points=0)
     print(f'Result: {res.x} achieved in {perf_counter() - start} s')

### 2024-05-16 experiments/10_end_to_end/10_end_to_end.py
Final testing
so hype

@@ -6,7 +6,8 @@ import torch
 from blase.emulator import SparseLinearEmulator as SLE
 from blase.optimizer import default_clean
 from collections import defaultdict
-from gollum.phoenix import PHOENIXSpectrum, PHOENIXGrid
+from gollum.phoenix import PHOENIXSpectrum
+from itertools import product
 from os import listdir
 from pickle import dump, load
 from re import split
@@ -98,14 +99,20 @@ def inference_test():
 
 
 if __name__ == '__main__':
+    sys.stderr = open('log.txt', 'w')
+    sys.stdout = open('out.txt', 'w')
     start = perf_counter()
-    sys.stdout = open('log.txt', 'w')
     interpolator_list = load(open('interpolator_list.pkl', 'rb'))
-    G = np.random.default_rng()
-    point_random = G.uniform([5200, 2, -0.5], [6200, 6, 0], (100, 3))
-    spec = default_clean(PHOENIXSpectrum(teff=5700, logg=4.5, Z=0, download=True))
-    reconstructions = reconstructn(spec.wavelength.value, point_random, interpolator_list)
-    print('Random reconstructions complete')
-    rms_array = np.sqrt(((reconstructions - spec.flux.value)**2).mean(axis=1))
-    res = gp_minimize(rms_loss(spec.wavelength.value, spec.flux.value, interpolator_list), dimensions=[(5200.0, 6200), (2.0, 6), (-0.5, 0)], n_calls=20, x0=[list(array) for array in point_random], y0=list(rms_array), n_initial_points=0)
-    print(f'Result: {res.x} achieved in {perf_counter() - start} s')
+    R = np.random.default_rng()
+    teff_samples = np.arange(3000, 11000.1, 1000)
+    logg_samples = np.arange(2, 6.1, 2)
+    Z_samples = np.arange(-0.5, 0.1, 0.5)
+    print(f'Initializations complete in {perf_counter() - start} s')
+    for T, G, Z in product(teff_samples, logg_samples, Z_samples):
+        spec = default_clean(PHOENIXSpectrum(teff=T, logg=G, Z=Z, download=True))
+        start = perf_counter()
+        point_random = R.uniform([T-500, 2, -0.5], [T+500, 6, 0], (100, 3))
+        reconstructions = reconstructn(spec.wavelength.value, point_random, interpolator_list)
[…]

### 2024-05-17 experiments/10_end_to_end/10_end_to_end.py
Complete testing

@@ -109,7 +109,7 @@ if __name__ == '__main__':
     Z_samples = np.arange(-0.5, 0.1, 0.5)
     print(f'Initializations complete in {perf_counter() - start} s')
     for T, G, Z in product(teff_samples, logg_samples, Z_samples):
-        spec = default_clean(PHOENIXSpectrum(teff=T, logg=G, Z=Z, download=True))
+        spec = default_clean(PHOENIXSpectrum(teff=int(T), logg=G, Z=Z, download=True))
         start = perf_counter()
         point_random = R.uniform([T-500, 2, -0.5], [T+500, 6, 0], (100, 3))
         reconstructions = reconstructn(spec.wavelength.value, point_random, interpolator_list)

### 2024-05-17 paper/paper2_neo/paper.tex
Start of Gully edit of the paper prose
had to direct copy-paste text from VScode into the web to circumvent annoying git permissions issues

@@
 
+%token commit by gully!
+
 \section{Introduction}
 Stellar spectra are exceedingly rich sources of information about the stars
-which create them. They contain information about the fundamental properties
-of the star such as its temperature and surface gravity, as well as
-countless absorption lines reflecting the star's chemical composition. On 
-top of that, extrinsic factors such as the radial velocity of the star or 
-the rotation of the star itself further transform spectra. When these spectra
-are observed, even the location of the observing instrument and the 
-capabilities of the instrument itself introduce further modifications. 
-In short, stellar spectra represent extremely complex, high-dimensional 
-data that has gone through multiple transformations before reaching our 
-eyes as a deceptively simple graph of flux versus wavelength.
+that produce them. Spectra encode fundamental properties such as temperature, surface gravity, and chemical composition via their
+seemingly countless absorption lines. Extrinsic properties such as the stellar radial velocity or 
+the projected equatorial rotation further transform spectra. Our observations of the spectra
+degrade the otherwise pure astronomical signals, and so the resolution the observing instrument 
+and other instrumental properties limit our ability to extract fundamental properties perfectly. 
+Stellar spectra represent extremely complex, high-dimensional data that has gone through multiple transformations before reaching our detectors.
+\authorcomment1{I like that you mention high-dimensional.  It recommend expanding on the meaning of high dimensional.  ``High'' in which sense?}
 
-Scientists want to extract as much information as possible from
-a star's spectra to get back the properties that lie at the root of it all. 
+The challenge of extracting as much information as possible from
+a stellar spectrum to get back the properties that lie at the root of it all. 
 However, this has proven to be no easy task. Spectroscopic surveys such as 

### 2024-05-17 paper/paper2_neo/paper.tex
Edit paper figure order and add testing results

@@
 accuracy and to increase the scope to include substellar objects with grids
-such as Sonora \citep{bobcat, cholla, diamondback, elfowl}, for now we limit our scope to a subset of the PHOENIX grid.
+such as Sonora \citep{bobcat, cholla, diamondback, elfowl}, for now we limit 
+our scope to a subset of the PHOENIX grid.
+
+Our study is divided into three main parts. First, we preprocess the PHOENIX
+subset and detect spectral lines with \texttt{blase}. Second, we interpolate
+the spectral line parameters, enabling us to evaluate the PHOENIX grid at 
+any point within the bounds of the subset. Lastly, we use Bayesian optimization 
+to power our inference algorithm and test its performance. An overview of this
+process is presented in Figure 1.
+\begin{figure*}
+    \centering
+    \includegraphics[width=\textwidth]{figure1}
+    \caption{Overview of the process used in this study.}
+\end{figure*}
 
@@
 There is also a second issue, and that is that spectral lines were often
-only detected in some spectra from the PHOENIX subset. In Figure 1, we show
+only detected in some spectra from the PHOENIX subset. In Figure 2, we show
 that different grid points sport differing counts of detected spectral lines.
+\begin{figure*}
+    \centering
+    \includegraphics[width=\textwidth]{figure2}
+    \caption{Number of detected spectral lines at each grid point of 
+    a slice of the PHOENIX subset at solar metallicity. We can see that
+    the number of detected lines decreases with increasing $T_{\mathrm{eff}}$.
+    Also note that from $T_{\mathrm{eff}} = 7000$ K onward, the spacing 
+    between grid points increases from 100 K to 200 K.}
+\end{figure*}
 This means that in theory, interpolation would be inaccurate in regions where 
@@
 
-\begin{figure*}
-    \centering
-    \includegraphics[width=\textwidth]{figure1}
-    \caption{The number of detected spectral lines at each grid point of 
-    a slice of the PHOENIX subset at solar metallicity. We can see that
-    the number of detected lines decreases with increasing $T_{\mathrm{eff}}$.
-    Also note that from $T_{\mathrm{eff}} = 7000$ K onward, the spacing 
-    between grid points increases from 100 K to 200 K.}
-\end{figure*}
-
 To solve this, we artificially
@@
 nullified the line when evaluated. An example of the appearance of missing 
-sections in heatmaps where a line does not appear is shown in Figure 2. 
+sections in heatmaps where a line does not appear is shown in Figure 3. 
 In total, across the entire PHOENIX subset, \texttt{blase} detected 
@@
     \centering
-    \includegraphics[width=\textwidth]{figure2}
+    \includegraphics[width=\textwidth]{figure3}
     \caption{Heatmap showing how $\ln(a)$ and $\sigma$ vary over the PHOENIX subset
@@
 interpolate a new spectrum within the domain of the PHOENIX subset,
-we call this the PHOENIX generator. In Figure 3, we show the same 
-spectral line as in Figure 2, but now supersampled using the PHOENIX
+we call this the PHOENIX generator. In Figure 4, we show the same 
+spectral line as in Figure 3, but now supersampled using the PHOENIX
 generator evaluated over the same slice.
@@
     \centering
-    \includegraphics[width=\textwidth]{figure3}
+    \includegraphics[width=\textwidth]{figure4}
     \caption{Heatmap showing how $\ln(a)$ and $\sigma$ vary over the PHOENIX 
@@
 state dictionary is done serially, leading to what is actually more of 
-a psuedo-parallel evaluation. Performance results are shown in Figure 4,
+a psuedo-parallel evaluation. Performance results are shown in Figure 5,
 and we can see that the pseudo-parallel implementation is much faster.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\columnwidth]{figure5}
+    \caption{Line chart showing the time taken to reconstruct varying
+    numbers of spectra using the PHOENIX generator (lower is better). 
+    We can see that the time taken per spectra for the serial 
+    implementation hovers around 15 seconds within run-to-run variance, 
+    while the pseudo-parallel implementation continually decreases in time 
+    taken per spectrum as the number of inputs increases. The speedup factor 
+    (higher is better) increases as more spectra are generated, which is 
+    also a desirable outcome.}
+\end{figure}
 
@@
 \end{gather}
-The optimizer was configured to first run 30 random 
+The optimizer was configured to first run 100 random 
 evaluations to seed the surrogate model, then run 20 more evaluations now guided 
-by the surrogate model. This totals to 50 evaluations, which was deemed 
-sufficient for this study. One inference run took us an average of 15 minutes to 
-complete.
+by the surrogate model. This totals to 120 evaluations, which was deemed 
+sufficient for this study. One inference run takes us on average, just under 
+7.5 minutes to complete.
 
@@
 
-\begin{figure}
-    \centering
-    \includegraphics[width=\columnwidth]{figure4}
-    \caption{Line chart showing the time taken to reconstruct varying
-    numbers of spectra using the PHOENIX generator (lower is better). 
-    We can see that the time taken per spectra for the serial 
-    implementation hovers around 15 seconds within run-to-run variance, 
-    while the pseudo-parallel implementation continually decreases in time 
-    taken per spectrum as the number of inputs increases. The speedup factor 
-    (higher is better) increases as more spectra are generated, which is 
-    also a desirable outcome.}
-\end{figure}
+We know that $T_{\mathrm{eff}}$ tends to be fairly well-constrained prior to 
+even looking at the spectra, so when testing the inference algorithm, we limited
+its search space to only include $T_{\mathrm{eff}}$ which lay within 500 K of the
+true value on either side. $\log(g)$ and [Fe/H] were allowed to vary freely. 
+The test sample consisted of 9 unique $T_{\mathrm{eff}}$ values, 3 unique $\log(g)$ 
+values, and 2 unique [Fe/H] values, totaling 54 unique spectra. $T_{\mathrm{eff}}$
+ranged from 3000 K to 11000 K in increments of 1000 K, $\log(g)$ ranged from 2 to 6
+in increments of 2, and [Fe/H] ranged from -0.5 to 0 in increments of 0.5.
+
+The results of the inference algorithm are as follows: $T_{\mathrm{eff}}$ was
+off from the true result by an average of 185 K or 2.6\%. $\log(g)$ was off 
+by an average of 0.19 or 6.8\%. [Fe/H] was off by an average of 0.12 dex (no 
+percentage, because you can't divide by zero).
 
@@
     \end{tabular}
-    \caption{This computer is owned by Caroline Morley and is available to
+    \caption{Triton is owned by Caroline Morley and is available to
     members of her research group. It was used for all computations, but not 
@@
 
-This study combines backpropagation, interpolation, and Bayesian optimization
-to develop a HML tool for stellar parameter inference. A summative overview 
-of the end-to-end process is shown in Figure 5.
-\begin{figure*}
-    \centering
-    \includegraphics[width=\textwidth]{figure5}
-    \caption{A simple diagram showing the essence of the process used in 
-    this paper.}
-\end{figure*}
 It should be noted that this study represents a proof of concept, and that 

### 2024-05-29 paper/paper2/paper.tex
Gully edits in Overleaf

--- 2024-05-17 paper/paper2_neo/paper.tex
@@
 \begin{abstract}
-    The advent of machine learning methods has revolutionized 
-    numerous scientific methods, among which is spectroscopic 
-    inference. Studies can and have shown that machine learning 
-    methods prove effective in inferring what we call
-    fundamental stellar parameters. These include $T_{\mathrm{eff}}$,
-    $\log(g)$, and [Fe/H] for the purposes of this study but can 
-    also encompass other parameters such as [$\alpha$/Fe], C/O, and
-    $f_{sed}$. While current methods tend to treat spectra purely
-    from a data perspective, we propose to use a hybrid machine 
-    learning (HML) method that uses the semi-empirical approach 
-    of treating spectra as a set of spectral lines put forth by 
-    \texttt{blase} \citep{blase}, combined with traditional linear interpolation 
-    as well as Bayesian optimization. We interpolate the four 
-    line properties ($\mu'$, $\ln(a)$, $\sigma$, $\gamma$) of 
+    The advent of machine learning methods is revolutionizing 
+    numerous scientific disciplines, with a growing number of examples in astronomical spectroscopic inference.
+Machine learning applied to spectroscopy may be expected to infer fundamental stellar parameters at greater speed, accuracy, and/or precision than conventional techniques.
+Here we introduce a new scalable ML technique to infer $T_{\mathrm{eff}}$,
+    $\log(g)$, and [Fe/H] parameters, and the extensible method holds promise for other properties such as [$\alpha$/Fe], C/O, and
+    $f_{sed}$. Our method contrasts with traditional data-driven
+    approaches by introducing hybrid machine 
+    learning (HML), a semi-empirical approach 
+    to treating spectra as differentiable spectral lines described in 
+    \texttt{blase} \citep{blase}.  
+    \authorcomment1{Say how many spectra you analyze in the ``PHOENIX subset''.  State that an outcome of this work is the precomputed training.}
+    We combine traditional linear interpolation 
+    and Bayesian optimization to create a manifold mapping between 
+    line properties ($\mu'$, $\ln(a)$, $\sigma$, $\gamma$) to fundamental paramerts.  of 
     128,723 unique spectral lines across a subset of the PHOENIX
@@
 Stellar spectra are exceedingly rich sources of information about the stars
-that produce them. Spectra encode fundamental properties such as temperature, surface gravity, and chemical composition via their
-seemingly countless absorption lines. Extrinsic properties such as the stellar radial velocity or 
-the projected equatorial rotation further transform spectra. Our observations of the spectra
-degrade the otherwise pure astronomical signals, and so the resolution the observing instrument 
-and other instrumental properties limit our ability to extract fundamental properties perfectly. 
-Stellar spectra represent extremely complex, high-dimensional data that has gone through multiple transformations before reaching our detectors.
+that produce them. Spectra encode fundamental properties such as temperature, 
+surface gravity, and chemical composition via their seemingly countless 
+absorption lines. Extrinsic properties such as the stellar radial velocity or 
+the projected equatorial rotation further transform spectra. Our observations 
+of the spectra degrade the otherwise pure astronomical signals, and so the 
+resolution the observing instrument and other instrumental properties limit 
+our ability to extract fundamental properties perfectly. Stellar spectra 
+represent extremely complex, high-dimensional data that has gone through 
+multiple transformations before reaching our detectors.
 \authorcomment1{I like that you mention high-dimensional.  It recommend expanding on the meaning of high dimensional.  ``High'' in which sense?}
 
-The challenge of extracting as much information as possible from
-a stellar spectrum to get back the properties that lie at the root of it all. 
-However, this has proven to be no easy task. Spectroscopic surveys such as 
-APOGEE develop their own in-house pipelines to extract stellar parameters
-from spectra (ASPCAP in the case of APOGEE), and these pipelines are fairly
+The challenge lies in extracting as much information as possible from
+a stellar spectrum to get back the stellar fundamental properties. 
+Spectroscopic surveys such as APOGEE develop their own in-house pipelines to extract stellar parameters
+from spectra (ASPCAP in the case of APOGEE), and these pipelines appear
 effective \citep{APOGEE, ASPCAP}. However, they all work on the core assumption that the data to be 
-analyzed is a list of pixels. In addition, the pipelines that surveys 
-develop are typically closed-source and usually limited to the scope of the 
-survey itself, causing a sort of fragmentation. This motivates the 
-development of a more universal, open-source tool that can do what spectral 
-analysis pipelines do, but in an alternative way.
+analyzed is a list of pixels. Analysis pipelines may be closed-source or limited to the scope of the survey itself, causing a sort of siloing effect among surveys. This motivates the 
+development of a more universal, instrument-agnostic open-source framework that can apply broadly to a range of spectral observations with relatively little tuning.
 
@@
 different ways. The standard practice is to treat the wavelength and flux as 
-simply two arrays and implement any from a myriad of different algorithms to 
-inference fundamental stellar parameters. Others take a more mathematical 
-approach by decomposing spectra into some eigenbasis and treating it as a 
-regression problem, such as \texttt{starfish} \citep{starfish}. However, \texttt{blase} takes 
-a different approach. It treats spectra not as a set of pixels or a set of 
-eigenbasis coefficients but as a set of spectral lines, specifically Voigt 
-profiles. There is no metric of this approach being better or worse than the 
-others mentioned so far, but it is a different paradigm and is worth 
-exploring as a semi-empirical tool. This study aims to take advantage of 
-\texttt{blase}'s ability to change a spectrum's basis from flux versus 
-wavelength into spectral line versus spectral line properties, and use that 
-to develop a robust tool for stellar parameter inference.
+simply two data arrays and bespoke algorithms tailored to a small number of well-calibrated spectral lines to obtain fundamental stellar parameters and chemical compositions.
+Other whole-spectrum fitting abstractions decompose model spectra into an eigenbasis, simplifying the data-model comparision stage nto a tractable regression, such as \texttt{starfish} \citep{starfish}.
+However, \texttt{blase} takes
+an intermediate approach \authorcomment1{cite}. 
+It treats spectra not as a set of pixels or a set of  eigenbasis coefficients but as a set of spectral lines, specifically Voigt profiles.
+Each of these approaches has tradeoffs, but one key advantage places blase in a new category: the ability to fit the non-linear spectral line parameters with automatic differentiation.
+This autodiff paradigm positions blase a promising semi-empirical tool.\authorcomment1{consider citing exojax and wobble, which also use autodiff for spectra}
+\authorcomment1{Say a bit more on what blase paper 1 limitations-- only demonstrated on a few spectra with hand-chosen fiducial models, not scaled out to the entire grid.}
 
-However, this is only a proof of concept. Here we develop a tool that 
-represents a very basic realization of the idea of line-by-line inference. 
-We make no claims as to the performance of this tool, but rather aim to show 
-that the idea is sound and able to be implemented, even if rudimentary, and 
-that with further research and development, this tool has the potential to 
-become useful for spectral inference.
+Ideally this process could be end-to-end differentiable, with all spectral inference code living inside a single autodiffable machine learning framework, like \texttt{PyTorch} or \texttt{JAX}.
+Here we separate the problem into two pieces.
+First, the scaling out of the blase method to over 1000 precomputed synthetic spectral model clones, yielding a downloadable cache of pretrained machine learning models with over xx0,000 spectral lines.
+This product alone serves as a useful asset to the community, since it can serve as a lookup for individual lines.
+Second, we explore the properties of the manifold mapping between fundamental stellar properties and the underlying spectral lines.
+Finally, we show how reconstructions of the spectra through the manifold can be adapted for end-to-end spectral inference.
+Extensions to this underlying framework have the potential to be the best in class with further research and development.
 
-We chose the PHOENIX synthetic spectral model grid as the basis for this 
-study, because it is fairly well-known. While we plan to extend this tool to 
-leverage information across multiple model grids in the future, both for 
-accuracy and to increase the scope to include substellar objects with grids
-such as Sonora \citep{bobcat, cholla, diamondback, elfowl}, for now we limit 
-our scope to a subset of the PHOENIX grid.
+We chose the widely-adopted PHOENIX synthetic spectral model grid \authorcomment{cite?} as the basis for this study.
+The framework can be straightforwardly extended to any other model grid in the future, including substellar atmosphere grids such as Sonora \citep{bobcat, cholla, diamondback, elfowl}, for now we limit our scope to a subset of the PHOENIX grid.
 
+\authorcomment1{whoops, I repeated this above.  Key idea is to put some numbers here.}
 Our study is divided into three main parts. First, we preprocess the PHOENIX
@@
 the spectral line parameters, enabling us to evaluate the PHOENIX grid at 
-any point within the bounds of the subset. Lastly, we use Bayesian optimization 
-to power our inference algorithm and test its performance. An overview of this
-process is presented in Figure 1.
+any point within the bounds of the subset. Lastly, we use Bayesian optimization to power our inference algorithm and test its performance. An overview of this process is presented in Figure 1.
 \begin{figure*}
@@
 PHOENIX synthetic spectral model grid. Instead, we focused on a subset of 
-the grid, whose parameter ranges are given in Table 1.
+the grid, with the parameter ranges given in Table 1.
+\authorcomment1{state the number of spectra in the subset if you have not already done so.  optionally state how many teff, logg, fe/h spacings.}
 
@@
 \subsection{Preprocessing with \texttt{gollum}}
-First, the PHOENIX subset was accessed directly from the PHOENIX website
-using \texttt{gollum}'s download option \citep{gollum}. The spectra were then put through a 
-three-step preprocessing pipeline.
+First, the PHOENIX subset was programmatically retrieved with \texttt{gollum} \citep{gollum}. The spectra were then put through a 
+three-step preprocessing pipeline, identical to \authorcomment1{blase paper 1}.
 \begin{enumerate}
@@
 \subsection{Line Identification with \texttt{blase}}
-The next step was to convert the PHOENIX subset into an interpretable format.
-We wanted to represent the spectra as a list of spectral lines rather than a
-list of fluxes. This was done using \texttt{blase}, which detects spectral 
-lines as Voigt profiles and tunes the profiles to mimic the original 
-PHOENIX spectrum with back propagation. Four parameters were optimized: the 
+The next step was to convert the PHOENIX subset into a physically interpretable intermediate representation: a table of spectral lines properties rather than a list of fluxes.
+This was done using \texttt{blase}, which detects spectral lines as Voigt profiles and tunes the profiles to mimic the original PHOENIX spectrum with back propagation.
+Four parameters were optimized: the 
 line center $\mu$, the log-amplitude $\ln(a)$, the Gaussian width $\sigma$, and the 
@@
 that determines the extent of the Voigt profile to evaluate, saving
-computational resources by not evaluating small numbers. Prominence sets a
+computational resources by not evaluating negligible line wings. Prominence sets a
 lower limit for the amplitude of detected lines, which also saves resources 
@@
 smaller values for prominence both increase the accuracy of using
-\texttt{blase}'s line detection and optimization at the expense of an
+\texttt{blase}'s spectral reconstruction at the expense of an
 increased computational cost. In addition, it should be mentioned that
@@
 Gaussian profiles, respectively. Notice that we use $\mu'$ instead of $\mu$ 
-in the formula. This is because \texttt{blase} allows the line center to shift 
+in the formula. This is because \texttt{blase} optionally allows the line center to shift 
 slightly during optimization, and it is this shifted center which is used in 
@@
 cross-model line identification, explained in the next section. Once optimization 
-was complete, the list of identified lines, present in the state dictionary of 
-\texttt{blase}'s model, was saved to a \texttt{.pt} file for each PHOENIX subset 
-grid point \citep{pytorch}. The total disk space these files took up is 465 MB.
+was complete, the list of identified lines, was saved to a so-called ``state dictionary'' a common representation for pre-trained machine learning models that can be stored and reopened from this static cache.
+These are stored as the \texttt{.pt} file format for each of the $N=xx,000$ PHOENIX subset
+grid point \citep{pytorch}. The total disk space these files took up is 465 MB.\authorcomment1{Restate the number of grid points for good measure, }.
 
@@
 \subsection{Cross-Model Line Identification}
-Earlier we mentioned that \texttt{blase} tuned the line centers of
+As previously mentioned, \texttt{blase} tunes the line centers of
 detected lines. This means that from one PHOENIX spectrum to the next, the
@@
 study is to interpolate the properties of each line, we needed to identify the
-presence of a particular line across the PHOENIX subset. We decided to do
+presence of a particular line across the PHOENIX subset, associating the same line with every occurrence. We decided to do
 this by using the line centers $\mu$ of the detected lines pre-optimization. 
@@
 the parameters of Voigt profiles, we can see in Equation 2 that even if the 
-interpolation method is linear, the actual result of interpolation varies 
-non-linearly.
+interpolation method is linear, a final spectral reconstruction will vary 
+non-linearly in the flux dimension.
 
-There is also a second issue, and that is that spectral lines were often
+There is also a second issue.  Spectral lines were often
 only detected in some spectra from the PHOENIX subset. In Figure 2, we show
@@
 \end{figure*}
-This means that in theory, interpolation would be inaccurate in regions where 
-a line does not appear, and also breaks in practice because \texttt{scipy}'s 
-\texttt{RegularGridInterpolator} relies on a rectilinear grid, which we would not 
-have if regions of grid points were missing.
+
+This missing-lines phenomenon can arise for three different reasons: Stellar atmospherse genuinely do not produce that line at detectable strength at the given temperature and surface gravity.
+Second, PHOENIX missed it, possibly accidentally.
+Or third, our line-finding and line-association algorithms missed it.
+Whatever the cause, these missing lines have immediate practical consequences.
+In theory, interpolation schemes may be inaccurate in regions where
+a line does not appear, or could break algorithms that demand a rectilinear grid, such as \texttt{scipy}'s \texttt{RegularGridInterpolator}.
 
@@
 populated missing grid points with log-amplitudes of -1000, which retained
-interpolator stability by not being an infinity, but also essentially 
+interpolator stability by not being infinity, but also essentially 
 nullified the line when evaluated. An example of the appearance of missing 
@@
 In total, across the entire PHOENIX subset, \texttt{blase} detected 
-128,723 individual spectral lines, meaning we have that many manifolds to 
-interpolate.
+128,723 individual spectral lines.
+
+Every one of these lines can be visualized as a manifold: a mapping between one 4-dimensional heatmap of line properties to a 3-dimensional heatmap of fundamental properties.
+All of these 128,723 manifolds have to be interpolated.
 
@@
 interpolator objects was created using linear interpolation, and these 
-interpolators were aggregated into a single list, which was when pickled 
-and written to a \texttt{.pkl} file. These interpolators generate multiple 
+interpolators were aggregated into a single list, which was then
+written to disk in the \texttt{.pkl} file storage format.  These interpolators generate multiple 
 manifolds with the following mapping:
@@
 domain of the PHOENIX subset, turning a discretely sampled PHOENIX subset 
-into a continuous one. With the given size of the PHOENIX subset, the 
+into a continuously evaluable function, sometimes called a spectral emulator or surrogate model. With the given size of the PHOENIX subset, the 
 interpolator list takes up 13.2 GB of disk space. This evaluation 
@@
 log-amplitude of the line is less than -100, the line is excluded from the 
-state dictionary. Why -100 instead of -1000? This is because near the 
-areas that we artificially populated, the manifold experiences artifacts due 
-the sudden decline. We have to anticipate that decline and cut off the 
-behavior in advance. This saves on computational time. Then, the state
-dictionary is fed into \texttt{blase}'s \texttt{SparseLinearEmulator},
+state dictionary. The rationale for -100 instead of -1000 originates from the fact that at wavelength regions near artificially populated lines, the manifold experiences artifacts due 
+a sudden decline. We have to anticipate that decline and cut off the 
+behavior in advance. This saves on computational time. 
+\authorcomment1{I do not understand this.  Maybe consider removing?}
+
+Finally, the state dictionary is fed into \texttt{blase}'s \texttt{SparseLinearEmulator},
 which reconstructs the spectrum by constructing a forward model based on 
@@
 The typical use case for the PHOENIX generator may be to batch 
-reconstruct spectra from an array of input coordinates. Therefore, there
-is some motivation to attempt to reduce the computational cost of this
+reconstruct spectra from an array of input coordinates \authorcomment1{which input coordinates-- fundamental properties or line properties?}. Therefore, there
+is some motivation to reduce the computational cost of this
 procedure to tractable levels. We evaluate the computational time needed
@@
 by the surrogate model. This totals to 120 evaluations, which was deemed 
-sufficient for this study. One inference run takes us on average, just under 
+sufficient for this study. One inference run takes on average just under 
 7.5 minutes to complete.
@@
 To test the performance of the inference algorithm, we used the PHOENIX subset
-itself. At first glance, this may seem like a useless endeavor, as the 
-PHOENIX generator has indeed memorized the PHOENIX subset, being able to
+itself. At first glance, this may seem circular, as the 
+PHOENIX generator has memorized the PHOENIX subset, being able to
 reconstruct a PHOENIX spectrum when evaluated at that grid point. However,
-that is precisely what allows us to use the PHOENIX subset as test data.
+that stategy allows us to use the PHOENIX subset as test data.
 The Bayesian optimizer's surrogate model is seeded by random continuously 
@@
 
-We know that $T_{\mathrm{eff}}$ tends to be fairly well-constrained prior to 
+We know that in typical observational settings, a coarse estimate for $T_{\mathrm{eff}}$ tends to be fairly well-constrained from ancillary data prior to 
 even looking at the spectra, so when testing the inference algorithm, we limited
@@
     \end{tabular}
-    \caption{Triton is owned by Caroline Morley and is available to
-    members of her research group. It was used for all computations, but not 
+    \caption{This machine was used for all computations, but not 
     for generating visualizations. The EPYC 7513 is a 32c/64t CPU with a 
@@
 
-It should be noted that this study represents a proof of concept, and that 
-there are numerous design considerations that could be improved upon with 
+The interpolation scheme presented here represents a proof of concept, showing that leveraging the mapping between synthetic spectral lines and their inputs can yield a semi-empirical basis for data-model comparisons.
+There are numerous design considerations that could be improved upon with
 future work. These considerations include but are not limited to the 
@@
     \item \textit{Limited PHOENIX Subset}: The PHOENIX subset used in this 
-    study was just that, a subset. The full PHOENIX grid not only expands 
+    study did not include the full PHOENIX grid, which expands 
     the [Fe/H] range to [-4.0, 1.0] dex and the $\log(g)$ range to [0, 6], 
-    but also includes the alpha element abundance parameter, which we 
+    and also includes the alpha element abundance parameter, which we 
     elected to fix at 0 for this study. In addition to the actual fundamental 

### 2024-05-30 paper/paper2/paper.tex
Paper changes and Ca I stuff

@@
 \usepackage{enumitem}
+\usepackage[all]{hypcap}
 
@@
 \begin{abstract}
-    The advent of machine learning methods is revolutionizing 
-    numerous scientific disciplines, with a growing number of examples in astronomical spectroscopic inference.
-Machine learning applied to spectroscopy may be expected to infer fundamental stellar parameters at greater speed, accuracy, and/or precision than conventional techniques.
-Here we introduce a new scalable ML technique to infer $T_{\mathrm{eff}}$,
-    $\log(g)$, and [Fe/H] parameters, and the extensible method holds promise for other properties such as [$\alpha$/Fe], C/O, and
-    $f_{sed}$. Our method contrasts with traditional data-driven
-    approaches by introducing hybrid machine 
-    learning (HML), a semi-empirical approach 
-    to treating spectra as differentiable spectral lines described in 
-    \texttt{blase} \citep{blase}.  
-    \authorcomment1{Say how many spectra you analyze in the ``PHOENIX subset''.  State that an outcome of this work is the precomputed training.}
-    We combine traditional linear interpolation 
-    and Bayesian optimization to create a manifold mapping between 
-    line properties ($\mu'$, $\ln(a)$, $\sigma$, $\gamma$) to fundamental paramerts.  of 
-    128,723 unique spectral lines across a subset of the PHOENIX
-    synthetic model grid \citep{PHOENIX} to create the PHOENIX generator, allowing
-    us to reconstruct spectra at any continuously valued point 
-    lying within the domain of the PHOENIX subset, and even reconstruct 
-    batches of spectra in pseudo-parallel fashion, with the speedup 
-    exceeding 20 times once more than 60 spectra are reconstructed. 
-    We then combine the PHOENIX generator with Bayesian optimization to 
-    efficiently infer fundamental stellar parameters. This study is 
-    however a proof of concept, as we have not yet included various
-    features that would be necessary for a full-fledged tool for 
-    spectroscopic inference, however we show that semi-empirical 
-    HML methods are a viable, albeit fledgling alternative to 
-    more traditional approaches.
+    The advent of machine learning (ML) methods has and continues to revolutionize numerous scientific disciplines, with a growing number of examples in astronomical spectroscopic inference.
+    ML methods in spectroscopy are expected to perform with greater efficacy than conventional computational techniques.
+    In this study we introduce a hybrid ML (HML) method that combines automatic differentiation (autodiff), interpolation, and Bayesian optimization to infer stellar parameters given stellar spectra.
+    The stellar parameters we study are $T_{\mathrm{eff}}$, $\log(g)$, and [Fe/H], but this method has potential for extension to other properties such as [$\alpha$/Fe], C/O, and $f_{sed}$.
+    We first use the semi-empirical approach to treating spectra as sets of tunable spectral lines introduced in \texttt{blase} \citep{blase}, contrasting with traditional data-driven approaches. 
+    \texttt{blase} is used on 1,314 spectra from the PHOENIX synthetic spectral model grid \citep{PHOENIX}, identifying 128,723 unique spectral lines.
+    For each of these lines, a manifold mapping stellar parameters to spectral line parameters is created using regular grid linear interpolation.
+    These manifolds are aggregated to to create the PHOENIX generator, sporting parallelized reconstruction of spectra at continuously valued points within the PHOENIX subset.
+    Gaussian Process minimization is then used to infer stellar parameters by minimizing the root-mean-square (RMS) loss between the input spectrum and PHOENIX generator spectra.
+    From testing, the inference error in $T_{\mathrm{eff}}$ was 185 K, $\log(g)$ was 0.19, and [Fe/H] was 0.12 dex.
+    Our products are threefold: first an archive of the state dictionaries of the \texttt{blase} clones of the PHOENIX subset, second the PHOENIX generator itself, and third, the reconstruction and inference algorithm.
+    This study represents a rudimentary proof of concept showing that semi-empirical HML is a viable, albeit fledgling alternative to traditional approaches.
 \end{abstract}
 
-%token commit by gully!
+\section{Introduction}
+Stellar spectra are exceedingly rich sources of information about the stars that produce them. 
+Spectra encode fundamental properties such as temperature, surface gravity, and chemical composition via their seemingly countless absorption lines. 
+Extrinsic properties such as the stellar radial velocity or the projected equatorial rotation further transform spectra. 
+Our observations of the spectra degrade the otherwise pure astronomical signals, and so the resolution the observing instrument and other instrumental properties limit our ability to extract fundamental properties perfectly.
+Stellar spectra represent extremely complex, data, influenced by multiple parameters, that have gone through multiple transformations before reaching our detectors.
 
-\section{Introduction}
-Stellar spectra are exceedingly rich sources of information about the stars
-that produce them. Spectra encode fundamental properties such as temperature, 
-surface gravity, and chemical composition via their seemingly countless 
-absorption lines. Extrinsic properties such as the stellar radial velocity or 
-the projected equatorial rotation further transform spectra. Our observations 
-of the spectra degrade the otherwise pure astronomical signals, and so the 
-resolution the observing instrument and other instrumental properties limit 
-our ability to extract fundamental properties perfectly. Stellar spectra 
-represent extremely complex, high-dimensional data that has gone through 
-multiple transformations before reaching our detectors.
-\authorcomment1{I like that you mention high-dimensional.  It recommend expanding on the meaning of high dimensional.  ``High'' in which sense?}
+The challenge lies in extracting as much information as possible from a stellar spectrum to get back its stellar parameters. 
+Spectroscopic surveys such as APOGEE develop their own in-house pipelines to extract stellar parameters from spectra (ASPCAP in the case of APOGEE), and these pipelines appear effective \citep{APOGEE, ASPCAP}.
+However, they all work on the core assumption that the data to be analyzed is a list of pixels. 
+Analysis pipelines may be closed-source or limited to the scope of the survey itself, causing a sort of siloing effect among surveys. 
+This motivates the development of more universal, instrument-agnostic open-source frameworks that can apply broadly to a range of spectral observations with relatively little tuning.
 
-The challenge lies in extracting as much information as possible from
-a stellar spectrum to get back the stellar fundamental properties. 
-Spectroscopic surveys such as APOGEE develop their own in-house pipelines to extract stellar parameters
-from spectra (ASPCAP in the case of APOGEE), and these pipelines appear
-effective \citep{APOGEE, ASPCAP}. However, they all work on the core assumption that the data to be 
-analyzed is a list of pixels. Analysis pipelines may be closed-source or limited to the scope of the survey itself, causing a sort of siloing effect among surveys. This motivates the 
-development of a more universal, instrument-agnostic open-source framework that can apply broadly to a range of spectral observations with relatively little tuning.
+Multiple efforts have been made in this direction, treating spectra in different ways. 
+The standard practice is to treat the wavelength and flux as simply two arrays and use bespoke algorithms tailored to a small number of well-calibrated spectral lines to obtain fundamental stellar parameters and chemical compositions.
+Other whole-spectrum fitting abstractions decompose model spectra into an eigenbasis, implementing the data-model comparison stage as a tractable regression, such as \texttt{starfish} \citep{starfish}.
+However, \texttt{blase} takes an intermediate approach \citep{blase}, treating spectra not as a set of pixels or a set of eigenbasis coefficients but as a set of spectral lines, specifically Voigt profiles.
+Each of these approaches has tradeoffs, but one key distinction places \texttt{blase} in another category defined by the ability to fit the nonlinear spectral line parameters with autodiff.
+We see autodiff used in tools such as \texttt{exojax} and \texttt{wobble} \citep{exojax, wobble}, but the recasting of spectra into sets of inherently nonlinear spectral lines positions \texttt{blase} as a unique and promising semi-empirical tool.
+The original \texttt{blase} paper demonstrated the ability to tune spectral lines with autodiff, but only on a few hand-picked fiducial models, not scaled out to an entire model grid.
 
-Multiple efforts have been made in this direction, treating spectra in 
-different ways. The standard practice is to treat the wavelength and flux as 
-simply two data arrays and bespoke algorithms tailored to a small number of well-calibrated spectral lines to obtain fundamental stellar parameters and chemical compositions.
-Other whole-spectrum fitting abstractions decompose model spectra into an eigenbasis, simplifying the data-model comparision stage nto a tractable regression, such as \texttt{starfish} \citep{starfish}.
-However, \texttt{blase} takes
-an intermediate approach \authorcomment1{cite}. 
-It treats spectra not as a set of pixels or a set of  eigenbasis coefficients but as a set of spectral lines, specifically Voigt profiles.
-Each of these approaches has tradeoffs, but one key advantage places blase in a new category: the ability to fit the non-linear spectral line parameters with automatic differentiation.
-This autodiff paradigm positions blase a promising semi-empirical tool.\authorcomment1{consider citing exojax and wobble, which also use autodiff for spectra}
-\authorcomment1{Say a bit more on what blase paper 1 limitations-- only demonstrated on a few spectra with hand-chosen fiducial models, not scaled out to the entire grid.}
-
-Ideally this process could be end-to-end differentiable, with all spectral inference code living inside a single autodiffable machine learning framework, like \texttt{PyTorch} or \texttt{JAX}.
-Here we separate the problem into two pieces.
-First, the scaling out of the blase method to over 1000 precomputed synthetic spectral model clones, yielding a downloadable cache of pretrained machine learning models with over xx0,000 spectral lines.
-This product alone serves as a useful asset to the community, since it can serve as a lookup for individual lines.
-Second, we explore the properties of the manifold mapping between fundamental stellar properties and the underlying spectral lines.
-Finally, we show how reconstructions of the spectra through the manifold can be adapted for end-to-end spectral inference.
-Extensions to this underlying framework have the potential to be the best in class with further research and development.
-
-We chose the widely-adopted PHOENIX synthetic spectral model grid \authorcomment{cite?} as the basis for this study.
-The framework can be straightforwardly extended to any other model grid in the future, including substellar atmosphere grids such as Sonora \citep{bobcat, cholla, diamondback, elfowl}, for now we limit our scope to a subset of the PHOENIX grid.
-
-\authorcomment1{whoops, I repeated this above.  Key idea is to put some numbers here.}
-Our study is divided into three main parts. First, we preprocess the PHOENIX
-subset and detect spectral lines with \texttt{blase}. Second, we interpolate
-the spectral line parameters, enabling us to evaluate the PHOENIX grid at 
-any point within the bounds of the subset. Lastly, we use Bayesian optimization to power our inference algorithm and test its performance. An overview of this process is presented in Figure 1.
+Ideally this process could be monolithic, with the end-to-end spectral inference code powered by a single autodiffable machine learning framework, like PyTorch or JAX \citep{pytorch, jax}. 
+However, here we separate the problem into three pieces. 
+First, the scaling out of the blase method to 1,314 precomputed synthetic spectral model clones, yielding a downloadable archive of pretrained machine learning models with 128,723 unique spectral lines.
+Second, the creation of manifolds mapping stellar parameters to spectral line parameters using regular grid linear interpolation. 
+Finally, we show how reconstructions of the spectra using said manifolds can be used for spectral inference.
+An overview of this process is shown in \autoref{fig:figure1}.
 \begin{figure*}
@@
     \caption{Overview of the process used in this study.}
+    \label{fig:figure1}
 \end{figure*}
+Advancements to this underlying basis have the potential to be top tier options with further research and development.
+
+We chose the widely-adopted PHOENIX synthetic spectral model grid \citep{PHOENIX} as the basis for this study.
+Our approach can be straightforwardly applied to any other model grid in the future, including substellar atmosphere grids such as Sonora \citep{bobcat, cholla, diamondback, elfowl}, but for now we limit our scope to a subset of the PHOENIX grid.
 
@@
 \subsection{The PHOENIX Subset}
-For the purposes of this study, we did not consider the full range of the
-PHOENIX synthetic spectral model grid. Instead, we focused on a subset of 
-the grid, with the parameter ranges given in Table 1.
-\authorcomment1{state the number of spectra in the subset if you have not already done so.  optionally state how many teff, logg, fe/h spacings.}
+For the purposes of this study, we did not consider the full range of the PHOENIX synthetic spectral model grid. 
+Instead, we focused on a subset of the grid, with details given in \autoref{tab:table1}.
 
@@
     \end{tabular}
-    \caption{The subset of the PHOENIX grid used in this study. These limits 
-    were imposed to reduce the computational cost of the algorithms and to 
-    ensure a rectilinear parameter space in order to work with \texttt{scipy}'s
-    \texttt{RegularGridInterpolator} \citep{scipy}. The wavelength limits in particular 
-    roughly line up with that of the Habitable Zone Planet Finder (HPF) 
-    spectrograph \citep{HPF}. This subset is comprised of 1314 individual spectra.}
+    \caption{The subset of the PHOENIX grid used in this study. 
+    These limits were imposed to reduce the computational cost of the algorithms and to ensure a rectilinear parameter space in order to work with \texttt{scipy}'s \texttt{RegularGridInterpolator} \citep{scipy}.
+    The wavelength limits in particular roughly line up with that of the Habitable Zone Planet Finder (HPF) spectrograph \citep{HPF}. 
+    This subset is comprised of 1,314 individual spectra: 73 unique $T_{\mathrm{eff}}$ values, 9 unique $\log(g)$ values, and 2 unique [Fe/H] values.}
+    \label{tab:table1}
 \end{table}
@@
 First, the PHOENIX subset was programmatically retrieved with \texttt{gollum} \citep{gollum}. The spectra were then put through a 
-three-step preprocessing pipeline, identical to \authorcomment1{blase paper 1}.
+three-step preprocessing pipeline similar to that from \citealt{blase}.
 \begin{enumerate}
-    \item \textit{Blackbody Division}: Since the $T_{\mathrm{eff}}$ of each spectrum 
-    is known, the according blackbody spectrum was divided out.
-    \item \textit{Percentile Normalization}: The spectra were normalized by dividing
-    them by their 99th percentile.
-    \item \textit{Continuum Normalization}: The spectra were further normalized by 
-    dividing them by a \nth{5} order polynomial continuum.
+    \item \textit{Blackbody Division}: Since the $T_{\mathrm{eff}}$ of each spectrum is known, the according blackbody spectrum was divided out.
+    \item \textit{Percentile Normalization}: The spectra were normalized by dividing them by their 99th percentile.
+    \item \textit{Continuum Normalization}: The spectra were further normalized by dividing them by a \nth{5} order polynomial continuum.
 \end{enumerate}
@@
 \end{gather}
-where $\mathsf{\bar{S}}$ is the preprocessed spectrum, $\mathsf{S}$ is the 
-original spectrum,  $\mathsf{B}$ is the blackbody spectrum, $\mathsf{Q}_n$ 
-is the $n^\mathrm{th}$ order polynomial continuum fit, and $P_n$ is the 
-$n^\mathrm{th}$ percentile function. Arithmetic operations between arrays 
-are assumed to be elementwise in all notation from here on out.
+where $\mathsf{\bar{S}}$ is the preprocessed spectrum, $\mathsf{S}$ is the original spectrum,  $\mathsf{B}$ is the blackbody spectrum, $\mathsf{Q}_n$ is the $n^\mathrm{th}$ order polynomial continuum fit, and $P_n$ is the $n^\mathrm{th}$ percentile function.
+Arithmetic operations between arrays are assumed to be elementwise in all notation from here on out.
 
 \subsection{Line Identification with \texttt{blase}}
-The next step was to convert the PHOENIX subset into a physically interpretable intermediate representation: a table of spectral lines properties rather than a list of fluxes.
+The next step was to convert the PHOENIX subset into a physically interpretable intermediate representation: a table of spectral line properties rather than a list of fluxes. 
 This was done using \texttt{blase}, which detects spectral lines as Voigt profiles and tunes the profiles to mimic the original PHOENIX spectrum with back propagation. 
-Four parameters were optimized: the 
-line center $\mu$, the log-amplitude $\ln(a)$, the Gaussian width $\sigma$, and the 
-Lorentzian width $\gamma$. The optimization used the Adam optimizer with a
-learning rate of 0.05 over 100 epochs \citep{adam}. In addition, we limited two custom 
-parameters: wing cut to 6000 and prominence to 0.005. Wing cut is a parameter
-that determines the extent of the Voigt profile to evaluate, saving
-computational resources by not evaluating negligible line wings. Prominence sets a
-lower limit for the amplitude of detected lines, which also saves resources 
-by disregarding shallow lines. In short, larger values for wing cut and 
-smaller values for prominence both increase the accuracy of using
-\texttt{blase}'s spectral reconstruction at the expense of an
-increased computational cost. In addition, it should be mentioned that
-\texttt{blase} uses the pseudo-Voigt approximation, which saves on 
-computational cost while remaining accurate to about 1\% \citep{pseudovoigt1, pseudovoigt2}. It is a 
-weighted average of a Gaussian and Lorentzian as opposed to a convolution.
+Four parameters were optimized: the line center $\mu$, the log-amplitude $\ln(a)$, the Gaussian width $\sigma$, and the Lorentzian width $\gamma$. 
+The optimization used the Adam optimizer with a learning rate of 0.05 over 100 epochs \citep{adam}. 
+In addition, we limited two custom parameters: wing cut to 6000 and prominence to 0.005. 
+Wing cut is a parameter that determines the extent of the Voigt profile to evaluate, saving computational resources by not evaluating negligible line wings. 
+Prominence sets a lower limit for the amplitude of detected lines, which also saves resources by disregarding shallow lines. 
+In short, smaller values for wing cut and smaller values for prominence both decrease the computational cost of \texttt{blase}'s cloning process at the expense of decreasing its accuracy slightly. 
+In addition, it should be mentioned that \texttt{blase} uses the pseudo-Voigt approximation, which saves on computational cost while remaining accurate to about 1\% \citep{pseudovoigt1, pseudovoigt2}.
+It is a weighted average of a Gaussian and Lorentzian as opposed to a convolution.
 \texttt{blase}'s pseudo-Voight profile implementation uses the following:
@@
     \nonumber \mathbf{u} = \mqty[1.36603 \\ -0.47719 \\ 0.11116] \quad \mathbf{v} = \mqty[1 \\ 2.69269 \\ 2.42843 \\ 4.47163 \\ 0.07842 \\ 1]
+    \label{eq:equation2}
 \end{gather}
 \end{subequations}
-where $\mathbf{L}$ and $\mathbf{G}$ are abbreviations for Lorentzian and 
-Gaussian profiles, respectively. Notice that we use $\mu'$ instead of $\mu$ 
-in the formula. This is because \texttt{blase} optionally allows the line center to shift 
-slightly during optimization, and it is this shifted center which is used in 
-computation. The individual Voigt profiles are still indexed by $\mu$ for 
-cross-model line identification, explained in the next section. Once optimization 
-was complete, the list of identified lines, was saved to a so-called ``state dictionary'' a common representation for pre-trained machine learning models that can be stored and reopened from this static cache.
-These are stored as the \texttt{.pt} file format for each of the $N=xx,000$ PHOENIX subset
-grid point \citep{pytorch}. The total disk space these files took up is 465 MB.\authorcomment1{Restate the number of grid points for good measure, }.
+where $\mathbf{L}$ and $\mathbf{G}$ are abbreviations for Lorentzian and Gaussian profiles, respectively. 
+Notice that we use $\mu'$ instead of $\mu$ in the formula. 
+This is because \texttt{blase} optionally allows the line center to shift slightly during optimization, and it is this shifted center which is used in computation. 
+The individual Voigt profiles are still indexed by $\mu$ for cross-model line identification, explained in the next section. 
+Once optimization was complete, the list of identified lines, was saved to a `state dictionary': a common representation for pre-trained machine learning models that can be stored to disk for reuse later.
+These are stored in the \texttt{.pt} file format for each of the 1,314 PHOENIX subset grid points. 
+The total disk space these files take up is 465 MB (382 MB when downloaded as a zip archive).
 
@@
 \subsection{Cross-Model Line Identification}
-As previously mentioned, \texttt{blase} tunes the line centers of
-detected lines. This means that from one PHOENIX spectrum to the next, the
-same line could have a slightly different line center. Since the goal of this
-study is to interpolate the properties of each line, we needed to identify the
-presence of a particular line across the PHOENIX subset, associating the same line with every occurrence. We decided to do
-this by using the line centers $\mu$ of the detected lines pre-optimization. 
-Now with each spectral line indexed by $\mu$, we had four parameters to interpolate:
-$\mu'$, $\ln(a)$, $\sigma$, and $\gamma$. Note that since we are dealing with 
-the parameters of Voigt profiles, we can see in Equation 2 that even if the 
-interpolation method is linear, a final spectral reconstruction will vary 
-non-linearly in the flux dimension.
+As previously mentioned, \texttt{blase} tunes the line centers of detected lines. 
+This means that from one PHOENIX spectrum to the next, the same line could have a slightly different line center. 
+Since the goal of this study is to interpolate the properties of each line, we needed to identify the presence of a particular line across the PHOENIX subset, associating the same line with every occurrence.
+We decided to do this by using the line centers $\mu$ of the detected lines pre-optimization. 
+Now with each spectral line indexed by $\mu$, we had four parameters to interpolate: $\mu'$, $\ln(a)$, $\sigma$, and $\gamma$. 
+Note that since we are dealing with the parameters of Voigt profiles, we can see in \autoref{eq:equation2} that even if the interpolation method is linear, a final spectral reconstruction will vary nonlinearly in flux.
 
-There is also a second issue.  Spectral lines were often
-only detected in some spectra from the PHOENIX subset. In Figure 2, we show
-that different grid points sport differing counts of detected spectral lines.
+There is also a second issue. Spectral lines were often only detected in some spectra from the PHOENIX subset. 
+In \autoref{fig:figure2}, we show that different grid points sport differing counts of detected spectral lines.
 \begin{figure*}
@@
     between grid points increases from 100 K to 200 K.}
+    \label{fig:figure2}
 \end{figure*}
@@
 Whatever the cause, these missing lines have immediate practical consequences. 
-In theory, interpolation schemes may be inaccurate in regions where
-a line does not appear, or could break algorithms that demand a rectilinear grid, such as \texttt{scipy}'s \texttt{RegularGridInterpolator}.
+Rectilinear interpolation schemes break in regions where a line does not appear; you can't interpolate a quantity that simply doesn't exist.
 
-To solve this, we artificially
-populated missing grid points with log-amplitudes of -1000, which retained
-interpolator stability by not being infinity, but also essentially 
-nullified the line when evaluated. An example of the appearance of missing 
-sections in heatmaps where a line does not appear is shown in Figure 3. 
-In total, across the entire PHOENIX subset, \texttt{blase} detected 
-128,723 individual spectral lines.
-
-Every one of these lines can be visualized as a manifold: a mapping between one 4-dimensional heatmap of line properties to a 3-dimensional heatmap of fundamental properties.
-All of these 128,723 manifolds have to be interpolated.
-
+To solve this, we artificially populated missing grid points with log-amplitudes of -1000, which retained interpolator stability while nullifying the evaluated line. 
+An example of the appearance of missing sections in heatmaps where a line does not appear is shown in \autoref{fig:figure3}. 
+In total, across the entire PHOENIX subset, \texttt{blase} detected 128,723 individual spectral lines.  
+Every one of these lines can be visualized as a manifold mapping a 3D stellar parameter vector to a 4D spectral line parameter vector, and every one must be interpolated.
 \begin{figure*}
@@
     \includegraphics[width=\textwidth]{figure3}
-    \caption{Heatmap showing how $\ln(a)$ and $\sigma$ vary over the PHOENIX subset
-    slice at solar metallicity. Notice the missing chunk in the top left
-    of the figure; \texttt{blase} did not detect a spectral line here,
-    but we have to artificially populate those points with lines that 
-    have $\ln(a) = -1000$. The change in spacing on the x-axis is owing to
-    the step of $T_\mathrm{eff}$ in the PHOENIX grid changing from 100 K to 
-    200 K after $T_\mathrm{eff} = 7000$ K.}
+    \caption{Heatmap showing how $\ln(a)$ and $\sigma$ vary over the PHOENIX subset slice at solar metallicity. 
+    Notice the missing chunk in the top left of the figure; \texttt{blase} did not detect a spectral line here, but we have to artificially populate those points with lines that have $\ln(a) = -1000$.
+    The change in spacing on the x-axis is owing to the step of $T_\mathrm{eff}$ in the PHOENIX grid changing from 100 K to 200 K after $T_\mathrm{eff} = 7000$ K.}
+    \label{fig:figure3}
 \end{figure*}
@@
 \subsection{Continuously Evaluable Manifolds}
-For each line, the inputs to the interpolator were the three input parameters $T_{\mathrm{eff}}$,
-$\log(g)$, and [Fe/H], and the output was a list of four parameters, 
-$\mu'$, $\ln(a)$, $\sigma$, and $\gamma$. For each line, one of these 
-interpolator objects was created using linear interpolation, and these 
-interpolators were aggregated into a single list, which was then
-written to disk in the \texttt{.pkl} file storage format.  These interpolators generate multiple 
-manifolds with the following mapping:
+For each line, the inputs to the interpolator were the three input parameters $T_{\mathrm{eff}}$, $\log(g)$, and [Fe/H], and the output was a list of four parameters, $\mu'$, $\ln(a)$, $\sigma$, and $\gamma$.
+For each line, one of these interpolator objects was created using linear interpolation, and these interpolators were aggregated into a single list, which was then written to disk in the \texttt{.pkl} file storage format.
+These interpolators generate multiple manifolds representing the following mapping:
 \begin{gather}
@@
 \end{gather}
-These interpolators could now be evaluated at any point lying within the 
-domain of the PHOENIX subset, turning a discretely sampled PHOENIX subset 
-into a continuously evaluable function, sometimes called a spectral emulator or surrogate model. With the given size of the PHOENIX subset, the 
-interpolator list takes up 13.2 GB of disk space. This evaluation 
-is able to reconstruct an existing PHOENIX spectrum or alternatively 
-interpolate a new spectrum within the domain of the PHOENIX subset,
-we call this the PHOENIX generator. In Figure 4, we show the same 
-spectral line as in Figure 3, but now supersampled using the PHOENIX
-generator evaluated over the same slice.
+These interpolators could now be evaluated at any point lying within the domain of the PHOENIX subset, turning a discretely sampled PHOENIX subset into a continuously evaluable function, sometimes called a spectral emulator.
+With the given size of the PHOENIX subset, the interpolator list takes up 13.2 GB of disk space. 
+This evaluation is able to reconstruct an existing PHOENIX spectrum or alternatively interpolate a new spectrum within the domain of the PHOENIX subset, so we call this the PHOENIX generator. 
+In \autoref{fig:figure4}, we show the same spectral line as in \autoref{fig:figure3}, but now supersampled using the PHOENIX generator evaluated over the same slice.
 
-The spectral reconstruction process is done by iterating over the 
-PHOENIX generator, evaluating each interpolator at the given coordinates, 
-then reshaping the data into the same format that PyTorch uses for 
-state dictionaries. During the iteration, if the interpolated
-log-amplitude of the line is less than -100, the line is excluded from the 
-state dictionary. The rationale for -100 instead of -1000 originates from the fact that at wavelength regions near artificially populated lines, the manifold experiences artifacts due 
-a sudden decline. We have to anticipate that decline and cut off the 
-behavior in advance. This saves on computational time. 
-\authorcomment1{I do not understand this.  Maybe consider removing?}
+The spectral reconstruction process is done by iterating over the PHOENIX generator, evaluating each interpolator at the given coordinates, then reshaping the data into the same format that PyTorch uses for state dictionaries.
+During the iteration, if the interpolated log-amplitude of the line is less than -100, the line is excluded from the state dictionary. 
+We do this to avoid artifacts in the manifolds due to the artifical population of log-amplitudes of -1000 where grid points were missing.
 
-Finally, the state dictionary is fed into \texttt{blase}'s \texttt{SparseLinearEmulator},
-which reconstructs the spectrum by constructing a forward model based on 
-the input state dictionary. Any \texttt{nan} values are set to 1 (which we 
-can do because the spectra are all normalized), and the spectrum is 
-complete.
+Finally, the state dictionary is fed into \texttt{blase}'s \texttt{SparseLinearEmulator}, which reconstructs the spectrum by constructing a forward model based on the input state dictionary. 
+Any \texttt{nan} values are set to 1 (which we can do because the spectra are all normalized), and the spectral reconstruction is complete.
 
@@
     \includegraphics[width=\textwidth]{figure4}
-    \caption{Heatmap showing how $\ln(a)$ and $\sigma$ vary over the PHOENIX 
-    subset slice at solar metallicity, now supersampled with the PHOENIX 
-    generator. Notice that the missing chunk in the top left still exists 
-    and does not display any artifacts, as the artificially populated points 
-    are removed after interpolation to retain the model's integrity. Also
-    see that the x-axis spacing is now uniform, as the PHOENIX generator 
-    was evaluated at constant step.}
+    \caption{Heatmap showing how $\ln(a)$ and $\sigma$ vary over the PHOENIX subset slice at solar metallicity, now supersampled with the PHOENIX generator. 
+    Notice that the missing chunk in the top left still exists and does not display any artifacts, as the artificially populated points are removed after interpolation to retain the model's integrity.
+    Also see that the x-axis spacing is now uniform, as the PHOENIX generator was evaluated at constant step.}
+    \label{fig:figure4}
 \end{figure*}
@@
 \subsection{Spectral Reconstruction Time}
-The typical use case for the PHOENIX generator may be to batch 
-reconstruct spectra from an array of input coordinates \authorcomment1{which input coordinates-- fundamental properties or line properties?}. Therefore, there
-is some motivation to reduce the computational cost of this
-procedure to tractable levels. We evaluate the computational time needed
-to use the PHOENIX generator in two distinct ways. First, for a single
-input, which would be relevant in serial applications. Second, for an array
-of multiple inputs, which would be relevant in parallel applications.
-The \texttt{RegularGridInterpolator} API allows for the passing in of 
-an entire array of input coordinates to be evaluated at once, however 
-using \texttt{blase} to reconstruct the spectrum from our interpolated
-state dictionary is done serially, leading to what is actually more of 
-a psuedo-parallel evaluation. Performance results are shown in Figure 5,
-and we can see that the pseudo-parallel implementation is much faster.
+A typical use case for the PHOENIX generator may be to batch reconstruct spectra from an array of input stellar parameters. 
+Therefore, there is some motivation to reduce the computational cost of this procedure to tractable levels. 
+We evaluate the computational time needed to use the PHOENIX generator in two distinct ways. 
+First, for a single input, which would be relevant in serial applications. 
+Second, for an array of multiple inputs, which would be relevant in parallel applications.
+The \texttt{RegularGridInterpolator} API allows for the passing in of an entire array of input coordinates to be evaluated at once. 
+However using \texttt{blase} to reconstruct the spectrum from our interpolated state dictionary is always done serially, leading to what is actually more of a psuedo-parallel evaluation, but extremely performant nonetheless.
+Performance results are shown in \autoref{fig:figure5}, and we can see that the multi-reconstruction is much faster than a series of single reconstructions.
 
@@
     \includegraphics[width=\columnwidth]{figure5}
-    \caption{Line chart showing the time taken to reconstruct varying
-    numbers of spectra using the PHOENIX generator (lower is better). 
-    We can see that the time taken per spectra for the serial 
-    implementation hovers around 15 seconds within run-to-run variance, 
-    while the pseudo-parallel implementation continually decreases in time 
-    taken per spectrum as the number of inputs increases. The speedup factor 
-    (higher is better) increases as more spectra are generated, which is 
-    also a desirable outcome.}
+    \caption{Line chart showing the time taken to reconstruct varying numbers of spectra using the PHOENIX generator (lower is better). 
+    We can see that the time taken per spectra for the serial implementation hovers around 15 seconds within run-to-run variance, while the pseudo-parallel implementation continually decreases in time taken per spectrum as the number of inputs increases.
+    The speedup factor (higher is better) increases as more spectra are generated, which is also a desirable outcome.}
+    \label{fig:figure5}
 \end{figure}
@@
 \subsection{Inference Algorithm}
-We elected to use Bayesian optimization as the inference algorithm, specifically 
-the \texttt{gp\_minimize} function from the \texttt{scikit-optimize} library \citep{skopt}. 
-This algorithm uses a Gaussian Process to model the objective function, which in 
-this case was the RMS (Root-Mean-Square) loss between the interpolated spectrum 
-$\mathsf{M}$ and the true spectrum $\mathsf{D}$, defined as:
+We elected to use Bayesian optimization as the inference algorithm, specifically the \texttt{gp\_minimize} function from the \texttt{scikit-optimize} library \citep{skopt}. 
+This algorithm uses a Gaussian Process to model the objective function, which in this case was the RMS (Root-Mean-Square) loss between the interpolated spectrum $\mathsf{M}$ and the true spectrum $\mathsf{D}$, defined as:
 \begin{gather}
@@
 \end{gather}
-The optimizer was configured to first run 100 random 
-evaluations to seed the surrogate model, then run 20 more evaluations now guided 
-by the surrogate model. This totals to 120 evaluations, which was deemed 
-sufficient for this study. One inference run takes on average just under 
-7.5 minutes to complete.
+The optimizer was configured to first run 100 random evaluations to seed the surrogate model, then run 20 more evaluations now guided by the surrogate model. 
+This totals to 120 evaluations, which was deemed sufficient for this study. 
+One inference run takes on average just under 7.5 minutes to complete.
 
 \subsection{Bayesian Optimizer Performance}
-To test the performance of the inference algorithm, we used the PHOENIX subset
-itself. At first glance, this may seem circular, as the 
-PHOENIX generator has memorized the PHOENIX subset, being able to
-reconstruct a PHOENIX spectrum when evaluated at that grid point. However,
-that stategy allows us to use the PHOENIX subset as test data.
-The Bayesian optimizer's surrogate model is seeded by random continuously 
-sampled generator evaluations within the search space, \textit{not} grid 
-points of the PHOENIX subset, meaning the surrogate model has no `memorization' 
-to speak of. If the optimizer had been an iterative grid refinement search, 
-this would not have been possible, because grid methods rely on just that, 
-the grid.
+To test the performance of the inference algorithm, we used the PHOENIX subset itself. 
+At first glance, this may seem circular, as the PHOENIX generator has memorized the PHOENIX subset, being able to reconstruct a PHOENIX spectrum when evaluated at that grid point. 
+However, that stategy allows us to use the PHOENIX subset as test data in the context of Bayesian optimization.
+The \texttt{gp\_minimize} surrogate model is seeded by random continuously sampled generator evaluations within the search space, \textit{not} grid points of the PHOENIX subset, meaning the surrogate model has no memorization to speak of.
+If the optimizer had been a grid-based strategy, this would not have been possible, because then the surrogate model would be affected by memorization.
 
-We know that in typical observational settings, a coarse estimate for $T_{\mathrm{eff}}$ tends to be fairly well-constrained from ancillary data prior to 
-even looking at the spectra, so when testing the inference algorithm, we limited
-its search space to only include $T_{\mathrm{eff}}$ which lay within 500 K of the
-true value on either side. $\log(g)$ and [Fe/H] were allowed to vary freely. 
-The test sample consisted of 9 unique $T_{\mathrm{eff}}$ values, 3 unique $\log(g)$ 
-values, and 2 unique [Fe/H] values, totaling 54 unique spectra. $T_{\mathrm{eff}}$
-ranged from 3000 K to 11000 K in increments of 1000 K, $\log(g)$ ranged from 2 to 6
-in increments of 2, and [Fe/H] ranged from -0.5 to 0 in increments of 0.5.
+We know that in typical observational settings, a coarse estimate for $T_{\mathrm{eff}}$ tends to be fairly well-constrained from ancillary data prior to even looking at the spectra.
+So when testing the inference algorithm, we limited its search space to only include $T_{\mathrm{eff}}$ which lay within 500 K of the true value on either side. 
+$\log(g)$ and [Fe/H] were allowed to vary freely. 
+The test sample consisted of 9 unique $T_{\mathrm{eff}}$ values, 3 unique $\log(g)$ values, and 2 unique [Fe/H] values, totaling 54 unique spectra in the test set. 
+$T_{\mathrm{eff}}$ ranged from 3000 K to 11000 K in increments of 1000 K, $\log(g)$ ranged from 2 to 6 in increments of 2, and [Fe/H] ranged from -0.5 to 0 in increments of 0.5.
 
-The results of the inference algorithm are as follows: $T_{\mathrm{eff}}$ was
-off from the true result by an average of 185 K or 2.6\%. $\log(g)$ was off 
-by an average of 0.19 or 6.8\%. [Fe/H] was off by an average of 0.12 dex (no 
-percentage, because you can't divide by zero).
+The results of the inference algorithm are as follows: $T_{\mathrm{eff}}$ was off from the true result by an average of 185 K or 2.6\%. 
+$\log(g)$ was off by an average of 0.19 or 6.8\%. [Fe/H] was off by an average of 0.12 dex, which is 24\% of our search range.
+From this, we can see that $T_{\mathrm{eff}}$ was the most accurately inferred parameter, followed by $\log(g)$, and then [Fe/H].
 
@@
     \end{tabular}
-    \caption{This machine was used for all computations, but not 
-    for generating visualizations. The EPYC 7513 is a 32c/64t CPU with a 
-    boost clock of 3.65 GHz. The A100 has 6912 CUDA cores.}
+    \caption{This machine was used for all computations, but not for generating visualizations. The EPYC 7513 is a 32c/64t CPU with a boost clock of 3.65 GHz. The A100 has 6912 CUDA cores.}
 \end{table}
@@
 The interpolation scheme presented here represents a proof of concept, showing that leveraging the mapping between synthetic spectral lines and their inputs can yield a semi-empirical basis for data-model comparisons.
-There are numerous design considerations that could be improved upon with
-future work. These considerations include but are not limited to the 
-following:
+There are numerous design considerations that could be improved upon with future work. 
+These include but are not limited to the following:
 \begin{enumerate}[label=-]
-    \item \textit{Limited PHOENIX Subset}: The PHOENIX subset used in this 
-    study did not include the full PHOENIX grid, which expands 
-    the [Fe/H] range to [-4.0, 1.0] dex and the $\log(g)$ range to [0, 6], 
-    and also includes the alpha element abundance parameter, which we 
-    elected to fix at 0 for this study. In addition to the actual fundamental 
-    stellar parameters, we also took a subset of the PHOENIX wavelength range, 
-    with the full [500, 55000] \AA \ wavelength range also being left to 
-    future work.
-    \item \textit{Strict Wavelength Range}: Currently, the generator only 
-    supports inference on spectra whose wavelength limits are either equal 
-    to it, or encompass that of the generator and have been truncated to 
-    match. However, when the spectrum in question has a smaller wavelength 
-    range than the generator, currently there is no functionality to truncate 
-    the generator. This would require externally indexing the generator's 
-    individual interpolators by line center position and selectively
-    evaluating those to eliminate wasteful computation.
-    \item \textit{Single Model Grid}: The PHOENIX grid is not the only model
-    grid of synthetic spectra available, and it does not apply to all types 
-    of stars. Future work would extend the reach of this study's algorithm 
-    to encompass other model grids such as the Sonora series for substellar
-    models. \texttt{blase} should be able to have an option for the user to 
-    input which model grid they would like to base the inference on, and 
-    to get even more advanced, perhaps even have the ability to intelligently
-    determine which model grid to use automatically.
-    \item \textit{Memorization vs.\ Generalization}: The current design of
-    the algorithm constructs manifolds using interpolation. This means that
-    performance is good at points close to PHOENIX subset grid points, but
-    is highly dependent on the type of interpolation used. As interpolators
-    require memorization of the data, advanced interpolation becomes extremely
-    expensive in terms of disk utilization. Future work would involve 
-    constructing manifolds using regression, which would allow for much
-    better generalization and lower disk utilization at the expense of
-    some accuracy.
-    \item \textit{Extrinsic Absence}: The current design of our algorithm
-    does not account for extrinsic parameters that modify the appearance of 
-    spectra such as rotational broadening and doppler shifting. Future work
-    would need to develop ways to tune these extrinsic parameters alongside
-    the fundamental stellar parameters.
-    \item \textit{Framework Overhead}: As this algorithm is currently more 
-    proof of concept than practical, it uses convenience functions from 
-    various libraries, which naturally introduces some level of overhead and 
-    leaves performance on the table. Future work would involve writing 
-    custom functions expressly designed for \texttt{blase}, most likely a 
-    complete rewrite of the library from the ground up.
-    \item \textit{Pseudo-Interpretability}: Our algorithm boasts interpretability
-    by considering spectral lines as the objects of interest as opposed to 
-    the rather uninterpretable flux values of other approaches. However, this
-    is only a step in the direction of interpretability. True interpretability
-    would decompose a spectrum not into a set of spectral lines, but into a
-    set of species component spectra, which requires a much more advanced 
-    understanding of different species and their behavior, as well as
-    direct access to a radiative transfer code as opposed to an off-the-shelf
-    model grid. This approach would also extend the inference from just
-    fundamental stellar parameters defined by a grid to any set of parameters
-    accounted for in the radiative transfer model, down to specific species
-    abundances. So while readers may be tempted to try to identify certain 
-    spectral lines such as the one we use for our figures, this is not 
-    the point. \texttt{blase} is \textit{agnostic} to the identity of the line
-    that it is optimizing. We study these lines as \texttt{blase} sees them
-    (i.e. their four shape parameters), because for the purposes of this study, 
-    that is the only information that is useful.
-    \item \textit{The Continuum Black Box}: Continuum normalization is a
-    process that is not yet completely understood, and is currently done as 
-    a preprocessing step with a fairly simple algorithm. Future work would
-    dive deeper into the science of continuums and develop more advanced 
-    methods that can discern continuums with greater accuracy and less
-    modeling restrictions.
-    \item \textit{One Voigt Fits All}: The current assumption of \texttt{blase}
-    is that every spectral line is a Voigt profile. This assumption is largely
-    true, but there are situations where that is simply not enough. Future
-    studies need to account for more advanced spectral line profiles and 
-    procedures to deal with phenomena such as ro-vibrational bands.
+    \item \textit{Limited PHOENIX Subset}: The PHOENIX subset used in this study did not include the full PHOENIX grid, which expands the [Fe/H] range to [-4.0, 1.0] dex and the $\log(g)$ range to [0, 6], and also includes the alpha element abundance parameter, which we elected to fix at 0 for this study.
+    In addition to the actual stellar parameters, we also took a subset of the PHOENIX wavelength range, with the full [500, 55000] \AA \ wavelength range also being left to future work.
+    \item \textit{Strict Wavelength Range}: Currently, the generator only supports inference on spectra whose wavelength limits are either equal to it, or encompass that of the generator and have been truncated to match.
+    However, when the spectrum in question has a smaller wavelength range than the generator, currently there is no functionality to truncate the generator. 
+    This would require externally indexing the generator's individual interpolators by line center position and selectively evaluating those to eliminate wasteful computation.
+    \item \textit{Single Model Grid}: The PHOENIX grid is not the only model grid of synthetic spectra available, and it does not apply to all types of stars. 
+    Future work would extend the reach of this study's algorithm to encompass other model grids such as the Sonora series of substellar models. 
+    \texttt{blase} should be able to have an option for the user to input which model grid they would like to base the inference on, and to get even more advanced, perhaps even have the ability to intelligently determine which model grid to use automatically.
+    \item \textit{Memorization vs.\ Generalization}: The current design of the algorithm constructs manifolds using interpolation. 
+    This means that performance is good at points close to PHOENIX subset grid points, but is highly dependent on the type of interpolation used. 
+    As interpolators require memorization of the data, advanced interpolation becomes extremely expensive in terms of disk utilization. 
+    Future work would involve constructing manifolds using more advanced methods, which would allow for much better generalization and lower disk utilization at the expense of some accuracy.
+    \item \textit{Extrinsic Absence}: The current design of our algorithm does not account for extrinsic parameters that modify the appearance of spectra such as rotational broadening and Doppler shifting.
+    Future work would need to develop ways to tune these extrinsic parameters alongside stellar parameters.
+    \item \textit{Framework Overhead}: As this algorithm is currently more proof of concept than practical, it uses convenience functions from various libraries, which naturally introduces some level of overhead and leaves performance on the table.
+    Future work would involve writing custom functions expressly designed for \texttt{blase}, most likely a complete rewrite of the library from the ground up.
+    \item \textit{Pseudo-Interpretability}: Our algorithm boasts interpretability by considering spectral lines as the objects of interest as opposed to the rather uninterpretable flux values of other approaches.
+    However, this is only a step in the direction of interpretability. 
+    True interpretability would decompose a spectrum not into a set of spectral lines, but into a set of species component spectra, which requires a much more advanced understanding of different species and their behavior, as well as direct access to a radiative transfer code as opposed to an off-the-shelf model grid.
+    This approach would also extend the inference from just stellar parameters defined by a grid to any set of parameters accounted for in the radiative transfer model, down to specific species abundances.
+    So while readers may be tempted to try to identify certain spectral lines seen in our figures, this is not the point. 
+    \texttt{blase} is \textit{agnostic} to the identity of the line that it is optimizing. 
+    We study these lines as \texttt{blase} sees them (i.e. their four shape parameters), because for the purposes of this study, that is the only information that is useful.
+    \item \textit{The Continuum Black Box}: Continuum normalization is a process that is not yet completely understood, and is currently done as a preprocessing step with a fairly simple algorithm. 
+    Future work would dive deeper into the science of continuums and develop more advanced methods that can discern continuums with greater accuracy and less modeling restrictions.
+    \item \textit{One Voigt Fits All}: The current assumption of \texttt{blase} is that every spectral line is a Voigt profile. 
+    This assumption is largely true, but there are situations where that is simply not enough. 
+    Future studies need to account for more advanced spectral line profiles and procedures to deal with phenomena such as ro-vibrational bands.
 \end{enumerate}
-In short, it is quite clear that \texttt{blase} is a long way from being a
-powerhouse in spectral inference, however it represents a step down a road
-not yet traveled, and the potential for future growth is immense. At the 
-end of the day, we aim to develop a tool that can analyze spectra with the 
-growing power of physics-informed machine learning.
+In short, it is quite clear that \texttt{blase} is a long way from being a powerhouse in spectral inference, however it represents a step down a road not yet traveled, and the potential for future growth is immense.
+At the end of the day, we aim to develop a tool that can analyze spectra with the growing power of physics-informed machine learning.
 

### 2024-06-10 paper/paper2/paper.tex
Updates from Overleaf

@@
     For each of these lines, a manifold mapping stellar parameters to spectral line parameters is created using regular grid linear interpolation.
-    These manifolds are aggregated to to create the PHOENIX generator, sporting parallelized reconstruction of spectra at continuously valued points within the PHOENIX subset.
+    These manifolds are aggregated to create the PHOENIX generator, sporting parallelized reconstruction of spectra at continuously valued points within the PHOENIX subset.
     Gaussian Process minimization is then used to infer stellar parameters by minimizing the root-mean-square (RMS) loss between the input spectrum and PHOENIX generator spectra.
@@
 Stellar spectra are exceedingly rich sources of information about the stars that produce them. 
-Spectra encode fundamental properties such as temperature, surface gravity, and chemical composition via their seemingly countless absorption lines. 
-Extrinsic properties such as the stellar radial velocity or the projected equatorial rotation further transform spectra. 
-Our observations of the spectra degrade the otherwise pure astronomical signals, and so the resolution the observing instrument and other instrumental properties limit our ability to extract fundamental properties perfectly.
-Stellar spectra represent extremely complex, data, influenced by multiple parameters, that have gone through multiple transformations before reaching our detectors.
+Spectra encode fundamental properties such as temperature, surface gravity, and chemical composition via their numerous absorption lines. 
+Extrinsic properties such as the stellar radial velocity or the projected equatorial rotation shift the wavelengths of lines and broaden their widths, respectively. 
+Observing the spectrum alters it again; the resolution, bandwidth, and other properties of the instrument change the fidelity at which we observe the spectrum, and can limit our ability to extract fundamental properties precisely.
+Observed stellar spectra thus represent extremely complex, data, influenced by multiple parameters, that have gone through multiple transformations before reaching our detectors.
 
 The challenge lies in extracting as much information as possible from a stellar spectrum to get back its stellar parameters. 
-Spectroscopic surveys such as APOGEE develop their own in-house pipelines to extract stellar parameters from spectra (ASPCAP in the case of APOGEE), and these pipelines appear effective \citep{APOGEE, ASPCAP}.
-However, they all work on the core assumption that the data to be analyzed is a list of pixels. 
+Spectroscopic surveys such as APOGEE use their own in-house pipelines to extract stellar parameters from spectra (ASPCAP in the case of APOGEE), and these pipelines appear effective \citep{APOGEE, ASPCAP}.
+The core assumption for the vast majority of pipelines is that the data to be analyzed is a list of pixels. 
 Analysis pipelines may be closed-source or limited to the scope of the survey itself, causing a sort of siloing effect among surveys. 
@@
 Multiple efforts have been made in this direction, treating spectra in different ways. 
-The standard practice is to treat the wavelength and flux as simply two arrays and use bespoke algorithms tailored to a small number of well-calibrated spectral lines to obtain fundamental stellar parameters and chemical compositions.
+The standard practice is to treat the wavelength and flux as simply two arrays and use bespoke algorithms tailored to a small number of well-calibrated spectral lines to obtain fundamental stellar parameters and chemical compositions \citep{pipeline1, pipeline2}.
 Other whole-spectrum fitting abstractions decompose model spectra into an eigenbasis, implementing the data-model comparison stage as a tractable regression, such as \texttt{starfish} \citep{starfish}.
-However, \texttt{blase} takes an intermediate approach \citep{blase}, treating spectra not as a set of pixels or a set of eigenbasis coefficients but as a set of spectral lines, specifically Voigt profiles.
-Each of these approaches has tradeoffs, but one key distinction places \texttt{blase} in another category defined by the ability to fit the nonlinear spectral line parameters with autodiff.
-We see autodiff used in tools such as \texttt{exojax} and \texttt{wobble} \citep{exojax, wobble}, but the recasting of spectra into sets of inherently nonlinear spectral lines positions \texttt{blase} as a unique and promising semi-empirical tool.
+
+\texttt{blase}, first presented in \citealt{blase}, takes an intermediate approach, treating spectra not as a set of pixels or a set of eigenbasis coefficients but as a set of spectral lines, specifically Voigt profiles.
+Each of these approaches has tradeoffs, but one key computational advantage of  \texttt{blase} is that it can fit the nonlinear spectral line parameters with automatic differentiation (autodiff).
+In a nutshell, autodiff is technology that tracks transformations made to data using the chain rule, even being able to differentiate control flow transformations involving if-else blocks and the like.
+With the gradient obtained from autodiff's chain rule, we can optimize model parameters by just going against the direction of the gradient (because we usually want to minimize something such as a loss function, we go in the direction of greatest decrease, which is always the negative of the gradient).
+We see autodiff used in other astrophysical software such as \texttt{exojax} and \texttt{wobble} \citep{exojax, wobble}, but the recasting of spectra into sets of inherently nonlinear spectral lines positions \texttt{blase} as a unique and promising semi-empirical tool.
 The original \texttt{blase} paper demonstrated the ability to tune spectral lines with autodiff, but only on a few hand-picked fiducial models, not scaled out to an entire model grid.
@@
 \end{figure*}
-Advancements to this underlying basis have the potential to be top tier options with further research and development.
 
@@
 For the purposes of this study, we did not consider the full range of the PHOENIX synthetic spectral model grid. 
-Instead, we focused on a subset of the grid, with details given in \autoref{tab:table1}.
+Instead, we focused on a subset of the grid, focusing on near solar metallicities and a broad range of effective temperature and surface gravity, with details given in \autoref{tab:table1}.
 
@@
         \hline
-        \bf{Parameter} & \bf{Symbol} & \bf{Range}\\
+        \bf{Parameter} & \bf{Symbol} & \bf{Interval}\\
         \hline
@@
 \subsection{Preprocessing with \texttt{gollum}}
-First, the PHOENIX subset was programmatically retrieved with \texttt{gollum} \citep{gollum}. The spectra were then put through a 
+First, the PHOENIX subset was programmatically retrieved with \texttt{gollum}, which downloaded them from the PHOENIX FTP server \citep{gollum}. The spectra were then put through a 
 three-step preprocessing pipeline similar to that from \citealt{blase}.
@@
     \item \textit{Blackbody Division}: Since the $T_{\mathrm{eff}}$ of each spectrum is known, the according blackbody spectrum was divided out.
-    \item \textit{Percentile Normalization}: The spectra were normalized by dividing them by their 99th percentile.
-    \item \textit{Continuum Normalization}: The spectra were further normalized by dividing them by a \nth{5} order polynomial continuum.
+    \item \textit{Percentile Normalization}: The spectra were normalized by dividing them by their 99th percentile in order to collapse the dynamic range of flux and only look at relative features.
+    \item \textit{Continuum Normalization}: The spectra were further normalized by dividing them by a \nth{5} order polynomial continuum fit using a peak-finding filter in order to eliminate curvature that would inhibit line modeling.
 \end{enumerate}
@@
 where $\mathsf{\bar{S}}$ is the preprocessed spectrum, $\mathsf{S}$ is the original spectrum,  $\mathsf{B}$ is the blackbody spectrum, $\mathsf{Q}_n$ is the $n^\mathrm{th}$ order polynomial continuum fit, and $P_n$ is the $n^\mathrm{th}$ percentile function.
-Arithmetic operations between arrays are assumed to be elementwise in all notation from here on out.
+Arithmetic operations between arrays are assumed to be elementwise in all following notation.
 
 \subsection{Line Identification with \texttt{blase}}
-The next step was to convert the PHOENIX subset into a physically interpretable intermediate representation: a table of spectral line properties rather than a list of fluxes. 
-This was done using \texttt{blase}, which detects spectral lines as Voigt profiles and tunes the profiles to mimic the original PHOENIX spectrum with back propagation. 
+The next step was to convert the PHOENIX subset into a physically interpretable intermediate representation: a table of spectral line properties rather than a array of fluxes. 
+We used \texttt{blase}, which models spectral lines as Voigt profiles and tunes the profiles to mimic the original PHOENIX spectrum with back propagation.
+Back propagation, put simply, is the process of moving in the autodiff gradient field to update ML model parameters (like mentioned earlier, usually against the gradient because we minimize loss functions).
 Four parameters were optimized: the line center $\mu$, the log-amplitude $\ln(a)$, the Gaussian width $\sigma$, and the Lorentzian width $\gamma$. 
@@
 Wing cut is a parameter that determines the extent of the Voigt profile to evaluate, saving computational resources by not evaluating negligible line wings. 
-Prominence sets a lower limit for the amplitude of detected lines, which also saves resources by disregarding shallow lines. 
+Prominence sets a lower limit for the amplitude of detected lines, which also saves resources by disregarding shallow lines, so in our case we disregard lines that affect the spectrum by $<$ 0.5\%.
 In short, smaller values for wing cut and smaller values for prominence both decrease the computational cost of \texttt{blase}'s cloning process at the expense of decreasing its accuracy slightly. 
-In addition, it should be mentioned that \texttt{blase} uses the pseudo-Voigt approximation, which saves on computational cost while remaining accurate to about 1\% \citep{pseudovoigt1, pseudovoigt2}.
-It is a weighted average of a Gaussian and Lorentzian as opposed to a convolution.
-\texttt{blase}'s pseudo-Voight profile implementation uses the following:
+We note that \texttt{blase} uses the pseudo-Voigt approximation, which saves on computational cost while remaining accurate to about 1\% \citep{pseudovoigt1, pseudovoigt2}. 
+The pseudo-Voigt approximation uses a weighted average of a Gaussian and Lorentzian as opposed to a convolution.
+\texttt{blase}'s pseudo-Voigt profile implementation uses the following:
 \begin{subequations}
@@
 The total disk space these files take up is 465 MB (382 MB when downloaded as a zip archive). 
+For reference, the storage space the PHOENIX subset takes up on disk is approximately 8.1 GB.
+This represents a data compression factor of around 20 just by recasting the spectrum with \texttt{blase}.
 
@@
 
-There is also a second issue. Spectral lines were often only detected in some spectra from the PHOENIX subset. 
+Spectral lines were often only detected in some spectra from the PHOENIX subset. 
 In \autoref{fig:figure2}, we show that different grid points sport differing counts of detected spectral lines.
@@
 
-This missing-lines phenomenon can arise for three different reasons: Stellar atmospherse genuinely do not produce that line at detectable strength at the given temperature and surface gravity. 
-Second, PHOENIX missed it, possibly accidentally. 
-Or third, our line-finding and line-association algorithms missed it. 
+This missing-lines phenomenon can arise for different reasons.
+Perhaps for astrophysical reasons: stellar atmospheres genuinely do not produce that line at detectable strength at the given temperature and surface gravity. 
+Or alternatively, our line-finding and line-association algorithms missed it. 
 Whatever the cause, these missing lines have immediate practical consequences. 
@@
 The \texttt{RegularGridInterpolator} API allows for the passing in of an entire array of input coordinates to be evaluated at once. 
-However using \texttt{blase} to reconstruct the spectrum from our interpolated state dictionary is always done serially, leading to what is actually more of a psuedo-parallel evaluation, but extremely performant nonetheless.
+However using \texttt{blase} to reconstruct the spectrum from our interpolated state dictionary is always done serially, leading to what is actually more of a pseudo-parallel evaluation, but extremely performant nonetheless.
 Performance results are shown in \autoref{fig:figure5}, and we can see that the multi-reconstruction is much faster than a series of single reconstructions.
@@
 The optimizer was configured to first run 100 random evaluations to seed the surrogate model, then run 20 more evaluations now guided by the surrogate model. 
-This totals to 120 evaluations, which was deemed sufficient for this study. 
+This totals to 120 evaluations, a large sample to create a fairly detailed surrogate model, then a moderately precise guided evaluation phase, which was deemed sufficient for this study. 
+Fine-tuning these numbers is possible, but simply not warranted for a proof-of concept method.
 One inference run takes on average just under 7.5 minutes to complete.
@@
 At first glance, this may seem circular, as the PHOENIX generator has memorized the PHOENIX subset, being able to reconstruct a PHOENIX spectrum when evaluated at that grid point. 
-However, that stategy allows us to use the PHOENIX subset as test data in the context of Bayesian optimization.
+However, that strategy allows us to use the PHOENIX subset as test data in the context of Bayesian optimization.
 The \texttt{gp\_minimize} surrogate model is seeded by random continuously sampled generator evaluations within the search space, \textit{not} grid points of the PHOENIX subset, meaning the surrogate model has no memorization to speak of.
@@
 
-The results of the inference algorithm are as follows: $T_{\mathrm{eff}}$ was off from the true result by an average of 185 K or 2.6\%. 
-$\log(g)$ was off by an average of 0.19 or 6.8\%. [Fe/H] was off by an average of 0.12 dex, which is 24\% of our search range.
+The results of the inference algorithm are as follows: $T_{\mathrm{eff}}$ differed from the true result by an average of 185 K or 2.6\%. 
+$\log(g)$ differed by an average of 0.19 or 6.8\%. [Fe/H] differed by an average of 0.12 dex, which is 24\% of our search range.
 From this, we can see that $T_{\mathrm{eff}}$ was the most accurately inferred parameter, followed by $\log(g)$, and then [Fe/H].
 
-\section{Final Considerations}
+\section{Conclusions}
 The computer used for this study, Triton, has the following specifications:
@@
     In addition to the actual stellar parameters, we also took a subset of the PHOENIX wavelength range, with the full [500, 55000] \AA \ wavelength range also being left to future work.
+    Users would be able to fit a greater variety of stellar spectra in many different wavelength regimes.
     \item \textit{Strict Wavelength Range}: Currently, the generator only supports inference on spectra whose wavelength limits are either equal to it, or encompass that of the generator and have been truncated to match.
@@
     This would require externally indexing the generator's individual interpolators by line center position and selectively evaluating those to eliminate wasteful computation.
+    This takes burden off the users to truncate their data to the PHOENIX generators, making use simpler.
     \item \textit{Single Model Grid}: The PHOENIX grid is not the only model grid of synthetic spectra available, and it does not apply to all types of stars. 
-    Future work would extend the reach of this study's algorithm to encompass other model grids such as the Sonora series of substellar models. 
+    Future work would extend the reach of this study's algorithm to encompass other model grids such as the Sonora series of substellar models, ATLAS \citep{atlas}, and coolTLUSTY \citep{coolTLUSTY}, reaching practitioners studying various types of stars and spectra.
     \texttt{blase} should be able to have an option for the user to input which model grid they would like to base the inference on, and to get even more advanced, perhaps even have the ability to intelligently determine which model grid to use automatically.
@@
     As interpolators require memorization of the data, advanced interpolation becomes extremely expensive in terms of disk utilization. 
-    Future work would involve constructing manifolds using more advanced methods, which would allow for much better generalization and lower disk utilization at the expense of some accuracy.
+    Future work would involve constructing manifolds using more advanced methods, which would allow for much better generalization, high speed, and lower disk utilization at the expense of some accuracy.
     \item \textit{Extrinsic Absence}: The current design of our algorithm does not account for extrinsic parameters that modify the appearance of spectra such as rotational broadening and Doppler shifting.
-    Future work would need to develop ways to tune these extrinsic parameters alongside stellar parameters.
+    Future work would need to develop ways to tune these extrinsic parameters alongside stellar parameters, enabling users to optimize these frequently-observed extrinsic parameters on top of the base stellar parameters.
     \item \textit{Framework Overhead}: As this algorithm is currently more proof of concept than practical, it uses convenience functions from various libraries, which naturally introduces some level of overhead and leaves performance on the table.
     Future work would involve writing custom functions expressly designed for \texttt{blase}, most likely a complete rewrite of the library from the ground up.
+    This has the potential to greatly increase the speed of this algorithm, depending on how much overhead is avoided with a bespoke implementation.
     \item \textit{Pseudo-Interpretability}: Our algorithm boasts interpretability by considering spectral lines as the objects of interest as opposed to the rather uninterpretable flux values of other approaches.
@@
     We study these lines as \texttt{blase} sees them (i.e. their four shape parameters), because for the purposes of this study, that is the only information that is useful.
+    Having more interpretability would let scientists actually study certain species and their spectral lines.
     \item \textit{The Continuum Black Box}: Continuum normalization is a process that is not yet completely understood, and is currently done as a preprocessing step with a fairly simple algorithm. 
     Future work would dive deeper into the science of continuums and develop more advanced methods that can discern continuums with greater accuracy and less modeling restrictions.
+    This would increase accuracy for end users.
     \item \textit{One Voigt Fits All}: The current assumption of \texttt{blase} is that every spectral line is a Voigt profile. 
@@
     Future studies need to account for more advanced spectral line profiles and procedures to deal with phenomena such as ro-vibrational bands.
+    This would increase accuracy for end users.
 \end{enumerate}
-In short, it is quite clear that \texttt{blase} is a long way from being a powerhouse in spectral inference, however it represents a step down a road not yet traveled, and the potential for future growth is immense.
-At the end of the day, we aim to develop a tool that can analyze spectra with the growing power of physics-informed machine learning.
+In this study, we have presented a proof-of-concept algorithm that interpolates a subset of the PHOENIX spectral model grid and then uses GP minimization for an inference algorithm.
+We create state dictionaries for all spectra in the PHOENIX subset, lossily distilling the spectra with a ~20 data compression factor.
+We create the PHOENIX generator, enabling anyone to create reconstructions of PHOENIX spectra with continuously valued stellar parameters.
+We introduce and test our inference algorithm, with average absolute deviations from true values of 185 K in 
+$T_\mathrm{eff}$, 0.19 in $\log(g)$, and 0.12 dex in [Fe/H].
+The methods discussed here represent a step down a road not traveled in spectral inference, and have the potential to become more advanced in the future by fully utilizing the strengths of physics-informed machine learning.
 

### 2024-06-12 experiments/10_end_to_end/10_end_to_end.py
Let's find the Ca I line

@@ -99,7 +99,7 @@ def inference_test():
 
 
 if __name__ == '__main__':
-    sys.stderr = open('log.txt', 'w')
+    """sys.stderr = open('log.txt', 'w')
     sys.stdout = open('out.txt', 'w')
     start = perf_counter()
     interpolator_list = load(open('interpolator_list.pkl', 'rb'))
@@ -115,4 +115,13 @@ if __name__ == '__main__':
         reconstructions = reconstructn(spec.wavelength.value, point_random, interpolator_list)
         rms_array = np.sqrt(((reconstructions - spec.flux.value)**2).mean(axis=1))
         res = gp_minimize(rms_loss(spec.wavelength.value, spec.flux.value, interpolator_list), dimensions=[(T-500.0, T+500), (2.0, 6), (-0.5, 0)], n_calls=20, x0=[list(array) for array in point_random], y0=list(rms_array), n_initial_points=0, n_jobs=16)
-        print(f'{(T, G, Z)} -> {res.x} achieved in {perf_counter() - start} s')
+        print(f'{(T, G, Z)} -> {res.x} achieved in {perf_counter() - start} s')"""
+    df = read_state_dicts('/home/sujays/github/blase/experiments/08_blase3D_HPC_test/emulator_states')
+    df_gp = df[['teff', 'logg', 'Z']]
+    df = df.explode(['center', 'amp', 'sigma', 'gamma', 'shift_center']).convert_dtypes(dtype_backend='numpy_nullable')
+    optimize_memory(df)
+    df_line = df.query('center <= 8545 and center <= 8544', engine='python').merge(df_gp, how='right', on=['teff', 'logg', 'Z']).fillna(-1000)
+    df.sort_values(['teff', 'logg', 'Z'], inplace=True)
+    for i, line in enumerate(df.value_counts('center').index):
+        if line <= 8545 and line >= 8544:
+            print(f'{i}: {line}')

### 2024-06-13 paper/paper2/paper.tex
Checkcheckcheckcheckcheck

@@
     \includegraphics[width=\textwidth]{figure2}
-    \caption{Number of detected spectral lines at each grid point of 
-    a slice of the PHOENIX subset at solar metallicity. We can see that
-    the number of detected lines decreases with increasing $T_{\mathrm{eff}}$.
-    Also note that from $T_{\mathrm{eff}} = 7000$ K onward, the spacing 
-    between grid points increases from 100 K to 200 K.}
+    \caption{Number of detected spectral lines at each grid point of a slice of the PHOENIX subset at solar metallicity. 
+    We can see that the number of detected lines decreases with increasing $T_{\mathrm{eff}}$.
+    Also note that from $T_{\mathrm{eff}} = 7000$ K onward, PHOENIX's sampling increment changes from 100 K to 200 K.}
     \label{fig:figure2}
@@
 To solve this, we artificially populated missing grid points with log-amplitudes of -1000, which retained interpolator stability while nullifying the evaluated line. 
-An example of the appearance of missing sections in heatmaps where a line does not appear is shown in \autoref{fig:figure3}. 
+Examples of the appearance of missing sections in heatmaps where a line does not appear are shown in \autoref{fig:figure3} and \autoref{fig:figure4}. 
 In total, across the entire PHOENIX subset, \texttt{blase} detected 128,723 individual spectral lines.  
@@
     \includegraphics[width=\textwidth]{figure3}
-    \caption{Heatmap showing how $\ln(a)$ and $\sigma$ vary over the PHOENIX subset slice at solar metallicity. 
-    Notice the missing chunk in the top left of the figure; \texttt{blase} did not detect a spectral line here, but we have to artificially populate those points with lines that have $\ln(a) = -1000$.
-    The change in spacing on the x-axis is owing to the step of $T_\mathrm{eff}$ in the PHOENIX grid changing from 100 K to 200 K after $T_\mathrm{eff} = 7000$ K.}
+    \caption{Heatmap showing how $\ln(a)$ and $\sigma$ vary over the PHOENIX subset slice at solar metallicity of an unknown spectral line. 
+    Notice the missing chunk in the top left of the figure; \texttt{blase} did not detect a spectral line here, but we have to artificially populate those points with lines that have $\ln(a) = -1000$.}
     \label{fig:figure3}
+\end{figure*}
+\begin{figure*}
+    \centering
+    \includegraphics[width=\textwidth]{figure4}
+    \caption{Heatmap showing how $\ln(a)$ and $\sigma$ vary over the PHOENIX subset slice at solar metallicity of a Helium II spectral line.
+    We can see that \texttt{blase} detects this line at only a select chunk of grid points in PHOENIX, leading to the large amount of missing data for the line.}
+    \label{fig:figure4}
 \end{figure*}
@@
 This evaluation is able to reconstruct an existing PHOENIX spectrum or alternatively interpolate a new spectrum within the domain of the PHOENIX subset, so we call this the PHOENIX generator. 
-In \autoref{fig:figure4}, we show the same spectral line as in \autoref{fig:figure3}, but now supersampled using the PHOENIX generator evaluated over the same slice.
+In \autoref{fig:figure5} and \autoref{fig:figure6}, we show the same spectral lines as in \autoref{fig:figure3} and \autoref{fig:figure4}, but now supersampled using the PHOENIX generator evaluated over the same slice.
+\begin{figure*}
+    \centering
+    \includegraphics[width=\textwidth]{figure5}
+    \caption{Heatmap showing how $\ln(a)$ and $\sigma$ of our unknown spectral line vary over the PHOENIX subset slice at solar metallicity, now supersampled with the PHOENIX generator. 
+    Notice that the missing chunk in the top left still exists and does not display any artifacts, as the artificially populated points are removed after interpolation to retain the model's integrity.
+    Also see that the x-axis spacing is now uniform, as the PHOENIX generator was evaluated at constant step.}
+    \label{fig:figure5}
+\end{figure*}
+\begin{figure*}
+    \centering
+    \includegraphics[width=\textwidth]{figure6}
+    \caption{Heatmap showing how $\ln(a)$ and $\sigma$ of our Helium II spectral line vary over the PHOENIX subset slice at solar metallicity, now supersampled with the PHOENIX generator.}
+    \label{fig:figure6}
+\end{figure*}
 
@@
 Any \texttt{nan} values are set to 1 (which we can do because the spectra are all normalized), and the spectral reconstruction is complete.
-
 \begin{figure*}
     \centering
-    \includegraphics[width=\textwidth]{figure4}
-    \caption{Heatmap showing how $\ln(a)$ and $\sigma$ vary over the PHOENIX subset slice at solar metallicity, now supersampled with the PHOENIX generator. 
-    Notice that the missing chunk in the top left still exists and does not display any artifacts, as the artificially populated points are removed after interpolation to retain the model's integrity.
-    Also see that the x-axis spacing is now uniform, as the PHOENIX generator was evaluated at constant step.}
-    \label{fig:figure4}
+    \includegraphics[width=\textwidth]{figure7}
+    \caption{Plot of an Fe I spectral line shown at the closest PHOENIX grid points to the Sun's stellar parameters, as well as a solar spectrum reconstructed with the PHOENIX generator.
+    We can see the spectral line shape be reconstructed mostly between the two native PHOENIX spectra, being closer to the 5800 K spectrum as the Sun's $T_{\mathrm{eff}}$ used here is 5772 K.}
+    \label{fig:figure7}
 \end{figure*}
+We can observe in \autoref{fig:figure7} that the reconstructed solar spectrum is not simply a pixel interpolation between the nearest grid points.
+It is interpolating the spectral line properties using hundreds of thousands of manifolds, each representing a nonlinear parameter in the shape of the spectral line.
 
@@
 However using \texttt{blase} to reconstruct the spectrum from our interpolated state dictionary is always done serially, leading to what is actually more of a pseudo-parallel evaluation, but extremely performant nonetheless.
-Performance results are shown in \autoref{fig:figure5}, and we can see that the multi-reconstruction is much faster than a series of single reconstructions.
+Performance results are shown in \autoref{fig:figure8}, and we can see that the multi-reconstruction is much faster than a series of single reconstructions.
 
-\begin{figure}
+\begin{figure}[t!]
     \centering
-    \includegraphics[width=\columnwidth]{figure5}
+    \includegraphics[width=\columnwidth]{figure8}
     \caption{Line chart showing the time taken to reconstruct varying numbers of spectra using the PHOENIX generator (lower is better). 
@@
     The speedup factor (higher is better) increases as more spectra are generated, which is also a desirable outcome.}
-    \label{fig:figure5}
+    \label{fig:figure8}
 \end{figure}
@@
     This approach would also extend the inference from just stellar parameters defined by a grid to any set of parameters accounted for in the radiative transfer model, down to specific species abundances.
-    So while readers may be tempted to try to identify certain spectral lines seen in our figures, this is not the point. 
+    So while we were able to identify the spectral lines used in our figures, it is not necessarily valuable to try to identify all 128,723 lines that we identify as unique with our algorithm. 
     \texttt{blase} is \textit{agnostic} to the identity of the line that it is optimizing. 
@@
 We create the PHOENIX generator, enabling anyone to create reconstructions of PHOENIX spectra with continuously valued stellar parameters.
-We introduce and test our inference algorithm, with average absolute deviations from true values of 185 K in 
-$T_\mathrm{eff}$, 0.19 in $\log(g)$, and 0.12 dex in [Fe/H].
+We introduce and test our inference algorithm, with average absolute deviations from true values of 185 K in $T_\mathrm{eff}$, 0.19 in $\log(g)$, and 0.12 dex in [Fe/H].
 The methods discussed here represent a step down a road not traveled in spectral inference, and have the potential to become more advanced in the future by fully utilizing the strengths of physics-informed machine learning.

### 2024-07-24 paper/paper2/paper.tex
Paper (hopefully) final edits

@@
-\documentclass[twocolumn]{aastex631}
+\documentclass[twocolumn, linenumbers]{aastex631}
 \usepackage{graphicx}
@@
 
-\graphicspath{{./images/}}
 \makeatletter
@@
 \begin{document}
-\title{Using Hybrid Machine Learning for Fundamental Stellar Parameter Inference}
+\title{A New Hybrid Machine Learning Method for Stellar Parameter Inference}
 \author[0000-0002-2290-6810]{Sujay Shankar}
-\author[0000-0002-4020-3457]{Michael Gully-Santiago}
+\correspondingauthor{Sujay Shankar}
+\email{sujays2001@gmail.com}
+\author[0000-0002-4020-3457]{Michael A. Gully-Santiago}
 \author[0000-0002-4404-0456]{Caroline V. Morley}
@@
 
-\shortauthors{Shankar \& Gully-Santiago \& Morley}
+\shorttitle{HML for Stellar Parameter Inference}
 
 \begin{abstract}
-    The advent of machine learning (ML) methods has and continues to revolutionize numerous scientific disciplines, with a growing number of examples in astronomical spectroscopic inference.
-    ML methods in spectroscopy are expected to perform with greater efficacy than conventional computational techniques.
-    In this study we introduce a hybrid ML (HML) method that combines automatic differentiation (autodiff), interpolation, and Bayesian optimization to infer stellar parameters given stellar spectra.
+    The advent of machine learning (ML) methods has and continues to revolutionize numerous scientific disciplines, with a growing number of examples in astronomical spectroscopic inference, as ML is expected to be more powerful than traditional techniques.
+    Here we introduce a hybrid ML (HML) method that combines automatic differentiation (autodiff), interpolation, and Bayesian optimization to infer stellar parameters given stellar spectra.
     The stellar parameters we study are $T_{\mathrm{eff}}$, $\log(g)$, and [Fe/H], but this method has potential for extension to other properties such as [$\alpha$/Fe], C/O, and $f_{sed}$.
-    We first use the semi-empirical approach to treating spectra as sets of tunable spectral lines introduced in \texttt{blase} \citep{blase}, contrasting with traditional data-driven approaches. 
-    \texttt{blase} is used on 1,314 spectra from the PHOENIX synthetic spectral model grid \citep{PHOENIX}, identifying 128,723 unique spectral lines.
-    For each of these lines, a manifold mapping stellar parameters to spectral line parameters is created using regular grid linear interpolation.
-    These manifolds are aggregated to create the PHOENIX generator, sporting parallelized reconstruction of spectra at continuously valued points within the PHOENIX subset.
+    We first use \texttt{blase}'s \citep{blase} nontraditional semi-empirical approach to treating spectra as sets of tunable spectral lines. 
+    \texttt{blase} is used on 1,314 spectra from the PHOENIX synthetic spectral model grid \citep{PHOENIX}, identifying 128,723 spectral lines.
+    For each of these lines, we continuously map stellar parameters to spectral line parameters using regular grid linear interpolation.
+    These manifolds are aggregated to create the PHOENIX generator, sporting parallelized reconstruction of spectra given continuously valued stellar parameters.
     Gaussian Process minimization is then used to infer stellar parameters by minimizing the root-mean-square (RMS) loss between the input spectrum and PHOENIX generator spectra.
     From testing, the inference error in $T_{\mathrm{eff}}$ was 185 K, $\log(g)$ was 0.19, and [Fe/H] was 0.12 dex.
-    Our products are threefold: first an archive of the state dictionaries of the \texttt{blase} clones of the PHOENIX subset, second the PHOENIX generator itself, and third, the reconstruction and inference algorithm.
-    This study represents a rudimentary proof of concept showing that semi-empirical HML is a viable, albeit fledgling alternative to traditional approaches.
+    Our products are an archive of the state dictionaries of the \texttt{blase} clones of the PHOENIX subset, as well as the spectral reconstruction and inference algorithms themselves.
+    This study is a proof of concept showing that semi-empirical HML is a viable alternative to traditional approaches.
 \end{abstract}
@@
 
-The challenge lies in extracting as much information as possible from a stellar spectrum to get back its stellar parameters. 
+Much of modern astronomical spectroscopy would benefit from an easy-to-use, fast, accurate, interpretable, and precise spectral model. 
+The determination of fundamental properties, the creation of Extreme Precision Radial Velocity (EPRV) templates, and the application of composite spectral fitting could all leverage such an innovation.
+This ambitious aim may be broadly referred to as a ``foundational spectroscopy model for astrophysics'', in reference to the same category of models used for large language models (LLMs) in artificial intelligence (AI).
+There are many challenges with creating such a foundational spectral model for astrophysics. 
+First, we know our physical inputs into the spectral modeling process are imperfect. 
+Second, simplifying assumptions in stellar (and substellar) atmospheres are necessarily inexact. 
+Ultimately computational costs have prevented the training of such models.  
+Most prevailing solutions have therefore had to choose either a model driven approach, in which the precomputed models are taken as gospel, or a data driven/empirical approach, in which our knowledge of stellar physics is ignored or treated phenomenologically, depending on the application.
+
 Spectroscopic surveys such as APOGEE use their own in-house pipelines to extract stellar parameters from spectra (ASPCAP in the case of APOGEE), and these pipelines appear effective \citep{APOGEE, ASPCAP}.
@@
 
-\texttt{blase}, first presented in \citealt{blase}, takes an intermediate approach, treating spectra not as a set of pixels or a set of eigenbasis coefficients but as a set of spectral lines, specifically Voigt profiles.
-Each of these approaches has tradeoffs, but one key computational advantage of  \texttt{blase} is that it can fit the nonlinear spectral line parameters with automatic differentiation (autodiff).
-In a nutshell, autodiff is technology that tracks transformations made to data using the chain rule, even being able to differentiate control flow transformations involving if-else blocks and the like.
+Ideally, we want a system that can self-consistently learn, a genuine AI foundational spectral model. 
+Such a system would enable the assignment of accurate stellar parameters, and could yield re-usable interpretable spectral models.
+\texttt{blase}, first presented in \citealt{blase}, took an important, albeit limited, step in this direction, treating spectra not as a set of pixels or a set of eigenbasis coefficients but as a set of interpretable and traceable spectral lines, specifically Voigt profiles.
+Each of these approaches has tradeoffs, but one key scientific advantage of blase comes from its ability to adapt to new information, while preserving some adherence to physics-based models. 
+This intelligent capability stems from its ability to fit a theoretically unlimited number of nonlinear spectral line parameters with automatic differentiation (autodiff).
+Autodiff is a technology that tracks transformations made to data using the chain rule, even being able to differentiate control flow transformations involving if-else blocks, for example.
 With the gradient obtained from autodiff's chain rule, we can optimize model parameters by just going against the direction of the gradient (because we usually want to minimize something such as a loss function, we go in the direction of greatest decrease, which is always the negative of the gradient).
-We see autodiff used in other astrophysical software such as \texttt{exojax} and \texttt{wobble} \citep{exojax, wobble}, but the recasting of spectra into sets of inherently nonlinear spectral lines positions \texttt{blase} as a unique and promising semi-empirical tool.
-The original \texttt{blase} paper demonstrated the ability to tune spectral lines with autodiff, but only on a few hand-picked fiducial models, not scaled out to an entire model grid.
+Autodiff has been used successfully in other astrophysical contexts such as \texttt{exojax} and \texttt{wobble} \citep{exojax, wobble}, but the recasting of spectra into sets of inherently nonlinear spectral lines positions \texttt{blase} as a unique and promising semi-empirical tool.
+The original \texttt{blase} paper demonstrated the ability to tune spectral lines with autodiff, but restricted to a pre-selected static synthetic model.
 
-Ideally this process could be monolithic, with the end-to-end spectral inference code powered by a single autodiffable machine learning framework, like PyTorch or JAX \citep{pytorch, jax}. 
-However, here we separate the problem into three pieces. 
-First, the scaling out of the blase method to 1,314 precomputed synthetic spectral model clones, yielding a downloadable archive of pretrained machine learning models with 128,723 unique spectral lines.
-Second, the creation of manifolds mapping stellar parameters to spectral line parameters using regular grid linear interpolation. 
-Finally, we show how reconstructions of the spectra using said manifolds can be used for spectral inference.
+Here we introduce the next logical step in the sequence of expanding interpretable spectroscopic machine learning from operating on a single grid point and towards an entire 3D grid of precomputed spectra.
+We rebrand this augmentation as ``blase3D''.
+
+Ideally this process would be monolithic, with the end-to-end spectral inference code powered by a single autodiffable machine learning framework, like PyTorch or JAX \citep{pytorch, jax}. 
+However, here we separate the problem into three pieces, only the first of which is currently differentiable.
+First, we scale out of the blase method to 1,314 precomputed synthetic spectral model clones, yielding a downloadable archive of pretrained machine learning models with 128,723 unique spectral lines. 
+Second, we fit manifolds mapping stellar parameters to uniformly-derived spectral line parameters using regular grid linear interpolation. 
+Finally, we show how reconstructions of the spectra using said manifolds can be used for inferring stellar fundamental properties from spectra. 
+This final step resembles the aims of ``atmospheric retrievals'', but in principle should be computationally much faster, more accurate, adaptive, and reusable.
 An overview of this process is shown in \autoref{fig:figure1}.
+
 \begin{figure*}
@@
 Prominence sets a lower limit for the amplitude of detected lines, which also saves resources by disregarding shallow lines, so in our case we disregard lines that affect the spectrum by $<$ 0.5\%.
-In short, smaller values for wing cut and smaller values for prominence both decrease the computational cost of \texttt{blase}'s cloning process at the expense of decreasing its accuracy slightly. 
+In short, our choices for wing cut and prominence decrease the computational cost of \texttt{blase}'s cloning process at the expense of decreasing its accuracy slightly. 
 We note that \texttt{blase} uses the pseudo-Voigt approximation, which saves on computational cost while remaining accurate to about 1\% \citep{pseudovoigt1, pseudovoigt2}. 
@@
     \caption{Heatmap showing how $\ln(a)$ and $\sigma$ vary over the PHOENIX subset slice at solar metallicity of an unknown spectral line. 
-    Notice the missing chunk in the top left of the figure; \texttt{blase} did not detect a spectral line here, but we have to artificially populate those points with lines that have $\ln(a) = -1000$.}
+    Notice the missing chunk in the top left of the figure; \texttt{blase} did not detect a spectral line here, but we have to artificially populate those points with lines that have $\ln(a) = -1000$.
+    This and all spectral lines shown in this paper were identified using the NIST spectral line database \citep{NIST}.}
     \label{fig:figure3}
@@
 
-\section{Conclusions}
-The computer used for this study, Triton, has the following specifications:
-\begin{table}[h!]
-    \centering
-    \begin{tabular}{ll}
-        \hline
-        CPU & AMD EPYC 7513\\
-        RAM & 256 GB\\
-        GPU & Nvidia A100 40GB ($\times$2)\\
-        \hline
-    \end{tabular}
-    \caption{This machine was used for all computations, but not for generating visualizations. The EPYC 7513 is a 32c/64t CPU with a boost clock of 3.65 GHz. The A100 has 6912 CUDA cores.}
-\end{table}
+\section{Discussion}
+\subsection{Scientific Applications}
+This study's paradigm of spectral inference can enable scientists to adopt self-consistent model grids and analyze their spectral line behavior in an interpretable fashion, which has been fairly uncommon practice thus far.
+Notably, this system tracks the shifting of spectral lines as a function of stellar parameters, which traditionally has been uncommon for algorithms to assess systematically. 
+We balance rigidity and flexibility where we want them while maintaining interpretability.
+
+To summarize, we bring to scientists the ability to understand exactly what spectral properties we consider mathematically and how we expect those properties to behave on an incredibly detailed level (the 4 line parameters of all spectral lines and their corresponding interpolating manifolds), and introduce the concept of postulating one precomputed model grid as the ground truth to base all further analysis on, converting the grid into a generator that can then interface with our inference algorithm.
+
+\subsection{Technical Considerations}
+To reiterate, the manifold fitting steps are not end-to-end autodifferentiable.  
+As seen in \autoref{fig:figure1}, these steps rely on \texttt{scipy}, which is not equipped with autodiff.  
+Without autodifferentiability, the ability to ``machine learn'' is significantly reduced compared to a hypothetical monolithic JAX or PyTorch system.  
+We see three reasons for for developing a non-autodifferentiable system.
+First, the familiar \texttt{scipy}-based system will serve as an easy entry point for most practitioners who are unfamiliar with autodiff, and may still benefit from and modify the code without expert ML knowledge.
+Second, this non-autodiff version serves as a pathfinder and benchmark against which an inevitable autodifferentiable version may be compared. 
+Finally, the inventory of familiar interpolation algorithms have not yet been ported to PyTorch or JAX, since machine learning or Gaussian Process fitting schemes are generally preferred within the ML community.
 
@@
 \end{enumerate}
+The computer used for this study, Triton, has the following specifications:
+\begin{table}[h!]
+    \centering
+    \begin{tabular}{ll}
+        \hline
+        CPU & AMD EPYC 7513\\
+        RAM & 256 GB\\
+        GPU & Nvidia A100 40GB ($\times$2)\\
+        \hline
+    \end{tabular}
+    \caption{This machine was used for all computations, but not for generating visualizations. The EPYC 7513 is a 32c/64t CPU with a boost clock of 3.65 GHz. The A100 has 6912 CUDA cores.}
+\end{table}
+
+\section{Conclusion}
 In this study, we have presented a proof-of-concept algorithm that interpolates a subset of the PHOENIX spectral model grid and then uses GP minimization for an inference algorithm.
-We create state dictionaries for all spectra in the PHOENIX subset, lossily distilling the spectra with a ~20 data compression factor.
-We create the PHOENIX generator, enabling anyone to create reconstructions of PHOENIX spectra with continuously valued stellar parameters.
+We create state dictionaries for all spectra in the PHOENIX subset, lossily distilling the spectra with a data compression factor of around 20.
+We create the PHOENIX generator and implement a performant spectral reconstruction algorithm, enabling anyone to create reconstructions of PHOENIX spectra with continuously valued stellar parameters.
 We introduce and test our inference algorithm, with average absolute deviations from true values of 185 K in $T_\mathrm{eff}$, 0.19 in $\log(g)$, and 0.12 dex in [Fe/H].
+In its current state, our algorithm operates on spectra within the PHOENIX subset parameter ranges in \autoref{tab:table1}, requiring that the spectra not contain noticeable Doppler shifting, rotational broadening, or other confounding factors.
 The methods discussed here represent a step down a road not traveled in spectral inference, and have the potential to become more advanced in the future by fully utilizing the strengths of physics-informed machine learning.
 
-\begin{acknowledgements}
-    Text
-\end{acknowledgements}
+\vspace{0.5cm}
+This material is based on work supported by the National Aeronautics and Space Administration under grant No.
+80NSSC21K0650 for the NNH20ZDA001N-ADAP:D.2 program.
+C.V.M.
+acknowledges support from the Alfred P.
+Sloan Foundation under grant number FG-2021-16592 and support from the National Science Foundation under grant number 1910969.
+\vspace{0.5cm}
 

### 2024-07-26 paper/paper2/paper.tex
Cite Zenodo dataset
Word count (bruh)

@@
 \begin{abstract}
-    The advent of machine learning (ML) methods has and continues to revolutionize numerous scientific disciplines, with a growing number of examples in astronomical spectroscopic inference, as ML is expected to be more powerful than traditional techniques.
+    The advent of machine learning (ML) is revolutionary to numerous scientific disciplines, with a growing number of examples in astronomical spectroscopic inference, as ML is expected to be more powerful than traditional techniques.
     Here we introduce a hybrid ML (HML) method that combines automatic differentiation (autodiff), interpolation, and Bayesian optimization to infer stellar parameters given stellar spectra.
@@
 We create state dictionaries for all spectra in the PHOENIX subset, lossily distilling the spectra with a data compression factor of around 20.
+These state dictionaries have been uploaded to Zenodo at \url{https://zenodo.org/records/11246174} \citep{zenodo}.
 We create the PHOENIX generator and implement a performant spectral reconstruction algorithm, enabling anyone to create reconstructions of PHOENIX spectra with continuously valued stellar parameters.

### 2025-01-19 paper/paper2/paper.tex
Move files

@@
     Here we introduce a hybrid ML (HML) method that combines automatic differentiation (autodiff), interpolation, and Bayesian optimization to infer stellar parameters given stellar spectra.
-    The stellar parameters we study are $T_{\mathrm{eff}}$, $\log(g)$, and [Fe/H], but this method has potential for extension to other properties such as [$\alpha$/Fe], C/O, and $f_{sed}$.
+    The stellar parameters we study are $T_{\mathrm{eff}}$, $\log(g)$, and [Fe/H], but this method has potential for extension to other properties such as [$\alpha$/Fe] \textbf{(alpha element abundance)}, C/O \textbf{(carbon-oxygen ratio)}, and $f_{sed}$ \textbf{(sedimentation efficiency)}.
     We first use \texttt{blase}'s \citep{blase} nontraditional semi-empirical approach to treating spectra as sets of tunable spectral lines.

### 2025-01-21 paper/paper2/paper.tex
Work on revisions

@@
     The stellar parameters we study are $T_{\mathrm{eff}}$, $\log(g)$, and [Fe/H], but this method has potential for extension to other properties such as [$\alpha$/Fe] \textbf{(alpha element abundance)}, C/O \textbf{(carbon-oxygen ratio)}, and $f_{sed}$ \textbf{(sedimentation efficiency)}.
-    We first use \texttt{blase}'s \citep{blase} nontraditional semi-empirical approach to treating spectra as sets of tunable spectral lines.
-    \texttt{blase} is used on 1,314 spectra from the PHOENIX synthetic spectral model grid \citep{PHOENIX}, identifying 128,723 spectral lines.
+    We first use \texttt{blase}'s nontraditional semi-empirical approach to treating spectra as sets of tunable spectral lines.
+    \texttt{blase} is used on 1,314 spectra from the PHOENIX synthetic spectral model grid, identifying 128,723 spectral lines.
     For each of these lines, we continuously map stellar parameters to spectral line parameters using regular grid linear interpolation.
-    These manifolds are aggregated to create the PHOENIX generator, sporting parallelized reconstruction of spectra given continuously valued stellar parameters.
+    These manifolds are aggregated to create the PHOENIX generator, \textbf{enabling} parallelized reconstruction of spectra given continuously valued stellar parameters.
     Gaussian Process minimization is then used to infer stellar parameters by minimizing the root-mean-square (RMS) loss between the input spectrum and PHOENIX generator spectra.
     From testing, the inference error in $T_{\mathrm{eff}}$ was 185 K, $\log(g)$ was 0.19, and [Fe/H] was 0.12 dex.
-    Our products are an archive of the state dictionaries of the \texttt{blase} clones of the PHOENIX subset, as well as the spectral reconstruction and inference algorithms themselves.
+    Our products are an archive of the \texttt{blase} \textbf{models} of the PHOENIX subset, as well as the spectral reconstruction and inference algorithms themselves.
     This study is a proof of concept showing that semi-empirical HML is a viable alternative to traditional approaches.
@@
 
-Much of modern astronomical spectroscopy would benefit from an easy-to-use, fast, accurate, interpretable, and precise spectral model.
+Modern astronomical spectroscopy \textbf{takes advantage of intuitive, performant spectral models such as KORG \citep{korg}, however it} would \textbf{further} benefit from \textbf{the paradigm of interpretability}.
 The determination of fundamental properties, the creation of Extreme Precision Radial Velocity (EPRV) templates, and the application of composite spectral fitting could all leverage such an innovation.
@@
 
-Spectroscopic surveys such as APOGEE use their own in-house pipelines to extract stellar parameters from spectra (ASPCAP in the case of APOGEE), and these pipelines appear effective \citep{APOGEE, ASPCAP}.
+Spectroscopic surveys such as APOGEE \textbf{\citep{APOGEE}} use their own in-house pipelines to extract stellar parameters from spectra (ASPCAP \textbf{\citep{ASPCAP}} in the case of APOGEE), and these pipelines appear effective.
 The core assumption for the vast majority of pipelines is that the data to be analyzed is a list of pixels.
@@
 \texttt{blase}, first presented in \citealt{blase}, took an important, albeit limited, step in this direction, treating spectra not as a set of pixels or a set of eigenbasis coefficients but as a set of interpretable and traceable spectral lines, specifically Voigt profiles.
-Each of these approaches has tradeoffs, but one key scientific advantage of blase comes from its ability to adapt to new information, while preserving some adherence to physics-based models.
+Each of these approaches has tradeoffs, but one key scientific advantage of \texttt{blase} comes from its ability to adapt to new information, while preserving some adherence to physics-based models.
 This intelligent capability stems from its ability to fit a theoretically unlimited number of nonlinear spectral line parameters with automatic differentiation (autodiff).
@@
 With the gradient obtained from autodiff's chain rule, we can optimize model parameters by just going against the direction of the gradient (because we usually want to minimize something such as a loss function, we go in the direction of greatest decrease, which is always the negative of the gradient).
-Autodiff has been used successfully in other astrophysical contexts such as \texttt{exojax} and \texttt{wobble} \citep{exojax, wobble}, but the recasting of spectra into sets of inherently nonlinear spectral lines positions \texttt{blase} as a unique and promising semi-empirical tool.
+Autodiff has been used successfully in other astrophysical contexts such as \texttt{exojax} \textbf{\citep{exojax}} and \texttt{wobble} \citep{wobble}, but the recasting of spectra into sets of inherently nonlinear spectral lines positions \texttt{blase} as a unique and promising semi-empirical tool.
 The original \texttt{blase} paper demonstrated the ability to tune spectral lines with autodiff, but restricted to a pre-selected static synthetic model.
@@
 
-Ideally this process would be monolithic, with the end-to-end spectral inference code powered by a single autodiffable machine learning framework, like PyTorch or JAX \citep{pytorch, jax}.
+Ideally this process would be monolithic, with the end-to-end spectral inference code powered by a single autodiffable machine learning framework, like PyTorch \textbf{\citep{pytorch}} or JAX \citep{jax}.
 However, here we separate the problem into three pieces, only the first of which is currently differentiable.

### 2025-01-22 paper/paper2/paper.tex
Up to 20/81 revisions

@@
 Autodiff has been used successfully in other astrophysical contexts such as \texttt{exojax} \textbf{\citep{exojax}} and \texttt{wobble} \citep{wobble}, but the recasting of spectra into sets of inherently nonlinear spectral lines positions \texttt{blase} as a unique and promising semi-empirical tool.
-The original \texttt{blase} paper demonstrated the ability to tune spectral lines with autodiff, but restricted to a pre-selected static synthetic model.
+The original \texttt{blase} paper demonstrated the ability to tune spectral lines with autodiff \textbf{and `clone' spectra, recasting them as ML models defined by interpretable sets of tuned spectral lines, however it was} restricted to a pre-selected static synthetic model.
 
 Here we introduce the next logical step in the sequence of expanding interpretable spectroscopic machine learning from operating on a single grid point and towards an entire 3D grid of precomputed spectra.
-We rebrand this augmentation as ``blase3D''.
+We rebrand this augmentation as \texttt{blase3D}.
 
@@
 However, here we separate the problem into three pieces, only the first of which is currently differentiable.
-First, we scale out of the blase method to 1,314 precomputed synthetic spectral model clones, yielding a downloadable archive of pretrained machine learning models with 128,723 unique spectral lines.
-Second, we fit manifolds mapping stellar parameters to uniformly-derived spectral line parameters using regular grid linear interpolation.
-Finally, we show how reconstructions of the spectra using said manifolds can be used for inferring stellar fundamental properties from spectra.
-This final step resembles the aims of ``atmospheric retrievals'', but in principle should be computationally much faster, more accurate, adaptive, and reusable.
+\textbf{In \autoref{sec:section2},} we scale out the \texttt{blase} method to 1,314 precomputed synthetic spectral model clones, yielding a downloadable archive of pretrained machine learning models with 128,723 unique spectral lines.
+\textbf{In \autoref{sec:section3},} we \textbf{then} fit manifolds mapping stellar parameters to uniformly-derived spectral line parameters using regular grid linear interpolation.
+Finally, \textbf{in \autoref{sec:section4},} we show how reconstructions of the spectra using said manifolds can be used for inferring stellar fundamental properties from spectra.
+This final step resembles the aims of ``atmospheric retrievals'', but in principle should be faster, adaptive, and reusable.
 An overview of this process is shown in \autoref{fig:figure1}.
@@
 
-We chose the widely-adopted PHOENIX synthetic spectral model grid \citep{PHOENIX} as the basis for this study.
-Our approach can be straightforwardly applied to any other model grid in the future, including substellar atmosphere grids such as Sonora \citep{bobcat, cholla, diamondback, elfowl}, but for now we limit our scope to a subset of the PHOENIX grid.
+\section{Cloning the PHOENIX Model Grid}
+\label{sec:section2}
+\subsection{The PHOENIX Subset}
+\textbf{For the purposes of this study, w}e chose the widely-adopted PHOENIX synthetic spectral model grid \citep{PHOENIX}.
+Our approach can be straightforwardly applied to any other model grid in the future, including substellar atmosphere grids such as Sonora \citep{bobcat, cholla, diamondback, elfowl}, but for now we limit our scope to a \textbf{rectilinear} subset of the PHOENIX grid, focusing on near solar metallicities and a broad range of effective temperature and surface gravity, with details given in \autoref{tab:table1}.
+\textbf{This approach is taken due to the computational cost of interpolation algorithms.
+Future versions with more advanced models will be able to bypass the limitation of a rectilinear subset.}
 
-\section{Cloning the PHOENIX Model Grid}
-\subsection{The PHOENIX Subset}
-For the purposes of this study, we did not consider the full range of the PHOENIX synthetic spectral model grid.
-Instead, we focused on a subset of the grid, focusing on near solar metallicities and a broad range of effective temperature and surface gravity, with details given in \autoref{tab:table1}.
-
-\begin{table}[h!]
-    \centering
-    \begin{tabular}{lll}
+\begin{table*}
+    \hspace*{0.6cm}\begin{tabular}{llll}
         \hline
-        \bf{Parameter} & \bf{Symbol} & \bf{Interval}\\
+        \bf{Parameter} & \bf{Symbol} & \bf{Interval} & \bf{Sampling}\\
         \hline
-        Alpha Element Abundance & $\alpha$ & [0] dex\\
-        Iron Abundance & [Fe/H] & [-0.5, 0] dex\\
-        Effective Temperature & $T_{\mathrm{eff}}$ & [2300, 12000] K\\
-        Surface Gravity & $\log(g)$ & [2, 6]\\
-        Wavelength & $\lambda$ & [8038, 12849] \AA\\
+        Alpha Element Abundance & $\alpha$ & [0] dex & \bf{N/A}\\
+        Iron Abundance & [Fe/H] & [-0.5, 0] dex & \bf{0.5 dex}\\
+        Effective Temperature & $T_{\mathrm{eff}}$ & [2300, 12000] K & \bf{100 K until 7000 K, then 200 K}\\
+        Surface Gravity & $\log(g)$ & [2, 6] & \bf{0.5} \\
+        Wavelength & $\lambda$ & [8038, 12849] \AA & $\mathbf{R = 500,000}$\\
         \hline
@@
     The wavelength limits in particular roughly line up with that of the Habitable Zone Planet Finder (HPF) spectrograph \citep{HPF}.
-    This subset is comprised of 1,314 individual spectra: 73 unique $T_{\mathrm{eff}}$ values, 9 unique $\log(g)$ values, and 2 unique [Fe/H] values.}
+    This subset is comprised of 1,314 individual spectra: 73 $T_{\mathrm{eff}}$ values, 9 $\log(g)$ values, and 2 [Fe/H] values.}
     \label{tab:table1}
-\end{table}
+\end{table*}
 
 \subsection{Preprocessing with \texttt{gollum}}
-First, the PHOENIX subset was programmatically retrieved with \texttt{gollum}, which downloaded them from the PHOENIX FTP server \citep{gollum}. The spectra were then put through a
+First, the PHOENIX subset was programmatically retrieved with \texttt{gollum} \textbf{\citep{gollum}}, which downloaded \textbf{spectra} from the PHOENIX FTP server.
+The spectra were then put through a
 three-step preprocessing pipeline similar to that from \citealt{blase}.
 \begin{enumerate}
-    \item \textit{Blackbody Division}: Since the $T_{\mathrm{eff}}$ of each spectrum is known, the according blackbody spectrum was divided out.
+    \item \textit{Blackbody Division}: Since the $T_{\mathrm{eff}}$ of each spectrum is known, the \textbf{corresponding} blackbody spectrum was divided out.
     \item \textit{Percentile Normalization}: The spectra were normalized by dividing them by their 99th percentile in order to collapse the dynamic range of flux and only look at relative features.
@@
 Mathematically, we can express the preprocessing as follows:
+\begin{linenomath*}
 \begin{gather}
@@
 \end{gather}
+\end{linenomath*}
 where $\mathsf{\bar{S}}$ is the preprocessed spectrum, $\mathsf{S}$ is the original spectrum,  $\mathsf{B}$ is the blackbody spectrum, $\mathsf{Q}_n$ is the $n^\mathrm{th}$ order polynomial continuum fit, and $P_n$ is the $n^\mathrm{th}$ percentile function.
@@
 \subsection{Line Identification with \texttt{blase}}
-The next step was to convert the PHOENIX subset into a physically interpretable intermediate representation: a table of spectral line properties rather than a array of fluxes.
+The next step was to convert the PHOENIX subset into a physically interpretable intermediate representation: a table of spectral line properties rather than a\textbf{n} array of fluxes.
 We used \texttt{blase}, which models spectral lines as Voigt profiles and tunes the profiles to mimic the original PHOENIX spectrum with back propagation.
@@
 Four parameters were optimized: the line center $\mu$, the log-amplitude $\ln(a)$, the Gaussian width $\sigma$, and the Lorentzian width $\gamma$.
-The optimization used the Adam optimizer with a learning rate of 0.05 over 100 epochs \citep{adam}.
+The optimization used the Adam optimizer \textbf{\citep{adam}} with a learning rate of 0.05 over 100 epochs.
 In addition, we limited two custom parameters: wing cut to 6000 and prominence to 0.005.
-Wing cut is a parameter that determines the extent of the Voigt profile to evaluate, saving computational resources by not evaluating negligible line wings.
-Prominence sets a lower limit for the amplitude of detected lines, which also saves resources by disregarding shallow lines, so in our case we disregard lines that affect the spectrum by $<$ 0.5\%.
+Wing cut \textbf{(in pixels)} is a parameter that determines the extent of the Voigt profile to evaluate, saving computational resources by not evaluating negligible line wings.
+Prominence \textbf{(in normalized flux counts)} sets a lower limit for the amplitude of detected lines, which saves resources by disregarding shallow lines, so in our case we disregard lines that affect the spectrum by $<$ 0.5\%.
 In short, our choices for wing cut and prominence decrease the computational cost of \texttt{blase}'s cloning process at the expense of decreasing its accuracy slightly.
-We note that \texttt{blase} uses the pseudo-Voigt approximation, which saves on computational cost while remaining accurate to about 1\% \citep{pseudovoigt1, pseudovoigt2}.
+We note that \texttt{blase} uses the pseudo-Voigt approximation, which saves on computational cost while remaining accurate to about 1\% \citep{pseudovoigt}.
 The pseudo-Voigt approximation uses a weighted average of a Gaussian and Lorentzian as opposed to a convolution.
 \texttt{blase}'s pseudo-Voigt profile implementation uses the following:
+\begin{linenomath*}
 \begin{subequations}
@@
 \end{subequations}
+\end{linenomath*}
 where $\mathbf{L}$ and $\mathbf{G}$ are abbreviations for Lorentzian and Gaussian profiles, respectively.
@@
 \section{Interpolating Manifolds}
+\label{sec:section3}
 \subsection{Cross-Model Line Identification}
@@
 \section{Bayesian Inference and Testing}
+\label{sec:section4}
 \subsection{Spectral Reconstruction Time}

### 2025-01-26 paper/paper2/paper.tex
Up to 40/81 revisions

@@
 The optimization used the Adam optimizer \textbf{\citep{adam}} with a learning rate of 0.05 over 100 epochs.
+\textbf{\citealt{blase} recommends this setup as the minimum, and for a proof-of-concept implementation, we found it best to leave it as is.}
 In addition, we limited two custom parameters: wing cut to 6000 and prominence to 0.005.
 Wing cut \textbf{(in pixels)} is a parameter that determines the extent of the Voigt profile to evaluate, saving computational resources by not evaluating negligible line wings.
-Prominence \textbf{(in normalized flux counts)} sets a lower limit for the amplitude of detected lines, which saves resources by disregarding shallow lines, so in our case we disregard lines that affect the spectrum by $<$ 0.5\%.
+Prominence \textbf{(in normalized flux counts)} sets a lower limit for the amplitude of detected lines, which saves resources by disregarding shallow lines, so in our case we disregard lines \textbf{with amplitude} $<$ 0.05.
 In short, our choices for wing cut and prominence decrease the computational cost of \texttt{blase}'s cloning process at the expense of decreasing its accuracy slightly.
-We note that \texttt{blase} uses the pseudo-Voigt approximation, which saves on computational cost while remaining accurate to about 1\% \citep{pseudovoigt}.
+\texttt{blase} uses the pseudo-Voigt approximation, which saves on computational cost while remaining accurate to about 1\% \citep{pseudovoigt}.
 The pseudo-Voigt approximation uses a weighted average of a Gaussian and Lorentzian as opposed to a convolution.
@@
 The individual Voigt profiles are still indexed by $\mu$ for cross-model line identification, explained in the next section.
-Once optimization was complete, the list of identified lines, was saved to a `state dictionary': a common representation for pre-trained machine learning models that can be stored to disk for reuse later.
+Once optimization was complete, the list of identified lines was saved to a `state dictionary': a common representation for pre-trained machine learning models that can be stored to disk for reuse later.
 These are stored in the \texttt{.pt} file format for each of the 1,314 PHOENIX subset grid points.
@@
 This represents a data compression factor of around 20 just by recasting the spectrum with \texttt{blase}.
+\textbf{The state dictionaries are available on Zenodo at \url{https://zenodo.org/records/11246174} \citep{zenodo}.}
 
@@
 Spectral lines were often only detected in some spectra from the PHOENIX subset.
-In \autoref{fig:figure2}, we show that different grid points sport differing counts of detected spectral lines.
+In \autoref{fig:figure2}, we show that different grid points \textbf{show} differing counts of detected spectral lines.
 \begin{figure*}
@@
 
-This missing-lines phenomenon can arise for different reasons.
+Th\textbf{e number of detected spectral lines changes as a function of stellar parameter} for different reasons.
 Perhaps for astrophysical reasons: stellar atmospheres genuinely do not produce that line at detectable strength at the given temperature and surface gravity.
@@
 In total, across the entire PHOENIX subset, \texttt{blase} detected 128,723 individual spectral lines.
-Every one of these lines can be visualized as a manifold mapping a 3D stellar parameter vector to a 4D spectral line parameter vector, and every one must be interpolated.
+Every one of these lines can be visualized as a manifold mapping a 3D stellar parameter vector to a 4D spectral line parameter vector, and every one \textbf{was} interpolated \textbf{to map stellar parameters to spectral line properties}.
 \begin{figure*}
@@
     \includegraphics[width=\textwidth]{figure3}
-    \caption{Heatmap showing how $\ln(a)$ and $\sigma$ vary over the PHOENIX subset slice at solar metallicity of an unknown spectral line.
+    \caption{Heatmap showing how $\ln(a)$ and $\sigma$ vary over the PHOENIX subset slice at solar metallicity of a \textbf{C I} spectral line.
     Notice the missing chunk in the top left of the figure; \texttt{blase} did not detect a spectral line here, but we have to artificially populate those points with lines that have $\ln(a) = -1000$.
@@
     \includegraphics[width=\textwidth]{figure4}
-    \caption{Heatmap showing how $\ln(a)$ and $\sigma$ vary over the PHOENIX subset slice at solar metallicity of a Helium II spectral line.
+    \caption{Heatmap showing how $\ln(a)$ and $\sigma$ vary over the PHOENIX subset slice at solar metallicity of a He II spectral line.
     We can see that \texttt{blase} detects this line at only a select chunk of grid points in PHOENIX, leading to the large amount of missing data for the line.}
@@
     \includegraphics[width=\textwidth]{figure5}
-    \caption{Heatmap showing how $\ln(a)$ and $\sigma$ of our unknown spectral line vary over the PHOENIX subset slice at solar metallicity, now supersampled with the PHOENIX generator.
+    \caption{Heatmap showing how $\ln(a)$ and $\sigma$ of our \textbf{C I} line vary over the PHOENIX subset slice at solar metallicity, now supersampled with the PHOENIX generator.
     Notice that the missing chunk in the top left still exists and does not display any artifacts, as the artificially populated points are removed after interpolation to retain the model's integrity.
@@
     \includegraphics[width=\textwidth]{figure6}
-    \caption{Heatmap showing how $\ln(a)$ and $\sigma$ of our Helium II spectral line vary over the PHOENIX subset slice at solar metallicity, now supersampled with the PHOENIX generator.}
+    \caption{Heatmap showing how $\ln(a)$ and $\sigma$ of our He II spectral line vary over the PHOENIX subset slice at solar metallicity, now supersampled with the PHOENIX generator.}
     \label{fig:figure6}
@@
 
-\section{Bayesian Inference and Testing}
-\label{sec:section4}
 \subsection{Spectral Reconstruction Time}
@@
 Second, for an array of multiple inputs, which would be relevant in parallel applications.
-The \texttt{RegularGridInterpolator} API allows for the passing in of an entire array of input coordinates to be evaluated at once.
+\textbf{\texttt{scipy}'s} \texttt{RegularGridInterpolator} API allows for the passing in of an entire array of input coordinates to be evaluated at once.
 However using \texttt{blase} to reconstruct the spectrum from our interpolated state dictionary is always done serially, leading to what is actually more of a pseudo-parallel evaluation, but extremely performant nonetheless.
 Performance results are shown in \autoref{fig:figure8}, and we can see that the multi-reconstruction is much faster than a series of single reconstructions.
+\textbf{The computer used for this test was a desktop with an AMD Ryzen 9 7950X3D and 32 GB of RAM (spectral reconstruction does not currently leverage the GPU).}
 
@@
     \includegraphics[width=\columnwidth]{figure8}
-    \caption{Line chart showing the time taken to reconstruct varying numbers of spectra using the PHOENIX generator (lower is better).
-    We can see that the time taken per spectra for the serial implementation hovers around 15 seconds within run-to-run variance, while the pseudo-parallel implementation continually decreases in time taken per spectrum as the number of inputs increases.
+    \caption{\textbf{Plots} showing the time taken to reconstruct varying numbers of spectra using the PHOENIX generator (lower is better).
+    We can see that the time taken per spectr\textbf{um} for the serial implementation hovers around 15 seconds within run-to-run variance, while the pseudo-parallel implementation continually decreases in time taken per spectrum as the number of inputs increases.
     The speedup factor (higher is better) increases as more spectra are generated, which is also a desirable outcome.}
@@
 \end{figure}
+
+\section{Bayesian Inference and Testing}
+\label{sec:section4}
 
@@
 \begin{gather}
-    \mathcal{L} = \ev{(\mathsf{M} - \mathsf{D})^{\circ 2}}^{1/2}
+    \mathcal{L} = \ev{(\mathsf{M} - \mathsf{D})^2}^{1/2}
 \end{gather}
@@
 
-We know that in typical observational settings, a coarse estimate for $T_{\mathrm{eff}}$ tends to be fairly well-constrained from ancillary data prior to even looking at the spectra.
+We know that in typical observational settings, a coarse estimate for $T_{\mathrm{eff}}$ tends to be fairly well-constrained from ancillary data \textbf{such as photometric color index}.
 So when testing the inference algorithm, we limited its search space to only include $T_{\mathrm{eff}}$ which lay within 500 K of the true value on either side.
@@
 First, the familiar \texttt{scipy}-based system will serve as an easy entry point for most practitioners who are unfamiliar with autodiff, and may still benefit from and modify the code without expert ML knowledge.
-Second, this non-autodiff version serves as a pathfinder and benchmark against which an inevitable autodifferentiable version may be compared.
+Second, this non-autodiff version serves as a\textbf{n initial} benchmark against which an inevitable autodifferentiable version may be compared.
 Finally, the inventory of familiar interpolation algorithms have not yet been ported to PyTorch or JAX, since machine learning or Gaussian Process fitting schemes are generally preferred within the ML community.
@@
     \item \textit{Single Model Grid}: The PHOENIX grid is not the only model grid of synthetic spectra available, and it does not apply to all types of stars.
-    Future work would extend the reach of this study's algorithm to encompass other model grids such as the Sonora series of substellar models, ATLAS \citep{atlas}, and coolTLUSTY \citep{coolTLUSTY}, reaching practitioners studying various types of stars and spectra.
-    \texttt{blase} should be able to have an option for the user to input which model grid they would like to base the inference on, and to get even more advanced, perhaps even have the ability to intelligently determine which model grid to use automatically.
+    Future work would extend the reach of this study's algorithm to encompass other model grids such as the Sonora series of substellar models \textbf{\citep{bobcat, cholla, diamondback, elfowl}}, ATLAS \citep{atlas}, and coolTLUSTY \citep{coolTLUSTY}, reaching practitioners studying various types of stars and spectra.
+    \textbf{Future} \texttt{blase} \textbf{versions will} be able to have an option for the user to input which model grid they would like to base the inference on, and to get even more advanced, perhaps even have the ability to intelligently determine which model grid to use automatically.
     \item \textit{Memorization vs.\ Generalization}: The current design of the algorithm constructs manifolds using interpolation.
@@
     \item \textit{The Continuum Black Box}: Continuum normalization is a process that is not yet completely understood, and is currently done as a preprocessing step with a fairly simple algorithm.
-    Future work would dive deeper into the science of continuums and develop more advanced methods that can discern continuums with greater accuracy and less modeling restrictions.
+    Future work would dive deeper into the science of continu\textbf{a} and develop more advanced methods that can discern continu\textbf{a} with greater accuracy and less modeling restrictions.
     This would increase accuracy for end users.
@@
 \section{Conclusion}
-In this study, we have presented a proof-of-concept algorithm that interpolates a subset of the PHOENIX spectral model grid and then uses GP minimization for an inference algorithm.
+In this study, we have presented a proof-of-concept algorithm that interpolates a subset of the PHOENIX spectral model grid and then uses GP minimization \textbf{to infer stellar paremeters}.
 We create state dictionaries for all spectra in the PHOENIX subset, lossily distilling the spectra with a data compression factor of around 20. 
-These state dictionaries have been uploaded to Zenodo at \url{https://zenodo.org/records/11246174} \citep{zenodo}.
+The\textbf{y are available on} Zenodo at \url{https://zenodo.org/records/11246174} \citep{zenodo}.
 We create the PHOENIX generator and implement a performant spectral reconstruction algorithm, enabling anyone to create reconstructions of PHOENIX spectra with continuously valued stellar parameters.

### 2025-02-05 paper/paper2/paper.tex
More paper revisions [45/81]

@@
     The advent of machine learning (ML) is revolutionary to numerous scientific disciplines, with a growing number of examples in astronomical spectroscopic inference, as ML is expected to be more powerful than traditional techniques.
-    Here we introduce a hybrid ML (HML) method that combines automatic differentiation (autodiff), interpolation, and Bayesian optimization to infer stellar parameters given stellar spectra.
-    The stellar parameters we study are $T_{\mathrm{eff}}$, $\log(g)$, and [Fe/H], but this method has potential for extension to other properties such as [$\alpha$/Fe] \textbf{(alpha element abundance)}, C/O \textbf{(carbon-oxygen ratio)}, and $f_{sed}$ \textbf{(sedimentation efficiency)}.
-    We first use \texttt{blase}'s nontraditional semi-empirical approach to treating spectra as sets of tunable spectral lines.
-    \texttt{blase} is used on 1,314 spectra from the PHOENIX synthetic spectral model grid, identifying 128,723 spectral lines.
-    For each of these lines, we continuously map stellar parameters to spectral line parameters using regular grid linear interpolation.
-    These manifolds are aggregated to create the PHOENIX generator, \textbf{enabling} parallelized reconstruction of spectra given continuously valued stellar parameters.
-    Gaussian Process minimization is then used to infer stellar parameters by minimizing the root-mean-square (RMS) loss between the input spectrum and PHOENIX generator spectra.
+    Here we introduce a hybrid ML (HML) method combin{ing} automatic differentiation, interpolation, and Bayesian optimization to infer stellar parameters given stellar spectra.
+    We study $T_{\mathrm{eff}}$, $\log(g)$, and [Fe/H], but this method \textbf{could be extended} to other \textbf{parameters} such as [$\alpha$/Fe] \textbf{(alpha element abundance)}, C/O \textbf{(carbon-oxygen ratio)}, and $f_{sed}$ \textbf{(sedimentation efficiency)}.
+    We first use \texttt{blase}'s nontraditional semi-empirical approach, \textbf{recasting} spectra \textbf{into} sets of tunable spectral lines.
+    \texttt{blase} is \textbf{run} on 1,314 spectra from \textbf{a rectilinear subset of} the PHOENIX synthetic spectral model grid \textbf{([Fe/H]: [-0.5, 0] dex, $T_\text{eff}$: [2300, 12000] K, $\log(g)$: [2, 6])}.
+    For each of the \textbf{128,723} lines, we continuously map stellar parameters to spectral line parameters using regular grid linear interpolation.
+    These manifolds are aggregated to create the PHOENIX generator, \textbf{enabling} parallelized reconstruction of spectra given stellar parameters.
+    Gaussian Process minimization is then used to infer stellar parameters by minimizing the root-mean-square (RMS) loss between input and PHOENIX generator spectra.
     From testing, the inference error in $T_{\mathrm{eff}}$ was 185 K, $\log(g)$ was 0.19, and [Fe/H] was 0.12 dex.
@@
 There are many challenges with creating such a foundational spectral model for astrophysics.
-First, we know our physical inputs into the spectral modeling process are imperfect.
-Second, simplifying assumptions in stellar (and substellar) atmospheres are necessarily inexact.
-Ultimately computational costs have prevented the training of such models.
-Most prevailing solutions have therefore had to choose either a model driven approach, in which the precomputed models are taken as gospel, or a data driven/empirical approach, in which our knowledge of stellar physics is ignored or treated phenomenologically, depending on the application.
+Physical inputs into the spectral modeling process are imperfect\textbf{, and} simplifying assumptions in stellar (and substellar) atmospheres are necessarily inexact \textbf{due to factors such as asymmetries and unknown physics}.
+Computational costs \textbf{make} the training of such models \textbf{challenging; however, advances have been made with attention-based \citep{foundation_attention} and transformer-based \citep{foundation_transformer} models, among others}.
+Most prevailing solutions have had to choose either a model driven approach \textbf{(such as this work)}, in which precomputed models are taken as gospel, or a data driven/empirical approach \textbf{(such as Lux \citep{lux})}, in which our knowledge of stellar physics is ignored or treated phenomenologically, depending on the application.
+\textbf{Hybrid solutions exist and have been implemented, such as those presented in \citealt{leung2019} and \citealt{rains2024}.
+Although a lingering challenge is always how to ideally balance the fusion of model driven and data driven paradigms.}
 
@@
 Performance results are shown in \autoref{fig:figure8}, and we can see that the multi-reconstruction is much faster than a series of single reconstructions.
-\textbf{The computer used for this test was a desktop with an AMD Ryzen 9 7950X3D and 32 GB of RAM (spectral reconstruction does not currently leverage the GPU).}
+\textbf{The computer used for this test has the following specifications (note that spectral reconstruction does not currently leverage the GPU):}
+\begin{table}[h!]
+    \centering
+    \begin{tabular}{ll}
+        \hline
+        CPU & AMD EPYC 7513 (32c/64t, 3.65 GHz boost)\\
+        RAM & 256 GB\\
+        GPU & Nvidia A100 40GB ($\times$2)\\
+        \hline
+    \end{tabular}
+    \caption{This machine was used for all computations, but not for generating visualizations.}
+\end{table}
 
@@
     As interpolators require memorization of the data, advanced interpolation becomes extremely expensive in terms of disk utilization.
-    Future work would involve constructing manifolds using more advanced methods, which would allow for much better generalization, high speed, and lower disk utilization at the expense of some accuracy.
+    Future work would involve constructing manifolds using \textbf{more generalizable ML} methods \textbf{such as lasso or ridge regression}, which would allow for much better generalization, high speed, and lower disk utilization at the expense of some accuracy.
     \item \textit{Extrinsic Absence}: The current design of our algorithm does not account for extrinsic parameters that modify the appearance of spectra such as rotational broadening and Doppler shifting.
@@
 \end{enumerate}
-The computer used for this study, Triton, has the following specifications:
-\begin{table}[h!]
-    \centering
-    \begin{tabular}{ll}
-        \hline
-        CPU & AMD EPYC 7513\\
-        RAM & 256 GB\\
-        GPU & Nvidia A100 40GB ($\times$2)\\
-        \hline
-    \end{tabular}
-    \caption{This machine was used for all computations, but not for generating visualizations. The EPYC 7513 is a 32c/64t CPU with a boost clock of 3.65 GHz. The A100 has 6912 CUDA cores.}
-\end{table}
 

### 2025-02-07 paper/paper2/paper.tex
More paper revisions: up to 50/81

@@
 Multiple efforts have been made in this direction, treating spectra in different ways.
-The standard practice is to treat the wavelength and flux as simply two arrays and use bespoke algorithms tailored to a small number of well-calibrated spectral lines to obtain fundamental stellar parameters and chemical compositions \citep{pipeline1, pipeline2}.
+The standard practice is to treat the wavelength and flux as simply two arrays and use bespoke algorithms tailored to a small number of well-calibrated spectral lines to obtain fundamental stellar parameters and chemical compositions \textbf{(e.g.
+the IGRINS YSO Survey \citep{pipeline1} and the IRTF Spectral Library \citep{pipeline2})}.
 Other whole-spectrum fitting abstractions decompose model spectra into an eigenbasis, implementing the data-model comparison stage as a tractable regression, such as \texttt{starfish} \citep{starfish}.
@@
 We decided to do this by using the line centers $\mu$ of the detected lines pre-optimization.
+
+\textbf{Pre-optimized line centers were required to be equal in order to group lines together.
+The values of $\mu$ pre-optimization are recorded by \texttt{blase} to 0.01 \AA (picometer) precision; if they instead used the full available floating point precision, this method would likely detect singular (ungroupable) instances of spectral lines across the grid, completely negating the premise of this study.
+Of course, this method can be confused by species that produce spectral lines at very close wavelengths and other confounding factors, likely resulting in some false positives and negatives.
+However, we deem it sufficient to approximate this way, as a proper treatment would be time-consuming and computationally expensive.}
+
 Now with each spectral line indexed by $\mu$, we had four parameters to interpolate: $\mu'$, $\ln(a)$, $\sigma$, and $\gamma$.
@@
     \includegraphics[width=\textwidth]{figure2}
-    \caption{Number of detected spectral lines at each grid point of a slice of the PHOENIX subset at solar metallicity.
+    \caption{Number of detected spectral lines at each grid point of the PHOENIX subset.
     We can see that the number of detected lines decreases with increasing $T_{\mathrm{eff}}$.
-    Also note that from $T_{\mathrm{eff}} = 7000$ K onward, PHOENIX's sampling increment changes from 100 K to 200 K.}
+    \textbf{Remember} that from $T_{\mathrm{eff}} = 7000$ K onward, PHOENIX's sampling increment changes from 100 K to 200 K.
+    \textbf{Due to some combination of computational artifacts from our line identification assumptions or innate behavior of PHOENIX grid spectra themselves, outliers such as the obvious one here at $T_{\mathrm{eff}} = 7000$ K and $\log(g) = 6$ appear.
+    Further exploration into why this happens is not in the scope of this study; for inference we are concerned with the lines themselves, not overall grid point behavior.}}
     \label{fig:figure2}
@@
 Perhaps for astrophysical reasons: stellar atmospheres genuinely do not produce that line at detectable strength at the given temperature and surface gravity.
-Or alternatively, our line-finding and line-association algorithms missed it.
+Alternatively, \textbf{\texttt{blase}'s fitting or our line identification assumptions led to artifacts}.
 Whatever the cause, these missing lines have immediate practical consequences.
@@
 These interpolators generate multiple manifolds representing the following mapping:
+\begin{linenomath*}
 \begin{gather}
@@
 \end{gather}
+\end{linenomath*}
 These interpolators could now be evaluated at any point lying within the domain of the PHOENIX subset, turning a discretely sampled PHOENIX subset into a continuously evaluable function, sometimes called a spectral emulator.

### 2025-02-07 paper/paper2/paper.tex
Table 2 autoref

@@
 Performance results are shown in \autoref{fig:figure8}, and we can see that the multi-reconstruction is much faster than a series of single reconstructions.
-\textbf{The computer used for this test has the following specifications (note that spectral reconstruction does not currently leverage the GPU):}
+\textbf{The computer used for this test has specifications shown in \autoref{tab:table2} (note that spectral reconstruction does not currently leverage the GPU):}
 \begin{table}[h!]
@@
     \end{tabular}
-    \caption{This machine was used for all computations, but not for generating visualizations.}
+    \caption{\textbf{Specifications of the} machine used for computations, but not for generating visualizations.}
+    \label{tab:table2}
 \end{table}

### 2025-02-10 paper/paper2/paper.tex
Work on paper revisions: [57/81]

@@
 \textbf{\citealt{blase} recommends this setup as the minimum, and for a proof-of-concept implementation, we found it best to leave it as is.}
-In addition, we limited two custom parameters: wing cut to 6000 and prominence to 0.005.
+In addition, we limited two custom parameters: wing cut to 6,000 and prominence to 0.005.
 Wing cut \textbf{(in pixels)} is a parameter that determines the extent of the Voigt profile to evaluate, saving computational resources by not evaluating negligible line wings.
@@
 Of course, this method can be confused by species that produce spectral lines at very close wavelengths and other confounding factors, likely resulting in some false positives and negatives.
-However, we deem it sufficient to approximate this way, as a proper treatment would be time-consuming and computationally expensive.}
+However, we deem it sufficient to approximate this way, as an in-depth treatment would be time-consuming and computationally expensive.}
 
@@
 Examples of the appearance of missing sections in heatmaps where a line does not appear are shown in \autoref{fig:figure3} and \autoref{fig:figure4}.
-In total, across the entire PHOENIX subset, \texttt{blase} detected 128,723 individual spectral lines.
+Across the entire PHOENIX subset, \texttt{blase} detected 128,723 spectral lines \textbf{with unique $\mu$}.
+\textbf{On average, a PHOENIX spectrum has 9,167 detected spectral lines, with the minimum being 252 ($T_\text{eff}$ = 12,000, $\log(g)$ = 6, [Fe/H] = -0.5) and the maximum being 34,551 ($T_\text{eff}$ = 2,300, $\log(g)$ = 3, [Fe/H] = 0).}
 Every one of these lines can be visualized as a manifold mapping a 3D stellar parameter vector to a 4D spectral line parameter vector, and every one \textbf{was} interpolated \textbf{to map stellar parameters to spectral line properties}.
@@
     \includegraphics[width=\textwidth]{figure3}
-    \caption{Heatmap showing how $\ln(a)$ and $\sigma$ vary over the PHOENIX subset slice at solar metallicity of a \textbf{C I} spectral line.
-    Notice the missing chunk in the top left of the figure; \texttt{blase} did not detect a spectral line here, but we have to artificially populate those points with lines that have $\ln(a) = -1000$.
+    \caption{Heatmap\textbf{s} showing how $\ln(a)$, $\sigma$\textbf{, $\gamma$, and $\mu' - \mu$} vary over the PHOENIX subset slice at solar metallicity of a \textbf{C I} spectral line.
+    Notice the missing chunk in the top left; \texttt{blase} did not detect a spectral line here, \textbf{so} we artificially populate those points with lines that have $\ln(a) = -1000$ \textbf{during interpolation}.
     This and all spectral lines shown in this paper were identified using the NIST spectral line database \citep{NIST}.}
@@
 With the given size of the PHOENIX subset, the interpolator list takes up 13.2 GB of disk space.
+\textbf{This is larger than the original PHOENIX subset by about 63\% and than the state dictionary clone archive by a factor of over 34.
+This is due to the fact that we are now pickling entire function objects that must memorize the data.
+Future work with more space-efficient models and clever storage strategies could reduce storage space requirements in exchange for I/O performance.}
 This evaluation is able to reconstruct an existing PHOENIX spectrum or alternatively interpolate a new spectrum within the domain of the PHOENIX subset, so we call this the PHOENIX generator.
-In \autoref{fig:figure5} and \autoref{fig:figure6}, we show the same spectral lines as in \autoref{fig:figure3} and \autoref{fig:figure4}, but now supersampled using the PHOENIX generator evaluated over the same slice.
-\begin{figure*}
-    \centering
-    \includegraphics[width=\textwidth]{figure5}
-    \caption{Heatmap showing how $\ln(a)$ and $\sigma$ of our \textbf{C I} line vary over the PHOENIX subset slice at solar metallicity, now supersampled with the PHOENIX generator.
-    Notice that the missing chunk in the top left still exists and does not display any artifacts, as the artificially populated points are removed after interpolation to retain the model's integrity.
-    Also see that the x-axis spacing is now uniform, as the PHOENIX generator was evaluated at constant step.}
-    \label{fig:figure5}
-\end{figure*}
-\begin{figure*}
-    \centering
-    \includegraphics[width=\textwidth]{figure6}
-    \caption{Heatmap showing how $\ln(a)$ and $\sigma$ of our He II spectral line vary over the PHOENIX subset slice at solar metallicity, now supersampled with the PHOENIX generator.}
-    \label{fig:figure6}
-\end{figure*}
 
@@
 Any \texttt{nan} values are set to 1 (which we can do because the spectra are all normalized), and the spectral reconstruction is complete.
+We can observe in \autoref{fig:figure5} that \textbf{a Fe I line from a} reconstructed solar spectrum is not simply a pixel interpolation between the nearest \textbf{PHOENIX} grid point \textbf{spectra}.
 \begin{figure*}
     \centering
-    \includegraphics[width=\textwidth]{figure7}
-    \caption{Plot of an Fe I spectral line shown at the closest PHOENIX grid points to the Sun's stellar parameters, as well as a solar spectrum reconstructed with the PHOENIX generator.
+    \includegraphics[width=\textwidth]{figure5}
+    \caption{Plot of an Fe I spectral line shown at the closest PHOENIX grid points to the Sun's stellar parameters \textbf{(solid lines)}, as well as a solar spectrum reconstructed with the PHOENIX generator \textbf{(dotted line)}.
     We can see the spectral line shape be reconstructed mostly between the two native PHOENIX spectra, being closer to the 5800 K spectrum as the Sun's $T_{\mathrm{eff}}$ used here is 5772 K.}
-    \label{fig:figure7}
+    \label{fig:figure5}
 \end{figure*}
-We can observe in \autoref{fig:figure7} that the reconstructed solar spectrum is not simply a pixel interpolation between the nearest grid points.
 It is interpolating the spectral line properties using hundreds of thousands of manifolds, each representing a nonlinear parameter in the shape of the spectral line.
@@
 However using \texttt{blase} to reconstruct the spectrum from our interpolated state dictionary is always done serially, leading to what is actually more of a pseudo-parallel evaluation, but extremely performant nonetheless.
-Performance results are shown in \autoref{fig:figure8}, and we can see that the multi-reconstruction is much faster than a series of single reconstructions.
+Performance results are shown in \autoref{fig:figure6}, and we can see that the multi-reconstruction is much faster than a series of single reconstructions.
 \textbf{The computer used for this test has specifications shown in \autoref{tab:table2} (note that spectral reconstruction does not currently leverage the GPU):}
@@
     \centering
-    \includegraphics[width=\columnwidth]{figure8}
+    \includegraphics[width=\columnwidth]{figure6}
     \caption{\textbf{Plots} showing the time taken to reconstruct varying numbers of spectra using the PHOENIX generator (lower is better).
@@
     The speedup factor (higher is better) increases as more spectra are generated, which is also a desirable outcome.}
-    \label{fig:figure8}
+    \label{fig:figure6}
 \end{figure}
@@
 This algorithm uses a Gaussian Process to model the objective function, which in this case was the RMS (Root-Mean-Square) loss between the interpolated spectrum $\mathsf{M}$ and the true spectrum $\mathsf{D}$, defined as:
+\begin{linenomath*}
 \begin{gather}
@@
 \end{gather}
+\end{linenomath*}
 The optimizer was configured to first run 100 random evaluations to seed the surrogate model, then run 20 more evaluations now guided by the surrogate model.

### 2025-02-19 paper/paper2/paper.tex
Work on Revisions [60/81]

@@
 Performance results are shown in \autoref{fig:figure6}, and we can see that the multi-reconstruction is much faster than a series of single reconstructions.
-\textbf{The computer used for this test has specifications shown in \autoref{tab:table2} (note that spectral reconstruction does not currently leverage the GPU):}
+\textbf{While the generator takes around 15 seconds per spectrum serially, this drops to 1 second per spectrum as soon as 30 reconstructions.
+The speedup factor between the two methods increases as more spectra are generated, reaching a factor of 20 at 60 reconstructions.
+The computer used for this test has specifications shown in \autoref{tab:table2} (note that spectral reconstruction does not currently leverage the GPU):}
 \begin{table}[h!]
@@
 \subsection{Inference Algorithm}
-We elected to use Bayesian optimization as the inference algorithm, specifically the \texttt{gp\_minimize} function from the \texttt{scikit-optimize} library \citep{skopt}.
+\textbf{The goals of this study's inference algorithm are to infer the stellar parameters $T_\text{eff}$, $\log(g)$, and [Fe/H] by comparing a spectrum with our reconstructions and solving the optimization problem.}
+We elected to use Bayesian optimization \textbf{for this component}, specifically the \texttt{gp\_minimize} function from the \texttt{scikit-optimize} library \citep{skopt}.
 This algorithm uses a Gaussian Process to model the objective function, which in this case was the RMS (Root-Mean-Square) loss between the interpolated spectrum $\mathsf{M}$ and the true spectrum $\mathsf{D}$, defined as:
@@
 The optimizer was configured to first run 100 random evaluations to seed the surrogate model, then run 20 more evaluations now guided by the surrogate model.
+\textbf{A random evaluation is simply a randomly generated tuple of stellar parameters within the parameter space of the PHOENIX generator, which is then evaluated with the generator to get a reconstruction and then compared with the inference target to retrieve its RMS error.
+The surrogate model is the approximation to the true objective function that the optimizer constructs based on the set of random evaluations (in this case defined by a set of 100 points rather than a fully continuous, infinitely resolved function)
+Guided evaluations are different, they are not randomly generated tuples of stellar parameters.
+They use the surrogate model in order to select new test points for the minimum of the objective function.}
 This totals to 120 evaluations, a large sample to create a fairly detailed surrogate model, then a moderately precise guided evaluation phase, which was deemed sufficient for this study.

### 2025-02-20 paper/paper2/paper.tex
Revisions more stuff

@@
 They use the surrogate model in order to select new test points for the minimum of the objective function.}
-This totals to 120 evaluations, a large sample to create a fairly detailed surrogate model, then a moderately precise guided evaluation phase, which was deemed sufficient for this study.
+This totals to 120 evaluations, \textbf{100} to create \textbf{the} surrogate model, then \textbf{20} guided evaluation\textbf{s}, which was deemed sufficient for this study.
 Fine-tuning these numbers is possible, but simply not warranted for a proof-of concept method.
+\textbf{Tweaking these numbers could lead to a more accurate inference, however it would incur additional computational cost as well.}
 One inference run takes on average just under 7.5 minutes to complete.

### 2025-02-21 paper/paper2/paper.tex
Paper revisions [69/81]

@@
     We first use \texttt{blase}'s nontraditional semi-empirical approach, \textbf{recasting} spectra \textbf{into} sets of tunable spectral lines.
-    \texttt{blase} is \textbf{run} on 1,314 spectra from \textbf{a rectilinear subset of} the PHOENIX synthetic spectral model grid \textbf{([Fe/H]: [-0.5, 0] dex, $T_\text{eff}$: [2300, 12000] K, $\log(g)$: [2, 6])}.
+    \texttt{blase} is \textbf{run} on 1,314 spectra from \textbf{a rectilinear subset of} the PHOENIX synthetic spectral model grid \textbf{([Fe/H]: [-0.5, 0] dex, $T_\text{eff}$: [2300, 12000] K, $\log(g)$: [2, 6], $\lambda$: [8038, 12849] \AA)}.
     For each of the \textbf{128,723} lines, we continuously map stellar parameters to spectral line parameters using regular grid linear interpolation.
@@
 To test the performance of the inference algorithm, we used the PHOENIX subset itself.
-At first glance, this may seem circular, as the PHOENIX generator has memorized the PHOENIX subset, being able to reconstruct a PHOENIX spectrum when evaluated at that grid point.
-However, that strategy allows us to use the PHOENIX subset as test data in the context of Bayesian optimization.
-The \texttt{gp\_minimize} surrogate model is seeded by random continuously sampled generator evaluations within the search space, \textit{not} grid points of the PHOENIX subset, meaning the surrogate model has no memorization to speak of.
-If the optimizer had been a grid-based strategy, this would not have been possible, because then the surrogate model would be affected by memorization.
+\textbf{This is possible for two reasons: first, the PHOENIX generator has not truly memorized the PHOENIX subset due to its Voigt profile basis and the ignoring of weaker features, and secondly, random samples within the parameter space are extremely unlikely to be PHOENIX grid points themselves, avoiding any overlap.}
 
@@
 Notably, this system tracks the shifting of spectral lines as a function of stellar parameters, which traditionally has been uncommon for algorithms to assess systematically.
-We balance rigidity and flexibility where we want them while maintaining interpretability.
+We balance rigidity \textbf{in the Voigt profile basis} and flexibility \textbf{in line parameter tunability, resulting in an} interpretab\textbf{le system}.
 
-To summarize, we bring to scientists the ability to understand exactly what spectral properties we consider mathematically and how we expect those properties to behave on an incredibly detailed level (the 4 line parameters of all spectral lines and their corresponding interpolating manifolds), and introduce the concept of postulating one precomputed model grid as the ground truth to base all further analysis on, converting the grid into a generator that can then interface with our inference algorithm.
+To summarize, we bring to scientists the ability to understand exactly what spectral properties we consider mathematically and how we expect those properties to behave on an incredibly detailed level (the 4 line parameters of all spectral lines and their corresponding interpolating manifolds).
+\textbf{We treat the PHOENIX} precomputed model grid as the ground truth, \textbf{constructing the PHOENIX} generator that then interface\textbf{s} with our inference algorithm.
 
@@
 First, the familiar \texttt{scipy}-based system will serve as an easy entry point for most practitioners who are unfamiliar with autodiff, and may still benefit from and modify the code without expert ML knowledge.
-Second, this non-autodiff version serves as a\textbf{n initial} benchmark against which an inevitable autodifferentiable version may be compared.
+Second, this non-autodiff version serves as a\textbf{n initial} benchmark against which an inevitable autodifferentiable version \textbf{of \texttt{blase}} may be compared.
 Finally, the inventory of familiar interpolation algorithms have not yet been ported to PyTorch or JAX, since machine learning or Gaussian Process fitting schemes are generally preferred within the ML community.
@@
 \section{Conclusion}
-In this study, we have presented a proof-of-concept algorithm that interpolates a subset of the PHOENIX spectral model grid and then uses GP minimization \textbf{to infer stellar paremeters}.
-We create state dictionaries for all spectra in the PHOENIX subset, lossily distilling the spectra with a data compression factor of around 20. 
+In this study, we have presented a proof-of-concept algorithm that \textbf{deals with} a subset of the PHOENIX spectral model grid \textbf{([Fe/H]: [-0.5, 0] dex, $T_\text{eff}$: [2300, 12000] K, $\log(g)$: [2, 6], $\lambda$: [8038, 12849] \AA)}.
+We create state dictionaries for all spectra in the PHOENIX subset \textbf{using \texttt{blase}, recasting spectra into lists of Voigt profiles, and} lossily compress \textbf{the spectral data by a} factor of around 20.
 The\textbf{y are available on} Zenodo at \url{https://zenodo.org/records/11246174} \citep{zenodo}.
 We create the PHOENIX generator and implement a performant spectral reconstruction algorithm, enabling anyone to create reconstructions of PHOENIX spectra with continuously valued stellar parameters.
-We introduce and test our inference algorithm, with average absolute deviations from true values of 185 K in $T_\mathrm{eff}$, 0.19 in $\log(g)$, and 0.12 dex in [Fe/H].
+We introduce and test our \textbf{GP minimization} algorithm \textbf{to infer stellar parameters}, with average absolute deviations from true values of 185 K in $T_\mathrm{eff}$, 0.19 in $\log(g)$, and 0.12 dex in [Fe/H].
 In its current state, our algorithm operates on spectra within the PHOENIX subset parameter ranges in \autoref{tab:table1}, requiring that the spectra not contain noticeable Doppler shifting, rotational broadening, or other confounding factors.

### 2025-02-26 paper/paper2/paper.tex
Work on paper revisions [72/81]

@@
 First, the PHOENIX subset was programmatically retrieved with \texttt{gollum} \textbf{\citep{gollum}}, which downloaded \textbf{spectra} from the PHOENIX FTP server.
-The spectra were then put through a
-three-step preprocessing pipeline similar to that from \citealt{blase}.
+The spectra were then put through a \textbf{short} preprocessing pipeline similar to that from \citealt{blase}.
 \begin{enumerate}
     \item \textit{Blackbody Division}: Since the $T_{\mathrm{eff}}$ of each spectrum is known, the \textbf{corresponding} blackbody spectrum was divided out.
-    \item \textit{Percentile Normalization}: The spectra were normalized by dividing them by their 99th percentile in order to collapse the dynamic range of flux and only look at relative features.
     \item \textit{Continuum Normalization}: The spectra were further normalized by dividing them by a \nth{5} order polynomial continuum fit using a peak-finding filter in order to eliminate curvature that would inhibit line modeling.
@@
 \begin{gather}
-    \mathsf{\bar{S}} = \dfrac{\mathsf{S}}{\mathsf{B}\mathsf{Q}_5P_{99}}
+    \mathsf{\bar{S}} = \dfrac{\mathsf{S}}{\mathsf{B}\mathsf{Q}_5}
 \end{gather}
 \end{linenomath*}
-where $\mathsf{\bar{S}}$ is the preprocessed spectrum, $\mathsf{S}$ is the original spectrum,  $\mathsf{B}$ is the blackbody spectrum, $\mathsf{Q}_n$ is the $n^\mathrm{th}$ order polynomial continuum fit, and $P_n$ is the $n^\mathrm{th}$ percentile function.
+where $\mathsf{\bar{S}}$ is the preprocessed spectrum, $\mathsf{S}$ is the original spectrum,  $\mathsf{B}$ is the blackbody spectrum, \textbf{and} $\mathsf{Q}_n$ is the $n^\mathrm{th}$ order polynomial continuum fit.
 Arithmetic operations between arrays are assumed to be elementwise in all following notation.
@@
 To solve this, we artificially populated missing grid points with log-amplitudes of -1000, which retained interpolator stability while nullifying the evaluated line.
-Examples of the appearance of missing sections in heatmaps where a line does not appear are shown in \autoref{fig:figure3} and \autoref{fig:figure4}.
+\textbf{To prove the feasibility of this computational trick, we can do a simple thought experiment.
+Moving 1 K in $T_\text{eff}$ away from a grid point on the boundary of detectability would reduce a line's amplitude by a factor of over 20,000 (or more precisely, $e^{10}$).
+This is an extremely large reduction for an extremely small deviation from a grid point, meaning that to any realistic precision, the line immediately vanishes as soon as \texttt{blase} no longer detects it.}
+
+Examples of the appearance of missing sections in heatmaps where a line does not appear \textbf{can be seen} in \autoref{fig:figure3} and \autoref{fig:figure4}\textbf{, which show heatmaps of the 11617.66\AA \ C I and the 10123.66\AA \ He II lines \citep{NIST} varying over a solar metallicity slice of the PHOENIX subset stellar parameter space}.
 Across the entire PHOENIX subset, \texttt{blase} detected 128,723 spectral lines \textbf{with unique $\mu$}.
@@
     \caption{Heatmap\textbf{s} showing how $\ln(a)$, $\sigma$\textbf{, $\gamma$, and $\mu' - \mu$} vary over the PHOENIX subset slice at solar metallicity of a \textbf{C I} spectral line.
-    Notice the missing chunk in the top left; \texttt{blase} did not detect a spectral line here, \textbf{so} we artificially populate those points with lines that have $\ln(a) = -1000$ \textbf{during interpolation}.
+    Notice the missing chunk in the top left; \textbf{this could perhaps be due to the line center shifting to the point that \texttt{blase} identified the chunk as a different line entirely (notice the line center offset increasing in magnitude approaching the chunk border).}
+    \textbf{Since} \texttt{blase} did not detect a spectral line here, \textbf{so} we artificially populate those points with lines that have $\ln(a) = -1000$ \textbf{during interpolation}.
     This and all spectral lines shown in this paper were identified using the NIST spectral line database \citep{NIST}.}
@@
 \end{figure*}
-\begin{figure*}
+\begin{figure}
     \centering
-    \includegraphics[width=\textwidth]{figure4}
+    \includegraphics[width=\columnwidth]{figure4}
     \caption{Heatmap showing how $\ln(a)$ and $\sigma$ vary over the PHOENIX subset slice at solar metallicity of a He II spectral line.
-    We can see that \texttt{blase} detects this line at only a select chunk of grid points in PHOENIX, leading to the large amount of missing data for the line.}
+    \textbf{This plot is much smaller than \autoref{fig:figure3} because} \texttt{blase} detects this line at a select chunk of grid points, leading to \textbf{data only existing between $T_\text{eff}$ 3600 K and 6600 K}.}
     \label{fig:figure4}
-\end{figure*}
+\end{figure}
 

### 2025-03-05 paper/paper2/paper.tex
More revisions [78/81]

@@
     We first use \texttt{blase}'s nontraditional semi-empirical approach, \textbf{recasting} spectra \textbf{into} sets of tunable spectral lines.
-    \texttt{blase} is \textbf{run} on 1,314 spectra from \textbf{a rectilinear subset of} the PHOENIX synthetic spectral model grid \textbf{([Fe/H]: [-0.5, 0] dex, $T_\text{eff}$: [2300, 12000] K, $\log(g)$: [2, 6], $\lambda$: [8038, 12849] \AA)}.
-    For each of the \textbf{128,723} lines, we continuously map stellar parameters to spectral line parameters using regular grid linear interpolation.
+    \texttt{blase} is \textbf{run} on 1,314 spectra from \textbf{a rectilinear subset of} the PHOENIX synthetic spectral model grid \textbf{($T_\text{eff}$: [2300, 12000] K, $\log(g)$: [2, 6], [Fe/H]: [-0.5, 0] dex, $\lambda$: [8038, 12849] \AA)}.
+    For \textbf{128,723 detected features}, we map stellar parameters to spectral line parameters using linear interpolation.
     These manifolds are aggregated to create the PHOENIX generator, \textbf{enabling} parallelized reconstruction of spectra given stellar parameters.
-    Gaussian Process minimization is then used to infer stellar parameters by minimizing the root-mean-square (RMS) loss between input and PHOENIX generator spectra.
-    From testing, the inference error in $T_{\mathrm{eff}}$ was 185 K, $\log(g)$ was 0.19, and [Fe/H] was 0.12 dex.
-    Our products are an archive of the \texttt{blase} \textbf{models} of the PHOENIX subset, as well as the spectral reconstruction and inference algorithms themselves.
+    Gaussian Process minimization is then used to infer stellar parameters by minimizing the root-mean-square loss between input and PHOENIX generator spectra.
+    \textbf{In} testing \textbf{210 noise-free synthetic models ($T_{\mathrm{eff}}$: [3000, 11000] K, $\log(g)$: [2, 6], [Fe/H]: [-0.5, 0] dex), we find inference errors of:} $T_{\mathrm{eff}}$\textbf{:} 185 K, $\log(g)$\textbf{:} 0.19, and [Fe/H]\textbf{:} 0.12 dex.
+    \textbf{We also publish online} an archive of the \texttt{blase} \textbf{models} of the PHOENIX subset.
     This study is a proof of concept showing that semi-empirical HML is a viable alternative to traditional approaches.
@@
 Wing cut \textbf{(in pixels)} is a parameter that determines the extent of the Voigt profile to evaluate, saving computational resources by not evaluating negligible line wings.
-Prominence \textbf{(in normalized flux counts)} sets a lower limit for the amplitude of detected lines, which saves resources by disregarding shallow lines, so in our case we disregard lines \textbf{with amplitude} $<$ 0.05.
+Prominence \textbf{(in normalized flux counts)} sets a lower limit for the amplitude of detected lines, which saves resources by disregarding shallow lines, so in our case we disregard lines \textbf{with amplitude less than} 0.005.
 In short, our choices for wing cut and prominence decrease the computational cost of \texttt{blase}'s cloning process at the expense of decreasing its accuracy slightly.
@@
     \includegraphics[width=\textwidth]{figure3}
-    \caption{Heatmap\textbf{s} showing how $\ln(a)$, $\sigma$\textbf{, $\gamma$, and $\mu' - \mu$} vary over the PHOENIX subset slice at solar metallicity of a \textbf{C I} spectral line.
+    \caption{Heatmap\textbf{s} showing how $\ln(a)$, $\sigma$\textbf{, $\gamma$, and $\mu' - \mu$} vary over the PHOENIX subset slice at solar metallicity of a \textbf{C I} spectral line\textbf{, notable for being detected at nearly all subset grid points}.
     Notice the missing chunk in the top left; \textbf{this could perhaps be due to the line center shifting to the point that \texttt{blase} identified the chunk as a different line entirely (notice the line center offset increasing in magnitude approaching the chunk border).}
-    \textbf{Since} \texttt{blase} did not detect a spectral line here, \textbf{so} we artificially populate those points with lines that have $\ln(a) = -1000$ \textbf{during interpolation}.
+    \textbf{Since} \texttt{blase} did not detect a spectral line here, we artificially populate those points with lines that have $\ln(a) = -1000$ \textbf{during interpolation}.
     This and all spectral lines shown in this paper were identified using the NIST spectral line database \citep{NIST}.}
@@
     \includegraphics[width=\columnwidth]{figure4}
-    \caption{Heatmap showing how $\ln(a)$ and $\sigma$ vary over the PHOENIX subset slice at solar metallicity of a He II spectral line.
+    \caption{Heatmap\textbf{s} showing how $\ln(a)$, $\sigma$\textbf{, $\gamma$, and $\mu' - \mu$} vary over the PHOENIX subset slice at solar metallicity of a He II spectral line.
     \textbf{This plot is much smaller than \autoref{fig:figure3} because} \texttt{blase} detects this line at a select chunk of grid points, leading to \textbf{data only existing between $T_\text{eff}$ 3600 K and 6600 K}.}
@@
 Any \texttt{nan} values are set to 1 (which we can do because the spectra are all normalized), and the spectral reconstruction is complete.
-We can observe in \autoref{fig:figure5} that \textbf{a Fe I line from a} reconstructed solar spectrum is not simply a pixel interpolation between the nearest \textbf{PHOENIX} grid point \textbf{spectra}.
-\begin{figure*}
+\textbf{In \autoref{fig:figure5}, we see a comparison between a PHOENIX generator solar spectrum and two nearby native PHOENIX grid point spectra.}
+\begin{figure}
     \centering
-    \includegraphics[width=\textwidth]{figure5}
-    \caption{Plot of an Fe I spectral line shown at the closest PHOENIX grid points to the Sun's stellar parameters \textbf{(solid lines)}, as well as a solar spectrum reconstructed with the PHOENIX generator \textbf{(dotted line)}.
-    We can see the spectral line shape be reconstructed mostly between the two native PHOENIX spectra, being closer to the 5800 K spectrum as the Sun's $T_{\mathrm{eff}}$ used here is 5772 K.}
+    \includegraphics[width=\columnwidth]{figure5}
+    \caption{Plot of \textbf{the 10123.66 \AA \ He II} spectral line shown at the closest PHOENIX grid points to the Sun's stellar parameters \textbf{(solid lines)}, as well as a solar spectrum reconstructed with the PHOENIX generator \textbf{(dotted line)}.
+    We can see the spectral line shape be reconstructed mostly between the two native PHOENIX spectra.}
     \label{fig:figure5}
-\end{figure*}
+\end{figure}
+
+We can observe in \autoref{fig:figure6} that \textbf{a Fe I line from a} reconstructed solar spectrum is not simply a pixel interpolation between the nearest \textbf{PHOENIX} grid point \textbf{spectra}.
+\begin{figure}
+    \centering
+    \includegraphics[width=\columnwidth]{figure6}
+    \caption{\textbf{Comparison between the PHOENIX generator's line interpolation and a standard pixel intrpolation of the 10123.66 \AA \ He II spectral line.}}
+    \label{fig:figure6}
+\end{figure}
 It is interpolating the spectral line properties using hundreds of thousands of manifolds, each representing a nonlinear parameter in the shape of the spectral line.
@@
 However using \texttt{blase} to reconstruct the spectrum from our interpolated state dictionary is always done serially, leading to what is actually more of a pseudo-parallel evaluation, but extremely performant nonetheless.
-Performance results are shown in \autoref{fig:figure6}, and we can see that the multi-reconstruction is much faster than a series of single reconstructions.
-\textbf{While the generator takes around 15 seconds per spectrum serially, this drops to 1 second per spectrum as soon as 30 reconstructions.
+
+Performance results are shown in \autoref{fig:figure7}, and we can see that \textbf{the PHOENIX generator's parallel mode} is much faster than \textbf{serial mode}.
+\textbf{While the generator takes around 15 seconds per spectrum in serial mode, this drops to 1 second per spectrum in parallel mode as soon as 30 reconstructions.
 The speedup factor between the two methods increases as more spectra are generated, reaching a factor of 20 at 60 reconstructions.
-The computer used for this test has specifications shown in \autoref{tab:table2} (note that spectral reconstruction does not currently leverage the GPU):}
+Assuming linear time scaling with wavelength range, according to KORG benchmarks \citep{korg}, which generate spectra over a 500 \AA \ wavelength range, the PHOENIX generator in serial mode is approximately as fast as KORG and $\sim$30 times faster than Turbospectrum \citep{turbospectrum}.
+The speedup factor in parallel mode compared to KORG is then the same speedup factor from the serial mode.}
+
+\textbf{The computer used for this test has specifications shown in \autoref{tab:table2} (note that spectral reconstruction does not currently leverage the GPU):}
 \begin{table}[h!]
@@
     \end{tabular}
-    \caption{\textbf{Specifications of the} machine used for computations, but not for generating visualizations.}
+    \caption{\textbf{Specifications of the} machine used for computation.
+    \textbf{This machine is owned by Caroline Morley.}}
     \label{tab:table2}
@@
     \centering
-    \includegraphics[width=\columnwidth]{figure6}
+    \includegraphics[width=\columnwidth]{figure7}
     \caption{\textbf{Plots} showing the time taken to reconstruct varying numbers of spectra using the PHOENIX generator (lower is better).
@@
     The speedup factor (higher is better) increases as more spectra are generated, which is also a desirable outcome.}
-    \label{fig:figure6}
+    \label{fig:figure7}
 \end{figure}
@@
 We create the PHOENIX generator and implement a performant spectral reconstruction algorithm, enabling anyone to create reconstructions of PHOENIX spectra with continuously valued stellar parameters.
-We introduce and test our \textbf{GP minimization} algorithm \textbf{to infer stellar parameters}, with average absolute deviations from true values of 185 K in $T_\mathrm{eff}$, 0.19 in $\log(g)$, and 0.12 dex in [Fe/H].
+We introduce and test our \textbf{GP minimization} algorithm \textbf{to infer stellar parameters.} 
+\textbf{Testing on 210 noise-free synthetic models ($T_{\mathrm{eff}}$: [3000, 11000] K, $\log(g)$: [2, 6], [Fe/H]: [-0.5, 0] dex) yields} average absolute deviations from true values of 185 K in $T_\mathrm{eff}$, 0.19 in $\log(g)$, and 0.12 dex in [Fe/H].
 In its current state, our algorithm operates on spectra within the PHOENIX subset parameter ranges in \autoref{tab:table1}, requiring that the spectra not contain noticeable Doppler shifting, rotational broadening, or other confounding factors.

### 2025-03-05 experiments/08_blase3D_HPC_test/08_blase3D_HPC_test.py

@@ -8,7 +8,7 @@ from sys import argv
 
 
 def main(wl_lo: int, wl_hi: int, device: str | torch.device):
-    grid = PHOENIXGrid(wl_lo=wl_lo, wl_hi=wl_hi, path="/data/libraries/raw/PHOENIX/")
+    grid = PHOENIXGrid(wl_lo=wl_lo, wl_hi=wl_hi, path="/mnt/c/Users/sujay/Downloads/PHOENIX/")
     for spec in grid:
         run_emulator(spec, device=device)
 
@@ -24,6 +24,6 @@ if __name__ == "__main__":
                 device = torch.device("mps")
             except RuntimeError:
                 device = torch.device("cpu") 
-    main(int(argv[1]), int(argv[2]), device)
+    main(8038, 12849, device)
     end_time = dt.now()
     system(f"echo {f'{start_time} to {end_time}: {end_time - start_time}'} >> delta.txt")

### 2025-03-05 experiments/10_end_to_end/10_end_to_end.py

@@ -13,6 +13,7 @@ from pickle import dump, load
 from re import split
 from skopt import gp_minimize
 from scipy.interpolate import RegularGridInterpolator
+from sys import getsizeof
 from time import perf_counter
 from tqdm import tqdm
 from typing import Callable
@@ -99,29 +100,18 @@ def inference_test():
 
 
 if __name__ == '__main__':
-    """sys.stderr = open('log.txt', 'w')
-    sys.stdout = open('out.txt', 'w')
     start = perf_counter()
     interpolator_list = load(open('interpolator_list.pkl', 'rb'))
     R = np.random.default_rng()
-    teff_samples = np.arange(3000, 11000.1, 1000)
-    logg_samples = np.arange(2, 6.1, 2)
+    teff_samples = np.arange(3000, 11000.1, 400)
+    logg_samples = np.arange(2, 6.1, 1)
     Z_samples = np.arange(-0.5, 0.1, 0.5)
     print(f'Initializations complete in {perf_counter() - start} s')
     for T, G, Z in product(teff_samples, logg_samples, Z_samples):
-        spec = default_clean(PHOENIXSpectrum(teff=int(T), logg=G, Z=Z, download=True))
+        spec = default_clean(PHOENIXSpectrum(teff=int(T), logg=G, Z=Z, wl_lo=8038, wl_hi=12849))
         start = perf_counter()
         point_random = R.uniform([T-500, 2, -0.5], [T+500, 6, 0], (100, 3))
         reconstructions = reconstructn(spec.wavelength.value, point_random, interpolator_list)
         rms_array = np.sqrt(((reconstructions - spec.flux.value)**2).mean(axis=1))
-        res = gp_minimize(rms_loss(spec.wavelength.value, spec.flux.value, interpolator_list), dimensions=[(T-500.0, T+500), (2.0, 6), (-0.5, 0)], n_calls=20, x0=[list(array) for array in point_random], y0=list(rms_array), n_initial_points=0, n_jobs=16)
-        print(f'{(T, G, Z)} -> {res.x} achieved in {perf_counter() - start} s')"""
-    df = read_state_dicts('/home/sujays/github/blase/experiments/08_blase3D_HPC_test/emulator_states')
-    df_gp = df[['teff', 'logg', 'Z']]
-    df = df.explode(['center', 'amp', 'sigma', 'gamma', 'shift_center']).convert_dtypes(dtype_backend='numpy_nullable')
[…]

### 2025-03-05 src/blase/optimizer.py

@@ -9,7 +9,7 @@ extract = lambda x: x.detach().cpu().numpy()
 
 
 def default_clean(spec: PHOENIXSpectrum):
-    norm_spec = spec.divide_by_blackbody().normalize(99) # B in paper
+    norm_spec = spec.divide_by_blackbody() # B in paper
     continuum = norm_spec.fit_continuum(polyorder=5) # P in paper
     return norm_spec.divide(continuum, handle_meta="ff") # S in paper
 

### 2025-03-05 paper/paper2/paper.tex
Nearly finished! Just need to cite like one more thing!

@@
 \begin{abstract}
-    The advent of machine learning (ML) is revolutionary to numerous scientific disciplines, with a growing number of examples in astronomical spectroscopic inference, as ML is expected to be more powerful than traditional techniques.
-    Here we introduce a hybrid ML (HML) method combin{ing} automatic differentiation, interpolation, and Bayesian optimization to infer stellar parameters given stellar spectra.
+    The advent of machine learning (ML) is revolutionary to numerous scientific disciplines, with a growing number of examples in astronomical spectroscopic inference, as ML is more powerful than traditional techniques.
+    Here we introduce a hybrid ML (HML) method combin{ing} automatic differentiation, interpolation, and Bayesian optimization to infer stellar parameters \textbf{of} stellar spectra.
     We study $T_{\mathrm{eff}}$, $\log(g)$, and [Fe/H], but this method \textbf{could be extended} to other \textbf{parameters} such as [$\alpha$/Fe] \textbf{(alpha element abundance)}, C/O \textbf{(carbon-oxygen ratio)}, and $f_{sed}$ \textbf{(sedimentation efficiency)}.
-    We first use \texttt{blase}'s nontraditional semi-empirical approach, \textbf{recasting} spectra \textbf{into} sets of tunable spectral lines.
-    \texttt{blase} is \textbf{run} on 1,314 spectra from \textbf{a rectilinear subset of} the PHOENIX synthetic spectral model grid \textbf{($T_\text{eff}$: [2300, 12000] K, $\log(g)$: [2, 6], [Fe/H]: [-0.5, 0] dex, $\lambda$: [8038, 12849] \AA)}.
+    We first use \texttt{blase} \textbf{to} semi-empirical\textbf{ly recast} spectra \textbf{into} sets of \textbf{Voigt profiles}.
+    \texttt{blase} is \textbf{run} on 1,314 \textbf{synthetic} spectra from \textbf{a rectilinear subset of} the PHOENIX model grid \textbf{($T_\text{eff}$: [2300, 12000] K, $\log(g)$: [2, 6] dex, [Fe/H]: [-0.5, 0] dex, $\lambda$: [8038, 12849] \AA)}.
     For \textbf{128,723 detected features}, we map stellar parameters to spectral line parameters using linear interpolation.
-    These manifolds are aggregated to create the PHOENIX generator, \textbf{enabling} parallelized reconstruction of spectra given stellar parameters.
-    Gaussian Process minimization is then used to infer stellar parameters by minimizing the root-mean-square loss between input and PHOENIX generator spectra.
-    \textbf{In} testing \textbf{210 noise-free synthetic models ($T_{\mathrm{eff}}$: [3000, 11000] K, $\log(g)$: [2, 6], [Fe/H]: [-0.5, 0] dex), we find inference errors of:} $T_{\mathrm{eff}}$\textbf{:} 185 K, $\log(g)$\textbf{:} 0.19, and [Fe/H]\textbf{:} 0.12 dex.
-    \textbf{We also publish online} an archive of the \texttt{blase} \textbf{models} of the PHOENIX subset.
-    This study is a proof of concept showing that semi-empirical HML is a viable alternative to traditional approaches.
+    \textbf{This} create\textbf{s} the PHOENIX generator, \textbf{enabling parallelized spectral synthesis}.
+    Gaussian Process minimization is used to infer stellar parameters by minimizing \textbf{a} root-mean-square loss \textbf{function}.
+    Testing \textbf{210 noise-free models ($T_{\mathrm{eff}}$: [3000, 11000] K, $\log(g)$: [2, 6] dex, [Fe/H]: [-0.5, 0] dex), we find inference errors:} $T_{\mathrm{eff}}$\textbf{: 93 K, $\log(g)$: 0.24 dex, and [Fe/H]: 0.056 dex for $T_\text{eff}$ < 7000 K, and $T_{\mathrm{eff}}$: 347 K, $\log(g)$: 0.26 dex, and [Fe/H]: 0.16 dex for $T_\text{eff} \ge$ 7000 K}.
+    \textbf{We also upload online} an archive of \texttt{blase} \textbf{models} of the PHOENIX subset.
+    This proof-of-concept \textbf{study} shows that semi-empirical HML is a viable alternative to traditional approaches \textbf{in spectroscopic inference}.
 \end{abstract}
@@
         Effective Temperature & $T_{\mathrm{eff}}$ & [2300, 12000] K & \bf{100 K until 7000 K, then 200 K}\\
-        Surface Gravity & $\log(g)$ & [2, 6] & \bf{0.5} \\
+        Surface Gravity & $\log(g)$ & [2, 6] dex & \bf{0.5 dex} \\
         Wavelength & $\lambda$ & [8038, 12849] \AA & $\mathbf{R = 500,000}$\\
@@
 In \autoref{fig:figure2}, we show that different grid points \textbf{show} differing counts of detected spectral lines.
+
 \begin{figure*}
@@
     \textbf{Remember} that from $T_{\mathrm{eff}} = 7000$ K onward, PHOENIX's sampling increment changes from 100 K to 200 K.
-    \textbf{Due to some combination of computational artifacts from our line identification assumptions or innate behavior of PHOENIX grid spectra themselves, outliers such as the obvious one here at $T_{\mathrm{eff}} = 7000$ K and $\log(g) = 6$ appear.
+    \textbf{Due to some combination of computational artifacts from our line identification assumptions or innate behavior of PHOENIX grid spectra themselves, outliers such as the obvious one here at $T_{\mathrm{eff}} = 7000$ K and $\log(g) = 6$ dex appear.
     Further exploration into why this happens is not in the scope of this study; for inference we are concerned with the lines themselves, not overall grid point behavior.}}
@@
 
-Examples of the appearance of missing sections in heatmaps where a line does not appear \textbf{can be seen} in \autoref{fig:figure3} and \autoref{fig:figure4}\textbf{, which show heatmaps of the 11617.66\AA \ C I and the 10123.66\AA \ He II lines \citep{NIST} varying over a solar metallicity slice of the PHOENIX subset stellar parameter space}.
+Examples of the appearance of missing sections in heatmaps where a line does not appear \textbf{can be seen} in \autoref{fig:figure3} and \autoref{fig:figure4}\textbf{, which show heatmaps of the 11617.66 \AA \ C I and the 10123.66 \AA \ He II lines (identified with NIST \citep{NIST}) varying over a solar metallicity slice of the PHOENIX subset stellar parameter space}.
 Across the entire PHOENIX subset, \texttt{blase} detected 128,723 spectral lines \textbf{with unique $\mu$}.
-\textbf{On average, a PHOENIX spectrum has 9,167 detected spectral lines, with the minimum being 252 ($T_\text{eff}$ = 12,000, $\log(g)$ = 6, [Fe/H] = -0.5) and the maximum being 34,551 ($T_\text{eff}$ = 2,300, $\log(g)$ = 3, [Fe/H] = 0).}
+\textbf{On average, a PHOENIX spectrum has 9,167 detected spectral lines, with the minimum being 252 ($T_\text{eff}$ = 12000 K, $\log(g)$ = 6 dex, [Fe/H] = -0.5 dex) and the maximum being 34,551 ($T_\text{eff}$ = 2300 K, $\log(g)$ = 3 dex, [Fe/H] = 0 dex).}
 Every one of these lines can be visualized as a manifold mapping a 3D stellar parameter vector to a 4D spectral line parameter vector, and every one \textbf{was} interpolated \textbf{to map stellar parameters to spectral line properties}.
@@
     Notice the missing chunk in the top left; \textbf{this could perhaps be due to the line center shifting to the point that \texttt{blase} identified the chunk as a different line entirely (notice the line center offset increasing in magnitude approaching the chunk border).}
-    \textbf{Since} \texttt{blase} did not detect a spectral line here, we artificially populate those points with lines that have $\ln(a) = -1000$ \textbf{during interpolation}.
-    This and all spectral lines shown in this paper were identified using the NIST spectral line database \citep{NIST}.}
+    \textbf{Since} \texttt{blase} did not detect a spectral line here, we artificially populate those points with lines that have $\ln(a) = -1000$ \textbf{during interpolation}.}
     \label{fig:figure3}
@@
     \caption{Heatmap\textbf{s} showing how $\ln(a)$, $\sigma$\textbf{, $\gamma$, and $\mu' - \mu$} vary over the PHOENIX subset slice at solar metallicity of a He II spectral line.
-    \textbf{This plot is much smaller than \autoref{fig:figure3} because} \texttt{blase} detects this line at a select chunk of grid points, leading to \textbf{data only existing between $T_\text{eff}$ 3600 K and 6600 K}.}
+    \texttt{blase} detects this line at a select chunk of grid points, leading to \textbf{data only existing between $T_\text{eff}$ 3600 K and 6600 K}.}
     \label{fig:figure4}
@@
 Any \texttt{nan} values are set to 1 (which we can do because the spectra are all normalized), and the spectral reconstruction is complete.
-\textbf{In \autoref{fig:figure5}, we see a comparison between a PHOENIX generator solar spectrum and two nearby native PHOENIX grid point spectra.}
+\textbf{In \autoref{fig:figure5}, we see a comparison between a PHOENIX generator solar spectrum and two nearby native PHOENIX grid point spectra at the 10123.66 \AA \ He II spectral line.}
 \begin{figure}
@@
     \includegraphics[width=\columnwidth]{figure5}
-    \caption{Plot of \textbf{the 10123.66 \AA \ He II} spectral line shown at the closest PHOENIX grid points to the Sun's stellar parameters \textbf{(solid lines)}, as well as a solar spectrum reconstructed with the PHOENIX generator \textbf{(dotted line)}.
-    We can see the spectral line shape be reconstructed mostly between the two native PHOENIX spectra.}
+    \caption{Plot of \textbf{the 10123.66 \AA \ He II} spectral line shown at the closest PHOENIX grid points to the Sun's stellar parameters \textbf{(solid lines)}, as well as a solar spectrum reconstructed with the PHOENIX generator \textbf{(dotted line)}.}
     \label{fig:figure5}
@@
 
-We can observe in \autoref{fig:figure6} that \textbf{a Fe I line from a} reconstructed solar spectrum is not simply a pixel interpolation between the nearest \textbf{PHOENIX} grid point \textbf{spectra}.
+We can observe in \autoref{fig:figure6} that \textbf{our He II line from a} reconstructed solar spectrum is not simply a pixel interpolation between the nearest \textbf{PHOENIX} grid point \textbf{spectra}.
 \begin{figure}
@@
     \includegraphics[width=\columnwidth]{figure6}
-    \caption{\textbf{Comparison between the PHOENIX generator's line interpolation and a standard pixel intrpolation of the 10123.66 \AA \ He II spectral line.}}
+    \caption{\textbf{Comparison between the PHOENIX generator's line interpolation and a standard pixel interpolation of the 10123.66 \AA \ He II spectral line.
+    Notice that line interpolation results in more symmetrical line wings due to the use of Voigt profiles.}}
     \label{fig:figure6}
@@
 Assuming linear time scaling with wavelength range, according to KORG benchmarks \citep{korg}, which generate spectra over a 500 \AA \ wavelength range, the PHOENIX generator in serial mode is approximately as fast as KORG and $\sim$30 times faster than Turbospectrum \citep{turbospectrum}.
-The speedup factor in parallel mode compared to KORG is then the same speedup factor from the serial mode.}
+The speedup factor in parallel mode compared to KORG is then the same variable speedup factor from serial mode (20 times faster at 60 reconstructions).}
 
@@
         RAM & 256 GB\\
-        GPU & Nvidia A100 40GB ($\times$2)\\
+        GPU & Nvidia A100 [40GB] ($\times$2)\\
         \hline
     \end{tabular}
-    \caption{\textbf{Specifications of the} machine used for computation.
+    \caption{\textbf{Specifications of the} machine used for \textbf{cloning and reconstruction testing}.
     \textbf{This machine is owned by Caroline Morley.}}
@@
 \textbf{Tweaking these numbers could lead to a more accurate inference, however it would incur additional computational cost as well.}
-One inference run takes on average just under 7.5 minutes to complete.
+One inference takes on average \textbf{200 seconds, using the computer with specifications in \autoref{tab:table3}}.
+
+\textbf{Inference algorithms that output more information and produce posterior distributions such as \texttt{starfish} \citep{starfish} or MINESweeper \citep{minesweeper} take on the order of an hour to complete an inference on supercomputers.
+\texttt{astroNN} \citep{leung2019} is a more direct comparison, and represents an extremely powerful inference tool that performs inference thousands of times faster thanks to its neural network architecture.}
+\begin{table}[h!]
+    \centering
+    \begin{tabular}{ll}
+        \hline
+        CPU & AMD Ryzen 9 7950X3D (16c/32t, 5.7 GHz boost)\\
+        RAM & 32 GB\\
+        GPU & Nvidia GeForce RTX 4080 SUPER [16GB]\\
+        \hline
+    \end{tabular}
+    \caption{\textbf{Specifications of the machine used for general testing and for generating visualizations.}
+    \textbf{This machine is owned by Sujay Shankar.}}
+    \label{tab:table3}
+\end{table}
 
@@
 $\log(g)$ and [Fe/H] were allowed to vary freely.
-The test sample consisted of 9 unique $T_{\mathrm{eff}}$ values, 3 unique $\log(g)$ values, and 2 unique [Fe/H] values, totaling 54 unique spectra in the test set.
-$T_{\mathrm{eff}}$ ranged from 3000 K to 11000 K in increments of 1000 K, $\log(g)$ ranged from 2 to 6 in increments of 2, and [Fe/H] ranged from -0.5 to 0 in increments of 0.5.
+The test sample consisted of 21 $T_{\mathrm{eff}}$ values\textbf{: [3000, 11000] K in steps of 400 K}, 5 $\log(g)$ values\textbf{: [2, 6] dex in steps of 0.5 dex} , and 2 [Fe/H] values\textbf{: [-0.5, 0] dex in steps of 0.5 dex}, totaling \textbf{210} unique spectra in the test set.
+\textbf{Two test passes were run, meaning that 420 inference tests were run in total.}
 
-The results of the inference algorithm are as follows: $T_{\mathrm{eff}}$ differed from the true result by an average of 185 K or 2.6\%.
-$\log(g)$ differed by an average of 0.19 or 6.8\%. [Fe/H] differed by an average of 0.12 dex, which is 24\% of our search range.
+In \autoref{fig:figure8}, we show the residuals of the inferred stellar parameters with respect to the true values.
+\begin{figure*}
+    \centering
+    \includegraphics[width=\textwidth]{figure8}
+    \caption{\textbf{Residual plots for the GP minimization inference algorithm, all plotted with respect to $T_\text{eff}$ (values aggregated across $\log(g)$ and [Fe/H]).
+    Notice the interesting change in behavior of all residuals at around $T_\text{eff} = 7000$ K.}}
+    \label{fig:figure8}
+\end{figure*}
+\textbf{We see that there is a distinction between the algorithm's behavior above and below $T_{\mathrm{eff}} = 7000$ K, so we define low and high temperature regimes accordingly.
+For $T_{\mathrm{eff}}$ < 7000 K, the inferred} $T_{\mathrm{eff}}$ differed from the true \textbf{value} by an average of \textbf{93} K, $\log(g)$ \textbf{0.24 dex}, and [Fe/H] \textbf{0.056} dex \textbf{(standard deviations of 139 K, 0.40 dex, and 0.15 dex, respectively)}.
+\textbf{For $T_{\mathrm{eff}} \ge$ 7000 K, the inferred $T_{\mathrm{eff}}$ differed from the true value by an average of 347 K, $\log(g)$ 0.26 dex, and [Fe/H] 0.16 dex (standard deviations of 293 K, 0.32 dex, and 0.23 dex, respectively).}
 From this, we can see that $T_{\mathrm{eff}}$ was the most accurately inferred parameter, followed by $\log(g)$, and then [Fe/H].
@@
 \section{Conclusion}
-In this study, we have presented a proof-of-concept algorithm that \textbf{deals with} a subset of the PHOENIX spectral model grid \textbf{([Fe/H]: [-0.5, 0] dex, $T_\text{eff}$: [2300, 12000] K, $\log(g)$: [2, 6], $\lambda$: [8038, 12849] \AA)}.
+In this study, we have presented a proof-of-concept algorithm that \textbf{deals with} a subset of the PHOENIX spectral model grid \textbf{($T_\text{eff}$: [2300, 12000] K, $\log(g)$: [2, 6] dex, [Fe/H]: [-0.5, 0] dex, $\lambda$: [8038, 12849] \AA)}.
 We create state dictionaries for all spectra in the PHOENIX subset \textbf{using \texttt{blase}, recasting spectra into lists of Voigt profiles, and} lossily compress \textbf{the spectral data by a} factor of around 20.
 The\textbf{y are available on} Zenodo at \url{https://zenodo.org/records/11246174} \citep{zenodo}.
+
 We create the PHOENIX generator and implement a performant spectral reconstruction algorithm, enabling anyone to create reconstructions of PHOENIX spectra with continuously valued stellar parameters.
-We introduce and test our \textbf{GP minimization} algorithm \textbf{to infer stellar parameters.} 
-\textbf{Testing on 210 noise-free synthetic models ($T_{\mathrm{eff}}$: [3000, 11000] K, $\log(g)$: [2, 6], [Fe/H]: [-0.5, 0] dex) yields} average absolute deviations from true values of 185 K in $T_\mathrm{eff}$, 0.19 in $\log(g)$, and 0.12 dex in [Fe/H].
-In its current state, our algorithm operates on spectra within the PHOENIX subset parameter ranges in \autoref{tab:table1}, requiring that the spectra not contain noticeable Doppler shifting, rotational broadening, or other confounding factors.
+
+We \textbf{then} introduce and test our \textbf{GP minimization} algorithm \textbf{to infer stellar parameters.} 
+\textbf{Two test passes on 210 noise-free synthetic models ($T_{\mathrm{eff}}$: [3000, 11000] K, $\log(g)$: [2, 6] dex, [Fe/H]: [-0.5, 0] dex) yield} average absolute deviations from true values of \textbf{93} K in $T_\mathrm{eff}$, \textbf{0.24 dex} in $\log(g)$, and \textbf{0.056} dex in [Fe/H] \textbf{in low-temperature ($T_\mathrm{eff}$ < 7000) regimes.
+In high-temperature regimes ($T_\text{eff} \ge$ 7000), the errors increase to 347 K in $T_\mathrm{eff}$, 0.26 dex in $\log(g)$, and 0.16 dex in [Fe/H]}.
+
+In its current state, our algorithm requir\textbf{es} that the spectra not contain noticeable Doppler shifting, rotational broadening, \textbf{noise}, or other confounding factors.
 The methods discussed here represent a step down a road not traveled in spectral inference, and have the potential to become more advanced in the future by fully utilizing the strengths of physics-informed machine learning.

### 2025-03-05 paper/paper2/paper.tex
GG?!

@@
 
-Ideally this process would be monolithic, with the end-to-end spectral inference code powered by a single autodiffable machine learning framework, like PyTorch \textbf{\citep{pytorch}} or JAX \citep{jax}.
-However, here we separate the problem into three pieces, only the first of which is currently differentiable.
+Ideally this process would be monolithic, with the end-to-end spectral inference code powered by a single autodiff\textbf{erenti}able machine learning framework, like PyTorch \textbf{\citep{pytorch}} or JAX \citep{jax}.
+However, here we separate the problem into three pieces, only the first \textbf{currently being auto}differentiable.
 \textbf{In \autoref{sec:section2},} we scale out the \texttt{blase} method to 1,314 precomputed synthetic spectral model clones, yielding a downloadable archive of pretrained machine learning models with 128,723 unique spectral lines.
@@
 Our approach can be straightforwardly applied to any other model grid in the future, including substellar atmosphere grids such as Sonora \citep{bobcat, cholla, diamondback, elfowl}, but for now we limit our scope to a \textbf{rectilinear} subset of the PHOENIX grid, focusing on near solar metallicities and a broad range of effective temperature and surface gravity, with details given in \autoref{tab:table1}.
-\textbf{This approach is taken due to the computational cost of interpolation algorithms.
-Future versions with more advanced models will be able to bypass the limitation of a rectilinear subset.}
+\textbf{This approach is taken due to the computational cost and storage impact of interpolation algorithms.
+Future versions with more advanced models will be able to cover a broader scope and bypass the limitation of a rectilinear subset.}
 
@@
 
-\subsection{Inference Algorithm}
+\subsection{\textbf{Gaussian Process Minimization}}
 \textbf{The goals of this study's inference algorithm are to infer the stellar parameters $T_\text{eff}$, $\log(g)$, and [Fe/H] by comparing a spectrum with our reconstructions and solving the optimization problem.}
@@
 
-\subsection{Bayesian Optimizer Performance}
+\subsection{\textbf{Inference Accuracy}}
 To test the performance of the inference algorithm, we used the PHOENIX subset itself.
@@
 For $T_{\mathrm{eff}}$ < 7000 K, the inferred} $T_{\mathrm{eff}}$ differed from the true \textbf{value} by an average of \textbf{93} K, $\log(g)$ \textbf{0.24 dex}, and [Fe/H] \textbf{0.056} dex \textbf{(standard deviations of 139 K, 0.40 dex, and 0.15 dex, respectively)}.
-\textbf{For $T_{\mathrm{eff}} \ge$ 7000 K, the inferred $T_{\mathrm{eff}}$ differed from the true value by an average of 347 K, $\log(g)$ 0.26 dex, and [Fe/H] 0.16 dex (standard deviations of 293 K, 0.32 dex, and 0.23 dex, respectively).}
-From this, we can see that $T_{\mathrm{eff}}$ was the most accurately inferred parameter, followed by $\log(g)$, and then [Fe/H].
+\textbf{For $T_{\mathrm{eff}} \ge$ 7000 K, the inferred $T_{\mathrm{eff}}$ differed from the true value by an average of 347 K, $\log(g)$ 0.26 dex, and [Fe/H] 0.16 dex (standard deviations of 293 K, 0.32 dex, and 0.23 dex, respectively).
+We see that \texttt{blase3D} is much more effective at predicting stellar parameters of lower temperature spectra, presumably due to the larger amount of spectral lines that can be detected.}
+
+\textbf{In the low-temperature regime, our median $T_\text{eff}$ residual is about 10 K, about half that of MINESweeper \citep{minesweeper} and \texttt{astroNN} \citep{leung2019}, and about ten times larger than ASPCAP \citep{ASPCAP}.
+Median $\log(g)$ and [Fe/H] residuals are 0 dex, which are comparable to all three aforementioned methods, however \texttt{blase3D} is overall less consistent as the standard deviation of the residuals are about an order of magnitude larger.
+At high temperatures, the median $T_\text{eff}$ residual changes drastically to about -434 K, which is significantly worse than the other methods.
+$\log(g)$ becomes a few times less accurate, and both $\log(g)$ and [Fe/H] have large standard deviations, all leading to large inconsistencies and worse performance in this temperature regime.
+\texttt{blase3D} at this stage does not quite reach the level of contemporary methods that can deal with more realistic data and cover much larger parameter spaces, but the fact that, in its best case, it can perform to within an order of magnitude using a drastically simpler implementation, sparks promise for future improvements.}
 

### 2025-03-05 paper/paper2/paper.tex
Fix references running off page

@@
     \includegraphics[width=\columnwidth]{figure5}
-    \caption{Plot of \textbf{the 10123.66 \AA \ He II} spectral line shown at the closest PHOENIX grid points to the Sun's stellar parameters \textbf{(solid lines)}, as well as a solar spectrum reconstructed with the PHOENIX generator \textbf{(dotted line)}.}
+    \caption{Plot of \textbf{the 10123.66 \AA \ He II} spectral line shown at PHOENIX grid points \textbf{in the neighborhood of} the Sun's stellar parameters \textbf{(solid lines)}, as well as a \textbf{reconstruction at solar parameters} with the PHOENIX generator \textbf{(dotted line)}.}
     \label{fig:figure5}
@@
 
+\pagebreak
 \bibliography{paper}

### 2025-03-07 paper/paper2/paper.tex
Adjust paper figures

@@
     We first use \texttt{blase} \textbf{to} semi-empirical\textbf{ly recast} spectra \textbf{into} sets of \textbf{Voigt profiles}.
-    \texttt{blase} is \textbf{run} on 1,314 \textbf{synthetic} spectra from \textbf{a rectilinear subset of} the PHOENIX model grid \textbf{($T_\text{eff}$: [2300, 12000] K, $\log(g)$: [2, 6] dex, [Fe/H]: [-0.5, 0] dex, $\lambda$: [8038, 12849] \AA)}.
+    \texttt{blase} is \textbf{run} on 1,314 \textbf{synthetic} spectra from \textbf{a rectilinear subset of} the PHOENIX model grid \textbf{($T_\text{eff}$: [2300, 12000] K, $\log(g)$: [2, 6], [Fe/H]: [-0.5, 0], $\lambda$: [8038, 12849] \AA)}.
     For \textbf{128,723 detected features}, we map stellar parameters to spectral line parameters using linear interpolation.
@@
     Gaussian Process minimization is used to infer stellar parameters by minimizing \textbf{a} root-mean-square loss \textbf{function}.
-    Testing \textbf{210 noise-free models ($T_{\mathrm{eff}}$: [3000, 11000] K, $\log(g)$: [2, 6] dex, [Fe/H]: [-0.5, 0] dex), we find inference errors:} $T_{\mathrm{eff}}$\textbf{: 93 K, $\log(g)$: 0.24 dex, and [Fe/H]: 0.056 dex for $T_\text{eff}$ < 7000 K, and $T_{\mathrm{eff}}$: 347 K, $\log(g)$: 0.26 dex, and [Fe/H]: 0.16 dex for $T_\text{eff} \ge$ 7000 K}.
+    Testing \textbf{210 noise-free models ($T_{\mathrm{eff}}$: [3000, 11000] K, $\log(g)$: [2, 6], [Fe/H]: [-0.5, 0]), we find inference errors:} $T_{\mathrm{eff}}$\textbf{: 93 K, $\log(g)$: 0.24, and [Fe/H]: 0.056 for $T_\text{eff}$ < 7000 K, and $T_{\mathrm{eff}}$: 347 K, $\log(g)$: 0.26, and [Fe/H]: 0.16 for $T_\text{eff} \ge$ 7000 K}.
     \textbf{We also upload online} an archive of \texttt{blase} \textbf{models} of the PHOENIX subset.
@@
         \hline
-        Alpha Element Abundance & $\alpha$ & [0] dex & \bf{N/A}\\
-        Iron Abundance & [Fe/H] & [-0.5, 0] dex & \bf{0.5 dex}\\
+        Alpha Element Abundance & $\alpha$ & [0]& \bf{N/A}\\
+        Iron Abundance & [Fe/H] & [-0.5, 0] & \bf{0.5}\\
         Effective Temperature & $T_{\mathrm{eff}}$ & [2300, 12000] K & \bf{100 K until 7000 K, then 200 K}\\
-        Surface Gravity & $\log(g)$ & [2, 6] dex & \bf{0.5 dex} \\
+        Surface Gravity & $\log(g)$ & [2, 6] & \bf{0.5} \\
         Wavelength & $\lambda$ & [8038, 12849] \AA & $\mathbf{R = 500,000}$\\
@@
 \textbf{Pre-optimized line centers were required to be equal in order to group lines together.
-The values of $\mu$ pre-optimization are recorded by \texttt{blase} to 0.01 \AA (picometer) precision; if they instead used the full available floating point precision, this method would likely detect singular (ungroupable) instances of spectral lines across the grid, completely negating the premise of this study.
+The values of $\mu$ pre-optimization are recorded by \texttt{blase} to 0.01 \AA \ (picometer) precision; if they instead used the full available floating point precision, this method would likely detect singular (ungroupable) instances of spectral lines across the grid, completely negating the premise of this study.
 Of course, this method can be confused by species that produce spectral lines at very close wavelengths and other confounding factors, likely resulting in some false positives and negatives.
@@
 Across the entire PHOENIX subset, \texttt{blase} detected 128,723 spectral lines \textbf{with unique $\mu$}.
-\textbf{On average, a PHOENIX spectrum has 9,167 detected spectral lines, with the minimum being 252 ($T_\text{eff}$ = 12000 K, $\log(g)$ = 6 dex, [Fe/H] = -0.5 dex) and the maximum being 34,551 ($T_\text{eff}$ = 2300 K, $\log(g)$ = 3 dex, [Fe/H] = 0 dex).}
+\textbf{On average, a PHOENIX spectrum has 9,167 detected spectral lines, with the minimum being 252 ($T_\text{eff}$ = 12000 K, $\log(g)$ = 6, [Fe/H] = -0.5) and the maximum being 34,551 ($T_\text{eff}$ = 2300 K, $\log(g)$ = 3, [Fe/H] = 0).}
 Every one of these lines can be visualized as a manifold mapping a 3D stellar parameter vector to a 4D spectral line parameter vector, and every one \textbf{was} interpolated \textbf{to map stellar parameters to spectral line properties}.
@@
 Performance results are shown in \autoref{fig:figure7}, and we can see that \textbf{the PHOENIX generator's parallel mode} is much faster than \textbf{serial mode}.
+\begin{figure}[t!]
+    \centering
+    \includegraphics[width=\columnwidth]{figure7}
+    \caption{\textbf{Plots} showing the time taken to reconstruct varying numbers of spectra using the PHOENIX generator (lower is better).
+    We can see that the time taken per spectr\textbf{um} for the serial implementation hovers around 15 seconds within run-to-run variance, while the pseudo-parallel implementation continually decreases in time taken per spectrum as the number of inputs increases.
+    The speedup factor (higher is better) increases as more spectra are generated, which is also a desirable outcome.}
+    \label{fig:figure7}
+\end{figure}
 \textbf{While the generator takes around 15 seconds per spectrum in serial mode, this drops to 1 second per spectrum in parallel mode as soon as 30 reconstructions.
@@
 \end{table}
-
-\begin{figure}[t!]
-    \centering
-    \includegraphics[width=\columnwidth]{figure7}
-    \caption{\textbf{Plots} showing the time taken to reconstruct varying numbers of spectra using the PHOENIX generator (lower is better).
-    We can see that the time taken per spectr\textbf{um} for the serial implementation hovers around 15 seconds within run-to-run variance, while the pseudo-parallel implementation continually decreases in time taken per spectrum as the number of inputs increases.
-    The speedup factor (higher is better) increases as more spectra are generated, which is also a desirable outcome.}
-    \label{fig:figure7}
-\end{figure}
 
@@
 $\log(g)$ and [Fe/H] were allowed to vary freely.
-The test sample consisted of 21 $T_{\mathrm{eff}}$ values\textbf{: [3000, 11000] K in steps of 400 K}, 5 $\log(g)$ values\textbf{: [2, 6] dex in steps of 0.5 dex} , and 2 [Fe/H] values\textbf{: [-0.5, 0] dex in steps of 0.5 dex}, totaling \textbf{210} unique spectra in the test set.
+The test sample consisted of 21 $T_{\mathrm{eff}}$ values\textbf{: [3000, 11000] K in steps of 400 K}, 5 $\log(g)$ values\textbf{: [2, 6] in steps of 0.5} , and 2 [Fe/H] values\textbf{: [-0.5, 0] in steps of 0.5}, totaling \textbf{210} unique spectra in the test set.
 \textbf{Two test passes were run, meaning that 420 inference tests were run in total.}
@@
 \textbf{We see that there is a distinction between the algorithm's behavior above and below $T_{\mathrm{eff}} = 7000$ K, so we define low and high temperature regimes accordingly.
-For $T_{\mathrm{eff}}$ < 7000 K, the inferred} $T_{\mathrm{eff}}$ differed from the true \textbf{value} by an average of \textbf{93} K, $\log(g)$ \textbf{0.24 dex}, and [Fe/H] \textbf{0.056} dex \textbf{(standard deviations of 139 K, 0.40 dex, and 0.15 dex, respectively)}.
-\textbf{For $T_{\mathrm{eff}} \ge$ 7000 K, the inferred $T_{\mathrm{eff}}$ differed from the true value by an average of 347 K, $\log(g)$ 0.26 dex, and [Fe/H] 0.16 dex (standard deviations of 293 K, 0.32 dex, and 0.23 dex, respectively).
-We see that \texttt{blase3D} is much more effective at predicting stellar parameters of lower temperature spectra, presumably due to the larger amount of spectral lines that can be detected.}
+It is interesting that this observed change occurs where the PHOENIX grid changes in $T_\text{eff}$ sampling from 100 K to 200 K.}
+
+\textbf{For $T_{\mathrm{eff}}$ < 7000 K, the inferred} $T_{\mathrm{eff}}$ differed from the true \textbf{value} by an average of \textbf{93} K, $\log(g)$ \textbf{0.24}, and [Fe/H] \textbf{0.056} \textbf{(standard deviations of 139 K, 0.40, and 0.15, respectively)}.
+\textbf{For $T_{\mathrm{eff}} \ge$ 7000 K, the inferred $T_{\mathrm{eff}}$ differed from the true value by an average of 347 K, $\log(g)$ 0.26, and [Fe/H] 0.16 (standard deviations of 293 K, 0.32, and 0.23, respectively).
+We see that \texttt{blase3D} is more effective and generally more self-consistent when predicting stellar parameters of lower temperature spectra.}
 
@@
 
-\vspace{0.5cm}
+\textbf{All figures in this paper are publicly available on GitHub at \url{https://github.com/Sujay-Shankar/blase3D} under \texttt{paper/paper2}.}
+
+\section*{}
 This material is based on work supported by the National Aeronautics and Space Administration under grant No.
@@
 Sloan Foundation under grant number FG-2021-16592 and support from the National Science Foundation under grant number 1910969.
-\vspace{0.5cm}
+\vspace*{0.5cm}
 
@@
 
-\pagebreak
 \bibliography{paper}

### 2025-03-07 paper/paper2/paper.tex
Microscopic change
Microscopic change 2
Add teff=7000K discussion

@@
     \tilde{\mathsf{V}}_\mu(\lambda) = a\left[\eta \mathbf{L}(\lambda - \mu'; f) + (1 - \eta)\mathbf{G}(\lambda - \mu'; f)\right]\\
-    \eta = \sum_{n=1}^3 \mathbf{u}_n\left(\dfrac{2\gamma}{f}\right)^n\\
+    \eta = \sum_{n=0}^2 \mathbf{u}_{n}\left(\dfrac{2\gamma}{f}\right)^{n+1}\\
     f = 32\sum_{n=0}^5 \mathbf{v}_n \left(\sqrt{2\ln(2)}\sigma\right)^{5-n}(\gamma)^n\\
@@
 \textbf{We see that there is a distinction between the algorithm's behavior above and below $T_{\mathrm{eff}} = 7000$ K, so we define low and high temperature regimes accordingly.
-It is interesting that this observed change occurs where the PHOENIX grid changes in $T_\text{eff}$ sampling from 100 K to 200 K.}
+We know that as $T_\text{eff}$ increases, we detect less lines with \texttt{blase}, as shown in \autoref{fig:figure2}, and lines tend to spread out, possibly resulting in artifacts from our wing cut approximation.
+Any number of physical or computational factors could explain this behavior, although the apparent sharpness of the transition leads us to postulate that it could be due to PHOENIX grid $T_\text{eff}$ sampling changing from 100 K to 200 K at this point.}
 
@@
 
-\textbf{In the low-temperature regime, our median $T_\text{eff}$ residual is about 10 K, about half that of MINESweeper \citep{minesweeper} and \texttt{astroNN} \citep{leung2019}, and about ten times larger than ASPCAP \citep{ASPCAP}.
+\textbf{In the low-temperature regime, our median $T_\text{eff}$ residual is about 10 K, about half that of MINESweeper \citep{minesweeper} and \texttt{astroNN} \citep{leung2019}, but about ten times larger than ASPCAP \citep{ASPCAP}.
 Median $\log(g)$ and [Fe/H] residuals are 0 dex, which are comparable to all three aforementioned methods, however \texttt{blase3D} is overall less consistent as the standard deviation of the residuals are about an order of magnitude larger.
 At high temperatures, the median $T_\text{eff}$ residual changes drastically to about -434 K, which is significantly worse than the other methods.
-$\log(g)$ becomes a few times less accurate, and both $\log(g)$ and [Fe/H] have large standard deviations, all leading to large inconsistencies and worse performance in this temperature regime.
+$\log(g)$ becomes a few times less accurate, and both $\log(g)$ and [Fe/H] have larger standard deviations, all leading to large inconsistencies and worse performance in this temperature regime.
 \texttt{blase3D} at this stage does not quite reach the level of contemporary methods that can deal with more realistic data and cover much larger parameter spaces, but the fact that, in its best case, it can perform to within an order of magnitude using a drastically simpler implementation, sparks promise for future improvements.}
@@
     In addition to the actual stellar parameters, we also took a subset of the PHOENIX wavelength range, with the full [500, 55000] \AA \ wavelength range also being left to future work.
-    Users would be able to fit a greater variety of stellar spectra in many different wavelength regimes.
+    \textbf{With this, u}sers would be able to fit a greater variety of stellar spectra in many different wavelength regimes.
     \item \textit{Strict Wavelength Range}: Currently, the generator only supports inference on spectra whose wavelength limits are either equal to it, or encompass that of the generator and have been truncated to match.
@@
     \item \textit{Extrinsic Absence}: The current design of our algorithm does not account for extrinsic parameters that modify the appearance of spectra such as rotational broadening and Doppler shifting.
-    Future work would need to develop ways to tune these extrinsic parameters alongside stellar parameters, enabling users to optimize these frequently-observed extrinsic parameters on top of the base stellar parameters.
+    Future work would need to develop ways to tune these extrinsic parameters alongside stellar parameters.
     \item \textit{Framework Overhead}: As this algorithm is currently more proof of concept than practical, it uses convenience functions from various libraries, which naturally introduces some level of overhead and leaves performance on the table.

### 2025-03-07 paper/paper2/paper.tex
Fix <

@@
     Gaussian Process minimization is used to infer stellar parameters by minimizing \textbf{a} root-mean-square loss \textbf{function}.
-    Testing \textbf{210 noise-free models ($T_{\mathrm{eff}}$: [3000, 11000] K, $\log(g)$: [2, 6], [Fe/H]: [-0.5, 0]), we find inference errors:} $T_{\mathrm{eff}}$\textbf{: 93 K, $\log(g)$: 0.24, and [Fe/H]: 0.056 for $T_\text{eff}$ < 7000 K, and $T_{\mathrm{eff}}$: 347 K, $\log(g)$: 0.26, and [Fe/H]: 0.16 for $T_\text{eff} \ge$ 7000 K}.
+    Testing \textbf{210 noise-free models ($T_{\mathrm{eff}}$: [3000, 11000] K, $\log(g)$: [2, 6], [Fe/H]: [-0.5, 0]), we find inference errors:} $T_{\mathrm{eff}}$\textbf{: 93 K, $\log(g)$: 0.24, and [Fe/H]: 0.056 for $T_\text{eff} < $ 7000 K, and $T_{\mathrm{eff}}$: 347 K, $\log(g)$: 0.26, and [Fe/H]: 0.16 for $T_\text{eff} \ge$ 7000 K}.
     \textbf{We also upload online} an archive of \texttt{blase} \textbf{models} of the PHOENIX subset.
@@
 
-\textbf{For $T_{\mathrm{eff}}$ < 7000 K, the inferred} $T_{\mathrm{eff}}$ differed from the true \textbf{value} by an average of \textbf{93} K, $\log(g)$ \textbf{0.24}, and [Fe/H] \textbf{0.056} \textbf{(standard deviations of 139 K, 0.40, and 0.15, respectively)}.
+\textbf{For $T_{\mathrm{eff}} < $ 7000 K, the inferred} $T_{\mathrm{eff}}$ differed from the true \textbf{value} by an average of \textbf{93} K, $\log(g)$ \textbf{0.24}, and [Fe/H] \textbf{0.056} \textbf{(standard deviations of 139 K, 0.40, and 0.15, respectively)}.
 \textbf{For $T_{\mathrm{eff}} \ge$ 7000 K, the inferred $T_{\mathrm{eff}}$ differed from the true value by an average of 347 K, $\log(g)$ 0.26, and [Fe/H] 0.16 (standard deviations of 293 K, 0.32, and 0.23, respectively).
@@
 We \textbf{then} introduce and test our \textbf{GP minimization} algorithm \textbf{to infer stellar parameters.} 
-\textbf{Two test passes on 210 noise-free synthetic models ($T_{\mathrm{eff}}$: [3000, 11000] K, $\log(g)$: [2, 6] dex, [Fe/H]: [-0.5, 0] dex) yield} average absolute deviations from true values of \textbf{93} K in $T_\mathrm{eff}$, \textbf{0.24 dex} in $\log(g)$, and \textbf{0.056} dex in [Fe/H] \textbf{in low-temperature ($T_\mathrm{eff}$ < 7000) regimes.
+\textbf{Two test passes on 210 noise-free synthetic models ($T_{\mathrm{eff}}$: [3000, 11000] K, $\log(g)$: [2, 6] dex, [Fe/H]: [-0.5, 0] dex) yield} average absolute deviations from true values of \textbf{93} K in $T_\mathrm{eff}$, \textbf{0.24 dex} in $\log(g)$, and \textbf{0.056} dex in [Fe/H] \textbf{in low-temperature ($T_\mathrm{eff} < $ 7000) regimes.
 In high-temperature regimes ($T_\text{eff} \ge$ 7000), the errors increase to 347 K in $T_\mathrm{eff}$, 0.26 dex in $\log(g)$, and 0.16 dex in [Fe/H]}.

### 2025-04-03 paper/paper2/paper.tex
Begin rev2 stage

@@
     The advent of machine learning (ML) is revolutionary to numerous scientific disciplines, with a growing number of examples in astronomical spectroscopic inference, as ML is more powerful than traditional techniques.
-    Here we introduce a hybrid ML (HML) method combin{ing} automatic differentiation, interpolation, and Bayesian optimization to infer stellar parameters \textbf{of} stellar spectra.
+    Here we introduce a hybrid ML (HML) method combin\textbf{ing} automatic differentiation, interpolation, and Bayesian optimization to infer stellar parameters \textbf{of} stellar spectra.
     We study $T_{\mathrm{eff}}$, $\log(g)$, and [Fe/H], but this method \textbf{could be extended} to other \textbf{parameters} such as [$\alpha$/Fe] \textbf{(alpha element abundance)}, C/O \textbf{(carbon-oxygen ratio)}, and $f_{sed}$ \textbf{(sedimentation efficiency)}.
@@
 The speedup factor between the two methods increases as more spectra are generated, reaching a factor of 20 at 60 reconstructions.
-Assuming linear time scaling with wavelength range, according to KORG benchmarks \citep{korg}, which generate spectra over a 500 \AA \ wavelength range, the PHOENIX generator in serial mode is approximately as fast as KORG and $\sim$30 times faster than Turbospectrum \citep{turbospectrum}.
-The speedup factor in parallel mode compared to KORG is then the same variable speedup factor from serial mode (20 times faster at 60 reconstructions).}
+Assuming linear time scaling with wavelength range, according to KORG benchmarks \citep{korg}, which generate spectra over a 500 \AA \ wavelength range, the PHOENIX generator in serial mode is approximately as fast as KORG and $\sim$30 times faster than Turbospectrum \citep{turbospectrum}.}
 
@@
     \includegraphics[width=\textwidth]{figure8}
-    \caption{\textbf{Residual plots for the GP minimization inference algorithm, all plotted with respect to $T_\text{eff}$ (values aggregated across $\log(g)$ and [Fe/H]).
+    \caption{\textbf{Residual plots for the GP minimization inference algorithm, all plotted with respect to $T_\text{eff}$ (values averaged across $\log(g)$ and [Fe/H]).
     Notice the interesting change in behavior of all residuals at around $T_\text{eff} = 7000$ K.}}

### 2025-04-03 paper/paper2/paper.tex
Finish revision 2

@@
     \includegraphics[width=\textwidth]{figure8}
-    \caption{\textbf{Residual plots for the GP minimization inference algorithm, all plotted with respect to $T_\text{eff}$ (values averaged across $\log(g)$ and [Fe/H]).
+    \caption{\textbf{Residual plots for the GP minimization inference algorithm, all plotted with respect to $T_\text{eff}$.
+    Values are averaged across $\log(g)$ and [Fe/H], and the vertical bars indicate the spread of the 20 data points behind each average via the standard deviation.
+    These bars' size is not surprising; these should not be considered as proper error bars.
+    The 20 data points being averaged at each $T_\text{eff}$ are actually 10 differently distributed pairs of points, with each pair having unique $\log(g)$ and [Fe/H] (the two points within a pair should be considered as identically distributed).
     Notice the interesting change in behavior of all residuals at around $T_\text{eff} = 7000$ K.}}
@@
 
-\textbf{In the low-temperature regime, our median $T_\text{eff}$ residual is about 10 K, about half that of MINESweeper \citep{minesweeper} and \texttt{astroNN} \citep{leung2019}, but about ten times larger than ASPCAP \citep{ASPCAP}.
-Median $\log(g)$ and [Fe/H] residuals are 0 dex, which are comparable to all three aforementioned methods, however \texttt{blase3D} is overall less consistent as the standard deviation of the residuals are about an order of magnitude larger.
-At high temperatures, the median $T_\text{eff}$ residual changes drastically to about -434 K, which is significantly worse than the other methods.
-$\log(g)$ becomes a few times less accurate, and both $\log(g)$ and [Fe/H] have larger standard deviations, all leading to large inconsistencies and worse performance in this temperature regime.
+\textbf{In the low-temperature regime, our median absolute $T_\text{eff}$ residual is 60 K, about three times that of MINESweeper \citep{minesweeper} and \texttt{astroNN} \citep{leung2019}, being within an order of magnitude, but about 60 times larger than ASPCAP \citep{ASPCAP}.
+For $\log(g)$, this quantity is 0.09, about an order of magnitude larger than the other methods, and for [Fe/H], 0, comparable to all three aforementioned methods.
+It should be noted that \texttt{blase3D} is generally less consistent as the standard deviation of its residuals are about an order of magnitude larger than the other methods.
+At high temperatures, the median absolute $T_\text{eff}$ residual changes drastically to 500 K (meaning that more often than not, inference in this regime runs into the edge of the search space), which is significantly worse than both itself in the low temperature regime and the other methods.
+For $\log(g)$, this quantity becomes 0.13, and [Fe/H] remains at 0, although once again, they have larger standard deviations, all leading to large inconsistencies and generally worse performance in this temperature regime.
 \texttt{blase3D} at this stage does not quite reach the level of contemporary methods that can deal with more realistic data and cover much larger parameter spaces, but the fact that, in its best case, it can perform to within an order of magnitude using a drastically simpler implementation, sparks promise for future improvements.}

### 2025-04-08 paper/paper2/paper.tex
Unbold stuff

@@
-\documentclass[twocolumn, linenumbers]{aastex631}
+\documentclass[twocolumn]{aastex631}
 \usepackage{graphicx}
@@
     The advent of machine learning (ML) is revolutionary to numerous scientific disciplines, with a growing number of examples in astronomical spectroscopic inference, as ML is more powerful than traditional techniques.
-    Here we introduce a hybrid ML (HML) method combin\textbf{ing} automatic differentiation, interpolation, and Bayesian optimization to infer stellar parameters \textbf{of} stellar spectra.
-    We study $T_{\mathrm{eff}}$, $\log(g)$, and [Fe/H], but this method \textbf{could be extended} to other \textbf{parameters} such as [$\alpha$/Fe] \textbf{(alpha element abundance)}, C/O \textbf{(carbon-oxygen ratio)}, and $f_{sed}$ \textbf{(sedimentation efficiency)}.
-    We first use \texttt{blase} \textbf{to} semi-empirical\textbf{ly recast} spectra \textbf{into} sets of \textbf{Voigt profiles}.
-    \texttt{blase} is \textbf{run} on 1,314 \textbf{synthetic} spectra from \textbf{a rectilinear subset of} the PHOENIX model grid \textbf{($T_\text{eff}$: [2300, 12000] K, $\log(g)$: [2, 6], [Fe/H]: [-0.5, 0], $\lambda$: [8038, 12849] \AA)}.
-    For \textbf{128,723 detected features}, we map stellar parameters to spectral line parameters using linear interpolation.
-    \textbf{This} create\textbf{s} the PHOENIX generator, \textbf{enabling parallelized spectral synthesis}.
-    Gaussian Process minimization is used to infer stellar parameters by minimizing \textbf{a} root-mean-square loss \textbf{function}.
-    Testing \textbf{210 noise-free models ($T_{\mathrm{eff}}$: [3000, 11000] K, $\log(g)$: [2, 6], [Fe/H]: [-0.5, 0]), we find inference errors:} $T_{\mathrm{eff}}$\textbf{: 93 K, $\log(g)$: 0.24, and [Fe/H]: 0.056 for $T_\text{eff} < $ 7000 K, and $T_{\mathrm{eff}}$: 347 K, $\log(g)$: 0.26, and [Fe/H]: 0.16 for $T_\text{eff} \ge$ 7000 K}.
-    \textbf{We also upload online} an archive of \texttt{blase} \textbf{models} of the PHOENIX subset.
-    This proof-of-concept \textbf{study} shows that semi-empirical HML is a viable alternative to traditional approaches \textbf{in spectroscopic inference}.
+    Here we introduce a hybrid ML (HML) method combining automatic differentiation, interpolation, and Bayesian optimization to infer stellar parameters of stellar spectra.
+    We study $T_{\mathrm{eff}}$, $\log(g)$, and [Fe/H], but this method could be extended to other parameters such as [$\alpha$/Fe] (alpha element abundance), C/O (carbon-oxygen ratio), and $f_{sed}$ (sedimentation efficiency).
+    We first use \texttt{blase} to semi-empirically recast spectra into sets of Voigt profiles.
+    \texttt{blase} is run on 1,314 synthetic spectra from a rectilinear subset of the PHOENIX model grid ($T_\text{eff}$: [2300, 12000] K, $\log(g)$: [2, 6], [Fe/H]: [-0.5, 0], $\lambda$: [8038, 12849] \AA).
+    For 128,723 detected features, we map stellar parameters to spectral line parameters using linear interpolation.
+    This creates the PHOENIX generator, enabling parallelized spectral synthesis.
+    Gaussian Process minimization is used to infer stellar parameters by minimizing a root-mean-square loss function.
+    Testing 210 noise-free models ($T_{\mathrm{eff}}$: [3000, 11000] K, $\log(g)$: [2, 6], [Fe/H]: [-0.5, 0]), we find inference errors: $T_{\mathrm{eff}}$: 93 K, $\log(g)$: 0.24, and [Fe/H]: 0.056 for $T_\text{eff} < $ 7000 K, and $T_{\mathrm{eff}}$: 347 K, $\log(g)$: 0.26, and [Fe/H]: 0.16 for $T_\text{eff} \ge$ 7000 K.
+    We also upload online an archive of \texttt{blase} models of the PHOENIX subset.
+    This proof-of-concept study shows that semi-empirical HML is a viable alternative to traditional approaches in spectroscopic inference.
 \end{abstract}
@@
 
-Modern astronomical spectroscopy \textbf{takes advantage of intuitive, performant spectral models such as KORG \citep{korg}, however it} would \textbf{further} benefit from \textbf{the paradigm of interpretability}.
+Modern astronomical spectroscopy takes advantage of intuitive, performant spectral models such as KORG \citep{korg}, however it would further benefit from the paradigm of interpretability.
 The determination of fundamental properties, the creation of Extreme Precision Radial Velocity (EPRV) templates, and the application of composite spectral fitting could all leverage such an innovation.
@@
 There are many challenges with creating such a foundational spectral model for astrophysics.
-Physical inputs into the spectral modeling process are imperfect\textbf{, and} simplifying assumptions in stellar (and substellar) atmospheres are necessarily inexact \textbf{due to factors such as asymmetries and unknown physics}.
-Computational costs \textbf{make} the training of such models \textbf{challenging; however, advances have been made with attention-based \citep{foundation_attention} and transformer-based \citep{foundation_transformer} models, among others}.
-Most prevailing solutions have had to choose either a model driven approach \textbf{(such as this work)}, in which precomputed models are taken as gospel, or a data driven/empirical approach \textbf{(such as Lux \citep{lux})}, in which our knowledge of stellar physics is ignored or treated phenomenologically, depending on the application.
-\textbf{Hybrid solutions exist and have been implemented, such as those presented in \citealt{leung2019} and \citealt{rains2024}.
-Although a lingering challenge is always how to ideally balance the fusion of model driven and data driven paradigms.}
+Physical inputs into the spectral modeling process are imperfect, and simplifying assumptions in stellar (and substellar) atmospheres are necessarily inexact due to factors such as asymmetries and unknown physics.
+Computational costs make the training of such models challenging; however, advances have been made with attention-based \citep{foundation_attention} and transformer-based \citep{foundation_transformer} models, among others.
+Most prevailing solutions have had to choose either a model driven approach (such as this work), in which precomputed models are taken as gospel, or a data driven/empirical approach (such as Lux \citep{lux}), in which our knowledge of stellar physics is ignored or treated phenomenologically, depending on the application.
+Hybrid solutions exist and have been implemented, such as those presented in \citealt{leung2019} and \citealt{rains2024}.
+Although a lingering challenge is always how to ideally balance the fusion of model driven and data driven paradigms.
 
-Spectroscopic surveys such as APOGEE \textbf{\citep{APOGEE}} use their own in-house pipelines to extract stellar parameters from spectra (ASPCAP \textbf{\citep{ASPCAP}} in the case of APOGEE), and these pipelines appear effective.
+Spectroscopic surveys such as APOGEE \citep{APOGEE} use their own in-house pipelines to extract stellar parameters from spectra (ASPCAP \citep{ASPCAP} in the case of APOGEE), and these pipelines appear effective.
 The core assumption for the vast majority of pipelines is that the data to be analyzed is a list of pixels.
@@
 Multiple efforts have been made in this direction, treating spectra in different ways.
-The standard practice is to treat the wavelength and flux as simply two arrays and use bespoke algorithms tailored to a small number of well-calibrated spectral lines to obtain fundamental stellar parameters and chemical compositions \textbf{(e.g.
-the IGRINS YSO Survey \citep{pipeline1} and the IRTF Spectral Library \citep{pipeline2})}.
+The standard practice is to treat the wavelength and flux as simply two arrays and use bespoke algorithms tailored to a small number of well-calibrated spectral lines to obtain fundamental stellar parameters and chemical compositions (e.g.
+the IGRINS YSO Survey \citep{pipeline1} and the IRTF Spectral Library \citep{pipeline2}).
 Other whole-spectrum fitting abstractions decompose model spectra into an eigenbasis, implementing the data-model comparison stage as a tractable regression, such as \texttt{starfish} \citep{starfish}.
@@
 With the gradient obtained from autodiff's chain rule, we can optimize model parameters by just going against the direction of the gradient (because we usually want to minimize something such as a loss function, we go in the direction of greatest decrease, which is always the negative of the gradient).
-Autodiff has been used successfully in other astrophysical contexts such as \texttt{exojax} \textbf{\citep{exojax}} and \texttt{wobble} \citep{wobble}, but the recasting of spectra into sets of inherently nonlinear spectral lines positions \texttt{blase} as a unique and promising semi-empirical tool.
-The original \texttt{blase} paper demonstrated the ability to tune spectral lines with autodiff \textbf{and `clone' spectra, recasting them as ML models defined by interpretable sets of tuned spectral lines, however it was} restricted to a pre-selected static synthetic model.
+Autodiff has been used successfully in other astrophysical contexts such as \texttt{exojax} \citep{exojax} and \texttt{wobble} \citep{wobble}, but the recasting of spectra into sets of inherently nonlinear spectral lines positions \texttt{blase} as a unique and promising semi-empirical tool.
+The original \texttt{blase} paper demonstrated the ability to tune spectral lines with autodiff and `clone' spectra, recasting them as ML models defined by interpretable sets of tuned spectral lines, however it was restricted to a pre-selected static synthetic model.
 
@@
 
-Ideally this process would be monolithic, with the end-to-end spectral inference code powered by a single autodiff\textbf{erenti}able machine learning framework, like PyTorch \textbf{\citep{pytorch}} or JAX \citep{jax}.
-However, here we separate the problem into three pieces, only the first \textbf{currently being auto}differentiable.
-\textbf{In \autoref{sec:section2},} we scale out the \texttt{blase} method to 1,314 precomputed synthetic spectral model clones, yielding a downloadable archive of pretrained machine learning models with 128,723 unique spectral lines.
-\textbf{In \autoref{sec:section3},} we \textbf{then} fit manifolds mapping stellar parameters to uniformly-derived spectral line parameters using regular grid linear interpolation.
-Finally, \textbf{in \autoref{sec:section4},} we show how reconstructions of the spectra using said manifolds can be used for inferring stellar fundamental properties from spectra.
+Ideally this process would be monolithic, with the end-to-end spectral inference code powered by a single autodifferentiable machine learning framework, like PyTorch \citep{pytorch} or JAX \citep{jax}.
+However, here we separate the problem into three pieces, only the first currently being autodifferentiable.
+In \autoref{sec:section2}, we scale out the \texttt{blase} method to 1,314 precomputed synthetic spectral model clones, yielding a downloadable archive of pretrained machine learning models with 128,723 unique spectral lines.
+In \autoref{sec:section3}, we then fit manifolds mapping stellar parameters to uniformly-derived spectral line parameters using regular grid linear interpolation.
+Finally, in \autoref{sec:section4}, we show how reconstructions of the spectra using said manifolds can be used for inferring stellar fundamental properties from spectra.
 This final step resembles the aims of ``atmospheric retrievals'', but in principle should be faster, adaptive, and reusable.
@@
 \subsection{The PHOENIX Subset}
-\textbf{For the purposes of this study, w}e chose the widely-adopted PHOENIX synthetic spectral model grid \citep{PHOENIX}.
-Our approach can be straightforwardly applied to any other model grid in the future, including substellar atmosphere grids such as Sonora \citep{bobcat, cholla, diamondback, elfowl}, but for now we limit our scope to a \textbf{rectilinear} subset of the PHOENIX grid, focusing on near solar metallicities and a broad range of effective temperature and surface gravity, with details given in \autoref{tab:table1}.
-\textbf{This approach is taken due to the computational cost and storage impact of interpolation algorithms.
-Future versions with more advanced models will be able to cover a broader scope and bypass the limitation of a rectilinear subset.}
+For the purposes of this study, we chose the widely-adopted PHOENIX synthetic spectral model grid \citep{PHOENIX}.
+Our approach can be straightforwardly applied to any other model grid in the future, including substellar atmosphere grids such as Sonora \citep{bobcat, cholla, diamondback, elfowl}, but for now we limit our scope to a rectilinear subset of the PHOENIX grid, focusing on near solar metallicities and a broad range of effective temperature and surface gravity, with details given in \autoref{tab:table1}.
+This approach is taken due to the computational cost and storage impact of interpolation algorithms.
+Future versions with more advanced models will be able to cover a broader scope and bypass the limitation of a rectilinear subset.
 
 \begin{table*}
-    \hspace*{0.6cm}\begin{tabular}{llll}
+    \hspace*{1cm}\begin{tabular}{llll}
         \hline
@@
         \hline
-        Alpha Element Abundance & $\alpha$ & [0]& \bf{N/A}\\
-        Iron Abundance & [Fe/H] & [-0.5, 0] & \bf{0.5}\\
-        Effective Temperature & $T_{\mathrm{eff}}$ & [2300, 12000] K & \bf{100 K until 7000 K, then 200 K}\\
-        Surface Gravity & $\log(g)$ & [2, 6] & \bf{0.5} \\
-        Wavelength & $\lambda$ & [8038, 12849] \AA & $\mathbf{R = 500,000}$\\
+        Alpha Element Abundance & $\alpha$ & [0]& N/A\\
+        Iron Abundance & [Fe/H] & [-0.5, 0] & 0.5\\
+        Effective Temperature & $T_{\mathrm{eff}}$ & [2300, 12000] K & 100 K until 7000 K, then 200 K\\
+        Surface Gravity & $\log(g)$ & [2, 6] & 0.5 \\
+        Wavelength & $\lambda$ & [8038, 12849] \AA & $R = 500,000$\\
         \hline
@@
 \subsection{Preprocessing with \texttt{gollum}}
-First, the PHOENIX subset was programmatically retrieved with \texttt{gollum} \textbf{\citep{gollum}}, which downloaded \textbf{spectra} from the PHOENIX FTP server.
-The spectra were then put through a \textbf{short} preprocessing pipeline similar to that from \citealt{blase}.
+First, the PHOENIX subset was programmatically retrieved with \texttt{gollum} \citep{gollum}, which downloaded spectra from the PHOENIX FTP server.
+The spectra were then put through a short preprocessing pipeline similar to that from \citealt{blase}.
 \begin{enumerate}
-    \item \textit{Blackbody Division}: Since the $T_{\mathrm{eff}}$ of each spectrum is known, the \textbf{corresponding} blackbody spectrum was divided out.
+    \item \textit{Blackbody Division}: Since the $T_{\mathrm{eff}}$ of each spectrum is known, the corresponding blackbody spectrum was divided out.
     \item \textit{Continuum Normalization}: The spectra were further normalized by dividing them by a \nth{5} order polynomial continuum fit using a peak-finding filter in order to eliminate curvature that would inhibit line modeling.
@@
 \end{linenomath*}
-where $\mathsf{\bar{S}}$ is the preprocessed spectrum, $\mathsf{S}$ is the original spectrum,  $\mathsf{B}$ is the blackbody spectrum, \textbf{and} $\mathsf{Q}_n$ is the $n^\mathrm{th}$ order polynomial continuum fit.
+where $\mathsf{\bar{S}}$ is the preprocessed spectrum, $\mathsf{S}$ is the original spectrum,  $\mathsf{B}$ is the blackbody spectrum, and $\mathsf{Q}_n$ is the $n^\mathrm{th}$ order polynomial continuum fit.
 Arithmetic operations between arrays are assumed to be elementwise in all following notation.
@@
 \subsection{Line Identification with \texttt{blase}}
-The next step was to convert the PHOENIX subset into a physically interpretable intermediate representation: a table of spectral line properties rather than a\textbf{n} array of fluxes.
+The next step was to convert the PHOENIX subset into a physically interpretable intermediate representation: a table of spectral line properties rather than an array of fluxes.
 We used \texttt{blase}, which models spectral lines as Voigt profiles and tunes the profiles to mimic the original PHOENIX spectrum with back propagation.
@@
 Four parameters were optimized: the line center $\mu$, the log-amplitude $\ln(a)$, the Gaussian width $\sigma$, and the Lorentzian width $\gamma$.
-The optimization used the Adam optimizer \textbf{\citep{adam}} with a learning rate of 0.05 over 100 epochs.
-\textbf{\citealt{blase} recommends this setup as the minimum, and for a proof-of-concept implementation, we found it best to leave it as is.}
+The optimization used the Adam optimizer \citep{adam} with a learning rate of 0.05 over 100 epochs.
+\citealt{blase} recommends this setup as the minimum, and for a proof-of-concept implementation, we found it best to leave it as is.
 In addition, we limited two custom parameters: wing cut to 6,000 and prominence to 0.005.
-Wing cut \textbf{(in pixels)} is a parameter that determines the extent of the Voigt profile to evaluate, saving computational resources by not evaluating negligible line wings.
-Prominence \textbf{(in normalized flux counts)} sets a lower limit for the amplitude of detected lines, which saves resources by disregarding shallow lines, so in our case we disregard lines \textbf{with amplitude less than} 0.005.
+Wing cut (in pixels) is a parameter that determines the extent of the Voigt profile to evaluate, saving computational resources by not evaluating negligible line wings.
+Prominence (in normalized flux counts) sets a lower limit for the amplitude of detected lines, which saves resources by disregarding shallow lines, so in our case we disregard lines with amplitude less than 0.005.
 In short, our choices for wing cut and prominence decrease the computational cost of \texttt{blase}'s cloning process at the expense of decreasing its accuracy slightly.
@@
 This represents a data compression factor of around 20 just by recasting the spectrum with \texttt{blase}.
-\textbf{The state dictionaries are available on Zenodo at \url{https://zenodo.org/records/11246174} \citep{zenodo}.}
+The state dictionaries are available on Zenodo at \url{https://zenodo.org/records/11246174} \citep{zenodo}.
 
@@
 
-\textbf{Pre-optimized line centers were required to be equal in order to group lines together.
+Pre-optimized line centers were required to be equal in order to group lines together.
 The values of $\mu$ pre-optimization are recorded by \texttt{blase} to 0.01 \AA \ (picometer) precision; if they instead used the full available floating point precision, this method would likely detect singular (ungroupable) instances of spectral lines across the grid, completely negating the premise of this study.
 Of course, this method can be confused by species that produce spectral lines at very close wavelengths and other confounding factors, likely resulting in some false positives and negatives.
-However, we deem it sufficient to approximate this way, as an in-depth treatment would be time-consuming and computationally expensive.}
+However, we deem it sufficient to approximate this way, as an in-depth treatment would be time-consuming and computationally expensive.
 
@@
 Spectral lines were often only detected in some spectra from the PHOENIX subset.
-In \autoref{fig:figure2}, we show that different grid points \textbf{show} differing counts of detected spectral lines.
+In \autoref{fig:figure2}, we show that different grid points show differing counts of detected spectral lines.
 
@@
     We can see that the number of detected lines decreases with increasing $T_{\mathrm{eff}}$.
-    \textbf{Remember} that from $T_{\mathrm{eff}} = 7000$ K onward, PHOENIX's sampling increment changes from 100 K to 200 K.
-    \textbf{Due to some combination of computational artifacts from our line identification assumptions or innate behavior of PHOENIX grid spectra themselves, outliers such as the obvious one here at $T_{\mathrm{eff}} = 7000$ K and $\log(g) = 6$ dex appear.
-    Further exploration into why this happens is not in the scope of this study; for inference we are concerned with the lines themselves, not overall grid point behavior.}}
+    Remember that from $T_{\mathrm{eff}} = 7000$ K onward, PHOENIX's sampling increment changes from 100 K to 200 K.
+    Due to some combination of computational artifacts from our line identification assumptions or innate behavior of PHOENIX grid spectra themselves, outliers such as the obvious one here at $T_{\mathrm{eff}} = 7000$ K and $\log(g) = 6$ dex appear.
+    Further exploration into why this happens is not in the scope of this study; for inference we are concerned with the lines themselves, not overall grid point behavior.}
     \label{fig:figure2}
@@
 
-Th\textbf{e number of detected spectral lines changes as a function of stellar parameter} for different reasons.
+The number of detected spectral lines changes as a function of stellar parameter for different reasons.
 Perhaps for astrophysical reasons: stellar atmospheres genuinely do not produce that line at detectable strength at the given temperature and surface gravity.
-Alternatively, \textbf{\texttt{blase}'s fitting or our line identification assumptions led to artifacts}.
+Alternatively, \texttt{blase}'s fitting or our line identification assumptions led to artifacts.
 Whatever the cause, these missing lines have immediate practical consequences.
@@
 To solve this, we artificially populated missing grid points with log-amplitudes of -1000, which retained interpolator stability while nullifying the evaluated line.
-\textbf{To prove the feasibility of this computational trick, we can do a simple thought experiment.
+To prove the feasibility of this computational trick, we can do a simple thought experiment.
 Moving 1 K in $T_\text{eff}$ away from a grid point on the boundary of detectability would reduce a line's amplitude by a factor of over 20,000 (or more precisely, $e^{10}$).
-This is an extremely large reduction for an extremely small deviation from a grid point, meaning that to any realistic precision, the line immediately vanishes as soon as \texttt{blase} no longer detects it.}
+This is an extremely large reduction for an extremely small deviation from a grid point, meaning that to any realistic precision, the line immediately vanishes as soon as \texttt{blase} no longer detects it.
 
-Examples of the appearance of missing sections in heatmaps where a line does not appear \textbf{can be seen} in \autoref{fig:figure3} and \autoref{fig:figure4}\textbf{, which show heatmaps of the 11617.66 \AA \ C I and the 10123.66 \AA \ He II lines (identified with NIST \citep{NIST}) varying over a solar metallicity slice of the PHOENIX subset stellar parameter space}.
-Across the entire PHOENIX subset, \texttt{blase} detected 128,723 spectral lines \textbf{with unique $\mu$}.
-\textbf{On average, a PHOENIX spectrum has 9,167 detected spectral lines, with the minimum being 252 ($T_\text{eff}$ = 12000 K, $\log(g)$ = 6, [Fe/H] = -0.5) and the maximum being 34,551 ($T_\text{eff}$ = 2300 K, $\log(g)$ = 3, [Fe/H] = 0).}
-Every one of these lines can be visualized as a manifold mapping a 3D stellar parameter vector to a 4D spectral line parameter vector, and every one \textbf{was} interpolated \textbf{to map stellar parameters to spectral line properties}.
+Examples of the appearance of missing sections in heatmaps where a line does not appear can be seen in \autoref{fig:figure3} and \autoref{fig:figure4}, which show heatmaps of the 11617.66 \AA \ C I and the 10123.66 \AA \ He II lines (identified with NIST \citep{NIST}) varying over a solar metallicity slice of the PHOENIX subset stellar parameter space.
+Across the entire PHOENIX subset, \texttt{blase} detected 128,723 spectral lines with unique $\mu$.
+On average, a PHOENIX spectrum has 9,167 detected spectral lines, with the minimum being 252 ($T_\text{eff}$ = 12000 K, $\log(g)$ = 6, [Fe/H] = -0.5) and the maximum being 34,551 ($T_\text{eff}$ = 2300 K, $\log(g)$ = 3, [Fe/H] = 0).
+Every one of these lines can be visualized as a manifold mapping a 3D stellar parameter vector to a 4D spectral line parameter vector, and every one was interpolated to map stellar parameters to spectral line properties.
 \begin{figure*}
@@
     \includegraphics[width=\textwidth]{figure3}
-    \caption{Heatmap\textbf{s} showing how $\ln(a)$, $\sigma$\textbf{, $\gamma$, and $\mu' - \mu$} vary over the PHOENIX subset slice at solar metallicity of a \textbf{C I} spectral line\textbf{, notable for being detected at nearly all subset grid points}.
-    Notice the missing chunk in the top left; \textbf{this could perhaps be due to the line center shifting to the point that \texttt{blase} identified the chunk as a different line entirely (notice the line center offset increasing in magnitude approaching the chunk border).}
-    \textbf{Since} \texttt{blase} did not detect a spectral line here, we artificially populate those points with lines that have $\ln(a) = -1000$ \textbf{during interpolation}.}
+    \caption{Heatmaps showing how $\ln(a)$, $\sigma$, $\gamma$, and $\mu' - \mu$ vary over the PHOENIX subset slice at solar metallicity of a C I spectral line, notable for being detected at nearly all subset grid points.
+    Notice the missing chunk in the top left; this could perhaps be due to the line center shifting to the point that \texttt{blase} identified the chunk as a different line entirely (notice the line center offset increasing in magnitude approaching the chunk border).
+    Since \texttt{blase} did not detect a spectral line here, we artificially populate those points with lines that have $\ln(a) = -1000$ during interpolation.}
     \label{fig:figure3}
@@
     \includegraphics[width=\columnwidth]{figure4}
-    \caption{Heatmap\textbf{s} showing how $\ln(a)$, $\sigma$\textbf{, $\gamma$, and $\mu' - \mu$} vary over the PHOENIX subset slice at solar metallicity of a He II spectral line.
-    \texttt{blase} detects this line at a select chunk of grid points, leading to \textbf{data only existing between $T_\text{eff}$ 3600 K and 6600 K}.}
+    \caption{Heatmaps showing how $\ln(a)$, $\sigma$, $\gamma$, and $\mu' - \mu$ vary over the PHOENIX subset slice at solar metallicity of a He II spectral line.
+    \texttt{blase} detects this line at a select chunk of grid points, leading to data only existing between $T_\text{eff}$ 3600 K and 6600 K}.
     \label{fig:figure4}
@@
 With the given size of the PHOENIX subset, the interpolator list takes up 13.2 GB of disk space.
-\textbf{This is larger than the original PHOENIX subset by about 63\% and than the state dictionary clone archive by a factor of over 34.
+This is larger than the original PHOENIX subset by about 63\% and than the state dictionary clone archive by a factor of over 34.
 This is due to the fact that we are now pickling entire function objects that must memorize the data.
-Future work with more space-efficient models and clever storage strategies could reduce storage space requirements in exchange for I/O performance.}
+Future work with more space-efficient models and clever storage strategies could reduce storage space requirements in exchange for I/O performance.
 This evaluation is able to reconstruct an existing PHOENIX spectrum or alternatively interpolate a new spectrum within the domain of the PHOENIX subset, so we call this the PHOENIX generator.
@@
 Any \texttt{nan} values are set to 1 (which we can do because the spectra are all normalized), and the spectral reconstruction is complete.
-\textbf{In \autoref{fig:figure5}, we see a comparison between a PHOENIX generator solar spectrum and two nearby native PHOENIX grid point spectra at the 10123.66 \AA \ He II spectral line.}
+In \autoref{fig:figure5}, we see a comparison between a PHOENIX generator solar spectrum and two nearby native PHOENIX grid point spectra at the 10123.66 \AA \ He II spectral line.
 \begin{figure}
@@
     \includegraphics[width=\columnwidth]{figure5}
-    \caption{Plot of \textbf{the 10123.66 \AA \ He II} spectral line shown at PHOENIX grid points \textbf{in the neighborhood of} the Sun's stellar parameters \textbf{(solid lines)}, as well as a \textbf{reconstruction at solar parameters} with the PHOENIX generator \textbf{(dotted line)}.}
+    \caption{Plot of the 10123.66 \AA \ He II spectral line shown at PHOENIX grid points in the neighborhood of the Sun's stellar parameters (solid lines), as well as a reconstruction at solar parameters with the PHOENIX generator (dotted line)}.
     \label{fig:figure5}
@@
 
-We can observe in \autoref{fig:figure6} that \textbf{our He II line from a} reconstructed solar spectrum is not simply a pixel interpolation between the nearest \textbf{PHOENIX} grid point \textbf{spectra}.
+We can observe in \autoref{fig:figure6} that our He II line from a reconstructed solar spectrum is not simply a pixel interpolation between the nearest PHOENIX grid point spectra.
 \begin{figure}
@@
     \includegraphics[width=\columnwidth]{figure6}
-    \caption{\textbf{Comparison between the PHOENIX generator's line interpolation and a standard pixel interpolation of the 10123.66 \AA \ He II spectral line.
-    Notice that line interpolation results in more symmetrical line wings due to the use of Voigt profiles.}}
+    \caption{Comparison between the PHOENIX generator's line interpolation and a standard pixel interpolation of the 10123.66 \AA \ He II spectral line.
+    Notice that line interpolation results in more symmetrical line wings due to the use of Voigt profiles.}
     \label{fig:figure6}
@@
 Second, for an array of multiple inputs, which would be relevant in parallel applications.
-\textbf{\texttt{scipy}'s} \texttt{RegularGridInterpolator} API allows for the passing in of an entire array of input coordinates to be evaluated at once.
+\texttt{scipy}'s \texttt{RegularGridInterpolator} API allows for the passing in of an entire array of input coordinates to be evaluated at once.
 However using \texttt{blase} to reconstruct the spectrum from our interpolated state dictionary is always done serially, leading to what is actually more of a pseudo-parallel evaluation, but extremely performant nonetheless.
 
-Performance results are shown in \autoref{fig:figure7}, and we can see that \textbf{the PHOENIX generator's parallel mode} is much faster than \textbf{serial mode}.
+Performance results are shown in \autoref{fig:figure7}, and we can see that the PHOENIX generator's parallel mode is much faster than serial mode.
 \begin{figure}[t!]
@@
     \includegraphics[width=\columnwidth]{figure7}
-    \caption{\textbf{Plots} showing the time taken to reconstruct varying numbers of spectra using the PHOENIX generator (lower is better).
-    We can see that the time taken per spectr\textbf{um} for the serial implementation hovers around 15 seconds within run-to-run variance, while the pseudo-parallel implementation continually decreases in time taken per spectrum as the number of inputs increases.
+    \caption{Plots showing the time taken to reconstruct varying numbers of spectra using the PHOENIX generator (lower is better).
+    We can see that the time taken per spectrum for the serial implementation hovers around 15 seconds within run-to-run variance, while the pseudo-parallel implementation continually decreases in time taken per spectrum as the number of inputs increases.
     The speedup factor (higher is better) increases as more spectra are generated, which is also a desirable outcome.}
@@
 \end{figure}
-\textbf{While the generator takes around 15 seconds per spectrum in serial mode, this drops to 1 second per spectrum in parallel mode as soon as 30 reconstructions.
+While the generator takes around 15 seconds per spectrum in serial mode, this drops to 1 second per spectrum in parallel mode as soon as 30 reconstructions.
 The speedup factor between the two methods increases as more spectra are generated, reaching a factor of 20 at 60 reconstructions.
-Assuming linear time scaling with wavelength range, according to KORG benchmarks \citep{korg}, which generate spectra over a 500 \AA \ wavelength range, the PHOENIX generator in serial mode is approximately as fast as KORG and $\sim$30 times faster than Turbospectrum \citep{turbospectrum}.}
+Assuming linear time scaling with wavelength range, according to KORG benchmarks \citep{korg}, which generate spectra over a 500 \AA \ wavelength range, the PHOENIX generator in serial mode is approximately as fast as KORG and $\sim$30 times faster than Turbospectrum \citep{turbospectrum}.
 
-\textbf{The computer used for this test has specifications shown in \autoref{tab:table2} (note that spectral reconstruction does not currently leverage the GPU):}
+The computer used for this test has specifications shown in \autoref{tab:table2} (note that spectral reconstruction does not currently leverage the GPU):
 \begin{table}[h!]
@@
     \end{tabular}
-    \caption{\textbf{Specifications of the} machine used for \textbf{cloning and reconstruction testing}.
-    \textbf{This machine is owned by Caroline Morley.}}
+    \caption{Specifications of the machine used for cloning and reconstruction testing.
+    This machine is owned by Caroline Morley.}
     \label{tab:table2}
@@
 
-\subsection{\textbf{Gaussian Process Minimization}}
-\textbf{The goals of this study's inference algorithm are to infer the stellar parameters $T_\text{eff}$, $\log(g)$, and [Fe/H] by comparing a spectrum with our reconstructions and solving the optimization problem.}
-We elected to use Bayesian optimization \textbf{for this component}, specifically the \texttt{gp\_minimize} function from the \texttt{scikit-optimize} library \citep{skopt}.
+\subsection{Gaussian Process Minimization}
+The goals of this study's inference algorithm are to infer the stellar parameters $T_\text{eff}$, $\log(g)$, and [Fe/H] by comparing a spectrum with our reconstructions and solving the optimization problem.
+We elected to use Bayesian optimization for this component, specifically the \texttt{gp\_minimize} function from the \texttt{scikit-optimize} library \citep{skopt}.
 This algorithm uses a Gaussian Process to model the objective function, which in this case was the RMS (Root-Mean-Square) loss between the interpolated spectrum $\mathsf{M}$ and the true spectrum $\mathsf{D}$, defined as:
@@
 The optimizer was configured to first run 100 random evaluations to seed the surrogate model, then run 20 more evaluations now guided by the surrogate model.
-\textbf{A random evaluation is simply a randomly generated tuple of stellar parameters within the parameter space of the PHOENIX generator, which is then evaluated with the generator to get a reconstruction and then compared with the inference target to retrieve its RMS error.
+A random evaluation is simply a randomly generated tuple of stellar parameters within the parameter space of the PHOENIX generator, which is then evaluated with the generator to get a reconstruction and then compared with the inference target to retrieve its RMS error.
 The surrogate model is the approximation to the true objective function that the optimizer constructs based on the set of random evaluations (in this case defined by a set of 100 points rather than a fully continuous, infinitely resolved function)
 Guided evaluations are different, they are not randomly generated tuples of stellar parameters.
-They use the surrogate model in order to select new test points for the minimum of the objective function.}
-This totals to 120 evaluations, \textbf{100} to create \textbf{the} surrogate model, then \textbf{20} guided evaluation\textbf{s}, which was deemed sufficient for this study.
+They use the surrogate model in order to select new test points for the minimum of the objective function.
+This totals to 120 evaluations, 100 to create the surrogate model, then 20 guided evaluations, which was deemed sufficient for this study.
 Fine-tuning these numbers is possible, but simply not warranted for a proof-of concept method.
-\textbf{Tweaking these numbers could lead to a more accurate inference, however it would incur additional computational cost as well.}
-One inference takes on average \textbf{200 seconds, using the computer with specifications in \autoref{tab:table3}}.
+Tweaking these numbers could lead to a more accurate inference, however it would incur additional computational cost as well.
+One inference takes on average 200 seconds, using the computer with specifications in \autoref{tab:table3}.
 
-\textbf{Inference algorithms that output more information and produce posterior distributions such as \texttt{starfish} \citep{starfish} or MINESweeper \citep{minesweeper} take on the order of an hour to complete an inference on supercomputers.
-\texttt{astroNN} \citep{leung2019} is a more direct comparison, and represents an extremely powerful inference tool that performs inference thousands of times faster thanks to its neural network architecture.}
+Inference algorithms that output more information and produce posterior distributions such as \texttt{starfish} \citep{starfish} or MINESweeper \citep{minesweeper} take on the order of an hour to complete an inference on supercomputers.
+\texttt{astroNN} \citep{leung2019} is a more direct comparison, and represents an extremely powerful inference tool that performs inference thousands of times faster thanks to its neural network architecture.
 \begin{table}[h!]
@@
     \end{tabular}
-    \caption{\textbf{Specifications of the machine used for general testing and for generating visualizations.}
-    \textbf{This machine is owned by Sujay Shankar.}}
+    \caption{Specifications of the machine used for general testing and for generating visualizations.
+    This machine is owned by Sujay Shankar.}
     \label{tab:table3}
@@
 
-\subsection{\textbf{Inference Accuracy}}
+\subsection{Inference Accuracy}
 To test the performance of the inference algorithm, we used the PHOENIX subset itself.
-\textbf{This is possible for two reasons: first, the PHOENIX generator has not truly memorized the PHOENIX subset due to its Voigt profile basis and the ignoring of weaker features, and secondly, random samples within the parameter space are extremely unlikely to be PHOENIX grid points themselves, avoiding any overlap.}
+This is possible for two reasons: first, the PHOENIX generator has not truly memorized the PHOENIX subset due to its Voigt profile basis and the ignoring of weaker features, and secondly, random samples within the parameter space are extremely unlikely to be PHOENIX grid points themselves, avoiding any overlap.
 
-We know that in typical observational settings, a coarse estimate for $T_{\mathrm{eff}}$ tends to be fairly well-constrained from ancillary data \textbf{such as photometric color index}.
+We know that in typical observational settings, a coarse estimate for $T_{\mathrm{eff}}$ tends to be fairly well-constrained from ancillary data such as photometric color index.
 So when testing the inference algorithm, we limited its search space to only include $T_{\mathrm{eff}}$ which lay within 500 K of the true value on either side.
 $\log(g)$ and [Fe/H] were allowed to vary freely.
-The test sample consisted of 21 $T_{\mathrm{eff}}$ values\textbf{: [3000, 11000] K in steps of 400 K}, 5 $\log(g)$ values\textbf{: [2, 6] in steps of 0.5} , and 2 [Fe/H] values\textbf{: [-0.5, 0] in steps of 0.5}, totaling \textbf{210} unique spectra in the test set.
-\textbf{Two test passes were run, meaning that 420 inference tests were run in total.}
+The test sample consisted of 21 $T_{\mathrm{eff}}$ values: [3000, 11000] K in steps of 400 K, 5 $\log(g)$ values: [2, 6] in steps of 0.5 , and 2 [Fe/H] values: [-0.5, 0] in steps of 0.5, totaling 210 unique spectra in the test set.
+Two test passes were run, meaning that 420 inference tests were run in total.
 
@@
 \end{figure*}
-\textbf{We see that there is a distinction between the algorithm's behavior above and below $T_{\mathrm{eff}} = 7000$ K, so we define low and high temperature regimes accordingly.
+We see that there is a distinction between the algorithm's behavior above and below $T_{\mathrm{eff}} = 7000$ K, so we define low and high temperature regimes accordingly.
 We know that as $T_\text{eff}$ increases, we detect less lines with \texttt{blase}, as shown in \autoref{fig:figure2}, and lines tend to spread out, possibly resulting in artifacts from our wing cut approximation.
-Any number of physical or computational factors could explain this behavior, although the apparent sharpness of the transition leads us to postulate that it could be due to PHOENIX grid $T_\text{eff}$ sampling changing from 100 K to 200 K at this point.}
+Any number of physical or computational factors could explain this behavior, although the apparent sharpness of the transition leads us to postulate that it could be due to PHOENIX grid $T_\text{eff}$ sampling changing from 100 K to 200 K at this point.
 
-\textbf{For $T_{\mathrm{eff}} < $ 7000 K, the inferred} $T_{\mathrm{eff}}$ differed from the true \textbf{value} by an average of \textbf{93} K, $\log(g)$ \textbf{0.24}, and [Fe/H] \textbf{0.056} \textbf{(standard deviations of 139 K, 0.40, and 0.15, respectively)}.
-\textbf{For $T_{\mathrm{eff}} \ge$ 7000 K, the inferred $T_{\mathrm{eff}}$ differed from the true value by an average of 347 K, $\log(g)$ 0.26, and [Fe/H] 0.16 (standard deviations of 293 K, 0.32, and 0.23, respectively).
-We see that \texttt{blase3D} is more effective and generally more self-consistent when predicting stellar parameters of lower temperature spectra.}
+For $T_{\mathrm{eff}} < $ 7000 K, the inferred $T_{\mathrm{eff}}$ differed from the true value by an average of 93 K, $\log(g)$ 0.24, and [Fe/H] 0.056 (standard deviations of 139 K, 0.40, and 0.15, respectively).
+For $T_{\mathrm{eff}} \ge$ 7000 K, the inferred $T_{\mathrm{eff}}$ differed from the true value by an average of 347 K, $\log(g)$ 0.26, and [Fe/H] 0.16 (standard deviations of 293 K, 0.32, and 0.23, respectively).
+We see that \texttt{blase3D} is more effective and generally more self-consistent when predicting stellar parameters of lower temperature spectra.
 
@@
 Notably, this system tracks the shifting of spectral lines as a function of stellar parameters, which traditionally has been uncommon for algorithms to assess systematically.
-We balance rigidity \textbf{in the Voigt profile basis} and flexibility \textbf{in line parameter tunability, resulting in an} interpretab\textbf{le system}.
+We balance rigidity in the Voigt profile basis and flexibility in line parameter tunability, resulting in an interpretable system.
 
 To summarize, we bring to scientists the ability to understand exactly what spectral properties we consider mathematically and how we expect those properties to behave on an incredibly detailed level (the 4 line parameters of all spectral lines and their corresponding interpolating manifolds).
-\textbf{We treat the PHOENIX} precomputed model grid as the ground truth, \textbf{constructing the PHOENIX} generator that then interface\textbf{s} with our inference algorithm.
+We treat the PHOENIX precomputed model grid as the ground truth, constructing the PHOENIX generator that then interfaces with our inference algorithm.
 
@@
 First, the familiar \texttt{scipy}-based system will serve as an easy entry point for most practitioners who are unfamiliar with autodiff, and may still benefit from and modify the code without expert ML knowledge.
-Second, this non-autodiff version serves as a\textbf{n initial} benchmark against which an inevitable autodifferentiable version \textbf{of \texttt{blase}} may be compared.
+Second, this non-autodiff version serves as an initial benchmark against which an inevitable autodifferentiable version of \texttt{blase} may be compared.
 Finally, the inventory of familiar interpolation algorithms have not yet been ported to PyTorch or JAX, since machine learning or Gaussian Process fitting schemes are generally preferred within the ML community.
@@
     In addition to the actual stellar parameters, we also took a subset of the PHOENIX wavelength range, with the full [500, 55000] \AA \ wavelength range also being left to future work.
-    \textbf{With this, u}sers would be able to fit a greater variety of stellar spectra in many different wavelength regimes.
+    With this, users would be able to fit a greater variety of stellar spectra in many different wavelength regimes.
     \item \textit{Strict Wavelength Range}: Currently, the generator only supports inference on spectra whose wavelength limits are either equal to it, or encompass that of the generator and have been truncated to match.
@@
     \item \textit{Single Model Grid}: The PHOENIX grid is not the only model grid of synthetic spectra available, and it does not apply to all types of stars.
-    Future work would extend the reach of this study's algorithm to encompass other model grids such as the Sonora series of substellar models \textbf{\citep{bobcat, cholla, diamondback, elfowl}}, ATLAS \citep{atlas}, and coolTLUSTY \citep{coolTLUSTY}, reaching practitioners studying various types of stars and spectra.
-    \textbf{Future} \texttt{blase} \textbf{versions will} be able to have an option for the user to input which model grid they would like to base the inference on, and to get even more advanced, perhaps even have the ability to intelligently determine which model grid to use automatically.
+    Future work would extend the reach of this study's algorithm to encompass other model grids such as the Sonora series of substellar models \citep{bobcat, cholla, diamondback, elfowl}, ATLAS \citep{atlas}, and coolTLUSTY \citep{coolTLUSTY}, reaching practitioners studying various types of stars and spectra.
+    Future \texttt{blase} versions will be able to have an option for the user to input which model grid they would like to base the inference on, and to get even more advanced, perhaps even have the ability to intelligently determine which model grid to use automatically.
     \item \textit{Memorization vs.\ Generalization}: The current design of the algorithm constructs manifolds using interpolation.
@@
     As interpolators require memorization of the data, advanced interpolation becomes extremely expensive in terms of disk utilization.
-    Future work would involve constructing manifolds using \textbf{more generalizable ML} methods \textbf{such as lasso or ridge regression}, which would allow for much better generalization, high speed, and lower disk utilization at the expense of some accuracy.
+    Future work would involve constructing manifolds using more generalizable ML methods such as lasso or ridge regression, which would allow for much better generalization, high speed, and lower disk utilization at the expense of some accuracy.
     \item \textit{Extrinsic Absence}: The current design of our algorithm does not account for extrinsic parameters that modify the appearance of spectra such as rotational broadening and Doppler shifting.
@@
     \item \textit{The Continuum Black Box}: Continuum normalization is a process that is not yet completely understood, and is currently done as a preprocessing step with a fairly simple algorithm.
-    Future work would dive deeper into the science of continu\textbf{a} and develop more advanced methods that can discern continu\textbf{a} with greater accuracy and less modeling restrictions.
+    Future work would dive deeper into the science of continua and develop more advanced methods that can discern continua with greater accuracy and less modeling restrictions.
     This would increase accuracy for end users.
@@
 \section{Conclusion}
-In this study, we have presented a proof-of-concept algorithm that \textbf{deals with} a subset of the PHOENIX spectral model grid \textbf{($T_\text{eff}$: [2300, 12000] K, $\log(g)$: [2, 6] dex, [Fe/H]: [-0.5, 0] dex, $\lambda$: [8038, 12849] \AA)}.
-We create state dictionaries for all spectra in the PHOENIX subset \textbf{using \texttt{blase}, recasting spectra into lists of Voigt profiles, and} lossily compress \textbf{the spectral data by a} factor of around 20.
-The\textbf{y are available on} Zenodo at \url{https://zenodo.org/records/11246174} \citep{zenodo}.
+In this study, we have presented a proof-of-concept algorithm that deals with a subset of the PHOENIX spectral model grid ($T_\text{eff}$: [2300, 12000] K, $\log(g)$: [2, 6] dex, [Fe/H]: [-0.5, 0] dex, $\lambda$: [8038, 12849] \AA).
+We create state dictionaries for all spectra in the PHOENIX subset using \texttt{blase}, recasting spectra into lists of Voigt profiles, and lossily compress the spectral data by a factor of around 20.
+They are available on Zenodo at \url{https://zenodo.org/records/11246174} \citep{zenodo}.
 
@@
 
-We \textbf{then} introduce and test our \textbf{GP minimization} algorithm \textbf{to infer stellar parameters.} 
-\textbf{Two test passes on 210 noise-free synthetic models ($T_{\mathrm{eff}}$: [3000, 11000] K, $\log(g)$: [2, 6] dex, [Fe/H]: [-0.5, 0] dex) yield} average absolute deviations from true values of \textbf{93} K in $T_\mathrm{eff}$, \textbf{0.24 dex} in $\log(g)$, and \textbf{0.056} dex in [Fe/H] \textbf{in low-temperature ($T_\mathrm{eff} < $ 7000) regimes.
-In high-temperature regimes ($T_\text{eff} \ge$ 7000), the errors increase to 347 K in $T_\mathrm{eff}$, 0.26 dex in $\log(g)$, and 0.16 dex in [Fe/H]}.
+We then introduce and test our GP minimization algorithm to infer stellar parameters. 
+Two test passes on 210 noise-free synthetic models ($T_{\mathrm{eff}}$: [3000, 11000] K, $\log(g)$: [2, 6] dex, [Fe/H]: [-0.5, 0] dex) yield average absolute deviations from true values of 93 K in $T_\mathrm{eff}$, 0.24 dex in $\log(g)$, and 0.056 dex in [Fe/H] in low-temperature ($T_\mathrm{eff} < $ 7000) regimes.
+In high-temperature regimes ($T_\text{eff} \ge$ 7000), the errors increase to 347 K in $T_\mathrm{eff}$, 0.26 dex in $\log(g)$, and 0.16 dex in [Fe/H].
 
-In its current state, our algorithm requir\textbf{es} that the spectra not contain noticeable Doppler shifting, rotational broadening, \textbf{noise}, or other confounding factors.
+In its current state, our algorithm requires that the spectra not contain noticeable Doppler shifting, rotational broadening, noise, or other confounding factors.
 The methods discussed here represent a step down a road not traveled in spectral inference, and have the potential to become more advanced in the future by fully utilizing the strengths of physics-informed machine learning.
 
-\textbf{All figures in this paper are publicly available on GitHub at \url{https://github.com/Sujay-Shankar/blase3D} under \texttt{paper/paper2}.}
+All figures in this paper are publicly available on GitHub at \url{https://github.com/Sujay-Shankar/blase3D} under \texttt{paper/paper2}.
 

### 2025-04-25 paper/paper2/paper.tex
Unbold some last things

@@
     \includegraphics[width=\textwidth]{figure8}
-    \caption{\textbf{Residual plots for the GP minimization inference algorithm, all plotted with respect to $T_\text{eff}$.
+    \caption{Residual plots for the GP minimization inference algorithm, all plotted with respect to $T_\text{eff}$.
     Values are averaged across $\log(g)$ and [Fe/H], and the vertical bars indicate the spread of the 20 data points behind each average via the standard deviation.
@@
     The 20 data points being averaged at each $T_\text{eff}$ are actually 10 differently distributed pairs of points, with each pair having unique $\log(g)$ and [Fe/H] (the two points within a pair should be considered as identically distributed).
-    Notice the interesting change in behavior of all residuals at around $T_\text{eff} = 7000$ K.}}
+    Notice the interesting change in behavior of all residuals at around $T_\text{eff} = 7000$ K.}
     \label{fig:figure8}
@@
 
-\textbf{In the low-temperature regime, our median absolute $T_\text{eff}$ residual is 60 K, about three times that of MINESweeper \citep{minesweeper} and \texttt{astroNN} \citep{leung2019}, being within an order of magnitude, but about 60 times larger than ASPCAP \citep{ASPCAP}.
+In the low-temperature regime, our median absolute $T_\text{eff}$ residual is 60 K, about three times that of MINESweeper \citep{minesweeper} and \texttt{astroNN} \citep{leung2019}, being within an order of magnitude, but about 60 times larger than ASPCAP \citep{ASPCAP}.
 For $\log(g)$, this quantity is 0.09, about an order of magnitude larger than the other methods, and for [Fe/H], 0, comparable to all three aforementioned methods.
@@
 For $\log(g)$, this quantity becomes 0.13, and [Fe/H] remains at 0, although once again, they have larger standard deviations, all leading to large inconsistencies and generally worse performance in this temperature regime.
-\texttt{blase3D} at this stage does not quite reach the level of contemporary methods that can deal with more realistic data and cover much larger parameter spaces, but the fact that, in its best case, it can perform to within an order of magnitude using a drastically simpler implementation, sparks promise for future improvements.}
+\texttt{blase3D} at this stage does not quite reach the level of contemporary methods that can deal with more realistic data and cover much larger parameter spaces, but the fact that, in its best case, it can perform to within an order of magnitude using a drastically simpler implementation, sparks promise for future improvements.
 
